### 2020-02-27 dfol_draft/Extras/mytemplate.tex
initial trials

<!DOCTYPE html>
<html lang="en">
  <head>
    <meta charset="utf-8">
  <link rel="dns-prefetch" href="https://github.githubassets.com">
  <link rel="dns-prefetch" href="https://avatars0.githubusercontent.com">
  <link rel="dns-prefetch" href="https://avatars1.githubusercontent.com">
  <link rel="dns-prefetch" href="https://avatars2.githubusercontent.com">
  <link rel="dns-prefetch" href="https://avatars3.githubusercontent.com">
  <link rel="dns-prefetch" href="https://github-cloud.s3.amazonaws.com">
  <link rel="dns-prefetch" href="https://user-images.githubusercontent.com/">

  <link crossorigin="anonymous" media="all" integrity="sha512-ZUjVod2EvYMDbGqRSyW0rpfgBq3i+gnR/4PfrzLsy5f20oIcRfgFQFVKgi3Ztp917bP1K/kdP5q8+nAlJ3+cFA==" rel="stylesheet" href="https://github.githubassets.com/assets/frameworks-6548d5a1dd84bd83036c6a914b25b4ae.css" />

    <link crossorigin="anonymous" media="all" integrity="sha512-MDyKkLoFH4URLSBid10ZAluaTFZmFNB1z9P8gQAKfkXORgZCf3t5kmOSxiOYOAdlt0Ok30kH1FEtdolXdyYcJw==" rel="stylesheet" href="https://github.githubassets.com/assets/github-303c8a90ba051f85112d2062775d1902.css" />

  <meta name="viewport" content="width=device-width">

  <title>Affiliations/mytemplate.tex at master · drdanholmes/Affiliations</title>
    <meta name="description" content="RMarkdown Templates for Managing Academic Affiliations for MS-Word and LaTeX - drdanholmes/Affiliations">
    <link rel="search" type="application/opensearchdescription+xml" href="/opensearch.xml" title="GitHub">
  <link rel="fluid-icon" href="https://github.com/fluidicon.png" title="GitHub">
  <meta property="fb:app_id" content="1401488693436528">

    <meta name="twitter:image:src" content="https://avatars3.githubusercontent.com/u/17508391?s=400&amp;v=4" /><meta name="twitter:site" content="@github" /><meta name="twitter:card" content="summary" /><meta name="twitter:title" content="drdanholmes/Affiliations" /><meta name="twitter:description" content="RMarkdown Templates for Managing Academic Affiliations for MS-Word and LaTeX - drdanholmes/Affiliations" />
    <meta property="og:image" content="https://avatars3.githubusercontent.com/u/17508391?s=400&amp;v=4" /><meta property="og:site_name" content="GitHub" /><meta property="og:type" content="object" /><meta property="og:title" content="drdanholmes/Affiliations" /><meta property="og:url" content="https://github.com/drdanholmes/Affiliations" /><meta property="og:description" content="RMarkdown Templates for Managing Academic Affiliations for MS-Word and LaTeX - drdanholmes/Affiliations" />

  <link rel="assets" href="https://github.githubassets.com/">
  <link rel="web-socket" href="wss://live.github.com/_sockets/VjI6NDg1MTMzNjY1OmNiZGQ5N2U1MmNlMWZiMDlkYTkzZjA5YTdhYjBjNTNkZDc0YTFmMTNhN2ZmMjk4YmVmNGU3MjdjNDM5ZGMyMTQ=--4d2fa2a817b173224c57a8283c337d71f5acd04c">
  <link rel="sudo-modal" href="/sessions/sudo_modal">

  <meta name="request-id" content="CB95:4C4D:11ECB4:1D8804:5E5690DB" data-pjax-transient="true" /><meta name="html-safe-nonce" content="e10583c265ecb6446b852a6b2c5b5a01cd997a34" data-pjax-transient="true" /><meta name="visitor-payload" content="eyJyZWZlcnJlciI6Imh0dHBzOi8vZ2l0aHViLmNvbS9kcmRhbmhvbG1lcy9BZmZpbGlhdGlvbnMvdHJlZS9tYXN0ZXIvRXh0cmFzIiwicmVxdWVzdF9pZCI6IkNCOTU6NEM0RDoxMUVDQjQ6MUQ4ODA0OjVFNTY5MERCIiwidmlzaXRvcl9pZCI6IjM4OTEyNDI0MDc1NjI1MTY0OTEiLCJyZWdpb25fZWRnZSI6ImlhZCIsInJlZ2lvbl9yZW5kZXIiOiJpYWQifQ==" data-pjax-transient="true" /><meta name="visitor-hmac" content="cd7c112e8258c3ed359b73ec6c0202c7e0907e277d645f7789285499e1c5c921" data-pjax-transient="true" />

  <meta name="github-keyboard-shortcuts" content="repository,source-code" data-pjax-transient="true" />

  <meta name="selected-link" value="repo_source" data-pjax-transient>

      <meta name="google-site-verification" content="KT5gs8h0wvaagLKAVWq8bbeNwnZZK1r1XQysX3xurLU">
    <meta name="google-site-verification" content="ZzhVyEFwb7w3e0-uOTltm8Jsck2F5StVihD0exw2fsA">
    <meta name="google-site-verification" content="GXs5KoUUkNCoaAZn7wPN-t01Pywp9M3sEjnt_3_ZWPc">

  <meta name="octolytics-host" content="collector.githubapp.com" /><meta name="octolytics-app-id" content="github" /><meta name="octolytics-event-url" content="https://collector.githubapp.com/github-external/browser_event" /><meta name="octolytics-dimension-ga_id" content="" class="js-octo-ga-id" /><meta name="octolytics-actor-id" content="15913450" /><meta name="octolytics-actor-login" content="chguiterman" /><meta name="octolytics-actor-hash" content="ac81b7290092b2c4a1347b423992342f7581f20c94fd8a35e3bf98004cc18b13" />
<meta name="analytics-location" content="/&lt;user-name&gt;/&lt;repo-name&gt;/blob/show" data-pjax-transient="true" />

    <meta name="google-analytics" content="UA-3769691-2">

  <meta class="js-ga-set" name="userId" content="aff5acac9f546acde3aae63df4076c10">

<meta class="js-ga-set" name="dimension1" content="Logged In">

      <meta name="hostname" content="github.com">
    <meta name="user-login" content="chguiterman">

      <meta name="expected-hostname" content="github.com">

      <meta name="js-proxy-site-detection-payload" content="MTAzNzdhZjAzZmFhNDJkY2IxZDQyYWIzOWU0M2NjOWM1MTgxNGYwNzE2Y2Y4ODZjMDY4MWEzYWY0NGU4ODE0MHx7InJlbW90ZV9hZGRyZXNzIjoiNjQuMjIzLjE1OC4xNTgiLCJyZXF1ZXN0X2lkIjoiQ0I5NTo0QzREOjExRUNCNDoxRDg4MDQ6NUU1NjkwREIiLCJ0aW1lc3RhbXAiOjE1ODI3MzE0ODgsImhvc3QiOiJnaXRodWIuY29tIn0=">

    <meta name="enabled-features" content="MARKETPLACE_FEATURED_BLOG_POSTS,MARKETPLACE_INVOICED_BILLING,MARKETPLACE_SOCIAL_PROOF_CUSTOMERS,MARKETPLACE_TRENDING_SOCIAL_PROOF,MARKETPLACE_RECOMMENDATIONS,MARKETPLACE_PENDING_INSTALLATIONS,RELATED_ISSUES,GHE_CLOUD_TRIAL">

  <meta http-equiv="x-pjax-version" content="f93c7a12e9ec02d97f76ef56eb7dd432">

      <link href="https://github.com/drdanholmes/Affiliations/commits/master.atom" rel="alternate" title="Recent Commits to Affiliations:master" type="application/atom+xml">

  <meta name="go-import" content="github.com/drdanholmes/Affiliations git https://github.com/drdanholmes/Affiliations.git">

  <meta name="octolytics-dimension-user_id" content="17508391" /><meta name="octolytics-dimension-user_login" content="drdanholmes" /><meta name="octolytics-dimension-repository_id" content="204400584" /><meta name="octolytics-dimension-repository_nwo" content="drdanholmes/Affiliations" /><meta name="octolytics-dimension-repository_public" content="true" /><meta name="octolytics-dimension-repository_is_fork" content="false" /><meta name="octolytics-dimension-repository_network_root_id" content="204400584" /><meta name="octolytics-dimension-repository_network_root_nwo" content="drdanholmes/Affiliations" /><meta name="octolytics-dimension-repository_explore_github_marketplace_ci_cta_shown" content="false" />

    <link rel="canonical" href="https://github.com/drdanholmes/Affiliations/blob/master/Extras/mytemplate.tex" data-pjax-transient>

  <meta name="browser-stats-url" content="https://api.github.com/_private/browser/stats">

  <meta name="browser-errors-url" content="https://api.github.com/_private/browser/errors">

  <link rel="mask-icon" href="https://github.githubassets.com/pinned-octocat.svg" color="#000000">
  <link rel="icon" type="image/x-icon" class="js-site-favicon" href="https://github.githubassets.com/favicon.ico">

<meta name="theme-color" content="#1e2327">

  <link rel="manifest" href="/manifest.json" crossOrigin="use-credentials">

  </head>

  <body class="logged-in env-production page-responsive page-blob">

  <div class="position-relative js-header-wrapper ">
    <a href="#start-of-content" tabindex="1" class="p-3 bg-blue text-white show-on-focus js-skip-to-content">Skip to content</a>
    <span class="Progress progress-pjax-loader position-fixed width-full js-pjax-loader-bar">
      <span class="progress-pjax-loader-bar top-0 left-0" style="width: 0%;"></span>
    </span>

          <header class="Header js-details-container Details flex-wrap flex-lg-nowrap p-responsive" role="banner">

    <div class="Header-item d-none d-lg-flex">
      <a class="Header-link" href="https://github.com/" data-hotkey="g d" aria-label="Homepage" data-ga-click="Header, go to dashboard, icon:logo">
  <svg class="octicon octicon-mark-github v-align-middle" height="32" viewBox="0 0 16 16" version="1.1" width="32" aria-hidden="true"><path fill-rule="evenodd" d="M8 0C3.58 0 0 3.58 0 8c0 3.54 2.29 6.53 5.47 7.59.4.07.55-.17.55-.38 0-.19-.01-.82-.01-1.49-2.01.37-2.53-.49-2.69-.94-.09-.23-.48-.94-.82-1.13-.28-.15-.68-.52-.01-.53.63-.01 1.08.58 1.23.82.72 1.21 1.87.87 2.33.66.07-.52.28-.87.51-1.07-1.78-.2-3.64-.89-3.64-3.95 0-.87.31-1.59.82-2.15-.08-.2-.36-1.02.08-2.12 0 0 .67-.21 2.2.82.64-.18 1.32-.27 2-.27.68 0 1.36.09 2 .27 1.53-1.04 2.2-.82 2.2-.82.44 1.1.16 1.92.08 2.12.51.56.82 1.27.82 2.15 0 3.07-1.87 3.75-3.65 3.95.29.25.54.73.54 1.48 0 1.07-.01 1.93-.01 2.2 0 .21.15.46.55.38A8.013 8.013 0 0016 8c0-4.42-3.58-8-8-8z"/></svg>
</a>

    </div>

    <div class="Header-item d-lg-none">
      <button class="Header-link btn-link js-details-target" type="button" aria-label="Toggle navigation" aria-expanded="false">
        <svg height="24" class="octicon octicon-three-bars" viewBox="0 0 12 16" version="1.1" width="18" aria-hidden="true"><path fill-rule="evenodd" d="M11.41 9H.59C0 9 0 8.59 0 8c0-.59 0-1 .59-1H11.4c.59 0 .59.41.59 1 0 .59 0 1-.59 1h.01zm0-4H.59C0 5 0 4.59 0 4c0-.59 0-1 .59-1H11.4c.59 0 .59.41.59 1 0 .59 0 1-.59 1h.01zM.59 11H11.4c.59 0 .59.41.59 1 0 .59 0 1-.59 1H.59C0 13 0 12.59 0 12c0-.59 0-1 .59-1z"/></svg>
      </button>
    </div>

    <div class="Header-item Header-item--full flex-column flex-lg-row width-full flex-order-2 flex-lg-order-none mr-0 mr-lg-3 mt-3 mt-lg-0 Details-content--hidden">
        <div class="header-search flex-self-stretch flex-lg-self-auto mr-0 mr-lg-3 mb-3 mb-lg-0 scoped-search site-scoped-search js-site-search position-relative js-jump-to"
  role="combobox"
  aria-owns="jump-to-results"
  aria-label="Search or jump to"
  aria-haspopup="listbox"
  aria-expanded="false"
>
  <div class="position-relative">
    <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="js-site-search-form" role="search" aria-label="Site" data-scope-type="Repository" data-scope-id="204400584" data-scoped-search-url="/drdanholmes/Affiliations/search" data-unscoped-search-url="/search" action="/drdanholmes/Affiliations/search" accept-charset="UTF-8" method="get"><input name="utf8" type="hidden" value="&#x2713;" />
      <label class="form-control input-sm header-search-wrapper p-0 header-search-wrapper-jump-to position-relative d-flex flex-justify-between flex-items-center js-chromeless-input-container">
        <input type="text"
          class="form-control input-sm header-search-input jump-to-field js-jump-to-field js-site-search-focus js-site-search-field is-clearable"
          data-hotkey="s,/"
          name="q"
          value=""
          placeholder="Search or jump to…"
          data-unscoped-placeholder="Search or jump to…"
          data-scoped-placeholder="Search or jump to…"
          autocapitalize="off"
          aria-autocomplete="list"
          aria-controls="jump-to-results"
          aria-label="Search or jump to…"
          data-jump-to-suggestions-path="/_graphql/GetSuggestedNavigationDestinations"
          spellcheck="false"
          autocomplete="off"
          >
          <input type="hidden" value="suScuHZvj2N/u6CAxbC0wb+EUdEuVYKiKxq7xDFgRXfFQ9zWp4Zp/ADf/ROVSQFtG2+GUKEkmaJ8xdkpUryDFA==" data-csrf="true" class="js-data-jump-to-suggestions-path-csrf" />
          <input type="hidden" class="js-site-search-type-field" name="type" >
            <img src="https://github.githubassets.com/images/search-key-slash.svg" alt="" class="mr-2 header-search-key-slash">

            <div class="Box position-absolute overflow-hidden d-none jump-to-suggestions js-jump-to-suggestions-container">

<ul class="d-none js-jump-to-suggestions-template-container">

<li class="d-flex flex-justify-start flex-items-center p-0 f5 navigation-item js-navigation-item js-jump-to-suggestion" role="option">
  <a tabindex="-1" class="no-underline d-flex flex-auto flex-items-center jump-to-suggestions-path js-jump-to-suggestion-path js-navigation-open p-2" href="">
    <div class="jump-to-octicon js-jump-to-octicon flex-shrink-0 mr-2 text-center d-none">
      <svg height="16" width="16" class="octicon octicon-repo flex-shrink-0 js-jump-to-octicon-repo d-none" title="Repository" aria-label="Repository" viewBox="0 0 12 16" version="1.1" role="img"><path fill-rule="evenodd" d="M4 9H3V8h1v1zm0-3H3v1h1V6zm0-2H3v1h1V4zm0-2H3v1h1V2zm8-1v12c0 .55-.45 1-1 1H6v2l-1.5-1.5L3 16v-2H1c-.55 0-1-.45-1-1V1c0-.55.45-1 1-1h10c.55 0 1 .45 1 1zm-1 10H1v2h2v-1h3v1h5v-2zm0-10H2v9h9V1z"/></svg>
      <svg height="16" width="16" class="octicon octicon-project flex-shrink-0 js-jump-to-octicon-project d-none" title="Project" aria-label="Project" viewBox="0 0 15 16" version="1.1" role="img"><path fill-rule="evenodd" d="M10 12h3V2h-3v10zm-4-2h3V2H6v8zm-4 4h3V2H2v12zm-1 1h13V1H1v14zM14 0H1a1 1 0 00-1 1v14a1 1 0 001 1h13a1 1 0 001-1V1a1 1 0 00-1-1z"/></svg>
      <svg height="16" width="16" class="octicon octicon-search flex-shrink-0 js-jump-to-octicon-search d-none" title="Search" aria-label="Search" viewBox="0 0 16 16" version="1.1" role="img"><path fill-rule="evenodd" d="M15.7 13.3l-3.81-3.83A5.93 5.93 0 0013 6c0-3.31-2.69-6-6-6S1 2.69 1 6s2.69 6 6 6c1.3 0 2.48-.41 3.47-1.11l3.83 3.81c.19.2.45.3.7.3.25 0 .52-.09.7-.3a.996.996 0 000-1.41v.01zM7 10.7c-2.59 0-4.7-2.11-4.7-4.7 0-2.59 2.11-4.7 4.7-4.7 2.59 0 4.7 2.11 4.7 4.7 0 2.59-2.11 4.7-4.7 4.7z"/></svg>
    </div>

    <img class="avatar mr-2 flex-shrink-0 js-jump-to-suggestion-avatar d-none" alt="" aria-label="Team" src="" width="28" height="28">

    <div class="jump-to-suggestion-name js-jump-to-suggestion-name flex-auto overflow-hidden text-left no-wrap css-truncate css-truncate-target">
    </div>

    <div class="border rounded-1 flex-shrink-0 bg-gray px-1 text-gray-light ml-1 f6 d-none js-jump-to-badge-search">
      <span class="js-jump-to-badge-search-text-default d-none" aria-label="in this repository">
        In this repository
      </span>
      <span class="js-jump-to-badge-search-text-global d-none" aria-label="in all of GitHub">
        All GitHub
      </span>
      <span aria-hidden="true" class="d-inline-block ml-1 v-align-middle">↵</span>
    </div>

    <div aria-hidden="true" class="border rounded-1 flex-shrink-0 bg-gray px-1 text-gray-light ml-1 f6 d-none d-on-nav-focus js-jump-to-badge-jump">
      Jump to
      <span class="d-inline-block ml-1 v-align-middle">↵</span>
    </div>
  </a>
</li>

</ul>

<ul class="d-none js-jump-to-no-results-template-container">
  <li class="d-flex flex-justify-center flex-items-center f5 d-none js-jump-to-suggestion p-2">
    <span class="text-gray">No suggested jump to results</span>
  </li>
</ul>

<ul id="jump-to-results" role="listbox" class="p-0 m-0 js-navigation-container jump-to-suggestions-results-container js-jump-to-suggestions-results-container">

<li class="d-flex flex-justify-start flex-items-center p-0 f5 navigation-item js-navigation-item js-jump-to-scoped-search d-none" role="option">
  <a tabindex="-1" class="no-underline d-flex flex-auto flex-items-center jump-to-suggestions-path js-jump-to-suggestion-path js-navigation-open p-2" href="">
    <div class="jump-to-octicon js-jump-to-octicon flex-shrink-0 mr-2 text-center d-none">
      <svg height="16" width="16" class="octicon octicon-repo flex-shrink-0 js-jump-to-octicon-repo d-none" title="Repository" aria-label="Repository" viewBox="0 0 12 16" version="1.1" role="img"><path fill-rule="evenodd" d="M4 9H3V8h1v1zm0-3H3v1h1V6zm0-2H3v1h1V4zm0-2H3v1h1V2zm8-1v12c0 .55-.45 1-1 1H6v2l-1.5-1.5L3 16v-2H1c-.55 0-1-.45-1-1V1c0-.55.45-1 1-1h10c.55 0 1 .45 1 1zm-1 10H1v2h2v-1h3v1h5v-2zm0-10H2v9h9V1z"/></svg>
      <svg height="16" width="16" class="octicon octicon-project flex-shrink-0 js-jump-to-octicon-project d-none" title="Project" aria-label="Project" viewBox="0 0 15 16" version="1.1" role="img"><path fill-rule="evenodd" d="M10 12h3V2h-3v10zm-4-2h3V2H6v8zm-4 4h3V2H2v12zm-1 1h13V1H1v14zM14 0H1a1 1 0 00-1 1v14a1 1 0 001 1h13a1 1 0 001-1V1a1 1 0 00-1-1z"/></svg>
      <svg height="16" width="16" class="octicon octicon-search flex-shrink-0 js-jump-to-octicon-search d-none" title="Search" aria-label="Search" viewBox="0 0 16 16" version="1.1" role="img"><path fill-rule="evenodd" d="M15.7 13.3l-3.81-3.83A5.93 5.93 0 0013 6c0-3.31-2.69-6-6-6S1 2.69 1 6s2.69 6 6 6c1.3 0 2.48-.41 3.47-1.11l3.83 3.81c.19.2.45.3.7.3.25 0 .52-.09.7-.3a.996.996 0 000-1.41v.01zM7 10.7c-2.59 0-4.7-2.11-4.7-4.7 0-2.59 2.11-4.7 4.7-4.7 2.59 0 4.7 2.11 4.7 4.7 0 2.59-2.11 4.7-4.7 4.7z"/></svg>
    </div>

    <img class="avatar mr-2 flex-shrink-0 js-jump-to-suggestion-avatar d-none" alt="" aria-label="Team" src="" width="28" height="28">

    <div class="jump-to-suggestion-name js-jump-to-suggestion-name flex-auto overflow-hidden text-left no-wrap css-truncate css-truncate-target">
    </div>

    <div class="border rounded-1 flex-shrink-0 bg-gray px-1 text-gray-light ml-1 f6 d-none js-jump-to-badge-search">
      <span class="js-jump-to-badge-search-text-default d-none" aria-label="in this repository">
        In this repository
      </span>
      <span class="js-jump-to-badge-search-text-global d-none" aria-label="in all of GitHub">
        All GitHub
      </span>
      <span aria-hidden="true" class="d-inline-block ml-1 v-align-middle">↵</span>
    </div>

    <div aria-hidden="true" class="border rounded-1 flex-shrink-0 bg-gray px-1 text-gray-light ml-1 f6 d-none d-on-nav-focus js-jump-to-badge-jump">
      Jump to
      <span class="d-inline-block ml-1 v-align-middle">↵</span>
    </div>
  </a>
</li>

<li class="d-flex flex-justify-start flex-items-center p-0 f5 navigation-item js-navigation-item js-jump-to-global-search d-none" role="option">
  <a tabindex="-1" class="no-underline d-flex flex-auto flex-items-center jump-to-suggestions-path js-jump-to-suggestion-path js-navigation-open p-2" href="">
    <div class="jump-to-octicon js-jump-to-octicon flex-shrink-0 mr-2 text-center d-none">
      <svg height="16" width="16" class="octicon octicon-repo flex-shrink-0 js-jump-to-octicon-repo d-none" title="Repository" aria-label="Repository" viewBox="0 0 12 16" version="1.1" role="img"><path fill-rule="evenodd" d="M4 9H3V8h1v1zm0-3H3v1h1V6zm0-2H3v1h1V4zm0-2H3v1h1V2zm8-1v12c0 .55-.45 1-1 1H6v2l-1.5-1.5L3 16v-2H1c-.55 0-1-.45-1-1V1c0-.55.45-1 1-1h10c.55 0 1 .45 1 1zm-1 10H1v2h2v-1h3v1h5v-2zm0-10H2v9h9V1z"/></svg>
      <svg height="16" width="16" class="octicon octicon-project flex-shrink-0 js-jump-to-octicon-project d-none" title="Project" aria-label="Project" viewBox="0 0 15 16" version="1.1" role="img"><path fill-rule="evenodd" d="M10 12h3V2h-3v10zm-4-2h3V2H6v8zm-4 4h3V2H2v12zm-1 1h13V1H1v14zM14 0H1a1 1 0 00-1 1v14a1 1 0 001 1h13a1 1 0 001-1V1a1 1 0 00-1-1z"/></svg>
      <svg height="16" width="16" class="octicon octicon-search flex-shrink-0 js-jump-to-octicon-search d-none" title="Search" aria-label="Search" viewBox="0 0 16 16" version="1.1" role="img"><path fill-rule="evenodd" d="M15.7 13.3l-3.81-3.83A5.93 5.93 0 0013 6c0-3.31-2.69-6-6-6S1 2.69 1 6s2.69 6 6 6c1.3 0 2.48-.41 3.47-1.11l3.83 3.81c.19.2.45.3.7.3.25 0 .52-.09.7-.3a.996.996 0 000-1.41v.01zM7 10.7c-2.59 0-4.7-2.11-4.7-4.7 0-2.59 2.11-4.7 4.7-4.7 2.59 0 4.7 2.11 4.7 4.7 0 2.59-2.11 4.7-4.7 4.7z"/></svg>
    </div>

    <img class="avatar mr-2 flex-shrink-0 js-jump-to-suggestion-avatar d-none" alt="" aria-label="Team" src="" width="28" height="28">

    <div class="jump-to-suggestion-name js-jump-to-suggestion-name flex-auto overflow-hidden text-left no-wrap css-truncate css-truncate-target">
    </div>

    <div class="border rounded-1 flex-shrink-0 bg-gray px-1 text-gray-light ml-1 f6 d-none js-jump-to-badge-search">
      <span class="js-jump-to-badge-search-text-default d-none" aria-label="in this repository">
        In this repository
      </span>
      <span class="js-jump-to-badge-search-text-global d-none" aria-label="in all of GitHub">
        All GitHub
      </span>
      <span aria-hidden="true" class="d-inline-block ml-1 v-align-middle">↵</span>
    </div>

    <div aria-hidden="true" class="border rounded-1 flex-shrink-0 bg-gray px-1 text-gray-light ml-1 f6 d-none d-on-nav-focus js-jump-to-badge-jump">
      Jump to
      <span class="d-inline-block ml-1 v-align-middle">↵</span>
    </div>
  </a>
</li>

    <li class="d-flex flex-justify-center flex-items-center p-0 f5 js-jump-to-suggestion">
      <img src="https://github.githubassets.com/images/spinners/octocat-spinner-128.gif" alt="Octocat Spinner Icon" class="m-2" width="28">
    </li>
</ul>

            </div>
      </label>
</form>  </div>
</div>

      <nav class="d-flex flex-column flex-lg-row flex-self-stretch flex-lg-self-auto" aria-label="Global">
    <a class="Header-link d-block d-lg-none py-2 py-lg-0 border-top border-lg-top-0 border-white-fade-15" data-ga-click="Header, click, Nav menu - item:dashboard:user" aria-label="Dashboard" href="/dashboard">
      Dashboard
</a>

  <a class="js-selected-navigation-item Header-link  mr-0 mr-lg-3 py-2 py-lg-0 border-top border-lg-top-0 border-white-fade-15" data-hotkey="g p" data-ga-click="Header, click, Nav menu - item:pulls context:user" aria-label="Pull requests you created" data-selected-links="/pulls /pulls/assigned /pulls/mentioned /pulls" href="/pulls">
    Pull requests
</a>
  <a class="js-selected-navigation-item Header-link  mr-0 mr-lg-3 py-2 py-lg-0 border-top border-lg-top-0 border-white-fade-15" data-hotkey="g i" data-ga-click="Header, click, Nav menu - item:issues context:user" aria-label="Issues you created" data-selected-links="/issues /issues/assigned /issues/mentioned /issues" href="/issues">
    Issues
</a>
    <div class="mr-0 mr-lg-3 py-2 py-lg-0 border-top border-lg-top-0 border-white-fade-15">
      <a class="js-selected-navigation-item Header-link" data-ga-click="Header, click, Nav menu - item:marketplace context:user" data-octo-click="marketplace_click" data-octo-dimensions="location:nav_bar" data-selected-links=" /marketplace" href="/marketplace">
        Marketplace
</a>

    </div>

  <a class="js-selected-navigation-item Header-link  mr-0 mr-lg-3 py-2 py-lg-0 border-top border-lg-top-0 border-white-fade-15" data-ga-click="Header, click, Nav menu - item:explore" data-selected-links="/explore /trending /trending/developers /integrations /integrations/feature/code /integrations/feature/collaborate /integrations/feature/ship showcases showcases_search showcases_landing /explore" href="/explore">
    Explore
</a>

    <a class="Header-link d-block d-lg-none mr-0 mr-lg-3 py-2 py-lg-0 border-top border-lg-top-0 border-white-fade-15" href="https://github.com/chguiterman">
      <img class="avatar" height="20" width="20" alt="@chguiterman" src="https://avatars1.githubusercontent.com/u/15913450?s=60&amp;v=4" />
      chguiterman
</a>
    <!-- '"` --><!-- </textarea></xmp> --></option></form><form action="/logout" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="3dQsKJCOR7gJ6HzhNTYpou4zwOWDvS/d48BUc6qFHs9xqfQyoiqYJbzfgzN6oEM6Pht/9373XN2kzMiBnG9DVg==" />
      <button type="submit" class="Header-link mr-0 mr-lg-3 py-2 py-lg-0 border-top border-lg-top-0 border-white-fade-15 d-lg-none btn-link d-block width-full text-left" data-ga-click="Header, sign out, icon:logout" style="padding-left: 2px;">
        <svg class="octicon octicon-sign-out v-align-middle" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M12 9V7H8V5h4V3l4 3-4 3zm-2 3H6V3L2 1h8v3h1V1c0-.55-.45-1-1-1H1C.45 0 0 .45 0 1v11.38c0 .39.22.73.55.91L6 16.01V13h4c.55 0 1-.45 1-1V8h-1v4z"/></svg>
        Sign out
      </button>
</form></nav>

    </div>

    <div class="Header-item Header-item--full flex-justify-center d-lg-none position-relative">
      <div class="css-truncate css-truncate-target width-fit position-absolute left-0 right-0 text-center">
              <svg class="octicon octicon-repo" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M4 9H3V8h1v1zm0-3H3v1h1V6zm0-2H3v1h1V4zm0-2H3v1h1V2zm8-1v12c0 .55-.45 1-1 1H6v2l-1.5-1.5L3 16v-2H1c-.55 0-1-.45-1-1V1c0-.55.45-1 1-1h10c.55 0 1 .45 1 1zm-1 10H1v2h2v-1h3v1h5v-2zm0-10H2v9h9V1z"/></svg>
    <a class="Header-link" href="/drdanholmes">drdanholmes</a>
    /
    <a class="Header-link" href="/drdanholmes/Affiliations">Affiliations</a>

</div>
    </div>

    <div class="Header-item mr-0 mr-lg-3 flex-order-1 flex-lg-order-none">

    <a aria-label="You have no unread notifications" class="Header-link notification-indicator position-relative tooltipped tooltipped-sw js-socket-channel js-notification-indicator" data-hotkey="g n" data-ga-click="Header, go to notifications, icon:read" data-channel="notification-changed:15913450" href="/notifications/beta">
        <span class="js-indicator-modifier mail-status "></span>
        <svg class="octicon octicon-bell" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M14 12v1H0v-1l.73-.58c.77-.77.81-2.55 1.19-4.42C2.69 3.23 6 2 6 2c0-.55.45-1 1-1s1 .45 1 1c0 0 3.39 1.23 4.16 5 .38 1.88.42 3.66 1.19 4.42l.66.58H14zm-7 4c1.11 0 2-.89 2-2H5c0 1.11.89 2 2 2z"/></svg>
</a>
    </div>

    <div class="Header-item position-relative d-none d-lg-flex">
      <details class="details-overlay details-reset">
  <summary class="Header-link"
      aria-label="Create new…"
      data-ga-click="Header, create new, icon:add">
    <svg class="octicon octicon-plus" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M12 9H7v5H5V9H0V7h5V2h2v5h5v2z"/></svg> <span class="dropdown-caret"></span>
  </summary>
  <details-menu class="dropdown-menu dropdown-menu-sw">

<a role="menuitem" class="dropdown-item" href="/new" data-ga-click="Header, create new repository">
  New repository
</a>

  <a role="menuitem" class="dropdown-item" href="/new/import" data-ga-click="Header, import a repository">
    Import repository
  </a>

<a role="menuitem" class="dropdown-item" href="https://gist.github.com/" data-ga-click="Header, create new gist">
  New gist
</a>

  <a role="menuitem" class="dropdown-item" href="/organizations/new" data-ga-click="Header, create new organization">
    New organization
  </a>

  <div role="none" class="dropdown-divider"></div>
  <div class="dropdown-header">
    <span title="drdanholmes/Affiliations">This repository</span>
  </div>
    <a role="menuitem" class="dropdown-item" href="/drdanholmes/Affiliations/issues/new" data-ga-click="Header, create new issue" data-skip-pjax>
      New issue
    </a>

  </details-menu>
</details>

    </div>

    <div class="Header-item position-relative mr-0 d-none d-lg-flex">

  <details class="details-overlay details-reset js-feature-preview-indicator-container" data-feature-preview-indicator-src="/users/chguiterman/feature_preview/indicator_check.json">

  <summary class="Header-link"
    aria-label="View profile and more"
    data-ga-click="Header, show menu, icon:avatar">
    <img class="avatar " alt="@chguiterman" width="20" height="20" src="https://avatars1.githubusercontent.com/u/15913450?s=60&amp;v=4">

      <span class="feature-preview-indicator js-feature-preview-indicator" hidden></span>
    <span class="dropdown-caret"></span>
  </summary>
  <details-menu class="dropdown-menu dropdown-menu-sw mt-2" style="width: 180px">
    <div class="header-nav-current-user css-truncate"><a role="menuitem" class="no-underline user-profile-link px-3 pt-2 pb-2 mb-n2 mt-n1 d-block" href="/chguiterman" data-ga-click="Header, go to profile, text:Signed in as">Signed in as <strong class="css-truncate-target">chguiterman</strong></a></div>
    <div role="none" class="dropdown-divider"></div>

      <div class="pl-3 pr-3 f6 user-status-container js-user-status-context pb-1" data-url="/users/status?compact=1&amp;link_mentions=0&amp;truncate=1">

<div class="js-user-status-container
    user-status-compact rounded-1 px-2 py-1 mt-2
    border
  " data-team-hovercards-enabled>
  <details class="js-user-status-details details-reset details-overlay details-overlay-dark">
    <summary class="btn-link btn-block link-gray no-underline js-toggle-user-status-edit toggle-user-status-edit "
      role="menuitem" data-hydro-click="{&quot;event_type&quot;:&quot;user_profile.click&quot;,&quot;payload&quot;:{&quot;profile_user_id&quot;:17508391,&quot;target&quot;:&quot;EDIT_USER_STATUS&quot;,&quot;user_id&quot;:15913450,&quot;originating_url&quot;:&quot;https://github.com/drdanholmes/Affiliations/blob/master/Extras/mytemplate.tex&quot;}}" data-hydro-click-hmac="5855263bf07cab84c8ec40746be94007531f86b86d3772657f99d9b1a1f28ae1">
      <div class="d-flex">
        <div class="f6 lh-condensed user-status-header
          d-inline-block v-align-middle
            user-status-emoji-only-header circle
            pr-2
"
            style="max-width: 29px"
          >
          <div class="user-status-emoji-container flex-shrink-0 mr-1 mt-1 lh-condensed-ultra v-align-bottom" style="">
            <svg class="octicon octicon-smiley" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8 0C3.58 0 0 3.58 0 8s3.58 8 8 8 8-3.58 8-8-3.58-8-8-8zm4.81 12.81a6.72 6.72 0 01-2.17 1.45c-.83.36-1.72.53-2.64.53-.92 0-1.81-.17-2.64-.53-.81-.34-1.55-.83-2.17-1.45a6.773 6.773 0 01-1.45-2.17A6.59 6.59 0 011.21 8c0-.92.17-1.81.53-2.64.34-.81.83-1.55 1.45-2.17.62-.62 1.36-1.11 2.17-1.45A6.59 6.59 0 018 1.21c.92 0 1.81.17 2.64.53.81.34 1.55.83 2.17 1.45.62.62 1.11 1.36 1.45 2.17.36.83.53 1.72.53 2.64 0 .92-.17 1.81-.53 2.64-.34.81-.83 1.55-1.45 2.17zM4 6.8v-.59c0-.66.53-1.19 1.2-1.19h.59c.66 0 1.19.53 1.19 1.19v.59c0 .67-.53 1.2-1.19 1.2H5.2C4.53 8 4 7.47 4 6.8zm5 0v-.59c0-.66.53-1.19 1.2-1.19h.59c.66 0 1.19.53 1.19 1.19v.59c0 .67-.53 1.2-1.19 1.2h-.59C9.53 8 9 7.47 9 6.8zm4 3.2c-.72 1.88-2.91 3-5 3s-4.28-1.13-5-3c-.14-.39.23-1 .66-1h8.59c.41 0 .89.61.75 1z"/></svg>
          </div>
        </div>
        <div class="
          d-inline-block v-align-middle

           css-truncate css-truncate-target 
           user-status-message-wrapper f6"
           style="line-height: 20px;" >
          <div class="d-inline-block text-gray-dark v-align-text-top text-left">
              <span class="text-gray ml-2">Set status</span>
          </div>
        </div>
      </div>
    </summary>
    <details-dialog class="details-dialog rounded-1 anim-fade-in fast Box Box--overlay" role="dialog" tabindex="-1">
      <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="position-relative flex-auto js-user-status-form" action="/users/status?compact=1&amp;link_mentions=0&amp;truncate=1" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="_method" value="put" /><input type="hidden" name="authenticity_token" value="i2/QNvK7qRXY4wv8i6SpXu5fLU9YrO0nPWxKA7R03qnJLW9vVXerq3BF/z1HykGMjB5K+2ss69K79Q0KTu/ihA==" />
        <div class="Box-header bg-gray border-bottom p-3">
          <button class="Box-btn-octicon js-toggle-user-status-edit btn-octicon float-right" type="reset" aria-label="Close dialog" data-close-dialog>
            <svg class="octicon octicon-x" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48L7.48 8z"/></svg>
          </button>
          <h3 class="Box-title f5 text-bold text-gray-dark">Edit status</h3>
        </div>
        <input type="hidden" name="emoji" class="js-user-status-emoji-field" value="">
        <input type="hidden" name="organization_id" class="js-user-status-org-id-field" value="">
        <div class="px-3 py-2 text-gray-dark">
          <div class="js-characters-remaining-container position-relative mt-2">
            <div class="input-group d-table form-group my-0 js-user-status-form-group">
              <span class="input-group-button d-table-cell v-align-middle" style="width: 1%">
                <button type="button" aria-label="Choose an emoji" class="btn-outline btn js-toggle-user-status-emoji-picker btn-open-emoji-picker p-0">
                  <span class="js-user-status-original-emoji" hidden></span>
                  <span class="js-user-status-custom-emoji"></span>
                  <span class="js-user-status-no-emoji-icon" >
                    <svg class="octicon octicon-smiley" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8 0C3.58 0 0 3.58 0 8s3.58 8 8 8 8-3.58 8-8-3.58-8-8-8zm4.81 12.81a6.72 6.72 0 01-2.17 1.45c-.83.36-1.72.53-2.64.53-.92 0-1.81-.17-2.64-.53-.81-.34-1.55-.83-2.17-1.45a6.773 6.773 0 01-1.45-2.17A6.59 6.59 0 011.21 8c0-.92.17-1.81.53-2.64.34-.81.83-1.55 1.45-2.17.62-.62 1.36-1.11 2.17-1.45A6.59 6.59 0 018 1.21c.92 0 1.81.17 2.64.53.81.34 1.55.83 2.17 1.45.62.62 1.11 1.36 1.45 2.17.36.83.53 1.72.53 2.64 0 .92-.17 1.81-.53 2.64-.34.81-.83 1.55-1.45 2.17zM4 6.8v-.59c0-.66.53-1.19 1.2-1.19h.59c.66 0 1.19.53 1.19 1.19v.59c0 .67-.53 1.2-1.19 1.2H5.2C4.53 8 4 7.47 4 6.8zm5 0v-.59c0-.66.53-1.19 1.2-1.19h.59c.66 0 1.19.53 1.19 1.19v.59c0 .67-.53 1.2-1.19 1.2h-.59C9.53 8 9 7.47 9 6.8zm4 3.2c-.72 1.88-2.91 3-5 3s-4.28-1.13-5-3c-.14-.39.23-1 .66-1h8.59c.41 0 .89.61.75 1z"/></svg>
                  </span>
                </button>
              </span>
              <text-expander keys=": @" data-mention-url="/autocomplete/user-suggestions" data-emoji-url="/autocomplete/emoji">
                <input
                  type="text"
                  autocomplete="off"
                  data-no-org-url="/autocomplete/user-suggestions"
                  data-org-url="/suggestions?mention_suggester=1"
                  data-maxlength="80"
                  class="d-table-cell width-full form-control js-user-status-message-field js-characters-remaining-field"
                  placeholder="What's happening?"
                  name="message"
                  value=""
                  aria-label="What is your current status?">
              </text-expander>
              <div class="error">Could not update your status, please try again.</div>
            </div>
            <div style="margin-left: 53px" class="my-1 text-small label-characters-remaining js-characters-remaining" data-suffix="remaining" hidden>
              80 remaining
            </div>
          </div>
          <include-fragment class="js-user-status-emoji-picker" data-url="/users/status/emoji"></include-fragment>
          <div class="overflow-auto ml-n3 mr-n3 px-3 border-bottom" style="max-height: 33vh">
            <div class="user-status-suggestions js-user-status-suggestions collapsed overflow-hidden">
              <h4 class="f6 text-normal my-3">Suggestions:</h4>
              <div class="mx-3 mt-2 clearfix">
                  <div class="float-left col-6">
                      <button type="button" value=":palm_tree:" class="d-flex flex-items-baseline flex-items-stretch lh-condensed f6 btn-link link-gray no-underline js-predefined-user-status mb-1">
                        <div class="emoji-status-width mr-2 v-align-middle js-predefined-user-status-emoji">
                          <g-emoji alias="palm_tree" fallback-src="https://github.githubassets.com/images/icons/emoji/unicode/1f334.png">🌴</g-emoji>
                        </div>
                        <div class="d-flex flex-items-center no-underline js-predefined-user-status-message ws-normal text-left" style="border-left: 1px solid transparent">
                          On vacation
                        </div>
                      </button>
                      <button type="button" value=":face_with_thermometer:" class="d-flex flex-items-baseline flex-items-stretch lh-condensed f6 btn-link link-gray no-underline js-predefined-user-status mb-1">
                        <div class="emoji-status-width mr-2 v-align-middle js-predefined-user-status-emoji">
                          <g-emoji alias="face_with_thermometer" fallback-src="https://github.githubassets.com/images/icons/emoji/unicode/1f912.png">🤒</g-emoji>
                        </div>
                        <div class="d-flex flex-items-center no-underline js-predefined-user-status-message ws-normal text-left" style="border-left: 1px solid transparent">
                          Out sick
                        </div>
                      </button>
                  </div>
                  <div class="float-left col-6">
                      <button type="button" value=":house:" class="d-flex flex-items-baseline flex-items-stretch lh-condensed f6 btn-link link-gray no-underline js-predefined-user-status mb-1">
                        <div class="emoji-status-width mr-2 v-align-middle js-predefined-user-status-emoji">
                          <g-emoji alias="house" fallback-src="https://github.githubassets.com/images/icons/emoji/unicode/1f3e0.png">🏠</g-emoji>
                        </div>
                        <div class="d-flex flex-items-center no-underline js-predefined-user-status-message ws-normal text-left" style="border-left: 1px solid transparent">
                          Working from home
                        </div>
                      </button>
                      <button type="button" value=":dart:" class="d-flex flex-items-baseline flex-items-stretch lh-condensed f6 btn-link link-gray no-underline js-predefined-user-status mb-1">
                        <div class="emoji-status-width mr-2 v-align-middle js-predefined-user-status-emoji">
                          <g-emoji alias="dart" fallback-src="https://github.githubassets.com/images/icons/emoji/unicode/1f3af.png">🎯</g-emoji>
                        </div>
                        <div class="d-flex flex-items-center no-underline js-predefined-user-status-message ws-normal text-left" style="border-left: 1px solid transparent">
                          Focusing
                        </div>
                      </button>
                  </div>
              </div>
            </div>
            <div class="user-status-limited-availability-container">
              <div class="form-checkbox my-0">
                <input type="checkbox" name="limited_availability" value="1" class="js-user-status-limited-availability-checkbox" data-default-message="I may be slow to respond." aria-describedby="limited-availability-help-text-truncate-true-compact-true" id="limited-availability-truncate-true-compact-true">
                <label class="d-block f5 text-gray-dark mb-1" for="limited-availability-truncate-true-compact-true">
                  Busy
                </label>
                <p class="note" id="limited-availability-help-text-truncate-true-compact-true">
                  When others mention you, assign you, or request your review,
                  GitHub will let them know that you have limited availability.
                </p>
              </div>
            </div>
          </div>
          <div class="d-inline-block f5 mr-2 pt-3 pb-2" >
  <div class="d-inline-block mr-1">
    Clear status
  </div>

  <details class="js-user-status-expire-drop-down f6 dropdown details-reset details-overlay d-inline-block mr-2">
    <summary class="f5 btn-link link-gray-dark border px-2 py-1 rounded-1" aria-haspopup="true">
      <div class="js-user-status-expiration-interval-selected d-inline-block v-align-baseline">
        Never
      </div>
      <div class="dropdown-caret"></div>
    </summary>

    <ul class="dropdown-menu dropdown-menu-se pl-0 overflow-auto" style="width: 220px; max-height: 15.5em">
      <li>
        <button type="button" class="btn-link dropdown-item js-user-status-expire-button ws-normal" title="Never">
          <span class="d-inline-block text-bold mb-1">Never</span>
          <div class="f6 lh-condensed">Keep this status until you clear your status or edit your status.</div>
        </button>
      </li>
      <li class="dropdown-divider" role="none"></li>
        <li>
          <button type="button" class="btn-link dropdown-item ws-normal js-user-status-expire-button" title="in 30 minutes" value="2020-02-26T11:08:08-05:00">
            in 30 minutes
          </button>
        </li>
        <li>
          <button type="button" class="btn-link dropdown-item ws-normal js-user-status-expire-button" title="in 1 hour" value="2020-02-26T11:38:08-05:00">
            in 1 hour
          </button>
        </li>
        <li>
          <button type="button" class="btn-link dropdown-item ws-normal js-user-status-expire-button" title="in 4 hours" value="2020-02-26T14:38:08-05:00">
            in 4 hours
          </button>
        </li>
        <li>
          <button type="button" class="btn-link dropdown-item ws-normal js-user-status-expire-button" title="today" value="2020-02-26T23:59:59-05:00">
            today
          </button>
        </li>
        <li>
          <button type="button" class="btn-link dropdown-item ws-normal js-user-status-expire-button" title="this week" value="2020-03-01T23:59:59-05:00">
            this week
          </button>
        </li>
    </ul>
  </details>
  <input class="js-user-status-expiration-date-input" type="hidden" name="expires_at" value="">
</div>

          <include-fragment class="js-user-status-org-picker" data-url="/users/status/organizations"></include-fragment>
        </div>
        <div class="d-flex flex-items-center flex-justify-between p-3 border-top">
          <button type="submit" disabled class="width-full btn btn-primary mr-2 js-user-status-submit">
            Set status
          </button>
          <button type="button" disabled class="width-full js-clear-user-status-button btn ml-2 ">
            Clear status
          </button>
        </div>
</form>    </details-dialog>
  </details>
</div>

      </div>
      <div role="none" class="dropdown-divider"></div>

    <a role="menuitem" class="dropdown-item" href="/chguiterman" data-ga-click="Header, go to profile, text:your profile">Your profile</a>

    <a role="menuitem" class="dropdown-item" href="/chguiterman?tab=repositories" data-ga-click="Header, go to repositories, text:your repositories">Your repositories</a>

    <a role="menuitem" class="dropdown-item" href="/chguiterman?tab=projects" data-ga-click="Header, go to projects, text:your projects">Your projects</a>

    <a role="menuitem" class="dropdown-item" href="/chguiterman?tab=stars" data-ga-click="Header, go to starred repos, text:your stars">Your stars</a>
      <a role="menuitem" class="dropdown-item" href="https://gist.github.com/mine" data-ga-click="Header, your gists, text:your gists">Your gists</a>

    <div role="none" class="dropdown-divider"></div>

<div id="feature-enrollment-toggle" class="hide-sm hide-md feature-preview-details position-relative">
  <button
    type="button"
    class="dropdown-item btn-link"
    role="menuitem"
    data-feature-preview-trigger-url="/users/chguiterman/feature_previews"
    data-feature-preview-close-details="{&quot;event_type&quot;:&quot;feature_preview.clicks.close_modal&quot;,&quot;payload&quot;:{&quot;originating_url&quot;:&quot;https://github.com/drdanholmes/Affiliations/blob/master/Extras/mytemplate.tex&quot;,&quot;user_id&quot;:15913450}}"
    data-feature-preview-close-hmac="9fa64e36bc3cd20eda341a2346687a7f95c08290f70eb1020887882769082356"
    data-hydro-click="{&quot;event_type&quot;:&quot;feature_preview.clicks.open_modal&quot;,&quot;payload&quot;:{&quot;link_location&quot;:&quot;user_dropdown&quot;,&quot;originating_url&quot;:&quot;https://github.com/drdanholmes/Affiliations/blob/master/Extras/mytemplate.tex&quot;,&quot;user_id&quot;:15913450}}"
    data-hydro-click-hmac="b6b31416f04188b53c719adf18048ce1293deae60f2f4082bc6e6c53af165bc8"
  >
    Feature preview
  </button>
    <span class="feature-preview-indicator js-feature-preview-indicator" hidden></span>
</div>

    <a role="menuitem" class="dropdown-item" href="https://help.github.com" data-ga-click="Header, go to help, text:help">Help</a>
    <a role="menuitem" class="dropdown-item" href="/settings/profile" data-ga-click="Header, go to settings, icon:settings">Settings</a>
    <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="logout-form" action="/logout" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="iixfsjbwt2rwOfmtnKWuDgM6j5K1NJbFjqC7/iA0juwmUYeoBFRo90UOBn/TM8SW0xIwgEh+5cXJrCcMFt7TdQ==" />

      <button type="submit" class="dropdown-item dropdown-signout" data-ga-click="Header, sign out, icon:logout" role="menuitem">
        Sign out
      </button>
      <input type="text" name="required_field_db29" hidden="hidden" class="form-control" /><input type="hidden" name="timestamp" value="1582731488642" class="form-control" /><input type="hidden" name="timestamp_secret" value="5b00f4ada03dd93c418cc0be7ca882b1f9c94e2ae2e62f9c52a45d527679e2ac" class="form-control" />
</form>  </details-menu>
</details>

    </div>

  </header>

  </div>

  <div id="start-of-content" class="show-on-focus"></div>

    <div id="js-flash-container">

</div>

  <include-fragment class="js-notification-shelf-include-fragment" data-base-src="https://github.com/notifications/beta/shelf"></include-fragment>

  <div class="application-main " data-commit-hovercards-enabled>
        <div itemscope itemtype="http://schema.org/SoftwareSourceCode" class="">
    <main  >

  <div class="pagehead repohead hx_repohead readability-menu bg-gray-light pb-0 pt-0 pt-lg-3">

    <div class="container-lg mb-4 p-responsive d-none d-lg-flex">

      <div class="flex-auto min-width-0 width-fit mr-3">
        <h1 class="public  d-flex flex-wrap flex-items-center break-word float-none ">
    <svg class="octicon octicon-repo" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M4 9H3V8h1v1zm0-3H3v1h1V6zm0-2H3v1h1V4zm0-2H3v1h1V2zm8-1v12c0 .55-.45 1-1 1H6v2l-1.5-1.5L3 16v-2H1c-.55 0-1-.45-1-1V1c0-.55.45-1 1-1h10c.55 0 1 .45 1 1zm-1 10H1v2h2v-1h3v1h5v-2zm0-10H2v9h9V1z"/></svg>
  <span class="author ml-1 flex-self-stretch" itemprop="author">
    <a class="url fn" rel="author" data-hovercard-type="user" data-hovercard-url="/users/drdanholmes/hovercard" data-octo-click="hovercard-link-click" data-octo-dimensions="link_type:self" href="/drdanholmes">drdanholmes</a>
  </span>
  <span class="path-divider flex-self-stretch">/</span>
  <strong itemprop="name" class="mr-2 flex-self-stretch">
    <a data-pjax="#js-repo-pjax-container" href="/drdanholmes/Affiliations">Affiliations</a>
  </strong>

</h1>

      </div>

      <ul class="pagehead-actions flex-shrink-0"  >

  <li>

    <!-- '"` --><!-- </textarea></xmp> --></option></form><form data-remote="true" class="clearfix js-social-form js-social-container" action="/notifications/subscribe" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="u7qwDfSZLxsweTh9dTOvnyT5RsOzSC2amVaQBOXzcLcsnxKLJG+/6+fLdlaYQNlJ6ckijVAnhmDBL5WCNuQHUQ==" />      <input type="hidden" name="repository_id" value="204400584">

      <details class="details-reset details-overlay select-menu float-left">
        <summary class="select-menu-button float-left btn btn-sm btn-with-count" data-hydro-click="{&quot;event_type&quot;:&quot;repository.click&quot;,&quot;payload&quot;:{&quot;target&quot;:&quot;WATCH_BUTTON&quot;,&quot;repository_id&quot;:204400584,&quot;originating_url&quot;:&quot;https://github.com/drdanholmes/Affiliations/blob/master/Extras/mytemplate.tex&quot;,&quot;user_id&quot;:15913450}}" data-hydro-click-hmac="c0cbc47b583f68f67838302c7cea5da82ec23fb1314cd61b0467a9c4b40aa057" data-ga-click="Repository, click Watch settings, action:blob#show">          <span data-menu-button>
              <svg class="octicon octicon-eye v-align-text-bottom" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.06 2C3 2 0 8 0 8s3 6 8.06 6C13 14 16 8 16 8s-3-6-7.94-6zM8 12c-2.2 0-4-1.78-4-4 0-2.2 1.8-4 4-4 2.22 0 4 1.8 4 4 0 2.22-1.78 4-4 4zm2-4c0 1.11-.89 2-2 2-1.11 0-2-.89-2-2 0-1.11.89-2 2-2 1.11 0 2 .89 2 2z"/></svg>
              Watch
          </span>
</summary>        <details-menu
          class="select-menu-modal position-absolute mt-5"
          style="z-index: 99;">
          <div class="select-menu-header">
            <span class="select-menu-title">Notifications</span>
          </div>
          <div class="select-menu-list">
            <button type="submit" name="do" value="included" class="select-menu-item width-full" aria-checked="true" role="menuitemradio">
              <svg class="octicon octicon-check select-menu-item-icon" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M12 5l-8 8-4-4 1.5-1.5L4 10l6.5-6.5L12 5z"/></svg>
              <div class="select-menu-item-text">
                <span class="select-menu-item-heading">Not watching</span>
                <span class="description">Be notified only when participating or @mentioned.</span>
                <span class="hidden-select-button-text" data-menu-button-contents>
                  <svg class="octicon octicon-eye v-align-text-bottom" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.06 2C3 2 0 8 0 8s3 6 8.06 6C13 14 16 8 16 8s-3-6-7.94-6zM8 12c-2.2 0-4-1.78-4-4 0-2.2 1.8-4 4-4 2.22 0 4 1.8 4 4 0 2.22-1.78 4-4 4zm2-4c0 1.11-.89 2-2 2-1.11 0-2-.89-2-2 0-1.11.89-2 2-2 1.11 0 2 .89 2 2z"/></svg>
                  Watch
                </span>
              </div>
            </button>

            <button type="submit" name="do" value="release_only" class="select-menu-item width-full" aria-checked="false" role="menuitemradio">
              <svg class="octicon octicon-check select-menu-item-icon" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M12 5l-8 8-4-4 1.5-1.5L4 10l6.5-6.5L12 5z"/></svg>
              <div class="select-menu-item-text">
                <span class="select-menu-item-heading">Releases only</span>
                <span class="description">Be notified of new releases, and when participating or @mentioned.</span>
                <span class="hidden-select-button-text" data-menu-button-contents>
                  <svg class="octicon octicon-eye v-align-text-bottom" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.06 2C3 2 0 8 0 8s3 6 8.06 6C13 14 16 8 16 8s-3-6-7.94-6zM8 12c-2.2 0-4-1.78-4-4 0-2.2 1.8-4 4-4 2.22 0 4 1.8 4 4 0 2.22-1.78 4-4 4zm2-4c0 1.11-.89 2-2 2-1.11 0-2-.89-2-2 0-1.11.89-2 2-2 1.11 0 2 .89 2 2z"/></svg>
                  Unwatch releases
                </span>
              </div>
            </button>

            <button type="submit" name="do" value="subscribed" class="select-menu-item width-full" aria-checked="false" role="menuitemradio">
              <svg class="octicon octicon-check select-menu-item-icon" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M12 5l-8 8-4-4 1.5-1.5L4 10l6.5-6.5L12 5z"/></svg>
              <div class="select-menu-item-text">
                <span class="select-menu-item-heading">Watching</span>
                <span class="description">Be notified of all conversations.</span>
                <span class="hidden-select-button-text" data-menu-button-contents>
                  <svg class="octicon octicon-eye v-align-text-bottom" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.06 2C3 2 0 8 0 8s3 6 8.06 6C13 14 16 8 16 8s-3-6-7.94-6zM8 12c-2.2 0-4-1.78-4-4 0-2.2 1.8-4 4-4 2.22 0 4 1.8 4 4 0 2.22-1.78 4-4 4zm2-4c0 1.11-.89 2-2 2-1.11 0-2-.89-2-2 0-1.11.89-2 2-2 1.11 0 2 .89 2 2z"/></svg>
                  Unwatch
                </span>
              </div>
            </button>

            <button type="submit" name="do" value="ignore" class="select-menu-item width-full" aria-checked="false" role="menuitemradio">
              <svg class="octicon octicon-check select-menu-item-icon" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M12 5l-8 8-4-4 1.5-1.5L4 10l6.5-6.5L12 5z"/></svg>
              <div class="select-menu-item-text">
                <span class="select-menu-item-heading">Ignoring</span>
                <span class="description">Never be notified.</span>
                <span class="hidden-select-button-text" data-menu-button-contents>
                  <svg class="octicon octicon-mute v-align-text-bottom" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8 2.81v10.38c0 .67-.81 1-1.28.53L3 10H1c-.55 0-1-.45-1-1V7c0-.55.45-1 1-1h2l3.72-3.72C7.19 1.81 8 2.14 8 2.81zm7.53 3.22l-1.06-1.06-1.97 1.97-1.97-1.97-1.06 1.06L11.44 8 9.47 9.97l1.06 1.06 1.97-1.97 1.97 1.97 1.06-1.06L13.56 8l1.97-1.97z"/></svg>
                  Stop ignoring
                </span>
              </div>
            </button>
          </div>
        </details-menu>
      </details>
        <a class="social-count js-social-count"
          href="/drdanholmes/Affiliations/watchers"
          aria-label="1 user is watching this repository">
          1
        </a>
</form>
  </li>

  <li>
      <div class="js-toggler-container js-social-container starring-container ">
    <form class="starred js-social-form" action="/drdanholmes/Affiliations/unstar" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="AMmYUIQmdnEZ8IBXllUAaNDls0AhU3FKW+I/lXUmDkmHGIjB/XZXWvjFgYc5Fz8wJnVoqgl/q6z9jc6df+K5zQ==" />
      <input type="hidden" name="context" value="repository"></input>
      <button type="submit" class="btn btn-sm btn-with-count js-toggler-target" aria-label="Unstar this repository" title="Unstar drdanholmes/Affiliations" data-hydro-click="{&quot;event_type&quot;:&quot;repository.click&quot;,&quot;payload&quot;:{&quot;target&quot;:&quot;UNSTAR_BUTTON&quot;,&quot;repository_id&quot;:204400584,&quot;originating_url&quot;:&quot;https://github.com/drdanholmes/Affiliations/blob/master/Extras/mytemplate.tex&quot;,&quot;user_id&quot;:15913450}}" data-hydro-click-hmac="ebb4ae74084b742cab361091a82075c706bcf74992916948a6a159b8cdb5cd38" data-ga-click="Repository, click unstar button, action:blob#show; text:Unstar">        <svg aria-label="star" height="16" class="octicon octicon-star v-align-text-bottom" viewBox="0 0 14 16" version="1.1" width="14" role="img"><path fill-rule="evenodd" d="M14 6l-4.9-.64L7 1 4.9 5.36 0 6l3.6 3.26L2.67 14 7 11.67 11.33 14l-.93-4.74L14 6z"/></svg>

        Unstar
</button>        <a class="social-count js-social-count" href="/drdanholmes/Affiliations/stargazers"
           aria-label="6 users starred this repository">
           6
        </a>
</form>
    <form class="unstarred js-social-form" action="/drdanholmes/Affiliations/star" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="Y6qCZMZoFVJ9g5+f/2QCZt1s4soh3fDAIwe17aJpzK2wcRPl/dz7U4HPMgeyPL7pIp5h4p5msD7yowBktWWrlg==" />
      <input type="hidden" name="context" value="repository"></input>
      <button type="submit" class="btn btn-sm btn-with-count js-toggler-target" aria-label="Unstar this repository" title="Star drdanholmes/Affiliations" data-hydro-click="{&quot;event_type&quot;:&quot;repository.click&quot;,&quot;payload&quot;:{&quot;target&quot;:&quot;STAR_BUTTON&quot;,&quot;repository_id&quot;:204400584,&quot;originating_url&quot;:&quot;https://github.com/drdanholmes/Affiliations/blob/master/Extras/mytemplate.tex&quot;,&quot;user_id&quot;:15913450}}" data-hydro-click-hmac="8b8fd815e5494bf008a73b9c9bc2d4e5ed9c0009e2f9fccffd91421f668c88e0" data-ga-click="Repository, click star button, action:blob#show; text:Star">        <svg aria-label="star" height="16" class="octicon octicon-star v-align-text-bottom" viewBox="0 0 14 16" version="1.1" width="14" role="img"><path fill-rule="evenodd" d="M14 6l-4.9-.64L7 1 4.9 5.36 0 6l3.6 3.26L2.67 14 7 11.67 11.33 14l-.93-4.74L14 6z"/></svg>

        Star
</button>        <a class="social-count js-social-count" href="/drdanholmes/Affiliations/stargazers"
           aria-label="6 users starred this repository">
          6
        </a>
</form>  </div>

  </li>

  <li>
          <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="btn-with-count" action="/drdanholmes/Affiliations/fork" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="3BQRxBzQIcEsw9ZkbVvKIXJ7ew8E2nhcrLlzyHUu8qhZBpxLWZhbxDILQhBVTLHsMxRvh3/Cjgvmn+LFiF3D8Q==" />
            <button class="btn btn-sm btn-with-count" data-hydro-click="{&quot;event_type&quot;:&quot;repository.click&quot;,&quot;payload&quot;:{&quot;target&quot;:&quot;FORK_BUTTON&quot;,&quot;repository_id&quot;:204400584,&quot;originating_url&quot;:&quot;https://github.com/drdanholmes/Affiliations/blob/master/Extras/mytemplate.tex&quot;,&quot;user_id&quot;:15913450}}" data-hydro-click-hmac="d2c4436086174caf7a6b1a6dab3e84714751211132ae49dfd7e49fcd6ad97468" data-ga-click="Repository, show fork modal, action:blob#show; text:Fork" type="submit" title="Fork your own copy of drdanholmes/Affiliations to your account" aria-label="Fork your own copy of drdanholmes/Affiliations to your account">              <svg class="octicon octicon-repo-forked v-align-text-bottom" viewBox="0 0 10 16" version="1.1" width="10" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8 1a1.993 1.993 0 00-1 3.72V6L5 8 3 6V4.72A1.993 1.993 0 002 1a1.993 1.993 0 00-1 3.72V6.5l3 3v1.78A1.993 1.993 0 005 15a1.993 1.993 0 001-3.72V9.5l3-3V4.72A1.993 1.993 0 008 1zM2 4.2C1.34 4.2.8 3.65.8 3c0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2zm3 10c-.66 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2zm3-10c-.66 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2z"/></svg>
              Fork
</button></form>
    <a href="/drdanholmes/Affiliations/network/members" class="social-count"
       aria-label="2 users forked this repository">
      2
    </a>
  </li>
</ul>

    </div>

<nav class="hx_reponav reponav js-repo-nav js-sidenav-container-pjax clearfix container-lg p-responsive d-none d-lg-block"
     itemscope
     itemtype="http://schema.org/BreadcrumbList"
    aria-label="Repository"
     data-pjax="#js-repo-pjax-container">

  <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
    <a class="js-selected-navigation-item selected reponav-item" itemprop="url" data-hotkey="g c" aria-current="page" data-selected-links="repo_source repo_downloads repo_commits repo_releases repo_tags repo_branches repo_packages /drdanholmes/Affiliations" href="/drdanholmes/Affiliations">
      <div class="d-inline"><svg class="octicon octicon-code" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M9.5 3L8 4.5 11.5 8 8 11.5 9.5 13 14 8 9.5 3zm-5 0L0 8l4.5 5L6 11.5 2.5 8 6 4.5 4.5 3z"/></svg></div>
      <span itemprop="name">Code</span>
      <meta itemprop="position" content="1">
</a>  </span>

    <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
      <a itemprop="url" data-hotkey="g i" class="js-selected-navigation-item reponav-item" data-selected-links="repo_issues repo_labels repo_milestones /drdanholmes/Affiliations/issues" href="/drdanholmes/Affiliations/issues">
        <div class="d-inline"><svg class="octicon octicon-issue-opened" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M7 2.3c3.14 0 5.7 2.56 5.7 5.7s-2.56 5.7-5.7 5.7A5.71 5.71 0 011.3 8c0-3.14 2.56-5.7 5.7-5.7zM7 1C3.14 1 0 4.14 0 8s3.14 7 7 7 7-3.14 7-7-3.14-7-7-7zm1 3H6v5h2V4zm0 6H6v2h2v-2z"/></svg></div>
        <span itemprop="name">Issues</span>
        <span class="Counter">0</span>
        <meta itemprop="position" content="2">
</a>    </span>

  <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
    <a data-hotkey="g p" data-skip-pjax="true" itemprop="url" class="js-selected-navigation-item reponav-item" data-selected-links="repo_pulls checks /drdanholmes/Affiliations/pulls" href="/drdanholmes/Affiliations/pulls">
      <div class="d-inline"><svg class="octicon octicon-git-pull-request" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M11 11.28V5c-.03-.78-.34-1.47-.94-2.06C9.46 2.35 8.78 2.03 8 2H7V0L4 3l3 3V4h1c.27.02.48.11.69.31.21.2.3.42.31.69v6.28A1.993 1.993 0 0010 15a1.993 1.993 0 001-3.72zm-1 2.92c-.66 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2zM4 3c0-1.11-.89-2-2-2a1.993 1.993 0 00-1 3.72v6.56A1.993 1.993 0 002 15a1.993 1.993 0 001-3.72V4.72c.59-.34 1-.98 1-1.72zm-.8 10c0 .66-.55 1.2-1.2 1.2-.65 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2zM2 4.2C1.34 4.2.8 3.65.8 3c0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2z"/></svg></div>
      <span itemprop="name">Pull requests</span>
      <span class="Counter">0</span>
      <meta itemprop="position" content="4">
</a>  </span>

    <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement" class="position-relative float-left">
      <a data-hotkey="g w" data-skip-pjax="true" class="js-selected-navigation-item reponav-item" data-selected-links="repo_actions /drdanholmes/Affiliations/actions" href="/drdanholmes/Affiliations/actions">
        <div class="d-inline"><svg class="octicon octicon-play" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M14 8A7 7 0 110 8a7 7 0 0114 0zm-8.223 3.482l4.599-3.066a.5.5 0 000-.832L5.777 4.518A.5.5 0 005 4.934v6.132a.5.5 0 00.777.416z"/></svg></div>
        Actions
</a>
    </span>

    <a data-hotkey="g b" class="js-selected-navigation-item reponav-item" data-selected-links="repo_projects new_repo_project repo_project /drdanholmes/Affiliations/projects" href="/drdanholmes/Affiliations/projects">
      <div class="d-inline"><svg class="octicon octicon-project" viewBox="0 0 15 16" version="1.1" width="15" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M10 12h3V2h-3v10zm-4-2h3V2H6v8zm-4 4h3V2H2v12zm-1 1h13V1H1v14zM14 0H1a1 1 0 00-1 1v14a1 1 0 001 1h13a1 1 0 001-1V1a1 1 0 00-1-1z"/></svg></div>
      Projects
      <span class="Counter">0</span>
</a>
    <a class="js-selected-navigation-item reponav-item" data-hotkey="g w" data-selected-links="repo_wiki /drdanholmes/Affiliations/wiki" href="/drdanholmes/Affiliations/wiki">
      <div class="d-inline"><svg class="octicon octicon-book" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M3 5h4v1H3V5zm0 3h4V7H3v1zm0 2h4V9H3v1zm11-5h-4v1h4V5zm0 2h-4v1h4V7zm0 2h-4v1h4V9zm2-6v9c0 .55-.45 1-1 1H9.5l-1 1-1-1H2c-.55 0-1-.45-1-1V3c0-.55.45-1 1-1h5.5l1 1 1-1H15c.55 0 1 .45 1 1zm-8 .5L7.5 3H2v9h6V3.5zm7-.5H9.5l-.5.5V12h6V3z"/></svg></div>
      Wiki
</a>
    <a data-skip-pjax="true" class="js-selected-navigation-item reponav-item" data-selected-links="security alerts policy token_scanning code_scanning /drdanholmes/Affiliations/security/advisories" href="/drdanholmes/Affiliations/security/advisories">
      <div class="d-inline"><svg class="octicon octicon-shield" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M0 2l7-2 7 2v6.02C14 12.69 8.69 16 7 16c-1.69 0-7-3.31-7-7.98V2zm1 .75L7 1l6 1.75v5.268C13 12.104 8.449 15 7 15c-1.449 0-6-2.896-6-6.982V2.75zm1 .75L7 2v12c-1.207 0-5-2.482-5-5.985V3.5z"/></svg></div>
      Security
</a>
    <a class="js-selected-navigation-item reponav-item" data-selected-links="repo_graphs repo_contributors dependency_graph pulse people /drdanholmes/Affiliations/pulse" href="/drdanholmes/Affiliations/pulse">
      <div class="d-inline"><svg class="octicon octicon-graph" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M16 14v1H0V0h1v14h15zM5 13H3V8h2v5zm4 0H7V3h2v10zm4 0h-2V6h2v7z"/></svg></div>
      Insights
</a>

</nav>

  <div class="reponav-wrapper reponav-small d-lg-none">
  <nav class="reponav js-reponav text-center no-wrap"
       itemscope
       itemtype="http://schema.org/BreadcrumbList">

    <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
      <a class="js-selected-navigation-item selected reponav-item" itemprop="url" aria-current="page" data-selected-links="repo_source repo_downloads repo_commits repo_releases repo_tags repo_branches repo_packages /drdanholmes/Affiliations" href="/drdanholmes/Affiliations">
        <span itemprop="name">Code</span>
        <meta itemprop="position" content="1">
</a>    </span>

      <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
        <a itemprop="url" class="js-selected-navigation-item reponav-item" data-selected-links="repo_issues repo_labels repo_milestones /drdanholmes/Affiliations/issues" href="/drdanholmes/Affiliations/issues">
          <span itemprop="name">Issues</span>
          <span class="Counter">0</span>
          <meta itemprop="position" content="2">
</a>      </span>

    <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
      <a itemprop="url" class="js-selected-navigation-item reponav-item" data-selected-links="repo_pulls checks /drdanholmes/Affiliations/pulls" href="/drdanholmes/Affiliations/pulls">
        <span itemprop="name">Pull requests</span>
        <span class="Counter">0</span>
        <meta itemprop="position" content="4">
</a>    </span>

      <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
        <a itemprop="url" class="js-selected-navigation-item reponav-item" data-selected-links="repo_projects new_repo_project repo_project /drdanholmes/Affiliations/projects" href="/drdanholmes/Affiliations/projects">
          <span itemprop="name">Projects</span>
          <span class="Counter">0</span>
          <meta itemprop="position" content="5">
</a>      </span>

      <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
        <a itemprop="url" class="js-selected-navigation-item reponav-item" data-selected-links="repo_actions /drdanholmes/Affiliations/actions" href="/drdanholmes/Affiliations/actions">
          <span itemprop="name">Actions</span>
          <meta itemprop="position" content="6">
</a>      </span>

      <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
        <a itemprop="url" class="js-selected-navigation-item reponav-item" data-selected-links="repo_wiki /drdanholmes/Affiliations/wiki" href="/drdanholmes/Affiliations/wiki">
          <span itemprop="name">Wiki</span>
          <meta itemprop="position" content="7">
</a>      </span>

      <a itemprop="url" class="js-selected-navigation-item reponav-item" data-selected-links="security alerts policy token_scanning code_scanning /drdanholmes/Affiliations/security/advisories" href="/drdanholmes/Affiliations/security/advisories">
        <span itemprop="name">Security</span>
        <meta itemprop="position" content="8">
</a>
      <a class="js-selected-navigation-item reponav-item" data-selected-links="pulse /drdanholmes/Affiliations/pulse" href="/drdanholmes/Affiliations/pulse">
        Pulse
</a>
      <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
        <a itemprop="url" class="js-selected-navigation-item reponav-item" data-selected-links="community /drdanholmes/Affiliations/community" href="/drdanholmes/Affiliations/community">
          Community
</a>      </span>

  </nav>
</div>

  </div>

  <include-fragment class="js-notification-shelf-include-fragment" data-base-src="https://github.com/notifications/beta/shelf"></include-fragment>

<div class="container-lg clearfix new-discussion-timeline  p-responsive">
  <div class="repository-content ">

    <a class="d-none js-permalink-shortcut" data-hotkey="y" href="/drdanholmes/Affiliations/blob/c2313b057f54bba751b84530b4b6c47911d73922/Extras/mytemplate.tex">Permalink</a>

    <!-- blob contrib key: blob_contributors:v21:522cfca4182adff5ba4b60fc6ef42e13 -->

    <div class="d-flex flex-items-start flex-shrink-0 flex-column flex-md-row pb-3">
      <span class="d-flex flex-justify-between width-full width-md-auto">

<details class="details-reset details-overlay branch-select-menu " id="branch-select-menu">
  <summary class="btn btn-sm css-truncate"
           data-hotkey="w"
           title="Switch branches or tags">
    <i>Branch:</i>
    <span class="css-truncate-target" data-menu-button>master</span>
    <span class="dropdown-caret"></span>
  </summary>

  <details-menu class="SelectMenu SelectMenu--hasFilter" src="/drdanholmes/Affiliations/refs/master/Extras/mytemplate.tex?source_action=show&amp;source_controller=blob" preload>
    <div class="SelectMenu-modal">
      <include-fragment class="SelectMenu-loading" aria-label="Menu is loading">
        <svg class="octicon octicon-octoface anim-pulse" height="32" viewBox="0 0 16 16" version="1.1" width="32" aria-hidden="true"><path fill-rule="evenodd" d="M14.7 5.34c.13-.32.55-1.59-.13-3.31 0 0-1.05-.33-3.44 1.3-1-.28-2.07-.32-3.13-.32s-2.13.04-3.13.32c-2.39-1.64-3.44-1.3-3.44-1.3-.68 1.72-.26 2.99-.13 3.31C.49 6.21 0 7.33 0 8.69 0 13.84 3.33 15 7.98 15S16 13.84 16 8.69c0-1.36-.49-2.48-1.3-3.35zM8 14.02c-3.3 0-5.98-.15-5.98-3.35 0-.76.38-1.48 1.02-2.07 1.07-.98 2.9-.46 4.96-.46 2.07 0 3.88-.52 4.96.46.65.59 1.02 1.3 1.02 2.07 0 3.19-2.68 3.35-5.98 3.35zM5.49 9.01c-.66 0-1.2.8-1.2 1.78s.54 1.79 1.2 1.79c.66 0 1.2-.8 1.2-1.79s-.54-1.78-1.2-1.78zm5.02 0c-.66 0-1.2.79-1.2 1.78s.54 1.79 1.2 1.79c.66 0 1.2-.8 1.2-1.79s-.53-1.78-1.2-1.78z"/></svg>
      </include-fragment>
    </div>
  </details-menu>
</details>

        <div class="BtnGroup flex-shrink-0 d-md-none">
          <a href="/drdanholmes/Affiliations/find/master"
                class="js-pjax-capture-input btn btn-sm BtnGroup-item"
                data-pjax
                data-hotkey="t">
            Find file
          </a>
          <clipboard-copy value="Extras/mytemplate.tex" class="btn btn-sm BtnGroup-item">
            Copy path
          </clipboard-copy>
        </div>
      </span>
      <h2 id="blob-path" class="breadcrumb flex-auto min-width-0 text-normal flex-md-self-center ml-md-2 mr-md-3 my-2 my-md-0">
        <span class="js-repo-root text-bold"><span class="js-path-segment"><a data-pjax="true" href="/drdanholmes/Affiliations"><span>Affiliations</span></a></span></span><span class="separator">/</span><span class="js-path-segment"><a data-pjax="true" href="/drdanholmes/Affiliations/tree/master/Extras"><span>Extras</span></a></span><span class="separator">/</span><strong class="final-path">mytemplate.tex</strong>
      </h2>

      <div class="BtnGroup flex-shrink-0 d-none d-md-inline-block">
        <a href="/drdanholmes/Affiliations/find/master"
              class="js-pjax-capture-input btn btn-sm BtnGroup-item"
              data-pjax
              data-hotkey="t">
          Find file
        </a>
        <clipboard-copy value="Extras/mytemplate.tex" class="btn btn-sm BtnGroup-item">
          Copy path
        </clipboard-copy>
      </div>
    </div>

    <include-fragment src="/drdanholmes/Affiliations/contributors/master/Extras/mytemplate.tex" class="Box Box--condensed commit-loader">
      <div class="Box-body bg-blue-light f6">
        Fetching contributors&hellip;
      </div>

      <div class="Box-body d-flex flex-items-center" >
        <img alt="" class="loader-loading mr-2" src="https://github.githubassets.com/images/spinners/octocat-spinner-32-EAF2F5.gif" width="16" height="16" />
        <span class="text-red h6 loader-error">Cannot retrieve contributors at this time</span>
      </div>
</include-fragment>

    <div class="Box mt-3 position-relative">

<div class="Box-header py-2 d-flex flex-column flex-shrink-0 flex-md-row flex-md-items-center">
  <div class="text-mono f6 flex-auto pr-3 flex-order-2 flex-md-order-1 mt-2 mt-md-0">

      512 lines (501 sloc)
      <span class="file-info-divider"></span>
    12.3 KB
  </div>

  <div class="d-flex py-1 py-md-0 flex-auto flex-order-1 flex-md-order-2 flex-sm-grow-0 flex-justify-between">

    <div class="BtnGroup">
      <a id="raw-url" class="btn btn-sm BtnGroup-item" href="/drdanholmes/Affiliations/raw/master/Extras/mytemplate.tex">Raw</a>
        <a class="btn btn-sm js-update-url-with-hash BtnGroup-item" data-hotkey="b" href="/drdanholmes/Affiliations/blame/master/Extras/mytemplate.tex">Blame</a>
      <a rel="nofollow" class="btn btn-sm BtnGroup-item" href="/drdanholmes/Affiliations/commits/master/Extras/mytemplate.tex">History</a>
    </div>

    <div>
          <a class="btn-octicon tooltipped tooltipped-nw js-remove-unless-platform"
             data-platforms="windows,mac"
             href="x-github-client://openRepo/https://github.com/drdanholmes/Affiliations?branch=master&amp;filepath=Extras%2Fmytemplate.tex"
             aria-label="Open this file in GitHub Desktop"
             data-ga-click="Repository, open with desktop">
              <svg class="octicon octicon-device-desktop" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M15 2H1c-.55 0-1 .45-1 1v9c0 .55.45 1 1 1h5.34c-.25.61-.86 1.39-2.34 2h8c-1.48-.61-2.09-1.39-2.34-2H15c.55 0 1-.45 1-1V3c0-.55-.45-1-1-1zm0 9H1V3h14v8z"/></svg>
          </a>

          <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="inline-form js-update-url-with-hash" action="/drdanholmes/Affiliations/edit/master/Extras/mytemplate.tex" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="FkqcmkViFJ8yLr2BRh9DspHRYSGH7MwcONhpBDifwkNY32iPzkUhkJw2nPAW8wpJ6g8BSuzP4I6HqoVaEqbsCA==" />
            <button class="btn-octicon tooltipped tooltipped-nw" type="submit"
              aria-label="Fork this project and edit the file" data-hotkey="e" data-disable-with>
              <svg class="octicon octicon-pencil" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M0 12v3h3l8-8-3-3-8 8zm3 2H1v-2h1v1h1v1zm10.3-9.3L12 6 9 3l1.3-1.3a.996.996 0 011.41 0l1.59 1.59c.39.39.39 1.02 0 1.41z"/></svg>
            </button>
</form>
          <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="inline-form" action="/drdanholmes/Affiliations/delete/master/Extras/mytemplate.tex" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="MwsRKDsPnkNg45DjlxL1ASMGxMy32Xt9Y2b7OSufkCdAiddYmXYVdwju+6PkURt2oPF9H7/SnV6QiEcedasvLQ==" />
            <button class="btn-octicon btn-octicon-danger tooltipped tooltipped-nw" type="submit"
              aria-label="Fork this project and delete the file" data-disable-with>
              <svg class="octicon octicon-trashcan" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M11 2H9c0-.55-.45-1-1-1H5c-.55 0-1 .45-1 1H2c-.55 0-1 .45-1 1v1c0 .55.45 1 1 1v9c0 .55.45 1 1 1h7c.55 0 1-.45 1-1V5c.55 0 1-.45 1-1V3c0-.55-.45-1-1-1zm-1 12H3V5h1v8h1V5h1v8h1V5h1v8h1V5h1v9zm1-10H2V3h9v1z"/></svg>
            </button>
</form>    </div>
  </div>
</div>

  <div itemprop="text" class="Box-body p-0 blob-wrapper data type-tex ">

<table class="highlight tab-size js-file-line-container" data-tab-size="8">
      <tr>
        <td id="L1" class="blob-num js-line-number" data-line-number="1"></td>
        <td id="LC1" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> Options for packages loaded elsewhere</span></td>
      </tr>
      <tr>
        <td id="L2" class="blob-num js-line-number" data-line-number="2"></td>
        <td id="LC2" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\PassOptionsToPackage</span>{unicode<span class="pl-s"><span class="pl-pds">$</span>for(hyperrefoptions)<span class="pl-pds">$</span></span>,<span class="pl-s"><span class="pl-pds">$</span>hyperrefoptions<span class="pl-pds">$</span><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span>}{hyperref}</td>
      </tr>
      <tr>
        <td id="L3" class="blob-num js-line-number" data-line-number="3"></td>
        <td id="LC3" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\PassOptionsToPackage</span>{hyphens}{url}</td>
      </tr>
      <tr>
        <td id="L4" class="blob-num js-line-number" data-line-number="4"></td>
        <td id="LC4" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(colorlinks)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L5" class="blob-num js-line-number" data-line-number="5"></td>
        <td id="LC5" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\PassOptionsToPackage</span>{dvipsnames,svgnames*,x11names*}{xcolor}</td>
      </tr>
      <tr>
        <td id="L6" class="blob-num js-line-number" data-line-number="6"></td>
        <td id="LC6" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L7" class="blob-num js-line-number" data-line-number="7"></td>
        <td id="LC7" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(dir)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L8" class="blob-num js-line-number" data-line-number="8"></td>
        <td id="LC8" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(latex-dir-rtl)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L9" class="blob-num js-line-number" data-line-number="9"></td>
        <td id="LC9" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\PassOptionsToPackage</span>{RTLdocument}{bidi}</td>
      </tr>
      <tr>
        <td id="L10" class="blob-num js-line-number" data-line-number="10"></td>
        <td id="LC10" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L11" class="blob-num js-line-number" data-line-number="11"></td>
        <td id="LC11" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L12" class="blob-num js-line-number" data-line-number="12"></td>
        <td id="LC12" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span></span></td>
      </tr>
      <tr>
        <td id="L13" class="blob-num js-line-number" data-line-number="13"></td>
        <td id="LC13" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\documentclass</span>[</td>
      </tr>
      <tr>
        <td id="L14" class="blob-num js-line-number" data-line-number="14"></td>
        <td id="LC14" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(fontsize)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L15" class="blob-num js-line-number" data-line-number="15"></td>
        <td id="LC15" class="blob-code blob-code-inner js-file-line">  <span class="pl-s"><span class="pl-pds">$</span>fontsize<span class="pl-pds">$</span></span>,</td>
      </tr>
      <tr>
        <td id="L16" class="blob-num js-line-number" data-line-number="16"></td>
        <td id="LC16" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L17" class="blob-num js-line-number" data-line-number="17"></td>
        <td id="LC17" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(lang)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L18" class="blob-num js-line-number" data-line-number="18"></td>
        <td id="LC18" class="blob-code blob-code-inner js-file-line">  <span class="pl-s"><span class="pl-pds">$</span>babel-lang<span class="pl-pds">$</span></span>,</td>
      </tr>
      <tr>
        <td id="L19" class="blob-num js-line-number" data-line-number="19"></td>
        <td id="LC19" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L20" class="blob-num js-line-number" data-line-number="20"></td>
        <td id="LC20" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(papersize)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L21" class="blob-num js-line-number" data-line-number="21"></td>
        <td id="LC21" class="blob-code blob-code-inner js-file-line">  <span class="pl-s"><span class="pl-pds">$</span>papersize<span class="pl-pds">$</span></span>paper,</td>
      </tr>
      <tr>
        <td id="L22" class="blob-num js-line-number" data-line-number="22"></td>
        <td id="LC22" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L23" class="blob-num js-line-number" data-line-number="23"></td>
        <td id="LC23" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(beamer)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L24" class="blob-num js-line-number" data-line-number="24"></td>
        <td id="LC24" class="blob-code blob-code-inner js-file-line">  ignorenonframetext,</td>
      </tr>
      <tr>
        <td id="L25" class="blob-num js-line-number" data-line-number="25"></td>
        <td id="LC25" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(handout)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L26" class="blob-num js-line-number" data-line-number="26"></td>
        <td id="LC26" class="blob-code blob-code-inner js-file-line">  handout,</td>
      </tr>
      <tr>
        <td id="L27" class="blob-num js-line-number" data-line-number="27"></td>
        <td id="LC27" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L28" class="blob-num js-line-number" data-line-number="28"></td>
        <td id="LC28" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(aspectratio)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L29" class="blob-num js-line-number" data-line-number="29"></td>
        <td id="LC29" class="blob-code blob-code-inner js-file-line">  aspectratio=<span class="pl-s"><span class="pl-pds">$</span>aspectratio<span class="pl-pds">$</span></span>,</td>
      </tr>
      <tr>
        <td id="L30" class="blob-num js-line-number" data-line-number="30"></td>
        <td id="LC30" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L31" class="blob-num js-line-number" data-line-number="31"></td>
        <td id="LC31" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L32" class="blob-num js-line-number" data-line-number="32"></td>
        <td id="LC32" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>for(classoption)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L33" class="blob-num js-line-number" data-line-number="33"></td>
        <td id="LC33" class="blob-code blob-code-inner js-file-line">  <span class="pl-s"><span class="pl-pds">$</span>classoption<span class="pl-pds">$</span><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span>,</td>
      </tr>
      <tr>
        <td id="L34" class="blob-num js-line-number" data-line-number="34"></td>
        <td id="LC34" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L35" class="blob-num js-line-number" data-line-number="35"></td>
        <td id="LC35" class="blob-code blob-code-inner js-file-line">]{<span class="pl-s"><span class="pl-pds">$</span>documentclass<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L36" class="blob-num js-line-number" data-line-number="36"></td>
        <td id="LC36" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(beamer)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L37" class="blob-num js-line-number" data-line-number="37"></td>
        <td id="LC37" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(background-image)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L38" class="blob-num js-line-number" data-line-number="38"></td>
        <td id="LC38" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\usebackgroundtemplate</span>{<span class="pl-c"><span class="pl-c">%</span></span></td>
      </tr>
      <tr>
        <td id="L39" class="blob-num js-line-number" data-line-number="39"></td>
        <td id="LC39" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\includegraphics</span>[width=<span class="pl-c1">\paperwidth</span>]{<span class="pl-s"><span class="pl-pds">$</span>background-image<span class="pl-pds">$</span></span>}<span class="pl-c"><span class="pl-c">%</span></span></td>
      </tr>
      <tr>
        <td id="L40" class="blob-num js-line-number" data-line-number="40"></td>
        <td id="LC40" class="blob-code blob-code-inner js-file-line">}</td>
      </tr>
      <tr>
        <td id="L41" class="blob-num js-line-number" data-line-number="41"></td>
        <td id="LC41" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L42" class="blob-num js-line-number" data-line-number="42"></td>
        <td id="LC42" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">pgfpages</span>}</td>
      </tr>
      <tr>
        <td id="L43" class="blob-num js-line-number" data-line-number="43"></td>
        <td id="LC43" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\setbeamertemplate</span>{caption}[numbered]</td>
      </tr>
      <tr>
        <td id="L44" class="blob-num js-line-number" data-line-number="44"></td>
        <td id="LC44" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\setbeamertemplate</span>{caption label separator}{: }</td>
      </tr>
      <tr>
        <td id="L45" class="blob-num js-line-number" data-line-number="45"></td>
        <td id="LC45" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\setbeamercolor</span>{caption name}{fg=normal text.fg}</td>
      </tr>
      <tr>
        <td id="L46" class="blob-num js-line-number" data-line-number="46"></td>
        <td id="LC46" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\beamertemplatenavigationsymbols</span><span class="pl-s"><span class="pl-pds">$</span>if(navigation)<span class="pl-pds">$</span><span class="pl-pds">$</span>navigation<span class="pl-pds">$</span><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span>empty<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L47" class="blob-num js-line-number" data-line-number="47"></td>
        <td id="LC47" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>for(beameroption)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L48" class="blob-num js-line-number" data-line-number="48"></td>
        <td id="LC48" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\setbeameroption</span>{<span class="pl-s"><span class="pl-pds">$</span>beameroption<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L49" class="blob-num js-line-number" data-line-number="49"></td>
        <td id="LC49" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L50" class="blob-num js-line-number" data-line-number="50"></td>
        <td id="LC50" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> Prevent slide breaks in the middle of a paragraph</span></td>
      </tr>
      <tr>
        <td id="L51" class="blob-num js-line-number" data-line-number="51"></td>
        <td id="LC51" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\widowpenalties</span> 1 10000</td>
      </tr>
      <tr>
        <td id="L52" class="blob-num js-line-number" data-line-number="52"></td>
        <td id="LC52" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\raggedbottom</span></td>
      </tr>
      <tr>
        <td id="L53" class="blob-num js-line-number" data-line-number="53"></td>
        <td id="LC53" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(section-titles)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L54" class="blob-num js-line-number" data-line-number="54"></td>
        <td id="LC54" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\setbeamertemplate</span>{part page}{</td>
      </tr>
      <tr>
        <td id="L55" class="blob-num js-line-number" data-line-number="55"></td>
        <td id="LC55" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\centering</span></td>
      </tr>
      <tr>
        <td id="L56" class="blob-num js-line-number" data-line-number="56"></td>
        <td id="LC56" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\begin</span>{<span class="pl-smi">beamercolorbox</span>}[sep=16pt,center]{part title}</td>
      </tr>
      <tr>
        <td id="L57" class="blob-num js-line-number" data-line-number="57"></td>
        <td id="LC57" class="blob-code blob-code-inner js-file-line">    <span class="pl-c1">\usebeamerfont</span>{part title}<span class="pl-c1">\insertpart\par</span></td>
      </tr>
      <tr>
        <td id="L58" class="blob-num js-line-number" data-line-number="58"></td>
        <td id="LC58" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\end</span>{<span class="pl-smi">beamercolorbox</span>}</td>
      </tr>
      <tr>
        <td id="L59" class="blob-num js-line-number" data-line-number="59"></td>
        <td id="LC59" class="blob-code blob-code-inner js-file-line">}</td>
      </tr>
      <tr>
        <td id="L60" class="blob-num js-line-number" data-line-number="60"></td>
        <td id="LC60" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\setbeamertemplate</span>{section page}{</td>
      </tr>
      <tr>
        <td id="L61" class="blob-num js-line-number" data-line-number="61"></td>
        <td id="LC61" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\centering</span></td>
      </tr>
      <tr>
        <td id="L62" class="blob-num js-line-number" data-line-number="62"></td>
        <td id="LC62" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\begin</span>{<span class="pl-smi">beamercolorbox</span>}[sep=12pt,center]{part title}</td>
      </tr>
      <tr>
        <td id="L63" class="blob-num js-line-number" data-line-number="63"></td>
        <td id="LC63" class="blob-code blob-code-inner js-file-line">    <span class="pl-c1">\usebeamerfont</span>{section title}<span class="pl-c1">\insertsection\par</span></td>
      </tr>
      <tr>
        <td id="L64" class="blob-num js-line-number" data-line-number="64"></td>
        <td id="LC64" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\end</span>{<span class="pl-smi">beamercolorbox</span>}</td>
      </tr>
      <tr>
        <td id="L65" class="blob-num js-line-number" data-line-number="65"></td>
        <td id="LC65" class="blob-code blob-code-inner js-file-line">}</td>
      </tr>
      <tr>
        <td id="L66" class="blob-num js-line-number" data-line-number="66"></td>
        <td id="LC66" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\setbeamertemplate</span>{subsection page}{</td>
      </tr>
      <tr>
        <td id="L67" class="blob-num js-line-number" data-line-number="67"></td>
        <td id="LC67" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\centering</span></td>
      </tr>
      <tr>
        <td id="L68" class="blob-num js-line-number" data-line-number="68"></td>
        <td id="LC68" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\begin</span>{<span class="pl-smi">beamercolorbox</span>}[sep=8pt,center]{part title}</td>
      </tr>
      <tr>
        <td id="L69" class="blob-num js-line-number" data-line-number="69"></td>
        <td id="LC69" class="blob-code blob-code-inner js-file-line">    <span class="pl-c1">\usebeamerfont</span>{subsection title}<span class="pl-c1">\insertsubsection\par</span></td>
      </tr>
      <tr>
        <td id="L70" class="blob-num js-line-number" data-line-number="70"></td>
        <td id="LC70" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\end</span>{<span class="pl-smi">beamercolorbox</span>}</td>
      </tr>
      <tr>
        <td id="L71" class="blob-num js-line-number" data-line-number="71"></td>
        <td id="LC71" class="blob-code blob-code-inner js-file-line">}</td>
      </tr>
      <tr>
        <td id="L72" class="blob-num js-line-number" data-line-number="72"></td>
        <td id="LC72" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\AtBeginPart</span>{</td>
      </tr>
      <tr>
        <td id="L73" class="blob-num js-line-number" data-line-number="73"></td>
        <td id="LC73" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\frame</span>{<span class="pl-c1">\partpage</span>}</td>
      </tr>
      <tr>
        <td id="L74" class="blob-num js-line-number" data-line-number="74"></td>
        <td id="LC74" class="blob-code blob-code-inner js-file-line">}</td>
      </tr>
      <tr>
        <td id="L75" class="blob-num js-line-number" data-line-number="75"></td>
        <td id="LC75" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\AtBeginSection</span>{</td>
      </tr>
      <tr>
        <td id="L76" class="blob-num js-line-number" data-line-number="76"></td>
        <td id="LC76" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\ifbibliography</span></td>
      </tr>
      <tr>
        <td id="L77" class="blob-num js-line-number" data-line-number="77"></td>
        <td id="LC77" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\else</span></td>
      </tr>
      <tr>
        <td id="L78" class="blob-num js-line-number" data-line-number="78"></td>
        <td id="LC78" class="blob-code blob-code-inner js-file-line">    <span class="pl-c1">\frame</span>{<span class="pl-c1">\sectionpage</span>}</td>
      </tr>
      <tr>
        <td id="L79" class="blob-num js-line-number" data-line-number="79"></td>
        <td id="LC79" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\fi</span></td>
      </tr>
      <tr>
        <td id="L80" class="blob-num js-line-number" data-line-number="80"></td>
        <td id="LC80" class="blob-code blob-code-inner js-file-line">}</td>
      </tr>
      <tr>
        <td id="L81" class="blob-num js-line-number" data-line-number="81"></td>
        <td id="LC81" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\AtBeginSubsection</span>{</td>
      </tr>
      <tr>
        <td id="L82" class="blob-num js-line-number" data-line-number="82"></td>
        <td id="LC82" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\frame</span>{<span class="pl-c1">\subsectionpage</span>}</td>
      </tr>
      <tr>
        <td id="L83" class="blob-num js-line-number" data-line-number="83"></td>
        <td id="LC83" class="blob-code blob-code-inner js-file-line">}</td>
      </tr>
      <tr>
        <td id="L84" class="blob-num js-line-number" data-line-number="84"></td>
        <td id="LC84" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L85" class="blob-num js-line-number" data-line-number="85"></td>
        <td id="LC85" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L86" class="blob-num js-line-number" data-line-number="86"></td>
        <td id="LC86" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(beamerarticle)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L87" class="blob-num js-line-number" data-line-number="87"></td>
        <td id="LC87" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">beamerarticle</span>} <span class="pl-c"><span class="pl-c">%</span> needs to be loaded first</span></td>
      </tr>
      <tr>
        <td id="L88" class="blob-num js-line-number" data-line-number="88"></td>
        <td id="LC88" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L89" class="blob-num js-line-number" data-line-number="89"></td>
        <td id="LC89" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(fontfamily)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L90" class="blob-num js-line-number" data-line-number="90"></td>
        <td id="LC90" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>[<span class="pl-v">$for(fontfamilyoptions)$$fontfamilyoptions$$sep$,$endfor$</span>]{<span class="pl-c1"><span class="pl-s"><span class="pl-pds">$</span>fontfamily<span class="pl-pds">$</span></span></span>}</td>
      </tr>
      <tr>
        <td id="L91" class="blob-num js-line-number" data-line-number="91"></td>
        <td id="LC91" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L92" class="blob-num js-line-number" data-line-number="92"></td>
        <td id="LC92" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">lmodern</span>}</td>
      </tr>
      <tr>
        <td id="L93" class="blob-num js-line-number" data-line-number="93"></td>
        <td id="LC93" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L94" class="blob-num js-line-number" data-line-number="94"></td>
        <td id="LC94" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(linestretch)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L95" class="blob-num js-line-number" data-line-number="95"></td>
        <td id="LC95" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">setspace</span>}</td>
      </tr>
      <tr>
        <td id="L96" class="blob-num js-line-number" data-line-number="96"></td>
        <td id="LC96" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L97" class="blob-num js-line-number" data-line-number="97"></td>
        <td id="LC97" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">amssymb,amsmath</span>}</td>
      </tr>
      <tr>
        <td id="L98" class="blob-num js-line-number" data-line-number="98"></td>
        <td id="LC98" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">ifxetex,ifluatex</span>}</td>
      </tr>
      <tr>
        <td id="L99" class="blob-num js-line-number" data-line-number="99"></td>
        <td id="LC99" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\ifnum</span> 0<span class="pl-c1">\ifxetex</span> 1<span class="pl-k">\fi</span><span class="pl-c1">\ifluatex</span> 1<span class="pl-k">\fi</span>=0 <span class="pl-c"><span class="pl-c">%</span> if pdftex</span></td>
      </tr>
      <tr>
        <td id="L100" class="blob-num js-line-number" data-line-number="100"></td>
        <td id="LC100" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\usepackage</span>[<span class="pl-v">$if(fontenc)$$fontenc$$else$T1$endif$</span>]{<span class="pl-c1">fontenc</span>}</td>
      </tr>
      <tr>
        <td id="L101" class="blob-num js-line-number" data-line-number="101"></td>
        <td id="LC101" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\usepackage</span>[<span class="pl-v">utf8</span>]{<span class="pl-c1">inputenc</span>}</td>
      </tr>
      <tr>
        <td id="L102" class="blob-num js-line-number" data-line-number="102"></td>
        <td id="LC102" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\usepackage</span>{<span class="pl-c1">textcomp</span>} <span class="pl-c"><span class="pl-c">%</span> provide euro and other symbols</span></td>
      </tr>
      <tr>
        <td id="L103" class="blob-num js-line-number" data-line-number="103"></td>
        <td id="LC103" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\else</span> <span class="pl-c"><span class="pl-c">%</span> if luatex or xetex</span></td>
      </tr>
      <tr>
        <td id="L104" class="blob-num js-line-number" data-line-number="104"></td>
        <td id="LC104" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(mathspec)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L105" class="blob-num js-line-number" data-line-number="105"></td>
        <td id="LC105" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\ifxetex</span></td>
      </tr>
      <tr>
        <td id="L106" class="blob-num js-line-number" data-line-number="106"></td>
        <td id="LC106" class="blob-code blob-code-inner js-file-line">    <span class="pl-k">\usepackage</span>{<span class="pl-c1">mathspec</span>}</td>
      </tr>
      <tr>
        <td id="L107" class="blob-num js-line-number" data-line-number="107"></td>
        <td id="LC107" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\else</span></td>
      </tr>
      <tr>
        <td id="L108" class="blob-num js-line-number" data-line-number="108"></td>
        <td id="LC108" class="blob-code blob-code-inner js-file-line">    <span class="pl-k">\usepackage</span>{<span class="pl-c1">unicode-math</span>}</td>
      </tr>
      <tr>
        <td id="L109" class="blob-num js-line-number" data-line-number="109"></td>
        <td id="LC109" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\fi</span></td>
      </tr>
      <tr>
        <td id="L110" class="blob-num js-line-number" data-line-number="110"></td>
        <td id="LC110" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L111" class="blob-num js-line-number" data-line-number="111"></td>
        <td id="LC111" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\usepackage</span>{<span class="pl-c1">unicode-math</span>}</td>
      </tr>
      <tr>
        <td id="L112" class="blob-num js-line-number" data-line-number="112"></td>
        <td id="LC112" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L113" class="blob-num js-line-number" data-line-number="113"></td>
        <td id="LC113" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\defaultfontfeatures</span>{Scale=MatchLowercase}</td>
      </tr>
      <tr>
        <td id="L114" class="blob-num js-line-number" data-line-number="114"></td>
        <td id="LC114" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\defaultfontfeatures</span>[<span class="pl-c1">\rmfamily</span>]{Ligatures=TeX,Scale=1}</td>
      </tr>
      <tr>
        <td id="L115" class="blob-num js-line-number" data-line-number="115"></td>
        <td id="LC115" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(mainfont)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L116" class="blob-num js-line-number" data-line-number="116"></td>
        <td id="LC116" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\setmainfont</span>[<span class="pl-s"><span class="pl-pds">$</span>for(mainfontoptions)<span class="pl-pds">$</span><span class="pl-pds">$</span>mainfontoptions<span class="pl-pds">$</span><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span>,<span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span>]{<span class="pl-s"><span class="pl-pds">$</span>mainfont<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L117" class="blob-num js-line-number" data-line-number="117"></td>
        <td id="LC117" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L118" class="blob-num js-line-number" data-line-number="118"></td>
        <td id="LC118" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(sansfont)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L119" class="blob-num js-line-number" data-line-number="119"></td>
        <td id="LC119" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\setsansfont</span>[<span class="pl-s"><span class="pl-pds">$</span>for(sansfontoptions)<span class="pl-pds">$</span><span class="pl-pds">$</span>sansfontoptions<span class="pl-pds">$</span><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span>,<span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span>]{<span class="pl-s"><span class="pl-pds">$</span>sansfont<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L120" class="blob-num js-line-number" data-line-number="120"></td>
        <td id="LC120" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L121" class="blob-num js-line-number" data-line-number="121"></td>
        <td id="LC121" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(monofont)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L122" class="blob-num js-line-number" data-line-number="122"></td>
        <td id="LC122" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\setmonofont</span>[<span class="pl-s"><span class="pl-pds">$</span>for(monofontoptions)<span class="pl-pds">$</span><span class="pl-pds">$</span>monofontoptions<span class="pl-pds">$</span><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span>,<span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span>]{<span class="pl-s"><span class="pl-pds">$</span>monofont<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L123" class="blob-num js-line-number" data-line-number="123"></td>
        <td id="LC123" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L124" class="blob-num js-line-number" data-line-number="124"></td>
        <td id="LC124" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>for(fontfamilies)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L125" class="blob-num js-line-number" data-line-number="125"></td>
        <td id="LC125" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\newfontfamily</span>{<span class="pl-s"><span class="pl-pds">$</span>fontfamilies.name<span class="pl-pds">$</span></span>}[<span class="pl-s"><span class="pl-pds">$</span>for(fontfamilies.options)<span class="pl-pds">$</span><span class="pl-pds">$</span>fontfamilies.options<span class="pl-pds">$</span><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span>,<span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span>]{<span class="pl-s"><span class="pl-pds">$</span>fontfamilies.font<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L126" class="blob-num js-line-number" data-line-number="126"></td>
        <td id="LC126" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L127" class="blob-num js-line-number" data-line-number="127"></td>
        <td id="LC127" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(mathfont)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L128" class="blob-num js-line-number" data-line-number="128"></td>
        <td id="LC128" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(mathspec)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L129" class="blob-num js-line-number" data-line-number="129"></td>
        <td id="LC129" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\ifxetex</span></td>
      </tr>
      <tr>
        <td id="L130" class="blob-num js-line-number" data-line-number="130"></td>
        <td id="LC130" class="blob-code blob-code-inner js-file-line">    <span class="pl-c1">\setmathfont</span>(Digits,Latin,Greek)[<span class="pl-s"><span class="pl-pds">$</span>for(mathfontoptions)<span class="pl-pds">$</span><span class="pl-pds">$</span>mathfontoptions<span class="pl-pds">$</span><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span>,<span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span>]{<span class="pl-s"><span class="pl-pds">$</span>mathfont<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L131" class="blob-num js-line-number" data-line-number="131"></td>
        <td id="LC131" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\else</span></td>
      </tr>
      <tr>
        <td id="L132" class="blob-num js-line-number" data-line-number="132"></td>
        <td id="LC132" class="blob-code blob-code-inner js-file-line">    <span class="pl-c1">\setmathfont</span>[<span class="pl-s"><span class="pl-pds">$</span>for(mathfontoptions)<span class="pl-pds">$</span><span class="pl-pds">$</span>mathfontoptions<span class="pl-pds">$</span><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span>,<span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span>]{<span class="pl-s"><span class="pl-pds">$</span>mathfont<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L133" class="blob-num js-line-number" data-line-number="133"></td>
        <td id="LC133" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\fi</span></td>
      </tr>
      <tr>
        <td id="L134" class="blob-num js-line-number" data-line-number="134"></td>
        <td id="LC134" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L135" class="blob-num js-line-number" data-line-number="135"></td>
        <td id="LC135" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\setmathfont</span>[<span class="pl-s"><span class="pl-pds">$</span>for(mathfontoptions)<span class="pl-pds">$</span><span class="pl-pds">$</span>mathfontoptions<span class="pl-pds">$</span><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span>,<span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span>]{<span class="pl-s"><span class="pl-pds">$</span>mathfont<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L136" class="blob-num js-line-number" data-line-number="136"></td>
        <td id="LC136" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L137" class="blob-num js-line-number" data-line-number="137"></td>
        <td id="LC137" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L138" class="blob-num js-line-number" data-line-number="138"></td>
        <td id="LC138" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(CJKmainfont)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L139" class="blob-num js-line-number" data-line-number="139"></td>
        <td id="LC139" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\ifxetex</span></td>
      </tr>
      <tr>
        <td id="L140" class="blob-num js-line-number" data-line-number="140"></td>
        <td id="LC140" class="blob-code blob-code-inner js-file-line">    <span class="pl-k">\usepackage</span>{<span class="pl-c1">xeCJK</span>}</td>
      </tr>
      <tr>
        <td id="L141" class="blob-num js-line-number" data-line-number="141"></td>
        <td id="LC141" class="blob-code blob-code-inner js-file-line">    <span class="pl-c1">\setCJKmainfont</span>[<span class="pl-s"><span class="pl-pds">$</span>for(CJKoptions)<span class="pl-pds">$</span><span class="pl-pds">$</span>CJKoptions<span class="pl-pds">$</span><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span>,<span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span>]{<span class="pl-s"><span class="pl-pds">$</span>CJKmainfont<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L142" class="blob-num js-line-number" data-line-number="142"></td>
        <td id="LC142" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\fi</span></td>
      </tr>
      <tr>
        <td id="L143" class="blob-num js-line-number" data-line-number="143"></td>
        <td id="LC143" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L144" class="blob-num js-line-number" data-line-number="144"></td>
        <td id="LC144" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(luatexjapresetoptions)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L145" class="blob-num js-line-number" data-line-number="145"></td>
        <td id="LC145" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\ifluatex</span></td>
      </tr>
      <tr>
        <td id="L146" class="blob-num js-line-number" data-line-number="146"></td>
        <td id="LC146" class="blob-code blob-code-inner js-file-line">    <span class="pl-k">\usepackage</span>[<span class="pl-v">$for(luatexjapresetoptions)$$luatexjapresetoptions$$sep$,$endfor$</span>]{<span class="pl-c1">luatexja-preset</span>}</td>
      </tr>
      <tr>
        <td id="L147" class="blob-num js-line-number" data-line-number="147"></td>
        <td id="LC147" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\fi</span></td>
      </tr>
      <tr>
        <td id="L148" class="blob-num js-line-number" data-line-number="148"></td>
        <td id="LC148" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L149" class="blob-num js-line-number" data-line-number="149"></td>
        <td id="LC149" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(CJKmainfont)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L150" class="blob-num js-line-number" data-line-number="150"></td>
        <td id="LC150" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\ifluatex</span></td>
      </tr>
      <tr>
        <td id="L151" class="blob-num js-line-number" data-line-number="151"></td>
        <td id="LC151" class="blob-code blob-code-inner js-file-line">    <span class="pl-k">\usepackage</span>[<span class="pl-v">$for(luatexjafontspecoptions)$$luatexjafontspecoptions$$sep$,$endfor$</span>]{<span class="pl-c1">luatexja-fontspec</span>}</td>
      </tr>
      <tr>
        <td id="L152" class="blob-num js-line-number" data-line-number="152"></td>
        <td id="LC152" class="blob-code blob-code-inner js-file-line">    <span class="pl-c1">\setmainjfont</span>[<span class="pl-s"><span class="pl-pds">$</span>for(CJKoptions)<span class="pl-pds">$</span><span class="pl-pds">$</span>CJKoptions<span class="pl-pds">$</span><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span>,<span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span>]{<span class="pl-s"><span class="pl-pds">$</span>CJKmainfont<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L153" class="blob-num js-line-number" data-line-number="153"></td>
        <td id="LC153" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\fi</span></td>
      </tr>
      <tr>
        <td id="L154" class="blob-num js-line-number" data-line-number="154"></td>
        <td id="LC154" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L155" class="blob-num js-line-number" data-line-number="155"></td>
        <td id="LC155" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\fi</span></td>
      </tr>
      <tr>
        <td id="L156" class="blob-num js-line-number" data-line-number="156"></td>
        <td id="LC156" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(beamer)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L157" class="blob-num js-line-number" data-line-number="157"></td>
        <td id="LC157" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(theme)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L158" class="blob-num js-line-number" data-line-number="158"></td>
        <td id="LC158" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\usetheme</span>[<span class="pl-s"><span class="pl-pds">$</span>for(themeoptions)<span class="pl-pds">$</span><span class="pl-pds">$</span>themeoptions<span class="pl-pds">$</span><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span>,<span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span>]{<span class="pl-s"><span class="pl-pds">$</span>theme<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L159" class="blob-num js-line-number" data-line-number="159"></td>
        <td id="LC159" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L160" class="blob-num js-line-number" data-line-number="160"></td>
        <td id="LC160" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(colortheme)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L161" class="blob-num js-line-number" data-line-number="161"></td>
        <td id="LC161" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\usecolortheme</span>{<span class="pl-s"><span class="pl-pds">$</span>colortheme<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L162" class="blob-num js-line-number" data-line-number="162"></td>
        <td id="LC162" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L163" class="blob-num js-line-number" data-line-number="163"></td>
        <td id="LC163" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(fonttheme)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L164" class="blob-num js-line-number" data-line-number="164"></td>
        <td id="LC164" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\usefonttheme</span>{<span class="pl-s"><span class="pl-pds">$</span>fonttheme<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L165" class="blob-num js-line-number" data-line-number="165"></td>
        <td id="LC165" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L166" class="blob-num js-line-number" data-line-number="166"></td>
        <td id="LC166" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(mainfont)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L167" class="blob-num js-line-number" data-line-number="167"></td>
        <td id="LC167" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\usefonttheme</span>{serif} <span class="pl-c"><span class="pl-c">%</span> use mainfont rather than sansfont for slide text</span></td>
      </tr>
      <tr>
        <td id="L168" class="blob-num js-line-number" data-line-number="168"></td>
        <td id="LC168" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L169" class="blob-num js-line-number" data-line-number="169"></td>
        <td id="LC169" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(innertheme)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L170" class="blob-num js-line-number" data-line-number="170"></td>
        <td id="LC170" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\useinnertheme</span>{<span class="pl-s"><span class="pl-pds">$</span>innertheme<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L171" class="blob-num js-line-number" data-line-number="171"></td>
        <td id="LC171" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L172" class="blob-num js-line-number" data-line-number="172"></td>
        <td id="LC172" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(outertheme)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L173" class="blob-num js-line-number" data-line-number="173"></td>
        <td id="LC173" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\useoutertheme</span>{<span class="pl-s"><span class="pl-pds">$</span>outertheme<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L174" class="blob-num js-line-number" data-line-number="174"></td>
        <td id="LC174" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L175" class="blob-num js-line-number" data-line-number="175"></td>
        <td id="LC175" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L176" class="blob-num js-line-number" data-line-number="176"></td>
        <td id="LC176" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> Use upquote if available, for straight quotes in verbatim environments</span></td>
      </tr>
      <tr>
        <td id="L177" class="blob-num js-line-number" data-line-number="177"></td>
        <td id="LC177" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\IfFileExists</span>{upquote.sty}{<span class="pl-k">\usepackage</span>{<span class="pl-c1">upquote</span>}}{}</td>
      </tr>
      <tr>
        <td id="L178" class="blob-num js-line-number" data-line-number="178"></td>
        <td id="LC178" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\IfFileExists</span>{microtype.sty}{<span class="pl-c"><span class="pl-c">%</span> use microtype if available</span></td>
      </tr>
      <tr>
        <td id="L179" class="blob-num js-line-number" data-line-number="179"></td>
        <td id="LC179" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\usepackage</span>[<span class="pl-v">$for(microtypeoptions)$$microtypeoptions$$sep$,$endfor$</span>]{<span class="pl-c1">microtype</span>}</td>
      </tr>
      <tr>
        <td id="L180" class="blob-num js-line-number" data-line-number="180"></td>
        <td id="LC180" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\UseMicrotypeSet</span>[protrusion]{basicmath} <span class="pl-c"><span class="pl-c">%</span> disable protrusion for tt fonts</span></td>
      </tr>
      <tr>
        <td id="L181" class="blob-num js-line-number" data-line-number="181"></td>
        <td id="LC181" class="blob-code blob-code-inner js-file-line">}{}</td>
      </tr>
      <tr>
        <td id="L182" class="blob-num js-line-number" data-line-number="182"></td>
        <td id="LC182" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(indent)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L183" class="blob-num js-line-number" data-line-number="183"></td>
        <td id="LC183" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L184" class="blob-num js-line-number" data-line-number="184"></td>
        <td id="LC184" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatletter</span></td>
      </tr>
      <tr>
        <td id="L185" class="blob-num js-line-number" data-line-number="185"></td>
        <td id="LC185" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\@ifundefined</span>{KOMAClassName}{<span class="pl-c"><span class="pl-c">%</span> if non-KOMA class</span></td>
      </tr>
      <tr>
        <td id="L186" class="blob-num js-line-number" data-line-number="186"></td>
        <td id="LC186" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\IfFileExists</span>{parskip.sty}{<span class="pl-c"><span class="pl-c">%</span></span></td>
      </tr>
      <tr>
        <td id="L187" class="blob-num js-line-number" data-line-number="187"></td>
        <td id="LC187" class="blob-code blob-code-inner js-file-line">    <span class="pl-k">\usepackage</span>{<span class="pl-c1">parskip</span>}</td>
      </tr>
      <tr>
        <td id="L188" class="blob-num js-line-number" data-line-number="188"></td>
        <td id="LC188" class="blob-code blob-code-inner js-file-line">  }{<span class="pl-c"><span class="pl-c">%</span> else</span></td>
      </tr>
      <tr>
        <td id="L189" class="blob-num js-line-number" data-line-number="189"></td>
        <td id="LC189" class="blob-code blob-code-inner js-file-line">    <span class="pl-c1">\setlength</span>{<span class="pl-c1">\parindent</span>}{0pt}</td>
      </tr>
      <tr>
        <td id="L190" class="blob-num js-line-number" data-line-number="190"></td>
        <td id="LC190" class="blob-code blob-code-inner js-file-line">    <span class="pl-c1">\setlength</span>{<span class="pl-c1">\parskip</span>}{6pt plus 2pt minus 1pt}}</td>
      </tr>
      <tr>
        <td id="L191" class="blob-num js-line-number" data-line-number="191"></td>
        <td id="LC191" class="blob-code blob-code-inner js-file-line">}{<span class="pl-c"><span class="pl-c">%</span> if KOMA class</span></td>
      </tr>
      <tr>
        <td id="L192" class="blob-num js-line-number" data-line-number="192"></td>
        <td id="LC192" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\KOMAoptions</span>{parskip=half}}</td>
      </tr>
      <tr>
        <td id="L193" class="blob-num js-line-number" data-line-number="193"></td>
        <td id="LC193" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatother</span></td>
      </tr>
      <tr>
        <td id="L194" class="blob-num js-line-number" data-line-number="194"></td>
        <td id="LC194" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L195" class="blob-num js-line-number" data-line-number="195"></td>
        <td id="LC195" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(verbatim-in-note)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L196" class="blob-num js-line-number" data-line-number="196"></td>
        <td id="LC196" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">fancyvrb</span>}</td>
      </tr>
      <tr>
        <td id="L197" class="blob-num js-line-number" data-line-number="197"></td>
        <td id="LC197" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L198" class="blob-num js-line-number" data-line-number="198"></td>
        <td id="LC198" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">xcolor</span>}</td>
      </tr>
      <tr>
        <td id="L199" class="blob-num js-line-number" data-line-number="199"></td>
        <td id="LC199" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\IfFileExists</span>{xurl.sty}{<span class="pl-k">\usepackage</span>{<span class="pl-c1">xurl</span>}}{} <span class="pl-c"><span class="pl-c">%</span> add URL line breaks if available</span></td>
      </tr>
      <tr>
        <td id="L200" class="blob-num js-line-number" data-line-number="200"></td>
        <td id="LC200" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\IfFileExists</span>{bookmark.sty}{<span class="pl-k">\usepackage</span>{<span class="pl-c1">bookmark</span>}}{<span class="pl-k">\usepackage</span>{<span class="pl-c1">hyperref</span>}}</td>
      </tr>
      <tr>
        <td id="L201" class="blob-num js-line-number" data-line-number="201"></td>
        <td id="LC201" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\hypersetup</span>{</td>
      </tr>
      <tr>
        <td id="L202" class="blob-num js-line-number" data-line-number="202"></td>
        <td id="LC202" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(title-meta)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L203" class="blob-num js-line-number" data-line-number="203"></td>
        <td id="LC203" class="blob-code blob-code-inner js-file-line">  pdftitle={<span class="pl-s"><span class="pl-pds">$</span>title-meta<span class="pl-pds">$</span></span>},</td>
      </tr>
      <tr>
        <td id="L204" class="blob-num js-line-number" data-line-number="204"></td>
        <td id="LC204" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L205" class="blob-num js-line-number" data-line-number="205"></td>
        <td id="LC205" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(author-meta)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L206" class="blob-num js-line-number" data-line-number="206"></td>
        <td id="LC206" class="blob-code blob-code-inner js-file-line">  pdfauthor={<span class="pl-s"><span class="pl-pds">$</span>author-meta<span class="pl-pds">$</span></span>},</td>
      </tr>
      <tr>
        <td id="L207" class="blob-num js-line-number" data-line-number="207"></td>
        <td id="LC207" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L208" class="blob-num js-line-number" data-line-number="208"></td>
        <td id="LC208" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(lang)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L209" class="blob-num js-line-number" data-line-number="209"></td>
        <td id="LC209" class="blob-code blob-code-inner js-file-line">  pdflang={<span class="pl-s"><span class="pl-pds">$</span>lang<span class="pl-pds">$</span></span>},</td>
      </tr>
      <tr>
        <td id="L210" class="blob-num js-line-number" data-line-number="210"></td>
        <td id="LC210" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L211" class="blob-num js-line-number" data-line-number="211"></td>
        <td id="LC211" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(subject)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L212" class="blob-num js-line-number" data-line-number="212"></td>
        <td id="LC212" class="blob-code blob-code-inner js-file-line">  pdfsubject={<span class="pl-s"><span class="pl-pds">$</span>subject<span class="pl-pds">$</span></span>},</td>
      </tr>
      <tr>
        <td id="L213" class="blob-num js-line-number" data-line-number="213"></td>
        <td id="LC213" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L214" class="blob-num js-line-number" data-line-number="214"></td>
        <td id="LC214" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(keywords)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L215" class="blob-num js-line-number" data-line-number="215"></td>
        <td id="LC215" class="blob-code blob-code-inner js-file-line">  pdfkeywords={<span class="pl-s"><span class="pl-pds">$</span>for(keywords)<span class="pl-pds">$</span><span class="pl-pds">$</span>keywords<span class="pl-pds">$</span><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span>, <span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span>},</td>
      </tr>
      <tr>
        <td id="L216" class="blob-num js-line-number" data-line-number="216"></td>
        <td id="LC216" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L217" class="blob-num js-line-number" data-line-number="217"></td>
        <td id="LC217" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(colorlinks)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L218" class="blob-num js-line-number" data-line-number="218"></td>
        <td id="LC218" class="blob-code blob-code-inner js-file-line">  colorlinks=true,</td>
      </tr>
      <tr>
        <td id="L219" class="blob-num js-line-number" data-line-number="219"></td>
        <td id="LC219" class="blob-code blob-code-inner js-file-line">  linkcolor=<span class="pl-s"><span class="pl-pds">$</span>if(linkcolor)<span class="pl-pds">$</span><span class="pl-pds">$</span>linkcolor<span class="pl-pds">$</span><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span>Maroon<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span>,</td>
      </tr>
      <tr>
        <td id="L220" class="blob-num js-line-number" data-line-number="220"></td>
        <td id="LC220" class="blob-code blob-code-inner js-file-line">  filecolor=<span class="pl-s"><span class="pl-pds">$</span>if(filecolor)<span class="pl-pds">$</span><span class="pl-pds">$</span>filecolor<span class="pl-pds">$</span><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span>Maroon<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span>,</td>
      </tr>
      <tr>
        <td id="L221" class="blob-num js-line-number" data-line-number="221"></td>
        <td id="LC221" class="blob-code blob-code-inner js-file-line">  citecolor=<span class="pl-s"><span class="pl-pds">$</span>if(citecolor)<span class="pl-pds">$</span><span class="pl-pds">$</span>citecolor<span class="pl-pds">$</span><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span>Blue<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span>,</td>
      </tr>
      <tr>
        <td id="L222" class="blob-num js-line-number" data-line-number="222"></td>
        <td id="LC222" class="blob-code blob-code-inner js-file-line">  urlcolor=<span class="pl-s"><span class="pl-pds">$</span>if(urlcolor)<span class="pl-pds">$</span><span class="pl-pds">$</span>urlcolor<span class="pl-pds">$</span><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span>Blue<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span>,</td>
      </tr>
      <tr>
        <td id="L223" class="blob-num js-line-number" data-line-number="223"></td>
        <td id="LC223" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L224" class="blob-num js-line-number" data-line-number="224"></td>
        <td id="LC224" class="blob-code blob-code-inner js-file-line">  hidelinks,</td>
      </tr>
      <tr>
        <td id="L225" class="blob-num js-line-number" data-line-number="225"></td>
        <td id="LC225" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L226" class="blob-num js-line-number" data-line-number="226"></td>
        <td id="LC226" class="blob-code blob-code-inner js-file-line">  pdfcreator={LaTeX via pandoc}}</td>
      </tr>
      <tr>
        <td id="L227" class="blob-num js-line-number" data-line-number="227"></td>
        <td id="LC227" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\urlstyle</span>{same} <span class="pl-c"><span class="pl-c">%</span> disable monospaced font for URLs</span></td>
      </tr>
      <tr>
        <td id="L228" class="blob-num js-line-number" data-line-number="228"></td>
        <td id="LC228" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(verbatim-in-note)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L229" class="blob-num js-line-number" data-line-number="229"></td>
        <td id="LC229" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\VerbatimFootnotes</span> <span class="pl-c"><span class="pl-c">%</span> allow verbatim text in footnotes</span></td>
      </tr>
      <tr>
        <td id="L230" class="blob-num js-line-number" data-line-number="230"></td>
        <td id="LC230" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L231" class="blob-num js-line-number" data-line-number="231"></td>
        <td id="LC231" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(geometry)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L232" class="blob-num js-line-number" data-line-number="232"></td>
        <td id="LC232" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(beamer)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L233" class="blob-num js-line-number" data-line-number="233"></td>
        <td id="LC233" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\geometry</span>{<span class="pl-s"><span class="pl-pds">$</span>for(geometry)<span class="pl-pds">$</span><span class="pl-pds">$</span>geometry<span class="pl-pds">$</span><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span>,<span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L234" class="blob-num js-line-number" data-line-number="234"></td>
        <td id="LC234" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L235" class="blob-num js-line-number" data-line-number="235"></td>
        <td id="LC235" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>[<span class="pl-v">$for(geometry)$$geometry$$sep$,$endfor$</span>]{<span class="pl-c1">geometry</span>}</td>
      </tr>
      <tr>
        <td id="L236" class="blob-num js-line-number" data-line-number="236"></td>
        <td id="LC236" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L237" class="blob-num js-line-number" data-line-number="237"></td>
        <td id="LC237" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L238" class="blob-num js-line-number" data-line-number="238"></td>
        <td id="LC238" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(beamer)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L239" class="blob-num js-line-number" data-line-number="239"></td>
        <td id="LC239" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\newif\ifbibliography</span></td>
      </tr>
      <tr>
        <td id="L240" class="blob-num js-line-number" data-line-number="240"></td>
        <td id="LC240" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L241" class="blob-num js-line-number" data-line-number="241"></td>
        <td id="LC241" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(listings)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L242" class="blob-num js-line-number" data-line-number="242"></td>
        <td id="LC242" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">listings</span>}</td>
      </tr>
      <tr>
        <td id="L243" class="blob-num js-line-number" data-line-number="243"></td>
        <td id="LC243" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\newcommand</span>{<span class="pl-c1">\passthrough</span>}[1]{#1}</td>
      </tr>
      <tr>
        <td id="L244" class="blob-num js-line-number" data-line-number="244"></td>
        <td id="LC244" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\lstset</span>{defaultdialect=[5.3]Lua}</td>
      </tr>
      <tr>
        <td id="L245" class="blob-num js-line-number" data-line-number="245"></td>
        <td id="LC245" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\lstset</span>{defaultdialect=[x86masm]Assembler}</td>
      </tr>
      <tr>
        <td id="L246" class="blob-num js-line-number" data-line-number="246"></td>
        <td id="LC246" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L247" class="blob-num js-line-number" data-line-number="247"></td>
        <td id="LC247" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(lhs)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L248" class="blob-num js-line-number" data-line-number="248"></td>
        <td id="LC248" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\lstnewenvironment</span>{code}{<span class="pl-c1">\lstset</span>{language=Haskell,basicstyle=<span class="pl-c1">\small\ttfamily</span>}}{}</td>
      </tr>
      <tr>
        <td id="L249" class="blob-num js-line-number" data-line-number="249"></td>
        <td id="LC249" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L250" class="blob-num js-line-number" data-line-number="250"></td>
        <td id="LC250" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(highlighting-macros)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L251" class="blob-num js-line-number" data-line-number="251"></td>
        <td id="LC251" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>highlighting-macros<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L252" class="blob-num js-line-number" data-line-number="252"></td>
        <td id="LC252" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L253" class="blob-num js-line-number" data-line-number="253"></td>
        <td id="LC253" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(tables)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L254" class="blob-num js-line-number" data-line-number="254"></td>
        <td id="LC254" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">longtable,booktabs</span>}</td>
      </tr>
      <tr>
        <td id="L255" class="blob-num js-line-number" data-line-number="255"></td>
        <td id="LC255" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(beamer)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L256" class="blob-num js-line-number" data-line-number="256"></td>
        <td id="LC256" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">caption</span>}</td>
      </tr>
      <tr>
        <td id="L257" class="blob-num js-line-number" data-line-number="257"></td>
        <td id="LC257" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> Make caption package work with longtable</span></td>
      </tr>
      <tr>
        <td id="L258" class="blob-num js-line-number" data-line-number="258"></td>
        <td id="LC258" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatletter</span></td>
      </tr>
      <tr>
        <td id="L259" class="blob-num js-line-number" data-line-number="259"></td>
        <td id="LC259" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\def\fnum@table</span>{<span class="pl-c1">\tablename</span>~<span class="pl-c1">\thetable</span>}</td>
      </tr>
      <tr>
        <td id="L260" class="blob-num js-line-number" data-line-number="260"></td>
        <td id="LC260" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatother</span></td>
      </tr>
      <tr>
        <td id="L261" class="blob-num js-line-number" data-line-number="261"></td>
        <td id="LC261" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L262" class="blob-num js-line-number" data-line-number="262"></td>
        <td id="LC262" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> Correct order of tables after \paragraph or \subparagraph</span></td>
      </tr>
      <tr>
        <td id="L263" class="blob-num js-line-number" data-line-number="263"></td>
        <td id="LC263" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">etoolbox</span>}</td>
      </tr>
      <tr>
        <td id="L264" class="blob-num js-line-number" data-line-number="264"></td>
        <td id="LC264" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatletter</span></td>
      </tr>
      <tr>
        <td id="L265" class="blob-num js-line-number" data-line-number="265"></td>
        <td id="LC265" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\patchcmd\longtable</span>{<span class="pl-c1">\par</span>}{<span class="pl-k">\if</span>@noskipsec<span class="pl-c1">\mbox</span>{}<span class="pl-k">\fi</span><span class="pl-c1">\par</span>}{}{}</td>
      </tr>
      <tr>
        <td id="L266" class="blob-num js-line-number" data-line-number="266"></td>
        <td id="LC266" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatother</span></td>
      </tr>
      <tr>
        <td id="L267" class="blob-num js-line-number" data-line-number="267"></td>
        <td id="LC267" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> Allow footnotes in longtable head/foot</span></td>
      </tr>
      <tr>
        <td id="L268" class="blob-num js-line-number" data-line-number="268"></td>
        <td id="LC268" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\IfFileExists</span>{footnotehyper.sty}{<span class="pl-k">\usepackage</span>{<span class="pl-c1">footnotehyper</span>}}{<span class="pl-k">\usepackage</span>{<span class="pl-c1">footnote</span>}}</td>
      </tr>
      <tr>
        <td id="L269" class="blob-num js-line-number" data-line-number="269"></td>
        <td id="LC269" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makesavenoteenv</span>{longtable}</td>
      </tr>
      <tr>
        <td id="L270" class="blob-num js-line-number" data-line-number="270"></td>
        <td id="LC270" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L271" class="blob-num js-line-number" data-line-number="271"></td>
        <td id="LC271" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L272" class="blob-num js-line-number" data-line-number="272"></td>
        <td id="LC272" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(graphics)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L273" class="blob-num js-line-number" data-line-number="273"></td>
        <td id="LC273" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">graphicx,grffile</span>}</td>
      </tr>
      <tr>
        <td id="L274" class="blob-num js-line-number" data-line-number="274"></td>
        <td id="LC274" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatletter</span></td>
      </tr>
      <tr>
        <td id="L275" class="blob-num js-line-number" data-line-number="275"></td>
        <td id="LC275" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\def\maxwidth</span>{<span class="pl-k">\ifdim</span><span class="pl-c1">\Gin@nat@width</span>&gt;<span class="pl-c1">\linewidth\linewidth</span><span class="pl-k">\else</span><span class="pl-c1">\Gin@nat@width</span><span class="pl-k">\fi</span>}</td>
      </tr>
      <tr>
        <td id="L276" class="blob-num js-line-number" data-line-number="276"></td>
        <td id="LC276" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\def\maxheight</span>{<span class="pl-k">\ifdim</span><span class="pl-c1">\Gin@nat@height</span>&gt;<span class="pl-c1">\textheight\textheight</span><span class="pl-k">\else</span><span class="pl-c1">\Gin@nat@height</span><span class="pl-k">\fi</span>}</td>
      </tr>
      <tr>
        <td id="L277" class="blob-num js-line-number" data-line-number="277"></td>
        <td id="LC277" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatother</span></td>
      </tr>
      <tr>
        <td id="L278" class="blob-num js-line-number" data-line-number="278"></td>
        <td id="LC278" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> Scale images if necessary, so that they will not overflow the page</span></td>
      </tr>
      <tr>
        <td id="L279" class="blob-num js-line-number" data-line-number="279"></td>
        <td id="LC279" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> margins by default, and it is still possible to overwrite the defaults</span></td>
      </tr>
      <tr>
        <td id="L280" class="blob-num js-line-number" data-line-number="280"></td>
        <td id="LC280" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> using explicit options in \includegraphics[width, height, ...]{}</span></td>
      </tr>
      <tr>
        <td id="L281" class="blob-num js-line-number" data-line-number="281"></td>
        <td id="LC281" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\setkeys</span>{Gin}{width=<span class="pl-c1">\maxwidth</span>,height=<span class="pl-c1">\maxheight</span>,keepaspectratio}</td>
      </tr>
      <tr>
        <td id="L282" class="blob-num js-line-number" data-line-number="282"></td>
        <td id="LC282" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> Set default figure placement to htbp</span></td>
      </tr>
      <tr>
        <td id="L283" class="blob-num js-line-number" data-line-number="283"></td>
        <td id="LC283" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatletter</span></td>
      </tr>
      <tr>
        <td id="L284" class="blob-num js-line-number" data-line-number="284"></td>
        <td id="LC284" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\def\fps@figure</span>{htbp}</td>
      </tr>
      <tr>
        <td id="L285" class="blob-num js-line-number" data-line-number="285"></td>
        <td id="LC285" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatother</span></td>
      </tr>
      <tr>
        <td id="L286" class="blob-num js-line-number" data-line-number="286"></td>
        <td id="LC286" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L287" class="blob-num js-line-number" data-line-number="287"></td>
        <td id="LC287" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(links-as-notes)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L288" class="blob-num js-line-number" data-line-number="288"></td>
        <td id="LC288" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> Make links footnotes instead of hotlinks:</span></td>
      </tr>
      <tr>
        <td id="L289" class="blob-num js-line-number" data-line-number="289"></td>
        <td id="LC289" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\DeclareRobustCommand</span>{<span class="pl-c1">\href</span>}[2]{#2<span class="pl-c1">\footnote</span>{<span class="pl-c1">\url</span>{#1}}}</td>
      </tr>
      <tr>
        <td id="L290" class="blob-num js-line-number" data-line-number="290"></td>
        <td id="LC290" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L291" class="blob-num js-line-number" data-line-number="291"></td>
        <td id="LC291" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(strikeout)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L292" class="blob-num js-line-number" data-line-number="292"></td>
        <td id="LC292" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>[<span class="pl-v">normalem</span>]{<span class="pl-c1">ulem</span>}</td>
      </tr>
      <tr>
        <td id="L293" class="blob-num js-line-number" data-line-number="293"></td>
        <td id="LC293" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> Avoid problems with \sout in headers with hyperref</span></td>
      </tr>
      <tr>
        <td id="L294" class="blob-num js-line-number" data-line-number="294"></td>
        <td id="LC294" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\pdfstringdefDisableCommands</span>{<span class="pl-k">\renewcommand</span>{<span class="pl-c1">\sout</span>}{}}</td>
      </tr>
      <tr>
        <td id="L295" class="blob-num js-line-number" data-line-number="295"></td>
        <td id="LC295" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L296" class="blob-num js-line-number" data-line-number="296"></td>
        <td id="LC296" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\setlength</span>{<span class="pl-c1">\emergencystretch</span>}{3em} <span class="pl-c"><span class="pl-c">%</span> prevent overfull lines</span></td>
      </tr>
      <tr>
        <td id="L297" class="blob-num js-line-number" data-line-number="297"></td>
        <td id="LC297" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\providecommand</span>{<span class="pl-c1">\tightlist</span>}{<span class="pl-c"><span class="pl-c">%</span></span></td>
      </tr>
      <tr>
        <td id="L298" class="blob-num js-line-number" data-line-number="298"></td>
        <td id="LC298" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\setlength</span>{<span class="pl-c1">\itemsep</span>}{0pt}<span class="pl-c1">\setlength</span>{<span class="pl-c1">\parskip</span>}{0pt}}</td>
      </tr>
      <tr>
        <td id="L299" class="blob-num js-line-number" data-line-number="299"></td>
        <td id="LC299" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(numbersections)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L300" class="blob-num js-line-number" data-line-number="300"></td>
        <td id="LC300" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\setcounter</span>{secnumdepth}{<span class="pl-s"><span class="pl-pds">$</span>if(secnumdepth)<span class="pl-pds">$</span><span class="pl-pds">$</span>secnumdepth<span class="pl-pds">$</span><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span>5<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L301" class="blob-num js-line-number" data-line-number="301"></td>
        <td id="LC301" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L302" class="blob-num js-line-number" data-line-number="302"></td>
        <td id="LC302" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\setcounter</span>{secnumdepth}{-<span class="pl-c1">\maxdimen</span>} <span class="pl-c"><span class="pl-c">%</span> remove section numbering</span></td>
      </tr>
      <tr>
        <td id="L303" class="blob-num js-line-number" data-line-number="303"></td>
        <td id="LC303" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L304" class="blob-num js-line-number" data-line-number="304"></td>
        <td id="LC304" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(beamer)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L305" class="blob-num js-line-number" data-line-number="305"></td>
        <td id="LC305" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L306" class="blob-num js-line-number" data-line-number="306"></td>
        <td id="LC306" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(block-headings)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L307" class="blob-num js-line-number" data-line-number="307"></td>
        <td id="LC307" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> Make \paragraph and \subparagraph free-standing</span></td>
      </tr>
      <tr>
        <td id="L308" class="blob-num js-line-number" data-line-number="308"></td>
        <td id="LC308" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\ifx</span><span class="pl-c1">\paragraph\undefined</span><span class="pl-k">\else</span></td>
      </tr>
      <tr>
        <td id="L309" class="blob-num js-line-number" data-line-number="309"></td>
        <td id="LC309" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\let\oldparagraph\paragraph</span></td>
      </tr>
      <tr>
        <td id="L310" class="blob-num js-line-number" data-line-number="310"></td>
        <td id="LC310" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\renewcommand</span>{<span class="pl-c1">\paragraph</span>}[1]{<span class="pl-c1">\oldparagraph</span>{#1}<span class="pl-c1">\mbox</span>{}}</td>
      </tr>
      <tr>
        <td id="L311" class="blob-num js-line-number" data-line-number="311"></td>
        <td id="LC311" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\fi</span></td>
      </tr>
      <tr>
        <td id="L312" class="blob-num js-line-number" data-line-number="312"></td>
        <td id="LC312" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\ifx</span><span class="pl-c1">\subparagraph\undefined</span><span class="pl-k">\else</span></td>
      </tr>
      <tr>
        <td id="L313" class="blob-num js-line-number" data-line-number="313"></td>
        <td id="LC313" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\let\oldsubparagraph\subparagraph</span></td>
      </tr>
      <tr>
        <td id="L314" class="blob-num js-line-number" data-line-number="314"></td>
        <td id="LC314" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\renewcommand</span>{<span class="pl-c1">\subparagraph</span>}[1]{<span class="pl-c1">\oldsubparagraph</span>{#1}<span class="pl-c1">\mbox</span>{}}</td>
      </tr>
      <tr>
        <td id="L315" class="blob-num js-line-number" data-line-number="315"></td>
        <td id="LC315" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\fi</span></td>
      </tr>
      <tr>
        <td id="L316" class="blob-num js-line-number" data-line-number="316"></td>
        <td id="LC316" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L317" class="blob-num js-line-number" data-line-number="317"></td>
        <td id="LC317" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L318" class="blob-num js-line-number" data-line-number="318"></td>
        <td id="LC318" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(pagestyle)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L319" class="blob-num js-line-number" data-line-number="319"></td>
        <td id="LC319" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\pagestyle</span>{<span class="pl-s"><span class="pl-pds">$</span>pagestyle<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L320" class="blob-num js-line-number" data-line-number="320"></td>
        <td id="LC320" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L321" class="blob-num js-line-number" data-line-number="321"></td>
        <td id="LC321" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>for(header-includes)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L322" class="blob-num js-line-number" data-line-number="322"></td>
        <td id="LC322" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>header-includes<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L323" class="blob-num js-line-number" data-line-number="323"></td>
        <td id="LC323" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L324" class="blob-num js-line-number" data-line-number="324"></td>
        <td id="LC324" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(lang)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L325" class="blob-num js-line-number" data-line-number="325"></td>
        <td id="LC325" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\ifxetex</span></td>
      </tr>
      <tr>
        <td id="L326" class="blob-num js-line-number" data-line-number="326"></td>
        <td id="LC326" class="blob-code blob-code-inner js-file-line">  <span class="pl-c"><span class="pl-c">%</span> Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)</span></td>
      </tr>
      <tr>
        <td id="L327" class="blob-num js-line-number" data-line-number="327"></td>
        <td id="LC327" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\usepackage</span>{<span class="pl-c1">polyglossia</span>}</td>
      </tr>
      <tr>
        <td id="L328" class="blob-num js-line-number" data-line-number="328"></td>
        <td id="LC328" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\setmainlanguage</span>[<span class="pl-s"><span class="pl-pds">$</span>polyglossia-lang.options<span class="pl-pds">$</span></span>]{<span class="pl-s"><span class="pl-pds">$</span>polyglossia-lang.name<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L329" class="blob-num js-line-number" data-line-number="329"></td>
        <td id="LC329" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>for(polyglossia-otherlangs)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L330" class="blob-num js-line-number" data-line-number="330"></td>
        <td id="LC330" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\setotherlanguage</span>[<span class="pl-s"><span class="pl-pds">$</span>polyglossia-otherlangs.options<span class="pl-pds">$</span></span>]{<span class="pl-s"><span class="pl-pds">$</span>polyglossia-otherlangs.name<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L331" class="blob-num js-line-number" data-line-number="331"></td>
        <td id="LC331" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L332" class="blob-num js-line-number" data-line-number="332"></td>
        <td id="LC332" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\else</span></td>
      </tr>
      <tr>
        <td id="L333" class="blob-num js-line-number" data-line-number="333"></td>
        <td id="LC333" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\usepackage</span>[<span class="pl-v">shorthands=off,$for(babel-otherlangs)$$babel-otherlangs$,$endfor$main=$babel-lang$</span>]{<span class="pl-c1">babel</span>}</td>
      </tr>
      <tr>
        <td id="L334" class="blob-num js-line-number" data-line-number="334"></td>
        <td id="LC334" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(babel-newcommands)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L335" class="blob-num js-line-number" data-line-number="335"></td>
        <td id="LC335" class="blob-code blob-code-inner js-file-line">  <span class="pl-s"><span class="pl-pds">$</span>babel-newcommands<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L336" class="blob-num js-line-number" data-line-number="336"></td>
        <td id="LC336" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L337" class="blob-num js-line-number" data-line-number="337"></td>
        <td id="LC337" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\fi</span></td>
      </tr>
      <tr>
        <td id="L338" class="blob-num js-line-number" data-line-number="338"></td>
        <td id="LC338" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L339" class="blob-num js-line-number" data-line-number="339"></td>
        <td id="LC339" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(dir)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L340" class="blob-num js-line-number" data-line-number="340"></td>
        <td id="LC340" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\ifxetex</span></td>
      </tr>
      <tr>
        <td id="L341" class="blob-num js-line-number" data-line-number="341"></td>
        <td id="LC341" class="blob-code blob-code-inner js-file-line">  <span class="pl-c"><span class="pl-c">%</span> Load bidi as late as possible as it modifies e.g. graphicx</span></td>
      </tr>
      <tr>
        <td id="L342" class="blob-num js-line-number" data-line-number="342"></td>
        <td id="LC342" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\usepackage</span>{<span class="pl-c1">bidi</span>}</td>
      </tr>
      <tr>
        <td id="L343" class="blob-num js-line-number" data-line-number="343"></td>
        <td id="LC343" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\fi</span></td>
      </tr>
      <tr>
        <td id="L344" class="blob-num js-line-number" data-line-number="344"></td>
        <td id="LC344" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\ifnum</span> 0<span class="pl-c1">\ifxetex</span> 1<span class="pl-k">\fi</span><span class="pl-c1">\ifluatex</span> 1<span class="pl-k">\fi</span>=0 <span class="pl-c"><span class="pl-c">%</span> if pdftex</span></td>
      </tr>
      <tr>
        <td id="L345" class="blob-num js-line-number" data-line-number="345"></td>
        <td id="LC345" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\TeXXeTstate</span>=1</td>
      </tr>
      <tr>
        <td id="L346" class="blob-num js-line-number" data-line-number="346"></td>
        <td id="LC346" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\newcommand</span>{<span class="pl-c1">\RL</span>}[1]{<span class="pl-c1">\beginR</span> #1<span class="pl-c1">\endR</span>}</td>
      </tr>
      <tr>
        <td id="L347" class="blob-num js-line-number" data-line-number="347"></td>
        <td id="LC347" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\newcommand</span>{<span class="pl-c1">\LR</span>}[1]{<span class="pl-c1">\beginL</span> #1<span class="pl-c1">\endL</span>}</td>
      </tr>
      <tr>
        <td id="L348" class="blob-num js-line-number" data-line-number="348"></td>
        <td id="LC348" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\newenvironment</span>{RTL}{<span class="pl-c1">\beginR</span>}{<span class="pl-c1">\endR</span>}</td>
      </tr>
      <tr>
        <td id="L349" class="blob-num js-line-number" data-line-number="349"></td>
        <td id="LC349" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\newenvironment</span>{LTR}{<span class="pl-c1">\beginL</span>}{<span class="pl-c1">\endL</span>}</td>
      </tr>
      <tr>
        <td id="L350" class="blob-num js-line-number" data-line-number="350"></td>
        <td id="LC350" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\fi</span></td>
      </tr>
      <tr>
        <td id="L351" class="blob-num js-line-number" data-line-number="351"></td>
        <td id="LC351" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L352" class="blob-num js-line-number" data-line-number="352"></td>
        <td id="LC352" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(natbib)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L353" class="blob-num js-line-number" data-line-number="353"></td>
        <td id="LC353" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>[<span class="pl-v">$natbiboptions$</span>]{<span class="pl-c1">natbib</span>}</td>
      </tr>
      <tr>
        <td id="L354" class="blob-num js-line-number" data-line-number="354"></td>
        <td id="LC354" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\bibliographystyle</span>{<span class="pl-s"><span class="pl-pds">$</span>if(biblio-style)<span class="pl-pds">$</span><span class="pl-pds">$</span>biblio-style<span class="pl-pds">$</span><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span>plainnat<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L355" class="blob-num js-line-number" data-line-number="355"></td>
        <td id="LC355" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L356" class="blob-num js-line-number" data-line-number="356"></td>
        <td id="LC356" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(biblatex)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L357" class="blob-num js-line-number" data-line-number="357"></td>
        <td id="LC357" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>[<span class="pl-v">$if(biblio-style)$style=$biblio-style$,$endif$$for(biblatexoptions)$$biblatexoptions$$sep$,$endfor$</span>]{<span class="pl-c1">biblatex</span>}</td>
      </tr>
      <tr>
        <td id="L358" class="blob-num js-line-number" data-line-number="358"></td>
        <td id="LC358" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>for(bibliography)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L359" class="blob-num js-line-number" data-line-number="359"></td>
        <td id="LC359" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\addbibresource</span>{<span class="pl-s"><span class="pl-pds">$</span>bibliography<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L360" class="blob-num js-line-number" data-line-number="360"></td>
        <td id="LC360" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L361" class="blob-num js-line-number" data-line-number="361"></td>
        <td id="LC361" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L362" class="blob-num js-line-number" data-line-number="362"></td>
        <td id="LC362" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L363" class="blob-num js-line-number" data-line-number="363"></td>
        <td id="LC363" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(title)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L364" class="blob-num js-line-number" data-line-number="364"></td>
        <td id="LC364" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\title</span>{<span class="pl-s"><span class="pl-pds">$</span>title<span class="pl-pds">$</span><span class="pl-pds">$</span>if(thanks)<span class="pl-pds">$</span></span><span class="pl-c1">\thanks</span>{<span class="pl-s"><span class="pl-pds">$</span>thanks<span class="pl-pds">$</span></span>}<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L365" class="blob-num js-line-number" data-line-number="365"></td>
        <td id="LC365" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L366" class="blob-num js-line-number" data-line-number="366"></td>
        <td id="LC366" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(subtitle)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L367" class="blob-num js-line-number" data-line-number="367"></td>
        <td id="LC367" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(beamer)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L368" class="blob-num js-line-number" data-line-number="368"></td>
        <td id="LC368" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L369" class="blob-num js-line-number" data-line-number="369"></td>
        <td id="LC369" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">etoolbox</span>}</td>
      </tr>
      <tr>
        <td id="L370" class="blob-num js-line-number" data-line-number="370"></td>
        <td id="LC370" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatletter</span></td>
      </tr>
      <tr>
        <td id="L371" class="blob-num js-line-number" data-line-number="371"></td>
        <td id="LC371" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\providecommand</span>{<span class="pl-c1">\subtitle</span>}[1]{<span class="pl-c"><span class="pl-c">%</span> add subtitle to \maketitle</span></td>
      </tr>
      <tr>
        <td id="L372" class="blob-num js-line-number" data-line-number="372"></td>
        <td id="LC372" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\apptocmd</span>{<span class="pl-c1">\@title</span>}{<span class="pl-c1">\par</span> {<span class="pl-c1">\large</span> #1 <span class="pl-c1">\par</span>}}{}{}</td>
      </tr>
      <tr>
        <td id="L373" class="blob-num js-line-number" data-line-number="373"></td>
        <td id="LC373" class="blob-code blob-code-inner js-file-line">}</td>
      </tr>
      <tr>
        <td id="L374" class="blob-num js-line-number" data-line-number="374"></td>
        <td id="LC374" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatother</span></td>
      </tr>
      <tr>
        <td id="L375" class="blob-num js-line-number" data-line-number="375"></td>
        <td id="LC375" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L376" class="blob-num js-line-number" data-line-number="376"></td>
        <td id="LC376" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\subtitle</span>{<span class="pl-s"><span class="pl-pds">$</span>subtitle<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L377" class="blob-num js-line-number" data-line-number="377"></td>
        <td id="LC377" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L378" class="blob-num js-line-number" data-line-number="378"></td>
        <td id="LC378" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L379" class="blob-num js-line-number" data-line-number="379"></td>
        <td id="LC379" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>$if(author)$</span></td>
      </tr>
      <tr>
        <td id="L380" class="blob-num js-line-number" data-line-number="380"></td>
        <td id="LC380" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>\author{$for(author)$$author$$sep$ \and $endfor$}</span></td>
      </tr>
      <tr>
        <td id="L381" class="blob-num js-line-number" data-line-number="381"></td>
        <td id="LC381" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>$endif$</span></td>
      </tr>
      <tr>
        <td id="L382" class="blob-num js-line-number" data-line-number="382"></td>
        <td id="LC382" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L383" class="blob-num js-line-number" data-line-number="383"></td>
        <td id="LC383" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(author)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L384" class="blob-num js-line-number" data-line-number="384"></td>
        <td id="LC384" class="blob-code blob-code-inner js-file-line">	<span class="pl-k">\usepackage</span>{<span class="pl-c1">authblk</span>}</td>
      </tr>
      <tr>
        <td id="L385" class="blob-num js-line-number" data-line-number="385"></td>
        <td id="LC385" class="blob-code blob-code-inner js-file-line">	<span class="pl-s"><span class="pl-pds">$</span>for(author)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L386" class="blob-num js-line-number" data-line-number="386"></td>
        <td id="LC386" class="blob-code blob-code-inner js-file-line">		<span class="pl-s"><span class="pl-pds">$</span>if(author.name)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L387" class="blob-num js-line-number" data-line-number="387"></td>
        <td id="LC387" class="blob-code blob-code-inner js-file-line">			<span class="pl-s"><span class="pl-pds">$</span>if(author.number)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L388" class="blob-num js-line-number" data-line-number="388"></td>
        <td id="LC388" class="blob-code blob-code-inner js-file-line">				<span class="pl-c1">\author</span>[<span class="pl-s"><span class="pl-pds">$</span>author.number<span class="pl-pds">$</span></span>]{<span class="pl-s"><span class="pl-pds">$</span>author.name<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L389" class="blob-num js-line-number" data-line-number="389"></td>
        <td id="LC389" class="blob-code blob-code-inner js-file-line">			<span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L390" class="blob-num js-line-number" data-line-number="390"></td>
        <td id="LC390" class="blob-code blob-code-inner js-file-line">				<span class="pl-c1">\author</span>[]{<span class="pl-s"><span class="pl-pds">$</span>author.name<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L391" class="blob-num js-line-number" data-line-number="391"></td>
        <td id="LC391" class="blob-code blob-code-inner js-file-line">			<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L392" class="blob-num js-line-number" data-line-number="392"></td>
        <td id="LC392" class="blob-code blob-code-inner js-file-line">			<span class="pl-s"><span class="pl-pds">$</span>if(author.affiliation)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L393" class="blob-num js-line-number" data-line-number="393"></td>
        <td id="LC393" class="blob-code blob-code-inner js-file-line">				<span class="pl-s"><span class="pl-pds">$</span>if(author.email)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L394" class="blob-num js-line-number" data-line-number="394"></td>
        <td id="LC394" class="blob-code blob-code-inner js-file-line">					<span class="pl-c1">\affil</span>{<span class="pl-s"><span class="pl-pds">$</span>author.affiliation<span class="pl-pds">$</span></span> <span class="pl-c1">\thanks</span>{<span class="pl-s"><span class="pl-pds">$</span>author.email<span class="pl-pds">$</span></span>}}</td>
      </tr>
      <tr>
        <td id="L395" class="blob-num js-line-number" data-line-number="395"></td>
        <td id="LC395" class="blob-code blob-code-inner js-file-line">				<span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L396" class="blob-num js-line-number" data-line-number="396"></td>
        <td id="LC396" class="blob-code blob-code-inner js-file-line">					<span class="pl-c1">\affil</span>{<span class="pl-s"><span class="pl-pds">$</span>author.affiliation<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L397" class="blob-num js-line-number" data-line-number="397"></td>
        <td id="LC397" class="blob-code blob-code-inner js-file-line">				<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L398" class="blob-num js-line-number" data-line-number="398"></td>
        <td id="LC398" class="blob-code blob-code-inner js-file-line">			<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L399" class="blob-num js-line-number" data-line-number="399"></td>
        <td id="LC399" class="blob-code blob-code-inner js-file-line">			<span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span>	</td>
      </tr>
      <tr>
        <td id="L400" class="blob-num js-line-number" data-line-number="400"></td>
        <td id="LC400" class="blob-code blob-code-inner js-file-line">			<span class="pl-c1">\author</span>{<span class="pl-s"><span class="pl-pds">$</span>author<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L401" class="blob-num js-line-number" data-line-number="401"></td>
        <td id="LC401" class="blob-code blob-code-inner js-file-line">		<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L402" class="blob-num js-line-number" data-line-number="402"></td>
        <td id="LC402" class="blob-code blob-code-inner js-file-line">	<span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L403" class="blob-num js-line-number" data-line-number="403"></td>
        <td id="LC403" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L404" class="blob-num js-line-number" data-line-number="404"></td>
        <td id="LC404" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L405" class="blob-num js-line-number" data-line-number="405"></td>
        <td id="LC405" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\date</span>{<span class="pl-s"><span class="pl-pds">$</span>date<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L406" class="blob-num js-line-number" data-line-number="406"></td>
        <td id="LC406" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(beamer)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L407" class="blob-num js-line-number" data-line-number="407"></td>
        <td id="LC407" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(institute)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L408" class="blob-num js-line-number" data-line-number="408"></td>
        <td id="LC408" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\institute</span>{<span class="pl-s"><span class="pl-pds">$</span>for(institute)<span class="pl-pds">$</span><span class="pl-pds">$</span>institute<span class="pl-pds">$</span><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span> <span class="pl-c1">\and</span> <span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L409" class="blob-num js-line-number" data-line-number="409"></td>
        <td id="LC409" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L410" class="blob-num js-line-number" data-line-number="410"></td>
        <td id="LC410" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(titlegraphic)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L411" class="blob-num js-line-number" data-line-number="411"></td>
        <td id="LC411" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\titlegraphic</span>{<span class="pl-c1">\includegraphics</span>{<span class="pl-s"><span class="pl-pds">$</span>titlegraphic<span class="pl-pds">$</span></span>}}</td>
      </tr>
      <tr>
        <td id="L412" class="blob-num js-line-number" data-line-number="412"></td>
        <td id="LC412" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L413" class="blob-num js-line-number" data-line-number="413"></td>
        <td id="LC413" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(logo)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L414" class="blob-num js-line-number" data-line-number="414"></td>
        <td id="LC414" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\logo</span>{<span class="pl-c1">\includegraphics</span>{<span class="pl-s"><span class="pl-pds">$</span>logo<span class="pl-pds">$</span></span>}}</td>
      </tr>
      <tr>
        <td id="L415" class="blob-num js-line-number" data-line-number="415"></td>
        <td id="LC415" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L416" class="blob-num js-line-number" data-line-number="416"></td>
        <td id="LC416" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L417" class="blob-num js-line-number" data-line-number="417"></td>
        <td id="LC417" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L418" class="blob-num js-line-number" data-line-number="418"></td>
        <td id="LC418" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\begin</span>{<span class="pl-smi">document</span>}</td>
      </tr>
      <tr>
        <td id="L419" class="blob-num js-line-number" data-line-number="419"></td>
        <td id="LC419" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(has-frontmatter)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L420" class="blob-num js-line-number" data-line-number="420"></td>
        <td id="LC420" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\frontmatter</span></td>
      </tr>
      <tr>
        <td id="L421" class="blob-num js-line-number" data-line-number="421"></td>
        <td id="LC421" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L422" class="blob-num js-line-number" data-line-number="422"></td>
        <td id="LC422" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(title)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L423" class="blob-num js-line-number" data-line-number="423"></td>
        <td id="LC423" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(beamer)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L424" class="blob-num js-line-number" data-line-number="424"></td>
        <td id="LC424" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\frame</span>{<span class="pl-c1">\titlepage</span>}</td>
      </tr>
      <tr>
        <td id="L425" class="blob-num js-line-number" data-line-number="425"></td>
        <td id="LC425" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L426" class="blob-num js-line-number" data-line-number="426"></td>
        <td id="LC426" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\maketitle</span></td>
      </tr>
      <tr>
        <td id="L427" class="blob-num js-line-number" data-line-number="427"></td>
        <td id="LC427" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L428" class="blob-num js-line-number" data-line-number="428"></td>
        <td id="LC428" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(abstract)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L429" class="blob-num js-line-number" data-line-number="429"></td>
        <td id="LC429" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\begin</span>{<span class="pl-smi">abstract</span>}</td>
      </tr>
      <tr>
        <td id="L430" class="blob-num js-line-number" data-line-number="430"></td>
        <td id="LC430" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>abstract<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L431" class="blob-num js-line-number" data-line-number="431"></td>
        <td id="LC431" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\end</span>{<span class="pl-smi">abstract</span>}</td>
      </tr>
      <tr>
        <td id="L432" class="blob-num js-line-number" data-line-number="432"></td>
        <td id="LC432" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L433" class="blob-num js-line-number" data-line-number="433"></td>
        <td id="LC433" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L434" class="blob-num js-line-number" data-line-number="434"></td>
        <td id="LC434" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L435" class="blob-num js-line-number" data-line-number="435"></td>
        <td id="LC435" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>for(include-before)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L436" class="blob-num js-line-number" data-line-number="436"></td>
        <td id="LC436" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>include-before<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L437" class="blob-num js-line-number" data-line-number="437"></td>
        <td id="LC437" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L438" class="blob-num js-line-number" data-line-number="438"></td>
        <td id="LC438" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L439" class="blob-num js-line-number" data-line-number="439"></td>
        <td id="LC439" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(toc)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L440" class="blob-num js-line-number" data-line-number="440"></td>
        <td id="LC440" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(toc-title)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L441" class="blob-num js-line-number" data-line-number="441"></td>
        <td id="LC441" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\renewcommand</span>*<span class="pl-c1">\contentsname</span>{<span class="pl-s"><span class="pl-pds">$</span>toc-title<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L442" class="blob-num js-line-number" data-line-number="442"></td>
        <td id="LC442" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L443" class="blob-num js-line-number" data-line-number="443"></td>
        <td id="LC443" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(beamer)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L444" class="blob-num js-line-number" data-line-number="444"></td>
        <td id="LC444" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\begin</span>{<span class="pl-smi">frame</span>}</td>
      </tr>
      <tr>
        <td id="L445" class="blob-num js-line-number" data-line-number="445"></td>
        <td id="LC445" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(toc-title)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L446" class="blob-num js-line-number" data-line-number="446"></td>
        <td id="LC446" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\frametitle</span>{<span class="pl-s"><span class="pl-pds">$</span>toc-title<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L447" class="blob-num js-line-number" data-line-number="447"></td>
        <td id="LC447" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L448" class="blob-num js-line-number" data-line-number="448"></td>
        <td id="LC448" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\tableofcontents</span>[hideallsubsections]</td>
      </tr>
      <tr>
        <td id="L449" class="blob-num js-line-number" data-line-number="449"></td>
        <td id="LC449" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\end</span>{<span class="pl-smi">frame</span>}</td>
      </tr>
      <tr>
        <td id="L450" class="blob-num js-line-number" data-line-number="450"></td>
        <td id="LC450" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L451" class="blob-num js-line-number" data-line-number="451"></td>
        <td id="LC451" class="blob-code blob-code-inner js-file-line">{</td>
      </tr>
      <tr>
        <td id="L452" class="blob-num js-line-number" data-line-number="452"></td>
        <td id="LC452" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(colorlinks)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L453" class="blob-num js-line-number" data-line-number="453"></td>
        <td id="LC453" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\hypersetup</span>{linkcolor=<span class="pl-s"><span class="pl-pds">$</span>if(toccolor)<span class="pl-pds">$</span><span class="pl-pds">$</span>toccolor<span class="pl-pds">$</span><span class="pl-pds">$</span>else<span class="pl-pds">$</span><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L454" class="blob-num js-line-number" data-line-number="454"></td>
        <td id="LC454" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L455" class="blob-num js-line-number" data-line-number="455"></td>
        <td id="LC455" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\setcounter</span>{tocdepth}{<span class="pl-s"><span class="pl-pds">$</span>toc-depth<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L456" class="blob-num js-line-number" data-line-number="456"></td>
        <td id="LC456" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\tableofcontents</span></td>
      </tr>
      <tr>
        <td id="L457" class="blob-num js-line-number" data-line-number="457"></td>
        <td id="LC457" class="blob-code blob-code-inner js-file-line">}</td>
      </tr>
      <tr>
        <td id="L458" class="blob-num js-line-number" data-line-number="458"></td>
        <td id="LC458" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L459" class="blob-num js-line-number" data-line-number="459"></td>
        <td id="LC459" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L460" class="blob-num js-line-number" data-line-number="460"></td>
        <td id="LC460" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(lot)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L461" class="blob-num js-line-number" data-line-number="461"></td>
        <td id="LC461" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\listoftables</span></td>
      </tr>
      <tr>
        <td id="L462" class="blob-num js-line-number" data-line-number="462"></td>
        <td id="LC462" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L463" class="blob-num js-line-number" data-line-number="463"></td>
        <td id="LC463" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(lof)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L464" class="blob-num js-line-number" data-line-number="464"></td>
        <td id="LC464" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\listoffigures</span></td>
      </tr>
      <tr>
        <td id="L465" class="blob-num js-line-number" data-line-number="465"></td>
        <td id="LC465" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L466" class="blob-num js-line-number" data-line-number="466"></td>
        <td id="LC466" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(linestretch)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L467" class="blob-num js-line-number" data-line-number="467"></td>
        <td id="LC467" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\setstretch</span>{<span class="pl-s"><span class="pl-pds">$</span>linestretch<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L468" class="blob-num js-line-number" data-line-number="468"></td>
        <td id="LC468" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L469" class="blob-num js-line-number" data-line-number="469"></td>
        <td id="LC469" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(has-frontmatter)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L470" class="blob-num js-line-number" data-line-number="470"></td>
        <td id="LC470" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\mainmatter</span></td>
      </tr>
      <tr>
        <td id="L471" class="blob-num js-line-number" data-line-number="471"></td>
        <td id="LC471" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L472" class="blob-num js-line-number" data-line-number="472"></td>
        <td id="LC472" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>body<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L473" class="blob-num js-line-number" data-line-number="473"></td>
        <td id="LC473" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L474" class="blob-num js-line-number" data-line-number="474"></td>
        <td id="LC474" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(has-frontmatter)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L475" class="blob-num js-line-number" data-line-number="475"></td>
        <td id="LC475" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\backmatter</span></td>
      </tr>
      <tr>
        <td id="L476" class="blob-num js-line-number" data-line-number="476"></td>
        <td id="LC476" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L477" class="blob-num js-line-number" data-line-number="477"></td>
        <td id="LC477" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(natbib)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L478" class="blob-num js-line-number" data-line-number="478"></td>
        <td id="LC478" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(bibliography)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L479" class="blob-num js-line-number" data-line-number="479"></td>
        <td id="LC479" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(biblio-title)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L480" class="blob-num js-line-number" data-line-number="480"></td>
        <td id="LC480" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(has-chapters)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L481" class="blob-num js-line-number" data-line-number="481"></td>
        <td id="LC481" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\renewcommand</span><span class="pl-c1">\bibname</span>{<span class="pl-s"><span class="pl-pds">$</span>biblio-title<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L482" class="blob-num js-line-number" data-line-number="482"></td>
        <td id="LC482" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L483" class="blob-num js-line-number" data-line-number="483"></td>
        <td id="LC483" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\renewcommand</span><span class="pl-c1">\refname</span>{<span class="pl-s"><span class="pl-pds">$</span>biblio-title<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L484" class="blob-num js-line-number" data-line-number="484"></td>
        <td id="LC484" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L485" class="blob-num js-line-number" data-line-number="485"></td>
        <td id="LC485" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L486" class="blob-num js-line-number" data-line-number="486"></td>
        <td id="LC486" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(beamer)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L487" class="blob-num js-line-number" data-line-number="487"></td>
        <td id="LC487" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\begin</span>{<span class="pl-smi">frame</span>}[allowframebreaks]{<span class="pl-s"><span class="pl-pds">$</span>biblio-title<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L488" class="blob-num js-line-number" data-line-number="488"></td>
        <td id="LC488" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\bibliographytrue</span></td>
      </tr>
      <tr>
        <td id="L489" class="blob-num js-line-number" data-line-number="489"></td>
        <td id="LC489" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L490" class="blob-num js-line-number" data-line-number="490"></td>
        <td id="LC490" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\bibliography</span>{<span class="pl-s"><span class="pl-pds">$</span>for(bibliography)<span class="pl-pds">$</span><span class="pl-pds">$</span>bibliography<span class="pl-pds">$</span><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span>,<span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L491" class="blob-num js-line-number" data-line-number="491"></td>
        <td id="LC491" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(beamer)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L492" class="blob-num js-line-number" data-line-number="492"></td>
        <td id="LC492" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\end</span>{<span class="pl-smi">frame</span>}</td>
      </tr>
      <tr>
        <td id="L493" class="blob-num js-line-number" data-line-number="493"></td>
        <td id="LC493" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L494" class="blob-num js-line-number" data-line-number="494"></td>
        <td id="LC494" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L495" class="blob-num js-line-number" data-line-number="495"></td>
        <td id="LC495" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L496" class="blob-num js-line-number" data-line-number="496"></td>
        <td id="LC496" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L497" class="blob-num js-line-number" data-line-number="497"></td>
        <td id="LC497" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(biblatex)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L498" class="blob-num js-line-number" data-line-number="498"></td>
        <td id="LC498" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(beamer)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L499" class="blob-num js-line-number" data-line-number="499"></td>
        <td id="LC499" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\begin</span>{<span class="pl-smi">frame</span>}[allowframebreaks]{<span class="pl-s"><span class="pl-pds">$</span>biblio-title<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L500" class="blob-num js-line-number" data-line-number="500"></td>
        <td id="LC500" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\bibliographytrue</span></td>
      </tr>
      <tr>
        <td id="L501" class="blob-num js-line-number" data-line-number="501"></td>
        <td id="LC501" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\printbibliography</span>[heading=none]</td>
      </tr>
      <tr>
        <td id="L502" class="blob-num js-line-number" data-line-number="502"></td>
        <td id="LC502" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\end</span>{<span class="pl-smi">frame</span>}</td>
      </tr>
      <tr>
        <td id="L503" class="blob-num js-line-number" data-line-number="503"></td>
        <td id="LC503" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L504" class="blob-num js-line-number" data-line-number="504"></td>
        <td id="LC504" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\printbibliography</span><span class="pl-s"><span class="pl-pds">$</span>if(biblio-title)<span class="pl-pds">$</span></span>[title=<span class="pl-s"><span class="pl-pds">$</span>biblio-title<span class="pl-pds">$</span></span>]<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L505" class="blob-num js-line-number" data-line-number="505"></td>
        <td id="LC505" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L506" class="blob-num js-line-number" data-line-number="506"></td>
        <td id="LC506" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L507" class="blob-num js-line-number" data-line-number="507"></td>
        <td id="LC507" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L508" class="blob-num js-line-number" data-line-number="508"></td>
        <td id="LC508" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>for(include-after)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L509" class="blob-num js-line-number" data-line-number="509"></td>
        <td id="LC509" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>include-after<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L510" class="blob-num js-line-number" data-line-number="510"></td>
        <td id="LC510" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L511" class="blob-num js-line-number" data-line-number="511"></td>
        <td id="LC511" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L512" class="blob-num js-line-number" data-line-number="512"></td>
        <td id="LC512" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\end</span>{<span class="pl-smi">document</span>}</td>
      </tr>
</table>

  <details class="details-reset details-overlay BlobToolbar position-absolute js-file-line-actions dropdown d-none" aria-hidden="true">
    <summary class="btn-octicon ml-0 px-2 p-0 bg-white border border-gray-dark rounded-1" aria-label="Inline file action toolbar">
      <svg class="octicon octicon-kebab-horizontal" viewBox="0 0 13 16" version="1.1" width="13" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M1.5 9a1.5 1.5 0 100-3 1.5 1.5 0 000 3zm5 0a1.5 1.5 0 100-3 1.5 1.5 0 000 3zM13 7.5a1.5 1.5 0 11-3 0 1.5 1.5 0 013 0z"/></svg>
    </summary>
    <details-menu>
      <ul class="BlobToolbar-dropdown dropdown-menu dropdown-menu-se mt-2" style="width:185px">
        <li>
          <clipboard-copy role="menuitem" class="dropdown-item" id="js-copy-lines" style="cursor:pointer;">
            Copy lines
          </clipboard-copy>
        </li>
        <li>
          <clipboard-copy role="menuitem" class="dropdown-item" id="js-copy-permalink" style="cursor:pointer;">
            Copy permalink
          </clipboard-copy>
        </li>
        <li><a class="dropdown-item js-update-url-with-hash" id="js-view-git-blame" role="menuitem" href="/drdanholmes/Affiliations/blame/c2313b057f54bba751b84530b4b6c47911d73922/Extras/mytemplate.tex">View git blame</a></li>
          <li><a class="dropdown-item" id="js-new-issue" role="menuitem" href="/drdanholmes/Affiliations/issues/new">Reference in new issue</a></li>
      </ul>
    </details-menu>
  </details>

  </div>

    </div>

  <details class="details-reset details-overlay details-overlay-dark">
    <summary data-hotkey="l" aria-label="Jump to line"></summary>
    <details-dialog class="Box Box--overlay d-flex flex-column anim-fade-in fast linejump" aria-label="Jump to line">
      <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="js-jump-to-line-form Box-body d-flex" action="" accept-charset="UTF-8" method="get"><input name="utf8" type="hidden" value="&#x2713;" />
        <input class="form-control flex-auto mr-3 linejump-input js-jump-to-line-field" type="text" placeholder="Jump to line&hellip;" aria-label="Jump to line" autofocus>
        <button type="submit" class="btn" data-close-dialog>Go</button>
</form>    </details-dialog>
  </details>

  </div>
</div>

    </main>
  </div>

  </div>

<div class="footer container-lg width-full p-responsive" role="contentinfo">
  <div class="position-relative d-flex flex-row-reverse flex-lg-row flex-wrap flex-lg-nowrap flex-justify-center flex-lg-justify-between pt-6 pb-2 mt-6 f6 text-gray border-top border-gray-light ">
    <ul class="list-style-none d-flex flex-wrap col-12 col-lg-5 flex-justify-center flex-lg-justify-between mb-2 mb-lg-0">
      <li class="mr-3 mr-lg-0">&copy; 2020 GitHub, Inc.</li>
        <li class="mr-3 mr-lg-0"><a data-ga-click="Footer, go to terms, text:terms" href="https://github.com/site/terms">Terms</a></li>
        <li class="mr-3 mr-lg-0"><a data-ga-click="Footer, go to privacy, text:privacy" href="https://github.com/site/privacy">Privacy</a></li>
        <li class="mr-3 mr-lg-0"><a data-ga-click="Footer, go to security, text:security" href="https://github.com/security">Security</a></li>
        <li class="mr-3 mr-lg-0"><a href="https://githubstatus.com/" data-ga-click="Footer, go to status, text:status">Status</a></li>
        <li><a data-ga-click="Footer, go to help, text:help" href="https://help.github.com">Help</a></li>
    </ul>

    <a aria-label="Homepage" title="GitHub" class="footer-octicon d-none d-lg-block mx-lg-4" href="https://github.com">
      <svg height="24" class="octicon octicon-mark-github" viewBox="0 0 16 16" version="1.1" width="24" aria-hidden="true"><path fill-rule="evenodd" d="M8 0C3.58 0 0 3.58 0 8c0 3.54 2.29 6.53 5.47 7.59.4.07.55-.17.55-.38 0-.19-.01-.82-.01-1.49-2.01.37-2.53-.49-2.69-.94-.09-.23-.48-.94-.82-1.13-.28-.15-.68-.52-.01-.53.63-.01 1.08.58 1.23.82.72 1.21 1.87.87 2.33.66.07-.52.28-.87.51-1.07-1.78-.2-3.64-.89-3.64-3.95 0-.87.31-1.59.82-2.15-.08-.2-.36-1.02.08-2.12 0 0 .67-.21 2.2.82.64-.18 1.32-.27 2-.27.68 0 1.36.09 2 .27 1.53-1.04 2.2-.82 2.2-.82.44 1.1.16 1.92.08 2.12.51.56.82 1.27.82 2.15 0 3.07-1.87 3.75-3.65 3.95.29.25.54.73.54 1.48 0 1.07-.01 1.93-.01 2.2 0 .21.15.46.55.38A8.013 8.013 0 0016 8c0-4.42-3.58-8-8-8z"/></svg>
</a>
   <ul class="list-style-none d-flex flex-wrap col-12 col-lg-5 flex-justify-center flex-lg-justify-between mb-2 mb-lg-0">
        <li class="mr-3 mr-lg-0"><a data-ga-click="Footer, go to contact, text:contact" href="https://github.com/contact">Contact GitHub</a></li>
        <li class="mr-3 mr-lg-0"><a href="https://github.com/pricing" data-ga-click="Footer, go to Pricing, text:Pricing">Pricing</a></li>
      <li class="mr-3 mr-lg-0"><a href="https://developer.github.com" data-ga-click="Footer, go to api, text:api">API</a></li>
      <li class="mr-3 mr-lg-0"><a href="https://training.github.com" data-ga-click="Footer, go to training, text:training">Training</a></li>
        <li class="mr-3 mr-lg-0"><a href="https://github.blog" data-ga-click="Footer, go to blog, text:blog">Blog</a></li>
        <li><a data-ga-click="Footer, go to about, text:about" href="https://github.com/about">About</a></li>

    </ul>
  </div>
  <div class="d-flex flex-justify-center pb-6">
    <span class="f6 text-gray-light"></span>
  </div>
</div>

  <div id="ajax-error-message" class="ajax-error-message flash flash-error">
    <svg class="octicon octicon-alert" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.893 1.5c-.183-.31-.52-.5-.887-.5s-.703.19-.886.5L.138 13.499a.98.98 0 000 1.001c.193.31.53.501.886.501h13.964c.367 0 .704-.19.877-.5a1.03 1.03 0 00.01-1.002L8.893 1.5zm.133 11.497H6.987v-2.003h2.039v2.003zm0-3.004H6.987V5.987h2.039v4.006z"/></svg>
    <button type="button" class="flash-close js-ajax-error-dismiss" aria-label="Dismiss error">
      <svg class="octicon octicon-x" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48L7.48 8z"/></svg>
    </button>
    You can’t perform that action at this time.
  </div>

    <script crossorigin="anonymous" async="async" integrity="sha512-o4vS4IKrjdy/HD+xr2+VhO6DxQmj5jikhHbEGrd8+JGhpmIOxRrpT1Qo5k3IhKimm8VXIu3pyYejLtOAkm+OsQ==" type="application/javascript" id="js-conditional-compat" data-src="https://github.githubassets.com/assets/compat-bootstrap-a38bd2e0.js"></script>
    <script crossorigin="anonymous" async="async" integrity="sha512-3yNijdFdVDBZQDWrBvMeD2J9gyXwI5MKUMJSWdEGP44DgS4NEPQw2TmVlDdNAWrseJO5C/sXBSTrL24DvGMDJw==" type="application/javascript" src="https://github.githubassets.com/assets/vendor-df23628d.js"></script>
    <script crossorigin="anonymous" integrity="sha512-YD0xPfLsUZUxjui//wwr0OgQEwI82YKRuwFtsU9iAJ20CVrvy875DrPPmTlphJwWHeikuQmoalVA8HbrN0dntQ==" type="application/javascript" src="https://github.githubassets.com/assets/frameworks-603d313d.js"></script>

    <script crossorigin="anonymous" async="async" integrity="sha512-kA1JCSWQ+QgqnPacgjlmyQlCjlCe6+HX0rdloGMJz9KG2OxjRbDewNuGabKiPodLlgGNk7LzwnjHepPZ8sYldg==" type="application/javascript" src="https://github.githubassets.com/assets/github-bootstrap-900d4909.js"></script>

  <div class="js-stale-session-flash flash flash-warn flash-banner" hidden
    >
    <svg class="octicon octicon-alert" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.893 1.5c-.183-.31-.52-.5-.887-.5s-.703.19-.886.5L.138 13.499a.98.98 0 000 1.001c.193.31.53.501.886.501h13.964c.367 0 .704-.19.877-.5a1.03 1.03 0 00.01-1.002L8.893 1.5zm.133 11.497H6.987v-2.003h2.039v2.003zm0-3.004H6.987V5.987h2.039v4.006z"/></svg>
    <span class="js-stale-session-flash-signed-in" hidden>You signed in with another tab or window. <a href="">Reload</a> to refresh your session.</span>
    <span class="js-stale-session-flash-signed-out" hidden>You signed out in another tab or window. <a href="">Reload</a> to refresh your session.</span>
  </div>
  <template id="site-details-dialog">
  <details class="details-reset details-overlay details-overlay-dark lh-default text-gray-dark hx_rsm" open>
    <summary role="button" aria-label="Close dialog"></summary>
    <details-dialog class="Box Box--overlay d-flex flex-column anim-fade-in fast hx_rsm-dialog hx_rsm-modal">
      <button class="Box-btn-octicon m-0 btn-octicon position-absolute right-0 top-0" type="button" aria-label="Close dialog" data-close-dialog>
        <svg class="octicon octicon-x" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48L7.48 8z"/></svg>
      </button>
      <div class="octocat-spinner my-6 js-details-dialog-spinner"></div>
    </details-dialog>
  </details>
</template>

  <div class="Popover js-hovercard-content position-absolute" style="display: none; outline: none;" tabindex="0">
  <div class="Popover-message Popover-message--bottom-left Popover-message--large Box box-shadow-large" style="width:360px;">
  </div>
</div>

  <div aria-live="polite" class="js-global-screen-reader-notice sr-only"></div>

  </body>
</html>

### 2020-02-27 dfol_draft/defol_draft2.Rmd
initial trials

---
title: "*dfoliatR*: An R package for identification and analyses of insect defoliation and growth suppression signals in tree rings"
authors:
  - Christopher H. Guiterman:
      email: chguiterman@email.arizona.edu
      affiliation: [LTRR]
      correspondence: true
    # footnote: 1
  - Ann M. Lynch:
      email: lyncha@email.arizona.edu
      affiliation: [RMRS, LTRR]
      correspondence: false
  - Jodi Axelson:
      email: jodi.axelson@berkeley.edu 
      affiliation: [UCB]
      correspondence: false
institute:
  - LTRR: Laboratory of Tree-Ring Research, University of Arizona, 1215 E Lowell St. Box 210045, Tucson, AZ, 85721
  - RMRS: U.S. Forest Service, Rocky Mountain Research Station, 1215 E Lowell St. Box 210045, Tucson, AZ, 85721
  - UCB: Department of Environmental Science, Policy & Management, University of California, Berkeley, 120 Mulford Hall, Berkeley, CA
abstract: For author review only. **Please do not share**
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  bookdown::word_document2:
    reference_docx: "../common/word-styles-reference-01.docx"
    toc: false
    toc_depth: 2
    fig_caption: yes
    df_print: kable
bibliography: "../common/Report.bib"
csl: "../common/ecology.csl"
keep_md: false
link-citations: false
---
```{r setup, comment=FALSE, message = FALSE, echo=FALSE, warning=FALSE}
rm(list=ls())           # Clean the environment
options(scipen=6)       # display digits proberly!! not the scientific version
options(digits.secs=6)  # use milliseconds in Date/Time data types
options(warning=FALSE)  # don't show warnings
options(figure_counter = TRUE)
library(knitr)          # set global knitr options of the document
library(dplyr)
# Here we set the R code invisible to not show when the document is rendered
opts_chunk$set(comment="", message = FALSE, echo=FALSE, error=FALSE, warning=FALSE,
               fig.show='asis', dev='pdf')
```

# Introduction

Some text.

### 2020-02-27 dfol_draft/dfol_draft.Rmd
initial trials

---
title: "*dfoliatR*: An R package for identification and analyses of insect defoliation and growth suppression signals in tree rings"
author:
  - Christopher H. Guiterman:
      email: chguiterman@email.arizona.edu
      affiliation: [LTRR]
      correspondence: true
    # footnote: 1
  - Ann M. Lynch:
      email: lyncha@email.arizona.edu
      affiliation: [RMRS, LTRR]
      correspondence: false
  - Jodi Axelson:
      email: jodi.axelson@berkeley.edu 
      affiliation: [UCB]
      correspondence: false
    # footnote: 2
# address:
#   - code: LTRR
#     address: Laboratory of Tree-Ring Research, University of Arizona, 1215 E Lowell St. Box 210045, Tucson, AZ, 85721
#   - code: RMRS
#     address: U.S. Forest Service, Rocky Mountain Research Station, 1215 E Lowell St. Box 210045, Tucson, AZ, 85721
#   - code: UCB
#     address: Department of Environmental Science, Policy & Management, 120 Mulford Hall, University of California, Berkeley
# Berkeley, CA 94720
institute:
  - LTRR: Laboratory of Tree-Ring Research, University of Arizona, 1215 E Lowell St. Box 210045, Tucson, AZ, 85721
  - RMRS: U.S. Forest Service, Rocky Mountain Research Station, 1215 E Lowell St. Box 210045, Tucson, AZ, 85721
  - UCB: Department of Environmental Science, Policy & Management, University of California, Berkeley, 120 Mulford Hall, Berkeley, CA 94720
# footnote:
#   - code: 1
#     text: "Corresponding Author (chguiterman@email.arizona.edu)"
#   # - code: 2
#   #   text: "Equal contribution"
abstract: |
  This is the abstract.

  It consists of two paragraphs.
keywords: "Defoliation, Insect outbreak"
journal: "Tree-Ring Research"
date: "`r Sys.Date()`"
# bibliography: mybibfile.bib
# linenumbers: true
#numbersections: true
# csl: elsevier-harvard.csl
# output: rticles::elsevier_article
# output: 
#     word_document:
#       toc: no
#       pandoc_args:
#         - '--lua-filter=scholarly-metadata.lua'
#         - '--lua-filter=author-info-blocks.lua'
#     pdf_document:
#       rticles::elsevier_article
#       pandoc_args:
#         - '--lua-filter=scholarly-metadata.lua'
#         - '--lua-filter=author-info-blocks.lua'
# output: rticles::elsevier_article
output:
 bookdown::word_document2:
   # reference_docx: "../extras/word-styles-reference-01.docx"
   # toc: false
   # toc_depth: 2
   # fig_caption: yes
   # df_print: kable
    pandoc_args:
      - --csl=Extras/clinical-mass-spectrometry.csl
      - --citation-abbreviations=Extras/abbreviations.json
      # - --filter=pandoc-crossref
      - --lua-filter=Extras/scholarly-metadata.lua
      - --lua-filter=Extras/author-info-blocks.lua
      - --reference-doc=Extras/Reference_Document.docx 
bibliography: "Extras/mybibfile.bib"
toc: false
---

_Text based on elsarticle sample manuscript, see [http://www.elsevier.com/author-schemas/latex-instructions#elsarticle](http://www.elsevier.com/author-schemas/latex-instructions#elsarticle)_

The Elsevier article class
==========================

#### Installation

If the document class *elsarticle* is not available on your computer,
you can download and install the system package *texlive-publishers*
(Linux) or install the LaTeX package *elsarticle* using the package
manager of your TeX installation, which is typically TeX Live or MikTeX.

#### Usage

Once the package is properly installed, you can use the document class
*elsarticle* to create a manuscript. Please make sure that your
manuscript follows the guidelines in the Guide for Authors of the
relevant journal. It is not necessary to typeset your manuscript in
exactly the same way as an article, unless you are submitting to a
camera-ready copy (CRC) journal.

#### Functionality

The Elsevier article class is based on the standard article class and
supports almost all of the functionality of that class. In addition, it
features commands and options to format the

-   document style

-   baselineskip

-   front matter

-   keywords and MSC codes

-   theorems, definitions and proofs

-   lables of enumerations

-   citation style and labeling.

Front matter
============

The author names and affiliations could be formatted in two ways:

(1) Group the authors per affiliation.

(2) Use footnotes to indicate the affiliations.

See the front matter of this document for examples. You are recommended
to conform your choice to the journal you are submitting to.

Bibliography styles
===================

There are various bibliography styles available. You can select the
style of your choice in the preamble of this document. These styles are
Elsevier styles based on standard styles like Harvard and Vancouver.
Please use BibTeX to generate your bibliography and include DOIs
whenever available.

Here are two sample references: @Feynman1963118 [@Dirac1953888].

References {#references .unnumbered}
==========

### 2020-02-27 dfol_draft/dfol_draft.tex
initial trials

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{Tree-Ring Research} % Sets Journal name

\usepackage{lineno} % add
  \linenumbers % turns line numbering on
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={dfoliatR: An R package for identification and analyses of insect defoliation and growth suppression signals in tree rings},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}

% Pandoc header

\begin{document}
\begin{frontmatter}

  \title{\emph{dfoliatR}: An R package for identification and analyses of insect
defoliation and growth suppression signals in tree rings}
    \author[]{}

    \author[]{}

    \author[]{}

  \begin{abstract}
  This is the abstract.

  It consists of two paragraphs.
  \end{abstract}

 \end{frontmatter}

\emph{Text based on elsarticle sample manuscript, see
\url{http://www.elsevier.com/author-schemas/latex-instructions\#elsarticle}}

\hypertarget{the-elsevier-article-class}{%
\section{The Elsevier article class}\label{the-elsevier-article-class}}

\hypertarget{installation}{%
\paragraph{Installation}\label{installation}}

If the document class \emph{elsarticle} is not available on your
computer, you can download and install the system package
\emph{texlive-publishers} (Linux) or install the LaTeX package
\emph{elsarticle} using the package manager of your TeX installation,
which is typically TeX Live or MikTeX.

\hypertarget{usage}{%
\paragraph{Usage}\label{usage}}

Once the package is properly installed, you can use the document class
\emph{elsarticle} to create a manuscript. Please make sure that your
manuscript follows the guidelines in the Guide for Authors of the
relevant journal. It is not necessary to typeset your manuscript in
exactly the same way as an article, unless you are submitting to a
camera-ready copy (CRC) journal.

\hypertarget{functionality}{%
\paragraph{Functionality}\label{functionality}}

The Elsevier article class is based on the standard article class and
supports almost all of the functionality of that class. In addition, it
features commands and options to format the

\begin{itemize}
\item
  document style
\item
  baselineskip
\item
  front matter
\item
  keywords and MSC codes
\item
  theorems, definitions and proofs
\item
  lables of enumerations
\item
  citation style and labeling.
\end{itemize}

\hypertarget{front-matter}{%
\section{Front matter}\label{front-matter}}

The author names and affiliations could be formatted in two ways:

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\item
  Group the authors per affiliation.
\item
  Use footnotes to indicate the affiliations.
\end{enumerate}

See the front matter of this document for examples. You are recommended
to conform your choice to the journal you are submitting to.

\hypertarget{bibliography-styles}{%
\section{Bibliography styles}\label{bibliography-styles}}

There are various bibliography styles available. You can select the
style of your choice in the preamble of this document. These styles are
Elsevier styles based on standard styles like Harvard and Vancouver.
Please use BibTeX~to generate your bibliography and include DOIs
whenever available.

Here are two sample references: Feynman and Vernon Jr. (1963; Dirac,
1953).

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Dirac1953888}{}%
Dirac, P.A., 1953. The lorentz transformation and absolute time. Physica
19, 888--896.
doi:\href{https://doi.org/10.1016/S0031-8914(53)80099-6}{10.1016/S0031-8914(53)80099-6}

\leavevmode\hypertarget{ref-Feynman1963118}{}%
Feynman, R.P., Vernon Jr., F.L., 1963. The theory of a general quantum
system interacting with a linear dissipative system. Annals of Physics
24, 118--173.
doi:\href{https://doi.org/10.1016/0003-4916(63)90068-X}{10.1016/0003-4916(63)90068-X}

\end{document}

### 2020-02-27 manuscript.Rmd
initial trials

---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    # template: svm-latex-ms.tex
bibliography: master.bib
header-includes:
  -  \usepackage{hyperref}
biblio-style: apsr
title: "A Pandoc Markdown Article Starter and Template"
thanks: "Replication files are available on the author's Github account (http://github.com/svmiller). **Current version**: `r format(Sys.time(), '%B %d, %Y')`; **Corresponding author**: svmille@clemson.edu."
author:
- name: Steven V. Miller
  affiliation: Clemson University
abstract: "This document provides an introduction to R Markdown, argues for its benefits, and presents a sample manuscript template intended for an academic audience. I include basic syntax to R Markdown and a minimal working example of how the analysis itself can be conducted within R with the `knitr` package."
keywords: "pandoc, r markdown, knitr"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
endnote: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE,
                      message=FALSE, warning=FALSE,
                      fig.path='figs/',
                      cache.path = '_cache/',
                      fig.process = function(x) {
                      x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
                      if (file.rename(x, x2)) x2 else x
                      })
```

## Introduction

Some text.

### 2020-02-27 manuscript.tex
initial trials

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  11pt,
]{article}
\usepackage[]{mathpazo}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={A Pandoc Markdown Article Starter and Template},
  pdfkeywords={pandoc, r markdown, knitr},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{hyperref}
\usepackage[]{natbib}
\bibliographystyle{apsr}

\title{A Pandoc Markdown Article Starter and Template\thanks{Replication files are available on the author's Github account
(\url{http://github.com/svmiller}). \textbf{Current version}: February
26, 2020; \textbf{Corresponding author}:
\href{mailto:svmille@clemson.edu}{\nolinkurl{svmille@clemson.edu}}.}}
\author{true}
\date{February 26, 2020}

\begin{document}
\maketitle
\begin{abstract}
This document provides an introduction to R Markdown, argues for its
benefits, and presents a sample manuscript template intended for an
academic audience. I include basic syntax to R Markdown and a minimal
working example of how the analysis itself can be conducted within R
with the \texttt{knitr} package.
\end{abstract}

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Some text.

  \bibliography{master.bib}

\end{document}

### 2020-02-27 paper/paper.Rmd
initial trials

---
title: Title here
subtitle: Do you have a subtitle? If so, write it here
titlerunning: Short form of title (if too long for head)
authorrunning: Short form of author list if too long for running head
thanks: | 
    Grants or other notes about the article that should go on the front 
    page should be placed here. General acknowledgments should be placed at the
    end of the article.

authors: 
- name: Äüthör 1
  address: Department of YYY, University of XXX
  email: abc@def

- name: Âuthóř 2
  address: Department of ZZZ, University of WWW
  email: djf@wef

keywords:
- key
- dictionary
- word

#PACS: 
#- PAC1
#- superPAC

MSC:
- MSC code 1
- MSC code 2

abstract: |
  The text of your abstract.  150 -- 250 words.

bibliography: bibliography.bib
bibstyle: spbasic
# bibstyle options spbasic(default), spphys, spmpsci
output: rticles::springer_article
---

# Introduction {#intro}

Your text comes here. Separate text sections with \cite{Mislevy06Cog}.

# Section title {#sec:1}

Text with citations by \cite{Galyardt14mmm}.

## Subsection title {#sec:2}

as required. Don't forget to give each section
and subsection a unique label (see Sect. \ref{sec:1}).

#### Paragraph headings

Use paragraph headings as needed.

\begin{align}
a^2+b^2=c^2
\end{align}

### 2020-02-27 paper/paper.tex
initial trials

% !TeX program = pdfLaTeX
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}

\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%

%% load any required packages here

\begin{document}

\title{Title here \thanks{Grants or other notes about the article that should go on the front page
should be placed here. General acknowledgments should be placed at the
end of the article.} }
 \subtitle{Do you have a subtitle? If so, write it here}

    \titlerunning{Short form of title (if too long for head)}

\author{  Äüthör 1 \and  Âuthóř 2 \and  }

    \authorrunning{ Short form of author list if too long for running head }

\institute{
        Äüthör 1 \at
     Department of YYY, University of XXX \\
     \email{\href{mailto:abc@def}{\nolinkurl{abc@def}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
        Âuthóř 2 \at
     Department of ZZZ, University of WWW \\
     \email{\href{mailto:djf@wef}{\nolinkurl{djf@wef}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
    }

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
The text of your abstract. 150 -- 250 words.
\\
\keywords{
        key \and
        dictionary \and
        word \and
    }

    \subclass{
                    MSC code 1 \and
                    MSC code 2 \and
            }

\end{abstract}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

\hypertarget{intro}{%
\section{Introduction}\label{intro}}

Your text comes here. Separate text sections with \cite{Mislevy06Cog}.

\hypertarget{sec:1}{%
\section{Section title}\label{sec:1}}

Text with citations by \cite{Galyardt14mmm}.

\hypertarget{sec:2}{%
\subsection{Subsection title}\label{sec:2}}

as required. Don't forget to give each section and subsection a unique
label (see Sect. \ref{sec:1}).

\hypertarget{paragraph-headings}{%
\paragraph{Paragraph headings}\label{paragraph-headings}}

Use paragraph headings as needed.

\begin{align}
a^2+b^2=c^2
\end{align}

\bibliographystyle{spbasic}
\bibliography{bibliography.bib}

\end{document}

### 2020-02-27 svm-latex-ms.tex
initial trials

<!DOCTYPE html>
<html lang="en">
  <head>
    <meta charset="utf-8">
  <link rel="dns-prefetch" href="https://github.githubassets.com">
  <link rel="dns-prefetch" href="https://avatars0.githubusercontent.com">
  <link rel="dns-prefetch" href="https://avatars1.githubusercontent.com">
  <link rel="dns-prefetch" href="https://avatars2.githubusercontent.com">
  <link rel="dns-prefetch" href="https://avatars3.githubusercontent.com">
  <link rel="dns-prefetch" href="https://github-cloud.s3.amazonaws.com">
  <link rel="dns-prefetch" href="https://user-images.githubusercontent.com/">

  <link crossorigin="anonymous" media="all" integrity="sha512-ZUjVod2EvYMDbGqRSyW0rpfgBq3i+gnR/4PfrzLsy5f20oIcRfgFQFVKgi3Ztp917bP1K/kdP5q8+nAlJ3+cFA==" rel="stylesheet" href="https://github.githubassets.com/assets/frameworks-6548d5a1dd84bd83036c6a914b25b4ae.css" />

    <link crossorigin="anonymous" media="all" integrity="sha512-aX4OkLpzulpadvOncEEPpJZnQyeKNm2npzJowbL5JxptkoZXNPPy61R059xmEa3YyVF4Y4YXB6g+5o08uvdWpA==" rel="stylesheet" href="https://github.githubassets.com/assets/github-697e0e90ba73ba5a5a76f3a770410fa4.css" />

  <meta name="viewport" content="width=device-width">

  <title>svm-r-markdown-templates/svm-latex-ms.tex at master · svmiller/svm-r-markdown-templates</title>
    <meta name="description" content="I have a suite of R Markdown templates for academic manuscripts, beamer presentations, and syllabi. I share them here. - svmiller/svm-r-markdown-templates">
    <link rel="search" type="application/opensearchdescription+xml" href="/opensearch.xml" title="GitHub">
  <link rel="fluid-icon" href="https://github.com/fluidicon.png" title="GitHub">
  <meta property="fb:app_id" content="1401488693436528">

    <meta name="twitter:image:src" content="https://avatars2.githubusercontent.com/u/4589542?s=400&amp;v=4" /><meta name="twitter:site" content="@github" /><meta name="twitter:card" content="summary" /><meta name="twitter:title" content="svmiller/svm-r-markdown-templates" /><meta name="twitter:description" content="I have a suite of R Markdown templates for academic manuscripts, beamer presentations, and syllabi. I share them here. - svmiller/svm-r-markdown-templates" />
    <meta property="og:image" content="https://avatars2.githubusercontent.com/u/4589542?s=400&amp;v=4" /><meta property="og:site_name" content="GitHub" /><meta property="og:type" content="object" /><meta property="og:title" content="svmiller/svm-r-markdown-templates" /><meta property="og:url" content="https://github.com/svmiller/svm-r-markdown-templates" /><meta property="og:description" content="I have a suite of R Markdown templates for academic manuscripts, beamer presentations, and syllabi. I share them here. - svmiller/svm-r-markdown-templates" />

  <link rel="assets" href="https://github.githubassets.com/">
  <link rel="web-socket" href="wss://live.github.com/_sockets/VjI6NDg1MTMzNjY1OjMzOWJkYmM4YzYzMmIxMGY4YjAwZDc2YTEzMjkxY2JkNWNkNDZjODc4ZjBlYmY3Mzg0YjMwMjY4MDgwMjU0M2U=--874b69e8645e339c452dd394f2d494999b4d3b82">
  <link rel="sudo-modal" href="/sessions/sudo_modal">

  <meta name="request-id" content="E1FC:4E2D:19E909:30B862:5E553558" data-pjax-transient="true" /><meta name="html-safe-nonce" content="e10583c265ecb6446b852a6b2c5b5a01cd997a34" data-pjax-transient="true" /><meta name="visitor-payload" content="eyJyZWZlcnJlciI6Imh0dHBzOi8vZ2l0aHViLmNvbS9zdm1pbGxlci9zdm0tci1tYXJrZG93bi10ZW1wbGF0ZXMiLCJyZXF1ZXN0X2lkIjoiRTFGQzo0RTJEOjE5RTkwOTozMEI4NjI6NUU1NTM1NTgiLCJ2aXNpdG9yX2lkIjoiMzg5MTI0MjQwNzU2MjUxNjQ5MSIsInJlZ2lvbl9lZGdlIjoiaWFkIiwicmVnaW9uX3JlbmRlciI6ImlhZCJ9" data-pjax-transient="true" /><meta name="visitor-hmac" content="31d5327f2a82638fd63670b242e5f5f7a825eb03fd8598fa55e7f74b6bcb7cb9" data-pjax-transient="true" />

  <meta name="github-keyboard-shortcuts" content="repository,source-code" data-pjax-transient="true" />

  <meta name="selected-link" value="repo_source" data-pjax-transient>

      <meta name="google-site-verification" content="KT5gs8h0wvaagLKAVWq8bbeNwnZZK1r1XQysX3xurLU">
    <meta name="google-site-verification" content="ZzhVyEFwb7w3e0-uOTltm8Jsck2F5StVihD0exw2fsA">
    <meta name="google-site-verification" content="GXs5KoUUkNCoaAZn7wPN-t01Pywp9M3sEjnt_3_ZWPc">

  <meta name="octolytics-host" content="collector.githubapp.com" /><meta name="octolytics-app-id" content="github" /><meta name="octolytics-event-url" content="https://collector.githubapp.com/github-external/browser_event" /><meta name="octolytics-dimension-ga_id" content="" class="js-octo-ga-id" /><meta name="octolytics-actor-id" content="15913450" /><meta name="octolytics-actor-login" content="chguiterman" /><meta name="octolytics-actor-hash" content="ac81b7290092b2c4a1347b423992342f7581f20c94fd8a35e3bf98004cc18b13" />
<meta name="analytics-location" content="/&lt;user-name&gt;/&lt;repo-name&gt;/blob/show" data-pjax-transient="true" />

    <meta name="google-analytics" content="UA-3769691-2">

  <meta class="js-ga-set" name="userId" content="aff5acac9f546acde3aae63df4076c10">

<meta class="js-ga-set" name="dimension1" content="Logged In">

      <meta name="hostname" content="github.com">
    <meta name="user-login" content="chguiterman">

      <meta name="expected-hostname" content="github.com">

      <meta name="js-proxy-site-detection-payload" content="NDllZTVmZTRhOTM5NzJlMzE2M2EzYTAyOWQwOTEwNDZhMjRkZGU0MzMxZDliNWMyYTYyNGNmYWJmMDRmNTNkMHx7InJlbW90ZV9hZGRyZXNzIjoiNjQuMjIzLjE1Ny4xNzYiLCJyZXF1ZXN0X2lkIjoiRTFGQzo0RTJEOjE5RTkwOTozMEI4NjI6NUU1NTM1NTgiLCJ0aW1lc3RhbXAiOjE1ODI2NDI1MjUsImhvc3QiOiJnaXRodWIuY29tIn0=">

    <meta name="enabled-features" content="MARKETPLACE_FEATURED_BLOG_POSTS,MARKETPLACE_INVOICED_BILLING,MARKETPLACE_SOCIAL_PROOF_CUSTOMERS,MARKETPLACE_TRENDING_SOCIAL_PROOF,MARKETPLACE_RECOMMENDATIONS,MARKETPLACE_PENDING_INSTALLATIONS,RELATED_ISSUES,GHE_CLOUD_TRIAL">

  <meta http-equiv="x-pjax-version" content="9f8e8c1805956d93c0f4a133eb96c5e3">

      <link href="https://github.com/svmiller/svm-r-markdown-templates/commits/master.atom" rel="alternate" title="Recent Commits to svm-r-markdown-templates:master" type="application/atom+xml">

  <meta name="go-import" content="github.com/svmiller/svm-r-markdown-templates git https://github.com/svmiller/svm-r-markdown-templates.git">

  <meta name="octolytics-dimension-user_id" content="4589542" /><meta name="octolytics-dimension-user_login" content="svmiller" /><meta name="octolytics-dimension-repository_id" content="51473841" /><meta name="octolytics-dimension-repository_nwo" content="svmiller/svm-r-markdown-templates" /><meta name="octolytics-dimension-repository_public" content="true" /><meta name="octolytics-dimension-repository_is_fork" content="false" /><meta name="octolytics-dimension-repository_network_root_id" content="51473841" /><meta name="octolytics-dimension-repository_network_root_nwo" content="svmiller/svm-r-markdown-templates" /><meta name="octolytics-dimension-repository_explore_github_marketplace_ci_cta_shown" content="false" />

    <link rel="canonical" href="https://github.com/svmiller/svm-r-markdown-templates/blob/master/svm-latex-ms.tex" data-pjax-transient>

  <meta name="browser-stats-url" content="https://api.github.com/_private/browser/stats">

  <meta name="browser-errors-url" content="https://api.github.com/_private/browser/errors">

  <link rel="mask-icon" href="https://github.githubassets.com/pinned-octocat.svg" color="#000000">
  <link rel="icon" type="image/x-icon" class="js-site-favicon" href="https://github.githubassets.com/favicon.ico">

<meta name="theme-color" content="#1e2327">

  <link rel="manifest" href="/manifest.json" crossOrigin="use-credentials">

  </head>

  <body class="logged-in env-production page-responsive page-blob">

  <div class="position-relative js-header-wrapper ">
    <a href="#start-of-content" tabindex="1" class="p-3 bg-blue text-white show-on-focus js-skip-to-content">Skip to content</a>
    <span class="Progress progress-pjax-loader position-fixed width-full js-pjax-loader-bar">
      <span class="progress-pjax-loader-bar top-0 left-0" style="width: 0%;"></span>
    </span>

          <header class="Header js-details-container Details flex-wrap flex-lg-nowrap p-responsive" role="banner">

    <div class="Header-item d-none d-lg-flex">
      <a class="Header-link" href="https://github.com/" data-hotkey="g d" aria-label="Homepage" data-ga-click="Header, go to dashboard, icon:logo">
  <svg class="octicon octicon-mark-github v-align-middle" height="32" viewBox="0 0 16 16" version="1.1" width="32" aria-hidden="true"><path fill-rule="evenodd" d="M8 0C3.58 0 0 3.58 0 8c0 3.54 2.29 6.53 5.47 7.59.4.07.55-.17.55-.38 0-.19-.01-.82-.01-1.49-2.01.37-2.53-.49-2.69-.94-.09-.23-.48-.94-.82-1.13-.28-.15-.68-.52-.01-.53.63-.01 1.08.58 1.23.82.72 1.21 1.87.87 2.33.66.07-.52.28-.87.51-1.07-1.78-.2-3.64-.89-3.64-3.95 0-.87.31-1.59.82-2.15-.08-.2-.36-1.02.08-2.12 0 0 .67-.21 2.2.82.64-.18 1.32-.27 2-.27.68 0 1.36.09 2 .27 1.53-1.04 2.2-.82 2.2-.82.44 1.1.16 1.92.08 2.12.51.56.82 1.27.82 2.15 0 3.07-1.87 3.75-3.65 3.95.29.25.54.73.54 1.48 0 1.07-.01 1.93-.01 2.2 0 .21.15.46.55.38A8.013 8.013 0 0016 8c0-4.42-3.58-8-8-8z"/></svg>
</a>

    </div>

    <div class="Header-item d-lg-none">
      <button class="Header-link btn-link js-details-target" type="button" aria-label="Toggle navigation" aria-expanded="false">
        <svg height="24" class="octicon octicon-three-bars" viewBox="0 0 12 16" version="1.1" width="18" aria-hidden="true"><path fill-rule="evenodd" d="M11.41 9H.59C0 9 0 8.59 0 8c0-.59 0-1 .59-1H11.4c.59 0 .59.41.59 1 0 .59 0 1-.59 1h.01zm0-4H.59C0 5 0 4.59 0 4c0-.59 0-1 .59-1H11.4c.59 0 .59.41.59 1 0 .59 0 1-.59 1h.01zM.59 11H11.4c.59 0 .59.41.59 1 0 .59 0 1-.59 1H.59C0 13 0 12.59 0 12c0-.59 0-1 .59-1z"/></svg>
      </button>
    </div>

    <div class="Header-item Header-item--full flex-column flex-lg-row width-full flex-order-2 flex-lg-order-none mr-0 mr-lg-3 mt-3 mt-lg-0 Details-content--hidden">
        <div class="header-search flex-self-stretch flex-lg-self-auto mr-0 mr-lg-3 mb-3 mb-lg-0 scoped-search site-scoped-search js-site-search position-relative js-jump-to"
  role="combobox"
  aria-owns="jump-to-results"
  aria-label="Search or jump to"
  aria-haspopup="listbox"
  aria-expanded="false"
>
  <div class="position-relative">
    <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="js-site-search-form" role="search" aria-label="Site" data-scope-type="Repository" data-scope-id="51473841" data-scoped-search-url="/svmiller/svm-r-markdown-templates/search" data-unscoped-search-url="/search" action="/svmiller/svm-r-markdown-templates/search" accept-charset="UTF-8" method="get"><input name="utf8" type="hidden" value="&#x2713;" />
      <label class="form-control input-sm header-search-wrapper p-0 header-search-wrapper-jump-to position-relative d-flex flex-justify-between flex-items-center js-chromeless-input-container">
        <input type="text"
          class="form-control input-sm header-search-input jump-to-field js-jump-to-field js-site-search-focus js-site-search-field is-clearable"
          data-hotkey="s,/"
          name="q"
          value=""
          placeholder="Search or jump to…"
          data-unscoped-placeholder="Search or jump to…"
          data-scoped-placeholder="Search or jump to…"
          autocapitalize="off"
          aria-autocomplete="list"
          aria-controls="jump-to-results"
          aria-label="Search or jump to…"
          data-jump-to-suggestions-path="/_graphql/GetSuggestedNavigationDestinations"
          spellcheck="false"
          autocomplete="off"
          >
          <input type="hidden" value="liLybTmxE7RjN5joI4KNt5ueTM9rYCjS0VtzQkwfLcHhhbID6Fj1KxxTxXtzezgbP3WbTuQRM9KGhBGvL8Prog==" data-csrf="true" class="js-data-jump-to-suggestions-path-csrf" />
          <input type="hidden" class="js-site-search-type-field" name="type" >
            <img src="https://github.githubassets.com/images/search-key-slash.svg" alt="" class="mr-2 header-search-key-slash">

            <div class="Box position-absolute overflow-hidden d-none jump-to-suggestions js-jump-to-suggestions-container">

<ul class="d-none js-jump-to-suggestions-template-container">

<li class="d-flex flex-justify-start flex-items-center p-0 f5 navigation-item js-navigation-item js-jump-to-suggestion" role="option">
  <a tabindex="-1" class="no-underline d-flex flex-auto flex-items-center jump-to-suggestions-path js-jump-to-suggestion-path js-navigation-open p-2" href="">
    <div class="jump-to-octicon js-jump-to-octicon flex-shrink-0 mr-2 text-center d-none">
      <svg height="16" width="16" class="octicon octicon-repo flex-shrink-0 js-jump-to-octicon-repo d-none" title="Repository" aria-label="Repository" viewBox="0 0 12 16" version="1.1" role="img"><path fill-rule="evenodd" d="M4 9H3V8h1v1zm0-3H3v1h1V6zm0-2H3v1h1V4zm0-2H3v1h1V2zm8-1v12c0 .55-.45 1-1 1H6v2l-1.5-1.5L3 16v-2H1c-.55 0-1-.45-1-1V1c0-.55.45-1 1-1h10c.55 0 1 .45 1 1zm-1 10H1v2h2v-1h3v1h5v-2zm0-10H2v9h9V1z"/></svg>
      <svg height="16" width="16" class="octicon octicon-project flex-shrink-0 js-jump-to-octicon-project d-none" title="Project" aria-label="Project" viewBox="0 0 15 16" version="1.1" role="img"><path fill-rule="evenodd" d="M10 12h3V2h-3v10zm-4-2h3V2H6v8zm-4 4h3V2H2v12zm-1 1h13V1H1v14zM14 0H1a1 1 0 00-1 1v14a1 1 0 001 1h13a1 1 0 001-1V1a1 1 0 00-1-1z"/></svg>
      <svg height="16" width="16" class="octicon octicon-search flex-shrink-0 js-jump-to-octicon-search d-none" title="Search" aria-label="Search" viewBox="0 0 16 16" version="1.1" role="img"><path fill-rule="evenodd" d="M15.7 13.3l-3.81-3.83A5.93 5.93 0 0013 6c0-3.31-2.69-6-6-6S1 2.69 1 6s2.69 6 6 6c1.3 0 2.48-.41 3.47-1.11l3.83 3.81c.19.2.45.3.7.3.25 0 .52-.09.7-.3a.996.996 0 000-1.41v.01zM7 10.7c-2.59 0-4.7-2.11-4.7-4.7 0-2.59 2.11-4.7 4.7-4.7 2.59 0 4.7 2.11 4.7 4.7 0 2.59-2.11 4.7-4.7 4.7z"/></svg>
    </div>

    <img class="avatar mr-2 flex-shrink-0 js-jump-to-suggestion-avatar d-none" alt="" aria-label="Team" src="" width="28" height="28">

    <div class="jump-to-suggestion-name js-jump-to-suggestion-name flex-auto overflow-hidden text-left no-wrap css-truncate css-truncate-target">
    </div>

    <div class="border rounded-1 flex-shrink-0 bg-gray px-1 text-gray-light ml-1 f6 d-none js-jump-to-badge-search">
      <span class="js-jump-to-badge-search-text-default d-none" aria-label="in this repository">
        In this repository
      </span>
      <span class="js-jump-to-badge-search-text-global d-none" aria-label="in all of GitHub">
        All GitHub
      </span>
      <span aria-hidden="true" class="d-inline-block ml-1 v-align-middle">↵</span>
    </div>

    <div aria-hidden="true" class="border rounded-1 flex-shrink-0 bg-gray px-1 text-gray-light ml-1 f6 d-none d-on-nav-focus js-jump-to-badge-jump">
      Jump to
      <span class="d-inline-block ml-1 v-align-middle">↵</span>
    </div>
  </a>
</li>

</ul>

<ul class="d-none js-jump-to-no-results-template-container">
  <li class="d-flex flex-justify-center flex-items-center f5 d-none js-jump-to-suggestion p-2">
    <span class="text-gray">No suggested jump to results</span>
  </li>
</ul>

<ul id="jump-to-results" role="listbox" class="p-0 m-0 js-navigation-container jump-to-suggestions-results-container js-jump-to-suggestions-results-container">

<li class="d-flex flex-justify-start flex-items-center p-0 f5 navigation-item js-navigation-item js-jump-to-scoped-search d-none" role="option">
  <a tabindex="-1" class="no-underline d-flex flex-auto flex-items-center jump-to-suggestions-path js-jump-to-suggestion-path js-navigation-open p-2" href="">
    <div class="jump-to-octicon js-jump-to-octicon flex-shrink-0 mr-2 text-center d-none">
      <svg height="16" width="16" class="octicon octicon-repo flex-shrink-0 js-jump-to-octicon-repo d-none" title="Repository" aria-label="Repository" viewBox="0 0 12 16" version="1.1" role="img"><path fill-rule="evenodd" d="M4 9H3V8h1v1zm0-3H3v1h1V6zm0-2H3v1h1V4zm0-2H3v1h1V2zm8-1v12c0 .55-.45 1-1 1H6v2l-1.5-1.5L3 16v-2H1c-.55 0-1-.45-1-1V1c0-.55.45-1 1-1h10c.55 0 1 .45 1 1zm-1 10H1v2h2v-1h3v1h5v-2zm0-10H2v9h9V1z"/></svg>
      <svg height="16" width="16" class="octicon octicon-project flex-shrink-0 js-jump-to-octicon-project d-none" title="Project" aria-label="Project" viewBox="0 0 15 16" version="1.1" role="img"><path fill-rule="evenodd" d="M10 12h3V2h-3v10zm-4-2h3V2H6v8zm-4 4h3V2H2v12zm-1 1h13V1H1v14zM14 0H1a1 1 0 00-1 1v14a1 1 0 001 1h13a1 1 0 001-1V1a1 1 0 00-1-1z"/></svg>
      <svg height="16" width="16" class="octicon octicon-search flex-shrink-0 js-jump-to-octicon-search d-none" title="Search" aria-label="Search" viewBox="0 0 16 16" version="1.1" role="img"><path fill-rule="evenodd" d="M15.7 13.3l-3.81-3.83A5.93 5.93 0 0013 6c0-3.31-2.69-6-6-6S1 2.69 1 6s2.69 6 6 6c1.3 0 2.48-.41 3.47-1.11l3.83 3.81c.19.2.45.3.7.3.25 0 .52-.09.7-.3a.996.996 0 000-1.41v.01zM7 10.7c-2.59 0-4.7-2.11-4.7-4.7 0-2.59 2.11-4.7 4.7-4.7 2.59 0 4.7 2.11 4.7 4.7 0 2.59-2.11 4.7-4.7 4.7z"/></svg>
    </div>

    <img class="avatar mr-2 flex-shrink-0 js-jump-to-suggestion-avatar d-none" alt="" aria-label="Team" src="" width="28" height="28">

    <div class="jump-to-suggestion-name js-jump-to-suggestion-name flex-auto overflow-hidden text-left no-wrap css-truncate css-truncate-target">
    </div>

    <div class="border rounded-1 flex-shrink-0 bg-gray px-1 text-gray-light ml-1 f6 d-none js-jump-to-badge-search">
      <span class="js-jump-to-badge-search-text-default d-none" aria-label="in this repository">
        In this repository
      </span>
      <span class="js-jump-to-badge-search-text-global d-none" aria-label="in all of GitHub">
        All GitHub
      </span>
      <span aria-hidden="true" class="d-inline-block ml-1 v-align-middle">↵</span>
    </div>

    <div aria-hidden="true" class="border rounded-1 flex-shrink-0 bg-gray px-1 text-gray-light ml-1 f6 d-none d-on-nav-focus js-jump-to-badge-jump">
      Jump to
      <span class="d-inline-block ml-1 v-align-middle">↵</span>
    </div>
  </a>
</li>

<li class="d-flex flex-justify-start flex-items-center p-0 f5 navigation-item js-navigation-item js-jump-to-global-search d-none" role="option">
  <a tabindex="-1" class="no-underline d-flex flex-auto flex-items-center jump-to-suggestions-path js-jump-to-suggestion-path js-navigation-open p-2" href="">
    <div class="jump-to-octicon js-jump-to-octicon flex-shrink-0 mr-2 text-center d-none">
      <svg height="16" width="16" class="octicon octicon-repo flex-shrink-0 js-jump-to-octicon-repo d-none" title="Repository" aria-label="Repository" viewBox="0 0 12 16" version="1.1" role="img"><path fill-rule="evenodd" d="M4 9H3V8h1v1zm0-3H3v1h1V6zm0-2H3v1h1V4zm0-2H3v1h1V2zm8-1v12c0 .55-.45 1-1 1H6v2l-1.5-1.5L3 16v-2H1c-.55 0-1-.45-1-1V1c0-.55.45-1 1-1h10c.55 0 1 .45 1 1zm-1 10H1v2h2v-1h3v1h5v-2zm0-10H2v9h9V1z"/></svg>
      <svg height="16" width="16" class="octicon octicon-project flex-shrink-0 js-jump-to-octicon-project d-none" title="Project" aria-label="Project" viewBox="0 0 15 16" version="1.1" role="img"><path fill-rule="evenodd" d="M10 12h3V2h-3v10zm-4-2h3V2H6v8zm-4 4h3V2H2v12zm-1 1h13V1H1v14zM14 0H1a1 1 0 00-1 1v14a1 1 0 001 1h13a1 1 0 001-1V1a1 1 0 00-1-1z"/></svg>
      <svg height="16" width="16" class="octicon octicon-search flex-shrink-0 js-jump-to-octicon-search d-none" title="Search" aria-label="Search" viewBox="0 0 16 16" version="1.1" role="img"><path fill-rule="evenodd" d="M15.7 13.3l-3.81-3.83A5.93 5.93 0 0013 6c0-3.31-2.69-6-6-6S1 2.69 1 6s2.69 6 6 6c1.3 0 2.48-.41 3.47-1.11l3.83 3.81c.19.2.45.3.7.3.25 0 .52-.09.7-.3a.996.996 0 000-1.41v.01zM7 10.7c-2.59 0-4.7-2.11-4.7-4.7 0-2.59 2.11-4.7 4.7-4.7 2.59 0 4.7 2.11 4.7 4.7 0 2.59-2.11 4.7-4.7 4.7z"/></svg>
    </div>

    <img class="avatar mr-2 flex-shrink-0 js-jump-to-suggestion-avatar d-none" alt="" aria-label="Team" src="" width="28" height="28">

    <div class="jump-to-suggestion-name js-jump-to-suggestion-name flex-auto overflow-hidden text-left no-wrap css-truncate css-truncate-target">
    </div>

    <div class="border rounded-1 flex-shrink-0 bg-gray px-1 text-gray-light ml-1 f6 d-none js-jump-to-badge-search">
      <span class="js-jump-to-badge-search-text-default d-none" aria-label="in this repository">
        In this repository
      </span>
      <span class="js-jump-to-badge-search-text-global d-none" aria-label="in all of GitHub">
        All GitHub
      </span>
      <span aria-hidden="true" class="d-inline-block ml-1 v-align-middle">↵</span>
    </div>

    <div aria-hidden="true" class="border rounded-1 flex-shrink-0 bg-gray px-1 text-gray-light ml-1 f6 d-none d-on-nav-focus js-jump-to-badge-jump">
      Jump to
      <span class="d-inline-block ml-1 v-align-middle">↵</span>
    </div>
  </a>
</li>

    <li class="d-flex flex-justify-center flex-items-center p-0 f5 js-jump-to-suggestion">
      <img src="https://github.githubassets.com/images/spinners/octocat-spinner-128.gif" alt="Octocat Spinner Icon" class="m-2" width="28">
    </li>
</ul>

            </div>
      </label>
</form>  </div>
</div>

      <nav class="d-flex flex-column flex-lg-row flex-self-stretch flex-lg-self-auto" aria-label="Global">
    <a class="Header-link d-block d-lg-none py-2 py-lg-0 border-top border-lg-top-0 border-white-fade-15" data-ga-click="Header, click, Nav menu - item:dashboard:user" aria-label="Dashboard" href="/dashboard">
      Dashboard
</a>

  <a class="js-selected-navigation-item Header-link  mr-0 mr-lg-3 py-2 py-lg-0 border-top border-lg-top-0 border-white-fade-15" data-hotkey="g p" data-ga-click="Header, click, Nav menu - item:pulls context:user" aria-label="Pull requests you created" data-selected-links="/pulls /pulls/assigned /pulls/mentioned /pulls" href="/pulls">
    Pull requests
</a>
  <a class="js-selected-navigation-item Header-link  mr-0 mr-lg-3 py-2 py-lg-0 border-top border-lg-top-0 border-white-fade-15" data-hotkey="g i" data-ga-click="Header, click, Nav menu - item:issues context:user" aria-label="Issues you created" data-selected-links="/issues /issues/assigned /issues/mentioned /issues" href="/issues">
    Issues
</a>
    <div class="mr-0 mr-lg-3 py-2 py-lg-0 border-top border-lg-top-0 border-white-fade-15">
      <a class="js-selected-navigation-item Header-link" data-ga-click="Header, click, Nav menu - item:marketplace context:user" data-octo-click="marketplace_click" data-octo-dimensions="location:nav_bar" data-selected-links=" /marketplace" href="/marketplace">
        Marketplace
</a>

    </div>

  <a class="js-selected-navigation-item Header-link  mr-0 mr-lg-3 py-2 py-lg-0 border-top border-lg-top-0 border-white-fade-15" data-ga-click="Header, click, Nav menu - item:explore" data-selected-links="/explore /trending /trending/developers /integrations /integrations/feature/code /integrations/feature/collaborate /integrations/feature/ship showcases showcases_search showcases_landing /explore" href="/explore">
    Explore
</a>

    <a class="Header-link d-block d-lg-none mr-0 mr-lg-3 py-2 py-lg-0 border-top border-lg-top-0 border-white-fade-15" href="https://github.com/chguiterman">
      <img class="avatar" height="20" width="20" alt="@chguiterman" src="https://avatars1.githubusercontent.com/u/15913450?s=60&amp;v=4" />
      chguiterman
</a>
    <!-- '"` --><!-- </textarea></xmp> --></option></form><form action="/logout" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="o+BRZjRQl/uTKGA3Jv6BzRg1Mj4Mur0BLBfZc5MWAE4PnYl8BvRIZiYfn+VpaOtVyB2NLPHwzgFrG0WBpfxd1w==" />
      <button type="submit" class="Header-link mr-0 mr-lg-3 py-2 py-lg-0 border-top border-lg-top-0 border-white-fade-15 d-lg-none btn-link d-block width-full text-left" data-ga-click="Header, sign out, icon:logout" style="padding-left: 2px;">
        <svg class="octicon octicon-sign-out v-align-middle" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M12 9V7H8V5h4V3l4 3-4 3zm-2 3H6V3L2 1h8v3h1V1c0-.55-.45-1-1-1H1C.45 0 0 .45 0 1v11.38c0 .39.22.73.55.91L6 16.01V13h4c.55 0 1-.45 1-1V8h-1v4z"/></svg>
        Sign out
      </button>
</form></nav>

    </div>

    <div class="Header-item Header-item--full flex-justify-center d-lg-none position-relative">
      <div class="css-truncate css-truncate-target width-fit position-absolute left-0 right-0 text-center">
              <svg class="octicon octicon-repo" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M4 9H3V8h1v1zm0-3H3v1h1V6zm0-2H3v1h1V4zm0-2H3v1h1V2zm8-1v12c0 .55-.45 1-1 1H6v2l-1.5-1.5L3 16v-2H1c-.55 0-1-.45-1-1V1c0-.55.45-1 1-1h10c.55 0 1 .45 1 1zm-1 10H1v2h2v-1h3v1h5v-2zm0-10H2v9h9V1z"/></svg>
    <a class="Header-link" href="/svmiller">svmiller</a>
    /
    <a class="Header-link" href="/svmiller/svm-r-markdown-templates">svm-r-markdown-templates</a>

</div>
    </div>

    <div class="Header-item mr-0 mr-lg-3 flex-order-1 flex-lg-order-none">

    <a aria-label="You have no unread notifications" class="Header-link notification-indicator position-relative tooltipped tooltipped-sw js-socket-channel js-notification-indicator" data-hotkey="g n" data-ga-click="Header, go to notifications, icon:read" data-channel="notification-changed:15913450" href="/notifications/beta">
        <span class="js-indicator-modifier mail-status "></span>
        <svg class="octicon octicon-bell" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M14 12v1H0v-1l.73-.58c.77-.77.81-2.55 1.19-4.42C2.69 3.23 6 2 6 2c0-.55.45-1 1-1s1 .45 1 1c0 0 3.39 1.23 4.16 5 .38 1.88.42 3.66 1.19 4.42l.66.58H14zm-7 4c1.11 0 2-.89 2-2H5c0 1.11.89 2 2 2z"/></svg>
</a>
    </div>

    <div class="Header-item position-relative d-none d-lg-flex">
      <details class="details-overlay details-reset">
  <summary class="Header-link"
      aria-label="Create new…"
      data-ga-click="Header, create new, icon:add">
    <svg class="octicon octicon-plus" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M12 9H7v5H5V9H0V7h5V2h2v5h5v2z"/></svg> <span class="dropdown-caret"></span>
  </summary>
  <details-menu class="dropdown-menu dropdown-menu-sw">

<a role="menuitem" class="dropdown-item" href="/new" data-ga-click="Header, create new repository">
  New repository
</a>

  <a role="menuitem" class="dropdown-item" href="/new/import" data-ga-click="Header, import a repository">
    Import repository
  </a>

<a role="menuitem" class="dropdown-item" href="https://gist.github.com/" data-ga-click="Header, create new gist">
  New gist
</a>

  <a role="menuitem" class="dropdown-item" href="/organizations/new" data-ga-click="Header, create new organization">
    New organization
  </a>

  <div role="none" class="dropdown-divider"></div>
  <div class="dropdown-header">
    <span title="svmiller/svm-r-markdown-templates">This repository</span>
  </div>
    <a role="menuitem" class="dropdown-item" href="/svmiller/svm-r-markdown-templates/issues/new" data-ga-click="Header, create new issue" data-skip-pjax>
      New issue
    </a>

  </details-menu>
</details>

    </div>

    <div class="Header-item position-relative mr-0 d-none d-lg-flex">

  <details class="details-overlay details-reset js-feature-preview-indicator-container" data-feature-preview-indicator-src="/users/chguiterman/feature_preview/indicator_check.json">

  <summary class="Header-link"
    aria-label="View profile and more"
    data-ga-click="Header, show menu, icon:avatar">
    <img class="avatar " alt="@chguiterman" width="20" height="20" src="https://avatars1.githubusercontent.com/u/15913450?s=60&amp;v=4">

      <span class="feature-preview-indicator js-feature-preview-indicator" hidden></span>
    <span class="dropdown-caret"></span>
  </summary>
  <details-menu class="dropdown-menu dropdown-menu-sw mt-2" style="width: 180px">
    <div class="header-nav-current-user css-truncate"><a role="menuitem" class="no-underline user-profile-link px-3 pt-2 pb-2 mb-n2 mt-n1 d-block" href="/chguiterman" data-ga-click="Header, go to profile, text:Signed in as">Signed in as <strong class="css-truncate-target">chguiterman</strong></a></div>
    <div role="none" class="dropdown-divider"></div>

      <div class="pl-3 pr-3 f6 user-status-container js-user-status-context pb-1" data-url="/users/status?compact=1&amp;link_mentions=0&amp;truncate=1">

<div class="js-user-status-container
    user-status-compact rounded-1 px-2 py-1 mt-2
    border
  " data-team-hovercards-enabled>
  <details class="js-user-status-details details-reset details-overlay details-overlay-dark">
    <summary class="btn-link btn-block link-gray no-underline js-toggle-user-status-edit toggle-user-status-edit "
      role="menuitem" data-hydro-click="{&quot;event_type&quot;:&quot;user_profile.click&quot;,&quot;payload&quot;:{&quot;profile_user_id&quot;:4589542,&quot;target&quot;:&quot;EDIT_USER_STATUS&quot;,&quot;user_id&quot;:15913450,&quot;originating_url&quot;:&quot;https://github.com/svmiller/svm-r-markdown-templates/blob/master/svm-latex-ms.tex&quot;}}" data-hydro-click-hmac="b58643669f9ed4ff6fff55e1a2e3353f42fe89d8e223233b7bdd445afb6b311f">
      <div class="d-flex">
        <div class="f6 lh-condensed user-status-header
          d-inline-block v-align-middle
            user-status-emoji-only-header circle
            pr-2
"
            style="max-width: 29px"
          >
          <div class="user-status-emoji-container flex-shrink-0 mr-1 mt-1 lh-condensed-ultra v-align-bottom" style="">
            <svg class="octicon octicon-smiley" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8 0C3.58 0 0 3.58 0 8s3.58 8 8 8 8-3.58 8-8-3.58-8-8-8zm4.81 12.81a6.72 6.72 0 01-2.17 1.45c-.83.36-1.72.53-2.64.53-.92 0-1.81-.17-2.64-.53-.81-.34-1.55-.83-2.17-1.45a6.773 6.773 0 01-1.45-2.17A6.59 6.59 0 011.21 8c0-.92.17-1.81.53-2.64.34-.81.83-1.55 1.45-2.17.62-.62 1.36-1.11 2.17-1.45A6.59 6.59 0 018 1.21c.92 0 1.81.17 2.64.53.81.34 1.55.83 2.17 1.45.62.62 1.11 1.36 1.45 2.17.36.83.53 1.72.53 2.64 0 .92-.17 1.81-.53 2.64-.34.81-.83 1.55-1.45 2.17zM4 6.8v-.59c0-.66.53-1.19 1.2-1.19h.59c.66 0 1.19.53 1.19 1.19v.59c0 .67-.53 1.2-1.19 1.2H5.2C4.53 8 4 7.47 4 6.8zm5 0v-.59c0-.66.53-1.19 1.2-1.19h.59c.66 0 1.19.53 1.19 1.19v.59c0 .67-.53 1.2-1.19 1.2h-.59C9.53 8 9 7.47 9 6.8zm4 3.2c-.72 1.88-2.91 3-5 3s-4.28-1.13-5-3c-.14-.39.23-1 .66-1h8.59c.41 0 .89.61.75 1z"/></svg>
          </div>
        </div>
        <div class="
          d-inline-block v-align-middle

           css-truncate css-truncate-target 
           user-status-message-wrapper f6"
           style="line-height: 20px;" >
          <div class="d-inline-block text-gray-dark v-align-text-top text-left">
              <span class="text-gray ml-2">Set status</span>
          </div>
        </div>
      </div>
    </summary>
    <details-dialog class="details-dialog rounded-1 anim-fade-in fast Box Box--overlay" role="dialog" tabindex="-1">
      <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="position-relative flex-auto js-user-status-form" action="/users/status?compact=1&amp;link_mentions=0&amp;truncate=1" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="_method" value="put" /><input type="hidden" name="authenticity_token" value="hhUlUz4xnfqkMYXOkI7PRKaN+Sg36Lb4bOIb/QGw+1jEV5oKmf2fRAyXcQ9c4CeWxMyenARosA3qe1z0+yvHdQ==" />
        <div class="Box-header bg-gray border-bottom p-3">
          <button class="Box-btn-octicon js-toggle-user-status-edit btn-octicon float-right" type="reset" aria-label="Close dialog" data-close-dialog>
            <svg class="octicon octicon-x" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48L7.48 8z"/></svg>
          </button>
          <h3 class="Box-title f5 text-bold text-gray-dark">Edit status</h3>
        </div>
        <input type="hidden" name="emoji" class="js-user-status-emoji-field" value="">
        <input type="hidden" name="organization_id" class="js-user-status-org-id-field" value="">
        <div class="px-3 py-2 text-gray-dark">
          <div class="js-characters-remaining-container position-relative mt-2">
            <div class="input-group d-table form-group my-0 js-user-status-form-group">
              <span class="input-group-button d-table-cell v-align-middle" style="width: 1%">
                <button type="button" aria-label="Choose an emoji" class="btn-outline btn js-toggle-user-status-emoji-picker btn-open-emoji-picker p-0">
                  <span class="js-user-status-original-emoji" hidden></span>
                  <span class="js-user-status-custom-emoji"></span>
                  <span class="js-user-status-no-emoji-icon" >
                    <svg class="octicon octicon-smiley" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8 0C3.58 0 0 3.58 0 8s3.58 8 8 8 8-3.58 8-8-3.58-8-8-8zm4.81 12.81a6.72 6.72 0 01-2.17 1.45c-.83.36-1.72.53-2.64.53-.92 0-1.81-.17-2.64-.53-.81-.34-1.55-.83-2.17-1.45a6.773 6.773 0 01-1.45-2.17A6.59 6.59 0 011.21 8c0-.92.17-1.81.53-2.64.34-.81.83-1.55 1.45-2.17.62-.62 1.36-1.11 2.17-1.45A6.59 6.59 0 018 1.21c.92 0 1.81.17 2.64.53.81.34 1.55.83 2.17 1.45.62.62 1.11 1.36 1.45 2.17.36.83.53 1.72.53 2.64 0 .92-.17 1.81-.53 2.64-.34.81-.83 1.55-1.45 2.17zM4 6.8v-.59c0-.66.53-1.19 1.2-1.19h.59c.66 0 1.19.53 1.19 1.19v.59c0 .67-.53 1.2-1.19 1.2H5.2C4.53 8 4 7.47 4 6.8zm5 0v-.59c0-.66.53-1.19 1.2-1.19h.59c.66 0 1.19.53 1.19 1.19v.59c0 .67-.53 1.2-1.19 1.2h-.59C9.53 8 9 7.47 9 6.8zm4 3.2c-.72 1.88-2.91 3-5 3s-4.28-1.13-5-3c-.14-.39.23-1 .66-1h8.59c.41 0 .89.61.75 1z"/></svg>
                  </span>
                </button>
              </span>
              <text-expander keys=": @" data-mention-url="/autocomplete/user-suggestions" data-emoji-url="/autocomplete/emoji">
                <input
                  type="text"
                  autocomplete="off"
                  data-no-org-url="/autocomplete/user-suggestions"
                  data-org-url="/suggestions?mention_suggester=1"
                  data-maxlength="80"
                  class="d-table-cell width-full form-control js-user-status-message-field js-characters-remaining-field"
                  placeholder="What's happening?"
                  name="message"
                  value=""
                  aria-label="What is your current status?">
              </text-expander>
              <div class="error">Could not update your status, please try again.</div>
            </div>
            <div style="margin-left: 53px" class="my-1 text-small label-characters-remaining js-characters-remaining" data-suffix="remaining" hidden>
              80 remaining
            </div>
          </div>
          <include-fragment class="js-user-status-emoji-picker" data-url="/users/status/emoji"></include-fragment>
          <div class="overflow-auto ml-n3 mr-n3 px-3 border-bottom" style="max-height: 33vh">
            <div class="user-status-suggestions js-user-status-suggestions collapsed overflow-hidden">
              <h4 class="f6 text-normal my-3">Suggestions:</h4>
              <div class="mx-3 mt-2 clearfix">
                  <div class="float-left col-6">
                      <button type="button" value=":palm_tree:" class="d-flex flex-items-baseline flex-items-stretch lh-condensed f6 btn-link link-gray no-underline js-predefined-user-status mb-1">
                        <div class="emoji-status-width mr-2 v-align-middle js-predefined-user-status-emoji">
                          <g-emoji alias="palm_tree" fallback-src="https://github.githubassets.com/images/icons/emoji/unicode/1f334.png">🌴</g-emoji>
                        </div>
                        <div class="d-flex flex-items-center no-underline js-predefined-user-status-message ws-normal text-left" style="border-left: 1px solid transparent">
                          On vacation
                        </div>
                      </button>
                      <button type="button" value=":face_with_thermometer:" class="d-flex flex-items-baseline flex-items-stretch lh-condensed f6 btn-link link-gray no-underline js-predefined-user-status mb-1">
                        <div class="emoji-status-width mr-2 v-align-middle js-predefined-user-status-emoji">
                          <g-emoji alias="face_with_thermometer" fallback-src="https://github.githubassets.com/images/icons/emoji/unicode/1f912.png">🤒</g-emoji>
                        </div>
                        <div class="d-flex flex-items-center no-underline js-predefined-user-status-message ws-normal text-left" style="border-left: 1px solid transparent">
                          Out sick
                        </div>
                      </button>
                  </div>
                  <div class="float-left col-6">
                      <button type="button" value=":house:" class="d-flex flex-items-baseline flex-items-stretch lh-condensed f6 btn-link link-gray no-underline js-predefined-user-status mb-1">
                        <div class="emoji-status-width mr-2 v-align-middle js-predefined-user-status-emoji">
                          <g-emoji alias="house" fallback-src="https://github.githubassets.com/images/icons/emoji/unicode/1f3e0.png">🏠</g-emoji>
                        </div>
                        <div class="d-flex flex-items-center no-underline js-predefined-user-status-message ws-normal text-left" style="border-left: 1px solid transparent">
                          Working from home
                        </div>
                      </button>
                      <button type="button" value=":dart:" class="d-flex flex-items-baseline flex-items-stretch lh-condensed f6 btn-link link-gray no-underline js-predefined-user-status mb-1">
                        <div class="emoji-status-width mr-2 v-align-middle js-predefined-user-status-emoji">
                          <g-emoji alias="dart" fallback-src="https://github.githubassets.com/images/icons/emoji/unicode/1f3af.png">🎯</g-emoji>
                        </div>
                        <div class="d-flex flex-items-center no-underline js-predefined-user-status-message ws-normal text-left" style="border-left: 1px solid transparent">
                          Focusing
                        </div>
                      </button>
                  </div>
              </div>
            </div>
            <div class="user-status-limited-availability-container">
              <div class="form-checkbox my-0">
                <input type="checkbox" name="limited_availability" value="1" class="js-user-status-limited-availability-checkbox" data-default-message="I may be slow to respond." aria-describedby="limited-availability-help-text-truncate-true-compact-true" id="limited-availability-truncate-true-compact-true">
                <label class="d-block f5 text-gray-dark mb-1" for="limited-availability-truncate-true-compact-true">
                  Busy
                </label>
                <p class="note" id="limited-availability-help-text-truncate-true-compact-true">
                  When others mention you, assign you, or request your review,
                  GitHub will let them know that you have limited availability.
                </p>
              </div>
            </div>
          </div>
          <div class="d-inline-block f5 mr-2 pt-3 pb-2" >
  <div class="d-inline-block mr-1">
    Clear status
  </div>

  <details class="js-user-status-expire-drop-down f6 dropdown details-reset details-overlay d-inline-block mr-2">
    <summary class="f5 btn-link link-gray-dark border px-2 py-1 rounded-1" aria-haspopup="true">
      <div class="js-user-status-expiration-interval-selected d-inline-block v-align-baseline">
        Never
      </div>
      <div class="dropdown-caret"></div>
    </summary>

    <ul class="dropdown-menu dropdown-menu-se pl-0 overflow-auto" style="width: 220px; max-height: 15.5em">
      <li>
        <button type="button" class="btn-link dropdown-item js-user-status-expire-button ws-normal" title="Never">
          <span class="d-inline-block text-bold mb-1">Never</span>
          <div class="f6 lh-condensed">Keep this status until you clear your status or edit your status.</div>
        </button>
      </li>
      <li class="dropdown-divider" role="none"></li>
        <li>
          <button type="button" class="btn-link dropdown-item ws-normal js-user-status-expire-button" title="in 30 minutes" value="2020-02-25T10:25:25-05:00">
            in 30 minutes
          </button>
        </li>
        <li>
          <button type="button" class="btn-link dropdown-item ws-normal js-user-status-expire-button" title="in 1 hour" value="2020-02-25T10:55:25-05:00">
            in 1 hour
          </button>
        </li>
        <li>
          <button type="button" class="btn-link dropdown-item ws-normal js-user-status-expire-button" title="in 4 hours" value="2020-02-25T13:55:25-05:00">
            in 4 hours
          </button>
        </li>
        <li>
          <button type="button" class="btn-link dropdown-item ws-normal js-user-status-expire-button" title="today" value="2020-02-25T23:59:59-05:00">
            today
          </button>
        </li>
        <li>
          <button type="button" class="btn-link dropdown-item ws-normal js-user-status-expire-button" title="this week" value="2020-03-01T23:59:59-05:00">
            this week
          </button>
        </li>
    </ul>
  </details>
  <input class="js-user-status-expiration-date-input" type="hidden" name="expires_at" value="">
</div>

          <include-fragment class="js-user-status-org-picker" data-url="/users/status/organizations"></include-fragment>
        </div>
        <div class="d-flex flex-items-center flex-justify-between p-3 border-top">
          <button type="submit" disabled class="width-full btn btn-primary mr-2 js-user-status-submit">
            Set status
          </button>
          <button type="button" disabled class="width-full js-clear-user-status-button btn ml-2 ">
            Clear status
          </button>
        </div>
</form>    </details-dialog>
  </details>
</div>

      </div>
      <div role="none" class="dropdown-divider"></div>

    <a role="menuitem" class="dropdown-item" href="/chguiterman" data-ga-click="Header, go to profile, text:your profile">Your profile</a>

    <a role="menuitem" class="dropdown-item" href="/chguiterman?tab=repositories" data-ga-click="Header, go to repositories, text:your repositories">Your repositories</a>

    <a role="menuitem" class="dropdown-item" href="/chguiterman?tab=projects" data-ga-click="Header, go to projects, text:your projects">Your projects</a>

    <a role="menuitem" class="dropdown-item" href="/chguiterman?tab=stars" data-ga-click="Header, go to starred repos, text:your stars">Your stars</a>
      <a role="menuitem" class="dropdown-item" href="https://gist.github.com/mine" data-ga-click="Header, your gists, text:your gists">Your gists</a>

    <div role="none" class="dropdown-divider"></div>

<div id="feature-enrollment-toggle" class="hide-sm hide-md feature-preview-details position-relative">
  <button
    type="button"
    class="dropdown-item btn-link"
    role="menuitem"
    data-feature-preview-trigger-url="/users/chguiterman/feature_previews"
    data-feature-preview-close-details="{&quot;event_type&quot;:&quot;feature_preview.clicks.close_modal&quot;,&quot;payload&quot;:{&quot;originating_url&quot;:&quot;https://github.com/svmiller/svm-r-markdown-templates/blob/master/svm-latex-ms.tex&quot;,&quot;user_id&quot;:15913450}}"
    data-feature-preview-close-hmac="02f102fae434dcaa4c6400a8dda03141d36f52ab222a13a8c6a6dfb2a7757ae6"
    data-hydro-click="{&quot;event_type&quot;:&quot;feature_preview.clicks.open_modal&quot;,&quot;payload&quot;:{&quot;link_location&quot;:&quot;user_dropdown&quot;,&quot;originating_url&quot;:&quot;https://github.com/svmiller/svm-r-markdown-templates/blob/master/svm-latex-ms.tex&quot;,&quot;user_id&quot;:15913450}}"
    data-hydro-click-hmac="6350418342ebebac195aa183d82ab88e098ce639a80ee1c5770c5837f86a5e1d"
  >
    Feature preview
  </button>
    <span class="feature-preview-indicator js-feature-preview-indicator" hidden></span>
</div>

    <a role="menuitem" class="dropdown-item" href="https://help.github.com" data-ga-click="Header, go to help, text:help">Help</a>
    <a role="menuitem" class="dropdown-item" href="/settings/profile" data-ga-click="Header, go to settings, icon:settings">Settings</a>
    <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="logout-form" action="/logout" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="FyHjUq30A1FCh0y0H3Ms/rwF+9+9CoWFYr0KECpa/yG7XDtIn1DczPews2ZQ5UZmbC1EzUBA9oUlsZbiHLCiuA==" />

      <button type="submit" class="dropdown-item dropdown-signout" data-ga-click="Header, sign out, icon:logout" role="menuitem">
        Sign out
      </button>
      <input type="text" name="required_field_4868" hidden="hidden" class="form-control" /><input type="hidden" name="timestamp" value="1582642525959" class="form-control" /><input type="hidden" name="timestamp_secret" value="200f0952d01cf9b2b6b89c5ea3f53935870c1a13f04de7aff92853019b5d8470" class="form-control" />
</form>  </details-menu>
</details>

    </div>

  </header>

  </div>

  <div id="start-of-content" class="show-on-focus"></div>

    <div id="js-flash-container">

</div>

  <include-fragment class="js-notification-shelf-include-fragment" data-base-src="https://github.com/notifications/beta/shelf"></include-fragment>

  <div class="application-main " data-commit-hovercards-enabled>
        <div itemscope itemtype="http://schema.org/SoftwareSourceCode" class="">
    <main  >

  <div class="pagehead repohead hx_repohead readability-menu bg-gray-light pb-0 pt-0 pt-lg-3">

    <div class="container-lg mb-4 p-responsive d-none d-lg-flex">

      <div class="flex-auto min-width-0 width-fit mr-3">
        <h1 class="public  d-flex flex-wrap flex-items-center break-word float-none ">
    <svg class="octicon octicon-repo" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M4 9H3V8h1v1zm0-3H3v1h1V6zm0-2H3v1h1V4zm0-2H3v1h1V2zm8-1v12c0 .55-.45 1-1 1H6v2l-1.5-1.5L3 16v-2H1c-.55 0-1-.45-1-1V1c0-.55.45-1 1-1h10c.55 0 1 .45 1 1zm-1 10H1v2h2v-1h3v1h5v-2zm0-10H2v9h9V1z"/></svg>
  <span class="author ml-1 flex-self-stretch" itemprop="author">
    <a class="url fn" rel="author" data-hovercard-type="user" data-hovercard-url="/users/svmiller/hovercard" data-octo-click="hovercard-link-click" data-octo-dimensions="link_type:self" href="/svmiller">svmiller</a>
  </span>
  <span class="path-divider flex-self-stretch">/</span>
  <strong itemprop="name" class="mr-2 flex-self-stretch">
    <a data-pjax="#js-repo-pjax-container" href="/svmiller/svm-r-markdown-templates">svm-r-markdown-templates</a>
  </strong>

</h1>

      </div>

      <ul class="pagehead-actions flex-shrink-0"  >

  <li>

    <!-- '"` --><!-- </textarea></xmp> --></option></form><form data-remote="true" class="clearfix js-social-form js-social-container" action="/notifications/subscribe" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="xbhbQ3heSpGzXBrsYjSrm5YJe1yw9q2NcLIiCvcVUUJSnfnFqKjaYWTuVMePR91NWzkfElOZBncoyyeMJAImpA==" />      <input type="hidden" name="repository_id" value="51473841">

      <details class="details-reset details-overlay select-menu float-left">
        <summary class="select-menu-button float-left btn btn-sm btn-with-count" data-hydro-click="{&quot;event_type&quot;:&quot;repository.click&quot;,&quot;payload&quot;:{&quot;target&quot;:&quot;WATCH_BUTTON&quot;,&quot;repository_id&quot;:51473841,&quot;originating_url&quot;:&quot;https://github.com/svmiller/svm-r-markdown-templates/blob/master/svm-latex-ms.tex&quot;,&quot;user_id&quot;:15913450}}" data-hydro-click-hmac="179a94da15c229e52314fef95f451ea36dea872d96443fbf90fb6e95328edc34" data-ga-click="Repository, click Watch settings, action:blob#show">          <span data-menu-button>
              <svg class="octicon octicon-eye v-align-text-bottom" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.06 2C3 2 0 8 0 8s3 6 8.06 6C13 14 16 8 16 8s-3-6-7.94-6zM8 12c-2.2 0-4-1.78-4-4 0-2.2 1.8-4 4-4 2.22 0 4 1.8 4 4 0 2.22-1.78 4-4 4zm2-4c0 1.11-.89 2-2 2-1.11 0-2-.89-2-2 0-1.11.89-2 2-2 1.11 0 2 .89 2 2z"/></svg>
              Watch
          </span>
</summary>        <details-menu
          class="select-menu-modal position-absolute mt-5"
          style="z-index: 99;">
          <div class="select-menu-header">
            <span class="select-menu-title">Notifications</span>
          </div>
          <div class="select-menu-list">
            <button type="submit" name="do" value="included" class="select-menu-item width-full" aria-checked="true" role="menuitemradio">
              <svg class="octicon octicon-check select-menu-item-icon" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M12 5l-8 8-4-4 1.5-1.5L4 10l6.5-6.5L12 5z"/></svg>
              <div class="select-menu-item-text">
                <span class="select-menu-item-heading">Not watching</span>
                <span class="description">Be notified only when participating or @mentioned.</span>
                <span class="hidden-select-button-text" data-menu-button-contents>
                  <svg class="octicon octicon-eye v-align-text-bottom" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.06 2C3 2 0 8 0 8s3 6 8.06 6C13 14 16 8 16 8s-3-6-7.94-6zM8 12c-2.2 0-4-1.78-4-4 0-2.2 1.8-4 4-4 2.22 0 4 1.8 4 4 0 2.22-1.78 4-4 4zm2-4c0 1.11-.89 2-2 2-1.11 0-2-.89-2-2 0-1.11.89-2 2-2 1.11 0 2 .89 2 2z"/></svg>
                  Watch
                </span>
              </div>
            </button>

            <button type="submit" name="do" value="release_only" class="select-menu-item width-full" aria-checked="false" role="menuitemradio">
              <svg class="octicon octicon-check select-menu-item-icon" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M12 5l-8 8-4-4 1.5-1.5L4 10l6.5-6.5L12 5z"/></svg>
              <div class="select-menu-item-text">
                <span class="select-menu-item-heading">Releases only</span>
                <span class="description">Be notified of new releases, and when participating or @mentioned.</span>
                <span class="hidden-select-button-text" data-menu-button-contents>
                  <svg class="octicon octicon-eye v-align-text-bottom" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.06 2C3 2 0 8 0 8s3 6 8.06 6C13 14 16 8 16 8s-3-6-7.94-6zM8 12c-2.2 0-4-1.78-4-4 0-2.2 1.8-4 4-4 2.22 0 4 1.8 4 4 0 2.22-1.78 4-4 4zm2-4c0 1.11-.89 2-2 2-1.11 0-2-.89-2-2 0-1.11.89-2 2-2 1.11 0 2 .89 2 2z"/></svg>
                  Unwatch releases
                </span>
              </div>
            </button>

            <button type="submit" name="do" value="subscribed" class="select-menu-item width-full" aria-checked="false" role="menuitemradio">
              <svg class="octicon octicon-check select-menu-item-icon" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M12 5l-8 8-4-4 1.5-1.5L4 10l6.5-6.5L12 5z"/></svg>
              <div class="select-menu-item-text">
                <span class="select-menu-item-heading">Watching</span>
                <span class="description">Be notified of all conversations.</span>
                <span class="hidden-select-button-text" data-menu-button-contents>
                  <svg class="octicon octicon-eye v-align-text-bottom" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.06 2C3 2 0 8 0 8s3 6 8.06 6C13 14 16 8 16 8s-3-6-7.94-6zM8 12c-2.2 0-4-1.78-4-4 0-2.2 1.8-4 4-4 2.22 0 4 1.8 4 4 0 2.22-1.78 4-4 4zm2-4c0 1.11-.89 2-2 2-1.11 0-2-.89-2-2 0-1.11.89-2 2-2 1.11 0 2 .89 2 2z"/></svg>
                  Unwatch
                </span>
              </div>
            </button>

            <button type="submit" name="do" value="ignore" class="select-menu-item width-full" aria-checked="false" role="menuitemradio">
              <svg class="octicon octicon-check select-menu-item-icon" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M12 5l-8 8-4-4 1.5-1.5L4 10l6.5-6.5L12 5z"/></svg>
              <div class="select-menu-item-text">
                <span class="select-menu-item-heading">Ignoring</span>
                <span class="description">Never be notified.</span>
                <span class="hidden-select-button-text" data-menu-button-contents>
                  <svg class="octicon octicon-mute v-align-text-bottom" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8 2.81v10.38c0 .67-.81 1-1.28.53L3 10H1c-.55 0-1-.45-1-1V7c0-.55.45-1 1-1h2l3.72-3.72C7.19 1.81 8 2.14 8 2.81zm7.53 3.22l-1.06-1.06-1.97 1.97-1.97-1.97-1.06 1.06L11.44 8 9.47 9.97l1.06 1.06 1.97-1.97 1.97 1.97 1.06-1.06L13.56 8l1.97-1.97z"/></svg>
                  Stop ignoring
                </span>
              </div>
            </button>
          </div>
        </details-menu>
      </details>
        <a class="social-count js-social-count"
          href="/svmiller/svm-r-markdown-templates/watchers"
          aria-label="31 users are watching this repository">
          31
        </a>
</form>
  </li>

  <li>
      <div class="js-toggler-container js-social-container starring-container ">
    <form class="starred js-social-form" action="/svmiller/svm-r-markdown-templates/unstar" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="olVK1zKHqHqWkP3PScq4Zn9512sKmrTeG/h37dOIXP+P4efdDpr68mvAXgrGjWbrKttS/SgWFd/laClxPFDKJw==" />
      <input type="hidden" name="context" value="repository"></input>
      <button type="submit" class="btn btn-sm btn-with-count js-toggler-target" aria-label="Unstar this repository" title="Unstar svmiller/svm-r-markdown-templates" data-hydro-click="{&quot;event_type&quot;:&quot;repository.click&quot;,&quot;payload&quot;:{&quot;target&quot;:&quot;UNSTAR_BUTTON&quot;,&quot;repository_id&quot;:51473841,&quot;originating_url&quot;:&quot;https://github.com/svmiller/svm-r-markdown-templates/blob/master/svm-latex-ms.tex&quot;,&quot;user_id&quot;:15913450}}" data-hydro-click-hmac="3101eb0638194f34d67df3bdf73d9a683a53bb1cdf419d9a24d3b6e9a0a658fc" data-ga-click="Repository, click unstar button, action:blob#show; text:Unstar">        <svg aria-label="star" height="16" class="octicon octicon-star v-align-text-bottom" viewBox="0 0 14 16" version="1.1" width="14" role="img"><path fill-rule="evenodd" d="M14 6l-4.9-.64L7 1 4.9 5.36 0 6l3.6 3.26L2.67 14 7 11.67 11.33 14l-.93-4.74L14 6z"/></svg>

        Unstar
</button>        <a class="social-count js-social-count" href="/svmiller/svm-r-markdown-templates/stargazers"
           aria-label="503 users starred this repository">
           503
        </a>
</form>
    <form class="unstarred js-social-form" action="/svmiller/svm-r-markdown-templates/star" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="u++gC0unU2W0rW5JR33F1dgRe6RzCezDDaB/gZRPGG9861Qaz3P/Bk1rpZdMxnHn0jUxRf93V4iL+2gSkQpvAQ==" />
      <input type="hidden" name="context" value="repository"></input>
      <button type="submit" class="btn btn-sm btn-with-count js-toggler-target" aria-label="Unstar this repository" title="Star svmiller/svm-r-markdown-templates" data-hydro-click="{&quot;event_type&quot;:&quot;repository.click&quot;,&quot;payload&quot;:{&quot;target&quot;:&quot;STAR_BUTTON&quot;,&quot;repository_id&quot;:51473841,&quot;originating_url&quot;:&quot;https://github.com/svmiller/svm-r-markdown-templates/blob/master/svm-latex-ms.tex&quot;,&quot;user_id&quot;:15913450}}" data-hydro-click-hmac="007904ea97d87d2a9a87db4e1ddcec7f69a58a78e0657f88b54d8ffb053643c4" data-ga-click="Repository, click star button, action:blob#show; text:Star">        <svg aria-label="star" height="16" class="octicon octicon-star v-align-text-bottom" viewBox="0 0 14 16" version="1.1" width="14" role="img"><path fill-rule="evenodd" d="M14 6l-4.9-.64L7 1 4.9 5.36 0 6l3.6 3.26L2.67 14 7 11.67 11.33 14l-.93-4.74L14 6z"/></svg>

        Star
</button>        <a class="social-count js-social-count" href="/svmiller/svm-r-markdown-templates/stargazers"
           aria-label="503 users starred this repository">
          503
        </a>
</form>  </div>

  </li>

  <li>
          <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="btn-with-count" action="/svmiller/svm-r-markdown-templates/fork" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="HYw/M4aTp32OVyrUwTEFm9a/0Diu19ONSKwhEwVh1jUWhKD5jknjyw1dqs5nWcVw6uDDINDmzyBpD8eOionR0A==" />
            <button class="btn btn-sm btn-with-count" data-hydro-click="{&quot;event_type&quot;:&quot;repository.click&quot;,&quot;payload&quot;:{&quot;target&quot;:&quot;FORK_BUTTON&quot;,&quot;repository_id&quot;:51473841,&quot;originating_url&quot;:&quot;https://github.com/svmiller/svm-r-markdown-templates/blob/master/svm-latex-ms.tex&quot;,&quot;user_id&quot;:15913450}}" data-hydro-click-hmac="601bd52e0727cbd47706f20bb76f7dba81ab51994dd1486fa09d92913bafc926" data-ga-click="Repository, show fork modal, action:blob#show; text:Fork" type="submit" title="Fork your own copy of svmiller/svm-r-markdown-templates to your account" aria-label="Fork your own copy of svmiller/svm-r-markdown-templates to your account">              <svg class="octicon octicon-repo-forked v-align-text-bottom" viewBox="0 0 10 16" version="1.1" width="10" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8 1a1.993 1.993 0 00-1 3.72V6L5 8 3 6V4.72A1.993 1.993 0 002 1a1.993 1.993 0 00-1 3.72V6.5l3 3v1.78A1.993 1.993 0 005 15a1.993 1.993 0 001-3.72V9.5l3-3V4.72A1.993 1.993 0 008 1zM2 4.2C1.34 4.2.8 3.65.8 3c0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2zm3 10c-.66 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2zm3-10c-.66 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2z"/></svg>
              Fork
</button></form>
    <a href="/svmiller/svm-r-markdown-templates/network/members" class="social-count"
       aria-label="405 users forked this repository">
      405
    </a>
  </li>
</ul>

    </div>

<nav class="hx_reponav reponav js-repo-nav js-sidenav-container-pjax clearfix container-lg p-responsive d-none d-lg-block"
     itemscope
     itemtype="http://schema.org/BreadcrumbList"
    aria-label="Repository"
     data-pjax="#js-repo-pjax-container">

  <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
    <a class="js-selected-navigation-item selected reponav-item" itemprop="url" data-hotkey="g c" aria-current="page" data-selected-links="repo_source repo_downloads repo_commits repo_releases repo_tags repo_branches repo_packages /svmiller/svm-r-markdown-templates" href="/svmiller/svm-r-markdown-templates">
      <div class="d-inline"><svg class="octicon octicon-code" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M9.5 3L8 4.5 11.5 8 8 11.5 9.5 13 14 8 9.5 3zm-5 0L0 8l4.5 5L6 11.5 2.5 8 6 4.5 4.5 3z"/></svg></div>
      <span itemprop="name">Code</span>
      <meta itemprop="position" content="1">
</a>  </span>

    <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
      <a itemprop="url" data-hotkey="g i" class="js-selected-navigation-item reponav-item" data-selected-links="repo_issues repo_labels repo_milestones /svmiller/svm-r-markdown-templates/issues" href="/svmiller/svm-r-markdown-templates/issues">
        <div class="d-inline"><svg class="octicon octicon-issue-opened" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M7 2.3c3.14 0 5.7 2.56 5.7 5.7s-2.56 5.7-5.7 5.7A5.71 5.71 0 011.3 8c0-3.14 2.56-5.7 5.7-5.7zM7 1C3.14 1 0 4.14 0 8s3.14 7 7 7 7-3.14 7-7-3.14-7-7-7zm1 3H6v5h2V4zm0 6H6v2h2v-2z"/></svg></div>
        <span itemprop="name">Issues</span>
        <span class="Counter">11</span>
        <meta itemprop="position" content="2">
</a>    </span>

  <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
    <a data-hotkey="g p" data-skip-pjax="true" itemprop="url" class="js-selected-navigation-item reponav-item" data-selected-links="repo_pulls checks /svmiller/svm-r-markdown-templates/pulls" href="/svmiller/svm-r-markdown-templates/pulls">
      <div class="d-inline"><svg class="octicon octicon-git-pull-request" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M11 11.28V5c-.03-.78-.34-1.47-.94-2.06C9.46 2.35 8.78 2.03 8 2H7V0L4 3l3 3V4h1c.27.02.48.11.69.31.21.2.3.42.31.69v6.28A1.993 1.993 0 0010 15a1.993 1.993 0 001-3.72zm-1 2.92c-.66 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2zM4 3c0-1.11-.89-2-2-2a1.993 1.993 0 00-1 3.72v6.56A1.993 1.993 0 002 15a1.993 1.993 0 001-3.72V4.72c.59-.34 1-.98 1-1.72zm-.8 10c0 .66-.55 1.2-1.2 1.2-.65 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2zM2 4.2C1.34 4.2.8 3.65.8 3c0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2z"/></svg></div>
      <span itemprop="name">Pull requests</span>
      <span class="Counter">0</span>
      <meta itemprop="position" content="4">
</a>  </span>

    <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement" class="position-relative float-left">
      <a data-hotkey="g w" data-skip-pjax="true" class="js-selected-navigation-item reponav-item" data-selected-links="repo_actions /svmiller/svm-r-markdown-templates/actions" href="/svmiller/svm-r-markdown-templates/actions">
        <div class="d-inline"><svg class="octicon octicon-play" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M14 8A7 7 0 110 8a7 7 0 0114 0zm-8.223 3.482l4.599-3.066a.5.5 0 000-.832L5.777 4.518A.5.5 0 005 4.934v6.132a.5.5 0 00.777.416z"/></svg></div>
        Actions
</a>
    </span>

    <a data-hotkey="g b" class="js-selected-navigation-item reponav-item" data-selected-links="repo_projects new_repo_project repo_project /svmiller/svm-r-markdown-templates/projects" href="/svmiller/svm-r-markdown-templates/projects">
      <div class="d-inline"><svg class="octicon octicon-project" viewBox="0 0 15 16" version="1.1" width="15" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M10 12h3V2h-3v10zm-4-2h3V2H6v8zm-4 4h3V2H2v12zm-1 1h13V1H1v14zM14 0H1a1 1 0 00-1 1v14a1 1 0 001 1h13a1 1 0 001-1V1a1 1 0 00-1-1z"/></svg></div>
      Projects
      <span class="Counter">0</span>
</a>
    <a class="js-selected-navigation-item reponav-item" data-hotkey="g w" data-selected-links="repo_wiki /svmiller/svm-r-markdown-templates/wiki" href="/svmiller/svm-r-markdown-templates/wiki">
      <div class="d-inline"><svg class="octicon octicon-book" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M3 5h4v1H3V5zm0 3h4V7H3v1zm0 2h4V9H3v1zm11-5h-4v1h4V5zm0 2h-4v1h4V7zm0 2h-4v1h4V9zm2-6v9c0 .55-.45 1-1 1H9.5l-1 1-1-1H2c-.55 0-1-.45-1-1V3c0-.55.45-1 1-1h5.5l1 1 1-1H15c.55 0 1 .45 1 1zm-8 .5L7.5 3H2v9h6V3.5zm7-.5H9.5l-.5.5V12h6V3z"/></svg></div>
      Wiki
</a>
    <a data-skip-pjax="true" class="js-selected-navigation-item reponav-item" data-selected-links="security alerts policy token_scanning code_scanning /svmiller/svm-r-markdown-templates/security/advisories" href="/svmiller/svm-r-markdown-templates/security/advisories">
      <div class="d-inline"><svg class="octicon octicon-shield" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M0 2l7-2 7 2v6.02C14 12.69 8.69 16 7 16c-1.69 0-7-3.31-7-7.98V2zm1 .75L7 1l6 1.75v5.268C13 12.104 8.449 15 7 15c-1.449 0-6-2.896-6-6.982V2.75zm1 .75L7 2v12c-1.207 0-5-2.482-5-5.985V3.5z"/></svg></div>
      Security
</a>
    <a class="js-selected-navigation-item reponav-item" data-selected-links="repo_graphs repo_contributors dependency_graph pulse people /svmiller/svm-r-markdown-templates/pulse" href="/svmiller/svm-r-markdown-templates/pulse">
      <div class="d-inline"><svg class="octicon octicon-graph" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M16 14v1H0V0h1v14h15zM5 13H3V8h2v5zm4 0H7V3h2v10zm4 0h-2V6h2v7z"/></svg></div>
      Insights
</a>

</nav>

  <div class="reponav-wrapper reponav-small d-lg-none">
  <nav class="reponav js-reponav text-center no-wrap"
       itemscope
       itemtype="http://schema.org/BreadcrumbList">

    <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
      <a class="js-selected-navigation-item selected reponav-item" itemprop="url" aria-current="page" data-selected-links="repo_source repo_downloads repo_commits repo_releases repo_tags repo_branches repo_packages /svmiller/svm-r-markdown-templates" href="/svmiller/svm-r-markdown-templates">
        <span itemprop="name">Code</span>
        <meta itemprop="position" content="1">
</a>    </span>

      <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
        <a itemprop="url" class="js-selected-navigation-item reponav-item" data-selected-links="repo_issues repo_labels repo_milestones /svmiller/svm-r-markdown-templates/issues" href="/svmiller/svm-r-markdown-templates/issues">
          <span itemprop="name">Issues</span>
          <span class="Counter">11</span>
          <meta itemprop="position" content="2">
</a>      </span>

    <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
      <a itemprop="url" class="js-selected-navigation-item reponav-item" data-selected-links="repo_pulls checks /svmiller/svm-r-markdown-templates/pulls" href="/svmiller/svm-r-markdown-templates/pulls">
        <span itemprop="name">Pull requests</span>
        <span class="Counter">0</span>
        <meta itemprop="position" content="4">
</a>    </span>

      <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
        <a itemprop="url" class="js-selected-navigation-item reponav-item" data-selected-links="repo_projects new_repo_project repo_project /svmiller/svm-r-markdown-templates/projects" href="/svmiller/svm-r-markdown-templates/projects">
          <span itemprop="name">Projects</span>
          <span class="Counter">0</span>
          <meta itemprop="position" content="5">
</a>      </span>

      <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
        <a itemprop="url" class="js-selected-navigation-item reponav-item" data-selected-links="repo_actions /svmiller/svm-r-markdown-templates/actions" href="/svmiller/svm-r-markdown-templates/actions">
          <span itemprop="name">Actions</span>
          <meta itemprop="position" content="6">
</a>      </span>

      <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
        <a itemprop="url" class="js-selected-navigation-item reponav-item" data-selected-links="repo_wiki /svmiller/svm-r-markdown-templates/wiki" href="/svmiller/svm-r-markdown-templates/wiki">
          <span itemprop="name">Wiki</span>
          <meta itemprop="position" content="7">
</a>      </span>

      <a itemprop="url" class="js-selected-navigation-item reponav-item" data-selected-links="security alerts policy token_scanning code_scanning /svmiller/svm-r-markdown-templates/security/advisories" href="/svmiller/svm-r-markdown-templates/security/advisories">
        <span itemprop="name">Security</span>
        <meta itemprop="position" content="8">
</a>
      <a class="js-selected-navigation-item reponav-item" data-selected-links="pulse /svmiller/svm-r-markdown-templates/pulse" href="/svmiller/svm-r-markdown-templates/pulse">
        Pulse
</a>
      <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
        <a itemprop="url" class="js-selected-navigation-item reponav-item" data-selected-links="community /svmiller/svm-r-markdown-templates/community" href="/svmiller/svm-r-markdown-templates/community">
          Community
</a>      </span>

  </nav>
</div>

  </div>

  <include-fragment class="js-notification-shelf-include-fragment" data-base-src="https://github.com/notifications/beta/shelf"></include-fragment>

<div class="container-lg clearfix new-discussion-timeline  p-responsive">
  <div class="repository-content ">

    <a class="d-none js-permalink-shortcut" data-hotkey="y" href="/svmiller/svm-r-markdown-templates/blob/6b678ff10ecb592214fb21f8f6842ae36d9cb8b9/svm-latex-ms.tex">Permalink</a>

    <!-- blob contrib key: blob_contributors:v21:073bee0626b85aaaa1eff874bb75f850 -->

    <div class="d-flex flex-items-start flex-shrink-0 flex-column flex-md-row pb-3">
      <span class="d-flex flex-justify-between width-full width-md-auto">

<details class="details-reset details-overlay branch-select-menu " id="branch-select-menu">
  <summary class="btn btn-sm css-truncate"
           data-hotkey="w"
           title="Switch branches or tags">
    <i>Branch:</i>
    <span class="css-truncate-target" data-menu-button>master</span>
    <span class="dropdown-caret"></span>
  </summary>

  <details-menu class="SelectMenu SelectMenu--hasFilter" src="/svmiller/svm-r-markdown-templates/refs/master/svm-latex-ms.tex?source_action=show&amp;source_controller=blob" preload>
    <div class="SelectMenu-modal">
      <include-fragment class="SelectMenu-loading" aria-label="Menu is loading">
        <svg class="octicon octicon-octoface anim-pulse" height="32" viewBox="0 0 16 16" version="1.1" width="32" aria-hidden="true"><path fill-rule="evenodd" d="M14.7 5.34c.13-.32.55-1.59-.13-3.31 0 0-1.05-.33-3.44 1.3-1-.28-2.07-.32-3.13-.32s-2.13.04-3.13.32c-2.39-1.64-3.44-1.3-3.44-1.3-.68 1.72-.26 2.99-.13 3.31C.49 6.21 0 7.33 0 8.69 0 13.84 3.33 15 7.98 15S16 13.84 16 8.69c0-1.36-.49-2.48-1.3-3.35zM8 14.02c-3.3 0-5.98-.15-5.98-3.35 0-.76.38-1.48 1.02-2.07 1.07-.98 2.9-.46 4.96-.46 2.07 0 3.88-.52 4.96.46.65.59 1.02 1.3 1.02 2.07 0 3.19-2.68 3.35-5.98 3.35zM5.49 9.01c-.66 0-1.2.8-1.2 1.78s.54 1.79 1.2 1.79c.66 0 1.2-.8 1.2-1.79s-.54-1.78-1.2-1.78zm5.02 0c-.66 0-1.2.79-1.2 1.78s.54 1.79 1.2 1.79c.66 0 1.2-.8 1.2-1.79s-.53-1.78-1.2-1.78z"/></svg>
      </include-fragment>
    </div>
  </details-menu>
</details>

        <div class="BtnGroup flex-shrink-0 d-md-none">
          <a href="/svmiller/svm-r-markdown-templates/find/master"
                class="js-pjax-capture-input btn btn-sm BtnGroup-item"
                data-pjax
                data-hotkey="t">
            Find file
          </a>
          <clipboard-copy value="svm-latex-ms.tex" class="btn btn-sm BtnGroup-item">
            Copy path
          </clipboard-copy>
        </div>
      </span>
      <h2 id="blob-path" class="breadcrumb flex-auto min-width-0 text-normal flex-md-self-center ml-md-2 mr-md-3 my-2 my-md-0">
        <span class="js-repo-root text-bold"><span class="js-path-segment"><a data-pjax="true" href="/svmiller/svm-r-markdown-templates"><span>svm-r-markdown-templates</span></a></span></span><span class="separator">/</span><strong class="final-path">svm-latex-ms.tex</strong>
      </h2>

      <div class="BtnGroup flex-shrink-0 d-none d-md-inline-block">
        <a href="/svmiller/svm-r-markdown-templates/find/master"
              class="js-pjax-capture-input btn btn-sm BtnGroup-item"
              data-pjax
              data-hotkey="t">
          Find file
        </a>
        <clipboard-copy value="svm-latex-ms.tex" class="btn btn-sm BtnGroup-item">
          Copy path
        </clipboard-copy>
      </div>
    </div>

  <div class="Box Box--condensed d-flex flex-column flex-shrink-0">
      <div class="Box-body d-flex flex-justify-between bg-blue-light flex-column flex-md-row flex-items-start flex-md-items-center">
        <span class="pr-md-4 f6">
          <a rel="author" data-skip-pjax="true" data-hovercard-type="user" data-hovercard-url="/users/svmiller/hovercard" data-octo-click="hovercard-link-click" data-octo-dimensions="link_type:self" href="/svmiller"><img class="avatar" src="https://avatars1.githubusercontent.com/u/4589542?s=40&amp;v=4" width="20" height="20" alt="@svmiller" /></a>
          <a class="text-bold link-gray-dark lh-default v-align-middle" rel="author" data-hovercard-type="user" data-hovercard-url="/users/svmiller/hovercard" data-octo-click="hovercard-link-click" data-octo-dimensions="link_type:self" href="/svmiller">svmiller</a>
            <span class="lh-default v-align-middle">
              <a data-pjax="true" title="make a host of changes to svm-latex-ms.tex" class="link-gray" href="/svmiller/svm-r-markdown-templates/commit/d73f698d7355fab224e4ce2509a0f191c5929dcc">make a host of changes to svm-latex-ms.tex</a>
            </span>
        </span>
        <span class="d-inline-block flex-shrink-0 v-align-bottom f6 mt-2 mt-md-0">
          <a class="pr-2 text-mono link-gray" href="/svmiller/svm-r-markdown-templates/commit/d73f698d7355fab224e4ce2509a0f191c5929dcc" data-pjax>d73f698</a>
          <relative-time datetime="2019-06-18T17:27:10Z" class="no-wrap">Jun 18, 2019</relative-time>
        </span>
      </div>

    <div class="Box-body d-flex flex-items-center flex-auto f6 border-bottom-0 flex-wrap" >
      <details class="details-reset details-overlay details-overlay-dark lh-default text-gray-dark float-left mr-2" id="blob_contributors_box">
        <summary class="btn-link">
          <span><strong>1</strong> contributor</span>
        </summary>
        <details-dialog
          class="Box Box--overlay d-flex flex-column anim-fade-in fast"
          aria-label="Users who have contributed to this file"
          src="/svmiller/svm-r-markdown-templates/contributors-list/master/svm-latex-ms.tex" preload>
          <div class="Box-header">
            <button class="Box-btn-octicon btn-octicon float-right" type="button" aria-label="Close dialog" data-close-dialog>
              <svg class="octicon octicon-x" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48L7.48 8z"/></svg>
            </button>
            <h3 class="Box-title">
              Users who have contributed to this file
            </h3>
          </div>
          <include-fragment class="octocat-spinner my-3" aria-label="Loading..."></include-fragment>
        </details-dialog>
      </details>
    </div>
  </div>

    <div class="Box mt-3 position-relative">

<div class="Box-header py-2 d-flex flex-column flex-shrink-0 flex-md-row flex-md-items-center">
  <div class="text-mono f6 flex-auto pr-3 flex-order-2 flex-md-order-1 mt-2 mt-md-0">

      291 lines (211 sloc)
      <span class="file-info-divider"></span>
    6.52 KB
  </div>

  <div class="d-flex py-1 py-md-0 flex-auto flex-order-1 flex-md-order-2 flex-sm-grow-0 flex-justify-between">

    <div class="BtnGroup">
      <a id="raw-url" class="btn btn-sm BtnGroup-item" href="/svmiller/svm-r-markdown-templates/raw/master/svm-latex-ms.tex">Raw</a>
        <a class="btn btn-sm js-update-url-with-hash BtnGroup-item" data-hotkey="b" href="/svmiller/svm-r-markdown-templates/blame/master/svm-latex-ms.tex">Blame</a>
      <a rel="nofollow" class="btn btn-sm BtnGroup-item" href="/svmiller/svm-r-markdown-templates/commits/master/svm-latex-ms.tex">History</a>
    </div>

    <div>
          <a class="btn-octicon tooltipped tooltipped-nw js-remove-unless-platform"
             data-platforms="windows,mac"
             href="x-github-client://openRepo/https://github.com/svmiller/svm-r-markdown-templates?branch=master&amp;filepath=svm-latex-ms.tex"
             aria-label="Open this file in GitHub Desktop"
             data-ga-click="Repository, open with desktop">
              <svg class="octicon octicon-device-desktop" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M15 2H1c-.55 0-1 .45-1 1v9c0 .55.45 1 1 1h5.34c-.25.61-.86 1.39-2.34 2h8c-1.48-.61-2.09-1.39-2.34-2H15c.55 0 1-.45 1-1V3c0-.55-.45-1-1-1zm0 9H1V3h14v8z"/></svg>
          </a>

          <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="inline-form js-update-url-with-hash" action="/svmiller/svm-r-markdown-templates/edit/master/svm-latex-ms.tex" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="wDOyCiV7naOW/iXs6dOZCLmo31LqprFGUXMSgJQ9JDZcC8UtKspOMapl+N9MkLlNbEZ/MK+mJ8sLKnMINO1FZw==" />
            <button class="btn-octicon tooltipped tooltipped-nw" type="submit"
              aria-label="Fork this project and edit the file" data-hotkey="e" data-disable-with>
              <svg class="octicon octicon-pencil" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M0 12v3h3l8-8-3-3-8 8zm3 2H1v-2h1v1h1v1zm10.3-9.3L12 6 9 3l1.3-1.3a.996.996 0 011.41 0l1.59 1.59c.39.39.39 1.02 0 1.41z"/></svg>
            </button>
</form>
          <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="inline-form" action="/svmiller/svm-r-markdown-templates/delete/master/svm-latex-ms.tex" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="WNbJLLZt65KTlHH4Uefdliitm10/r4dQGLhYb/o64G5AYKa9z1ABvQWQPlkO4paA1g8Fug51GVkjX5B41HAHOQ==" />
            <button class="btn-octicon btn-octicon-danger tooltipped tooltipped-nw" type="submit"
              aria-label="Fork this project and delete the file" data-disable-with>
              <svg class="octicon octicon-trashcan" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M11 2H9c0-.55-.45-1-1-1H5c-.55 0-1 .45-1 1H2c-.55 0-1 .45-1 1v1c0 .55.45 1 1 1v9c0 .55.45 1 1 1h7c.55 0 1-.45 1-1V5c.55 0 1-.45 1-1V3c0-.55-.45-1-1-1zm-1 12H3V5h1v8h1V5h1v8h1V5h1v8h1V5h1v9zm1-10H2V3h9v1z"/></svg>
            </button>
</form>    </div>
  </div>
</div>

  <div itemprop="text" class="Box-body p-0 blob-wrapper data type-tex ">

<table class="highlight tab-size js-file-line-container" data-tab-size="8">
      <tr>
        <td id="L1" class="blob-num js-line-number" data-line-number="1"></td>
        <td id="LC1" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\documentclass</span>[<span class="pl-v">$if(fontsize)$$fontsize$,$endif$$if(lang)$$babel-lang$,$endif$$if(papersize)$$papersize$,$endif$$for(classoption)$$classoption$$sep$,$endfor$</span>]{<span class="pl-c1">article</span>}</td>
      </tr>
      <tr>
        <td id="L2" class="blob-num js-line-number" data-line-number="2"></td>
        <td id="LC2" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>[<span class="pl-v">left=1in,top=1in,right=1in,bottom=1in</span>]{<span class="pl-c1">geometry</span>}</td>
      </tr>
      <tr>
        <td id="L3" class="blob-num js-line-number" data-line-number="3"></td>
        <td id="LC3" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\newcommand</span>*{<span class="pl-c1">\authorfont</span>}{<span class="pl-c1">\fontfamily</span>{phv}<span class="pl-c1">\selectfont</span>}</td>
      </tr>
      <tr>
        <td id="L4" class="blob-num js-line-number" data-line-number="4"></td>
        <td id="LC4" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(fontfamily)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L5" class="blob-num js-line-number" data-line-number="5"></td>
        <td id="LC5" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>[<span class="pl-v">$fontfamilyoptions$</span>]{<span class="pl-c1"><span class="pl-s"><span class="pl-pds">$</span>fontfamily<span class="pl-pds">$</span></span></span>}</td>
      </tr>
      <tr>
        <td id="L6" class="blob-num js-line-number" data-line-number="6"></td>
        <td id="LC6" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L7" class="blob-num js-line-number" data-line-number="7"></td>
        <td id="LC7" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">lmodern</span>}</td>
      </tr>
      <tr>
        <td id="L8" class="blob-num js-line-number" data-line-number="8"></td>
        <td id="LC8" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L9" class="blob-num js-line-number" data-line-number="9"></td>
        <td id="LC9" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L10" class="blob-num js-line-number" data-line-number="10"></td>
        <td id="LC10" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L11" class="blob-num js-line-number" data-line-number="11"></td>
        <td id="LC11" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\usepackage</span>[<span class="pl-v">$if(fontenc)$$fontenc$$else$T1$endif$</span>]{<span class="pl-c1">fontenc</span>}</td>
      </tr>
      <tr>
        <td id="L12" class="blob-num js-line-number" data-line-number="12"></td>
        <td id="LC12" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\usepackage</span>[<span class="pl-v">utf8</span>]{<span class="pl-c1">inputenc</span>}</td>
      </tr>
      <tr>
        <td id="L13" class="blob-num js-line-number" data-line-number="13"></td>
        <td id="LC13" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L14" class="blob-num js-line-number" data-line-number="14"></td>
        <td id="LC14" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L15" class="blob-num js-line-number" data-line-number="15"></td>
        <td id="LC15" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L16" class="blob-num js-line-number" data-line-number="16"></td>
        <td id="LC16" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L17" class="blob-num js-line-number" data-line-number="17"></td>
        <td id="LC17" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">abstract</span>}</td>
      </tr>
      <tr>
        <td id="L18" class="blob-num js-line-number" data-line-number="18"></td>
        <td id="LC18" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\renewcommand</span>{<span class="pl-c1">\abstractname</span>}{}    <span class="pl-c"><span class="pl-c">%</span> clear the title</span></td>
      </tr>
      <tr>
        <td id="L19" class="blob-num js-line-number" data-line-number="19"></td>
        <td id="LC19" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\renewcommand</span>{<span class="pl-c1">\absnamepos</span>}{empty} <span class="pl-c"><span class="pl-c">%</span> originally center</span></td>
      </tr>
      <tr>
        <td id="L20" class="blob-num js-line-number" data-line-number="20"></td>
        <td id="LC20" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L21" class="blob-num js-line-number" data-line-number="21"></td>
        <td id="LC21" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\renewenvironment</span>{abstract}</td>
      </tr>
      <tr>
        <td id="L22" class="blob-num js-line-number" data-line-number="22"></td>
        <td id="LC22" class="blob-code blob-code-inner js-file-line"> {{<span class="pl-c"><span class="pl-c">%</span></span></td>
      </tr>
      <tr>
        <td id="L23" class="blob-num js-line-number" data-line-number="23"></td>
        <td id="LC23" class="blob-code blob-code-inner js-file-line">    <span class="pl-c1">\setlength</span>{<span class="pl-c1">\leftmargin</span>}{0mm}</td>
      </tr>
      <tr>
        <td id="L24" class="blob-num js-line-number" data-line-number="24"></td>
        <td id="LC24" class="blob-code blob-code-inner js-file-line">    <span class="pl-c1">\setlength</span>{<span class="pl-c1">\rightmargin</span>}{<span class="pl-c1">\leftmargin</span>}<span class="pl-c"><span class="pl-c">%</span></span></td>
      </tr>
      <tr>
        <td id="L25" class="blob-num js-line-number" data-line-number="25"></td>
        <td id="LC25" class="blob-code blob-code-inner js-file-line">  }<span class="pl-c"><span class="pl-c">%</span></span></td>
      </tr>
      <tr>
        <td id="L26" class="blob-num js-line-number" data-line-number="26"></td>
        <td id="LC26" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\relax</span>}</td>
      </tr>
      <tr>
        <td id="L27" class="blob-num js-line-number" data-line-number="27"></td>
        <td id="LC27" class="blob-code blob-code-inner js-file-line"> {<span class="pl-c1">\endlist</span>}</td>
      </tr>
      <tr>
        <td id="L28" class="blob-num js-line-number" data-line-number="28"></td>
        <td id="LC28" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L29" class="blob-num js-line-number" data-line-number="29"></td>
        <td id="LC29" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatletter</span></td>
      </tr>
      <tr>
        <td id="L30" class="blob-num js-line-number" data-line-number="30"></td>
        <td id="LC30" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\def\@maketitle</span>{<span class="pl-c"><span class="pl-c">%</span></span></td>
      </tr>
      <tr>
        <td id="L31" class="blob-num js-line-number" data-line-number="31"></td>
        <td id="LC31" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\newpage</span></td>
      </tr>
      <tr>
        <td id="L32" class="blob-num js-line-number" data-line-number="32"></td>
        <td id="LC32" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>  \null</span></td>
      </tr>
      <tr>
        <td id="L33" class="blob-num js-line-number" data-line-number="33"></td>
        <td id="LC33" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>  \vskip 2em%</span></td>
      </tr>
      <tr>
        <td id="L34" class="blob-num js-line-number" data-line-number="34"></td>
        <td id="LC34" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>  \begin{center}%</span></td>
      </tr>
      <tr>
        <td id="L35" class="blob-num js-line-number" data-line-number="35"></td>
        <td id="LC35" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\let</span> <span class="pl-c1">\footnote</span> <span class="pl-c1">\thanks</span></td>
      </tr>
      <tr>
        <td id="L36" class="blob-num js-line-number" data-line-number="36"></td>
        <td id="LC36" class="blob-code blob-code-inner js-file-line">    {<span class="pl-c1">\fontsize</span>{18}{20}<span class="pl-c1">\selectfont\raggedright</span>  <span class="pl-c1">\setlength</span>{<span class="pl-c1">\parindent</span>}{0pt} <span class="pl-c1">\@title</span> <span class="pl-c1">\par</span>}<span class="pl-c"><span class="pl-c">%</span></span></td>
      </tr>
      <tr>
        <td id="L37" class="blob-num js-line-number" data-line-number="37"></td>
        <td id="LC37" class="blob-code blob-code-inner js-file-line">}</td>
      </tr>
      <tr>
        <td id="L38" class="blob-num js-line-number" data-line-number="38"></td>
        <td id="LC38" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>\fi</span></td>
      </tr>
      <tr>
        <td id="L39" class="blob-num js-line-number" data-line-number="39"></td>
        <td id="LC39" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatother</span></td>
      </tr>
      <tr>
        <td id="L40" class="blob-num js-line-number" data-line-number="40"></td>
        <td id="LC40" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L41" class="blob-num js-line-number" data-line-number="41"></td>
        <td id="LC41" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L42" class="blob-num js-line-number" data-line-number="42"></td>
        <td id="LC42" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(appendix)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L43" class="blob-num js-line-number" data-line-number="43"></td>
        <td id="LC43" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\renewcommand</span>*<span class="pl-c1">\thetable</span>{<span class="pl-s"><span class="pl-pds">$</span>appendixletter<span class="pl-pds">$</span></span>.<span class="pl-c1">\arabic</span>{table}}</td>
      </tr>
      <tr>
        <td id="L44" class="blob-num js-line-number" data-line-number="44"></td>
        <td id="LC44" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\renewcommand</span>*<span class="pl-c1">\thefigure</span>{<span class="pl-s"><span class="pl-pds">$</span>appendixletter<span class="pl-pds">$</span></span>.<span class="pl-c1">\arabic</span>{figure}}</td>
      </tr>
      <tr>
        <td id="L45" class="blob-num js-line-number" data-line-number="45"></td>
        <td id="LC45" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L46" class="blob-num js-line-number" data-line-number="46"></td>
        <td id="LC46" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L47" class="blob-num js-line-number" data-line-number="47"></td>
        <td id="LC47" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L48" class="blob-num js-line-number" data-line-number="48"></td>
        <td id="LC48" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\setcounter</span>{secnumdepth}{0}</td>
      </tr>
      <tr>
        <td id="L49" class="blob-num js-line-number" data-line-number="49"></td>
        <td id="LC49" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L50" class="blob-num js-line-number" data-line-number="50"></td>
        <td id="LC50" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(listings)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L51" class="blob-num js-line-number" data-line-number="51"></td>
        <td id="LC51" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">listings</span>}</td>
      </tr>
      <tr>
        <td id="L52" class="blob-num js-line-number" data-line-number="52"></td>
        <td id="LC52" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L53" class="blob-num js-line-number" data-line-number="53"></td>
        <td id="LC53" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(lhs)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L54" class="blob-num js-line-number" data-line-number="54"></td>
        <td id="LC54" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\lstnewenvironment</span>{code}{<span class="pl-c1">\lstset</span>{language=r,basicstyle=<span class="pl-c1">\small\ttfamily</span>}}{}</td>
      </tr>
      <tr>
        <td id="L55" class="blob-num js-line-number" data-line-number="55"></td>
        <td id="LC55" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L56" class="blob-num js-line-number" data-line-number="56"></td>
        <td id="LC56" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(highlighting-macros)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L57" class="blob-num js-line-number" data-line-number="57"></td>
        <td id="LC57" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>highlighting-macros<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L58" class="blob-num js-line-number" data-line-number="58"></td>
        <td id="LC58" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L59" class="blob-num js-line-number" data-line-number="59"></td>
        <td id="LC59" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(verbatim-in-note)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L60" class="blob-num js-line-number" data-line-number="60"></td>
        <td id="LC60" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">fancyvrb</span>}</td>
      </tr>
      <tr>
        <td id="L61" class="blob-num js-line-number" data-line-number="61"></td>
        <td id="LC61" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L62" class="blob-num js-line-number" data-line-number="62"></td>
        <td id="LC62" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(tables)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L63" class="blob-num js-line-number" data-line-number="63"></td>
        <td id="LC63" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">longtable,booktabs</span>}</td>
      </tr>
      <tr>
        <td id="L64" class="blob-num js-line-number" data-line-number="64"></td>
        <td id="LC64" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L65" class="blob-num js-line-number" data-line-number="65"></td>
        <td id="LC65" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L66" class="blob-num js-line-number" data-line-number="66"></td>
        <td id="LC66" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(graphics)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L67" class="blob-num js-line-number" data-line-number="67"></td>
        <td id="LC67" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">graphicx,grffile</span>}</td>
      </tr>
      <tr>
        <td id="L68" class="blob-num js-line-number" data-line-number="68"></td>
        <td id="LC68" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatletter</span></td>
      </tr>
      <tr>
        <td id="L69" class="blob-num js-line-number" data-line-number="69"></td>
        <td id="LC69" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\def\maxwidth</span>{<span class="pl-k">\ifdim</span><span class="pl-c1">\Gin@nat@width</span>&gt;<span class="pl-c1">\linewidth\linewidth</span><span class="pl-k">\else</span><span class="pl-c1">\Gin@nat@width</span><span class="pl-k">\fi</span>}</td>
      </tr>
      <tr>
        <td id="L70" class="blob-num js-line-number" data-line-number="70"></td>
        <td id="LC70" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\def\maxheight</span>{<span class="pl-k">\ifdim</span><span class="pl-c1">\Gin@nat@height</span>&gt;<span class="pl-c1">\textheight\textheight</span><span class="pl-k">\else</span><span class="pl-c1">\Gin@nat@height</span><span class="pl-k">\fi</span>}</td>
      </tr>
      <tr>
        <td id="L71" class="blob-num js-line-number" data-line-number="71"></td>
        <td id="LC71" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatother</span></td>
      </tr>
      <tr>
        <td id="L72" class="blob-num js-line-number" data-line-number="72"></td>
        <td id="LC72" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> Scale images if necessary, so that they will not overflow the page</span></td>
      </tr>
      <tr>
        <td id="L73" class="blob-num js-line-number" data-line-number="73"></td>
        <td id="LC73" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> margins by default, and it is still possible to overwrite the defaults</span></td>
      </tr>
      <tr>
        <td id="L74" class="blob-num js-line-number" data-line-number="74"></td>
        <td id="LC74" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> using explicit options in \includegraphics[width, height, ...]{}</span></td>
      </tr>
      <tr>
        <td id="L75" class="blob-num js-line-number" data-line-number="75"></td>
        <td id="LC75" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\setkeys</span>{Gin}{width=<span class="pl-c1">\maxwidth</span>,height=<span class="pl-c1">\maxheight</span>,keepaspectratio}</td>
      </tr>
      <tr>
        <td id="L76" class="blob-num js-line-number" data-line-number="76"></td>
        <td id="LC76" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L77" class="blob-num js-line-number" data-line-number="77"></td>
        <td id="LC77" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L78" class="blob-num js-line-number" data-line-number="78"></td>
        <td id="LC78" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L79" class="blob-num js-line-number" data-line-number="79"></td>
        <td id="LC79" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(title)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L80" class="blob-num js-line-number" data-line-number="80"></td>
        <td id="LC80" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\title</span>{<span class="pl-s"><span class="pl-pds">$</span>title<span class="pl-pds">$</span><span class="pl-pds">$</span>if(subtitle)<span class="pl-pds">$</span></span>: <span class="pl-s"><span class="pl-pds">$</span>subtitle<span class="pl-pds">$</span><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span> <span class="pl-s"><span class="pl-pds">$</span>if(anonymous)<span class="pl-pds">$</span><span class="pl-pds">$</span>else<span class="pl-pds">$</span><span class="pl-pds">$</span>if(thanks)<span class="pl-pds">$</span></span><span class="pl-c1">\thanks</span>{<span class="pl-s"><span class="pl-pds">$</span>thanks<span class="pl-pds">$</span></span>} <span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span> }</td>
      </tr>
      <tr>
        <td id="L81" class="blob-num js-line-number" data-line-number="81"></td>
        <td id="LC81" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span> </td>
      </tr>
      <tr>
        <td id="L82" class="blob-num js-line-number" data-line-number="82"></td>
        <td id="LC82" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L83" class="blob-num js-line-number" data-line-number="83"></td>
        <td id="LC83" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L84" class="blob-num js-line-number" data-line-number="84"></td>
        <td id="LC84" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L85" class="blob-num js-line-number" data-line-number="85"></td>
        <td id="LC85" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(anonymous)<span class="pl-pds">$</span><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span><span class="pl-c1">\author</span>{<span class="pl-s"><span class="pl-pds">$</span>for(author)<span class="pl-pds">$</span></span><span class="pl-c1">\Large</span> <span class="pl-s"><span class="pl-pds">$</span>author.name<span class="pl-pds">$</span></span><span class="pl-c1">\vspace</span>{0.05in} <span class="pl-c1">\newline\normalsize</span><span class="pl-c1">\emph</span>{<span class="pl-mi"><span class="pl-s"><span class="pl-pds">$</span>author.affiliation<span class="pl-pds">$</span></span></span>}  <span class="pl-s"><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span> <span class="pl-c1">\and</span> <span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span>}<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L86" class="blob-num js-line-number" data-line-number="86"></td>
        <td id="LC86" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L87" class="blob-num js-line-number" data-line-number="87"></td>
        <td id="LC87" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L88" class="blob-num js-line-number" data-line-number="88"></td>
        <td id="LC88" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\date</span>{}</td>
      </tr>
      <tr>
        <td id="L89" class="blob-num js-line-number" data-line-number="89"></td>
        <td id="LC89" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L90" class="blob-num js-line-number" data-line-number="90"></td>
        <td id="LC90" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">titlesec</span>}</td>
      </tr>
      <tr>
        <td id="L91" class="blob-num js-line-number" data-line-number="91"></td>
        <td id="LC91" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L92" class="blob-num js-line-number" data-line-number="92"></td>
        <td id="LC92" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\titleformat</span>*{<span class="pl-c1">\section</span>}{<span class="pl-c1">\normalsize\bfseries</span>}</td>
      </tr>
      <tr>
        <td id="L93" class="blob-num js-line-number" data-line-number="93"></td>
        <td id="LC93" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\titleformat</span>*{<span class="pl-c1">\subsection</span>}{<span class="pl-c1">\normalsize\itshape</span>}</td>
      </tr>
      <tr>
        <td id="L94" class="blob-num js-line-number" data-line-number="94"></td>
        <td id="LC94" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\titleformat</span>*{<span class="pl-c1">\subsubsection</span>}{<span class="pl-c1">\normalsize\itshape</span>}</td>
      </tr>
      <tr>
        <td id="L95" class="blob-num js-line-number" data-line-number="95"></td>
        <td id="LC95" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\titleformat</span>*{<span class="pl-c1">\paragraph</span>}{<span class="pl-c1">\normalsize\itshape</span>}</td>
      </tr>
      <tr>
        <td id="L96" class="blob-num js-line-number" data-line-number="96"></td>
        <td id="LC96" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\titleformat</span>*{<span class="pl-c1">\subparagraph</span>}{<span class="pl-c1">\normalsize\itshape</span>}</td>
      </tr>
      <tr>
        <td id="L97" class="blob-num js-line-number" data-line-number="97"></td>
        <td id="LC97" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L98" class="blob-num js-line-number" data-line-number="98"></td>
        <td id="LC98" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L99" class="blob-num js-line-number" data-line-number="99"></td>
        <td id="LC99" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(natbib)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L100" class="blob-num js-line-number" data-line-number="100"></td>
        <td id="LC100" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">natbib</span>}</td>
      </tr>
      <tr>
        <td id="L101" class="blob-num js-line-number" data-line-number="101"></td>
        <td id="LC101" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\bibliographystyle</span>{<span class="pl-s"><span class="pl-pds">$</span>if(biblio-style)<span class="pl-pds">$</span><span class="pl-pds">$</span>biblio-style<span class="pl-pds">$</span><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span>plainnat<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L102" class="blob-num js-line-number" data-line-number="102"></td>
        <td id="LC102" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>[<span class="pl-v">strings</span>]{<span class="pl-c1">underscore</span>} <span class="pl-c"><span class="pl-c">%</span> protect underscores in most circumstances</span></td>
      </tr>
      <tr>
        <td id="L103" class="blob-num js-line-number" data-line-number="103"></td>
        <td id="LC103" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L104" class="blob-num js-line-number" data-line-number="104"></td>
        <td id="LC104" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L105" class="blob-num js-line-number" data-line-number="105"></td>
        <td id="LC105" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(biblatex)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L106" class="blob-num js-line-number" data-line-number="106"></td>
        <td id="LC106" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\usepackage</span><span class="pl-s"><span class="pl-pds">$</span>if(biblio-style)<span class="pl-pds">$</span></span>[style=<span class="pl-s"><span class="pl-pds">$</span>biblio-style<span class="pl-pds">$</span></span>]<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span>{biblatex}</td>
      </tr>
      <tr>
        <td id="L107" class="blob-num js-line-number" data-line-number="107"></td>
        <td id="LC107" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(biblatexoptions)<span class="pl-pds">$</span></span><span class="pl-c1">\ExecuteBibliographyOptions</span>{<span class="pl-s"><span class="pl-pds">$</span>for(biblatexoptions)<span class="pl-pds">$</span><span class="pl-pds">$</span>biblatexoptions<span class="pl-pds">$</span><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span>,<span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span>}<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L108" class="blob-num js-line-number" data-line-number="108"></td>
        <td id="LC108" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>for(bibliography)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L109" class="blob-num js-line-number" data-line-number="109"></td>
        <td id="LC109" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\addbibresource</span>{<span class="pl-s"><span class="pl-pds">$</span>bibliography<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L110" class="blob-num js-line-number" data-line-number="110"></td>
        <td id="LC110" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L111" class="blob-num js-line-number" data-line-number="111"></td>
        <td id="LC111" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L112" class="blob-num js-line-number" data-line-number="112"></td>
        <td id="LC112" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(listings)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L113" class="blob-num js-line-number" data-line-number="113"></td>
        <td id="LC113" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">listings</span>}</td>
      </tr>
      <tr>
        <td id="L114" class="blob-num js-line-number" data-line-number="114"></td>
        <td id="LC114" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L115" class="blob-num js-line-number" data-line-number="115"></td>
        <td id="LC115" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L116" class="blob-num js-line-number" data-line-number="116"></td>
        <td id="LC116" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L117" class="blob-num js-line-number" data-line-number="117"></td>
        <td id="LC117" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\newtheorem</span>{hypothesis}{Hypothesis}</td>
      </tr>
      <tr>
        <td id="L118" class="blob-num js-line-number" data-line-number="118"></td>
        <td id="LC118" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">setspace</span>}</td>
      </tr>
      <tr>
        <td id="L119" class="blob-num js-line-number" data-line-number="119"></td>
        <td id="LC119" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L120" class="blob-num js-line-number" data-line-number="120"></td>
        <td id="LC120" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L121" class="blob-num js-line-number" data-line-number="121"></td>
        <td id="LC121" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> set default figure placement to htbp</span></td>
      </tr>
      <tr>
        <td id="L122" class="blob-num js-line-number" data-line-number="122"></td>
        <td id="LC122" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatletter</span></td>
      </tr>
      <tr>
        <td id="L123" class="blob-num js-line-number" data-line-number="123"></td>
        <td id="LC123" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\def\fps@figure</span>{htbp}</td>
      </tr>
      <tr>
        <td id="L124" class="blob-num js-line-number" data-line-number="124"></td>
        <td id="LC124" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatother</span></td>
      </tr>
      <tr>
        <td id="L125" class="blob-num js-line-number" data-line-number="125"></td>
        <td id="LC125" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L126" class="blob-num js-line-number" data-line-number="126"></td>
        <td id="LC126" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>for(header-includes)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L127" class="blob-num js-line-number" data-line-number="127"></td>
        <td id="LC127" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>header-includes<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L128" class="blob-num js-line-number" data-line-number="128"></td>
        <td id="LC128" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L129" class="blob-num js-line-number" data-line-number="129"></td>
        <td id="LC129" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L130" class="blob-num js-line-number" data-line-number="130"></td>
        <td id="LC130" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> move the hyperref stuff down here, after header-includes, to allow for - \usepackage{hyperref}</span></td>
      </tr>
      <tr>
        <td id="L131" class="blob-num js-line-number" data-line-number="131"></td>
        <td id="LC131" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L132" class="blob-num js-line-number" data-line-number="132"></td>
        <td id="LC132" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatletter</span></td>
      </tr>
      <tr>
        <td id="L133" class="blob-num js-line-number" data-line-number="133"></td>
        <td id="LC133" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\@ifpackageloaded</span>{hyperref}{}{<span class="pl-c"><span class="pl-c">%</span></span></td>
      </tr>
      <tr>
        <td id="L134" class="blob-num js-line-number" data-line-number="134"></td>
        <td id="LC134" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\ifxetex</span></td>
      </tr>
      <tr>
        <td id="L135" class="blob-num js-line-number" data-line-number="135"></td>
        <td id="LC135" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\PassOptionsToPackage</span>{hyphens}{url}<span class="pl-c1">\usepackage</span>[setpagesize=false, <span class="pl-c"><span class="pl-c">%</span> page size defined by xetex</span></td>
      </tr>
      <tr>
        <td id="L136" class="blob-num js-line-number" data-line-number="136"></td>
        <td id="LC136" class="blob-code blob-code-inner js-file-line">              unicode=false, <span class="pl-c"><span class="pl-c">%</span> unicode breaks when used with xetex</span></td>
      </tr>
      <tr>
        <td id="L137" class="blob-num js-line-number" data-line-number="137"></td>
        <td id="LC137" class="blob-code blob-code-inner js-file-line">              xetex]{hyperref}</td>
      </tr>
      <tr>
        <td id="L138" class="blob-num js-line-number" data-line-number="138"></td>
        <td id="LC138" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\else</span></td>
      </tr>
      <tr>
        <td id="L139" class="blob-num js-line-number" data-line-number="139"></td>
        <td id="LC139" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\PassOptionsToPackage</span>{hyphens}{url}<span class="pl-k">\usepackage</span>[<span class="pl-v">draft,unicode=true</span>]{<span class="pl-c1">hyperref</span>}</td>
      </tr>
      <tr>
        <td id="L140" class="blob-num js-line-number" data-line-number="140"></td>
        <td id="LC140" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\fi</span></td>
      </tr>
      <tr>
        <td id="L141" class="blob-num js-line-number" data-line-number="141"></td>
        <td id="LC141" class="blob-code blob-code-inner js-file-line">}</td>
      </tr>
      <tr>
        <td id="L142" class="blob-num js-line-number" data-line-number="142"></td>
        <td id="LC142" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L143" class="blob-num js-line-number" data-line-number="143"></td>
        <td id="LC143" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\@ifpackageloaded</span>{color}{</td>
      </tr>
      <tr>
        <td id="L144" class="blob-num js-line-number" data-line-number="144"></td>
        <td id="LC144" class="blob-code blob-code-inner js-file-line">    <span class="pl-c1">\PassOptionsToPackage</span>{usenames,dvipsnames}{color}</td>
      </tr>
      <tr>
        <td id="L145" class="blob-num js-line-number" data-line-number="145"></td>
        <td id="LC145" class="blob-code blob-code-inner js-file-line">}{<span class="pl-c"><span class="pl-c">%</span></span></td>
      </tr>
      <tr>
        <td id="L146" class="blob-num js-line-number" data-line-number="146"></td>
        <td id="LC146" class="blob-code blob-code-inner js-file-line">    <span class="pl-k">\usepackage</span>[<span class="pl-v">usenames,dvipsnames</span>]{<span class="pl-c1">color</span>}</td>
      </tr>
      <tr>
        <td id="L147" class="blob-num js-line-number" data-line-number="147"></td>
        <td id="LC147" class="blob-code blob-code-inner js-file-line">}</td>
      </tr>
      <tr>
        <td id="L148" class="blob-num js-line-number" data-line-number="148"></td>
        <td id="LC148" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatother</span></td>
      </tr>
      <tr>
        <td id="L149" class="blob-num js-line-number" data-line-number="149"></td>
        <td id="LC149" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\hypersetup</span>{breaklinks=true,</td>
      </tr>
      <tr>
        <td id="L150" class="blob-num js-line-number" data-line-number="150"></td>
        <td id="LC150" class="blob-code blob-code-inner js-file-line">            bookmarks=true,</td>
      </tr>
      <tr>
        <td id="L151" class="blob-num js-line-number" data-line-number="151"></td>
        <td id="LC151" class="blob-code blob-code-inner js-file-line">            pdfauthor={<span class="pl-s"><span class="pl-pds">$</span>if(anonymous)<span class="pl-pds">$</span><span class="pl-pds">$</span>else<span class="pl-pds">$</span><span class="pl-pds">$</span>for(author)<span class="pl-pds">$</span><span class="pl-pds">$</span>author.name<span class="pl-pds">$</span></span> (<span class="pl-s"><span class="pl-pds">$</span>author.affiliation<span class="pl-pds">$</span></span>)<span class="pl-s"><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span> and <span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span>},</td>
      </tr>
      <tr>
        <td id="L152" class="blob-num js-line-number" data-line-number="152"></td>
        <td id="LC152" class="blob-code blob-code-inner js-file-line">             pdfkeywords = {<span class="pl-s"><span class="pl-pds">$</span>if(keywords)<span class="pl-pds">$</span><span class="pl-pds">$</span>keywords<span class="pl-pds">$</span><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span>},  </td>
      </tr>
      <tr>
        <td id="L153" class="blob-num js-line-number" data-line-number="153"></td>
        <td id="LC153" class="blob-code blob-code-inner js-file-line">            pdftitle={<span class="pl-s"><span class="pl-pds">$</span>title<span class="pl-pds">$</span><span class="pl-pds">$</span>if(subtitle)<span class="pl-pds">$</span></span>: <span class="pl-s"><span class="pl-pds">$</span>subtitle<span class="pl-pds">$</span><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span>},</td>
      </tr>
      <tr>
        <td id="L154" class="blob-num js-line-number" data-line-number="154"></td>
        <td id="LC154" class="blob-code blob-code-inner js-file-line">            colorlinks=true,</td>
      </tr>
      <tr>
        <td id="L155" class="blob-num js-line-number" data-line-number="155"></td>
        <td id="LC155" class="blob-code blob-code-inner js-file-line">            citecolor=<span class="pl-s"><span class="pl-pds">$</span>if(citecolor)<span class="pl-pds">$</span><span class="pl-pds">$</span>citecolor<span class="pl-pds">$</span><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span>blue<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span>,</td>
      </tr>
      <tr>
        <td id="L156" class="blob-num js-line-number" data-line-number="156"></td>
        <td id="LC156" class="blob-code blob-code-inner js-file-line">            urlcolor=<span class="pl-s"><span class="pl-pds">$</span>if(urlcolor)<span class="pl-pds">$</span><span class="pl-pds">$</span>urlcolor<span class="pl-pds">$</span><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span>blue<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span>,</td>
      </tr>
      <tr>
        <td id="L157" class="blob-num js-line-number" data-line-number="157"></td>
        <td id="LC157" class="blob-code blob-code-inner js-file-line">            linkcolor=<span class="pl-s"><span class="pl-pds">$</span>if(linkcolor)<span class="pl-pds">$</span><span class="pl-pds">$</span>linkcolor<span class="pl-pds">$</span><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span>magenta<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span>,</td>
      </tr>
      <tr>
        <td id="L158" class="blob-num js-line-number" data-line-number="158"></td>
        <td id="LC158" class="blob-code blob-code-inner js-file-line">            pdfborder={0 0 0}}</td>
      </tr>
      <tr>
        <td id="L159" class="blob-num js-line-number" data-line-number="159"></td>
        <td id="LC159" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\urlstyle</span>{same}  <span class="pl-c"><span class="pl-c">%</span> don&#39;t use monospace font for urls</span></td>
      </tr>
      <tr>
        <td id="L160" class="blob-num js-line-number" data-line-number="160"></td>
        <td id="LC160" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L161" class="blob-num js-line-number" data-line-number="161"></td>
        <td id="LC161" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> Add an option for endnotes. -----</span></td>
      </tr>
      <tr>
        <td id="L162" class="blob-num js-line-number" data-line-number="162"></td>
        <td id="LC162" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L163" class="blob-num js-line-number" data-line-number="163"></td>
        <td id="LC163" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(endnotes)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L164" class="blob-num js-line-number" data-line-number="164"></td>
        <td id="LC164" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">endnotes</span>}</td>
      </tr>
      <tr>
        <td id="L165" class="blob-num js-line-number" data-line-number="165"></td>
        <td id="LC165" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\renewcommand</span>{<span class="pl-c1">\enotesize</span>}{<span class="pl-c1">\normalsize</span>}</td>
      </tr>
      <tr>
        <td id="L166" class="blob-num js-line-number" data-line-number="166"></td>
        <td id="LC166" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\let\footnote</span>=<span class="pl-c1">\endnote</span></td>
      </tr>
      <tr>
        <td id="L167" class="blob-num js-line-number" data-line-number="167"></td>
        <td id="LC167" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L168" class="blob-num js-line-number" data-line-number="168"></td>
        <td id="LC168" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L169" class="blob-num js-line-number" data-line-number="169"></td>
        <td id="LC169" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> add tightlist ----------</span></td>
      </tr>
      <tr>
        <td id="L170" class="blob-num js-line-number" data-line-number="170"></td>
        <td id="LC170" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\providecommand</span>{<span class="pl-c1">\tightlist</span>}{<span class="pl-c"><span class="pl-c">%</span></span></td>
      </tr>
      <tr>
        <td id="L171" class="blob-num js-line-number" data-line-number="171"></td>
        <td id="LC171" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\setlength</span>{<span class="pl-c1">\itemsep</span>}{0pt}<span class="pl-c1">\setlength</span>{<span class="pl-c1">\parskip</span>}{0pt}}</td>
      </tr>
      <tr>
        <td id="L172" class="blob-num js-line-number" data-line-number="172"></td>
        <td id="LC172" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L173" class="blob-num js-line-number" data-line-number="173"></td>
        <td id="LC173" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> add some other packages ----------</span></td>
      </tr>
      <tr>
        <td id="L174" class="blob-num js-line-number" data-line-number="174"></td>
        <td id="LC174" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L175" class="blob-num js-line-number" data-line-number="175"></td>
        <td id="LC175" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> \usepackage{multicol}</span></td>
      </tr>
      <tr>
        <td id="L176" class="blob-num js-line-number" data-line-number="176"></td>
        <td id="LC176" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> This should regulate where figures float</span></td>
      </tr>
      <tr>
        <td id="L177" class="blob-num js-line-number" data-line-number="177"></td>
        <td id="LC177" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> See: https://tex.stackexchange.com/questions/2275/keeping-tables-figures-close-to-where-they-are-mentioned</span></td>
      </tr>
      <tr>
        <td id="L178" class="blob-num js-line-number" data-line-number="178"></td>
        <td id="LC178" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>[<span class="pl-v">section</span>]{<span class="pl-c1">placeins</span>}</td>
      </tr>
      <tr>
        <td id="L179" class="blob-num js-line-number" data-line-number="179"></td>
        <td id="LC179" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L180" class="blob-num js-line-number" data-line-number="180"></td>
        <td id="LC180" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L181" class="blob-num js-line-number" data-line-number="181"></td>
        <td id="LC181" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\begin</span>{<span class="pl-smi">document</span>}</td>
      </tr>
      <tr>
        <td id="L182" class="blob-num js-line-number" data-line-number="182"></td>
        <td id="LC182" class="blob-code blob-code-inner js-file-line">	</td>
      </tr>
      <tr>
        <td id="L183" class="blob-num js-line-number" data-line-number="183"></td>
        <td id="LC183" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> \pagenumbering{arabic}% resets `page` counter to 1 </span></td>
      </tr>
      <tr>
        <td id="L184" class="blob-num js-line-number" data-line-number="184"></td>
        <td id="LC184" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>$if(appendix)$</span></td>
      </tr>
      <tr>
        <td id="L185" class="blob-num js-line-number" data-line-number="185"></td>
        <td id="LC185" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>\renewcommand*{\thepage}{$appendixletter$--\arabic{page}}</span></td>
      </tr>
      <tr>
        <td id="L186" class="blob-num js-line-number" data-line-number="186"></td>
        <td id="LC186" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>$endif$    </span></td>
      </tr>
      <tr>
        <td id="L187" class="blob-num js-line-number" data-line-number="187"></td>
        <td id="LC187" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L188" class="blob-num js-line-number" data-line-number="188"></td>
        <td id="LC188" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(removetitleabstract)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L189" class="blob-num js-line-number" data-line-number="189"></td>
        <td id="LC189" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L190" class="blob-num js-line-number" data-line-number="190"></td>
        <td id="LC190" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(title)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L191" class="blob-num js-line-number" data-line-number="191"></td>
        <td id="LC191" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> \maketitle</span></td>
      </tr>
      <tr>
        <td id="L192" class="blob-num js-line-number" data-line-number="192"></td>
        <td id="LC192" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L193" class="blob-num js-line-number" data-line-number="193"></td>
        <td id="LC193" class="blob-code blob-code-inner js-file-line">{<span class="pl-c"><span class="pl-c">%</span> \usefont{T1}{pnc}{m}{n}</span></td>
      </tr>
      <tr>
        <td id="L194" class="blob-num js-line-number" data-line-number="194"></td>
        <td id="LC194" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\setlength</span>{<span class="pl-c1">\parindent</span>}{0pt}</td>
      </tr>
      <tr>
        <td id="L195" class="blob-num js-line-number" data-line-number="195"></td>
        <td id="LC195" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\thispagestyle</span>{plain}</td>
      </tr>
      <tr>
        <td id="L196" class="blob-num js-line-number" data-line-number="196"></td>
        <td id="LC196" class="blob-code blob-code-inner js-file-line">{<span class="pl-c1">\fontsize</span>{18}{20}<span class="pl-c1">\selectfont\raggedright</span> </td>
      </tr>
      <tr>
        <td id="L197" class="blob-num js-line-number" data-line-number="197"></td>
        <td id="LC197" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\maketitle</span>  <span class="pl-c"><span class="pl-c">%</span> title \par  </span></td>
      </tr>
      <tr>
        <td id="L198" class="blob-num js-line-number" data-line-number="198"></td>
        <td id="LC198" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L199" class="blob-num js-line-number" data-line-number="199"></td>
        <td id="LC199" class="blob-code blob-code-inner js-file-line">}</td>
      </tr>
      <tr>
        <td id="L200" class="blob-num js-line-number" data-line-number="200"></td>
        <td id="LC200" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L201" class="blob-num js-line-number" data-line-number="201"></td>
        <td id="LC201" class="blob-code blob-code-inner js-file-line">{</td>
      </tr>
      <tr>
        <td id="L202" class="blob-num js-line-number" data-line-number="202"></td>
        <td id="LC202" class="blob-code blob-code-inner js-file-line">   <span class="pl-c1">\vskip</span> 13.5pt<span class="pl-c1">\relax</span> <span class="pl-c1">\normalsize\fontsize</span>{11}{12} </td>
      </tr>
      <tr>
        <td id="L203" class="blob-num js-line-number" data-line-number="203"></td>
        <td id="LC203" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(anonymous)<span class="pl-pds">$</span></span><span class="pl-c1">\hfill</span> <span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span><span class="pl-pds">$</span>for(author)<span class="pl-pds">$</span></span><span class="pl-c1">\textbf</span>{<span class="pl-mb"><span class="pl-c1">\authorfont</span> <span class="pl-s"><span class="pl-pds">$</span>author.name<span class="pl-pds">$</span></span></span>} <span class="pl-c1">\hskip</span> 15pt <span class="pl-c1">\emph</span>{<span class="pl-mi"><span class="pl-c1">\small</span> <span class="pl-s"><span class="pl-pds">$</span>author.affiliation<span class="pl-pds">$</span></span></span>}  <span class="pl-s"><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span> <span class="pl-c1">\par</span> <span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span> <span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L204" class="blob-num js-line-number" data-line-number="204"></td>
        <td id="LC204" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L205" class="blob-num js-line-number" data-line-number="205"></td>
        <td id="LC205" class="blob-code blob-code-inner js-file-line">}</td>
      </tr>
      <tr>
        <td id="L206" class="blob-num js-line-number" data-line-number="206"></td>
        <td id="LC206" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L207" class="blob-num js-line-number" data-line-number="207"></td>
        <td id="LC207" class="blob-code blob-code-inner js-file-line">}</td>
      </tr>
      <tr>
        <td id="L208" class="blob-num js-line-number" data-line-number="208"></td>
        <td id="LC208" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L209" class="blob-num js-line-number" data-line-number="209"></td>
        <td id="LC209" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L210" class="blob-num js-line-number" data-line-number="210"></td>
        <td id="LC210" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L211" class="blob-num js-line-number" data-line-number="211"></td>
        <td id="LC211" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L212" class="blob-num js-line-number" data-line-number="212"></td>
        <td id="LC212" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L213" class="blob-num js-line-number" data-line-number="213"></td>
        <td id="LC213" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L214" class="blob-num js-line-number" data-line-number="214"></td>
        <td id="LC214" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(abstract)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L215" class="blob-num js-line-number" data-line-number="215"></td>
        <td id="LC215" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L216" class="blob-num js-line-number" data-line-number="216"></td>
        <td id="LC216" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L217" class="blob-num js-line-number" data-line-number="217"></td>
        <td id="LC217" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L218" class="blob-num js-line-number" data-line-number="218"></td>
        <td id="LC218" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\begin</span>{<span class="pl-smi">abstract</span>}</td>
      </tr>
      <tr>
        <td id="L219" class="blob-num js-line-number" data-line-number="219"></td>
        <td id="LC219" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L220" class="blob-num js-line-number" data-line-number="220"></td>
        <td id="LC220" class="blob-code blob-code-inner js-file-line">    <span class="pl-c1">\hbox</span>{<span class="pl-c1">\vrule</span> height .2pt width 39.14pc}</td>
      </tr>
      <tr>
        <td id="L221" class="blob-num js-line-number" data-line-number="221"></td>
        <td id="LC221" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L222" class="blob-num js-line-number" data-line-number="222"></td>
        <td id="LC222" class="blob-code blob-code-inner js-file-line">    <span class="pl-c1">\vskip</span> 8.5pt <span class="pl-c"><span class="pl-c">%</span> \small </span></td>
      </tr>
      <tr>
        <td id="L223" class="blob-num js-line-number" data-line-number="223"></td>
        <td id="LC223" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L224" class="blob-num js-line-number" data-line-number="224"></td>
        <td id="LC224" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\noindent</span> <span class="pl-s"><span class="pl-pds">$</span>abstract<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L225" class="blob-num js-line-number" data-line-number="225"></td>
        <td id="LC225" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L226" class="blob-num js-line-number" data-line-number="226"></td>
        <td id="LC226" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(keywords)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L227" class="blob-num js-line-number" data-line-number="227"></td>
        <td id="LC227" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L228" class="blob-num js-line-number" data-line-number="228"></td>
        <td id="LC228" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\vskip</span> 8.5pt <span class="pl-c1">\noindent</span> <span class="pl-c1">\emph</span>{<span class="pl-mi">Keywords</span>}: <span class="pl-s"><span class="pl-pds">$</span>keywords<span class="pl-pds">$</span></span> <span class="pl-c1">\par</span></td>
      </tr>
      <tr>
        <td id="L229" class="blob-num js-line-number" data-line-number="229"></td>
        <td id="LC229" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L230" class="blob-num js-line-number" data-line-number="230"></td>
        <td id="LC230" class="blob-code blob-code-inner js-file-line">    <span class="pl-c1">\hbox</span>{<span class="pl-c1">\vrule</span> height .2pt width 39.14pc}</td>
      </tr>
      <tr>
        <td id="L231" class="blob-num js-line-number" data-line-number="231"></td>
        <td id="LC231" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L232" class="blob-num js-line-number" data-line-number="232"></td>
        <td id="LC232" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L233" class="blob-num js-line-number" data-line-number="233"></td>
        <td id="LC233" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L234" class="blob-num js-line-number" data-line-number="234"></td>
        <td id="LC234" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L235" class="blob-num js-line-number" data-line-number="235"></td>
        <td id="LC235" class="blob-code blob-code-inner js-file-line">    <span class="pl-c1">\hbox</span>{<span class="pl-c1">\vrule</span> height .2pt width 39.14pc}</td>
      </tr>
      <tr>
        <td id="L236" class="blob-num js-line-number" data-line-number="236"></td>
        <td id="LC236" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L237" class="blob-num js-line-number" data-line-number="237"></td>
        <td id="LC237" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L238" class="blob-num js-line-number" data-line-number="238"></td>
        <td id="LC238" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L239" class="blob-num js-line-number" data-line-number="239"></td>
        <td id="LC239" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\end</span>{<span class="pl-smi">abstract</span>}</td>
      </tr>
      <tr>
        <td id="L240" class="blob-num js-line-number" data-line-number="240"></td>
        <td id="LC240" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L241" class="blob-num js-line-number" data-line-number="241"></td>
        <td id="LC241" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L242" class="blob-num js-line-number" data-line-number="242"></td>
        <td id="LC242" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L243" class="blob-num js-line-number" data-line-number="243"></td>
        <td id="LC243" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\vskip</span> -8.5pt</td>
      </tr>
      <tr>
        <td id="L244" class="blob-num js-line-number" data-line-number="244"></td>
        <td id="LC244" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L245" class="blob-num js-line-number" data-line-number="245"></td>
        <td id="LC245" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(toc)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L246" class="blob-num js-line-number" data-line-number="246"></td>
        <td id="LC246" class="blob-code blob-code-inner js-file-line">{</td>
      </tr>
      <tr>
        <td id="L247" class="blob-num js-line-number" data-line-number="247"></td>
        <td id="LC247" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\hypersetup</span>{linkcolor=black}</td>
      </tr>
      <tr>
        <td id="L248" class="blob-num js-line-number" data-line-number="248"></td>
        <td id="LC248" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\setcounter</span>{tocdepth}{<span class="pl-s"><span class="pl-pds">$</span>toc-depth<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L249" class="blob-num js-line-number" data-line-number="249"></td>
        <td id="LC249" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\tableofcontents</span></td>
      </tr>
      <tr>
        <td id="L250" class="blob-num js-line-number" data-line-number="250"></td>
        <td id="LC250" class="blob-code blob-code-inner js-file-line">}</td>
      </tr>
      <tr>
        <td id="L251" class="blob-num js-line-number" data-line-number="251"></td>
        <td id="LC251" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L252" class="blob-num js-line-number" data-line-number="252"></td>
        <td id="LC252" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L253" class="blob-num js-line-number" data-line-number="253"></td>
        <td id="LC253" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span> <span class="pl-c"><span class="pl-c">%</span> removetitleabstract</span></td>
      </tr>
      <tr>
        <td id="L254" class="blob-num js-line-number" data-line-number="254"></td>
        <td id="LC254" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L255" class="blob-num js-line-number" data-line-number="255"></td>
        <td id="LC255" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\noindent</span> <span class="pl-s"><span class="pl-pds">$</span>if(anonymous)<span class="pl-pds">$</span></span><span class="pl-c1">\doublespacing</span><span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span><span class="pl-pds">$</span>if(spacing)<span class="pl-pds">$</span></span><span class="pl-cce">\$</span>spacing<span class="pl-s"><span class="pl-pds">$</span>spacing<span class="pl-pds">$</span></span>endif<span class="pl-s"><span class="pl-pds">$$</span>endif<span class="pl-s"><span class="pl-pds">$</span> </span></span></td>
      </tr>
      <tr>
        <td id="L256" class="blob-num js-line-number" data-line-number="256"></td>
        <td id="LC256" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>if(twocol)<span class="pl-s"><span class="pl-pds">$</span> <span class="pl-c1">\begin</span>{multicols}{2}<span class="pl-pds">$</span></span>endif<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L257" class="blob-num js-line-number" data-line-number="257"></td>
        <td id="LC257" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>body<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L258" class="blob-num js-line-number" data-line-number="258"></td>
        <td id="LC258" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>if(twocol)<span class="pl-s"><span class="pl-pds">$</span> <span class="pl-c1">\end</span>{multicols}<span class="pl-pds">$</span></span>endif<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L259" class="blob-num js-line-number" data-line-number="259"></td>
        <td id="LC259" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"></span></span></td>
      </tr>
      <tr>
        <td id="L260" class="blob-num js-line-number" data-line-number="260"></td>
        <td id="LC260" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>if(endnotes)<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L261" class="blob-num js-line-number" data-line-number="261"></td>
        <td id="LC261" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-c1">\newpage</span></span></span></td>
      </tr>
      <tr>
        <td id="L262" class="blob-num js-line-number" data-line-number="262"></td>
        <td id="LC262" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"></span></span></td>
      </tr>
      <tr>
        <td id="L263" class="blob-num js-line-number" data-line-number="263"></td>
        <td id="LC263" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-c1">\theendnotes</span></span></span></td>
      </tr>
      <tr>
        <td id="L264" class="blob-num js-line-number" data-line-number="264"></td>
        <td id="LC264" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>endif<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L265" class="blob-num js-line-number" data-line-number="265"></td>
        <td id="LC265" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"></span></span></td>
      </tr>
      <tr>
        <td id="L266" class="blob-num js-line-number" data-line-number="266"></td>
        <td id="LC266" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"></span></span></td>
      </tr>
      <tr>
        <td id="L267" class="blob-num js-line-number" data-line-number="267"></td>
        <td id="LC267" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"></span></span></td>
      </tr>
      <tr>
        <td id="L268" class="blob-num js-line-number" data-line-number="268"></td>
        <td id="LC268" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-c1">\newpage</span></span></span></td>
      </tr>
      <tr>
        <td id="L269" class="blob-num js-line-number" data-line-number="269"></td>
        <td id="LC269" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-c1">\singlespacing</span> </span></span></td>
      </tr>
      <tr>
        <td id="L270" class="blob-num js-line-number" data-line-number="270"></td>
        <td id="LC270" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>if(natbib)<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L271" class="blob-num js-line-number" data-line-number="271"></td>
        <td id="LC271" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>if(bibliography)<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L272" class="blob-num js-line-number" data-line-number="272"></td>
        <td id="LC272" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>if(biblio-title)<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L273" class="blob-num js-line-number" data-line-number="273"></td>
        <td id="LC273" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>if(book-class)<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L274" class="blob-num js-line-number" data-line-number="274"></td>
        <td id="LC274" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-c1">\renewcommand\bibname</span>{<span class="pl-s"><span class="pl-pds">$</span>biblio-title<span class="pl-pds">$</span></span>}</span></span></td>
      </tr>
      <tr>
        <td id="L275" class="blob-num js-line-number" data-line-number="275"></td>
        <td id="LC275" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>else<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L276" class="blob-num js-line-number" data-line-number="276"></td>
        <td id="LC276" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-c1">\renewcommand\refname</span>{<span class="pl-s"><span class="pl-pds">$</span>biblio-title<span class="pl-pds">$</span></span>}</span></span></td>
      </tr>
      <tr>
        <td id="L277" class="blob-num js-line-number" data-line-number="277"></td>
        <td id="LC277" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>endif<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L278" class="blob-num js-line-number" data-line-number="278"></td>
        <td id="LC278" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>endif<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L279" class="blob-num js-line-number" data-line-number="279"></td>
        <td id="LC279" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-c1">\bibliography</span>{<span class="pl-s"><span class="pl-pds">$</span>for(bibliography)<span class="pl-pds">$</span><span class="pl-pds">$</span>bibliography<span class="pl-pds">$</span><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span>,<span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span>}</span></span></td>
      </tr>
      <tr>
        <td id="L280" class="blob-num js-line-number" data-line-number="280"></td>
        <td id="LC280" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"></span></span></td>
      </tr>
      <tr>
        <td id="L281" class="blob-num js-line-number" data-line-number="281"></td>
        <td id="LC281" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>endif<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L282" class="blob-num js-line-number" data-line-number="282"></td>
        <td id="LC282" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>endif<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L283" class="blob-num js-line-number" data-line-number="283"></td>
        <td id="LC283" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>if(biblatex)<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L284" class="blob-num js-line-number" data-line-number="284"></td>
        <td id="LC284" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-c1">\printbibliography</span><span class="pl-pds">$</span></span>if(biblio-title)<span class="pl-s"><span class="pl-pds">$</span>[title=<span class="pl-pds">$</span></span>biblio-title<span class="pl-s"><span class="pl-pds">$</span>]<span class="pl-pds">$</span></span>endif<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L285" class="blob-num js-line-number" data-line-number="285"></td>
        <td id="LC285" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"></span></span></td>
      </tr>
      <tr>
        <td id="L286" class="blob-num js-line-number" data-line-number="286"></td>
        <td id="LC286" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>endif<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L287" class="blob-num js-line-number" data-line-number="287"></td>
        <td id="LC287" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>for(include-after)<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L288" class="blob-num js-line-number" data-line-number="288"></td>
        <td id="LC288" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>include-after<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L289" class="blob-num js-line-number" data-line-number="289"></td>
        <td id="LC289" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"></span></span></td>
      </tr>
      <tr>
        <td id="L290" class="blob-num js-line-number" data-line-number="290"></td>
        <td id="LC290" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>endfor<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L291" class="blob-num js-line-number" data-line-number="291"></td>
        <td id="LC291" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-c1">\end</span>{document}</span></span></td>
      </tr>
</table>

  <details class="details-reset details-overlay BlobToolbar position-absolute js-file-line-actions dropdown d-none" aria-hidden="true">
    <summary class="btn-octicon ml-0 px-2 p-0 bg-white border border-gray-dark rounded-1" aria-label="Inline file action toolbar">
      <svg class="octicon octicon-kebab-horizontal" viewBox="0 0 13 16" version="1.1" width="13" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M1.5 9a1.5 1.5 0 100-3 1.5 1.5 0 000 3zm5 0a1.5 1.5 0 100-3 1.5 1.5 0 000 3zM13 7.5a1.5 1.5 0 11-3 0 1.5 1.5 0 013 0z"/></svg>
    </summary>
    <details-menu>
      <ul class="BlobToolbar-dropdown dropdown-menu dropdown-menu-se mt-2" style="width:185px">
        <li>
          <clipboard-copy role="menuitem" class="dropdown-item" id="js-copy-lines" style="cursor:pointer;">
            Copy lines
          </clipboard-copy>
        </li>
        <li>
          <clipboard-copy role="menuitem" class="dropdown-item" id="js-copy-permalink" style="cursor:pointer;">
            Copy permalink
          </clipboard-copy>
        </li>
        <li><a class="dropdown-item js-update-url-with-hash" id="js-view-git-blame" role="menuitem" href="/svmiller/svm-r-markdown-templates/blame/6b678ff10ecb592214fb21f8f6842ae36d9cb8b9/svm-latex-ms.tex">View git blame</a></li>
          <li><a class="dropdown-item" id="js-new-issue" role="menuitem" href="/svmiller/svm-r-markdown-templates/issues/new">Reference in new issue</a></li>
      </ul>
    </details-menu>
  </details>

  </div>

    </div>

  <details class="details-reset details-overlay details-overlay-dark">
    <summary data-hotkey="l" aria-label="Jump to line"></summary>
    <details-dialog class="Box Box--overlay d-flex flex-column anim-fade-in fast linejump" aria-label="Jump to line">
      <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="js-jump-to-line-form Box-body d-flex" action="" accept-charset="UTF-8" method="get"><input name="utf8" type="hidden" value="&#x2713;" />
        <input class="form-control flex-auto mr-3 linejump-input js-jump-to-line-field" type="text" placeholder="Jump to line&hellip;" aria-label="Jump to line" autofocus>
        <button type="submit" class="btn" data-close-dialog>Go</button>
</form>    </details-dialog>
  </details>

  </div>
</div>

    </main>
  </div>

  </div>

<div class="footer container-lg width-full p-responsive" role="contentinfo">
  <div class="position-relative d-flex flex-row-reverse flex-lg-row flex-wrap flex-lg-nowrap flex-justify-center flex-lg-justify-between pt-6 pb-2 mt-6 f6 text-gray border-top border-gray-light ">
    <ul class="list-style-none d-flex flex-wrap col-12 col-lg-5 flex-justify-center flex-lg-justify-between mb-2 mb-lg-0">
      <li class="mr-3 mr-lg-0">&copy; 2020 GitHub, Inc.</li>
        <li class="mr-3 mr-lg-0"><a data-ga-click="Footer, go to terms, text:terms" href="https://github.com/site/terms">Terms</a></li>
        <li class="mr-3 mr-lg-0"><a data-ga-click="Footer, go to privacy, text:privacy" href="https://github.com/site/privacy">Privacy</a></li>
        <li class="mr-3 mr-lg-0"><a data-ga-click="Footer, go to security, text:security" href="https://github.com/security">Security</a></li>
        <li class="mr-3 mr-lg-0"><a href="https://githubstatus.com/" data-ga-click="Footer, go to status, text:status">Status</a></li>
        <li><a data-ga-click="Footer, go to help, text:help" href="https://help.github.com">Help</a></li>
    </ul>

    <a aria-label="Homepage" title="GitHub" class="footer-octicon d-none d-lg-block mx-lg-4" href="https://github.com">
      <svg height="24" class="octicon octicon-mark-github" viewBox="0 0 16 16" version="1.1" width="24" aria-hidden="true"><path fill-rule="evenodd" d="M8 0C3.58 0 0 3.58 0 8c0 3.54 2.29 6.53 5.47 7.59.4.07.55-.17.55-.38 0-.19-.01-.82-.01-1.49-2.01.37-2.53-.49-2.69-.94-.09-.23-.48-.94-.82-1.13-.28-.15-.68-.52-.01-.53.63-.01 1.08.58 1.23.82.72 1.21 1.87.87 2.33.66.07-.52.28-.87.51-1.07-1.78-.2-3.64-.89-3.64-3.95 0-.87.31-1.59.82-2.15-.08-.2-.36-1.02.08-2.12 0 0 .67-.21 2.2.82.64-.18 1.32-.27 2-.27.68 0 1.36.09 2 .27 1.53-1.04 2.2-.82 2.2-.82.44 1.1.16 1.92.08 2.12.51.56.82 1.27.82 2.15 0 3.07-1.87 3.75-3.65 3.95.29.25.54.73.54 1.48 0 1.07-.01 1.93-.01 2.2 0 .21.15.46.55.38A8.013 8.013 0 0016 8c0-4.42-3.58-8-8-8z"/></svg>
</a>
   <ul class="list-style-none d-flex flex-wrap col-12 col-lg-5 flex-justify-center flex-lg-justify-between mb-2 mb-lg-0">
        <li class="mr-3 mr-lg-0"><a data-ga-click="Footer, go to contact, text:contact" href="https://github.com/contact">Contact GitHub</a></li>
        <li class="mr-3 mr-lg-0"><a href="https://github.com/pricing" data-ga-click="Footer, go to Pricing, text:Pricing">Pricing</a></li>
      <li class="mr-3 mr-lg-0"><a href="https://developer.github.com" data-ga-click="Footer, go to api, text:api">API</a></li>
      <li class="mr-3 mr-lg-0"><a href="https://training.github.com" data-ga-click="Footer, go to training, text:training">Training</a></li>
        <li class="mr-3 mr-lg-0"><a href="https://github.blog" data-ga-click="Footer, go to blog, text:blog">Blog</a></li>
        <li><a data-ga-click="Footer, go to about, text:about" href="https://github.com/about">About</a></li>

    </ul>
  </div>
  <div class="d-flex flex-justify-center pb-6">
    <span class="f6 text-gray-light"></span>
  </div>
</div>

  <div id="ajax-error-message" class="ajax-error-message flash flash-error">
    <svg class="octicon octicon-alert" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.893 1.5c-.183-.31-.52-.5-.887-.5s-.703.19-.886.5L.138 13.499a.98.98 0 000 1.001c.193.31.53.501.886.501h13.964c.367 0 .704-.19.877-.5a1.03 1.03 0 00.01-1.002L8.893 1.5zm.133 11.497H6.987v-2.003h2.039v2.003zm0-3.004H6.987V5.987h2.039v4.006z"/></svg>
    <button type="button" class="flash-close js-ajax-error-dismiss" aria-label="Dismiss error">
      <svg class="octicon octicon-x" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48L7.48 8z"/></svg>
    </button>
    You can’t perform that action at this time.
  </div>

    <script crossorigin="anonymous" async="async" integrity="sha512-VTkKwyyXYz1e8w0v/7LXDKSa7yMy1qEQofgf/5bGrUv8wpbpaZxx5S3Uc6oYrvbOe432HJdJG5qsFdM9sbP+wg==" type="application/javascript" id="js-conditional-compat" data-src="https://github.githubassets.com/assets/compat-bootstrap-55390ac3.js"></script>
    <script crossorigin="anonymous" async="async" integrity="sha512-3yNijdFdVDBZQDWrBvMeD2J9gyXwI5MKUMJSWdEGP44DgS4NEPQw2TmVlDdNAWrseJO5C/sXBSTrL24DvGMDJw==" type="application/javascript" src="https://github.githubassets.com/assets/vendor-df23628d.js"></script>
    <script crossorigin="anonymous" integrity="sha512-pa+/BDovC8blIaZGNtVJ69/u0z2fhC5e/L9DzR8BO6t9Hl4djSDIShYmnFiMrfMSQ+T/H593bhxDcb3MBV/kTA==" type="application/javascript" src="https://github.githubassets.com/assets/frameworks-a5afbf04.js"></script>

    <script crossorigin="anonymous" async="async" integrity="sha512-3bq0hvgN3zABpTqliAmMBZcvVdvGiXaOuEXP1rDTfDoX5PaddILM/Sw1FCotmCXDHg8X/0GxDF+GYMHY1xC4aA==" type="application/javascript" src="https://github.githubassets.com/assets/github-bootstrap-ddbab486.js"></script>

  <div class="js-stale-session-flash flash flash-warn flash-banner" hidden
    >
    <svg class="octicon octicon-alert" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.893 1.5c-.183-.31-.52-.5-.887-.5s-.703.19-.886.5L.138 13.499a.98.98 0 000 1.001c.193.31.53.501.886.501h13.964c.367 0 .704-.19.877-.5a1.03 1.03 0 00.01-1.002L8.893 1.5zm.133 11.497H6.987v-2.003h2.039v2.003zm0-3.004H6.987V5.987h2.039v4.006z"/></svg>
    <span class="js-stale-session-flash-signed-in" hidden>You signed in with another tab or window. <a href="">Reload</a> to refresh your session.</span>
    <span class="js-stale-session-flash-signed-out" hidden>You signed out in another tab or window. <a href="">Reload</a> to refresh your session.</span>
  </div>
  <template id="site-details-dialog">
  <details class="details-reset details-overlay details-overlay-dark lh-default text-gray-dark hx_rsm" open>
    <summary role="button" aria-label="Close dialog"></summary>
    <details-dialog class="Box Box--overlay d-flex flex-column anim-fade-in fast hx_rsm-dialog hx_rsm-modal">
      <button class="Box-btn-octicon m-0 btn-octicon position-absolute right-0 top-0" type="button" aria-label="Close dialog" data-close-dialog>
        <svg class="octicon octicon-x" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48L7.48 8z"/></svg>
      </button>
      <div class="octocat-spinner my-6 js-details-dialog-spinner"></div>
    </details-dialog>
  </details>
</template>

  <div class="Popover js-hovercard-content position-absolute" style="display: none; outline: none;" tabindex="0">
  <div class="Popover-message Popover-message--bottom-left Popover-message--large Box box-shadow-large" style="width:360px;">
  </div>
</div>

  <div aria-live="polite" class="js-global-screen-reader-notice sr-only"></div>

  </body>
</html>

### 2020-02-27 paper_elsevier/paper_elsevier.Rmd
working manuscript files

---
title: "*dfoliatR*: An R package for detection and analysis of insect defoliation signals in tree rings"
author:
  - name: Christopher H. Guiterman
    email: chguiterman@email.arizona.edu
    affiliation: a,b
    footnote: 1
  - name: Ann M. Lynch
    email: lyncha@email.arizona.edu
    affiliation: a,c
    # footnote: [2, 3]
  - name: Jodi N. Axelson
    email: jodi.axelson@berkeley.edu
    affiliation: c
    # footnote: 4
address:
  - code: a
    address: "Laboratory of Tree-Ring Research, University of Arizona, 1215 E Lowell St. Box 210045, Tucson, AZ, 85721"
  - code: b
    address: "Three Pines Forest Research, LLC, PO Box 225, Etna, NH, 03750"
  - code: c
    address: "U.S. Forest Service, Rocky Mountain Research Station, 1215 E Lowell St. Box 210045, Tucson, AZ, 85721"
  - code: d
    address: "Dept of Environmental Science, Policy & Management, University of California, Berkeley, Berkeley, CA 94720"
footnote:
  - code: 1
    text: "Corresponding Author"
#   - code: 2
#     text: "Laboratory of Tree-Ring Research, University of Arizona, 1215 E Lowell St. Box 210045, Tucson, AZ, 85721"
#   - code: 3
#     text: "U.S. Forest Service, Rocky Mountain Research Station, 1215 E Lowell St. Box 210045, Tucson, AZ, 85721"
#   - code: 4
#     text: "Dept of Environmental Science, Policy & Management, University of California, Berkeley, Berkeley, CA 94720"
abstract: |
    We present a new R package to provide dendroecologists with tools to identify, quantify, analyze, and visualize growth
    suppression events in tree rings produced by insect defoliation. The 'dfoliatR' library is based on the Fortran V program 
    OUTBREAK, and builds on existing resources in the R computing environment. 'dfoliatR' expands on OUTBREAK to provide greater
    control of supression thresholds, additional output tables, and high-quality graphics. To use 'dfoliatR' requires 
    standardized ring-width measurements from insect host trees and an indexed tree-ring chronology from local non-host trees. 
    It performs an indexing procedure to remove the climatic signal represented in the non-host chronology from the host-tree
    series. It then infers defoliation events in individual trees. Site-level analyses identify outbreak events
    that synchronously affect a user-defined number or proportion of the host trees. Functions are available for
    summary statistics and graphics of tree- and site-level series.

    \hfill\break
keywords: |
  Dendroecology, spruce budworm, Choristoneura, pandora moth, Coloradia pandora Blake, larch-bud-moth
  \newpage
journal: "Tree-Ring Research"
date: "`r Sys.Date()`"
bibliography: mybibfile.bib
linenumbers: true
#numbersections: true
csl: elsevier-harvard.csl
output: rticles::elsevier_article
# output:
#   bookdown::pdf_book:
#     base_format: rticles::elsevier_article
#   bookdown::word_document2:
#     reference_docx: "../common/word-styles-reference-01.docx"
#     toc: false
#     toc_depth: 2
#     fig_caption: yes
#     df_print: kable
layout: review
---

Introduction
==========================

Overview of the software
==========================

The `dfoliatR` library requires two sets of tree-ring data to identify defoliation and outbreak events:

* Standardized ring-width series for individual trees of the host species
*	Standardized tree-ring chronology from a local non-host species

Users can develop these data sets in software of their choosing, such as in `dplR` or ARSTAN. It is important that the host-tree data include only one tree-ring series per tree. Both `dplR` and ARSTAN have options for averaging mutliple sample series into a tree-level series. The tree-ring series and chronology can be read into R via several available `dplR` functions.

`dfoliatR` begins to identify defoliation events in individual trees by removing the climatic signal as represented by the non-host chronology from the host tree series. This indexing procedure creates a series -- the "growth suppression index" -- in which disturbance is the predominant signal. Negative departures in the growth suppression index that surpass user-defined thresholds in duration, and magnitude are defined as *defoliation events*. The individual tree defoliation series are composited in an additional step to identify *outbreak events* that synchronously affect mutliple trees. Users have options to define the number and/or the proportion of trees required for a defoliation event to be considered an outbreak.

Note that these methods of separating tree- vs site-level disturbance categories is a major departure from the OUTBREAK program. In OUTBREAK the two levels of analysis are combined and users have more limited control of thresholds to define defoliation events versus outbreaks.

Availability and installation
==========================

The `dfoliatR` is provided free and open source [@chris_guiterman_2020_3626136]. It is provided to R users via the Comprehensive R Archive Network (CRAN; [https://cran.r-project.org/](https://cran.r-project.org/)). To install `dfoliatR` into R from CRAN use

```{r, echo=TRUE, eval=FALSE}
install.packages("dfoliatR")
```

In each R session, `dfoliatR` can be loaded via

```{r, echo=TRUE, eval=FALSE}
library(dfoliatR)
```

Development versions of `dfoliatR` are available on Github and installed using the `devtools` library in R,

```{r, echo=TRUE, eval=FALSE}
devtools::install_github("chguiterman/dfoliatR")
```

Support documentation in addition to this paper is provided within the package via help menus (accessed by typing `?` before a function name) and on the package website ([https://chguiterman.github.io/dfoliatR/](https://chguiterman.github.io/dfoliatR/)) which includes up-to-date vignettes. Code to execute a preprint of this manuscript and the code included below is available at (*insert later*).

Tree-level defoliation events
==========================

Site-level events
==========================

Evaluation
==========================

Extensions
==========================

References {#references .unnumbered}
==========================

### 2020-02-27 paper_elsevier/paper_elsevier.tex
working manuscript files

\documentclass[review]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{Tree-Ring Research} % Sets Journal name

\usepackage{lineno} % add
  \linenumbers % turns line numbering on
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={dfoliatR: An R package for detection and analysis of insect defoliation signals in tree rings},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}

% Pandoc header

\begin{document}
\begin{frontmatter}

  \title{\emph{dfoliatR}: An R package for detection and analysis of insect
defoliation signals in tree rings}
    \author[a,b]{Christopher H. Guiterman\corref{1}}
   \ead{chguiterman@email.arizona.edu} 
    \author[a,c]{Ann M. Lynch}
   \ead{lyncha@email.arizona.edu} 
    \author[c]{Jodi N. Axelson}
   \ead{jodi.axelson@berkeley.edu} 
      \address[a]{Laboratory of Tree-Ring Research, University of Arizona, 1215 E Lowell
St.~Box 210045, Tucson, AZ, 85721}
    \address[b]{Three Pines Forest Research, LLC, PO Box 225, Etna, NH, 03750}
    \address[c]{U.S. Forest Service, Rocky Mountain Research Station, 1215 E Lowell
St.~Box 210045, Tucson, AZ, 85721}
    \address[d]{Dept of Environmental Science, Policy \& Management, University of
California, Berkeley, Berkeley, CA 94720}
      \cortext[1]{Corresponding Author}

  \begin{abstract}
  We present a new R package to provide dendroecologists with tools to
  identify, quantify, analyze, and visualize growth suppression events in
  tree rings produced by insect defoliation. The `dfoliatR' library is
  based on the Fortran V program OUTBREAK, and builds on existing
  resources in the R computing environment. `dfoliatR' expands on OUTBREAK
  to provide greater control of supression thresholds, additional output
  tables, and high-quality graphics. To use `dfoliatR' requires
  standardized ring-width measurements from insect host trees and an
  indexed tree-ring chronology from local non-host trees. It performs an
  indexing procedure to remove the climatic signal represented in the
  non-host chronology from the host-tree series. It then infers
  defoliation events in individual trees. Site-level analyses identify
  outbreak events that synchronously affect a user-defined number or
  proportion of the host trees. Functions are available for summary
  statistics and graphics of tree- and site-level series.

  \hfill\break
  \end{abstract}
   \begin{keyword} Dendroecology, spruce budworm, Choristoneura, pandora moth, Coloradia
pandora Blake, larch-bud-moth \newpage\end{keyword}
 \end{frontmatter}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\hypertarget{overview-of-the-software}{%
\section{Overview of the software}\label{overview-of-the-software}}

The \texttt{dfoliatR} library requires two sets of tree-ring data to
identify defoliation and outbreak events:

\begin{itemize}
\tightlist
\item
  Standardized ring-width series for individual trees of the host
  species
\item
  Standardized tree-ring chronology from a local non-host species
\end{itemize}

Users can develop these data sets in software of their choosing, such as
in \texttt{dplR} or ARSTAN. It is important that the host-tree data
include only one tree-ring series per tree. Both \texttt{dplR} and
ARSTAN have options for averaging mutliple sample series into a
tree-level series. The tree-ring series and chronology can be read into
R via several available \texttt{dplR} functions.

\texttt{dfoliatR} begins to identify defoliation events in individual
trees by removing the climatic signal as represented by the non-host
chronology from the host tree series. This indexing procedure creates a
series -- the ``growth suppression index'' -- in which disturbance is
the predominant signal. Negative departures in the growth suppression
index that surpass user-defined thresholds in duration, and magnitude
will be defined as \emph{defoliation events}. The individual tree
defoliation series are composited in an additional step to identify
\emph{outbreak events} that synchronously affect mutliple trees. Users
have options to define the number and/or the proportion of trees
required for a defoliation event to be considered an outbreak.

Note that these methods of separating tree- vs site-level disturbance
categories is a major departure from the OUTBREAK program. In OUTBREAK
the two levels of analysis are combined and users have more limited
control of thresholds to define defoliation events versus outbreaks.

\hypertarget{availability-and-installation}{%
\section{Availability and
installation}\label{availability-and-installation}}

The \texttt{dfoliatR} is provided free and open source (Guiterman,
2020). It is provided to R users via the Comprehensive R Archive Network
(CRAN; \url{https://cran.r-project.org/}). To install \texttt{dfoliatR}
into R from CRAN use

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{install.packages}\NormalTok{(}\StringTok{"dfoliatR"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

In each R session, \texttt{dfoliatR} can be loaded via

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(dfoliatR)}
\end{Highlighting}
\end{Shaded}

Development versions of \texttt{dfoliatR} are available on Github and
installed using the \texttt{devtools} library in R,

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{devtools}\OperatorTok{::}\KeywordTok{install_github}\NormalTok{(}\StringTok{"chguiterman/dfoliatR"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Support documentation is provided within the package via help menus
(accessed by typing \texttt{?} before a function name), this paper, and
on the package website (\url{https://chguiterman.github.io/dfoliatR/})
which includes updated vignettes. Code to execute a preprint of this
manuscript and the code included below is available at (\emph{insert
later}).

\hypertarget{tree-level-defoliation-events}{%
\section{Tree-level defoliation
events}\label{tree-level-defoliation-events}}

\hypertarget{site-level-events}{%
\section{Site-level events}\label{site-level-events}}

\hypertarget{evaluation}{%
\section{Evaluation}\label{evaluation}}

\hypertarget{extensions}{%
\section{Extensions}\label{extensions}}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-chris_guiterman_2020_3626136}{}%
Guiterman, C., 2020. Chguiterman/dfoliatR: First release. Zenodo.
doi:\href{https://doi.org/10.5281/zenodo.3626136}{10.5281/zenodo.3626136}

\end{document}

### 2020-03-02 paper_elsevier/paper_elsevier.Rmd
paper edits

@@
 #numbersections: true
-csl: elsevier-harvard.csl
+csl: ecology.csl
 output: rticles::elsevier_article
@@
 layout: review
+preamble: |
+  \usepackage{amsmath}
 ---
@@
 ==========================
+
+The `dfoliatR` library is unique among a growing suite `R` packages designed for dendrochronology.
+Stemming from the `dplR` library [@Bunn2008] that provides `R` the ability to read and write an array of tree-ring data formats, standaridize ring widths, build and evaluate chronologies, perform qaulity control (to name a few), one can now also measure, perform and check crossdating, and perform many analytical tests [@Bunn2010; @Jevsenak2018; @Lara2015; @Zang2015].
+Tools for assessing stand dynamics and disturbance analyses are under rapid development, with new packages for assessing growth and release events [`TRADER`: @Altman2014], metrics of growth resilience [`pointRes`: @VanderMaaten-Theunissen2015], and fire history [`burnr`: @Malevich2018].
+The key objective of `dfoliatR` is to provide tools to identify and analyse insect defoliation and outbreak events by building on the methods employed the FORTRAN program OUTBREAK [@Swetnam1989].
+What sets `dfoliatR` apart from packages such as `TRADER` is that it explicitly performs an indexing procedue on host-tree series to remove climatic and other non-defoliation related signals represented by non-host tree species.
+Insect defoliation signals are identified in the disturbance index by the duration and magnitude of negative departures.
+
+`dfoliatR` draws upon data formats in `dplR` that are commonly employed by other tree-ring libraries.
+It uses and outputs data formats that faciliate the use packages embodied by `tidyverse` [@Wickham2019] that include efficient manipulation of data [`dplyr`: @Wickham2020dplyr] and graphics [`ggplot2`: @wickham2016ggplot2].
+
+In this paper, we describe the statistical methods employed by `dfoliatR`, its availability, and run through analyses for a signle site in New Mexico.
+Users need not have much experience in `R` to replicate the analyses and graphics below.
+All `R` code presented below is executatable in an `R` session once the required libraries are installed and loaded.
+Support documentation in addition to this paper is provided within the package via standard help menus (accessed by typing `?` before a function name) and on the package website ([https://chguiterman.github.io/dfoliatR/](https://chguiterman.github.io/dfoliatR/)), which includes up-to-date vignettes that describe in detail the functionality of the software.
+Code to create a preprint of this manuscript including the `R` scripts is available from [https://github.com/chguiterman/dfoliatR_paper](https://github.com/chguiterman/dfoliatR_paper).
 
@@
 Both `dplR` and ARSTAN have options for averaging mutliple sample series into a tree-level series.
-The tree-ring series and chronology can be read into R via several available `dplR` functions.
+The tree-ring series and chronology can be read into `R` via several available `dplR` functions.
 
 `dfoliatR` begins to identify defoliation events in individual trees by removing the climatic signal as represented by the non-host chronology from the host tree series.
-This indexing procedure creates a series -- the "growth suppression index" -- in which disturbance is the predominant signal.
-Negative departures in the growth suppression index that surpass user-defined thresholds in duration, and magnitude are defined as *defoliation events*.
+This indexing procedure creates a series -- the "growth suppression index" (GSI) -- in which disturbance is the predominant signal.
+The GSI is caluclated as
+\begin{align}
+\textrm{GSI}_{i} = \textrm{H}_{i} - \left( \textrm{NH}_{i} - \overline{\textrm{NH}} \right) \frac{\sigma_{\textrm{H}}}{\sigma_{\textrm{NH}}} \
+\end{align}
+where H and NH are the host tree series and the non-host chronology, in year i, respectively [@Swetnam1985; @Swetnam1989].
+Only years in which the individual host-tree series and the non-host chronology overlap are used in Equation 1.
+The non-host chronology is scaled by its mean ($\overline{\textrm{NH}} \approx 1.0$) and multiplied by the ratio of host and non-host standard deviations ($\frac{\sigma_{\textrm{H}}}{\sigma_{\textrm{NH}}}$), which approximates the variance of the host tree series.
+This "corrected" non-host chronology is subtracted from the host-tree series.
+
+Negative departures in the normalized growth suppression index that surpass user-defined thresholds in duration and magnitude will be defined as *defoliation events*.
+
 The individual tree defoliation series are composited in an additional step to identify *outbreak events* that synchronously affect mutliple trees.
@@
 The `dfoliatR` is provided free and open source [@chris_guiterman_2020_3626136].
-It is provided to R users via the Comprehensive R Archive Network (CRAN; [https://cran.r-project.org/](https://cran.r-project.org/)).
-To install `dfoliatR` into R from CRAN use
+It is provided to `R` users via the Comprehensive R Archive Network (CRAN; [https://cran.r-project.org/](https://cran.r-project.org/)).
+To install `dfoliatR` from CRAN use
 
@@
 
-In each R session, `dfoliatR` can be loaded via
+In each `R` session, `dfoliatR` can be loaded via
 
@@
 
-Development versions of `dfoliatR` are available on Github and installed using the `devtools` library in R,
+Development versions of `dfoliatR` are available on Github and installed using the `devtools` library,
 
@@
 
-Support documentation in addition to this paper is provided within the package via help menus (accessed by typing `?` before a function name) and on the package website ([https://chguiterman.github.io/dfoliatR/](https://chguiterman.github.io/dfoliatR/)) which includes up-to-date vignettes.
-Code to execute a preprint of this manuscript and the code included below is available at (*insert later*).
+Issues, bug reports, and ideas for improving `dfoliatR` can be posted to [https://github.com/chguiterman/dfoliatR/issues](https://github.com/chguiterman/dfoliatR/issues).
+As an Open Source library, we welcome and encourage community involvement in future development.
+The best ways to contribute to `dfoliatR` are through standard Github procedures or by contacting the first author.
 
-Tree-level defoliation events
+Usage
 ==========================
 
-Site-level events
-==========================
+## Tree-level defoliation events
+
+##Site-level events
 

### 2020-03-02 paper_elsevier/paper_elsevier.tex
paper edits

@@
 
+\usepackage{amsmath}
+
 \begin{document}
@@
 
+The \texttt{dfoliatR} library is unique among a growing suite \texttt{R}
+packages designed for dendrochronology. Stemming from the \texttt{dplR}
+library (Bunn 2008) that provides \texttt{R} the ability to read and
+write an array of tree-ring data formats, standaridize ring widths,
+build and evaluate chronologies, perform qaulity control (to name a
+few), one can now also measure, perform and check crossdating, and
+perform many analytical tests (Bunn 2010, Lara et al. 2015, Zang and
+Biondi 2015, Jevšenak and Levanič 2018). Tools for assessing stand
+dynamics and disturbance analyses are under rapid development, with new
+packages for assessing growth and release events (\texttt{TRADER}:
+Altman et al. 2014), metrics of growth resilience (\texttt{pointRes}:
+Maaten-Theunissen et al. 2015), and fire history (\texttt{burnr}:
+Malevich et al. 2018). The key objective of \texttt{dfoliatR} is to
+provide tools to identify and analyse insect defoliation and outbreak
+events by building on the methods employed the FORTRAN program OUTBREAK
+(Swetnam and Lynch 1989). What sets \texttt{dfoliatR} apart from
+packages such as \texttt{TRADER} is that it explicitly performs an
+indexing procedue on host-tree series to remove climatic and other
+non-defoliation related signals represented by non-host tree species.
+Insect defoliation signals are identified in the disturbance index by
+the duration and magnitude of negative departures.
+
+\texttt{dfoliatR} draws upon data formats in \texttt{dplR} that are
+commonly employed by other tree-ring libraries. It uses and outputs a
+data formats that faciliate the use packages embodied by
+\texttt{tidyverse} (Wickham et al. 2019) that include efficient data
+wrangling (\texttt{dplyr}: Wickham et al. 2020) and graphics
+(\texttt{ggplot2}: Wickham 2016).
+
+In this paper, we describe the statistical methods employed by
+\texttt{dfoliatR}, its availability, and run through analyses for a
+signle site in New Mexico. Users need not have much experience in
+\texttt{R} to replicate the analyses and graphics below. All \texttt{R}
+code presented below is executatable in an \texttt{R} session once the
+required libraries are installed and loaded. Support documentation in
+addition to this paper is provided within the package via standard help
+menus (accessed by typing \texttt{?} before a function name) and on the
+package website (\url{https://chguiterman.github.io/dfoliatR/}), which
+includes up-to-date vignettes that describe in detail the functionality
+of the software. Code to create a preprint of this manuscript including
+the \texttt{R} scripts is available from
+\url{https://github.com/chguiterman/dfoliatR_paper}.
+
 \hypertarget{overview-of-the-software}{%
@@
 tree-level series. The tree-ring series and chronology can be read into
-R via several available \texttt{dplR} functions.
+\texttt{R} via several available \texttt{dplR} functions.
 
@@
 chronology from the host tree series. This indexing procedure creates a
-series -- the ``growth suppression index'' -- in which disturbance is
-the predominant signal. Negative departures in the growth suppression
-index that surpass user-defined thresholds in duration, and magnitude
-will be defined as \emph{defoliation events}. The individual tree
-defoliation series are composited in an additional step to identify
-\emph{outbreak events} that synchronously affect mutliple trees. Users
-have options to define the number and/or the proportion of trees
-required for a defoliation event to be considered an outbreak.
+series -- the ``growth suppression index'' (GSI) -- in which disturbance
+is the predominant signal. The GSI is caluclated as \begin{align}
+\textrm{GSI}_{i} = \textrm{H}_{i} - \left( \textrm{NH}_{i} - \overline{\textrm{NH}} \right) \frac{\sigma_{\textrm{H}}}{\sigma_{\textrm{NH}}} \
+\end{align} where H and NH are the host tree series and the non-host
+chronology, in year i, respectively (Swetnam et al. 1985, Swetnam and
+Lynch 1989). Only years in which the individual host-tree series and the
+non-host chronology overlap are used in Equation 1. The non-host
+chronology is scaled by its mean
+(\(\overline{\textrm{NH}} \approx 1.0\)) and multiplied by the ratio of
+host and non-host standard deviations
+(\(\frac{\sigma_{\textrm{H}}}{\sigma_{\textrm{NH}}}\)), which
+approximates the variance of the host tree series. This ``corrected''
+non-host chronology is subtracted from the host-tree series.
+
+Negative departures in the normalized growth suppression index that
+surpass user-defined thresholds in duration and magnitude will be
+defined as \emph{defoliation events}.
+
+The individual tree defoliation series are composited in an additional
+step to identify \emph{outbreak events} that synchronously affect
+mutliple trees. Users have options to define the number and/or the
+proportion of trees required for a defoliation event to be considered an
+outbreak.
 
@@
 
-The \texttt{dfoliatR} is provided free and open source (Guiterman,
-2020). It is provided to R users via the Comprehensive R Archive Network
-(CRAN; \url{https://cran.r-project.org/}). To install \texttt{dfoliatR}
-into R from CRAN use
+The \texttt{dfoliatR} is provided free and open source (Guiterman 2020).
+It is provided to \texttt{R} users via the Comprehensive R Archive
+Network (CRAN; \url{https://cran.r-project.org/}). To install
+\texttt{dfoliatR} from CRAN use
 
@@
 
-In each R session, \texttt{dfoliatR} can be loaded via
+In each \texttt{R} session, \texttt{dfoliatR} can be loaded via
 
@@
 Development versions of \texttt{dfoliatR} are available on Github and
-installed using the \texttt{devtools} library in R,
+installed using the \texttt{devtools} library,
 
@@
 
-Support documentation is provided within the package via help menus
-(accessed by typing \texttt{?} before a function name), this paper, and
-on the package website (\url{https://chguiterman.github.io/dfoliatR/})
-which includes updated vignettes. Code to execute a preprint of this
-manuscript and the code included below is available at (\emph{insert
-later}).
+Issues, bug reports, and ideas for improving \texttt{dfoliatR} can be
+posted to \url{https://github.com/chguiterman/dfoliatR/issues}. As an
+Open Source library, we welcome and encourage community involvement in
+future development. The best ways to contribute to \texttt{dfoliatR} are
+through standard Github procedures or by contacting the first author.
+
+\hypertarget{usage}{%
+\section{Usage}\label{usage}}
 
 \hypertarget{tree-level-defoliation-events}{%
-\section{Tree-level defoliation
+\subsection{Tree-level defoliation
 events}\label{tree-level-defoliation-events}}
 
-\hypertarget{site-level-events}{%
-\section{Site-level events}\label{site-level-events}}
+\#\#Site-level events
 
@@
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-Altman2014}{}%
+Altman, J., P. Fibich, J. Dolezal, and T. Aakala. 2014. TRADER: A
+package for Tree Ring Analysis of Disturbance Events in R.
+Dendrochronologia 32:107--112.
+
+\leavevmode\hypertarget{ref-Bunn2008}{}%
+Bunn, A. G. 2008. A dendrochronology program library in R (dplR).
+Dendrochronologia 26:115--124.
+
+\leavevmode\hypertarget{ref-Bunn2010}{}%
+Bunn, A. G. 2010. Statistical and visual crossdating in R using the dplR
+library. Dendrochronologia 28:251--258.
+
 \leavevmode\hypertarget{ref-chris_guiterman_2020_3626136}{}%
-Guiterman, C., 2020. Chguiterman/dfoliatR: First release. Zenodo.
-doi:\href{https://doi.org/10.5281/zenodo.3626136}{10.5281/zenodo.3626136}
+Guiterman, C. 2020. Chguiterman/dfoliatR: First release. Zenodo.
+
+\leavevmode\hypertarget{ref-Jevsenak2018}{}%
+Jevšenak, J., and T. Levanič. 2018. dendroTools: R package for studying
+linear and nonlinear responses between tree-rings and daily
+environmental data. Dendrochronologia 48:32--39.
+
+\leavevmode\hypertarget{ref-Lara2015}{}%
+Lara, W., F. Bravo, and C. A. Sierra. 2015. MeasuRing: An R package to
+measure tree-ring widths from scanned images. Dendrochronologia
+34:43--50.
+
+\leavevmode\hypertarget{ref-VanderMaaten-Theunissen2015}{}%
+Maaten-Theunissen, M. van der, E. van der Maaten, and O. Bouriaud. 2015.
+PointRes: An R package to analyze pointer years and components of
+resilience. Dendrochronologia 35:34--38.
+
+\leavevmode\hypertarget{ref-Malevich2018}{}%
+Malevich, S. B., C. H. Guiterman, and E. Q. Margolis. 2018. burnr: Fire
+history analysis and graphics in R. Dendrochronologia 49:9--15.
+
+\leavevmode\hypertarget{ref-Swetnam1989}{}%
+Swetnam, T. W., and A. M. Lynch. 1989. A tree-ring reconstruction of
+western spruce budworm history in the southern Rocky Mountains. Forest
+Science 35:962--986.
+
+\leavevmode\hypertarget{ref-Swetnam1985}{}%
+Swetnam, T. W., M. A. Thompson, and E. K. Sutherland. 1985. Using
+dendrochronology to measure radial growth of defoliated trees. Page 39p.
+
+\leavevmode\hypertarget{ref-wickham2016ggplot2}{}%
+Wickham, H. 2016. Ggplot2: Elegant graphics for data analysis. Springer.
+
+\leavevmode\hypertarget{ref-Wickham2019}{}%
+Wickham, H., M. Averick, J. Bryan, W. Chang, L. McGowan, R. François, G.
+Grolemund, A. Hayes, L. Henry, J. Hester, M. Kuhn, T. Pedersen, E.
+Miller, S. Bache, K. Müller, J. Ooms, D. Robinson, D. Seidel, V. Spinu,
+K. Takahashi, D. Vaughan, C. Wilke, K. Woo, and H. Yutani. 2019. Welcome
+to the tidyverse. Journal of Open Source Software 4:1686.
+
+\leavevmode\hypertarget{ref-Wickham2020dplyr}{}%
+Wickham, H., R. François, L. Henry, and K. Müller. 2020. Dplyr: A
+grammar of data manipulation.
+
+\leavevmode\hypertarget{ref-Zang2015}{}%
+Zang, C., and F. Biondi. 2015. Treeclim: An R package for the numerical
+calibration of proxy-climate relationships. Ecography 38:431--436.
 

### 2020-03-05 paper_elsevier/paper_elsevier.Rmd
working on the tree overview

@@
 
-The `dfoliatR` library is unique among a growing suite `R` packages designed for dendrochronology.
-Stemming from the `dplR` library [@Bunn2008] that provides `R` the ability to read and write an array of tree-ring data formats, standaridize ring widths, build and evaluate chronologies, perform qaulity control (to name a few), one can now also measure, perform and check crossdating, and perform many analytical tests [@Bunn2010; @Jevsenak2018; @Lara2015; @Zang2015].
+The `dfoliatR` library adds to a growing suite `R` packages designed for dendrochronology.
+Stemming from the `dplR` library [@Bunn2008] that enables `R` users to read and write an array of tree-ring data formats, standaridize ring widths, build and evaluate chronologies, perform qaulity control (to name a few), one can now also measure ring widths [@Shi2019; @Lara2015], perform and check crossdating [@Bunn2010], and perform many analytical tests [@Jevsenak2018; @Zang2015].
 Tools for assessing stand dynamics and disturbance analyses are under rapid development, with new packages for assessing growth and release events [`TRADER`: @Altman2014], metrics of growth resilience [`pointRes`: @VanderMaaten-Theunissen2015], and fire history [`burnr`: @Malevich2018].
@@
 `dfoliatR` draws upon data formats in `dplR` that are commonly employed by other tree-ring libraries.
-It uses and outputs data formats that faciliate the use packages embodied by `tidyverse` [@Wickham2019] that include efficient manipulation of data [`dplyr`: @Wickham2020dplyr] and graphics [`ggplot2`: @wickham2016ggplot2].
+It uses and outputs data formats that faciliate the use packages embodied by `tidyverse` [@Wickham2019] that include efficient data manipulation [`dplyr`: @Wickham2020dplyr] and graphics [`ggplot2`: @wickham2016ggplot2].
 
@@
 
-Users can develop these data sets in software of their choosing, such as in `dplR` or ARSTAN.
+Users can develop these data sets in software of their choosing, such as `dplR` or ARSTAN.
 It is important that the host-tree data include only one tree-ring series per tree.
@@
 
-`dfoliatR` begins to identify defoliation events in individual trees by removing the climatic signal as represented by the non-host chronology from the host tree series.
-This indexing procedure creates a series -- the "growth suppression index" (GSI) -- in which disturbance is the predominant signal.
-The GSI is caluclated as
+At the heart of `dfoliatR` lies two functions: `defoliate_trees()` and `outbreak()`.
+These identify defoliation event son individual trees and thenc omposite across mutliple trees for a given site to identify outbreak events.
+It is up to the researcher to distinguish these separate spatial scales of analysis, `dfoliatR` then provides a set of functions to visualize and statistically summarize tree-level defoliation and site-level outbreak events, which will be discussed further.
+
+## Identifying Defoliation of Trees
+
+The `defoliate_trees()` function will often be the point of entry to the `dfoliatR` library.
+It performs two processes: removing the non-defoliatation growth signal from the host-tree series and identifying defoliation events for each host tree.
+The climatic or non-defoliation signals in each host-tree series is represented by the non-host chronology or a climate reconstruction.
+This indexing procedure generates the "growth suppression index" (GSI) in which disturbance is the predominant signal.
+The GSI is calculated as in @Swetnam1985 and @Swetnam1989,
 \begin{align}
@@
 \end{align}
-where H and NH are the host tree series and the non-host chronology, in year i, respectively [@Swetnam1985; @Swetnam1989].
-Only years in which the individual host-tree series and the non-host chronology overlap are used in Equation 1.
+where H and NH are the host-tree series and the non-host chronology, in year i, respectively.
+Only the common period between each host-tree series and the non-host chronology are used in Equation 1.
 The non-host chronology is scaled by its mean ($\overline{\textrm{NH}} \approx 1.0$) and multiplied by the ratio of host and non-host standard deviations ($\frac{\sigma_{\textrm{H}}}{\sigma_{\textrm{NH}}}$), which approximates the variance of the host tree series.
@@
 
-Negative departures in the normalized growth suppression index that surpass user-defined thresholds in duration and magnitude will be defined as *defoliation events*.
+Negative departures in the normalized GSI that surpass user-defined thresholds in duration and magnitude are defined as *defoliation events*.
+As in OUTBREAK, defoliation events have a single year of maximum departure (the variable 'max_reduction' in `defoliate_trees()`) that is used to assess the magnitude of growth suppression.
+The default setting is -1.28 (in units of standard deviation), which was previously determined as representative of WSBW effects (**cite**).
+Duration is assessed by examining sequences of negative GSI before and after the year of maximum departure.
+Each defoliation event is allowed one positive excersion on each side of the maximum departure year.
+Duration is assessed across the entire sequence that includes up to two positive excursions.
+The default duration is eight years, following OUTBREAK, as is commonly used in WSBW studies (**cite**).
+Defoliator insects vary in their durations of defoliations and outbreaks, so researchers should exmaine their data and perform sensitivity tests to to find a representative event duration.
+
+Researchers are given the options to extend defoliation events by bridging between successive events and allowing potentially short-duration events that occur at the end of the series.
+Bridging allows multiple events that are separated by a single year to be combined and considered as one defoliation event.
+We urge researchers to use this option carefully, as there is no setting to limit the number of potentially bridged events.
+Series end events is an option that can be used when the trees were sampled during a known defoliation event, which in many cases is what led the to the site being sampled for the reconstruction.
+This option eliminates the duration parameter, but all other thresholds, and bridging, still apply.
+These options are not available in OUTBREAK.
+
+## Identifying Outbreak Events
 
@@
 Users have options to define the number and/or the proportion of trees required for a defoliation event to be considered an outbreak.
-
-Note that these methods of separating tree- vs site-level disturbance categories is a major departure from the OUTBREAK program.
-In OUTBREAK the two levels of analysis are combined and users have more limited control of thresholds to define defoliation events versus outbreaks.
 
@@
 
-Usage
+Example Usage
 ==========================
@@
 
-##Site-level events
+## Site-level events
 

### 2020-03-05 paper_elsevier/paper_elsevier.tex
working on the tree overview

@@
 
-The \texttt{dfoliatR} library is unique among a growing suite \texttt{R}
+The \texttt{dfoliatR} library adds to a growing suite \texttt{R}
 packages designed for dendrochronology. Stemming from the \texttt{dplR}
-library (Bunn 2008) that provides \texttt{R} the ability to read and
-write an array of tree-ring data formats, standaridize ring widths,
-build and evaluate chronologies, perform qaulity control (to name a
-few), one can now also measure, perform and check crossdating, and
-perform many analytical tests (Bunn 2010, Lara et al. 2015, Zang and
-Biondi 2015, Jevšenak and Levanič 2018). Tools for assessing stand
-dynamics and disturbance analyses are under rapid development, with new
-packages for assessing growth and release events (\texttt{TRADER}:
-Altman et al. 2014), metrics of growth resilience (\texttt{pointRes}:
-Maaten-Theunissen et al. 2015), and fire history (\texttt{burnr}:
-Malevich et al. 2018). The key objective of \texttt{dfoliatR} is to
-provide tools to identify and analyse insect defoliation and outbreak
-events by building on the methods employed the FORTRAN program OUTBREAK
-(Swetnam and Lynch 1989). What sets \texttt{dfoliatR} apart from
-packages such as \texttt{TRADER} is that it explicitly performs an
-indexing procedue on host-tree series to remove climatic and other
-non-defoliation related signals represented by non-host tree species.
-Insect defoliation signals are identified in the disturbance index by
-the duration and magnitude of negative departures.
+library (Bunn 2008) that enables \texttt{R} users to read and write an
+array of tree-ring data formats, standaridize ring widths, build and
+evaluate chronologies, perform qaulity control (to name a few), one can
+now also measure ring widths (Lara et al. 2015, Shi and Xiang 2019),
+perform and check crossdating (Bunn 2010), and perform many analytical
+tests (Zang and Biondi 2015, Jevšenak and Levanič 2018). Tools for
+assessing stand dynamics and disturbance analyses are under rapid
+development, with new packages for assessing growth and release events
+(\texttt{TRADER}: Altman et al. 2014), metrics of growth resilience
+(\texttt{pointRes}: Maaten-Theunissen et al. 2015), and fire history
+(\texttt{burnr}: Malevich et al. 2018). The key objective of
+\texttt{dfoliatR} is to provide tools to identify and analyse insect
+defoliation and outbreak events by building on the methods employed the
+FORTRAN program OUTBREAK (Swetnam and Lynch 1989). What sets
+\texttt{dfoliatR} apart from packages such as \texttt{TRADER} is that it
+explicitly performs an indexing procedue on host-tree series to remove
+climatic and other non-defoliation related signals represented by
+non-host tree species. Insect defoliation signals are identified in the
+disturbance index by the duration and magnitude of negative departures.
 
 \texttt{dfoliatR} draws upon data formats in \texttt{dplR} that are
-commonly employed by other tree-ring libraries. It uses and outputs a
-data formats that faciliate the use packages embodied by
-\texttt{tidyverse} (Wickham et al. 2019) that include efficient data
-wrangling (\texttt{dplyr}: Wickham et al. 2020) and graphics
-(\texttt{ggplot2}: Wickham 2016).
+commonly employed by other tree-ring libraries. It uses and outputs data
+formats that faciliate the use packages embodied by \texttt{tidyverse}
+(Wickham et al. 2019) that include efficient data manipulation
+(\texttt{dplyr}: Wickham et al. 2020) and graphics (\texttt{ggplot2}:
+Wickham 2016).
 
@@
 Users can develop these data sets in software of their choosing, such as
-in \texttt{dplR} or ARSTAN. It is important that the host-tree data
-include only one tree-ring series per tree. Both \texttt{dplR} and
-ARSTAN have options for averaging mutliple sample series into a
-tree-level series. The tree-ring series and chronology can be read into
-\texttt{R} via several available \texttt{dplR} functions.
+\texttt{dplR} or ARSTAN. It is important that the host-tree data include
+only one tree-ring series per tree. Both \texttt{dplR} and ARSTAN have
+options for averaging mutliple sample series into a tree-level series.
+The tree-ring series and chronology can be read into \texttt{R} via
+several available \texttt{dplR} functions.
 
-\texttt{dfoliatR} begins to identify defoliation events in individual
-trees by removing the climatic signal as represented by the non-host
-chronology from the host tree series. This indexing procedure creates a
-series -- the ``growth suppression index'' (GSI) -- in which disturbance
-is the predominant signal. The GSI is caluclated as \begin{align}
+At the hear of \texttt{dfoliatR} lies two functions:
+\texttt{defoliate\_trees()} and \texttt{outbreak()}. These identify
+defoliation event son individual trees and thenc omposite across
+mutliple trees for a given site to identify outbreak events. It is up to
+the researcher to distinguish these separate spatial scales of analysis,
+\texttt{dfoliatR} then provides a set of functions to visualize and
+statistically summarize tree-level defoliation and site-level outbreak
+events, which will be discussed further.
+
+\hypertarget{identifying-defoliation-of-trees}{%
+\subsection{Identifying Defoliation of
+Trees}\label{identifying-defoliation-of-trees}}
+
+The \texttt{defoliate\_trees()} function will usually be used first. It
+performs two processes: removing the non-defoliatation growth signal
+from the host-tree series and identifying defoliation events for each
+host tree. The climatic or non-defoliation signals in each host-tree
+series is represented by the non-host chronology or a climate
+reconstruction. This indexing procedure generates the ``growth
+suppression index'' (GSI) in which disturbance is the predominant
+signal. The GSI is calculated as in Swetnam et al. (1985) and Swetnam
+and Lynch (1989), \begin{align}
 \textrm{GSI}_{i} = \textrm{H}_{i} - \left( \textrm{NH}_{i} - \overline{\textrm{NH}} \right) \frac{\sigma_{\textrm{H}}}{\sigma_{\textrm{NH}}} \
 \end{align} where H and NH are the host tree series and the non-host
-chronology, in year i, respectively (Swetnam et al. 1985, Swetnam and
-Lynch 1989). Only years in which the individual host-tree series and the
-non-host chronology overlap are used in Equation 1. The non-host
-chronology is scaled by its mean
+chronology, in year i, respectively. Only the common period between each
+host-tree series and the non-host chronology are used in Equation 1. The
+non-host chronology is scaled by its mean
 (\(\overline{\textrm{NH}} \approx 1.0\)) and multiplied by the ratio of
@@
 
-Negative departures in the normalized growth suppression index that
-surpass user-defined thresholds in duration and magnitude will be
-defined as \emph{defoliation events}.
+Negative departures in the normalized GSI that surpass user-defined
+thresholds in duration and magnitude are defined as \emph{defoliation
+events}. As in OUTBREAK, defoliation events have a single year of
+maximum departure (the variable `max\_reduction' in
+\texttt{defoliate\_trees()}) that is used to assess the magnitude of
+growth suppression. The default setting is -1.28 (in units of standard
+deviation), which was previously determined as representative of WSBW
+effects (\textbf{cite}). Duration is assessed by examining sequences of
+negative GSI before and after the year of maximum departure. Each
+defoliation event is allowed one positive excersion on each side of the
+maximum departure year. Duration is assessed across the entire sequence
+that includes up to two positive excursions. The default duration is
+eight years, following OUTBREAK, as is commonly used in WSBW studies
+(\textbf{cite}). Defoliator insects vary in their durations of
+defoliations and outbreaks, so researchers should exmaine their data and
+perform sensitivity tests to to find a representative event duration.
+
+Researchers are given the options to extend defoliation events by
+bridging between successive events and allowing potentially
+short-duration events that occur at the end of the series. Bridging
+allows multiple events that are separated by a single year to be
+combined and considered as one defolaition event. Each may have up to
+five years with positive growth (\textbf{CHECK}). Series end events is
+an option that can be used when the trees were sampled during a known
+defoliation event, which in many cases is what led the to the site being
+sampled for the reconstruction. This option eliminates the duration
+parameter, but all other thresholds, and bridging, still apply. These
+options are not available in OUTBREAK.
+
+\hypertarget{identifying-outbreak-events}{%
+\subsection{Identifying Outbreak
+Events}\label{identifying-outbreak-events}}
 
@@
 outbreak.
-
-Note that these methods of separating tree- vs site-level disturbance
-categories is a major departure from the OUTBREAK program. In OUTBREAK
-the two levels of analysis are combined and users have more limited
-control of thresholds to define defoliation events versus outbreaks.
 
@@
 
-\hypertarget{usage}{%
-\section{Usage}\label{usage}}
+\hypertarget{example-usage}{%
+\section{Example Usage}\label{example-usage}}
 
@@
 
-\#\#Site-level events
+\hypertarget{site-level-events}{%
+\subsection{Site-level events}\label{site-level-events}}
 
@@
 
+\leavevmode\hypertarget{ref-Shi2019}{}%
+Shi, J., and W. Xiang. 2019. MtreeRing: A shiny application for
+automatic measurements of tree-ring widths on digital images.
+
 \leavevmode\hypertarget{ref-Swetnam1989}{}%

### 2020-03-12 paper_elsevier/paper_elsevier.Rmd
2020-03-12 Edits/Additions following Ann's input

--- 2020-02-27 paper_elsevier/paper_elsevier.Rmd
@@
 abstract: |
-    We present a new R package to provide dendroecologists with tools to identify, quantify, analyze, and visualize growth
-    suppression events in tree rings produced by insect defoliation. The 'dfoliatR' library is based on the Fortran V program 
-    OUTBREAK, and builds on existing resources in the R computing environment. 'dfoliatR' expands on OUTBREAK to provide greater
-    control of supression thresholds, additional output tables, and high-quality graphics. To use 'dfoliatR' requires 
-    standardized ring-width measurements from insect host trees and an indexed tree-ring chronology from local non-host trees. 
-    It performs an indexing procedure to remove the climatic signal represented in the non-host chronology from the host-tree
-    series. It then infers defoliation events in individual trees. Site-level analyses identify outbreak events
-    that synchronously affect a user-defined number or proportion of the host trees. Functions are available for
-    summary statistics and graphics of tree- and site-level series.
+  We present a new R package to provide dendroecologists with tools to infer, quantify, analyze, and visualize growth suppression events in tree rings caused by insect defoliation.
+The ‘dfoliatR’ library is based on the FORTRAN V program OUTBREAK, and builds on existing resources in the R computing environment.
+‘dfoliatR’ expands on OUTBREAK to provide greater control of suppression thresholds, additional output tables, and high-quality graphics.
+To use ‘dfoliatR’ requires standardized ring-width measurements from insect host trees and an indexed tree-ring chronology from local non-host trees.
+It performs an indexing procedure to remove the climatic signal represented in the non-host chronology from the host-tree series.
+It then infers defoliation events in individual trees.
+Site-level analyses identify outbreak events that synchronously affect a user-defined number or proportion of the host trees.
+Functions are available for summary statistics and graphics of tree- and site-level series.
 
@@
 #numbersections: true
-csl: elsevier-harvard.csl
+csl: ecology.csl
 output: rticles::elsevier_article
 # output:
-#   bookdown::pdf_book:
-#     base_format: rticles::elsevier_article
 #   bookdown::word_document2:
@@
 
+Variation in the width and morphology of annual radial growth rings in wood permits dating and quantification of past forest insect defoliator outbreaks.
+Defoliation can be distinguished from climate- and other disturbance-related influences by comparing ring-width or other annually-resolved features in the wood of host species to non-host species or to climate records.
+The effect of defoliation on radial growth of trees has been recognized since the mid-1800s, but work prior to the 1980s was not cross-dated (based on annually-resolved and –dated measurements), not standardized (quantitatively controlled for tree geometry and age effects), and suffered from inadequate extraction of the defoliation-related variability from variability associated with climate and other factors [@Swetnam1985; @Speer2010; @Lynch2012].
+These and other problems may be circumvented by appropriate use of dendrochronological techniques.
+Quantitative methodology to infer forest defoliator outbreaks from cross-dated tree-ring records was developed in the 1980s by @Swetnam1985 for developing western spruce budworm chronologies [@Swetnam1989; @Swetnam1993].
+The methodology has since been successfully applied to a wide range of defoliator species, most of which are conifer herbivores, and has evolved in sophistication and application to a wide range of ecosystem situations [@Lynch2012].
+
+The main dendrochronological tool for inferring, dating, and quantifying defoliator outbreaks from tree-ring records has been the software routine OUTBREAK [@Swetnam1985; @outbreak; @Swetnam1989].
+OUTBREAK computes indices (described later in detail) of suppressed growth by subtracting a detrended and standardized climate series (a “control” chronology, usually a site chronology developed from non-host trees or a gridded climate data point series) from host individual-tree detrended and standardized radial growth series after the host and non-host series have been brought to a common variance.
+If the host and non-host species respond similarly to climate (which can be tested), the derived series retains variability that the host and non-host series do not have in common, generally the insect signal and noise (unexplained variability).
+The user defines a rule base specifying the magnitude and duration that a period of indexed growth suppression must meet or surpass for a period of suppressed growth to be inferred as a defoliation event at the tree level.
+
+Though powerful, OUTBREAK is outdated and increasingly difficult to use in modern computing environments.
+It was written in Fortran V with inherently severe restrictions, as RAM and disk space were limited at that time (256 kb and 10 MB, respectively) and Fortran conventions imposed very strict formatting, file naming, and output conventions.
+Windows-based execution operates in a DOS window, Apple-based xxx in xxx, and provides no graphical interface or capabilities.
+Barriers to batch operation impose burdens for analyses of larger data sets.
+We developed dfoliaR as an `R`- and `dplR`-based routine to overcome these issues.
+
+`dfoliatR` adds to a growing suite of dendrochronology packages the `R` computing environment [@RCore].
+Stemming from the `dplR` library [@Bunn2008] that enables `R` users to read and write an array of tree-ring data formats, standardize ring width series, build and evaluate chronologies, and perform quality control (to name a few), one can now also measure ring widths from scanned images of preparred samples [@Shi2019; @Lara2015], perform and check crossdating [@Bunn2010], and perform many analytical tests [@Jevsenak2018; @Zang2015].
+Tools for assessing stand dynamics and disturbance analyses are under rapid development, with new packages for assessing growth and release events [`TRADER`: @Altman2014], metrics of growth resilience [`pointRes`: @VanderMaaten-Theunissen2015], and fire history [`burnr`: @Malevich2018].
+The key objective of `dfoliatR` is to provide tools to identify and analyze insect defoliation and outbreak events by building on the methods employed by OUTBREAK.
+It capitilizes on the robust software already available in `R` by using `dplR` data formats for incoming tree-ring series and providing output data formats embodied by the `tidyverse` [@Wickham2019] that include efficient data manipulation [`dplyr`: @Wickham2020dplyr] and graphics [`ggplot2`: @wickham2016ggplot2].
+
+In this paper, we describe the statistical methods employed by `dfoliatR`, its availability, compare results to those produced by OUTBREAK, and present an example analyses.
+Users need not have much experience in `R` to replicate the analyses and graphics as presented.
+The `R` code presented below is executable in an `R` session once the required libraries are installed and loaded.
+Support documentation in addition to this paper is provided within the package via standard help menus (accessed by typing `?` before a function name) and on the package website ([https://chguiterman.github.io/dfoliatR/](https://chguiterman.github.io/dfoliatR/)), which includes up-to-date vignettes that describe in detail the functionality of the software.
+Code to create a preprint of this manuscript including the `R` scripts is available from [https://github.com/chguiterman/dfoliatR_paper](https://github.com/chguiterman/dfoliatR_paper).
+
 Overview of the software
@@
 
-The `dfoliatR` library requires two sets of tree-ring data to identify defoliation and outbreak events:
+The `dfoliatR` library requires two sets of tree-ring data to infer defoliation and outbreak events:
 
@@
 
-Users can develop these data sets in software of their choosing, such as in `dplR` or ARSTAN.
+Users can develop these data sets in software of their choosing, such as `dplR` [@Bunn2008] or ARSTAN [@arstan].
 It is important that the host-tree data include only one tree-ring series per tree.
-Both `dplR` and ARSTAN have options for averaging mutliple sample series into a tree-level series.
-The tree-ring series and chronology can be read into R via several available `dplR` functions.
+Both `dplR` and ARSTAN have options for averaging multiple sample series into a tree-level series.
 
-`dfoliatR` begins to identify defoliation events in individual trees by removing the climatic signal as represented by the non-host chronology from the host tree series.
-This indexing procedure creates a series -- the "growth suppression index" -- in which disturbance is the predominant signal.
-Negative departures in the growth suppression index that surpass user-defined thresholds in duration, and magnitude are defined as *defoliation events*.
-The individual tree defoliation series are composited in an additional step to identify *outbreak events* that synchronously affect mutliple trees.
-Users have options to define the number and/or the proportion of trees required for a defoliation event to be considered an outbreak.
+At the heart of `dfoliatR` lies two functions: `defoliate_trees()` and `outbreak()`.
+These identify defoliation events on individual trees and then composite across multiple trees to identify outbreak events.
 
-Note that these methods of separating tree- vs site-level disturbance categories is a major departure from the OUTBREAK program.
-In OUTBREAK the two levels of analysis are combined and users have more limited control of thresholds to define defoliation events versus outbreaks.
+## Identifying Defoliation of Trees
+
+The `defoliate_trees()` function is usually be the point of entry to the dfoliatR library.
+It performs two processes, removing climate-related growth signals from the host-tree series and then identifying tree-level defoliation events.
+The climatic or non-defoliation signals in each host-tree series are characterized by a non-host chronology or climate reconstruction.
+`dfoiatR` removes that non-defoliation signal by subtracting the non-host series from each host-tree series, which generates a residual index.
+In OUTBREAK, this residual index was termed the "corrected" index.
+We call it the "growth suppression index" (GSI).
+The GSI is calculated the same as in OUTBREAK [following @Swetnam1985; @Swetnam1989] for each host tree as
+\begin{align}
+\textrm{GSI}_{i} = \textrm{H}_{i} - \left( \textrm{NH}_{i} - \overline{\textrm{NH}} \right) \frac{\sigma_{\textrm{H}}}{\sigma_{\textrm{NH}}} \
+\end{align}
+where H and NH are the host-tree series and the non-host chronology, in year i, respectively.
+Only the common period between the host-tree series and the non-host chronology are used in Equation 1.
+The host and non-host chronologies are brought to common variance by scaling the non-host chronology by its mean ($\overline{\textrm{NH}} \approx 1.0$) and multiplying by the ratio of host and non-host standard deviations ($\frac{\sigma_{\textrm{H}}}{\sigma_{\textrm{NH}}}$), which approximates the variance of the host tree series.
+
+Negative departures in the normalized GSI that surpass user-defined thresholds in duration and magnitude are defined as *defoliation events*.
+As in OUTBREAK, magnitude is assessed on a single year within the departure sequence.
+The default setting is -1.28 (in units of standard deviation), which was previously determined to be representative of WSBW effects (**citation?**).
+Duration is assessed by examining sequences of negative GSI before and after the year of maximum departure.
+Each defoliation event is allowed one positive excursion on each side of the maximum departure year.
+Duration is assessed across the entire sequence that includes up to two positive excursions.
+The default duration is eight years, as is commonly used in WSBW studies (**citation?**).
+Different species of defoliation insects vary in the length of defoliation and the degree to which they can suppress tree growth.
+Researchers can, and should, adjust the duration and magnitude parameters accordingly and critically evaluate the results.
+
+Diverging from OUTBREAK, `dfoliatR` allows users to extend defoliation events by bridging successive events and also by allowing potentially short-duration events that occur at the end of the series.
+In cases where two defoliation events are separated by a single year, bridging will link them into a single event (**Show figure?**).
+We urge careful use of this option because there is no setting to limit the number or length of potentially bridged events.
+The series end option can be used in cases when the host trees were actively being defoliated at the time of sampling.
+This option eliminates the duration parameter for an event at the recent end of the series, but all other thresholds apply.
+The advantage of this parameter is that it can aid in identifying the start-year for the current defoliation event or outbreak, which is both useful in management and allows the current event to be included in return-interval estimates.
+
+## Inferring Outbreak Events
+
+Defoliation of one or a few trees does not constitute an outbreak.
+To determine when defoliation becomes an *outbreak event*, `dfoliatR` composites the individual tree defoliation series into a site-level chronology with the `outbreak()` function.
+Users have options to define the number and/or the proportion of trees required for an event to be considered an outbreak.
+Three parameters control the whether a defoliation event constitutes an outbreak: the minimum number of trees available, the minimum number of trees recording defoliation, and the percent of trees recording defoliation.
+The first allows the researcher to make a judgement call as to the confidence ascribed to reduced sample depth toward the ends of their chronologies, thus compensating for the "fading record problem" (**Swetnam and Fritts?**).
+The second two parameters adjust the scale of defoliation considered to be an outbreak.
+Absolute numbers of trees and percentages can be applied separately or in conjunction, following filtering conventions in tree ring fire history studies [@Malevich2018].
+We urge users to carefully consider the choice of absolute numbers in situations where the number of trees represented in the series varies with time, or the choice of percentages when sample size is small.
+
+Evaluation
+==========================
+
+## Approach
+
+We tested `dfoliatR` against OUTBREAK by comparing GSI to corrected indices for individual trees and years, outbreak status for individual trees and years, and percentage of trees recording outbreaks at the site level, using raw ring-width data from 8 sites in British Columbia, Colorado, and New Mexico and author-provided non-host site chronologies.
+
+We detrended host data for both dfoliatR and OUTBREAK using ARSTAN 6.1 (**downloaded 21 April 2002 from dpl; xxx Ann check the date, as 21 April 2002 is 6.05P**) with default double detrending (128 year wavelength and a 50% smoothing spline) and event thresholds of -1.28 normalized index and 8 years.
+Attempts to detrend with ARSTAN 44 in `dplR` were problematic for two reasons.
+Some versions of ARSTAN, including ARSTAN 44, do not include an option for producing tree-level averages from multiple cores from the same tree (.tre files).
+**More importantly, ring-width indices from ARSTAN 44 truncated the process at either end when sample size fell below 5 xxx (Jodi, what was the minimum?)**, and produced different values for years near the truncation point than did dpl-ARSTAN 6.05P.
+The majority of the detrended series values were identical, but not the tails, as the two routines do not use the same data to establish the tails.
+These differences produce different GSI (`dfoliatR`) and Corrected Indices (OUTBREAK) values at the tails, sometimes considerably, affecting whether or not outbreak events are inferred early or late in the tree series, timing of the first or last event onset, and subsequent computation of return intervals, duration, and periodicity.
+Note that there is nothing particularly important regarding the choice of ARSTAN 6.1 – it is simply the version that the second author used routinely and had available.
+
+## Findings
+
+`dfoliatR` and OUTBREAK compute growth suppresses indices and use them to infer identical tree-level outbreak events, including onset and termination dates and date and magnitude of the maximum growth suppression when `dfoliatR` bridging matches the OUTBREAK protocols.
+Once the issues with ARSTAN (described earlier) were resolved, dfoliatR and OUTBREAK produced identical indices at 0.000 precision.
+
+When inferring outbreak events at the tree level, our initial results did not always match OUTBREAK, leading to development of the bridging option.
+Identifying a minimum within a period while allowing for positive excursions is a tricky computation.
+OUTBREAK deliberately disallows back-to-back events – after a minimum value within a period of negative values, a second positive index terminates an event.
+OUTBREAK does not allow a period of highly negative values to be appended to an earlier outbreak if the intervening period is 1 year, even if those negative values fall below the minimum in the earlier period.
+Due in large part to reconstructions using OUTBREAK [see papers cited by @Lynch2012], we now know considerably more about forest defoliator outbreak regimes than we did in the 1980s when OUTBREAK was written.
+We think that two prolonged events separated by a single year should in some situations be considered a single outbreak at both the tree and site levels.
+This is particularly relevant to western spruce budworm and spruce budworm (*C.
+fumiferana*), for which regimes have been reconstructed for several geographic areas and for which researchers and forest health experts have gained considerable knowledge (**Sanders et al.
+1985, Brookes et al.
+1987 and many later publications**).
+We now know that the greatest growth suppression often occurs late in the outbreak due to cumulative effects (**rrr**).
+We developed the bridging option to permit these periods to be linked to preceding events separated by a single year.
+
+Bridging operates at the tree level in dfoliatR, and these effects generally have minor effects on site-level reconstructions (Fig.Bridging).
+Occasionally site-level events are inferred differently (Table.Bridging), and this can affect return interval and duration statistics.
+
+## Best Practice Recommendations
+
+We encourage authors to report software and library version numbers and download dates where applicable.
+This is standard practice for statistical packages, but has often been ignored for R- and dplR-based analyses.
+R-based software evolves over time, and results cannot always be replicated if the software versions are unknown.
+Similarly, significant changes (especially fixes) imply that re-analysis of data might differ from earlier published results and interpretations.
+Wherever possible, refereed publications supporting the software should be cited.
+Our experience with ARSTAN demonstrates the importance of adopting this practice.
+In the United States, analytic methods not documented thusly and supported by the literature might be questioned if the results are used in a National Environmental Protection Act (NEPA) process.
+
+We recommend that researchers compare dfoliatR results with and without bridging, and see which option agrees best with known insect biology and ecology.
+Bridging may be a logical choice for insects with prolonged non-outbreak periods, such as western spruce budworm, but not for situations where impacted stands barely recover from one outbreak before another begins, such as occurs with pine processionary caterpillars (*Thaumetopoea pityocampa* (Lepidoptera: Thaumetopoeidae)) (**Carus 2004, 2009**).
 
@@
 
-The `dfoliatR` is provided free and open source [@chris_guiterman_2020_3626136].
-It is provided to R users via the Comprehensive R Archive Network (CRAN; [https://cran.r-project.org/](https://cran.r-project.org/)).
-To install `dfoliatR` into R from CRAN use
+The `dfoliatR` library [@dfoliatR] is provided free and open source from the Comprehensive R Archive Network (CRAN; [https://cran.r-project.org/](https://cran.r-project.org/)).
+To install `dfoliatR` from CRAN use
 
@@
 
-Development versions of `dfoliatR` are available on Github and installed using the `devtools` library in R,
+Development versions of `dfoliatR` are available on GitHub and installed using the `devtools` library,
 
@@
 
-Support documentation in addition to this paper is provided within the package via help menus (accessed by typing `?` before a function name) and on the package website ([https://chguiterman.github.io/dfoliatR/](https://chguiterman.github.io/dfoliatR/)) which includes up-to-date vignettes.
-Code to execute a preprint of this manuscript and the code included below is available at (*insert later*).
+Issues, bug reports, and ideas for improving `dfoliatR` can be posted to [https://github.com/chguiterman/dfoliatR/issues](https://github.com/chguiterman/dfoliatR/issues).
+As an Open Source library, we welcome and encourage community involvement in future development.
+The best ways to contribute to `dfoliatR` are through standard GitHub procedures or by contacting the first author.
 
-Tree-level defoliation events
+Example Usage
 ==========================
 
-Site-level events
-==========================
+In `dfoliatR` we provide two sets of tree-ring data to aid users in exploring the functions, graphics, and outputs.
+Each set consists of Douglas-fir (*Pseudotsuga menziesii*) host-tree series, standardized with __________, and a local ponderosa pine (*Pinus ponderosa*) non-host chronology.
+The non-host ring-width data were standardized by ________ and the chronologies averaged following standard procedures [@Speer2010].
+Data from Demijohn Peak (DMJ; 2902 m asl), in the San Juan Mountains of southern Colorado, come from @Ryerson2003.
+Data from the East Fork site (EF; 2580 m asl) in the Jemez Mountains of northcentral New Mexico were presented by @Swetnam1993.
 
-Evaluation
-==========================
+## Tree-Level Defoliation Events
+
+## Site-Level Events
 
@@
 
+* Describe how dfoliatR can be combined with other R libraries
+  * Mapping, what else?
+
 References {#references .unnumbered}

### 2020-03-12 paper_elsevier/paper_elsevier.tex
2020-03-12 Edits/Additions following Ann's input

--- 2020-03-02 paper_elsevier/paper_elsevier.tex
@@
   We present a new R package to provide dendroecologists with tools to
-  identify, quantify, analyze, and visualize growth suppression events in
-  tree rings produced by insect defoliation. The `dfoliatR' library is
-  based on the Fortran V program OUTBREAK, and builds on existing
-  resources in the R computing environment. `dfoliatR' expands on OUTBREAK
-  to provide greater control of supression thresholds, additional output
-  tables, and high-quality graphics. To use `dfoliatR' requires
-  standardized ring-width measurements from insect host trees and an
-  indexed tree-ring chronology from local non-host trees. It performs an
-  indexing procedure to remove the climatic signal represented in the
-  non-host chronology from the host-tree series. It then infers
-  defoliation events in individual trees. Site-level analyses identify
-  outbreak events that synchronously affect a user-defined number or
-  proportion of the host trees. Functions are available for summary
-  statistics and graphics of tree- and site-level series.
+  infer, quantify, analyze, and visualize growth suppression events in
+  tree rings caused by insect defoliation. The `dfoliatR' library is based
+  on the FORTRAN V program OUTBREAK, and builds on existing resources in
+  the R computing environment. `dfoliatR' expands on OUTBREAK to provide
+  greater control of suppression thresholds, additional output tables, and
+  high-quality graphics. To use `dfoliatR' requires standardized
+  ring-width measurements from insect host trees and an indexed tree-ring
+  chronology from local non-host trees. It performs an indexing procedure
+  to remove the climatic signal represented in the non-host chronology
+  from the host-tree series. It then infers defoliation events in
+  individual trees. Site-level analyses identify outbreak events that
+  synchronously affect a user-defined number or proportion of the host
+  trees. Functions are available for summary statistics and graphics of
+  tree- and site-level series.
 
@@
 
-The \texttt{dfoliatR} library is unique among a growing suite \texttt{R}
-packages designed for dendrochronology. Stemming from the \texttt{dplR}
-library (Bunn 2008) that provides \texttt{R} the ability to read and
-write an array of tree-ring data formats, standaridize ring widths,
-build and evaluate chronologies, perform qaulity control (to name a
-few), one can now also measure, perform and check crossdating, and
-perform many analytical tests (Bunn 2010, Lara et al. 2015, Zang and
-Biondi 2015, Jevšenak and Levanič 2018). Tools for assessing stand
-dynamics and disturbance analyses are under rapid development, with new
-packages for assessing growth and release events (\texttt{TRADER}:
-Altman et al. 2014), metrics of growth resilience (\texttt{pointRes}:
-Maaten-Theunissen et al. 2015), and fire history (\texttt{burnr}:
-Malevich et al. 2018). The key objective of \texttt{dfoliatR} is to
-provide tools to identify and analyse insect defoliation and outbreak
-events by building on the methods employed the FORTRAN program OUTBREAK
-(Swetnam and Lynch 1989). What sets \texttt{dfoliatR} apart from
-packages such as \texttt{TRADER} is that it explicitly performs an
-indexing procedue on host-tree series to remove climatic and other
-non-defoliation related signals represented by non-host tree species.
-Insect defoliation signals are identified in the disturbance index by
-the duration and magnitude of negative departures.
+Variation in the width and morphology of annual radial growth rings in
+wood permits dating and quantification of past forest insect defoliator
+outbreaks. Defoliation can be distinguished from climate- and other
+disturbance-related influences by comparing ring-width or other
+annually-resolved features in the wood of host species to non-host
+species or to climate records. The effect of defoliation on radial
+growth of trees has been recognized since the mid-1800s, but work prior
+to the 1980s was not cross-dated (based on annually-resolved and --dated
+measurements), not standardized (quantitatively controlled for tree
+geometry and age effects), and suffered from inadequate extraction of
+the defoliation-related variability from variability associated with
+climate and other factors (Swetnam et al. 1985, Speer 2010, Lynch 2012).
+These and other problems may be circumvented by appropriate use of
+dendrochronological techniques. Quantitative methodology to infer forest
+defoliator outbreaks from cross-dated tree-ring records was developed in
+the 1980s by Swetnam et al. (1985) for developing western spruce budworm
+chronologies (Swetnam and Lynch 1989, 1993). The methodology has since
+been successfully applied to a wide range of defoliator species, most of
+which are conifer herbivores, and has evolved in sophistication and
+application to a wide range of ecosystem situations (Lynch 2012).
 
-\texttt{dfoliatR} draws upon data formats in \texttt{dplR} that are
-commonly employed by other tree-ring libraries. It uses and outputs a
-data formats that faciliate the use packages embodied by
+The main dendrochronological tool for inferring, dating, and quantifying
+defoliator outbreaks from tree-ring records has been the software
+routine OUTBREAK (Swetnam et al. 1985, Holmes and Swetnam 1986, Swetnam
+and Lynch 1989). OUTBREAK computes indices (described later in detail)
+of suppressed growth by subtracting a detrended and standardized climate
+series (a ``control'' chronology, usually a site chronology developed
+from non-host trees or a gridded climate data point series) from host
+individual-tree detrended and standardized radial growth series after
+the host and non-host series have been brought to a common variance. If
+the host and non-host species respond similarly to climate (which can be
+tested), the derived series retains variability that the host and
+non-host series do not have in common, generally the insect signal and
+noise (unexplained variability). The user defines a rule base specifying
+the magnitude and duration that a period of indexed growth suppression
+must meet or surpass for a period of suppressed growth to be inferred as
+a defoliation event at the tree level.
+
+Though powerful, OUTBREAK is outdated and increasingly difficult to use
+in modern computing environments. It was written in Fortran V with
+inherently severe restrictions, as RAM and disk space were limited at
+that time (256 kb and 10 MB, respectively) and Fortran conventions
+imposed very strict formatting, file naming, and output conventions.
+Windows-based execution operates in a DOS window, Apple-based xxx in
+xxx, and provides no graphical interface or capabilities. Barriers to
+batch operation impose burdens for analyses of larger data sets. We
+developed dfoliaR as an \texttt{R}- and \texttt{dplR}-based routine to
+overcome these issues.
+
+\texttt{dfoliatR} adds to a growing suite of dendrochronology packages
+the \texttt{R} computing environment (R Core Team 2019). Stemming from
+the \texttt{dplR} library (Bunn 2008) that enables \texttt{R} users to
+read and write an array of tree-ring data formats, standardize ring
+width series, build and evaluate chronologies, and perform quality
+control (to name a few), one can now also measure ring widths from
+scanned images of preparred samples (Lara et al. 2015, Shi and Xiang
+2019), perform and check crossdating (Bunn 2010), and perform many
+analytical tests (Zang and Biondi 2015, Jevšenak and Levanič 2018).
+Tools for assessing stand dynamics and disturbance analyses are under
+rapid development, with new packages for assessing growth and release
+events (\texttt{TRADER}: Altman et al. 2014), metrics of growth
+resilience (\texttt{pointRes}: Maaten-Theunissen et al. 2015), and fire
+history (\texttt{burnr}: Malevich et al. 2018). The key objective of
+\texttt{dfoliatR} is to provide tools to identify and analyze insect
+defoliation and outbreak events by building on the methods employed by
+OUTBREAK. It capitilizes on the robust software already available in
+\texttt{R} by using \texttt{dplR} data formats for incoming tree-ring
+series and providing output data formats embodied by the
 \texttt{tidyverse} (Wickham et al. 2019) that include efficient data
-wrangling (\texttt{dplyr}: Wickham et al. 2020) and graphics
+manipulation (\texttt{dplyr}: Wickham et al. 2020) and graphics
 (\texttt{ggplot2}: Wickham 2016).
@@
 In this paper, we describe the statistical methods employed by
-\texttt{dfoliatR}, its availability, and run through analyses for a
-signle site in New Mexico. Users need not have much experience in
-\texttt{R} to replicate the analyses and graphics below. All \texttt{R}
-code presented below is executatable in an \texttt{R} session once the
-required libraries are installed and loaded. Support documentation in
-addition to this paper is provided within the package via standard help
-menus (accessed by typing \texttt{?} before a function name) and on the
-package website (\url{https://chguiterman.github.io/dfoliatR/}), which
-includes up-to-date vignettes that describe in detail the functionality
-of the software. Code to create a preprint of this manuscript including
-the \texttt{R} scripts is available from
+\texttt{dfoliatR}, its availability, compare results to those produced
+by OUTBREAK, and present an example analyses. Users need not have much
+experience in \texttt{R} to replicate the analyses and graphics as
+presented. The \texttt{R} code presented below is executable in an
+\texttt{R} session once the required libraries are installed and loaded.
+Support documentation in addition to this paper is provided within the
+package via standard help menus (accessed by typing \texttt{?} before a
+function name) and on the package website
+(\url{https://chguiterman.github.io/dfoliatR/}), which includes
+up-to-date vignettes that describe in detail the functionality of the
+software. Code to create a preprint of this manuscript including the
+\texttt{R} scripts is available from
 \url{https://github.com/chguiterman/dfoliatR_paper}.
@@
 The \texttt{dfoliatR} library requires two sets of tree-ring data to
-identify defoliation and outbreak events:
+infer defoliation and outbreak events:
 
@@
 Users can develop these data sets in software of their choosing, such as
-in \texttt{dplR} or ARSTAN. It is important that the host-tree data
-include only one tree-ring series per tree. Both \texttt{dplR} and
-ARSTAN have options for averaging mutliple sample series into a
-tree-level series. The tree-ring series and chronology can be read into
-\texttt{R} via several available \texttt{dplR} functions.
+\texttt{dplR} (Bunn 2008) or ARSTAN (Cook and Holmes 1996). It is
+important that the host-tree data include only one tree-ring series per
+tree. Both \texttt{dplR} and ARSTAN have options for averaging multiple
+sample series into a tree-level series.
 
-\texttt{dfoliatR} begins to identify defoliation events in individual
-trees by removing the climatic signal as represented by the non-host
-chronology from the host tree series. This indexing procedure creates a
-series -- the ``growth suppression index'' (GSI) -- in which disturbance
-is the predominant signal. The GSI is caluclated as \begin{align}
+At the heart of \texttt{dfoliatR} lies two functions:
+\texttt{defoliate\_trees()} and \texttt{outbreak()}. These identify
+defoliation events on individual trees and then composite across
+multiple trees to identify outbreak events.
+
+\hypertarget{identifying-defoliation-of-trees}{%
+\subsection{Identifying Defoliation of
+Trees}\label{identifying-defoliation-of-trees}}
+
+The \texttt{defoliate\_trees()} function is usually be the point of
+entry to the dfoliatR library. It performs two processes, removing
+climate-related growth signals from the host-tree series and then
+identifying tree-level defoliation events. The climatic or
+non-defoliation signals in each host-tree series are characterized by a
+non-host chronology or climate reconstruction. \texttt{dfoiatR} removes
+that non-defoliation signal by subtracting the non-host series from each
+host-tree series, which generates a residual index. In OUTBREAK, this
+residual index was termed the ``corrected'' index. We call it the
+``growth suppression index'' (GSI). The GSI is calculated the same as in
+OUTBREAK (following Swetnam et al. 1985, Swetnam and Lynch 1989) for
+each host tree as \begin{align}
 \textrm{GSI}_{i} = \textrm{H}_{i} - \left( \textrm{NH}_{i} - \overline{\textrm{NH}} \right) \frac{\sigma_{\textrm{H}}}{\sigma_{\textrm{NH}}} \
-\end{align} where H and NH are the host tree series and the non-host
-chronology, in year i, respectively (Swetnam et al. 1985, Swetnam and
-Lynch 1989). Only years in which the individual host-tree series and the
-non-host chronology overlap are used in Equation 1. The non-host
-chronology is scaled by its mean
-(\(\overline{\textrm{NH}} \approx 1.0\)) and multiplied by the ratio of
+\end{align} where H and NH are the host-tree series and the non-host
+chronology, in year i, respectively. Only the common period between the
+host-tree series and the non-host chronology are used in Equation 1. The
+host and non-host chronologies are brought to common variance by scaling
+the non-host chronology by its mean
+(\(\overline{\textrm{NH}} \approx 1.0\)) and multiplying by the ratio of
 host and non-host standard deviations
 (\(\frac{\sigma_{\textrm{H}}}{\sigma_{\textrm{NH}}}\)), which
-approximates the variance of the host tree series. This ``corrected''
-non-host chronology is subtracted from the host-tree series.
+approximates the variance of the host tree series.
 
-Negative departures in the normalized growth suppression index that
-surpass user-defined thresholds in duration and magnitude will be
-defined as \emph{defoliation events}.
+Negative departures in the normalized GSI that surpass user-defined
+thresholds in duration and magnitude are defined as \emph{defoliation
+events}. As in OUTBREAK, magnitude is assessed on a single year within
+the departure sequence. The default setting is -1.28 (in units of
+standard deviation), which was previously determined to be
+representative of WSBW effects (\textbf{citation?}). Duration is
+assessed by examining sequences of negative GSI before and after the
+year of maximum departure. Each defoliation event is allowed one
+positive excursion on each side of the maximum departure year. Duration
+is assessed across the entire sequence that includes up to two positive
+excursions. The default duration is eight years, as is commonly used in
+WSBW studies (\textbf{citation?}). Different species of defoliation
+insects vary in the length of defoliation and the degree to which they
+can suppress tree growth. Researchers can, and should, adjust the
+duration and magnitude parameters accordingly and critically evaluate
+the results.
 
-The individual tree defoliation series are composited in an additional
-step to identify \emph{outbreak events} that synchronously affect
-mutliple trees. Users have options to define the number and/or the
-proportion of trees required for a defoliation event to be considered an
-outbreak.
+Diverging from OUTBREAK, \texttt{dfoliatR} allows users to extend
+defoliation events by bridging successive events and also by allowing
+potentially short-duration events that occur at the end of the series.
+In cases where two defoliation events are separated by a single year,
+bridging will link them into a single event (\textbf{Show figure?}). We
+urge careful use of this option because there is no setting to limit the
+number or length of potentially bridged events. The series end option
+can be used in cases when the host trees were actively being defoliated
+at the time of sampling. This option eliminates the duration parameter
+for an event at the recent end of the series, but all other thresholds
+apply. The advantage of this parameter is that it can aid in identifying
+the start-year for the current defoliation event or outbreak, which is
+both useful in management and allows the current event to be included in
+return-interval estimates.
 
-Note that these methods of separating tree- vs site-level disturbance
-categories is a major departure from the OUTBREAK program. In OUTBREAK
-the two levels of analysis are combined and users have more limited
-control of thresholds to define defoliation events versus outbreaks.
+\hypertarget{inferring-outbreak-events}{%
+\subsection{Inferring Outbreak Events}\label{inferring-outbreak-events}}
+
+Defoliation of one or a few trees does not constitute an outbreak. To
+determine when defoliation becomes an \emph{outbreak event},
+\texttt{dfoliatR} composites the individual tree defoliation series into
+a site-level chronology with the \texttt{outbreak()} function. Users
+have options to define the number and/or the proportion of trees
+required for an event to be considered an outbreak. Three parameters
+control the whether a defoliation event constitutes an outbreak: the
+minimum number of trees available, the minimum number of trees recording
+defoliation, and the percent of trees recording defoliation. The first
+allows the researcher to make a judgement call as to the confidence
+ascribed to reduced sample depth toward the ends of their chronologies,
+thus compensating for the ``fading record problem'' (\textbf{Swetnam and
+Fritts?}). The second two parameters adjust the scale of defoliation
+considered to be an outbreak. Absolute numbers of trees and percentages
+can be applied separately or in conjunction, following filtering
+conventions in tree ring fire history studies (Malevich et al. 2018). We
+urge users to carefully consider the choice of absolute numbers in
+situations where the number of trees represented in the series varies
+with time, or the choice of percentages when sample size is small.
+
+\hypertarget{evaluation}{%
+\section{Evaluation}\label{evaluation}}
+
+\hypertarget{approach}{%
+\subsection{Approach}\label{approach}}
+
+We tested \texttt{dfoliatR} against OUTBREAK by comparing GSI to
+corrected indices for individual trees and years, outbreak status for
+individual trees and years, and percentage of trees recording outbreaks
+at the site level, using raw ring-width data from 8 sites in British
+Columbia, Colorado, and New Mexico and author-provided non-host site
+chronologies.
+
+We detrended host data for both dfoliatR and OUTBREAK using ARSTAN 6.1
+(\textbf{downloaded 21 April 2002 from dpl; xxx Ann check the date, as
+21 April 2002 is 6.05P}) with default double detrending (128 year
+wavelength and a 50\% smoothing spline) and event thresholds of -1.28
+normalized index and 8 years. Attempts to detrend with ARSTAN 44 in
+\texttt{dplR} were problematic for two reasons. Some versions of ARSTAN,
+including ARSTAN 44, do not include an option for producing tree-level
+averages from multiple cores from the same tree (.tre files). More
+importantly, ring-width indices from ARSTAN 44 truncated the process at
+either end when sample size fell below 5 xxx (Jodi, what was the
+minimum?), and produced different values for years near the truncation
+point than did dpl-ARSTAN 6.05P. The majority of the detrended series
+values were identical, but not the tails, as the two routines do not use
+the same data to establish the tails. These differences produce
+different GSI (\texttt{dfoliatR}) and Corrected Indices (OUTBREAK)
+values at the tails, sometimes considerably, affecting whether or not
+outbreak events are inferred early or late in the tree series, timing of
+the first or last event onset, and subsequent computation of return
+intervals, duration, and periodicity. Note that there is nothing
+particularly important regarding the choice of ARSTAN 6.1 -- it is
+simply the version that the second author used routinely and had
+available.
+
+\hypertarget{findings}{%
+\subsection{Findings}\label{findings}}
+
+\texttt{dfoliatR} and OUTBREAK compute growth suppresses indices and use
+them to infer identical tree-level outbreak events, including onset and
+termination dates and date and magnitude of the maximum growth
+suppression when \texttt{dfoliatR} bridging matches the OUTBREAK
+protocols. Once the issues with ARSTAN (described earlier) were
+resolved, dfoliatR and OUTBREAK produced identical indices at 0.000
+precision.
+
+When inferring outbreak events at the tree level, our initial results
+did not always match OUTBREAK, leading to development of the bridging
+option. Identifying a minimum within a period while allowing for
+positive excursions is a tricky computation. OUTBREAK deliberately
+disallows back-to-back events -- after a minimum value within a period
+of negative values, a second positive index terminates an event.
+OUTBREAK does not allow a period of highly negative values to be
+appended to an earlier outbreak if the intervening period is 1 year,
+even if those negative values fall below the minimum in the earlier
+period. Due in large part to reconstructions using OUTBREAK (see papers
+cited by Lynch 2012), we now know considerably more about forest
+defoliator outbreak regimes than we did in the 1980s when OUTBREAK was
+written. We think that two prolonged events separated by a single year
+should in some situations be considered a single outbreak at both the
+tree and site levels. This is particularly relevant to western spruce
+budworm and spruce budworm (\emph{C. fumiferana}), for which regimes
+have been reconstructed for several geographic areas and for which
+researchers and forest health experts have gained considerable knowledge
+(\textbf{Sanders et al.~1985, Brookes et al.~1987 and many later
+publications}). We now know that the greatest growth suppression often
+occurs late in the outbreak due to cumulative effects (\textbf{rrr}). We
+developed the bridging option to permit these periods to be linked to
+preceding events separated by a single year.
+
+Bridging operates at the tree level in dfoliatR, and these effects
+generally have minor effects on site-level reconstructions
+(Fig.Bridging). Occasionally site-level events are inferred differently
+(Table.Bridging), and this can affect return interval and duration
+statistics.
+
+\hypertarget{best-practice-recommendations}{%
+\subsection{Best Practice
+Recommendations}\label{best-practice-recommendations}}
+
+We encourage authors to report software and library version numbers and
+download dates where applicable. This is standard practice for
+statistical packages, but has often been ignored for R- and dplR-based
+analyses. R-based software evolves over time, and results cannot always
+be replicated if the software versions are unknown. Similarly,
+significant changes (especially fixes) imply that re-analysis of data
+might differ from earlier published results and interpretations.
+Wherever possible, refereed publications supporting the software should
+be cited. Our experience with ARSTAN demonstrates the importance of
+adopting this practice. In the United States, analytic methods not
+documented thusly and supported by the literature might be questioned if
+the results are used in a National Environmental Protection Act (NEPA)
+process.
+
+We recommend that researchers compare dfoliatR results with and without
+bridging, and see which option agrees best with known insect biology and
+ecology. Bridging may be a logical choice for insects with prolonged
+non-outbreak periods, such as western spruce budworm, but not for
+situations where impacted stands barely recover from one outbreak before
+another begins, such as occurs with pine processionary caterpillars
+(\emph{Thaumetopoea pityocampa} (Lepidoptera: Thaumetopoeidae))
+(\textbf{Carus 2004, 2009}).
 
@@
 
-The \texttt{dfoliatR} is provided free and open source (Guiterman 2020).
-It is provided to \texttt{R} users via the Comprehensive R Archive
-Network (CRAN; \url{https://cran.r-project.org/}). To install
-\texttt{dfoliatR} from CRAN use
+The \texttt{dfoliatR} library (Guiterman et al. 2020) is provided free
+and open source from the Comprehensive R Archive Network (CRAN;
+\url{https://cran.r-project.org/}). To install \texttt{dfoliatR} from
+CRAN use
 
@@
 
-Development versions of \texttt{dfoliatR} are available on Github and
+Development versions of \texttt{dfoliatR} are available on GitHub and
 installed using the \texttt{devtools} library,
@@
 future development. The best ways to contribute to \texttt{dfoliatR} are
-through standard Github procedures or by contacting the first author.
+through standard GitHub procedures or by contacting the first author.
 
-\hypertarget{usage}{%
-\section{Usage}\label{usage}}
+\hypertarget{example-usage}{%
+\section{Example Usage}\label{example-usage}}
+
+In \texttt{dfoliatR} we provide two sets of tree-ring data to aid users
+in exploring the functions, graphics, and outputs. Each set consists of
+Douglas-fir (\emph{Pseudotsuga menziesii}) host-tree series,
+standardized with \_\_\_\_\_\_\_\_\_\_, and a local ponderosa pine
+(\emph{Pinus ponderosa}) non-host chronology. The non-host ring-width
+data were standardized by \_\_\_\_\_\_\_\_ and the chronologies averaged
+following standard procedures (Speer 2010). Data from Demijohn Peak
+(DMJ; 2902 m asl), in the San Juan Mountains of southern Colorado, come
+from Ryerson et al. (2003). Data from the East Fork site (EF; 2580 m
+asl) in the Jemez Mountains of northcentral New Mexico were presented by
+Swetnam and Lynch (1993).
 
 \hypertarget{tree-level-defoliation-events}{%
-\subsection{Tree-level defoliation
-events}\label{tree-level-defoliation-events}}
+\subsection{Tree-Level Defoliation
+Events}\label{tree-level-defoliation-events}}
 
-\#\#Site-level events
-
-\hypertarget{evaluation}{%
-\section{Evaluation}\label{evaluation}}
+\hypertarget{site-level-events}{%
+\subsection{Site-Level Events}\label{site-level-events}}
 
@@
 \section{Extensions}\label{extensions}}
+
+\begin{itemize}
+\tightlist
+\item
+  Describe how dfoliatR can be combined with other R libraries
+
+  \begin{itemize}
+  \tightlist
+  \item
+    Mapping, what else?
+  \end{itemize}
+\end{itemize}
 
@@
 
-\leavevmode\hypertarget{ref-chris_guiterman_2020_3626136}{}%
-Guiterman, C. 2020. Chguiterman/dfoliatR: First release. Zenodo.
+\leavevmode\hypertarget{ref-arstan}{}%
+Cook, E. R., and R. L. Holmes. 1996. Guide for computer program arstan.
+Pages 75--87 The international tree-ring data bank program library
+version.
+
+\leavevmode\hypertarget{ref-dfoliatR}{}%
+Guiterman, C., A. Lynch, and J. Axelson. 2020. DfoliatR: Detection and
+analysis of insect defoliation signals in tree rings.
+
+\leavevmode\hypertarget{ref-outbreak}{}%
+Holmes, R. L., and T. W. Swetnam. 1986. Dendroecology program library:
+Program outbreak user's manual. Laboratory of Tree-Ring Research,
+University of Arizona, Tucson.
 
@@
 
+\leavevmode\hypertarget{ref-Lynch2012}{}%
+Lynch, A. M. 2012. What Tree-Ring Reconstruction Tells Us about Conifer
+Defoliator Outbreaks. Pages 126--154 \emph{in} P. Barbosa, D. K.
+Letourneau, and A. A. Agrawal, editors. Insect outbreaks revisited.
+Blackwell Publishing Ltd.
+
 \leavevmode\hypertarget{ref-VanderMaaten-Theunissen2015}{}%
@@
 
+\leavevmode\hypertarget{ref-RCore}{}%
+R Core Team. 2019. R: A language and environment for statistical
+computing. R Foundation for Statistical Computing, Vienna, Austria.
+
+\leavevmode\hypertarget{ref-Ryerson2003}{}%
+Ryerson, D. E., T. W. Swetnam, and A. M. Lynch. 2003. A tree-ring
+reconstruction of western spruce budworm outbreaks in the San Juan
+Mountains, Colorado, U.S.A. Canadian Journal of Forest Research
+33:1010--1028.
+
+\leavevmode\hypertarget{ref-Shi2019}{}%
+Shi, J., and W. Xiang. 2019. MtreeRing: A shiny application for
+automatic measurements of tree-ring widths on digital images.
+
+\leavevmode\hypertarget{ref-Speer2010}{}%
+Speer, J. H. 2010. Fundamentals of tree-ring research. Page 333. The
+University of Arizona Press.
+
 \leavevmode\hypertarget{ref-Swetnam1989}{}%
@@
 Science 35:962--986.
+
+\leavevmode\hypertarget{ref-Swetnam1993}{}%
+Swetnam, T. W., and A. M. Lynch. 1993. Multicentury, Regional-Scale
+Patterns of Western Spruce Budworm Outbreaks. Ecological Monographs
+63:399--424.
 

### 2020-04-09 paper_elsevier/R/Compare_OUTBREAK_dfoliatR.R
comparative analysis

@@ -1,0 +1,175 @@
+# Compare results from OUTBREAK to dfoliatR
+
+library(dfoliatR)
+library(dplR)
+library(dplyr)
+library(purrr)
+library(tibble)
+library(tidyr)
+library(ggplot2)
+library(cowplot)
+# Load read_OUTBREAK function
+devtools::source_gist("https://gist.github.com/chguiterman/c88e9601dd085eb4215f0dc133a0e21c", sha1 = "fdb3cfefcd7e6ed3f7df3e20057c815bbd91dd0c")
+
+data_pth <- "paper_elsevier/Data/"
+
+## Master metadata file
+meta_in <- read.csv(paste0(data_pth, "Meta.csv"))
+
+## Helper function
+rm_sample_depth <- function(x){
+  if (any(names(x) %in% "samp.depth"))
+    df <- select(x, - samp.depth)
+  else df <- x
+  return(df)
+}
+
+
+# OUTBREAK data -----------------------------------------------------------
+
+outbr_results <- meta_in %>% 
+  transmute(site, 
+            outbr_cor = map(outbr_corrected, ~ read.compact(paste0(data_pth, .))),
+            outbr_col = map(outbr_site, ~ read.delim(paste0(data_pth, .))),
+            outbr_out = map2(outbr_file, insect_code, 
+                             ~ read_OUTREAK(fname = paste0(data_pth, .x),
+                                            level = "tree",
+                                            insect_code = .y)))
+
+# Tree-level table
+outbr_defol <- outbr_results %>% 
+  transmute(site,
+            outbr_cor = map(outbr_cor, ~ .[] %>%  
+                              select(! ends_with("NormCorr")) %>%
+                              rownames_to_column("year") %>% 
+                              pivot_longer(-year, names_to = "series", values_to = "ncor"))
+  ) %>% 
+  unnest(cols = outbr_cor) %>% 
+  na.omit() %>% 
+  arrange(series, year)
+
+outbr_out <- outbr_results %>% 
+  select(site, outbr_out) %>% 
+  unnest(cols = outbr_out)
+
+outbr_defol <- merge(outbr_defol, outbr_out, by = c("site", "series", "year"))
+
+# Site-level table
+outbr_obr <- outbr_results %>% 
+  transmute(site,
+            outbr_col = map(outbr_col, ~ .[] %>% 
+                              select(Year, 
+                                     contains("Trees", ignore.case = TRUE), 
[…]

### 2020-04-09 paper_elsevier/R/Example_Usage.R

@@ -1,0 +1,52 @@
+## Create tables and graphics for dfoliatR manuscript
+
+library(dfoliatR)
+library(ggplot2)
+library(patchwork)
+
+out_pth <- "paper_elsevier/Output/" 
+# Tree-level outputs ------------------------------------------------------
+
+data("dmj_h")
+data("dmj_nh")
+
+dmj_defol <- defoliate_trees(host_tree = dmj_h,
+                             nonhost_chron = dmj_nh,
+                             duration_years = 8,
+                             max_reduction = -1.28,
+                             bridge_events = TRUE,
+                             series_end_event = TRUE,
+                             list_output = FALSE)
+
+dmj_stats <- defol_stats(dmj_defol)
+write.csv(dmj_stats, paste0(out_pth, "defol-stats.csv"), row.names = FALSE)
+
+plot_defol(dmj_defol)
+ggsave(paste0(out_pth, "tree-plot-default.pdf"), width=5.75, dpi=300, units="in")
+
+p1 <- plot_defol(dmj_defol) +
+  theme(legend.position = c(0.5, -0.15),
+        legend.direction = "horizontal",
+        legend.background = element_blank(),
+        legend.text = element_text(size=8),
+        plot.margin = unit(c(0, 0, 15, 0), units = "points"),
+        plot.background = element_blank())
+  
+p2 <- p1 + 
+  scale_color_manual(values = c("red", "orange", "purple")) 
+
+p1 / p2 + plot_layout(guides = "keep") + plot_annotation(tag_levels = "A")
+ggsave(paste0(out_pth, "tree-plot-2panel.pdf"), width=5.75, dpi=300, units="in")
+
+# Site-level outputs ------------------------------------------------------
+
+dmj_obr <- outbreak(dmj_defol,
+                    filter_min_series = 3,
+                    filter_min_defol = 1,
+                    filter_perc = 25)
+
+plot_outbreak(dmj_obr, disp_index = "mean_gsi")
+ggsave(paste0(out_pth, "site-plot.pdf"), width=5.75, dpi=300, units="in")
+
+dmj_obr_stats <- outbreak_stats(dmj_obr)
+write.csv(dmj_obr_stats, paste0(out_pth, "obr-stats.csv"), row.names = FALSE)

### 2020-04-09 paper_elsevier/R/debug.R

@@ -1,0 +1,24 @@
+
+df_run <- meta_in %>% 
+  transmute(site,
+            host = map(host_file, ~ read.compact(paste0(data_pth, .))),
+            nonhost = map(nonhost_file, ~ read.crn(paste0(data_pth, .)) %>% 
+                            rm_sample_depth()),
+            defol = pmap(list(host, nonhost, ongoing_defol),
+                         ~ defoliate_trees(host_tree = ..1, 
+                                           nonhost_chron = ..2,
+                                           series_end_event = ..3,
+                                           duration_years = 8,
+                                           max_reduction = -1.28,
+                                           bridge_events = FALSE,
+                                           list_output = TRUE))
+  )
+
+fnl <- df_run$defol[[4]]           
+input_series <- fnl[[11]] 
+
+dmj <- df_run$defol[[1]]
+input_series <- dmj[[12]]
+
+fcc <- df_run$defol[[3]]
+input_series <- fcc[[28]]

### 2020-04-09 paper_elsevier/paper_elsevier.Rmd
writing progress [2020-04-09]

@@
 date: "`r Sys.Date()`"
-bibliography: mybibfile.bib
+bibliography: References.bib
 linenumbers: true
 #numbersections: true
-csl: ecology.csl
-output: rticles::elsevier_article
-# output:
+# csl: ecology.csl
+csl: elsevier-harvard.csl
+# output: rticles::elsevier_article
+output:
+  bookdown::pdf_book:
+      base_format: rticles::elsevier_article
   # bookdown::word_document2:
@@
   \usepackage{amsmath}
+  \usepackage{float}
+  \usepackage{booktabs}
 ---
+
+```{r setup, include=FALSE}
+library(dfoliatR)
+library(knitr)
+library(magrittr)
+# library(kableExtra)
+# options(kableExtra.latex.load_packages = FALSE)
+library(kableExtra)
+# # opts_chunk$set(echo = TRUE)
+# # knit_hooks$set(document = function(x) {sub('\\usepackage[]{color}', `\\usepackage[table]{xcolor}', x, fixed = TRUE)})`)})
+# knit_hooks$set(document = function(x) {sub('\\usepackage[]{color}', '\\usepackage{xcolor}', x, fixed = TRUE)})
+```
 
@@
 
-Variation in the width and morphology of annual radial growth rings in wood permits dating and quantification of past forest insect defoliator outbreaks.
+Variation in the width and morphology of annual radial growth rings in trees permits dating and quantification of past forest insect defoliator outbreaks.
 Defoliation can be distinguished from climate- and other disturbance-related influences by comparing ring-width or other annually-resolved features in the wood of host species to non-host species or to climate records.
-The effect of defoliation on radial growth of trees has been recognized since the mid-1800s, but work prior to the 1980s was not cross-dated (based on annually-resolved and –dated measurements), not standardized (quantitatively controlled for tree geometry and age effects), and suffered from inadequate extraction of the defoliation-related variability from variability associated with climate and other factors [@Swetnam1985; @Speer2010; @Lynch2012].
-These and other problems may be circumvented by appropriate use of dendrochronological techniques.
-Quantitative methodology to infer forest defoliator outbreaks from cross-dated tree-ring records was developed in the 1980s by @Swetnam1985 for developing western spruce budworm chronologies [@Swetnam1989; @Swetnam1993].
+The effect of defoliation on radial growth of trees has been recognized since the mid-1800s, but it was not until the 1980s that dendrochronology techniques were applied to identify and quantify defoliation events [@Swetnam1985; @Speer2010; @Lynch2012].
+The first studies [@Swetnam1985; @Swetnam1989; @Swetnam1993] focused on developing historical reconstructions of western spruce budworm (WSBW; *Choristoneura freemani*).
 The methodology has since been successfully applied to a wide range of defoliator species, most of which are conifer herbivores, and has evolved in sophistication and application to a wide range of ecosystem situations [@Lynch2012].
@@
 The main dendrochronological tool for inferring, dating, and quantifying defoliator outbreaks from tree-ring records has been the software routine OUTBREAK [@Swetnam1985; @outbreak; @Swetnam1989].
-OUTBREAK computes indices (described later in detail) of suppressed growth by subtracting a detrended and standardized climate series (a “control” chronology, usually a site chronology developed from non-host trees or a gridded climate data point series) from host individual-tree detrended and standardized radial growth series after the host and non-host series have been brought to a common variance.
-If the host and non-host species respond similarly to climate (which can be tested), the derived series retains variability that the host and non-host series do not have in common, generally the insect signal and noise (unexplained variability).
+OUTBREAK computes indices (described later in detail) of suppressed growth by subtracting a detrended and standardized climate series (a “control” chronology) from host individual-tree detrended and standardized radial growth series after the host and non-host series have been brought to a common variance.
+The non-host chronology often consisted of a site chronology developed from non-host tree species, but a gridded climate data point series, like the North American Drought Atlas [@Cook2004] also suffices.
+If the host and non-host species respond similarly to climate (which can be tested), the derived series retains variability that the host and non-host series do not have in common, generally the insect signal, along with some unexplained variability (noise).
 The user defines a rule base specifying the magnitude and duration that a period of indexed growth suppression must meet or surpass for a period of suppressed growth to be inferred as a defoliation event at the tree level.
@@
 Though powerful, OUTBREAK is outdated and increasingly difficult to use in modern computing environments.
-It was written in Fortran V with inherently severe restrictions, as RAM and disk space were limited at that time (256 kb and 10 MB, respectively) and Fortran conventions imposed very strict formatting, file naming, and output conventions.
-Windows-based execution operates in a DOS window, Apple-based xxx in xxx, and provides no graphical interface or capabilities.
-Barriers to batch operation impose burdens for analyses of larger data sets.
-We developed dfoliaR as an `R`- and `dplR`-based routine to overcome these issues.
+It was written in FORTRAN V with inherently severe restrictions, as RAM and disk space were limited at that time (256 kb and 10 MB, respectively) and FORTRAN conventions imposed very strict formatting, file naming, and output conventions.
+The program provides no graphical interface or capabilities, forcing users to import generated text files into spreadsheet or other software for assessing results and performing analyses.
+Furthermore, OUTBREAK can only handle one test at a time, creating barriers to batch operation and a large burden for researchers with a datasets including multiple sites.
+We developed the software library `dfoliaR` as an `R`- and `dplR`-based routine to overcome these issues.
 
 `dfoliatR` adds to a growing suite of dendrochronology packages the `R` computing environment [@RCore].
-Stemming from the `dplR` library [@Bunn2008] that enables `R` users to read and write an array of tree-ring data formats, standardize ring width series, build and evaluate chronologies, and perform quality control (to name a few), one can now also measure ring widths from scanned images of preparred samples [@Shi2019; @Lara2015], perform and check crossdating [@Bunn2010], and perform many analytical tests [@Jevsenak2018; @Zang2015].
+Stemming from the `dplR` library [@Bunn2008] that enables `R` users to read and write an array of tree-ring data formats, standardize ring width series, build and evaluate chronologies, and perform quality control (to name a few), one can now also measure ring widths from scanned images of prepared samples [@Shi2019; @Lara2015], perform and check crossdating [@Bunn2010], and perform many analytical tests [@Jevsenak2018; @Zang2015].
 Tools for assessing stand dynamics and disturbance analyses are under rapid development, with new packages for assessing growth and release events [`TRADER`: @Altman2014], metrics of growth resilience [`pointRes`: @VanderMaaten-Theunissen2015], and fire history [`burnr`: @Malevich2018].
 The key objective of `dfoliatR` is to provide tools to identify and analyze insect defoliation and outbreak events by building on the methods employed by OUTBREAK.
-It capitilizes on the robust software already available in `R` by using `dplR` data formats for incoming tree-ring series and providing output data formats embodied by the `tidyverse` [@Wickham2019] that include efficient data manipulation [`dplyr`: @Wickham2020dplyr] and graphics [`ggplot2`: @wickham2016ggplot2].
+It capitalizes on the robust software already available in `R` by using `dplR` data formats for incoming tree-ring series and providing output data formats embodied by the `tidyverse` [@Wickham2019] that include efficient data manipulation [@Wickham2020dplyr] and graphics [@wickham2016ggplot2].
 
-In this paper, we describe the statistical methods employed by `dfoliatR`, its availability, compare results to those produced by OUTBREAK, and present an example analyses.
+In this paper, we describe the statistical methods employed by `dfoliatR`, compare results to those produced by OUTBREAK, and present an example analysis.
 Users need not have much experience in `R` to replicate the analyses and graphics as presented.
-The `R` code presented below is executable in an `R` session once the required libraries are installed and loaded.
-Support documentation in addition to this paper is provided within the package via standard help menus (accessed by typing `?` before a function name) and on the package website ([https://chguiterman.github.io/dfoliatR/](https://chguiterman.github.io/dfoliatR/)), which includes up-to-date vignettes that describe in detail the functionality of the software.
-Code to create a preprint of this manuscript including the `R` scripts is available from [https://github.com/chguiterman/dfoliatR_paper](https://github.com/chguiterman/dfoliatR_paper).
+The `R` code below is executable in an `R` session once the required libraries are installed and loaded.
+Support documentation in addition to this paper is provided within the package via standard help menus (accessed by typing `?` before a function name) and on the package website ([https://chguiterman.github.io/dfoliatR/](https://chguiterman.github.io/dfoliatR/)), which includes up-to-date vignettes that describe various software routines.
+Code to generate a preprint of this manuscript including the `R` scripts, and tubular and graphical output is available from [https://github.com/chguiterman/dfoliatR_paper](https://github.com/chguiterman/dfoliatR_paper).
 
@@
 At the heart of `dfoliatR` lies two functions: `defoliate_trees()` and `outbreak()`.
-These identify defoliation events on individual trees and then composite across multiple trees to identify outbreak events.
+These identify defoliation events on individual trees (Figure \@ref(fig:fig-defol)) and then composite across multiple trees to identify outbreak events (Figure \@ref(fig:fig-obr)).
 
@@
 
-The `defoliate_trees()` function is usually be the point of entry to the dfoliatR library.
-It performs two processes, removing climate-related growth signals from the host-tree series and then identifying tree-level defoliation events.
+The `defoliate_trees()` function is the point of entry to the `dfoliatR` library.
+It performs two processes, removing climate-related growth signals from the host-tree series and identifying tree-level defoliation events.
 The climatic or non-defoliation signals in each host-tree series are characterized by a non-host chronology or climate reconstruction.
@@
 We call it the "growth suppression index" (GSI).
-The GSI is calculated the same as in OUTBREAK [following @Swetnam1985; @Swetnam1989] for each host tree as
+The GSI is calculated the same as in OUTBREAK for each host tree as
 \begin{align}
@@
 
-Negative departures in the normalized GSI that surpass user-defined thresholds in duration and magnitude are defined as *defoliation events*.
+Negative departures in the normalized GSI (NGSI) that surpass user-defined thresholds in duration and magnitude are defined as *defoliation events*.
 As in OUTBREAK, magnitude is assessed on a single year within the departure sequence.
-The default setting is -1.28 (in units of standard deviation), which was previously determined to be representative of WSBW effects (**citation?**).
-Duration is assessed by examining sequences of negative GSI before and after the year of maximum departure.
+The default setting is -1.28 (in units of standard deviation), which was previously determined to be representative of WSBW effects [@Swetnam1989].
+Duration is assessed by examining sequences of negative NGSI before and after the year of maximum departure.
 Each defoliation event is allowed one positive excursion on each side of the maximum departure year.
 Duration is assessed across the entire sequence that includes up to two positive excursions.
-The default duration is eight years, as is commonly used in WSBW studies (**citation?**).
+The default duration is eight years, as is commonly used in WSBW studies [@Swetnam1989].
 Different species of defoliation insects vary in the length of defoliation and the degree to which they can suppress tree growth.
@@
 Users have options to define the number and/or the proportion of trees required for an event to be considered an outbreak.
-Three parameters control the whether a defoliation event constitutes an outbreak: the minimum number of trees available, the minimum number of trees recording defoliation, and the percent of trees recording defoliation.
-The first allows the researcher to make a judgement call as to the confidence ascribed to reduced sample depth toward the ends of their chronologies, thus compensating for the "fading record problem" (**Swetnam and Fritts?**).
+Three parameters control whether a defoliation event constitutes an outbreak: the minimum number of trees available, the minimum number of trees recording defoliation, and the percent of trees recording defoliation.
+The first allows the researcher to make a judgement call as to the confidence ascribed to reduced sample depth toward the ends of their chronologies, thus compensating for the "fading record problem" [@Swetnam1999].
 The second two parameters adjust the scale of defoliation considered to be an outbreak.
@@
 
-We tested `dfoliatR` against OUTBREAK by comparing GSI to corrected indices for individual trees and years, outbreak status for individual trees and years, and percentage of trees recording outbreaks at the site level, using raw ring-width data from 8 sites in British Columbia, Colorado, and New Mexico and author-provided non-host site chronologies.
+We tested `dfoliatR` against OUTBREAK by comparing NGSI to corrected indices for individual trees and years, defoliation status for individual trees and years, and percentage of trees recording outbreaks at the site level, using standardized ring-width data from eight sites in British Columbia, Colorado, and New Mexico, and author-provided non-host site chronologies.
 
-We detrended host data for both dfoliatR and OUTBREAK using ARSTAN 6.1 (**downloaded 21 April 2002 from dpl; xxx Ann check the date, as 21 April 2002 is 6.05P**) with default double detrending (128 year wavelength and a 50% smoothing spline) and event thresholds of -1.28 normalized index and 8 years.
-Attempts to detrend with ARSTAN 44 in `dplR` were problematic for two reasons.
-Some versions of ARSTAN, including ARSTAN 44, do not include an option for producing tree-level averages from multiple cores from the same tree (.tre files).
-**More importantly, ring-width indices from ARSTAN 44 truncated the process at either end when sample size fell below 5 xxx (Jodi, what was the minimum?)**, and produced different values for years near the truncation point than did dpl-ARSTAN 6.05P.
-The majority of the detrended series values were identical, but not the tails, as the two routines do not use the same data to establish the tails.
-These differences produce different GSI (`dfoliatR`) and Corrected Indices (OUTBREAK) values at the tails, sometimes considerably, affecting whether or not outbreak events are inferred early or late in the tree series, timing of the first or last event onset, and subsequent computation of return intervals, duration, and periodicity.
-Note that there is nothing particularly important regarding the choice of ARSTAN 6.1 – it is simply the version that the second author used routinely and had available.
+We detrended host data for both `dfoliatR` and OUTBREAK using ARSTAN v6.1 [@arstan] with default double detrending (128 year wavelength and a 50% smoothing spline).
+In both `dfoliatR` and OUTBREAK we used  event thresholds of -1.28 normalized index, 8 years duration, and allowed for events at the end of series in seven of eight sites.
+These seven sites were sampled during ongoing outbreak events [@Swetnam1993].
+We found it necessary to be consistent in how we detrended and in what software (e.g., ARSTAN vs `dplR`) we employed because subtle differences in standardized ring-width indices created between the programs exacerbated differences between `dfoliatR` and OUTBREAK.
+In the end, we chose to only use the standardization output files from ARSTAN, which are easily read into `R` (and then `dfoliatR`) using the `dplR` package.
+
+The `R` code to replicate our comparisons is available from [https://github.com/chguiterman/dfoliatR_paper](https://github.com/chguiterman/dfoliatR_paper).
 
@@
 
-`dfoliatR` and OUTBREAK compute growth suppresses indices and use them to infer identical tree-level outbreak events, including onset and termination dates and date and magnitude of the maximum growth suppression when `dfoliatR` bridging matches the OUTBREAK protocols.
+Since both programs employ Equation 1, `dfoliatR` and OUTBREAK compute identical growth suppression indices at 0.000 precision (Figure \@ref(fig:df-obr-1-1)).
+In all there are 43,280 ring-width indices from 222 trees included in our test.
+Of these, the programs identified 11,530 years of defoliation.
+The programs agreed on 92% of the years, leaving 927 "difference" years in which only one program identified defoliation on an individual tree.
+The differences included 102 events on 85 trees.
+We carefully inspected each of these events in the full context of each tree's ring series, and categorized the differences as follows
+
+* *Series-end events* (40% of the total) in which we allowed OUTBREAK to include "truncated outbreaks" (for seven sites) at the end of each series.
+In `dfoliatR`, this option is controlled by the "series_end_events" parameter to `defoliate_trees()`.
+In OUTBREAK, the option appears while changing the duration parameter (option 3).
+When selected, OUTBREAK will include any sequences of negative indices at the either the beginning or the end of each tree series.
+No thresholds are applied in identifying these as defoliation events.
+In `dfoliatR`, the negative sequences are evaluated as with any other possible defoliation, but the duration threshold is omitted.
+Each of the 13 events included in these differences did not meet the "max_reduction" parameter (-1.28 NGSI) in `dfoliatR` and were excluded.
+In two cases, OUTBREAK included events at the beginning of the series where `dfoliatR` does not allow truncated events.
+In four cases, OUTBREAK omitted the last year of the series because it was positive.
+In two cases, `dfoliatR` omitted possible events because it had already included a positive NGSI excursion after the "max_reduction" year, and since it will only allow one excursion on either side of the max year, the events were omitted due to short duration.
+
+* *Sequential events* (36%) in which OUTBREAK omitted events that occur one year prior or one year following an identified event.
+OUTBREAK selects the event with the greatest negative index year.
+On two trees, OUTBREAK omitted two of three sequential events.
+`dfoliatR` allows all such events.
+While inspecting these differences, we added an option to `defoiate_trees()` that would "bridge" between sequential events, making them into single, long events.
+We felt that this was ecologically justified, especially for studies of WSBW, because the insect will occasionally maintain minor defoliation f individuals for decades even while not in an outbreak phase.
+
+* *Undetermined differences* (22%) occurred in cases where OUTBREAK omitted events without clear cause.
+
+* *Rounding* (2%) differences in the indices either omitted or cut short events on two trees. In both cases the indices were very close to zero.
+
+Our inspections revealed what we believe are short-comings in how OUTBREAK identifies defoliation events on individual trees.
+In every one of the 102 cases we inspected, we felt that `dfoliatR` provided a more biologically and statistically appropriate assessment of defoliation.
+
+and use them to infer identical tree-level outbreak events, including onset and termination dates and date and magnitude of the maximum growth suppression when `dfoliatR` bridging matches the OUTBREAK protocols.
 Once the issues with ARSTAN (described earlier) were resolved, dfoliatR and OUTBREAK produced identical indices at 0.000 precision.
@@
 
-Bridging operates at the tree level in dfoliatR, and these effects generally have minor effects on site-level reconstructions (Fig.Bridging).
-Occasionally site-level events are inferred differently (Table.Bridging), and this can affect return interval and duration statistics.
+Bridging operates at the tree level in `dfoliatR`, and these effects generally have minor effects on site-level reconstructions (**Fig.Bridging**).
+Occasionally site-level events are inferred differently (**Table.Bridging**), and this can affect return interval and duration statistics.
 
@@
 In `dfoliatR` we provide two sets of tree-ring data to aid users in exploring the functions, graphics, and outputs.
-Each set consists of Douglas-fir (*Pseudotsuga menziesii*) host-tree series, standardized with __________, and a local ponderosa pine (*Pinus ponderosa*) non-host chronology.
-The non-host ring-width data were standardized by ________ and the chronologies averaged following standard procedures [@Speer2010].
-Data from Demijohn Peak (DMJ; 2902 m asl), in the San Juan Mountains of southern Colorado, come from @Ryerson2003.
-Data from the East Fork site (EF; 2580 m asl) in the Jemez Mountains of northcentral New Mexico were presented by @Swetnam1993.
+Each set consists of individual host-tree series, standardized using 50-year splines with a 50% frequency response, and a local non-host chronology.
+The non-host ring-width data were standardized using 150-year splines with a 50% frequency response.
+The non-host chronologies were averaged following standard procedures [@Speer2010].
+Host trees from Demijohn Peak (DMJ; 2902 m asl) in the San Juan Mountains of southern Colorado include Douglas-fir (*Pseudotsuga menziesii*) compared against a local non-host ponderosa pine (*Pinus ponderosa*) chronology [@Ryerson2003].
+The East Fork site (EF; 2580 m asl) in the Jemez Mountains of northcentral New Mexico include Douglas-fir and white fire (*Abies concolor*) host trees and a ponderosa pine non-host chronology [@Swetnam1993].
 
@@
 
+To assess defoliation of trees, we begin with a set of standardized host-tree series and a local non-host chronology or climate reconstruction.
+First, the `dfoliatR` library needs to be loaded (after installation as above) and the included datasets can be loaded.
+The function `defoliate_trees()` performs the GSI indexing procedure on each host-tree series and then identifies defoliation events.
+
+``` {r, echo=TRUE, eval=FALSE}
+## Load the package
+library(dfoliatR)
+
+## Load data
+data("dmj_h") # host tree series
+data("dmj_nh") # non-host chronology
+# View data structure for any data object via 
+# "View(dmj_h)" or "head(dmj_h)"
+
+## Calculate indices, identify defoliation events
+dmj_defol <- defoliate_trees(host_tree = dmj_h,
+                             nonhost_chron = dmj_nh,
+                             duration_years = 8,  
+                             max_reduction = -1.28, 
+                             bridge_events = TRUE, 
+                             series_end_event = TRUE, 
+                             list_output = FALSE) 
+```
+
+The result of running `defoliate_trees()` is long-format (stacked) data frame with five variables: "year", "series", "gsi", "ngsi", and "defol_status." The "gsi" variable is the "corrected" tree-level growth suppression index.
+It was normalized or scaled into z-scores to produce "ngsi", which is used to identify defoliation events.
+The "defol_status" column provides a set of character strings, "nd" for non-defoliation year, "defol" for a defoliation year, "max_defol" for the year of maximum suppression (that acts as the basis for individual events), "bridge_defol" to identify years that link subsequent events (only one is present at DMJ), and "series_end_defol" to identify defoliation at the present-end of the series.
+Selecting `list_output = TRUE` provides a list-object of data frames, each with the `rwl` object that combines the host tree and non-host chronology and the columns created in `defoliate_trees()`.
+This option is not used by subsequent functions in `dfoliatR`, but researchers can examine it to check the results of the GSI calculation (Equation 1).
+
+We can assess the results of running `defoliate_trees()` through graphical and table outputs.
+The function `get_defol_events()` will provide a list of every defoliation event for every tree, with the corresponding mean "ngsi" value.
+A summary table of the results for each tree is produced by `defol_stats()` (Table \@ref(tab:tbl-defol)).
+
+```{r echo=TRUE, eval=FALSE}
+defol_stats(dmj_defol)
+View(dmj_defol)
+```
+
+The `plot_defol()` function produces a "ggplot" graphics object with line segments showing the measured sequence of each series and a filled segment for each identified defoliation event (Figure \@ref(fig:fig-defol)).
+The defoliation segments are colored by their relative severity based on their average NGSI value.
+The default cut-off values between "Severe" and "Moderate" is the overall mean across all events and between "Moderate"  and "Minor" is the first quartile.
+Users can define these breaks to suit their needs.
+
+```{r echo=TRUE, eval=FALSE}
+plot_defol(dmj_defol)
+## This creates a ggplot object, so additions can 
+## be made to adjust plotting themes and
+## aesthetics, like color.   
+## For more on the features of ggplot:
+## https://ggplot2.tidyverse.org/
+library(ggplot2)
+plot_defol(dmj_defol) +
+  scale_color_manual(values = c("red", "orange", "purple"))
+```
+
+These basic output functions aid in assessing the sensitivity of input parameters to `defoliate_trees()` including the duration and magnitude thresholds for identifying defoliation events.
+Using `plot_defol()` also provides a direct assessment of the between-tree variability in deflation.
+For further analyses, we composite the host-tree defoliation series to the site level using `outbreak()`.
+
 ## Site-Level Events
 
-Extensions
+Once we have identified defoliation events on individual trees, we can infer outbreak events at the site level.
+The function `outbreak()` acts to composite tree-level defoliation series into a single chronology, with input parameters that control thresholds in the number and proportions of trees recording a defoliation event.
+
+```{r echo=TRUE, eval=FALSE}
+## Use the defol object "dmj_defol" created above
+dmj_obr <- outbreak(dmj_defol,
+                    filter_min_series = 3,
+                    filter_min_defol = 1,
+                    filter_perc = 25)
+```
+
+Input parameters to `outbreak()` include "filter_min_series" to control the chronology cut-off points with regard to sample depth, "filter_min_defol" and "filter_perc" to control the minimum number and percent of trees, respectively, recording a defoliation event in a given year.
+`Outbreak()` produces a new data frame with eight variables: "year", "num_defol", "percent_defol", "num_max_defol", "mean_gsi", "mean_ngsi", and "outbreak_status." All of these variables are populated regardless of an inferred outbreak event, providing a continuous disturbance chronology.
+The "num_max_defol" variable counts the number of trees recording their maximum defoliation in a given year.
+The "mean_gsi" and "mean_ngsi" variables provide averages of these indices across all available trees.
+Finally, the "outbreak_status" column shows whether an outbreak event is inferred ("outbreak") or not ("not_obr").
+
+The default plotting function to visualize results form `outbreak()` creates a three-panel graph showing the mean site-level chronology, the percent of trees recording a defoliation, and the sample depth over time (Figure \@ref(fig:fig-obr)).
+
+```{r echo=TRUE, eval=FALSE}
+plot_outbreak(dmj_obr, disp_index = "mean_gsi")
+```
+
+Inferred outbreak events are shown in the top panel of Figure \@ref(fig:fig-obr) as the filled-in spaces.
+Users can change the time series in this panel with the "disp_index" parameter, choosing between the mean NGSI (the default) or GSI.
+
+A summary table of the inferred outbreak events is generated by the `outbreak_stats()` function (Table \@ref(tab:tbl-obr)).
+
+```{r echo=TRUE, eval=FALSE}
+dmj_obr_stats <- outbreak_stats(dmj_obr)
+```
+
+The output table from `outbreak_stats()` shows the start and end years of each event, along with their corresponding duration, the number and percent of trees in defoliation at the start of the event, the the number of trees recording the outbreak event (in part or full), the minimums of the mean GSI and NGSI indices, and two points of "peak defoliation." The first, "peak_outbreak_year," is the year in which the greatest number of trees is recording the outbreak, and the second, "peak_defol_year," is the year in with the lowest average NGSI.
+These output variables provide a array of options for assessing basic metrics of the insect defoliation regime, including interval analyses:
+
+```{r echo=TRUE, eval=FALSE}
+## calculate mean return interval for outbreak start years
+intervals <- diff(dmj_obr_stats$start)
+mean(intervals) ## 51.2
+```
+
+Conclusions
 ==========================
 
-* Describe how dfoliatR can be combined with other R libraries
-  * Mapping, what else?
+* Encourage use of `dfoliatR` over OUTBREAK -- more acruate, greater user control, more easily conduct sensitivity tests to evaluate paramters, data, and ecological patterns
+* Facilitates high-levelr statistical analyses by easily transfering to other libraries (MASS, etc.)
+* A community resource that can be easily maintained and updated into the future.
+
+Acknowledgments {-}
+==========================
+
+We are grateful to late Richard Holmes of the Laboratory of Tree-Ring Research (LTRR) for his achievements in developing OUTBREAK and many other tree-ring software programs, and we thank Thomas Swetnam for his insights and encouragement.
+The `dfoliatR` project was supported by the Agnes Haury Visiting Scholars Fellowship at the LTRR and by the University of California Berkeley.
+Additional support was provided by the LTTR and the United States Forest Service Rocky Mountain Research Station.
+Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S.
+Government.
 
@@
 ==========================
+
+<div id="refs"></div>
+\newpage
+
+# Tables & Figures {-}
+
+```{r tbl-defol, echo=FALSE}
+tree_dat <- read.csv("../paper_elsevier/Output/defol-stats.csv")
+kable(tree_dat, caption = "Summary statistics for individual tree defoliation events.", booktabs = TRUE) %>% 
+  kableExtra::kable_styling(latex_options = "hold_position")
+```
+
+```{r tbl-obr, echo=FALSE}
+obr_dat <- read.csv("../paper_elsevier/Output/obr-stats.csv")
+kable(obr_dat, format = "latex", caption = "Summary statistics for inferred outbreak events.", booktabs = TRUE) %>% 
+  kable_styling(font_size = 7) %>% 
+  kableExtra::landscape()
+```
+
+```{r fig-defol, echo=FALSE, fig.cap="Default dfoliatR graphics for individual trees, produced by the 'plot_defol()' function.
+Dotted lines represent the series length for each tree and colored segments show periods of defoliation.
+Users can define the Cut-off values to determine severe-moderate-minor defoliation intensities, and use standard ggplot2 graphics parameters to adjust styles and themes."}
+include_graphics("../paper_elsevier/Output/tree-plot-default.pdf")
+```
+
+```{r fig-obr, echo=FALSE, fig.cap="Default graphic to show outbreak events at the DMJ site.
+Top panel shows the mean GSI through time, with inferred outbreak events filled.
+The middle panel shows the percent of trees defoliated, and the bottom panel shows the number of trees through time."}
+include_graphics("../paper_elsevier/Output/site-plot.pdf")
+```
+
+``` {r df-obr-1-1, echo=FALSE, fig.cap="Comparison of disturbance indices between dfoliatR and OUTBREAK. Diagonal lines show equal values (1:1)."}
+include_graphics("../paper_elsevier/Output/plot-trees-df-obr.pdf")
+```

### 2020-04-09 paper_elsevier/paper_elsevier.tex
writing progress [2020-04-09]

@@
 \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\usepackage{longtable}
+\usepackage{graphicx}
+% We will generate all images so they have a width \maxwidth. This means
+% that they will get their normal width if they fit onto the page, but
+% are scaled down if they would overflow the margins.
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
+\else\Gin@nat@width\fi}
+\makeatother
+\let\Oldincludegraphics\includegraphics
+\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
 \ifxetex
@@
 
-\setcounter{secnumdepth}{0}
+\setcounter{secnumdepth}{5}
 % Pandoc toggle for numbering sections (defaults to be off)
-\setcounter{secnumdepth}{0}
 
 % Pandoc header
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 
 \usepackage{amsmath}
+\usepackage{float}
+\usepackage{booktabs}
 
@@
 
-  \title{\emph{dfoliatR}: An R package for detection and analysis of insect
-defoliation signals in tree rings}
+  \title{\emph{dfoliatR}: An R package for detection and analysis of insect defoliation signals in tree rings}
     \author[a,b]{Christopher H. Guiterman\corref{1}}
@@
    \ead{jodi.axelson@berkeley.edu} 
-      \address[a]{Laboratory of Tree-Ring Research, University of Arizona, 1215 E Lowell
-St.~Box 210045, Tucson, AZ, 85721}
+      \address[a]{Laboratory of Tree-Ring Research, University of Arizona, 1215 E Lowell St.~Box 210045, Tucson, AZ, 85721}
     \address[b]{Three Pines Forest Research, LLC, PO Box 225, Etna, NH, 03750}
-    \address[c]{U.S. Forest Service, Rocky Mountain Research Station, 1215 E Lowell
-St.~Box 210045, Tucson, AZ, 85721}
-    \address[d]{Dept of Environmental Science, Policy \& Management, University of
-California, Berkeley, Berkeley, CA 94720}
+    \address[c]{U.S. Forest Service, Rocky Mountain Research Station, 1215 E Lowell St.~Box 210045, Tucson, AZ, 85721}
+    \address[d]{Dept of Environmental Science, Policy \& Management, University of California, Berkeley, Berkeley, CA 94720}
       \cortext[1]{Corresponding Author}
@@
   \begin{abstract}
-  We present a new R package to provide dendroecologists with tools to
-  infer, quantify, analyze, and visualize growth suppression events in
-  tree rings caused by insect defoliation. The `dfoliatR' library is based
-  on the FORTRAN V program OUTBREAK, and builds on existing resources in
-  the R computing environment. `dfoliatR' expands on OUTBREAK to provide
-  greater control of suppression thresholds, additional output tables, and
-  high-quality graphics. To use `dfoliatR' requires standardized
-  ring-width measurements from insect host trees and an indexed tree-ring
-  chronology from local non-host trees. It performs an indexing procedure
-  to remove the climatic signal represented in the non-host chronology
-  from the host-tree series. It then infers defoliation events in
-  individual trees. Site-level analyses identify outbreak events that
-  synchronously affect a user-defined number or proportion of the host
-  trees. Functions are available for summary statistics and graphics of
-  tree- and site-level series.
+  We present a new R package to provide dendroecologists with tools to infer, quantify, analyze, and visualize growth suppression events in tree rings caused by insect defoliation.
+The `dfoliatR' library is based on the FORTRAN V program OUTBREAK, and builds on existing resources in the R computing environment.
+`dfoliatR' expands on OUTBREAK to provide greater control of suppression thresholds, additional output tables, and high-quality graphics.
+To use `dfoliatR' requires standardized ring-width measurements from insect host trees and an indexed tree-ring chronology from local non-host trees.
+It performs an indexing procedure to remove the climatic signal represented in the non-host chronology from the host-tree series.
+It then infers defoliation events in individual trees.
+Site-level analyses identify outbreak events that synchronously affect a user-defined number or proportion of the host trees.
+Functions are available for summary statistics and graphics of tree- and site-level series.
 
@@
   \end{abstract}
-   \begin{keyword} Dendroecology, spruce budworm, Choristoneura, pandora moth, Coloradia
-pandora Blake, larch-bud-moth \newpage\end{keyword}
+   \begin{keyword} Dendroecology, spruce budworm, Choristoneura, pandora moth, Coloradia pandora Blake, larch-bud-moth
+\newpage\end{keyword}
  \end{frontmatter}
@@
 
-Variation in the width and morphology of annual radial growth rings in
-wood permits dating and quantification of past forest insect defoliator
-outbreaks. Defoliation can be distinguished from climate- and other
-disturbance-related influences by comparing ring-width or other
-annually-resolved features in the wood of host species to non-host
-species or to climate records. The effect of defoliation on radial
-growth of trees has been recognized since the mid-1800s, but work prior
-to the 1980s was not cross-dated (based on annually-resolved and --dated
-measurements), not standardized (quantitatively controlled for tree
-geometry and age effects), and suffered from inadequate extraction of
-the defoliation-related variability from variability associated with
-climate and other factors (Swetnam et al. 1985, Speer 2010, Lynch 2012).
-These and other problems may be circumvented by appropriate use of
-dendrochronological techniques. Quantitative methodology to infer forest
-defoliator outbreaks from cross-dated tree-ring records was developed in
-the 1980s by Swetnam et al. (1985) for developing western spruce budworm
-chronologies (Swetnam and Lynch 1989, 1993). The methodology has since
-been successfully applied to a wide range of defoliator species, most of
-which are conifer herbivores, and has evolved in sophistication and
-application to a wide range of ecosystem situations (Lynch 2012).
+Variation in the width and morphology of annual radial growth rings in trees permits dating and quantification of past forest insect defoliator outbreaks.
+Defoliation can be distinguished from climate- and other disturbance-related influences by comparing ring-width or other annually-resolved features in the wood of host species to non-host species or to climate records.
+The effect of defoliation on radial growth of trees has been recognized since the mid-1800s, but it was not until the 1980s that dendrochronology techniques were applied to identify and quantify defoliation events (Lynch, 2012; Speer, 2010; Swetnam et al., 1985).
+The first studies (Swetnam and Lynch, 1993, 1989; Swetnam et al., 1985) focused on developing historical reconstructions of western spruce budworm (WSBW; \emph{Choristoneura freemani}).
+The methodology has since been successfully applied to a wide range of defoliator species, most of which are conifer herbivores, and has evolved in sophistication and application to a wide range of ecosystem situations (Lynch, 2012).
 
-The main dendrochronological tool for inferring, dating, and quantifying
-defoliator outbreaks from tree-ring records has been the software
-routine OUTBREAK (Swetnam et al. 1985, Holmes and Swetnam 1986, Swetnam
-and Lynch 1989). OUTBREAK computes indices (described later in detail)
-of suppressed growth by subtracting a detrended and standardized climate
-series (a ``control'' chronology, usually a site chronology developed
-from non-host trees or a gridded climate data point series) from host
-individual-tree detrended and standardized radial growth series after
-the host and non-host series have been brought to a common variance. If
-the host and non-host species respond similarly to climate (which can be
-tested), the derived series retains variability that the host and
-non-host series do not have in common, generally the insect signal and
-noise (unexplained variability). The user defines a rule base specifying
-the magnitude and duration that a period of indexed growth suppression
-must meet or surpass for a period of suppressed growth to be inferred as
-a defoliation event at the tree level.
+The main dendrochronological tool for inferring, dating, and quantifying defoliator outbreaks from tree-ring records has been the software routine OUTBREAK (Holmes and Swetnam, 1986; Swetnam and Lynch, 1989; Swetnam et al., 1985).
+OUTBREAK computes indices (described later in detail) of suppressed growth by subtracting a detrended and standardized climate series (a ``control'' chronology) from host individual-tree detrended and standardized radial growth series after the host and non-host series have been brought to a common variance.
+The non-host chronology often consisted of a site chronology developed from non-host tree species, but a gridded climate data point series, like the North American Drought Atlas (Cook and Krusic, 2004) also suffices.
+If the host and non-host species respond similarly to climate (which can be tested), the derived series retains variability that the host and non-host series do not have in common, generally the insect signal, along with some unexplained variability (noise).
+The user defines a rule base specifying the magnitude and duration that a period of indexed growth suppression must meet or surpass for a period of suppressed growth to be inferred as a defoliation event at the tree level.
 
-Though powerful, OUTBREAK is outdated and increasingly difficult to use
-in modern computing environments. It was written in Fortran V with
-inherently severe restrictions, as RAM and disk space were limited at
-that time (256 kb and 10 MB, respectively) and Fortran conventions
-imposed very strict formatting, file naming, and output conventions.
-Windows-based execution operates in a DOS window, Apple-based xxx in
-xxx, and provides no graphical interface or capabilities. Barriers to
-batch operation impose burdens for analyses of larger data sets. We
-developed dfoliaR as an \texttt{R}- and \texttt{dplR}-based routine to
-overcome these issues.
+Though powerful, OUTBREAK is outdated and increasingly difficult to use in modern computing environments.
+It was written in FORTRAN V with inherently severe restrictions, as RAM and disk space were limited at that time (256 kb and 10 MB, respectively) and FORTRAN conventions imposed very strict formatting, file naming, and output conventions.
+The program provides no graphical interface or capabilities, forcing users to import generated text files into spreadsheet or other software for assessing results and performing analyses.
+Furthermore, OUTBREAK can only handle one test at a time, creating barriers to batch operation and a large burden for researchers with a datasets including multiple sites.
+We developed the software library \texttt{dfoliaR} as an \texttt{R}- and \texttt{dplR}-based routine to overcome these issues.
 
-\texttt{dfoliatR} adds to a growing suite of dendrochronology packages
-the \texttt{R} computing environment (R Core Team 2019). Stemming from
-the \texttt{dplR} library (Bunn 2008) that enables \texttt{R} users to
-read and write an array of tree-ring data formats, standardize ring
-width series, build and evaluate chronologies, and perform quality
-control (to name a few), one can now also measure ring widths from
-scanned images of preparred samples (Lara et al. 2015, Shi and Xiang
-2019), perform and check crossdating (Bunn 2010), and perform many
-analytical tests (Zang and Biondi 2015, Jevšenak and Levanič 2018).
-Tools for assessing stand dynamics and disturbance analyses are under
-rapid development, with new packages for assessing growth and release
-events (\texttt{TRADER}: Altman et al. 2014), metrics of growth
-resilience (\texttt{pointRes}: Maaten-Theunissen et al. 2015), and fire
-history (\texttt{burnr}: Malevich et al. 2018). The key objective of
-\texttt{dfoliatR} is to provide tools to identify and analyze insect
-defoliation and outbreak events by building on the methods employed by
-OUTBREAK. It capitilizes on the robust software already available in
-\texttt{R} by using \texttt{dplR} data formats for incoming tree-ring
-series and providing output data formats embodied by the
-\texttt{tidyverse} (Wickham et al. 2019) that include efficient data
-manipulation (\texttt{dplyr}: Wickham et al. 2020) and graphics
-(\texttt{ggplot2}: Wickham 2016).
+\texttt{dfoliatR} adds to a growing suite of dendrochronology packages the \texttt{R} computing environment (R Core Team, 2019).
+Stemming from the \texttt{dplR} library (Bunn, 2008) that enables \texttt{R} users to read and write an array of tree-ring data formats, standardize ring width series, build and evaluate chronologies, and perform quality control (to name a few), one can now also measure ring widths from scanned images of prepared samples (Lara et al., 2015; Shi and Xiang, 2019), perform and check crossdating (Bunn, 2010), and perform many analytical tests (Jevšenak and Levanič, 2018; Zang and Biondi, 2015).
+Tools for assessing stand dynamics and disturbance analyses are under rapid development, with new packages for assessing growth and release events (\texttt{TRADER}: Altman et al., 2014), metrics of growth resilience (\texttt{pointRes}: Maaten-Theunissen et al., 2015), and fire history (\texttt{burnr}: Malevich et al., 2018).
+The key objective of \texttt{dfoliatR} is to provide tools to identify and analyze insect defoliation and outbreak events by building on the methods employed by OUTBREAK.
+It capitalizes on the robust software already available in \texttt{R} by using \texttt{dplR} data formats for incoming tree-ring series and providing output data formats embodied by the \texttt{tidyverse} (Wickham et al., 2019) that include efficient data manipulation (Wickham et al., 2020) and graphics (Wickham, 2016).
 
-In this paper, we describe the statistical methods employed by
-\texttt{dfoliatR}, its availability, compare results to those produced
-by OUTBREAK, and present an example analyses. Users need not have much
-experience in \texttt{R} to replicate the analyses and graphics as
-presented. The \texttt{R} code presented below is executable in an
-\texttt{R} session once the required libraries are installed and loaded.
-Support documentation in addition to this paper is provided within the
-package via standard help menus (accessed by typing \texttt{?} before a
-function name) and on the package website
-(\url{https://chguiterman.github.io/dfoliatR/}), which includes
-up-to-date vignettes that describe in detail the functionality of the
-software. Code to create a preprint of this manuscript including the
-\texttt{R} scripts is available from
-\url{https://github.com/chguiterman/dfoliatR_paper}.
+In this paper, we describe the statistical methods employed by \texttt{dfoliatR}, compare results to those produced by OUTBREAK, and present an example analysis.
+Users need not have much experience in \texttt{R} to replicate the analyses and graphics as presented.
+The \texttt{R} code below is executable in an \texttt{R} session once the required libraries are installed and loaded.
+Support documentation in addition to this paper is provided within the package via standard help menus (accessed by typing \texttt{?} before a function name) and on the package website (\url{https://chguiterman.github.io/dfoliatR/}), which includes up-to-date vignettes that describe various software routines.
+Code to generate a preprint of this manuscript including the \texttt{R} scripts, and tubular and graphical output is available from \url{https://github.com/chguiterman/dfoliatR_paper}.
 
@@
 
-The \texttt{dfoliatR} library requires two sets of tree-ring data to
-infer defoliation and outbreak events:
+The \texttt{dfoliatR} library requires two sets of tree-ring data to infer defoliation and outbreak events:
 
@@
 \item
-  Standardized ring-width series for individual trees of the host
-  species
+  Standardized ring-width series for individual trees of the host species
 \item
@@
 
-Users can develop these data sets in software of their choosing, such as
-\texttt{dplR} (Bunn 2008) or ARSTAN (Cook and Holmes 1996). It is
-important that the host-tree data include only one tree-ring series per
-tree. Both \texttt{dplR} and ARSTAN have options for averaging multiple
-sample series into a tree-level series.
+Users can develop these data sets in software of their choosing, such as \texttt{dplR} (Bunn, 2008) or ARSTAN (Cook and Holmes, 1996).
+It is important that the host-tree data include only one tree-ring series per tree.
+Both \texttt{dplR} and ARSTAN have options for averaging multiple sample series into a tree-level series.
 
-At the heart of \texttt{dfoliatR} lies two functions:
-\texttt{defoliate\_trees()} and \texttt{outbreak()}. These identify
-defoliation events on individual trees and then composite across
-multiple trees to identify outbreak events.
+At the heart of \texttt{dfoliatR} lies two functions: \texttt{defoliate\_trees()} and \texttt{outbreak()}.
+These identify defoliation events on individual trees (Figure \ref{fig:fig-defol}) and then composite across multiple trees to identify outbreak events (Figure \ref{fig:fig-obr}).
 
 \hypertarget{identifying-defoliation-of-trees}{%
-\subsection{Identifying Defoliation of
-Trees}\label{identifying-defoliation-of-trees}}
+\subsection{Identifying Defoliation of Trees}\label{identifying-defoliation-of-trees}}
 
-The \texttt{defoliate\_trees()} function is usually be the point of
-entry to the dfoliatR library. It performs two processes, removing
-climate-related growth signals from the host-tree series and then
-identifying tree-level defoliation events. The climatic or
-non-defoliation signals in each host-tree series are characterized by a
-non-host chronology or climate reconstruction. \texttt{dfoiatR} removes
-that non-defoliation signal by subtracting the non-host series from each
-host-tree series, which generates a residual index. In OUTBREAK, this
-residual index was termed the ``corrected'' index. We call it the
-``growth suppression index'' (GSI). The GSI is calculated the same as in
-OUTBREAK (following Swetnam et al. 1985, Swetnam and Lynch 1989) for
-each host tree as \begin{align}
+The \texttt{defoliate\_trees()} function is the point of entry to the \texttt{dfoliatR} library.
+It performs two processes, removing climate-related growth signals from the host-tree series and identifying tree-level defoliation events.
+The climatic or non-defoliation signals in each host-tree series are characterized by a non-host chronology or climate reconstruction.
+\texttt{dfoiatR} removes that non-defoliation signal by subtracting the non-host series from each host-tree series, which generates a residual index.
+In OUTBREAK, this residual index was termed the ``corrected'' index.
+We call it the ``growth suppression index'' (GSI).
+The GSI is calculated the same as in OUTBREAK for each host tree as
+\begin{align}
 \textrm{GSI}_{i} = \textrm{H}_{i} - \left( \textrm{NH}_{i} - \overline{\textrm{NH}} \right) \frac{\sigma_{\textrm{H}}}{\sigma_{\textrm{NH}}} \
-\end{align} where H and NH are the host-tree series and the non-host
-chronology, in year i, respectively. Only the common period between the
-host-tree series and the non-host chronology are used in Equation 1. The
-host and non-host chronologies are brought to common variance by scaling
-the non-host chronology by its mean
-(\(\overline{\textrm{NH}} \approx 1.0\)) and multiplying by the ratio of
-host and non-host standard deviations
-(\(\frac{\sigma_{\textrm{H}}}{\sigma_{\textrm{NH}}}\)), which
-approximates the variance of the host tree series.
+\end{align}
+where H and NH are the host-tree series and the non-host chronology, in year i, respectively.
+Only the common period between the host-tree series and the non-host chronology are used in Equation 1.
+The host and non-host chronologies are brought to common variance by scaling the non-host chronology by its mean (\(\overline{\textrm{NH}} \approx 1.0\)) and multiplying by the ratio of host and non-host standard deviations (\(\frac{\sigma_{\textrm{H}}}{\sigma_{\textrm{NH}}}\)), which approximates the variance of the host tree series.
 
-Negative departures in the normalized GSI that surpass user-defined
-thresholds in duration and magnitude are defined as \emph{defoliation
-events}. As in OUTBREAK, magnitude is assessed on a single year within
-the departure sequence. The default setting is -1.28 (in units of
-standard deviation), which was previously determined to be
-representative of WSBW effects (\textbf{citation?}). Duration is
-assessed by examining sequences of negative GSI before and after the
-year of maximum departure. Each defoliation event is allowed one
-positive excursion on each side of the maximum departure year. Duration
-is assessed across the entire sequence that includes up to two positive
-excursions. The default duration is eight years, as is commonly used in
-WSBW studies (\textbf{citation?}). Different species of defoliation
-insects vary in the length of defoliation and the degree to which they
-can suppress tree growth. Researchers can, and should, adjust the
-duration and magnitude parameters accordingly and critically evaluate
-the results.
+Negative departures in the normalized GSI (NGSI) that surpass user-defined thresholds in duration and magnitude are defined as \emph{defoliation events}.
+As in OUTBREAK, magnitude is assessed on a single year within the departure sequence.
+The default setting is -1.28 (in units of standard deviation), which was previously determined to be representative of WSBW effects (Swetnam and Lynch, 1989).
+Duration is assessed by examining sequences of negative NGSI before and after the year of maximum departure.
+Each defoliation event is allowed one positive excursion on each side of the maximum departure year.
+Duration is assessed across the entire sequence that includes up to two positive excursions.
+The default duration is eight years, as is commonly used in WSBW studies (Swetnam and Lynch, 1989).
+Different species of defoliation insects vary in the length of defoliation and the degree to which they can suppress tree growth.
+Researchers can, and should, adjust the duration and magnitude parameters accordingly and critically evaluate the results.
 
-Diverging from OUTBREAK, \texttt{dfoliatR} allows users to extend
-defoliation events by bridging successive events and also by allowing
-potentially short-duration events that occur at the end of the series.
-In cases where two defoliation events are separated by a single year,
-bridging will link them into a single event (\textbf{Show figure?}). We
-urge careful use of this option because there is no setting to limit the
-number or length of potentially bridged events. The series end option
-can be used in cases when the host trees were actively being defoliated
-at the time of sampling. This option eliminates the duration parameter
-for an event at the recent end of the series, but all other thresholds
-apply. The advantage of this parameter is that it can aid in identifying
-the start-year for the current defoliation event or outbreak, which is
-both useful in management and allows the current event to be included in
-return-interval estimates.
+Diverging from OUTBREAK, \texttt{dfoliatR} allows users to extend defoliation events by bridging successive events and also by allowing potentially short-duration events that occur at the end of the series.
+In cases where two defoliation events are separated by a single year, bridging will link them into a single event (\textbf{Show figure?}).
+We urge careful use of this option because there is no setting to limit the number or length of potentially bridged events.
+The series end option can be used in cases when the host trees were actively being defoliated at the time of sampling.
+This option eliminates the duration parameter for an event at the recent end of the series, but all other thresholds apply.
+The advantage of this parameter is that it can aid in identifying the start-year for the current defoliation event or outbreak, which is both useful in management and allows the current event to be included in return-interval estimates.
 
@@
 
-Defoliation of one or a few trees does not constitute an outbreak. To
-determine when defoliation becomes an \emph{outbreak event},
-\texttt{dfoliatR} composites the individual tree defoliation series into
-a site-level chronology with the \texttt{outbreak()} function. Users
-have options to define the number and/or the proportion of trees
-required for an event to be considered an outbreak. Three parameters
-control the whether a defoliation event constitutes an outbreak: the
-minimum number of trees available, the minimum number of trees recording
-defoliation, and the percent of trees recording defoliation. The first
-allows the researcher to make a judgement call as to the confidence
-ascribed to reduced sample depth toward the ends of their chronologies,
-thus compensating for the ``fading record problem'' (\textbf{Swetnam and
-Fritts?}). The second two parameters adjust the scale of defoliation
-considered to be an outbreak. Absolute numbers of trees and percentages
-can be applied separately or in conjunction, following filtering
-conventions in tree ring fire history studies (Malevich et al. 2018). We
-urge users to carefully consider the choice of absolute numbers in
-situations where the number of trees represented in the series varies
-with time, or the choice of percentages when sample size is small.
+Defoliation of one or a few trees does not constitute an outbreak.
+To determine when defoliation becomes an \emph{outbreak event}, \texttt{dfoliatR} composites the individual tree defoliation series into a site-level chronology with the \texttt{outbreak()} function.
+Users have options to define the number and/or the proportion of trees required for an event to be considered an outbreak.
+Three parameters control whether a defoliation event constitutes an outbreak: the minimum number of trees available, the minimum number of trees recording defoliation, and the percent of trees recording defoliation.
+The first allows the researcher to make a judgement call as to the confidence ascribed to reduced sample depth toward the ends of their chronologies, thus compensating for the ``fading record problem'' (Swetnam et al., 1999).
+The second two parameters adjust the scale of defoliation considered to be an outbreak.
+Absolute numbers of trees and percentages can be applied separately or in conjunction, following filtering conventions in tree ring fire history studies (Malevich et al., 2018).
+We urge users to carefully consider the choice of absolute numbers in situations where the number of trees represented in the series varies with time, or the choice of percentages when sample size is small.
 
@@
 
-We tested \texttt{dfoliatR} against OUTBREAK by comparing GSI to
-corrected indices for individual trees and years, outbreak status for
-individual trees and years, and percentage of trees recording outbreaks
-at the site level, using raw ring-width data from 8 sites in British
-Columbia, Colorado, and New Mexico and author-provided non-host site
-chronologies.
+We tested \texttt{dfoliatR} against OUTBREAK by comparing NGSI to corrected indices for individual trees and years, defoliation status for individual trees and years, and percentage of trees recording outbreaks at the site level, using standardized ring-width data from eight sites in British Columbia, Colorado, and New Mexico, and author-provided non-host site chronologies.
 
-We detrended host data for both dfoliatR and OUTBREAK using ARSTAN 6.1
-(\textbf{downloaded 21 April 2002 from dpl; xxx Ann check the date, as
-21 April 2002 is 6.05P}) with default double detrending (128 year
-wavelength and a 50\% smoothing spline) and event thresholds of -1.28
-normalized index and 8 years. Attempts to detrend with ARSTAN 44 in
-\texttt{dplR} were problematic for two reasons. Some versions of ARSTAN,
-including ARSTAN 44, do not include an option for producing tree-level
-averages from multiple cores from the same tree (.tre files). More
-importantly, ring-width indices from ARSTAN 44 truncated the process at
-either end when sample size fell below 5 xxx (Jodi, what was the
-minimum?), and produced different values for years near the truncation
-point than did dpl-ARSTAN 6.05P. The majority of the detrended series
-values were identical, but not the tails, as the two routines do not use
-the same data to establish the tails. These differences produce
-different GSI (\texttt{dfoliatR}) and Corrected Indices (OUTBREAK)
-values at the tails, sometimes considerably, affecting whether or not
-outbreak events are inferred early or late in the tree series, timing of
-the first or last event onset, and subsequent computation of return
-intervals, duration, and periodicity. Note that there is nothing
-particularly important regarding the choice of ARSTAN 6.1 -- it is
-simply the version that the second author used routinely and had
-available.
+We detrended host data for both \texttt{dfoliatR} and OUTBREAK using ARSTAN v6.1 (Cook and Holmes, 1996) with default double detrending (128 year wavelength and a 50\% smoothing spline).
+In both \texttt{dfoliatR} and OUTBREAK we used event thresholds of -1.28 normalized index, 8 years duration, and allowed for events at the end of series in seven of eight sites.
+These seven sites were sampled during ongoing outbreak events (Swetnam and Lynch, 1993).
+We found it necessary to be consistent in how we detrended and in what software (e.g., ARSTAN vs \texttt{dplR}) we employed because subtle differences in standardized ring-width indices created between the programs exacerbated differences between \texttt{dfoliatR} and OUTBREAK.
+In the end, we chose to only use the standardization output files from ARSTAN, which are easily read into \texttt{R} (and then \texttt{dfoliatR}) using the \texttt{dplR} package.
+
+The \texttt{R} code to replicate our comparisons is available from \url{https://github.com/chguiterman/dfoliatR_paper}.
 
@@
 
-\texttt{dfoliatR} and OUTBREAK compute growth suppresses indices and use
-them to infer identical tree-level outbreak events, including onset and
-termination dates and date and magnitude of the maximum growth
-suppression when \texttt{dfoliatR} bridging matches the OUTBREAK
-protocols. Once the issues with ARSTAN (described earlier) were
-resolved, dfoliatR and OUTBREAK produced identical indices at 0.000
-precision.
+Since both programs employ Equation 1, \texttt{dfoliatR} and OUTBREAK compute identical growth suppression indices at 0.000 precision (Figure \ref{fig:df-obr-1-1}).
+In all there are 43,280 ring-width indices from 222 trees included in our test.
+Of these, the programs identified 11,530 years of defoliation.
+The programs agreed on 92\% of the years, leaving 927 ``difference'' years in which only one program identified defoliation on an individual tree.
+The differences included 102 events on 85 trees.
+We carefully inspected each of these events in the full context of each tree's ring series, and categorized the differences as follows
 
-When inferring outbreak events at the tree level, our initial results
-did not always match OUTBREAK, leading to development of the bridging
-option. Identifying a minimum within a period while allowing for
-positive excursions is a tricky computation. OUTBREAK deliberately
-disallows back-to-back events -- after a minimum value within a period
-of negative values, a second positive index terminates an event.
-OUTBREAK does not allow a period of highly negative values to be
-appended to an earlier outbreak if the intervening period is 1 year,
-even if those negative values fall below the minimum in the earlier
-period. Due in large part to reconstructions using OUTBREAK (see papers
-cited by Lynch 2012), we now know considerably more about forest
-defoliator outbreak regimes than we did in the 1980s when OUTBREAK was
-written. We think that two prolonged events separated by a single year
-should in some situations be considered a single outbreak at both the
-tree and site levels. This is particularly relevant to western spruce
-budworm and spruce budworm (\emph{C. fumiferana}), for which regimes
-have been reconstructed for several geographic areas and for which
-researchers and forest health experts have gained considerable knowledge
-(\textbf{Sanders et al.~1985, Brookes et al.~1987 and many later
-publications}). We now know that the greatest growth suppression often
-occurs late in the outbreak due to cumulative effects (\textbf{rrr}). We
-developed the bridging option to permit these periods to be linked to
-preceding events separated by a single year.
+\begin{itemize}
+\item
+  \emph{Series-end events} (40\% of the total) in which we allowed OUTBREAK to include ``truncated outbreaks'' (for seven sites) at the end of each series.
+In \texttt{dfoliatR}, this option is controlled by the ``series\_end\_events'' parameter to \texttt{defoliate\_trees()}.
+In OUTBREAK, the option appears while changing the duration parameter (option 3).
+When selected, OUTBREAK will include any sequences of negative indices at the either the beginning or the end of each tree series.
+No thresholds are applied in identifying these as defoliation events.
+In \texttt{dfoliatR}, the negative sequences are evaluated as with any other possible defoliation, but the duration threshold is omitted.
+Each of the 13 events included in these differences did not meet the ``max\_reduction'' parameter (-1.28 NGSI) in \texttt{dfoliatR} and were excluded.
+In two cases, OUTBREAK included events at the beginning of the series where \texttt{dfoliatR} does not allow truncated events.
+In four cases, OUTBREAK omitted the last year of the series because it was positive.
+In two cases, \texttt{dfoliatR} omitted possible events because it had already included a positive NGSI excursion after the ``max\_reduction'' year, and since it will only allow one excursion on either side of the max year, the events were omitted due to short duration.
+\item
+  \emph{Sequential events} (36\%) in which OUTBREAK omitted events that occur one year prior or one year following an identified event.
+OUTBREAK selects the event with the greatest negative index year.
+On two trees, OUTBREAK omitted two of three sequential events.
+\texttt{dfoliatR} allows all such events.
+While inspecting these differences, we added an option to \texttt{defoiate\_trees()} that would ``bridge'' between sequential events, making them into single, long events.
+We felt that this was ecologically justified, especially for studies of WSBW, because the insect will occasionally maintain minor defoliation f individuals for decades even while not in an outbreak phase.
+\item
+  \emph{Undetermined differences} (22\%) occurred in cases where OUTBREAK omitted events without clear cause.
+\item
+  \emph{Rounding} (2\%) differences in the indices either omitted or cut short events on two trees. In both cases the indices were very close to zero.
+\end{itemize}
 
-Bridging operates at the tree level in dfoliatR, and these effects
-generally have minor effects on site-level reconstructions
-(Fig.Bridging). Occasionally site-level events are inferred differently
-(Table.Bridging), and this can affect return interval and duration
-statistics.
+Our inspections revealed what we believe are short-comings in how OUTBREAK identifies defoliation events on individual trees.
+In every one of the 102 cases we inspected, we felt that \texttt{dfoliatR} provided a more biologically and statistically appropriate assessment of defoliation.
+
+and use them to infer identical tree-level outbreak events, including onset and termination dates and date and magnitude of the maximum growth suppression when \texttt{dfoliatR} bridging matches the OUTBREAK protocols.
+Once the issues with ARSTAN (described earlier) were resolved, dfoliatR and OUTBREAK produced identical indices at 0.000 precision.
+
+When inferring outbreak events at the tree level, our initial results did not always match OUTBREAK, leading to development of the bridging option.
+Identifying a minimum within a period while allowing for positive excursions is a tricky computation.
+OUTBREAK deliberately disallows back-to-back events -- after a minimum value within a period of negative values, a second positive index terminates an event.
+OUTBREAK does not allow a period of highly negative values to be appended to an earlier outbreak if the intervening period is 1 year, even if those negative values fall below the minimum in the earlier period.
+Due in large part to reconstructions using OUTBREAK (see papers cited by Lynch, 2012), we now know considerably more about forest defoliator outbreak regimes than we did in the 1980s when OUTBREAK was written.
+We think that two prolonged events separated by a single year should in some situations be considered a single outbreak at both the tree and site levels.
+This is particularly relevant to western spruce budworm and spruce budworm (\emph{C.
+fumiferana}), for which regimes have been reconstructed for several geographic areas and for which researchers and forest health experts have gained considerable knowledge (\textbf{Sanders et al.~1985, Brookes et al.~1987 and many later publications}).
+We now know that the greatest growth suppression often occurs late in the outbreak due to cumulative effects (\textbf{rrr}).
+We developed the bridging option to permit these periods to be linked to preceding events separated by a single year.
+
+Bridging operates at the tree level in \texttt{dfoliatR}, and these effects generally have minor effects on site-level reconstructions (\textbf{Fig.Bridging}).
+Occasionally site-level events are inferred differently (\textbf{Table.Bridging}), and this can affect return interval and duration statistics.
 
 \hypertarget{best-practice-recommendations}{%
-\subsection{Best Practice
-Recommendations}\label{best-practice-recommendations}}
+\subsection{Best Practice Recommendations}\label{best-practice-recommendations}}
 
-We encourage authors to report software and library version numbers and
-download dates where applicable. This is standard practice for
-statistical packages, but has often been ignored for R- and dplR-based
-analyses. R-based software evolves over time, and results cannot always
-be replicated if the software versions are unknown. Similarly,
-significant changes (especially fixes) imply that re-analysis of data
-might differ from earlier published results and interpretations.
-Wherever possible, refereed publications supporting the software should
-be cited. Our experience with ARSTAN demonstrates the importance of
-adopting this practice. In the United States, analytic methods not
-documented thusly and supported by the literature might be questioned if
-the results are used in a National Environmental Protection Act (NEPA)
-process.
+We encourage authors to report software and library version numbers and download dates where applicable.
+This is standard practice for statistical packages, but has often been ignored for R- and dplR-based analyses.
+R-based software evolves over time, and results cannot always be replicated if the software versions are unknown.
+Similarly, significant changes (especially fixes) imply that re-analysis of data might differ from earlier published results and interpretations.
+Wherever possible, refereed publications supporting the software should be cited.
+Our experience with ARSTAN demonstrates the importance of adopting this practice.
+In the United States, analytic methods not documented thusly and supported by the literature might be questioned if the results are used in a National Environmental Protection Act (NEPA) process.
 
-We recommend that researchers compare dfoliatR results with and without
-bridging, and see which option agrees best with known insect biology and
-ecology. Bridging may be a logical choice for insects with prolonged
-non-outbreak periods, such as western spruce budworm, but not for
-situations where impacted stands barely recover from one outbreak before
-another begins, such as occurs with pine processionary caterpillars
-(\emph{Thaumetopoea pityocampa} (Lepidoptera: Thaumetopoeidae))
-(\textbf{Carus 2004, 2009}).
+We recommend that researchers compare dfoliatR results with and without bridging, and see which option agrees best with known insect biology and ecology.
+Bridging may be a logical choice for insects with prolonged non-outbreak periods, such as western spruce budworm, but not for situations where impacted stands barely recover from one outbreak before another begins, such as occurs with pine processionary caterpillars (\emph{Thaumetopoea pityocampa} (Lepidoptera: Thaumetopoeidae)) (\textbf{Carus 2004, 2009}).
 
 \hypertarget{availability-and-installation}{%
-\section{Availability and
-installation}\label{availability-and-installation}}
+\section{Availability and installation}\label{availability-and-installation}}
 
-The \texttt{dfoliatR} library (Guiterman et al. 2020) is provided free
-and open source from the Comprehensive R Archive Network (CRAN;
-\url{https://cran.r-project.org/}). To install \texttt{dfoliatR} from
-CRAN use
+The \texttt{dfoliatR} library (Guiterman et al., 2020) is provided free and open source from the Comprehensive R Archive Network (CRAN; \url{https://cran.r-project.org/}).
+To install \texttt{dfoliatR} from CRAN use
 
@@
 
-Development versions of \texttt{dfoliatR} are available on GitHub and
-installed using the \texttt{devtools} library,
+Development versions of \texttt{dfoliatR} are available on GitHub and installed using the \texttt{devtools} library,
 
@@
 
-Issues, bug reports, and ideas for improving \texttt{dfoliatR} can be
-posted to \url{https://github.com/chguiterman/dfoliatR/issues}. As an
-Open Source library, we welcome and encourage community involvement in
-future development. The best ways to contribute to \texttt{dfoliatR} are
-through standard GitHub procedures or by contacting the first author.
+Issues, bug reports, and ideas for improving \texttt{dfoliatR} can be posted to \url{https://github.com/chguiterman/dfoliatR/issues}.
+As an Open Source library, we welcome and encourage community involvement in future development.
+The best ways to contribute to \texttt{dfoliatR} are through standard GitHub procedures or by contacting the first author.
 
@@
 
-In \texttt{dfoliatR} we provide two sets of tree-ring data to aid users
-in exploring the functions, graphics, and outputs. Each set consists of
-Douglas-fir (\emph{Pseudotsuga menziesii}) host-tree series,
-standardized with \_\_\_\_\_\_\_\_\_\_, and a local ponderosa pine
-(\emph{Pinus ponderosa}) non-host chronology. The non-host ring-width
-data were standardized by \_\_\_\_\_\_\_\_ and the chronologies averaged
-following standard procedures (Speer 2010). Data from Demijohn Peak
-(DMJ; 2902 m asl), in the San Juan Mountains of southern Colorado, come
-from Ryerson et al. (2003). Data from the East Fork site (EF; 2580 m
-asl) in the Jemez Mountains of northcentral New Mexico were presented by
-Swetnam and Lynch (1993).
+In \texttt{dfoliatR} we provide two sets of tree-ring data to aid users in exploring the functions, graphics, and outputs.
+Each set consists of individual host-tree series, standardized using 50-year splines with a 50\% frequency response, and a local non-host chronology.
+The non-host ring-width data were standardized using 150-year splines with a 50\% frequency response.
+The non-host chronologies were averaged following standard procedures (Speer, 2010).
+Host trees from Demijohn Peak (DMJ; 2902 m asl) in the San Juan Mountains of southern Colorado include Douglas-fir (\emph{Pseudotsuga menziesii}) compared against a local non-host ponderosa pine (\emph{Pinus ponderosa}) chronology (Ryerson et al., 2003).
+The East Fork site (EF; 2580 m asl) in the Jemez Mountains of northcentral New Mexico include Douglas-fir and white fire (\emph{Abies concolor}) host trees and a ponderosa pine non-host chronology (Swetnam and Lynch, 1993).
 
 \hypertarget{tree-level-defoliation-events}{%
-\subsection{Tree-Level Defoliation
-Events}\label{tree-level-defoliation-events}}
+\subsection{Tree-Level Defoliation Events}\label{tree-level-defoliation-events}}
+
+To assess defoliation of trees, we begin with a set of standardized host-tree series and a local non-host chronology or climate reconstruction.
+First, the \texttt{dfoliatR} library needs to be loaded (after installation as above) and the included datasets can be loaded.
+The function \texttt{defoliate\_trees()} performs the GSI indexing procedure on each host-tree series and then identifies defoliation events.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{## Load the package}
+\KeywordTok{library}\NormalTok{(dfoliatR)}
+
+\CommentTok{## Load data}
+\KeywordTok{data}\NormalTok{(}\StringTok{"dmj_h"}\NormalTok{) }\CommentTok{# host tree series}
+\KeywordTok{data}\NormalTok{(}\StringTok{"dmj_nh"}\NormalTok{) }\CommentTok{# non-host chronology}
+\CommentTok{# View data structure for any data object via }
+\CommentTok{# "View(dmj_h)" or "head(dmj_h)"}
+
+\CommentTok{## Calculate indices, identify defoliation events}
+\NormalTok{dmj_defol <-}\StringTok{ }\KeywordTok{defoliate_trees}\NormalTok{(}\DataTypeTok{host_tree =}\NormalTok{ dmj_h,}
+                             \DataTypeTok{nonhost_chron =}\NormalTok{ dmj_nh,}
+                             \DataTypeTok{duration_years =} \DecValTok{8}\NormalTok{,  }
+                             \DataTypeTok{max_reduction =} \FloatTok{-1.28}\NormalTok{, }
+                             \DataTypeTok{bridge_events =} \OtherTok{TRUE}\NormalTok{, }
+                             \DataTypeTok{series_end_event =} \OtherTok{TRUE}\NormalTok{, }
+                             \DataTypeTok{list_output =} \OtherTok{FALSE}\NormalTok{) }
+\end{Highlighting}
+\end{Shaded}
+
+The result of running \texttt{defoliate\_trees()} is long-format (stacked) data frame with five variables: ``year'', ``series'', ``gsi'', ``ngsi'', and ``defol\_status.'' The ``gsi'' variable is the ``corrected'' tree-level growth suppression index.
+It was normalized or scaled into z-scores to produce ``ngsi'', which is used to identify defoliation events.
+The ``defol\_status'' column provides a set of character strings, ``nd'' for non-defoliation year, ``defol'' for a defoliation year, ``max\_defol'' for the year of maximum suppression (that acts as the basis for individual events), ``bridge\_defol'' to identify years that link subsequent events (only one is present at DMJ), and ``series\_end\_defol'' to identify defoliation at the present-end of the series.
+Selecting \texttt{list\_output\ =\ TRUE} provides a list-object of data frames, each with the \texttt{rwl} object that combines the host tree and non-host chronology and the columns created in \texttt{defoliate\_trees()}.
+This option is not used by subsequent functions in \texttt{dfoliatR}, but researchers can examine it to check the results of the GSI calculation (Equation 1).
+
+We can assess the results of running \texttt{defoliate\_trees()} through graphical and table outputs.
+The function \texttt{get\_defol\_events()} will provide a list of every defoliation event for every tree, with the corresponding mean ``ngsi'' value.
+A summary table of the results for each tree is produced by \texttt{defol\_stats()} (Table \ref{tab:tbl-defol}).
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{defol_stats}\NormalTok{(dmj_defol)}
+\KeywordTok{View}\NormalTok{(dmj_defol)}
+\end{Highlighting}
+\end{Shaded}
+
+The \texttt{plot\_defol()} function produces a ``ggplot'' graphics object with line segments showing the measured sequence of each series and a filled segment for each identified defoliation event (Figure \ref{fig:fig-defol}).
+The defoliation segments are colored by their relative severity based on their average NGSI value.
+The default cut-off values between ``Severe'' and ``Moderate'' is the overall mean across all events and between ``Moderate'' and ``Minor'' is the first quartile.
+Users can define these breaks to suit their needs.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{plot_defol}\NormalTok{(dmj_defol)}
+\CommentTok{## This creates a ggplot object, so additions can }
+\CommentTok{## be made to adjust plotting themes and}
+\CommentTok{## aesthetics, like color.   }
+\CommentTok{## For more on the features of ggplot:}
+\CommentTok{## https://ggplot2.tidyverse.org/}
+\KeywordTok{library}\NormalTok{(ggplot2)}
+\KeywordTok{plot_defol}\NormalTok{(dmj_defol) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{scale_color_manual}\NormalTok{(}\DataTypeTok{values =} \KeywordTok{c}\NormalTok{(}\StringTok{"red"}\NormalTok{, }\StringTok{"orange"}\NormalTok{, }\StringTok{"purple"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+These basic output functions aid in assessing the sensitivity of input parameters to \texttt{defoliate\_trees()} including the duration and magnitude thresholds for identifying defoliation events.
+Using \texttt{plot\_defol()} also provides a direct assessment of the between-tree variability in deflation.
+For further analyses, we composite the host-tree defoliation series to the site level using \texttt{outbreak()}.
 
@@
 
-\hypertarget{extensions}{%
-\section{Extensions}\label{extensions}}
+Once we have identified defoliation events on individual trees, we can infer outbreak events at the site level.
+The function \texttt{outbreak()} acts to composite tree-level defoliation series into a single chronology, with input parameters that control thresholds in the number and proportions of trees recording a defoliation event.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{## Use the defol object "dmj_defol" created above}
+\NormalTok{dmj_obr <-}\StringTok{ }\KeywordTok{outbreak}\NormalTok{(dmj_defol,}
+                    \DataTypeTok{filter_min_series =} \DecValTok{3}\NormalTok{,}
+                    \DataTypeTok{filter_min_defol =} \DecValTok{1}\NormalTok{,}
+                    \DataTypeTok{filter_perc =} \DecValTok{25}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+Input parameters to \texttt{outbreak()} include ``filter\_min\_series'' to control the chronology cut-off points with regard to sample depth, ``filter\_min\_defol'' and ``filter\_perc'' to control the minimum number and percent of trees, respectively, recording a defoliation event in a given year.
+\texttt{Outbreak()} produces a new data frame with eight variables: ``year'', ``num\_defol'', ``percent\_defol'', ``num\_max\_defol'', ``mean\_gsi'', ``mean\_ngsi'', and ``outbreak\_status.'' All of these variables are populated regardless of an inferred outbreak event, providing a continuous disturbance chronology.
+The ``num\_max\_defol'' variable counts the number of trees recording their maximum defoliation in a given year.
+The ``mean\_gsi'' and ``mean\_ngsi'' variables provide averages of these indices across all available trees.
+Finally, the ``outbreak\_status'' column shows whether an outbreak event is inferred (``outbreak'') or not (``not\_obr'').
+
+The default plotting function to visualize results form \texttt{outbreak()} creates a three-panel graph showing the mean site-level chronology, the percent of trees recording a defoliation, and the sample depth over time (Figure \ref{fig:fig-obr}).
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{plot_outbreak}\NormalTok{(dmj_obr, }\DataTypeTok{disp_index =} \StringTok{"mean_gsi"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+Inferred outbreak events are shown in the top panel of Figure \ref{fig:fig-obr} as the filled-in spaces.
+Users can change the time series in this panel with the ``disp\_index'' parameter, choosing between the mean NGSI (the default) or GSI.
+
+A summary table of the inferred outbreak events is generated by the \texttt{outbreak\_stats()} function (Table \ref{tab:tbl-obr}).
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{dmj_obr_stats <-}\StringTok{ }\KeywordTok{outbreak_stats}\NormalTok{(dmj_obr)}
+\end{Highlighting}
+\end{Shaded}
+
+The output table from \texttt{outbreak\_stats()} shows the start and end years of each event, along with their corresponding duration, the number and percent of trees in defoliation at the start of the event, the the number of trees recording the outbreak event (in part or full), the minimums of the mean GSI and NGSI indices, and two points of ``peak defoliation.'' The first, ``peak\_outbreak\_year,'' is the year in which the greatest number of trees is recording the outbreak, and the second, ``peak\_defol\_year,'' is the year in with the lowest average NGSI.
+These output variables provide a array of options for assessing basic metrics of the insect defoliation regime, including interval analyses:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{## calculate mean return interval for outbreak start years}
+\NormalTok{intervals <-}\StringTok{ }\KeywordTok{diff}\NormalTok{(dmj_obr_stats}\OperatorTok{$}\NormalTok{start)}
+\KeywordTok{mean}\NormalTok{(intervals) }\CommentTok{## 51.2}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{conclusions}{%
+\section{Conclusions}\label{conclusions}}
 
@@
 \item
-  Describe how dfoliatR can be combined with other R libraries
+  Encourage use of \texttt{dfoliatR} over OUTBREAK -- more acruate, greater user control, more easily conduct sensitivity tests to evaluate paramters, data, and ecological patterns
+\item
+  Facilitates high-levelr statistical analyses by easily transfering to other libraries (MASS, etc.)
+\item
+  A community resource that can be easily maintained and updated into the future.
+\end{itemize}
 
-  \begin{itemize}
-  \tightlist
-  \item
-    Mapping, what else?
-  \end{itemize}
-\end{itemize}
+\hypertarget{acknowledgments}{%
+\section*{Acknowledgments}\label{acknowledgments}}
+\addcontentsline{toc}{section}{Acknowledgments}
+
+We are grateful to late Richard Holmes of the Laboratory of Tree-Ring Research (LTRR) for his achievements in developing OUTBREAK and many other tree-ring software programs, and we thank Thomas Swetnam for his insights and encouragement.
+The \texttt{dfoliatR} project was supported by the Agnes Haury Visiting Scholars Fellowship at the LTRR and by the University of California Berkeley.
+Additional support was provided by the LTTR and the United States Forest Service Rocky Mountain Research Station.
+Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S.
+Government.
 
@@
 \leavevmode\hypertarget{ref-Altman2014}{}%
-Altman, J., P. Fibich, J. Dolezal, and T. Aakala. 2014. TRADER: A
-package for Tree Ring Analysis of Disturbance Events in R.
-Dendrochronologia 32:107--112.
+Altman, J., Fibich, P., Dolezal, J., Aakala, T., 2014.
+TRADER: A package for Tree Ring Analysis of Disturbance Events in R.
+Dendrochronologia 32, 107--112.
+doi:\href{https://doi.org/10.1016/j.dendro.2014.01.004}{10.1016/j.dendro.2014.01.004}
 
 \leavevmode\hypertarget{ref-Bunn2008}{}%
-Bunn, A. G. 2008. A dendrochronology program library in R (dplR).
-Dendrochronologia 26:115--124.
+Bunn, A.G., 2008. A dendrochronology program library in R (dplR). Dendrochronologia 26, 115--124. doi:\href{https://doi.org/10.1016/j.dendro.2008.01.002}{10.1016/j.dendro.2008.01.002}
 
 \leavevmode\hypertarget{ref-Bunn2010}{}%
-Bunn, A. G. 2010. Statistical and visual crossdating in R using the dplR
-library. Dendrochronologia 28:251--258.
+Bunn, A.G., 2010. Statistical and visual crossdating in R using the dplR library. Dendrochronologia 28, 251--258. doi:\href{https://doi.org/10.1016/j.dendro.2009.12.001}{10.1016/j.dendro.2009.12.001}
 
 \leavevmode\hypertarget{ref-arstan}{}%
-Cook, E. R., and R. L. Holmes. 1996. Guide for computer program arstan.
-Pages 75--87 The international tree-ring data bank program library
-version.
+Cook, E.R., Holmes, R.L., 1996. Guide for computer program arstan, The international tree-ring data bank program library version.
+
+\leavevmode\hypertarget{ref-Cook2004}{}%
+Cook, E.R., Krusic, P.J., 2004. The North American Drought Atlas.
 
 \leavevmode\hypertarget{ref-dfoliatR}{}%
-Guiterman, C., A. Lynch, and J. Axelson. 2020. DfoliatR: Detection and
-analysis of insect defoliation signals in tree rings.
+Guiterman, C., Lynch, A., Axelson, J., 2020. DfoliatR: Detection and analysis of insect defoliation signals in tree rings.
 
 \leavevmode\hypertarget{ref-outbreak}{}%
-Holmes, R. L., and T. W. Swetnam. 1986. Dendroecology program library:
-Program outbreak user's manual. Laboratory of Tree-Ring Research,
-University of Arizona, Tucson.
+Holmes, R.L., Swetnam, T.W., 1986. Dendroecology program library: Program outbreak user's manual. Laboratory of Tree-Ring Research, University of Arizona, Tucson.
 
 \leavevmode\hypertarget{ref-Jevsenak2018}{}%
-Jevšenak, J., and T. Levanič. 2018. dendroTools: R package for studying
-linear and nonlinear responses between tree-rings and daily
-environmental data. Dendrochronologia 48:32--39.
+Jevšenak, J., Levanič, T., 2018.
+dendroTools: R package for studying linear and nonlinear responses between tree-rings and daily environmental data.
+Dendrochronologia 48, 32--39.
+doi:\href{https://doi.org/10.1016/j.dendro.2018.01.005}{10.1016/j.dendro.2018.01.005}
 
 \leavevmode\hypertarget{ref-Lara2015}{}%
-Lara, W., F. Bravo, and C. A. Sierra. 2015. MeasuRing: An R package to
-measure tree-ring widths from scanned images. Dendrochronologia
-34:43--50.
+Lara, W., Bravo, F., Sierra, C.A., 2015.
+MeasuRing: An R package to measure tree-ring widths from scanned images.
+Dendrochronologia 34, 43--50.
+doi:\href{https://doi.org/10.1016/j.dendro.2015.04.002}{10.1016/j.dendro.2015.04.002}
 
 \leavevmode\hypertarget{ref-Lynch2012}{}%
-Lynch, A. M. 2012. What Tree-Ring Reconstruction Tells Us about Conifer
-Defoliator Outbreaks. Pages 126--154 \emph{in} P. Barbosa, D. K.
-Letourneau, and A. A. Agrawal, editors. Insect outbreaks revisited.
-Blackwell Publishing Ltd.
+Lynch, A.M., 2012.
+What Tree-Ring Reconstruction Tells Us about Conifer Defoliator Outbreaks, in: Barbosa, P., Letourneau, D.K., Agrawal, A.A.
+(Eds.), Insect Outbreaks Revisited.
+Blackwell Publishing Ltd, pp.
+126--154.
+doi:\href{https://doi.org/10.1002/9781118295205.ch7}{10.1002/9781118295205.ch7}
 
 \leavevmode\hypertarget{ref-VanderMaaten-Theunissen2015}{}%
-Maaten-Theunissen, M. van der, E. van der Maaten, and O. Bouriaud. 2015.
-PointRes: An R package to analyze pointer years and components of
-resilience. Dendrochronologia 35:34--38.
+Maaten-Theunissen, M.
+van der, Maaten, E.
+van der, Bouriaud, O., 2015.
+PointRes: An R package to analyze pointer years and components of resilience.
+Dendrochronologia 35, 34--38.
+doi:\href{https://doi.org/10.1016/j.dendro.2015.05.006}{10.1016/j.dendro.2015.05.006}
 
 \leavevmode\hypertarget{ref-Malevich2018}{}%
-Malevich, S. B., C. H. Guiterman, and E. Q. Margolis. 2018. burnr: Fire
-history analysis and graphics in R. Dendrochronologia 49:9--15.
+Malevich, S.B., Guiterman, C.H., Margolis, E.Q., 2018.
+burnr: Fire history analysis and graphics in R.
+Dendrochronologia 49, 9--15.
+doi:\href{https://doi.org/10.1016/j.dendro.2018.02.005}{10.1016/j.dendro.2018.02.005}
 
 \leavevmode\hypertarget{ref-RCore}{}%
-R Core Team. 2019. R: A language and environment for statistical
-computing. R Foundation for Statistical Computing, Vienna, Austria.
+R Core Team, 2019. R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria.
 
 \leavevmode\hypertarget{ref-Ryerson2003}{}%
-Ryerson, D. E., T. W. Swetnam, and A. M. Lynch. 2003. A tree-ring
-reconstruction of western spruce budworm outbreaks in the San Juan
-Mountains, Colorado, U.S.A. Canadian Journal of Forest Research
-33:1010--1028.
+Ryerson, D.E., Swetnam, T.W., Lynch, A.M., 2003.
+A tree-ring reconstruction of western spruce budworm outbreaks in the San Juan Mountains, Colorado, U.S.A.
+Canadian Journal of Forest Research 33, 1010--1028.
+doi:\href{https://doi.org/10.1139/X03-026}{10.1139/X03-026}
 
 \leavevmode\hypertarget{ref-Shi2019}{}%
-Shi, J., and W. Xiang. 2019. MtreeRing: A shiny application for
-automatic measurements of tree-ring widths on digital images.
+Shi, J., Xiang, W., 2019. MtreeRing: A shiny application for automatic measurements of tree-ring widths on digital images.
 
 \leavevmode\hypertarget{ref-Speer2010}{}%
-Speer, J. H. 2010. Fundamentals of tree-ring research. Page 333. The
-University of Arizona Press.
+Speer, J.H., 2010. Fundamentals of tree-ring research. The University of Arizona Press.
+
+\leavevmode\hypertarget{ref-Swetnam1999}{}%
+Swetnam, T., Allen, C., Betancourt, J., 1999. Applied historical ecology: using the past to manage for the future. Ecological applications 9, 1189--1206.
 
 \leavevmode\hypertarget{ref-Swetnam1989}{}%
-Swetnam, T. W., and A. M. Lynch. 1989. A tree-ring reconstruction of
-western spruce budworm history in the southern Rocky Mountains. Forest
-Science 35:962--986.
+Swetnam, T.W., Lynch, A.M., 1989.
+A tree-ring reconstruction of western spruce budworm history in the southern Rocky Mountains.
+Forest Science 35, 962--986.
+doi:\href{https://doi.org/10.1093/forestscience/35.4.962}{10.1093/forestscience/35.4.962}
 
 \leavevmode\hypertarget{ref-Swetnam1993}{}%
-Swetnam, T. W., and A. M. Lynch. 1993. Multicentury, Regional-Scale
-Patterns of Western Spruce Budworm Outbreaks. Ecological Monographs
-63:399--424.
+Swetnam, T.W., Lynch, A.M., 1993. Multicentury, Regional-Scale Patterns of Western Spruce Budworm Outbreaks. Ecological Monographs 63, 399--424.
 
 \leavevmode\hypertarget{ref-Swetnam1985}{}%
-Swetnam, T. W., M. A. Thompson, and E. K. Sutherland. 1985. Using
-dendrochronology to measure radial growth of defoliated trees. Page 39p.
+Swetnam, T.W., Thompson, M.A., Sutherland, E.K., 1985. Using dendrochronology to measure radial growth of defoliated trees.
 
 \leavevmode\hypertarget{ref-wickham2016ggplot2}{}%
-Wickham, H. 2016. Ggplot2: Elegant graphics for data analysis. Springer.
+Wickham, H., 2016. Ggplot2: Elegant graphics for data analysis. Springer.
 
 \leavevmode\hypertarget{ref-Wickham2019}{}%
-Wickham, H., M. Averick, J. Bryan, W. Chang, L. McGowan, R. François, G.
-Grolemund, A. Hayes, L. Henry, J. Hester, M. Kuhn, T. Pedersen, E.
-Miller, S. Bache, K. Müller, J. Ooms, D. Robinson, D. Seidel, V. Spinu,
-K. Takahashi, D. Vaughan, C. Wilke, K. Woo, and H. Yutani. 2019. Welcome
-to the tidyverse. Journal of Open Source Software 4:1686.
+Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L., François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M., Pedersen, T., Miller, E., Bache, S., Müller, K., Ooms, J., Robinson, D., Seidel, D., Spinu, V., Takahashi, K., Vaughan, D., Wilke, C., Woo, K., Yutani, H., 2019.
+Welcome to the tidyverse.
+Journal of Open Source Software 4, 1686.
+doi:\href{https://doi.org/10.21105/joss.01686}{10.21105/joss.01686}
 
 \leavevmode\hypertarget{ref-Wickham2020dplyr}{}%
-Wickham, H., R. François, L. Henry, and K. Müller. 2020. Dplyr: A
-grammar of data manipulation.
+Wickham, H., François, R., Henry, L., Müller, K., 2020. Dplyr: A grammar of data manipulation.
 
 \leavevmode\hypertarget{ref-Zang2015}{}%
-Zang, C., and F. Biondi. 2015. Treeclim: An R package for the numerical
-calibration of proxy-climate relationships. Ecography 38:431--436.
+Zang, C., Biondi, F., 2015.
+Treeclim: An R package for the numerical calibration of proxy-climate relationships.
+Ecography 38, 431--436.
+doi:\href{https://doi.org/10.1111/ecog.01335}{10.1111/ecog.01335}
+
+\newpage
+
+\hypertarget{tables-figures}{%
+\section*{Tables \& Figures}\label{tables-figures}}
+\addcontentsline{toc}{section}{Tables \& Figures}
+
+\begin{table}[!h]
+
+\caption{\label{tab:tbl-defol}Summary statistics for individual tree defoliation events.}
+\centering
+\begin{tabular}[t]{lrrrrrr}
+\toprule
+series & first & last & years & num\_events & tot\_years & mean\_duration\\
+\midrule
+DMJ01 & 1810 & 1996 & 187 & 4 & 50 & 12\\
+DMJ02 & 1750 & 1996 & 247 & 6 & 75 & 12\\
+DMJ03 & 1830 & 1996 & 167 & 4 & 42 & 10\\
+DMJ04 & 1720 & 1996 & 277 & 8 & 90 & 11\\
+DMJ06 & 1700 & 1996 & 297 & 6 & 81 & 14\\
+\addlinespace
+DMJ07 & 1710 & 1996 & 287 & 7 & 96 & 14\\
+DMJ11 & 1900 & 1997 & 98 & 2 & 19 & 10\\
+DMJ14 & 1675 & 1996 & 322 & 9 & 112 & 12\\
+DMJ15 & 1730 & 1996 & 267 & 4 & 64 & 16\\
+DMJ16 & 1746 & 1996 & 251 & 7 & 86 & 12\\
+\addlinespace
+DMJ17 & 1733 & 1996 & 264 & 6 & 85 & 14\\
+DMJ22 & 1720 & 1996 & 277 & 5 & 56 & 11\\
+DMJ23 & 1675 & 1997 & 323 & 8 & 96 & 12\\
+DMJ24 & 1895 & 1996 & 102 & 3 & 28 & 9\\
+DMJ25 & 1680 & 1996 & 317 & 5 & 70 & 14\\
+\addlinespace
+DMJ26 & 1700 & 1996 & 297 & 6 & 79 & 13\\
+DMJ27 & 1710 & 1996 & 287 & 3 & 34 & 11\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{landscape}\begin{table}
+
+\caption{\label{tab:tbl-obr}Summary statistics for inferred outbreak events.}
+\centering
+\fontsize{7}{9}\selectfont
+\begin{tabular}[t]{rrrrrrrrrr}
+\toprule
+start & end & duration & num\_trees\_start & perc\_trees\_start & peak\_outbreak\_year & num\_trees\_outbreak & peak\_defol\_year & min\_gsi & mean\_ngsi\\
+\midrule
+1680 & 1699 & 20 & 1 & 33.3 & 1690 & 3 & 1692 & 0.121 & -2.407\\
+1753 & 1769 & 17 & 6 & 46.2 & 1754 & 7 & 1755 & 0.343 & -1.611\\
+1825 & 1840 & 16 & 11 & 78.6 & 1831 & 12 & 1826 & 0.500 & -1.304\\
+1849 & 1865 & 17 & 7 & 46.7 & 1852 & 13 & 1853 & 0.252 & -1.994\\
+1881 & 1895 & 15 & 8 & 53.3 & 1886 & 14 & 1885 & 0.262 & -1.945\\
+\addlinespace
+1959 & 1970 & 12 & 7 & 41.2 & 1960 & 15 & 1965 & 0.328 & -1.830\\
+1987 & 1996 & 10 & 9 & 52.9 & 1989 & 15 & 1995 & 0.378 & -1.640\\
+\bottomrule
+\end{tabular}
+\end{table}
+\end{landscape}
+
+\begin{figure}
+\centering
+\includegraphics{../paper_elsevier/Output/tree-plot-default.pdf}
+\caption{\label{fig:fig-defol}Default dfoliatR graphics for individual trees, produced by the `plot\_defol()' function.
+Dotted lines represent the series length for each tree and colored segments show periods of defoliation.
+Users can define the Cut-off values to determine severe-moderate-minor defoliation intensities, and use standard ggplot2 graphics parameters to adjust styles and themes.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{../paper_elsevier/Output/site-plot.pdf}
+\caption{\label{fig:fig-obr}Default graphic to show outbreak events at the DMJ site.
+Top panel shows the mean GSI through time, with inferred outbreak events filled.
+The middle panel shows the percent of trees defoliated, and the bottom panel shows the number of trees through time.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{../paper_elsevier/Output/plot-trees-df-obr.pdf}
+\caption{\label{fig:df-obr-1-1}Comparison of disturbance indices between dfoliatR and OUTBREAK. Diagonal 1:1 lines show equal values.}
+\end{figure}
 

### 2020-05-05 paper_elsevier/paper_elsevier.Rmd
Ann edits and references

@@
   - name: Christopher H. Guiterman
-    email: chguiterman@email.arizona.edu
+    email: chguiterman@arizona.edu
     affiliation: a,b
@@
   - name: Ann M. Lynch
-    email: lyncha@email.arizona.edu
+    email: ann.m.lynch@usda.gov
     affiliation: a,c
@@
 abstract: |
-  We present a new R package to provide dendroecologists with tools to infer, quantify, analyze, and visualize growth suppression events in tree rings caused by insect defoliation.
-The ‘dfoliatR’ library is based on the FORTRAN V program OUTBREAK, and builds on existing resources in the R computing environment.
+  We present a new R package to provide dendroecologists with tools to infer, quantify, analyze, and visualize growth suppression events in tree rings caused by insect defoliation, and to reconstruct defoliator outbreak chronologies.
+‘dfoliatR’ is based on the OUTBREAK program and builds on existing resources in the R and dplR computing environments.
+It performs an indexing procedure to remove climatic signals, represented in non-host chronologies or other reconstructed climate series, from the host-tree series.
+It then infers defoliation events in individual trees based on user-specified thresholds.
+Site-level analyses identify outbreak events that synchronously affect user-defined numbers or proportions of involved host trees.
+Functions are provided for summary statistics and graphics of tree- and site-level series.
 ‘dfoliatR’ expands on OUTBREAK to provide greater control of suppression thresholds, additional output tables, and high-quality graphics.
-To use ‘dfoliatR’ requires standardized ring-width measurements from insect host trees and an indexed tree-ring chronology from local non-host trees.
-It performs an indexing procedure to remove the climatic signal represented in the non-host chronology from the host-tree series.
-It then infers defoliation events in individual trees.
-Site-level analyses identify outbreak events that synchronously affect a user-defined number or proportion of the host trees.
-Functions are available for summary statistics and graphics of tree- and site-level series.
+We provide two example data sets and script to enable users to gain familiarity with the package and its capabilities.
+The source code is available in the Comprehensive R Archive Network (CRAN) and on GitHub.
 
@@
   \newpage
-journal: "Tree-Ring Research"
+journal: "Dendrochronologia"
 date: "`r Sys.Date()`"
@@
 # csl: ecology.csl
-csl: elsevier-harvard.csl
+# csl: elsevier-harvard.csl
+# csl: elsevier-vancouver.csl
 # output: rticles::elsevier_article
@@
 #     df_print: kable
+# always_allow_html: true
 layout: review
@@
 Variation in the width and morphology of annual radial growth rings in trees permits dating and quantification of past forest insect defoliator outbreaks.
-Defoliation can be distinguished from climate- and other disturbance-related influences by comparing ring-width or other annually-resolved features in the wood of host species to non-host species or to climate records.
-The effect of defoliation on radial growth of trees has been recognized since the mid-1800s, but it was not until the 1980s that dendrochronology techniques were applied to identify and quantify defoliation events [@Swetnam1985; @Speer2010; @Lynch2012].
-The first studies [@Swetnam1985; @Swetnam1989; @Swetnam1993] focused on developing historical reconstructions of western spruce budworm (WSBW; *Choristoneura freemani*).
-The methodology has since been successfully applied to a wide range of defoliator species, most of which are conifer herbivores, and has evolved in sophistication and application to a wide range of ecosystem situations [@Lynch2012].
+Defoliation can be distinguished from climate- and other disturbance-related influences by comparing ring-width or other annually-resolved features in the wood of host species to that of non-host species or annually-resolved climate records.
+The effect of defoliation on radial growth of trees has been recognized since the 1860s, and used to reconstruct outbreak regimes since the 1950s [@Blais1954; @Lynch2012].
+It was not until the 1980s, however, that precise dendrochronological techniques were applied for inferring defoliation events and reconstructing defoliator outbreak regimes [@Swetnam1985; @Speer2010; @Lynch2012].
+The first studies [@Swetnam1985; @Swetnam1989; @Swetnam1993] focused on developing historical reconstructions of western spruce budworm (WSBW; *Choristoneura freemani* (Lepidoptera: Tortricidae); previously known as *C.
+occidentalis*) outbreak regimes.
+The methodology has since been successfully applied to a wide range of defoliator species, most of which are conifer herbivores, and has evolved in sophistication for a wide range of ecosystem situations [@Lynch2012].
 
-The main dendrochronological tool for inferring, dating, and quantifying defoliator outbreaks from tree-ring records has been the software routine OUTBREAK [@Swetnam1985; @outbreak; @Swetnam1989].
-OUTBREAK computes indices (described later in detail) of suppressed growth by subtracting a detrended and standardized climate series (a “control” chronology) from host individual-tree detrended and standardized radial growth series after the host and non-host series have been brought to a common variance.
-The non-host chronology often consisted of a site chronology developed from non-host tree species, but a gridded climate data point series, like the North American Drought Atlas [@Cook2004] also suffices.
-If the host and non-host species respond similarly to climate (which can be tested), the derived series retains variability that the host and non-host series do not have in common, generally the insect signal, along with some unexplained variability (noise).
+The main dendrochronological tool for inferring, dating, and characterizing defoliator outbreaks from tree-ring records has been the software routine OUTBREAK [@Swetnam1985; @outbreak; @Swetnam1989].
+OUTBREAK computes indices (described later in detail) of suppressed growth by subtracting a detrended and standardized climate series (a “control” chronology) from individual host-tree detrended and standardized radial growth series after the host and non-host series have been brought to a common variance.
+The non-host chronology usually consists of a site chronology developed from non-host tree species growing on a climate-sensitive site, but a gridded climate data point series, like the North American Drought Atlas [@Cook2004] also suffices.
+If the host and non-host species respond similarly to climate (which can and should be tested), the derived series retains variability that the host and non-host series do not have in common, generally the insect signal and some unexplained variability (noise).
 The user defines a rule base specifying the magnitude and duration that a period of indexed growth suppression must meet or surpass for a period of suppressed growth to be inferred as a defoliation event at the tree level.
@@
 It was written in FORTRAN V with inherently severe restrictions, as RAM and disk space were limited at that time (256 kb and 10 MB, respectively) and FORTRAN conventions imposed very strict formatting, file naming, and output conventions.
-The program provides no graphical interface or capabilities, forcing users to import generated text files into spreadsheet or other software for assessing results and performing analyses.
+The program lacks a graphical interface or capabilities, forcing users to import generated text files into spreadsheets or other software to assess results and performing analyses.
 Furthermore, OUTBREAK can only handle one test at a time, creating barriers to batch operation and a large burden for researchers with a datasets including multiple sites.
-We developed the software library `dfoliaR` as an `R`- and `dplR`-based routine to overcome these issues.
+We developed `dfoliatR` as an `R`- and `dplR`-based library to overcome these issues.
 
-`dfoliatR` adds to a growing suite of dendrochronology packages the `R` computing environment [@RCore].
-Stemming from the `dplR` library [@Bunn2008] that enables `R` users to read and write an array of tree-ring data formats, standardize ring width series, build and evaluate chronologies, and perform quality control (to name a few), one can now also measure ring widths from scanned images of prepared samples [@Shi2019; @Lara2015], perform and check crossdating [@Bunn2010], and perform many analytical tests [@Jevsenak2018; @Zang2015].
+`dfoliatR` adds to a growing suite of dendrochronology packages in the `R` computing environment [@RCore].
+Stemming from the `dplR` library [@Bunn2008] that enables `R` users to read and write an array of tree-ring data formats, standardize ring width series, build and evaluate chronologies, and perform quality control (to name a few), one can now also measure ring widths from scanned images of prepared samples [@Shi2019; @Lara2015], conduct and check crossdating [@Bunn2010], analyze sub-annual anotomical features [@Campelo2016; @Rathgeber2011], and perform many analytical tests [@Jevsenak2018; @Zang2015].
 Tools for assessing stand dynamics and disturbance analyses are under rapid development, with new packages for assessing growth and release events [`TRADER`: @Altman2014], metrics of growth resilience [`pointRes`: @VanderMaaten-Theunissen2015], and fire history [`burnr`: @Malevich2018].
@@
 
-In this paper, we describe the statistical methods employed by `dfoliatR`, compare results to those produced by OUTBREAK, and present an example analysis.
+In this paper, we describe the statistical methods employed by `dfoliatR`, compare results to those produced by OUTBREAK, and present an example analysis inlcuding test data sets and script.
 Users need not have much experience in `R` to replicate the analyses and graphics as presented.
 The `R` code below is executable in an `R` session once the required libraries are installed and loaded.
-Support documentation in addition to this paper is provided within the package via standard help menus (accessed by typing `?` before a function name) and on the package website ([https://chguiterman.github.io/dfoliatR/](https://chguiterman.github.io/dfoliatR/)), which includes up-to-date vignettes that describe various software routines.
-Code to generate a preprint of this manuscript including the `R` scripts, and tubular and graphical output is available from [https://github.com/chguiterman/dfoliatR_paper](https://github.com/chguiterman/dfoliatR_paper).
+Support documentation in addition to this paper is provided within the package via standard help menus and on the package website ([https://chguiterman.github.io/dfoliatR/](https://chguiterman.github.io/dfoliatR/)), which includes up-to-date vignettes that describe various routines.
+Code to generate a preprint of this manuscript, including the `R` scripts and tubular and graphical output is available from [https://github.com/chguiterman/dfoliatR_paper](https://github.com/chguiterman/dfoliatR_paper).
 
@@
 * Standardized ring-width series for individual trees of the host species
-*	Standardized tree-ring chronology from a local non-host species
+*	A standardized tree-ring chronology from a local non-host species, or a climate reconstruction
 
-Users can develop these data sets in software of their choosing, such as `dplR` [@Bunn2008] or ARSTAN [@arstan].
+Users can develop these data sets in the software of their choosing, such as `dplR` or ARSTAN [@arstan].
 It is important that the host-tree data include only one tree-ring series per tree.
@@
 The climatic or non-defoliation signals in each host-tree series are characterized by a non-host chronology or climate reconstruction.
-`dfoiatR` removes that non-defoliation signal by subtracting the non-host series from each host-tree series, which generates a residual index.
-In OUTBREAK, this residual index was termed the "corrected" index.
-We call it the "growth suppression index" (GSI).
+`dfoliatR` removes the non-defoliation signal by subtracting the non-host series from each host-tree series, which generates a residual index.
+In OUTBREAK, this residual index was termed the "corrected index." We call it the "growth suppression index" (GSI).
 The GSI is calculated the same as in OUTBREAK for each host tree as
@@
 Only the common period between the host-tree series and the non-host chronology are used in Equation 1.
-The host and non-host chronologies are brought to common variance by scaling the non-host chronology by its mean ($\overline{\textrm{NH}} \approx 1.0$) and multiplying by the ratio of host and non-host standard deviations ($\frac{\sigma_{\textrm{H}}}{\sigma_{\textrm{NH}}}$), which approximates the variance of the host tree series.
+The host and non-host chronologies are brought to common variance by scaling the non-host chronology by its mean ($\overline{\textrm{NH}}$) and multiplying by the ratio of host and non-host standard deviations ($\frac{\sigma_{\textrm{H}}}{\sigma_{\textrm{NH}}}$), which approximates the variance of the host tree series.
 
-Negative departures in the normalized GSI (NGSI) that surpass user-defined thresholds in duration and magnitude are defined as *defoliation events*.
-As in OUTBREAK, magnitude is assessed on a single year within the departure sequence.
-The default setting is -1.28 (in units of standard deviation), which was previously determined to be representative of WSBW effects [@Swetnam1989].
-Duration is assessed by examining sequences of negative NGSI before and after the year of maximum departure.
-Each defoliation event is allowed one positive excursion on each side of the maximum departure year.
-Duration is assessed across the entire sequence that includes up to two positive excursions.
-The default duration is eight years, as is commonly used in WSBW studies [@Swetnam1989].
-Different species of defoliation insects vary in the length of defoliation and the degree to which they can suppress tree growth.
-Researchers can, and should, adjust the duration and magnitude parameters accordingly and critically evaluate the results.
+Negative departures in the normalized GSI (NGSI; GSI converted to z-scores) that surpass user-specified thresholds in duration and magnitude are defined as *defoliation events*.
+As in OUTBREAK, NGSI must reach the magnitude threshold, assessed for each year within the departure sequence.
+The default setting is -1.28 (NGSI is in units of standard deviation), which was previously determined to be representative of WSBW effects (Swetnam and Lynch, 1989) and is commonly used for other species.
+Failure to reach the specified threshold results in sequences not being inferred as defoliation events.
 
-Diverging from OUTBREAK, `dfoliatR` allows users to extend defoliation events by bridging successive events and also by allowing potentially short-duration events that occur at the end of the series.
-In cases where two defoliation events are separated by a single year, bridging will link them into a single event (**Show figure?**).
-We urge careful use of this option because there is no setting to limit the number or length of potentially bridged events.
-The series end option can be used in cases when the host trees were actively being defoliated at the time of sampling.
-This option eliminates the duration parameter for an event at the recent end of the series, but all other thresholds apply.
-The advantage of this parameter is that it can aid in identifying the start-year for the current defoliation event or outbreak, which is both useful in management and allows the current event to be included in return-interval estimates.
+Event duration is assessed by examining sequences of negative NGSI (that exceed the magnitude threshold) before and after the year of maximum departure.
+Each defoliation event is allowed one single-year positive excursion on each side of the year of maximum departure.
+Duration is computed across the entire negative sequence that allows these two positive excursions.
+As in OUTBREAK, the user specifies a duration threshold (minimum number of years) for a departure sequence to be inferred as a defoliation event.
+The default threshold is eight years, as is commonly used in WSBW studies (Swetnam and Lynch, 1989).
+Failure to reach the specified duration threshold results in sequences not being inferred as defoliation events, regardless of NGSI values (i.e., both thresholds must be met).
+Researchers can, and should, adjust the duration and magnitude parameters accordingly and critically evaluate the results, as insect species vary in the length of their outbreaks and the degree to which they can suppress tree growth.
+
+Like OUTBREAK, users are provided an option to suspend the duration threshold at the recent end of the series in cases where an outbreak event is known to be ongoing.
+This should be used if the user has direct knowledge of defoliation at the site during the sampling campaign.
+The advantage of allowing potentially short, series-end events is that it allows the current event to be included in return-interval estimates, and can aid in identifying the start-year for the current defoliation event or outbreak.
+
+Diverging from OUTBREAK, `dfoliatR` allows users to extend defoliation events on individual trees by bridging between sequential events.
+In cases where two defoliation events are separated by a single year, bridging will link them into a single event.
+This option was added to `dfoliatR` during the testing phase of development, when we realized that OUTBREAK deliberately omits sequential, or back-to-back events, even when these events surpass thresholds in duration and magnitude.
+Instead, OUTBREAK will select the one sequential event with the greatest negative departure year.
+In every case we assessed (described below) we felt that the OUTBREAK-omitted defoliation events should have been maintained and recorded.
+Due in large part to reconstructions using OUTBREAK [see papers cited by @Lynch2012], we now know considerably more about forest defoliator outbreak regimes than we did in the 1980s when OUTBREAK was written.
+We think that two or more prolonged events separated by a single year should in some situations be considered a single event.
+This is particularly relevant to WSBW and spruce budworm (*C.
+fumiferana*), for which regimes have been reconstructed for several geographic areas and for which researchers and forest health experts have gained considerable knowledge [@Schmitt1984; @Sanders1985; @Brookes1987, and many later publications].
+In many cases, the greatest growth suppression often occurs late in the outbreak due to cumulative effects accrued on a tree's resources [@Swetnam1989; @Swetnam1993].
+We urge caution in using the bridging option, however, because it may not be appropriate for all studied insects, such as in situations where impacted stands barely recover from one outbreak before another begins, as with pine processionary caterpillars (*Thaumetopoea pityocampa* (Lepidoptera: Thaumetopoeidae)) [@Carus2004; @Carus2009].
 
@@
 The second two parameters adjust the scale of defoliation considered to be an outbreak.
-Absolute numbers of trees and percentages can be applied separately or in conjunction, following filtering conventions in tree ring fire history studies [@Malevich2018].
+Absolute numbers of trees and percentages can be applied separately or in conjunction, following filtering conventions in tree-ring fire history studies [@Malevich2018].
 We urge users to carefully consider the choice of absolute numbers in situations where the number of trees represented in the series varies with time, or the choice of percentages when sample size is small.
@@
 
-We tested `dfoliatR` against OUTBREAK by comparing NGSI to corrected indices for individual trees and years, defoliation status for individual trees and years, and percentage of trees recording outbreaks at the site level, using standardized ring-width data from eight sites in British Columbia, Colorado, and New Mexico, and author-provided non-host site chronologies.
+We tested `dfoliatR` against OUTBREAK by comparing NGSI to OUTBREAK's normalized corrected indices for individual trees and years, defoliation status for individual trees and years, and percentage of trees recording outbreaks at the site level, using standardized ring-width data from eight sites in British Columbia, Colorado, and New Mexico, and author-provided non-host site chronologies.
 
-We detrended host data for both `dfoliatR` and OUTBREAK using ARSTAN v6.1 [@arstan] with default double detrending (128 year wavelength and a 50% smoothing spline).
-In both `dfoliatR` and OUTBREAK we used  event thresholds of -1.28 normalized index, 8 years duration, and allowed for events at the end of series in seven of eight sites.
+We detrended host data for both `dfoliatR` and OUTBREAK using ARSTAN [v6.1, @arstan] with default double detrending (128 year wavelength and a 50% smoothing spline).
+In both `dfoliatR` and OUTBREAK we used event thresholds of -1.28 normalized indices, 8 years duration, and allowed for events at the end of series in seven of eight sites.
 These seven sites were sampled during ongoing outbreak events [@Swetnam1993].
-We found it necessary to be consistent in how we detrended and in what software (e.g., ARSTAN vs `dplR`) we employed because subtle differences in standardized ring-width indices created between the programs exacerbated differences between `dfoliatR` and OUTBREAK.
+We found it necessary to be consistent in how we detrended and in what software (e.g., ARSTAN vs `dplR`) we employed because subtle differences in standardized ring-width indices generated between the programs transferred into differences between `dfoliatR` and OUTBREAK.
 In the end, we chose to only use the standardization output files from ARSTAN, which are easily read into `R` (and then `dfoliatR`) using the `dplR` package.
@@
 
-Since both programs employ Equation 1, `dfoliatR` and OUTBREAK compute identical growth suppression indices at 0.000 precision (Figure \@ref(fig:df-obr-1-1)).
-In all there are 43,280 ring-width indices from 222 trees included in our test.
-Of these, the programs identified 11,530 years of defoliation.
-The programs agreed on 92% of the years, leaving 927 "difference" years in which only one program identified defoliation on an individual tree.
+Across the 43,280 ring-width indices from 222 trees included in our evaluation, we found that `dfoliatR` and OUTBREAK compute identical growth suppression indices at 0.000 precision.
+We expected this outcome because both programs employ Equation 1.
+At the tree-level, the programs identified 11,530 years of defoliation.
+The programs agreed on 97.9% of the years, leaving 927 "difference" years in which only one program identified defoliation on an individual tree.
 The differences included 102 events on 85 trees.
-We carefully inspected each of these events in the full context of each tree's ring series, and categorized the differences as follows
+We carefully inspected each of these events in the full context of each tree's ring-series, and categorized the differences as follows
 
-* *Series-end events* (40% of the total) in which we allowed OUTBREAK to include "truncated outbreaks" (for seven sites) at the end of each series.
+* *Series-end events* (40% of the total) in which OUTBREAK included "truncated outbreaks" (for seven sites) at the end of each series.
 In `dfoliatR`, this option is controlled by the "series_end_events" parameter to `defoliate_trees()`.
 In OUTBREAK, the option appears while changing the duration parameter (option 3).
-When selected, OUTBREAK will include any sequences of negative indices at the either the beginning or the end of each tree series.
-No thresholds are applied in identifying these as defoliation events.
-In `dfoliatR`, the negative sequences are evaluated as with any other possible defoliation, but the duration threshold is omitted.
+When selected, OUTBREAK will include any sequences of negative indices at the either the beginning or the end of each tree series as a defoliation event, without consideration of either duration or magnitude thresholds.
+In `dfoliatR`, the duration threshold is omitted and the magnitude threshold is retained in series-end-events.
 Each of the 13 events included in these differences did not meet the "max_reduction" parameter (-1.28 NGSI) in `dfoliatR` and were excluded.
 In two cases, OUTBREAK included events at the beginning of the series where `dfoliatR` does not allow truncated events.
-In four cases, OUTBREAK omitted the last year of the series because it was positive.
+In four cases, OUTBREAK omitted only the last year of the series because the index was positive.
 In two cases, `dfoliatR` omitted possible events because it had already included a positive NGSI excursion after the "max_reduction" year, and since it will only allow one excursion on either side of the max year, the events were omitted due to short duration.
@@
 * *Sequential events* (36%) in which OUTBREAK omitted events that occur one year prior or one year following an identified event.
-OUTBREAK selects the event with the greatest negative index year.
+When this occurs, OUTBREAK selects the event with the greatest negative index year.
 On two trees, OUTBREAK omitted two of three sequential events.
-`dfoliatR` allows all such events.
-While inspecting these differences, we added an option to `defoiate_trees()` that would "bridge" between sequential events, making them into single, long events.
-We felt that this was ecologically justified, especially for studies of WSBW, because the insect will occasionally maintain minor defoliation f individuals for decades even while not in an outbreak phase.
+While inspecting these differences, we added an option to `defoliate_trees()` that would "bridge" between sequential events (that surpass magnitude and duration thresholds)into single, long events.
+We felt that this was ecologically justified, especially for studies of WSBW, because the insect will occasionally defoliate individual trees for decades and reconstructed chronologies show outbreaks as long as **xxx (ANN CHECK)**.
 
-* *Undetermined differences* (22%) occurred in cases where OUTBREAK omitted events without clear cause.
+* *Undetermined differences* (22%) occurred in cases where OUTBREAK omitted events without clear cause that `dfoliatR` correctly identified as defoliations.
 
-* *Rounding* (2%) differences in the indices either omitted or cut short events on two trees. In both cases the indices were very close to zero.
+* *Rounding differences* (2%) in the indices either omitted or cut short events on two trees.
+In both cases the indices were very close to zero, and the difference was less than the precision of the raw data measurement.
 
-Our inspections revealed what we believe are short-comings in how OUTBREAK identifies defoliation events on individual trees.
+**(Bridging paragraph?)**
+
+At the site level, OUTBREAK and `dfoliatR` produce similar time series of percent trees defoliated (Figure \@ref(fig:df-obr-site)), which forms the basis for inferring outbreak occurrence, intensity, and duration.
+Nearly all of the differences between the programs include either more trees or a longer duration of inferred outbreak by `dfoliatR`.
+These arise from the inclusion of events omitted by OUTBREAK.
+These differences translate to greater numbers of trees in defoliation during certain events, or events recorded by single trees during periods of low sample depth.
+
+This evaluation revealed what we believe are short-comings in how OUTBREAK identifies defoliation events on individual trees.
 In every one of the 102 cases we inspected, we felt that `dfoliatR` provided a more biologically and statistically appropriate assessment of defoliation.
-
-and use them to infer identical tree-level outbreak events, including onset and termination dates and date and magnitude of the maximum growth suppression when `dfoliatR` bridging matches the OUTBREAK protocols.
-Once the issues with ARSTAN (described earlier) were resolved, dfoliatR and OUTBREAK produced identical indices at 0.000 precision.
-
-When inferring outbreak events at the tree level, our initial results did not always match OUTBREAK, leading to development of the bridging option.
-Identifying a minimum within a period while allowing for positive excursions is a tricky computation.
-OUTBREAK deliberately disallows back-to-back events – after a minimum value within a period of negative values, a second positive index terminates an event.
-OUTBREAK does not allow a period of highly negative values to be appended to an earlier outbreak if the intervening period is 1 year, even if those negative values fall below the minimum in the earlier period.
-Due in large part to reconstructions using OUTBREAK [see papers cited by @Lynch2012], we now know considerably more about forest defoliator outbreak regimes than we did in the 1980s when OUTBREAK was written.
-We think that two prolonged events separated by a single year should in some situations be considered a single outbreak at both the tree and site levels.
-This is particularly relevant to western spruce budworm and spruce budworm (*C.
-fumiferana*), for which regimes have been reconstructed for several geographic areas and for which researchers and forest health experts have gained considerable knowledge (**Sanders et al.
-1985, Brookes et al.
-1987 and many later publications**).
-We now know that the greatest growth suppression often occurs late in the outbreak due to cumulative effects (**rrr**).
-We developed the bridging option to permit these periods to be linked to preceding events separated by a single year.
-
-Bridging operates at the tree level in `dfoliatR`, and these effects generally have minor effects on site-level reconstructions (**Fig.Bridging**).
-Occasionally site-level events are inferred differently (**Table.Bridging**), and this can affect return interval and duration statistics.
-
-## Best Practice Recommendations
-
-We encourage authors to report software and library version numbers and download dates where applicable.
-This is standard practice for statistical packages, but has often been ignored for R- and dplR-based analyses.
-R-based software evolves over time, and results cannot always be replicated if the software versions are unknown.
-Similarly, significant changes (especially fixes) imply that re-analysis of data might differ from earlier published results and interpretations.
-Wherever possible, refereed publications supporting the software should be cited.
-Our experience with ARSTAN demonstrates the importance of adopting this practice.
-In the United States, analytic methods not documented thusly and supported by the literature might be questioned if the results are used in a National Environmental Protection Act (NEPA) process.
-
-We recommend that researchers compare dfoliatR results with and without bridging, and see which option agrees best with known insect biology and ecology.
-Bridging may be a logical choice for insects with prolonged non-outbreak periods, such as western spruce budworm, but not for situations where impacted stands barely recover from one outbreak before another begins, such as occurs with pine processionary caterpillars (*Thaumetopoea pityocampa* (Lepidoptera: Thaumetopoeidae)) (**Carus 2004, 2009**).
+Therefore, `dfoliatR` also did a better job inferring OUTBREAK events and associated statistics at the site level.
 
@@
 
-In `dfoliatR` we provide two sets of tree-ring data to aid users in exploring the functions, graphics, and outputs.
-Each set consists of individual host-tree series, standardized using 50-year splines with a 50% frequency response, and a local non-host chronology.
-The non-host ring-width data were standardized using 150-year splines with a 50% frequency response.
-The non-host chronologies were averaged following standard procedures [@Speer2010].
+Once `dfoliatR` is loaded into an `R` session (via `library(dfoliatR)`) users can access two sets of tree-ring data to aid in exploring the functions, graphics, and outputs.
+Each data set consists of individual host-tree series and a local non-host chronology.
+The host-tree series were standardized using 50-year splines with a 50% frequency response, while the non-host ring-width data were standardized using 150-year splines with a 50% frequency response and then averaged via Tukey's biweight robust mean procedure.
 Host trees from Demijohn Peak (DMJ; 2902 m asl) in the San Juan Mountains of southern Colorado include Douglas-fir (*Pseudotsuga menziesii*) compared against a local non-host ponderosa pine (*Pinus ponderosa*) chronology [@Ryerson2003].
-The East Fork site (EF; 2580 m asl) in the Jemez Mountains of northcentral New Mexico include Douglas-fir and white fire (*Abies concolor*) host trees and a ponderosa pine non-host chronology [@Swetnam1993].
+The East Fork site (EFK; 2580 m asl) in the Jemez Mountains of northcentral New Mexico includes Douglas-fir and white fir (*Abies concolor*) host trees and a ponderosa pine non-host chronology [@Swetnam1993].
+
+With `dfoliatR` loaded, the datasets are accessible using the `data()` function.
+The data object names are prefixed by their site codes.
+For instance, the `dmj_*` objects come from the DMJ site and include the host-tree series (`dmj_h`), the non-host chronology (`dmj_nh`), the defoliation series (`dmj_defol`), and the outbreak series (`dmj_obr`).
+The same suite of data are available for EFK using the `efk_*` prefix.
+
+In our example scripts below, `##` and `#` denote user comments, per standard `R` coding, which are colored in brown.
+Text in blue denote functions; black are loaded objects, and green are quoted variables and links.
 
@@
 
-To assess defoliation of trees, we begin with a set of standardized host-tree series and a local non-host chronology or climate reconstruction.
-First, the `dfoliatR` library needs to be loaded (after installation as above) and the included datasets can be loaded.
 The function `defoliate_trees()` performs the GSI indexing procedure on each host-tree series and then identifies defoliation events.
@@
 
-The result of running `defoliate_trees()` is long-format (stacked) data frame with five variables: "year", "series", "gsi", "ngsi", and "defol_status." The "gsi" variable is the "corrected" tree-level growth suppression index.
-It was normalized or scaled into z-scores to produce "ngsi", which is used to identify defoliation events.
-The "defol_status" column provides a set of character strings, "nd" for non-defoliation year, "defol" for a defoliation year, "max_defol" for the year of maximum suppression (that acts as the basis for individual events), "bridge_defol" to identify years that link subsequent events (only one is present at DMJ), and "series_end_defol" to identify defoliation at the present-end of the series.
-Selecting `list_output = TRUE` provides a list-object of data frames, each with the `rwl` object that combines the host tree and non-host chronology and the columns created in `defoliate_trees()`.
-This option is not used by subsequent functions in `dfoliatR`, but researchers can examine it to check the results of the GSI calculation (Equation 1).
+The result is long-format (stacked) data frame with five variables: "year", "series", "gsi", "ngsi", and "defol_status." The "defol_status" column indicates whether that year has defoliation or not, with a set of factors that include "nd" for non-defoliation year, "defol" for a defoliation year, "max_defol" for the year of maximum suppression (that acts as the basis for individual events), "bridge_defol" to identify years that link subsequent events (only one is present at DMJ), and "series_end_defol" to identify defoliation at the present-end of the series.
 
-We can assess the results of running `defoliate_trees()` through graphical and table outputs.
+Selecting `list_output = TRUE` in `defoliate_trees()` provides a list-object of data frames, each with an `rwl` object that combines the host tree and non-host series and the other columns created by `defoliate_trees()`.
+This option is not used by subsequent functions in `dfoliatR`, but researchers can examine it to check the results of the GSI calculation (Equation 1), such as the non-host series after scaling to a common variance with a particular host-tree series.
+
+The results of running `defoliate_trees()` can be assessed through graphical and table outputs.
 The function `get_defol_events()` will provide a list of every defoliation event for every tree, with the corresponding mean "ngsi" value.
@@
 The defoliation segments are colored by their relative severity based on their average NGSI value.
-The default cut-off values between "Severe" and "Moderate" is the overall mean across all events and between "Moderate"  and "Minor" is the first quartile.
-Users can define these breaks to suit their needs.
+The default cut-off values between "Severe" and "Moderate" is the overall mean across all events.
+The cut-off between "Moderate"  and "Minor" is the first quartile by default.
+Users can re-define the breaks to suit their needs via the "breaks" parameter in `plot_defol()`.
 
@@
 plot_defol(dmj_defol)
-## This creates a ggplot object, so additions can 
+## This creates a ggplot object. Additions can 
 ## be made to adjust plotting themes and
 ## aesthetics, like color.   
-## For more on the features of ggplot:
+## For more on the features of ggplot see
 ## https://ggplot2.tidyverse.org/
+## Adding 'scale_color_manual()' will
+## change the colors that denote 
+## the average NGSI value for each defoliation
+## event.
 library(ggplot2)
@@
 
-These basic output functions aid in assessing the sensitivity of input parameters to `defoliate_trees()` including the duration and magnitude thresholds for identifying defoliation events.
-Using `plot_defol()` also provides a direct assessment of the between-tree variability in deflation.
-For further analyses, we composite the host-tree defoliation series to the site level using `outbreak()`.
+These output functions aid in assessing the sensitivity of input parameters to `defoliate_trees()`, including the duration and magnitude thresholds for identifying defoliation events.
+Using `plot_defol()` also provides a direct assessment of the between-tree variability in defoliation.
 
@@
 
-Once we have identified defoliation events on individual trees, we can infer outbreak events at the site level.
-The function `outbreak()` acts to composite tree-level defoliation series into a single chronology, with input parameters that control thresholds in the number and proportions of trees recording a defoliation event.
+To infer outbreak events at the site level, the function `outbreak()` composites tree-level defoliation series into a single chronology, with input parameters that control thresholds in the number and proportions of trees recording a defoliation event.
 
@@
 
-Input parameters to `outbreak()` include "filter_min_series" to control the chronology cut-off points with regard to sample depth, "filter_min_defol" and "filter_perc" to control the minimum number and percent of trees, respectively, recording a defoliation event in a given year.
+Input parameters to `outbreak()` include "filter_min_series" to control the chronology cut-off points with regard to sample depth, "filter_min_defol" and "filter_perc" to control the minimum number and percent of trees recording a defoliation event in a given year.
 `Outbreak()` produces a new data frame with eight variables: "year", "num_defol", "percent_defol", "num_max_defol", "mean_gsi", "mean_ngsi", and "outbreak_status." All of these variables are populated regardless of an inferred outbreak event, providing a continuous disturbance chronology.
@@
 
-The default plotting function to visualize results form `outbreak()` creates a three-panel graph showing the mean site-level chronology, the percent of trees recording a defoliation, and the sample depth over time (Figure \@ref(fig:fig-obr)).
+The default plotting function to visualize results from `outbreak()` creates a three-panel graph showing the mean site-level chronology, the percent of trees recording a defoliation, and the sample depth over time (Figure \@ref(fig:fig-obr)).
 
@@
 dmj_obr_stats <- outbreak_stats(dmj_obr)
+View(dmj_obr_stats)
 ```
 
-The output table from `outbreak_stats()` shows the start and end years of each event, along with their corresponding duration, the number and percent of trees in defoliation at the start of the event, the the number of trees recording the outbreak event (in part or full), the minimums of the mean GSI and NGSI indices, and two points of "peak defoliation." The first, "peak_outbreak_year," is the year in which the greatest number of trees is recording the outbreak, and the second, "peak_defol_year," is the year in with the lowest average NGSI.
-These output variables provide a array of options for assessing basic metrics of the insect defoliation regime, including interval analyses:
+The output table from `outbreak_stats()` shows the start and end years of each event, along with their corresponding duration, the number and percent of trees in defoliation at the start of the event, the the number of trees recording the outbreak event (in part or full), the minimums of the mean GSI and NGSI indices, and two points of "peak defoliation." The first, "peak_outbreak_year" is the year in which the greatest number of trees are recording the outbreak, and the second, "peak_defol_year" is the year with the lowest average NGSI.
+These output variables provide a array of options for assessing metrics of the insect defoliation regime, including interval analyses:
 
@@
 intervals <- diff(dmj_obr_stats$start)
-mean(intervals) ## 51.2
+mean(intervals) ## 51.2 year mean return interval
 ```
@@
 
-* Encourage use of `dfoliatR` over OUTBREAK -- more acruate, greater user control, more easily conduct sensitivity tests to evaluate paramters, data, and ecological patterns
-* Facilitates high-levelr statistical analyses by easily transfering to other libraries (MASS, etc.)
-* A community resource that can be easily maintained and updated into the future.
+The `dfoliatR` package provides dendroecologists with tools to infer, quantify, analyze, and visualize tree-ring growth suppression events and to reconstruct forest insect defoliator outbreak regimes.
+It is built on the long-accepted host to nonhost comparison methodology used in the 1980s FORTRAN program OUTBREAK [@Swetnam1985; @Swetnam1989; @Swetnam1993].
+Key benefits are that `dfoliatR` provides easier control of suppression thresholds, provides additional output tables and high-quality graphics, defines defoliation events more consistently, and operates in the open-source `R` environment.
+
+Using `dfoliatR` requires standardized ring-width measurements from insect host trees and either an indexed tree-ring chronology from local non-host trees or suitable climate chronology.
+It performs an indexing procedure to remove the climatic signal represented in the non-host chronology from the host-tree series.
+It then infers defoliation events in individual trees.
+Site-level analyses identify outbreak events that synchronously affect a user-defined number or proportion of the host trees.
+Functions are provided for summary statistics and graphics of tree- and site-level series.
+The package produces publication-quality plots, and tabulates growth suppression indices and tree- and site-level outbreak event statistics for user-defined post-processing needs, including those suitable for charting and tabulating landscape- and regional-level results.
+
+`dfoliatR` adds a new option for dendroentimology to combine, or "bridge," sequential tree-level defoliation events into single events.
+In practice, we suggest that researchers carefully evaluate if bridging is ecologically applicable to study situations and insect ecology, and to carefully explore the data before deciding whether or not to use this option.
+It is probably not appropriate for insects with high-frequency, high–severity outbreaks, such as processionary caterpillars.
+Outbreak reconstructions of insects for which the interval is notably longer than typical outbreak duration, such as Douglas-fir tussock moth (*Orgyia pseudotsugata* (Lepidoptera: Erebidae)), are unlikely to be significantly affected.
+Species for which outbreak duration varies considerably, for which individual tree resilience to defoliation varies considerably, or which may chronically infest trees or sites, or alternate between chronic, outbreak, and minimal activity states, such as conifer-feeding *Choristoneura*, present more complicated challenges, and the researcher should use discretion with the bridge option.
+
+`dfoliatR` adds to the on-going open-source software development for dendrochronological methods [e.g., @Bunn2008; @Brewer2014; @Brewer2016].
+The `R` environment enables automation of analyses, allowing input/output processes to become routine, and empowering batch processing of large multi-site projects and sensitivity analyses.
+It also facilitates additional statistical analyses, such as spectral analyses and superposed epoch analyses [e.g., @Malevich2018], with easy transfer from `dfoliatR` and `dplR` to other libraries in `R`.
+Source code for `dfoliatR` is available in the Comprehensive R Archive Network (CRAN) and GitHub [https://github.com/chguiterman/dfoliatR](https://github.com/chguiterman/dfoliatR) with updated descriptions and helpful vignettes on the package website [https://chguiterman.github.io/dfoliatR/](https://chguiterman.github.io/dfoliatR/).
+Researchers wishing to contribute to further development of `dfoliatR` are encouraged to do so via the GitHub repository.
 
@@
 We are grateful to late Richard Holmes of the Laboratory of Tree-Ring Research (LTRR) for his achievements in developing OUTBREAK and many other tree-ring software programs, and we thank Thomas Swetnam for his insights and encouragement.
-The `dfoliatR` project was supported by the Agnes Haury Visiting Scholars Fellowship at the LTRR and by the University of California Berkeley.
-Additional support was provided by the LTTR and the United States Forest Service Rocky Mountain Research Station.
+The `dfoliatR` project was supported by the Agnes Haury Visiting Scholars Fellowship at the LTRR, the University of California Berkeley, and the Western Wildlands Environmental Threat Center (WWETAC).
+Additional support was provided by the LTRR and the United States Forest Service Rocky Mountain Research Station.
 Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S.
@@
 ```
+
+``` {r df-obr-site, echo=FALSE, fig.cap="Comparison of reconstructed western spruce budworm outbreaks computed by dfoliatR and OUTBREAK. Input parameters were identical between programs."}
+include_graphics("../paper_elsevier/Output/plot-sites-df-obr.pdf")
+```

### 2020-05-05 paper_elsevier/paper_elsevier.tex
Ann edits and references

--- 2020-03-05 paper_elsevier/paper_elsevier.tex
@@
 
-  \journal{Tree-Ring Research} % Sets Journal name
+  \journal{Dendrochronologia} % Sets Journal name
 
@@
 
-  \title{\emph{dfoliatR}: An R package for detection and analysis of insect
-defoliation signals in tree rings}
+  \title{\emph{dfoliatR}: An R package for detection and analysis of insect defoliation signals in tree rings}
     \author[a,b]{Christopher H. Guiterman\corref{1}}
-   \ead{chguiterman@email.arizona.edu} 
+   \ead{chguiterman@arizona.edu} 
     \author[a,c]{Ann M. Lynch}
-   \ead{lyncha@email.arizona.edu} 
+   \ead{ann.m.lynch@usda.gov} 
     \author[c]{Jodi N. Axelson}
    \ead{jodi.axelson@berkeley.edu} 
-      \address[a]{Laboratory of Tree-Ring Research, University of Arizona, 1215 E Lowell
-St.~Box 210045, Tucson, AZ, 85721}
+      \address[a]{Laboratory of Tree-Ring Research, University of Arizona, 1215 E Lowell St.~Box 210045, Tucson, AZ, 85721}
     \address[b]{Three Pines Forest Research, LLC, PO Box 225, Etna, NH, 03750}
-    \address[c]{U.S. Forest Service, Rocky Mountain Research Station, 1215 E Lowell
-St.~Box 210045, Tucson, AZ, 85721}
-    \address[d]{Dept of Environmental Science, Policy \& Management, University of
-California, Berkeley, Berkeley, CA 94720}
+    \address[c]{U.S. Forest Service, Rocky Mountain Research Station, 1215 E Lowell St.~Box 210045, Tucson, AZ, 85721}
+    \address[d]{Dept of Environmental Science, Policy \& Management, University of California, Berkeley, Berkeley, CA 94720}
       \cortext[1]{Corresponding Author}
@@
   \begin{abstract}
-  We present a new R package to provide dendroecologists with tools to
-  identify, quantify, analyze, and visualize growth suppression events in
-  tree rings produced by insect defoliation. The `dfoliatR' library is
-  based on the Fortran V program OUTBREAK, and builds on existing
-  resources in the R computing environment. `dfoliatR' expands on OUTBREAK
-  to provide greater control of supression thresholds, additional output
-  tables, and high-quality graphics. To use `dfoliatR' requires
-  standardized ring-width measurements from insect host trees and an
-  indexed tree-ring chronology from local non-host trees. It performs an
-  indexing procedure to remove the climatic signal represented in the
-  non-host chronology from the host-tree series. It then infers
-  defoliation events in individual trees. Site-level analyses identify
-  outbreak events that synchronously affect a user-defined number or
-  proportion of the host trees. Functions are available for summary
-  statistics and graphics of tree- and site-level series.
+  We present a new R package to provide dendroecologists with tools to infer, quantify, analyze, and visualize growth suppression events in tree rings caused by insect defoliation, and to reconstruct defoliator outbreak chronologies.
+`dfoliatR' is based on the OUTBREAK program and builds on existing resources in the R and dplR computing environments.
+It performs an indexing procedure to remove climatic signals, represented in non-host chronologies or other reconstructed climate series, from the host-tree series.
+It then infers defoliation events in individual trees based on user-specified thresholds.
+Site-level analyses identify outbreak events that synchronously affect user-defined numbers or proportions of involved host trees.
+Functions are provided for summary statistics and graphics of tree- and site-level series.
+`dfoliatR' expands on OUTBREAK to provide greater control of suppression thresholds, additional output tables, and high-quality graphics.
+We provide two example data sets and script to enable users to gain familiarity with the package and its capabilities.
+The source code is available in the Comprehensive R Archive Network (CRAN) and on GitHub.
 
@@
 
-The \texttt{dfoliatR} library adds to a growing suite \texttt{R}
-packages designed for dendrochronology. Stemming from the \texttt{dplR}
-library (Bunn 2008) that enables \texttt{R} users to read and write an
-array of tree-ring data formats, standaridize ring widths, build and
-evaluate chronologies, perform qaulity control (to name a few), one can
-now also measure ring widths (Lara et al. 2015, Shi and Xiang 2019),
-perform and check crossdating (Bunn 2010), and perform many analytical
-tests (Zang and Biondi 2015, Jevšenak and Levanič 2018). Tools for
-assessing stand dynamics and disturbance analyses are under rapid
-development, with new packages for assessing growth and release events
-(\texttt{TRADER}: Altman et al. 2014), metrics of growth resilience
-(\texttt{pointRes}: Maaten-Theunissen et al. 2015), and fire history
-(\texttt{burnr}: Malevich et al. 2018). The key objective of
-\texttt{dfoliatR} is to provide tools to identify and analyse insect
-defoliation and outbreak events by building on the methods employed the
-FORTRAN program OUTBREAK (Swetnam and Lynch 1989). What sets
-\texttt{dfoliatR} apart from packages such as \texttt{TRADER} is that it
-explicitly performs an indexing procedue on host-tree series to remove
-climatic and other non-defoliation related signals represented by
-non-host tree species. Insect defoliation signals are identified in the
-disturbance index by the duration and magnitude of negative departures.
+Variation in the width and morphology of annual radial growth rings in trees permits dating and quantification of past forest insect defoliator outbreaks.
+Defoliation can be distinguished from climate- and other disturbance-related influences by comparing ring-width or other annually-resolved features in the wood of host species to that of non-host species or annually-resolved climate records.
+The effect of defoliation on radial growth of trees has been recognized since the 1860s, and used to reconstruct outbreak regimes since the 1950s (Blais 1954; Lynch 2012).
+It was not until the 1980s, however, that precise dendrochronological techniques were applied for inferring defoliation events and reconstructing defoliator outbreak regimes (Swetnam, Thompson, and Sutherland 1985; Speer 2010; Lynch 2012).
+The first studies (Swetnam, Thompson, and Sutherland 1985; Swetnam and Lynch 1989, 1993) focused on developing historical reconstructions of western spruce budworm (WSBW; \emph{Choristoneura freemani} (Lepidoptera: Tortricidae); previously known as \emph{C.
+occidentalis}) outbreak regimes.
+The methodology has since been successfully applied to a wide range of defoliator species, most of which are conifer herbivores, and has evolved in sophistication for a wide range of ecosystem situations (Lynch 2012).
 
-\texttt{dfoliatR} draws upon data formats in \texttt{dplR} that are
-commonly employed by other tree-ring libraries. It uses and outputs data
-formats that faciliate the use packages embodied by \texttt{tidyverse}
-(Wickham et al. 2019) that include efficient data manipulation
-(\texttt{dplyr}: Wickham et al. 2020) and graphics (\texttt{ggplot2}:
-Wickham 2016).
+The main dendrochronological tool for inferring, dating, and characterizing defoliator outbreaks from tree-ring records has been the software routine OUTBREAK (Swetnam, Thompson, and Sutherland 1985; Holmes and Swetnam 1986; Swetnam and Lynch 1989).
+OUTBREAK computes indices (described later in detail) of suppressed growth by subtracting a detrended and standardized climate series (a ``control'' chronology) from individual host-tree detrended and standardized radial growth series after the host and non-host series have been brought to a common variance.
+The non-host chronology usually consists of a site chronology developed from non-host tree species growing on a climate-sensitive site, but a gridded climate data point series, like the North American Drought Atlas (Cook and Krusic 2004) also suffices.
+If the host and non-host species respond similarly to climate (which can and should be tested), the derived series retains variability that the host and non-host series do not have in common, generally the insect signal and some unexplained variability (noise).
+The user defines a rule base specifying the magnitude and duration that a period of indexed growth suppression must meet or surpass for a period of suppressed growth to be inferred as a defoliation event at the tree level.
 
-In this paper, we describe the statistical methods employed by
-\texttt{dfoliatR}, its availability, and run through analyses for a
-signle site in New Mexico. Users need not have much experience in
-\texttt{R} to replicate the analyses and graphics below. All \texttt{R}
-code presented below is executatable in an \texttt{R} session once the
-required libraries are installed and loaded. Support documentation in
-addition to this paper is provided within the package via standard help
-menus (accessed by typing \texttt{?} before a function name) and on the
-package website (\url{https://chguiterman.github.io/dfoliatR/}), which
-includes up-to-date vignettes that describe in detail the functionality
-of the software. Code to create a preprint of this manuscript including
-the \texttt{R} scripts is available from
-\url{https://github.com/chguiterman/dfoliatR_paper}.
+Though powerful, OUTBREAK is outdated and increasingly difficult to use in modern computing environments.
+It was written in FORTRAN V with inherently severe restrictions, as RAM and disk space were limited at that time (256 kb and 10 MB, respectively) and FORTRAN conventions imposed very strict formatting, file naming, and output conventions.
+The program lacks a graphical interface or capabilities, forcing users to import generated text files into spreadsheets or other software to assess results and performing analyses.
+Furthermore, OUTBREAK can only handle one test at a time, creating barriers to batch operation and a large burden for researchers with a datasets including multiple sites.
+We developed \texttt{dfoliatR} as an \texttt{R}- and \texttt{dplR}-based library to overcome these issues.
+
+\texttt{dfoliatR} adds to a growing suite of dendrochronology packages in the \texttt{R} computing environment (R Core Team 2019).
+Stemming from the \texttt{dplR} library (Bunn 2008) that enables \texttt{R} users to read and write an array of tree-ring data formats, standardize ring width series, build and evaluate chronologies, and perform quality control (to name a few), one can now also measure ring widths from scanned images of prepared samples (Shi et al.
+2019; Lara, Bravo, and Sierra 2015), conduct and check crossdating (Bunn 2010), analyze sub-annual anotomical features (Campelo et al.
+2016; Rathgeber et al.
+2011), and perform many analytical tests (Jevšenak and Levanič 2018; Zang and Biondi 2015).
+Tools for assessing stand dynamics and disturbance analyses are under rapid development, with new packages for assessing growth and release events (\texttt{TRADER}: Altman et al.
+2014), metrics of growth resilience (\texttt{pointRes}: van der Maaten-Theunissen, van der Maaten, and Bouriaud 2015), and fire history (\texttt{burnr}: Malevich, Guiterman, and Margolis 2018).
+The key objective of \texttt{dfoliatR} is to provide tools to identify and analyze insect defoliation and outbreak events by building on the methods employed by OUTBREAK.
+It capitalizes on the robust software already available in \texttt{R} by using \texttt{dplR} data formats for incoming tree-ring series and providing output data formats embodied by the \texttt{tidyverse} (Wickham et al.
+2019) that include efficient data manipulation (Wickham et al.
+2020) and graphics (Wickham 2016).
+
+In this paper, we describe the statistical methods employed by \texttt{dfoliatR}, compare results to those produced by OUTBREAK, and present an example analysis inlcuding test data sets and script.
+Users need not have much experience in \texttt{R} to replicate the analyses and graphics as presented.
+The \texttt{R} code below is executable in an \texttt{R} session once the required libraries are installed and loaded.
+Support documentation in addition to this paper is provided within the package via standard help menus and on the package website (\url{https://chguiterman.github.io/dfoliatR/}), which includes up-to-date vignettes that describe various routines.
+Code to generate a preprint of this manuscript, including the \texttt{R} scripts and tubular and graphical output is available from \url{https://github.com/chguiterman/dfoliatR_paper}.
 
@@
 
-The \texttt{dfoliatR} library requires two sets of tree-ring data to
-identify defoliation and outbreak events:
+The \texttt{dfoliatR} library requires two sets of tree-ring data to infer defoliation and outbreak events:
 
@@
 \item
-  Standardized ring-width series for individual trees of the host
-  species
+  Standardized ring-width series for individual trees of the host species
 \item
-  Standardized tree-ring chronology from a local non-host species
+  A standardized tree-ring chronology from a local non-host species, or a climate reconstruction
 \end{itemize}
 
-Users can develop these data sets in software of their choosing, such as
-\texttt{dplR} or ARSTAN. It is important that the host-tree data include
-only one tree-ring series per tree. Both \texttt{dplR} and ARSTAN have
-options for averaging mutliple sample series into a tree-level series.
-The tree-ring series and chronology can be read into \texttt{R} via
-several available \texttt{dplR} functions.
+Users can develop these data sets in the software of their choosing, such as \texttt{dplR} or ARSTAN (Cook and Holmes 1996).
+It is important that the host-tree data include only one tree-ring series per tree.
+Both \texttt{dplR} and ARSTAN have options for averaging multiple sample series into a tree-level series.
 
-At the hear of \texttt{dfoliatR} lies two functions:
-\texttt{defoliate\_trees()} and \texttt{outbreak()}. These identify
-defoliation event son individual trees and thenc omposite across
-mutliple trees for a given site to identify outbreak events. It is up to
-the researcher to distinguish these separate spatial scales of analysis,
-\texttt{dfoliatR} then provides a set of functions to visualize and
-statistically summarize tree-level defoliation and site-level outbreak
-events, which will be discussed further.
+At the heart of \texttt{dfoliatR} lies two functions: \texttt{defoliate\_trees()} and \texttt{outbreak()}.
+These identify defoliation events on individual trees (Figure \ref{fig:fig-defol}) and then composite across multiple trees to identify outbreak events (Figure \ref{fig:fig-obr}).
 
 \hypertarget{identifying-defoliation-of-trees}{%
-\subsection{Identifying Defoliation of
-Trees}\label{identifying-defoliation-of-trees}}
+\subsection{Identifying Defoliation of Trees}\label{identifying-defoliation-of-trees}}
 
-The \texttt{defoliate\_trees()} function will usually be used first. It
-performs two processes: removing the non-defoliatation growth signal
-from the host-tree series and identifying defoliation events for each
-host tree. The climatic or non-defoliation signals in each host-tree
-series is represented by the non-host chronology or a climate
-reconstruction. This indexing procedure generates the ``growth
-suppression index'' (GSI) in which disturbance is the predominant
-signal. The GSI is calculated as in Swetnam et al. (1985) and Swetnam
-and Lynch (1989), \begin{align}
+The \texttt{defoliate\_trees()} function is the point of entry to the \texttt{dfoliatR} library.
+It performs two processes, removing climate-related growth signals from the host-tree series and identifying tree-level defoliation events.
+The climatic or non-defoliation signals in each host-tree series are characterized by a non-host chronology or climate reconstruction.
+\texttt{dfoliatR} removes the non-defoliation signal by subtracting the non-host series from each host-tree series, which generates a residual index.
+In OUTBREAK, this residual index was termed the ``corrected index.'' We call it the ``growth suppression index'' (GSI).
+The GSI is calculated the same as in OUTBREAK for each host tree as
+\begin{align}
 \textrm{GSI}_{i} = \textrm{H}_{i} - \left( \textrm{NH}_{i} - \overline{\textrm{NH}} \right) \frac{\sigma_{\textrm{H}}}{\sigma_{\textrm{NH}}} \
-\end{align} where H and NH are the host tree series and the non-host
-chronology, in year i, respectively. Only the common period between each
-host-tree series and the non-host chronology are used in Equation 1. The
-non-host chronology is scaled by its mean
-(\(\overline{\textrm{NH}} \approx 1.0\)) and multiplied by the ratio of
-host and non-host standard deviations
-(\(\frac{\sigma_{\textrm{H}}}{\sigma_{\textrm{NH}}}\)), which
-approximates the variance of the host tree series. This ``corrected''
-non-host chronology is subtracted from the host-tree series.
+\end{align}
+where H and NH are the host-tree series and the non-host chronology, in year i, respectively.
+Only the common period between the host-tree series and the non-host chronology are used in Equation 1.
+The host and non-host chronologies are brought to common variance by scaling the non-host chronology by its mean (\(\overline{\textrm{NH}}\)) and multiplying by the ratio of host and non-host standard deviations (\(\frac{\sigma_{\textrm{H}}}{\sigma_{\textrm{NH}}}\)), which approximates the variance of the host tree series.
 
-Negative departures in the normalized GSI that surpass user-defined
-thresholds in duration and magnitude are defined as \emph{defoliation
-events}. As in OUTBREAK, defoliation events have a single year of
-maximum departure (the variable `max\_reduction' in
-\texttt{defoliate\_trees()}) that is used to assess the magnitude of
-growth suppression. The default setting is -1.28 (in units of standard
-deviation), which was previously determined as representative of WSBW
-effects (\textbf{cite}). Duration is assessed by examining sequences of
-negative GSI before and after the year of maximum departure. Each
-defoliation event is allowed one positive excersion on each side of the
-maximum departure year. Duration is assessed across the entire sequence
-that includes up to two positive excursions. The default duration is
-eight years, following OUTBREAK, as is commonly used in WSBW studies
-(\textbf{cite}). Defoliator insects vary in their durations of
-defoliations and outbreaks, so researchers should exmaine their data and
-perform sensitivity tests to to find a representative event duration.
+Negative departures in the normalized GSI (NGSI; GSI converted to z-scores) that surpass user-specified thresholds in duration and magnitude are defined as \emph{defoliation events}.
+As in OUTBREAK, NGSI must reach the magnitude threshold, assessed for each year within the departure sequence.
+The default setting is -1.28 (NGSI is in units of standard deviation), which was previously determined to be representative of WSBW effects (Swetnam and Lynch, 1989) and is commonly used for other species.
+Failure to reach the specified threshold results in sequences not being inferred as defoliation events.
 
-Researchers are given the options to extend defoliation events by
-bridging between successive events and allowing potentially
-short-duration events that occur at the end of the series. Bridging
-allows multiple events that are separated by a single year to be
-combined and considered as one defolaition event. Each may have up to
-five years with positive growth (\textbf{CHECK}). Series end events is
-an option that can be used when the trees were sampled during a known
-defoliation event, which in many cases is what led the to the site being
-sampled for the reconstruction. This option eliminates the duration
-parameter, but all other thresholds, and bridging, still apply. These
-options are not available in OUTBREAK.
+Event duration is assessed by examining sequences of negative NGSI (that exceed the magnitude threshold) before and after the year of maximum departure.
+Each defoliation event is allowed one single-year positive excursion on each side of the year of maximum departure.
+Duration is computed across the entire negative sequence that allows these two positive excursions.
+As in OUTBREAK, the user specifies a duration threshold (minimum number of years) for a departure sequence to be inferred as a defoliation event.
+The default threshold is eight years, as is commonly used in WSBW studies (Swetnam and Lynch, 1989).
+Failure to reach the specified duration threshold results in sequences not being inferred as defoliation events, regardless of NGSI values (i.e., both thresholds must be met).
+Researchers can, and should, adjust the duration and magnitude parameters accordingly and critically evaluate the results, as insect species vary in the length of their outbreaks and the degree to which they can suppress tree growth.
 
-\hypertarget{identifying-outbreak-events}{%
-\subsection{Identifying Outbreak
-Events}\label{identifying-outbreak-events}}
+Like OUTBREAK, users are provided an option to suspend the duration threshold at the recent end of the series in cases where an outbreak event is known to be ongoing.
+This should be used if the user has direct knowledge of defoliation at the site during the sampling campaign.
+The advantage of allowing potentially short, series-end events is that it allows the current event to be included in return-interval estimates, and can aid in identifying the start-year for the current defoliation event or outbreak.
 
-The individual tree defoliation series are composited in an additional
-step to identify \emph{outbreak events} that synchronously affect
-mutliple trees. Users have options to define the number and/or the
-proportion of trees required for a defoliation event to be considered an
-outbreak.
+Diverging from OUTBREAK, \texttt{dfoliatR} allows users to extend defoliation events on individual trees by bridging between sequential events.
+In cases where two defoliation events are separated by a single year, bridging will link them into a single event.
+This option was added to \texttt{dfoliatR} during the testing phase of development, when we realized that OUTBREAK deliberately omits sequential, or back-to-back events, even when these events surpass thresholds in duration and magnitude.
+Instead, OUTBREAK will select the one sequential event with the greatest negative departure year.
+In every case we assessed (described below) we felt that the OUTBREAK-omitted defoliation events should have been maintained and recorded.
+Due in large part to reconstructions using OUTBREAK (see papers cited by Lynch 2012), we now know considerably more about forest defoliator outbreak regimes than we did in the 1980s when OUTBREAK was written.
+We think that two or more prolonged events separated by a single year should in some situations be considered a single event.
+This is particularly relevant to WSBW and spruce budworm (\emph{C.
+fumiferana}), for which regimes have been reconstructed for several geographic areas and for which researchers and forest health experts have gained considerable knowledge (Schmitt, Grimble, and Searcy 1984; Sanders et al.
+1985; Brookes et al.
+1982, and many later publications).
+In many cases, the greatest growth suppression often occurs late in the outbreak due to cumulative effects accrued on a tree's resources (Swetnam and Lynch 1989, 1993).
+We urge caution in using the bridging option, however, because it may not be appropriate for all studied insects, such as in situations where impacted stands barely recover from one outbreak before another begins, as with pine processionary caterpillars (\emph{Thaumetopoea pityocampa} (Lepidoptera: Thaumetopoeidae)) (Carus 2004, 2009).
+
+\hypertarget{inferring-outbreak-events}{%
+\subsection{Inferring Outbreak Events}\label{inferring-outbreak-events}}
+
+Defoliation of one or a few trees does not constitute an outbreak.
+To determine when defoliation becomes an \emph{outbreak event}, \texttt{dfoliatR} composites the individual tree defoliation series into a site-level chronology with the \texttt{outbreak()} function.
+Users have options to define the number and/or the proportion of trees required for an event to be considered an outbreak.
+Three parameters control whether a defoliation event constitutes an outbreak: the minimum number of trees available, the minimum number of trees recording defoliation, and the percent of trees recording defoliation.
+The first allows the researcher to make a judgement call as to the confidence ascribed to reduced sample depth toward the ends of their chronologies, thus compensating for the ``fading record problem'' (Swetnam, Allen, and Betancourt 1999).
+The second two parameters adjust the scale of defoliation considered to be an outbreak.
+Absolute numbers of trees and percentages can be applied separately or in conjunction, following filtering conventions in tree-ring fire history studies (Malevich, Guiterman, and Margolis 2018).
+We urge users to carefully consider the choice of absolute numbers in situations where the number of trees represented in the series varies with time, or the choice of percentages when sample size is small.
+
+\hypertarget{evaluation}{%
+\section{Evaluation}\label{evaluation}}
+
+\hypertarget{approach}{%
+\subsection{Approach}\label{approach}}
+
+We tested \texttt{dfoliatR} against OUTBREAK by comparing NGSI to OUTBREAK's normalized corrected indices for individual trees and years, defoliation status for individual trees and years, and percentage of trees recording outbreaks at the site level, using standardized ring-width data from eight sites in British Columbia, Colorado, and New Mexico, and author-provided non-host site chronologies.
+
+We detrended host data for both \texttt{dfoliatR} and OUTBREAK using ARSTAN (v6.1, Cook and Holmes 1996) with default double detrending (128 year wavelength and a 50\% smoothing spline).
+In both \texttt{dfoliatR} and OUTBREAK we used event thresholds of -1.28 normalized indices, 8 years duration, and allowed for events at the end of series in seven of eight sites.
+These seven sites were sampled during ongoing outbreak events (Swetnam and Lynch 1993).
+We found it necessary to be consistent in how we detrended and in what software (e.g., ARSTAN vs \texttt{dplR}) we employed because subtle differences in standardized ring-width indices generated between the programs transferred into differences between \texttt{dfoliatR} and OUTBREAK.
+In the end, we chose to only use the standardization output files from ARSTAN, which are easily read into \texttt{R} (and then \texttt{dfoliatR}) using the \texttt{dplR} package.
+
+The \texttt{R} code to replicate our comparisons is available from \url{https://github.com/chguiterman/dfoliatR_paper}.
+
+\hypertarget{findings}{%
+\subsection{Findings}\label{findings}}
+
+Across the 43,280 ring-width indices from 222 trees included in our evaluation, we found that \texttt{dfoliatR} and OUTBREAK compute identical growth suppression indices at 0.000 precision.
+We expected this outcome because both programs employ Equation 1.
+At the tree-level, the programs identified 11,530 years of defoliation.
+The programs agreed on 97.9\% of the years, leaving 927 ``difference'' years in which only one program identified defoliation on an individual tree.
+The differences included 102 events on 85 trees.
+We carefully inspected each of these events in the full context of each tree's ring-series, and categorized the differences as follows
+
+\begin{itemize}
+\item
+  \emph{Series-end events} (40\% of the total) in which OUTBREAK included ``truncated outbreaks'' (for seven sites) at the end of each series.
+In \texttt{dfoliatR}, this option is controlled by the ``series\_end\_events'' parameter to \texttt{defoliate\_trees()}.
+In OUTBREAK, the option appears while changing the duration parameter (option 3).
+When selected, OUTBREAK will include any sequences of negative indices at the either the beginning or the end of each tree series as a defoliation event, without consideration of either duration or magnitude thresholds.
+In \texttt{dfoliatR}, the duration threshold is omitted and the magnitude threshold is retained in series-end-events.
+Each of the 13 events included in these differences did not meet the ``max\_reduction'' parameter (-1.28 NGSI) in \texttt{dfoliatR} and were excluded.
+In two cases, OUTBREAK included events at the beginning of the series where \texttt{dfoliatR} does not allow truncated events.
+In four cases, OUTBREAK omitted only the last year of the series because the index was positive.
+In two cases, \texttt{dfoliatR} omitted possible events because it had already included a positive NGSI excursion after the ``max\_reduction'' year, and since it will only allow one excursion on either side of the max year, the events were omitted due to short duration.
+\item
+  \emph{Sequential events} (36\%) in which OUTBREAK omitted events that occur one year prior or one year following an identified event.
+When this occurs, OUTBREAK selects the event with the greatest negative index year.
+On two trees, OUTBREAK omitted two of three sequential events.
+While inspecting these differences, we added an option to \texttt{defoliate\_trees()} that would ``bridge'' between sequential events (that surpass magnitude and duration thresholds)into single, long events.
+We felt that this was ecologically justified, especially for studies of WSBW, because the insect will occasionally defoliate individual trees for decades and reconstructed chronologies show outbreaks as long as \textbf{xxx (ANN CHECK)}.
+\item
+  \emph{Undetermined differences} (22\%) occurred in cases where OUTBREAK omitted events without clear cause that \texttt{dfoliatR} correctly identified as defoliations.
+\item
+  \emph{Rounding differences} (2\%) in the indices either omitted or cut short events on two trees.
+In both cases the indices were very close to zero, and the difference was less than the precision of the raw data measurement.
+\end{itemize}
+
+\textbf{(Bridging paragraph?)}
+
+At the site level, OUTBREAK and \texttt{dfoliatR} produce similar time series of percent trees defoliated (Figure \ref{fig:df-obr-site}), which forms the basis for inferring outbreak occurrence, intensity, and duration.
+Nearly all of the differences between the programs include either more trees or a longer duration of inferred outbreak by \texttt{dfoliatR}.
+These arise from the inclusion of events omitted by OUTBREAK.
+These differences translate to greater numbers of trees in defoliation during certain events, or events recorded by single trees during periods of low sample depth.
+
+This evaluation revealed what we believe are short-comings in how OUTBREAK identifies defoliation events on individual trees.
+In every one of the 102 cases we inspected, we felt that \texttt{dfoliatR} provided a more biologically and statistically appropriate assessment of defoliation.
+Therefore, \texttt{dfoliatR} also did a better job inferring OUTBREAK events and associated statistics at the site level.
 
 \hypertarget{availability-and-installation}{%
-\section{Availability and
-installation}\label{availability-and-installation}}
+\section{Availability and installation}\label{availability-and-installation}}
 
-The \texttt{dfoliatR} is provided free and open source (Guiterman 2020).
-It is provided to \texttt{R} users via the Comprehensive R Archive
-Network (CRAN; \url{https://cran.r-project.org/}). To install
-\texttt{dfoliatR} from CRAN use
+The \texttt{dfoliatR} library (Guiterman, Lynch, and Axelson 2020) is provided free and open source from the Comprehensive R Archive Network (CRAN; \url{https://cran.r-project.org/}).
+To install \texttt{dfoliatR} from CRAN use
 
@@
 
-Development versions of \texttt{dfoliatR} are available on Github and
-installed using the \texttt{devtools} library,
+Development versions of \texttt{dfoliatR} are available on GitHub and installed using the \texttt{devtools} library,
 
@@
 
-Issues, bug reports, and ideas for improving \texttt{dfoliatR} can be
-posted to \url{https://github.com/chguiterman/dfoliatR/issues}. As an
-Open Source library, we welcome and encourage community involvement in
-future development. The best ways to contribute to \texttt{dfoliatR} are
-through standard Github procedures or by contacting the first author.
+Issues, bug reports, and ideas for improving \texttt{dfoliatR} can be posted to \url{https://github.com/chguiterman/dfoliatR/issues}.
+As an Open Source library, we welcome and encourage community involvement in future development.
+The best ways to contribute to \texttt{dfoliatR} are through standard GitHub procedures or by contacting the first author.
 
@@
 
+Once \texttt{dfoliatR} is loaded into an \texttt{R} session (via \texttt{library(dfoliatR)}) users can access two sets of tree-ring data to aid in exploring the functions, graphics, and outputs.
+Each data set consists of individual host-tree series and a local non-host chronology.
+The host-tree series were standardized using 50-year splines with a 50\% frequency response, while the non-host ring-width data were standardized using 150-year splines with a 50\% frequency response and then averaged via Tukey's biweight robust mean procedure.
+Host trees from Demijohn Peak (DMJ; 2902 m asl) in the San Juan Mountains of southern Colorado include Douglas-fir (\emph{Pseudotsuga menziesii}) compared against a local non-host ponderosa pine (\emph{Pinus ponderosa}) chronology (Ryerson, Swetnam, and Lynch 2003).
+The East Fork site (EFK; 2580 m asl) in the Jemez Mountains of northcentral New Mexico includes Douglas-fir and white fir (\emph{Abies concolor}) host trees and a ponderosa pine non-host chronology (Swetnam and Lynch 1993).
+
+With \texttt{dfoliatR} loaded, the datasets are accessible using the \texttt{data()} function.
+The data object names are prefixed by their site codes.
+For instance, the \texttt{dmj\_*} objects come from the DMJ site and include the host-tree series (\texttt{dmj\_h}), the non-host chronology (\texttt{dmj\_nh}), the defoliation series (\texttt{dmj\_defol}), and the outbreak series (\texttt{dmj\_obr}).
+The same suite of data are available for EFK using the \texttt{efk\_*} prefix.
+
+In our example scripts below, \texttt{\#\#} and \texttt{\#} denote user comments, per standard \texttt{R} coding, which are colored in brown.
+Text in blue denote functions; black are loaded objects, and green are quoted variables and links.
+
 \hypertarget{tree-level-defoliation-events}{%
-\subsection{Tree-level defoliation
-events}\label{tree-level-defoliation-events}}
+\subsection{Tree-Level Defoliation Events}\label{tree-level-defoliation-events}}
+
+The function \texttt{defoliate\_trees()} performs the GSI indexing procedure on each host-tree series and then identifies defoliation events.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{## Load the package}
+\KeywordTok{library}\NormalTok{(dfoliatR)}
+
+\CommentTok{## Load data}
+\KeywordTok{data}\NormalTok{(}\StringTok{"dmj_h"}\NormalTok{) }\CommentTok{# host tree series}
+\KeywordTok{data}\NormalTok{(}\StringTok{"dmj_nh"}\NormalTok{) }\CommentTok{# non-host chronology}
+\CommentTok{# View data structure for any data object via }
+\CommentTok{# "View(dmj_h)" or "head(dmj_h)"}
+
+\CommentTok{## Calculate indices, identify defoliation events}
+\NormalTok{dmj_defol <-}\StringTok{ }\KeywordTok{defoliate_trees}\NormalTok{(}\DataTypeTok{host_tree =}\NormalTok{ dmj_h,}
+                             \DataTypeTok{nonhost_chron =}\NormalTok{ dmj_nh,}
+                             \DataTypeTok{duration_years =} \DecValTok{8}\NormalTok{,  }
+                             \DataTypeTok{max_reduction =} \FloatTok{-1.28}\NormalTok{, }
+                             \DataTypeTok{bridge_events =} \OtherTok{TRUE}\NormalTok{, }
+                             \DataTypeTok{series_end_event =} \OtherTok{TRUE}\NormalTok{, }
+                             \DataTypeTok{list_output =} \OtherTok{FALSE}\NormalTok{) }
+\end{Highlighting}
+\end{Shaded}
+
+The result is long-format (stacked) data frame with five variables: ``year'', ``series'', ``gsi'', ``ngsi'', and ``defol\_status.'' The ``defol\_status'' column indicates whether that year has defoliation or not, with a set of factors that include ``nd'' for non-defoliation year, ``defol'' for a defoliation year, ``max\_defol'' for the year of maximum suppression (that acts as the basis for individual events), ``bridge\_defol'' to identify years that link subsequent events (only one is present at DMJ), and ``series\_end\_defol'' to identify defoliation at the present-end of the series.
+
+Selecting \texttt{list\_output\ =\ TRUE} in \texttt{defoliate\_trees()} provides a list-object of data frames, each with an \texttt{rwl} object that combines the host tree and non-host series and the other columns created by \texttt{defoliate\_trees()}.
+This option is not used by subsequent functions in \texttt{dfoliatR}, but researchers can examine it to check the results of the GSI calculation (Equation 1), such as the non-host series after scaling to a common variance with a particular host-tree series.
+
+The results of running \texttt{defoliate\_trees()} can be assessed through graphical and table outputs.
+The function \texttt{get\_defol\_events()} will provide a list of every defoliation event for every tree, with the corresponding mean ``ngsi'' value.
+A summary table of the results for each tree is produced by \texttt{defol\_stats()} (Table \ref{tab:tbl-defol}).
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{defol_stats}\NormalTok{(dmj_defol)}
+\KeywordTok{View}\NormalTok{(dmj_defol)}
+\end{Highlighting}
+\end{Shaded}
+
+The \texttt{plot\_defol()} function produces a ``ggplot'' graphics object with line segments showing the measured sequence of each series and a filled segment for each identified defoliation event (Figure \ref{fig:fig-defol}).
+The defoliation segments are colored by their relative severity based on their average NGSI value.
+The default cut-off values between ``Severe'' and ``Moderate'' is the overall mean across all events.
+The cut-off between ``Moderate'' and ``Minor'' is the first quartile by default.
+Users can re-define the breaks to suit their needs via the ``breaks'' parameter in \texttt{plot\_defol()}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{plot_defol}\NormalTok{(dmj_defol)}
+\CommentTok{## This creates a ggplot object. Additions can }
+\CommentTok{## be made to adjust plotting themes and}
+\CommentTok{## aesthetics, like color.   }
+\CommentTok{## For more on the features of ggplot see}
+\CommentTok{## https://ggplot2.tidyverse.org/}
+\CommentTok{## Adding 'scale_color_manual()' will}
+\CommentTok{## change the colors that denote }
+\CommentTok{## the average NGSI value for each defoliation}
+\CommentTok{## event.}
+\KeywordTok{library}\NormalTok{(ggplot2)}
+\KeywordTok{plot_defol}\NormalTok{(dmj_defol) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{scale_color_manual}\NormalTok{(}\DataTypeTok{values =} \KeywordTok{c}\NormalTok{(}\StringTok{"red"}\NormalTok{, }\StringTok{"orange"}\NormalTok{, }\StringTok{"purple"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+These output functions aid in assessing the sensitivity of input parameters to \texttt{defoliate\_trees()}, including the duration and magnitude thresholds for identifying defoliation events.
+Using \texttt{plot\_defol()} also provides a direct assessment of the between-tree variability in defoliation.
 
 \hypertarget{site-level-events}{%
-\subsection{Site-level events}\label{site-level-events}}
+\subsection{Site-Level Events}\label{site-level-events}}
 
-\hypertarget{evaluation}{%
-\section{Evaluation}\label{evaluation}}
+To infer outbreak events at the site level, the function \texttt{outbreak()} composites tree-level defoliation series into a single chronology, with input parameters that control thresholds in the number and proportions of trees recording a defoliation event.
 
-\hypertarget{extensions}{%
-\section{Extensions}\label{extensions}}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{## Use the defol object "dmj_defol" created above}
+\NormalTok{dmj_obr <-}\StringTok{ }\KeywordTok{outbreak}\NormalTok{(dmj_defol,}
+                    \DataTypeTok{filter_min_series =} \DecValTok{3}\NormalTok{,}
+                    \DataTypeTok{filter_min_defol =} \DecValTok{1}\NormalTok{,}
+                    \DataTypeTok{filter_perc =} \DecValTok{25}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+Input parameters to \texttt{outbreak()} include ``filter\_min\_series'' to control the chronology cut-off points with regard to sample depth, ``filter\_min\_defol'' and ``filter\_perc'' to control the minimum number and percent of trees recording a defoliation event in a given year.
+\texttt{Outbreak()} produces a new data frame with eight variables: ``year'', ``num\_defol'', ``percent\_defol'', ``num\_max\_defol'', ``mean\_gsi'', ``mean\_ngsi'', and ``outbreak\_status.'' All of these variables are populated regardless of an inferred outbreak event, providing a continuous disturbance chronology.
+The ``num\_max\_defol'' variable counts the number of trees recording their maximum defoliation in a given year.
+The ``mean\_gsi'' and ``mean\_ngsi'' variables provide averages of these indices across all available trees.
+Finally, the ``outbreak\_status'' column shows whether an outbreak event is inferred (``outbreak'') or not (``not\_obr'').
+
+The default plotting function to visualize results from \texttt{outbreak()} creates a three-panel graph showing the mean site-level chronology, the percent of trees recording a defoliation, and the sample depth over time (Figure \ref{fig:fig-obr}).
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{plot_outbreak}\NormalTok{(dmj_obr, }\DataTypeTok{disp_index =} \StringTok{"mean_gsi"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+Inferred outbreak events are shown in the top panel of Figure \ref{fig:fig-obr} as the filled-in spaces.
+Users can change the time series in this panel with the ``disp\_index'' parameter, choosing between the mean NGSI (the default) or GSI.
+
+A summary table of the inferred outbreak events is generated by the \texttt{outbreak\_stats()} function (Table \ref{tab:tbl-obr}).
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{dmj_obr_stats <-}\StringTok{ }\KeywordTok{outbreak_stats}\NormalTok{(dmj_obr)}
+\KeywordTok{View}\NormalTok{(dmj_obr_stats)}
+\end{Highlighting}
+\end{Shaded}
+
+The output table from \texttt{outbreak\_stats()} shows the start and end years of each event, along with their corresponding duration, the number and percent of trees in defoliation at the start of the event, the the number of trees recording the outbreak event (in part or full), the minimums of the mean GSI and NGSI indices, and two points of ``peak defoliation.'' The first, ``peak\_outbreak\_year'' is the year in which the greatest number of trees are recording the outbreak, and the second, ``peak\_defol\_year'' is the year with the lowest average NGSI.
+These output variables provide a array of options for assessing metrics of the insect defoliation regime, including interval analyses:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{## calculate mean return interval for outbreak start years}
+\NormalTok{intervals <-}\StringTok{ }\KeywordTok{diff}\NormalTok{(dmj_obr_stats}\OperatorTok{$}\NormalTok{start)}
+\KeywordTok{mean}\NormalTok{(intervals) }\CommentTok{## 51.2 year mean return interval}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{conclusions}{%
+\section{Conclusions}\label{conclusions}}
+
+The \texttt{dfoliatR} package provides dendroecologists with tools to infer, quantify, analyze, and visualize tree-ring growth suppression events and to reconstruct forest insect defoliator outbreak regimes.
+It is built on the long-accepted host to nonhost comparison methodology used in the 1980s FORTRAN program OUTBREAK (Swetnam, Thompson, and Sutherland 1985; Swetnam and Lynch 1989, 1993).
+Key benefits are that \texttt{dfoliatR} provides easier control of suppression thresholds, provides additional output tables and high-quality graphics, defines defoliation events more consistently, and operates in the open-source \texttt{R} environment.
+
+Using \texttt{dfoliatR} requires standardized ring-width measurements from insect host trees and either an indexed tree-ring chronology from local non-host trees or suitable climate chronology.
+It performs an indexing procedure to remove the climatic signal represented in the non-host chronology from the host-tree series.
+It then infers defoliation events in individual trees.
+Site-level analyses identify outbreak events that synchronously affect a user-defined number or proportion of the host trees.
+Functions are provided for summary statistics and graphics of tree- and site-level series.
+The package produces publication-quality plots, and tabulates growth suppression indices and tree- and site-level outbreak event statistics for user-defined post-processing needs, including those suitable for charting and tabulating landscape- and regional-level results.
+
+\texttt{dfoliatR} adds a new option for dendroentimology to combine, or ``bridge,'' sequential tree-level defoliation events into single events.
+In practice, we suggest that researchers carefully evaluate if bridging is ecologically applicable to study situations and insect ecology, and to carefully explore the data before deciding whether or not to use this option.
+It is probably not appropriate for insects with high-frequency, high--severity outbreaks, such as processionary caterpillars.
+Outbreak reconstructions of insects for which the interval is notably longer than typical outbreak duration, such as Douglas-fir tussock moth (\emph{Orgyia pseudotsugata} (Lepidoptera: Erebidae)), are unlikely to be significantly affected.
+Species for which outbreak duration varies considerably, for which individual tree resilience to defoliation varies considerably, or which may chronically infest trees or sites, or alternate between chronic, outbreak, and minimal activity states, such as conifer-feeding \emph{Choristoneura}, present more complicated challenges, and the researcher should use discretion with the bridge option.
+
+\texttt{dfoliatR} adds to the on-going open-source software development for dendrochronological methods (e.g., Bunn 2008; Brewer 2014; Brewer and Guiterman 2016).
+The \texttt{R} environment enables automation of analyses, allowing input/output processes to become routine, and empowering batch processing of large multi-site projects and sensitivity analyses.
+It also facilitates additional statistical analyses, such as spectral analyses and superposed epoch analyses (e.g., Malevich, Guiterman, and Margolis 2018), with easy transfer from \texttt{dfoliatR} and \texttt{dplR} to other libraries in \texttt{R}.
+Source code for \texttt{dfoliatR} is available in the Comprehensive R Archive Network (CRAN) and GitHub \url{https://github.com/chguiterman/dfoliatR} with updated descriptions and helpful vignettes on the package website \url{https://chguiterman.github.io/dfoliatR/}.
+Researchers wishing to contribute to further development of \texttt{dfoliatR} are encouraged to do so via the GitHub repository.
+
+\hypertarget{acknowledgments}{%
+\section*{Acknowledgments}\label{acknowledgments}}
+\addcontentsline{toc}{section}{Acknowledgments}
+
+We are grateful to late Richard Holmes of the Laboratory of Tree-Ring Research (LTRR) for his achievements in developing OUTBREAK and many other tree-ring software programs, and we thank Thomas Swetnam for his insights and encouragement.
+The \texttt{dfoliatR} project was supported by the Agnes Haury Visiting Scholars Fellowship at the LTRR, the University of California Berkeley, and the Western Wildlands Environmental Threat Center (WWETAC).
+Additional support was provided by the LTRR and the United States Forest Service Rocky Mountain Research Station.
+Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S.
+Government.
 
@@
 \leavevmode\hypertarget{ref-Altman2014}{}%
-Altman, J., P. Fibich, J. Dolezal, and T. Aakala. 2014. TRADER: A
-package for Tree Ring Analysis of Disturbance Events in R.
-Dendrochronologia 32:107--112.
+Altman, J., P.
+Fibich, J.
+Dolezal, and T.
+Aakala.
+2014.
+``TRADER: A package for Tree Ring Analysis of Disturbance Events in R.'' \emph{Dendrochronologia} 32: 107--12.
+\url{https://doi.org/10.1016/j.dendro.2014.01.004}.
+
+\leavevmode\hypertarget{ref-Blais1954}{}%
+Blais, JR. 1954. ``The Recurrence of Spruce Budworm Infestations in the Past Century in the Lac Seul Area of Northwestern Ontario.'' \emph{Ecology} 35 (1): 62--71.
+
+\leavevmode\hypertarget{ref-Brewer2014}{}%
+Brewer, Peter W. 2014. ``Data Management in Dendroarchaeology Using Tellervo.'' \emph{Radiocarbon} 56 (4): S79--S83. \url{https://doi.org/10.2458/azu_rc.56.18320}.
+
+\leavevmode\hypertarget{ref-Brewer2016}{}%
+Brewer, Peter W., and Christopher H.
+Guiterman.
+2016.
+``A New Digital Field Data Collection System for Dendrochronology.'' \emph{Dendrochronologia} 38: 131--35.
+\url{https://doi.org/10.1016/j.dendro.2016.04.005}.
+
+\leavevmode\hypertarget{ref-Brookes1987}{}%
+Brookes, M.
+H., Robert W.
+Campbell, J.
+J.
+Colbert, R.
+G.
+Mitchell, and R.
+W.
+Stark (Technical Coordinators).
+1982.
+``Western Spruce Budworm.'' USDA Forest Service, Cooperative State Research Service, Technical Bulletin No.
+1694.
+198 p., Washington D.C.
 
 \leavevmode\hypertarget{ref-Bunn2008}{}%
-Bunn, A. G. 2008. A dendrochronology program library in R (dplR).
-Dendrochronologia 26:115--124.
+Bunn, Andrew G. 2008. ``A dendrochronology program library in R (dplR).'' \emph{Dendrochronologia} 26: 115--24. \url{https://doi.org/10.1016/j.dendro.2008.01.002}.
 
 \leavevmode\hypertarget{ref-Bunn2010}{}%
-Bunn, A. G. 2010. Statistical and visual crossdating in R using the dplR
-library. Dendrochronologia 28:251--258.
+---------. 2010. ``Statistical and visual crossdating in R using the dplR library.'' \emph{Dendrochronologia} 28: 251--58. \url{https://doi.org/10.1016/j.dendro.2009.12.001}.
 
-\leavevmode\hypertarget{ref-chris_guiterman_2020_3626136}{}%
-Guiterman, C. 2020. Chguiterman/dfoliatR: First release. Zenodo.
+\leavevmode\hypertarget{ref-Campelo2016}{}%
+Campelo, Filipe, Cristina Nabais, Ana Carvalho, and Joana Vieira.
+2016.
+``tracheideR--An R package to standardize tracheidograms.'' \emph{Dendrochronologia} 37: 64--68.
+\url{https://doi.org/10.1016/j.dendro.2015.12.006}.
+
+\leavevmode\hypertarget{ref-Carus2004}{}%
+Carus, Serdar.
+2004.
+``Impact of defoliation by the pine processionary moth (\emph{Thaumetopoea pityocampa}) on radial, height and volume growth of Calabrian pine (\emph{Pinus brutia}) trees in Turkey.'' \emph{Phytoparasitica} 32 (5): 459--69.
+
+\leavevmode\hypertarget{ref-Carus2009}{}%
+---------. 2009. ``Effects of Defoliation Caused by the Processionary Moth on Growth of Crimean Pines in Western Turkey.'' \emph{Phytoparasitica} 37 (2): 105--14.
+
+\leavevmode\hypertarget{ref-arstan}{}%
+Cook, Edward R, and Richard L Holmes. 1996. ``Guide for computer program ARSTAN.'' \emph{The International Tree-Ring Data Bank Program Library Version 2}, 75--87.
+
+\leavevmode\hypertarget{ref-Cook2004}{}%
+Cook, Edward R., and Paul J.
+Krusic.
+2004.
+\emph{The North American Drought Atlas}.
+(Lamont- Doherty Earth Observatory; the National Science Foundation: New York) Available at https://www.ncdc.noaa.gov/data-access/paleoclimatology-data/datasets/tree-ring/drought-variability {[}Last accessed May 2020{]}.
+
+\leavevmode\hypertarget{ref-dfoliatR}{}%
+Guiterman, Christopher, Ann Lynch, and Jodi Axelson.
+2020.
+\emph{DfoliatR: Detection and Analysis of Insect Defoliation Signals in Tree Rings}.
+R package version 0.1.0.
+\url{https://doi.org/10.5281/zenodo.3698163}.
+
+\leavevmode\hypertarget{ref-outbreak}{}%
+Holmes, Richard L, and Thomas W Swetnam. 1986. ``Dendroecology program library: program OUTBREAK user's manual.'' \emph{Laboratory of Tree-Ring Research, University of Arizona, Tucson}.
 
 \leavevmode\hypertarget{ref-Jevsenak2018}{}%
-Jevšenak, J., and T. Levanič. 2018. dendroTools: R package for studying
-linear and nonlinear responses between tree-rings and daily
-environmental data. Dendrochronologia 48:32--39.
+Jevšenak, Jernej, and Tom Levanič.
+2018.
+``dendroTools: R package for studying linear and nonlinear responses between tree-rings and daily environmental data.'' \emph{Dendrochronologia} 48: 32--39.
+\url{https://doi.org/10.1016/j.dendro.2018.01.005}.
 
 \leavevmode\hypertarget{ref-Lara2015}{}%
-Lara, W., F. Bravo, and C. A. Sierra. 2015. MeasuRing: An R package to
-measure tree-ring widths from scanned images. Dendrochronologia
-34:43--50.
+Lara, W., F.
+Bravo, and C.
+A.
+Sierra.
+2015.
+``MeasuRing: An R package to measure tree-ring widths from scanned images.'' \emph{Dendrochronologia} 34: 43--50.
+\url{https://doi.org/10.1016/j.dendro.2015.04.002}.
+
+\leavevmode\hypertarget{ref-Lynch2012}{}%
+Lynch, Ann M.
+2012.
+``What Tree-Ring Reconstruction Tells Us about Conifer Defoliator Outbreaks.'' In \emph{Insect Outbreaks Revisited}, edited by Pedro Barbosa, Deborah K.
+Letourneau, and Anurag A.
+Agrawal, 126--54.
+Blackwell Publishing Ltd.
+\url{https://doi.org/10.1002/9781118295205.ch7}.
+
+\leavevmode\hypertarget{ref-Malevich2018}{}%
+Malevich, Steven B., Christopher H.
+Guiterman, and Ellis Q.
+Margolis.
+2018.
+``burnr: Fire history analysis and graphics in R.'' \emph{Dendrochronologia} 49: 9--15.
+\url{https://doi.org/10.1016/j.dendro.2018.02.005}.
+
+\leavevmode\hypertarget{ref-Rathgeber2011}{}%
+Rathgeber, Cyrille B.
+K., Fleur Longuetaud, Frédéric Mothe, Henri Cuny, and Gilles Le Moguédec.
+2011.
+``Phenology of wood formation: Data processing, analysis and visualisation using R (package CAVIAR).'' \emph{Dendrochronologia} 29 (3): 139--49.
+\url{https://doi.org/10.1016/j.dendro.2011.01.004}.
+
+\leavevmode\hypertarget{ref-RCore}{}%
+R Core Team. 2019. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.
+
+\leavevmode\hypertarget{ref-Ryerson2003}{}%
+Ryerson, Daniel E, Thomas W Swetnam, and Ann M Lynch.
+2003.
+``A tree-ring reconstruction of western spruce budworm outbreaks in the San Juan Mountains, Colorado, U.S.A.'' \emph{Canadian Journal of Forest Research} 33: 1010--28.
+\url{https://doi.org/10.1139/X03-026}.
+
+\leavevmode\hypertarget{ref-Sanders1985}{}%
+Sanders, C.
+J., R.
+W.
+Stark, E.
+J.
+Mullins, and J.
+Murphy (eds).
+1985.
+``Recent Advances in Spruce Budworms Research.'' In \emph{Proceedings of the CANUSA Spruce Budworms Research Symposium, 16-20 September 1984, Bangor, Maine.
+Canadian Forestry Service, Ottawa, Ontario.
+527p.}
+
+\leavevmode\hypertarget{ref-Schmitt1984}{}%
+Schmitt, D.
+M., D.
+G.
+Grimble, and JL Searcy.
+1984.
+\emph{Managing the spruce budworm in eastern North America}.
+USDA Forest Service, Cooperative State Research Service, Agriculture Handbook No.
+620.
+192p., Washington D.C.
+
+\leavevmode\hypertarget{ref-Shi2019}{}%
+Shi, Jingning, Wei Xiang, Qijing Liu, and Sher Shah.
+2019.
+``MtreeRing: An R Package with Graphical User Interface for Automatic Measurement of Tree Ring Widths Using Image Processing Techniques.'' \emph{Dendrochronologia} 58: 125644.
+\url{https://doi.org/10.1016/j.dendro.2019.125644}.
+
+\leavevmode\hypertarget{ref-Speer2010}{}%
+Speer, James H. 2010. \emph{Fundamentals of Tree-Ring Research}. The University of Arizona Press.
+
+\leavevmode\hypertarget{ref-Swetnam1989}{}%
+Swetnam, Thomas W., and Ann M.
+Lynch.
+1989.
+``A tree-ring reconstruction of western spruce budworm history in the southern Rocky Mountains.'' \emph{Forest Science} 35 (4): 962--86.
+\url{https://doi.org/10.1093/forestscience/35.4.962}.
+
+\leavevmode\hypertarget{ref-Swetnam1993}{}%
+---------. 1993. ``Multicentury, Regional-Scale Patterns of Western Spruce Budworm Outbreaks.'' \emph{Ecological Monographs} 63 (4): 399--424. \url{https://doi.org/10.2307/2937153}.
+
+\leavevmode\hypertarget{ref-Swetnam1985}{}%
+Swetnam, Thomas W., Marna A.
+Thompson, and Elaine K.
+Sutherland.
+1985.
+\emph{Using dendrochronology to measure radial growth of defoliated trees}.
+Vol.
+639.
+United States Department of Agriculture handbook No.
+639.
+38 p., Washington D.C.
+\url{www.fs.usda.gov/treesearch/pubs/40898}.
+
+\leavevmode\hypertarget{ref-Swetnam1999}{}%
+Swetnam, TW, CD Allen, and JL Betancourt.
+1999.
+``Applied historical ecology: using the past to manage for the future.'' \emph{Ecological Applications} 9 (4): 1189--1206.
+\url{https://doi.org/10.1890/1051-0761(1999)009\%5B1189:AHEUTP\%5D2.0.CO;2}.
 
 \leavevmode\hypertarget{ref-VanderMaaten-Theunissen2015}{}%
-Maaten-Theunissen, M. van der, E. van der Maaten, and O. Bouriaud. 2015.
-PointRes: An R package to analyze pointer years and components of
-resilience. Dendrochronologia 35:34--38.
-
-\leavevmode\hypertarget{ref-Malevich2018}{}%
-Malevich, S. B., C. H. Guiterman, and E. Q. Margolis. 2018. burnr: Fire
-history analysis and graphics in R. Dendrochronologia 49:9--15.
-
-\leavevmode\hypertarget{ref-Shi2019}{}%
-Shi, J., and W. Xiang. 2019. MtreeRing: A shiny application for
-automatic measurements of tree-ring widths on digital images.
-
-\leavevmode\hypertarget{ref-Swetnam1989}{}%
-Swetnam, T. W., and A. M. Lynch. 1989. A tree-ring reconstruction of
-western spruce budworm history in the southern Rocky Mountains. Forest
-Science 35:962--986.
-
-\leavevmode\hypertarget{ref-Swetnam1985}{}%
-Swetnam, T. W., M. A. Thompson, and E. K. Sutherland. 1985. Using
-dendrochronology to measure radial growth of defoliated trees. Page 39p.
+van der Maaten-Theunissen, Marieke, Ernst van der Maaten, and Olivier Bouriaud.
+2015.
+``PointRes: An R package to analyze pointer years and components of resilience.'' \emph{Dendrochronologia} 35: 34--38.
+\url{https://doi.org/10.1016/j.dendro.2015.05.006}.
 
 \leavevmode\hypertarget{ref-wickham2016ggplot2}{}%
-Wickham, H. 2016. Ggplot2: Elegant graphics for data analysis. Springer.
+Wickham, Hadley. 2016. \emph{ggplot2: Elegant Graphics for Data Analysis}. Second edition. Springer-Verlag, New York. \url{https://ggplot2-book.org/}.
 
 \leavevmode\hypertarget{ref-Wickham2019}{}%
-Wickham, H., M. Averick, J. Bryan, W. Chang, L. McGowan, R. François, G.
-Grolemund, A. Hayes, L. Henry, J. Hester, M. Kuhn, T. Pedersen, E.
-Miller, S. Bache, K. Müller, J. Ooms, D. Robinson, D. Seidel, V. Spinu,
-K. Takahashi, D. Vaughan, C. Wilke, K. Woo, and H. Yutani. 2019. Welcome
-to the tidyverse. Journal of Open Source Software 4:1686.
+Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy McGowan, Romain François, Garrett Grolemund, et al.
+2019.
+``Welcome to the Tidyverse.'' \emph{Journal of Open Source Software} 4 (43): 1686.
+\url{https://doi.org/10.21105/joss.01686}.
 
 \leavevmode\hypertarget{ref-Wickham2020dplyr}{}%
-Wickham, H., R. François, L. Henry, and K. Müller. 2020. Dplyr: A
-grammar of data manipulation.
+Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2020. \emph{Dplyr: A Grammar of Data Manipulation}. R package version 0.8.4. \url{https://CRAN.R-project.org/package=dplyr}.
 
 \leavevmode\hypertarget{ref-Zang2015}{}%
-Zang, C., and F. Biondi. 2015. Treeclim: An R package for the numerical
-calibration of proxy-climate relationships. Ecography 38:431--436.
+Zang, Christian, and Franco Biondi.
+2015.
+``Treeclim: An R package for the numerical calibration of proxy-climate relationships.'' \emph{Ecography} 38 (4): 431--36.
+\url{https://doi.org/10.1111/ecog.01335}.
+
+\newpage
+
+\hypertarget{tables-figures}{%
+\section*{Tables \& Figures}\label{tables-figures}}
+\addcontentsline{toc}{section}{Tables \& Figures}
+
+\begin{table}[!h]
+
+\caption{\label{tab:tbl-defol}Summary statistics for individual tree defoliation events.}
+\centering
+\begin{tabular}[t]{lrrrrrr}
+\toprule
+series & first & last & years & num\_events & tot\_years & mean\_duration\\
+\midrule
+DMJ01 & 1810 & 1996 & 187 & 4 & 50 & 12\\
+DMJ02 & 1750 & 1996 & 247 & 6 & 75 & 12\\
+DMJ03 & 1830 & 1996 & 167 & 4 & 42 & 10\\
+DMJ04 & 1720 & 1996 & 277 & 8 & 90 & 11\\
+DMJ06 & 1700 & 1996 & 297 & 6 & 81 & 14\\
+\addlinespace
+DMJ07 & 1710 & 1996 & 287 & 7 & 96 & 14\\
+DMJ11 & 1900 & 1997 & 98 & 2 & 19 & 10\\
+DMJ14 & 1675 & 1996 & 322 & 9 & 112 & 12\\
+DMJ15 & 1730 & 1996 & 267 & 4 & 64 & 16\\
+DMJ16 & 1746 & 1996 & 251 & 7 & 86 & 12\\
+\addlinespace
+DMJ17 & 1733 & 1996 & 264 & 6 & 85 & 14\\
+DMJ22 & 1720 & 1996 & 277 & 5 & 56 & 11\\
+DMJ23 & 1675 & 1997 & 323 & 8 & 96 & 12\\
+DMJ24 & 1895 & 1996 & 102 & 3 & 28 & 9\\
+DMJ25 & 1680 & 1996 & 317 & 5 & 70 & 14\\
+\addlinespace
+DMJ26 & 1700 & 1996 & 297 & 6 & 79 & 13\\
+DMJ27 & 1710 & 1996 & 287 & 3 & 34 & 11\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{landscape}\begin{table}
+
+\caption{\label{tab:tbl-obr}Summary statistics for inferred outbreak events.}
+\centering
+\fontsize{7}{9}\selectfont
+\begin{tabular}[t]{rrrrrrrrrr}
+\toprule
+start & end & duration & num\_trees\_start & perc\_trees\_start & peak\_outbreak\_year & num\_trees\_outbreak & peak\_defol\_year & min\_gsi & mean\_ngsi\\
+\midrule
+1680 & 1699 & 20 & 1 & 33.3 & 1690 & 3 & 1692 & 0.121 & -2.407\\
+1753 & 1769 & 17 & 6 & 46.2 & 1754 & 7 & 1755 & 0.343 & -1.611\\
+1825 & 1840 & 16 & 11 & 78.6 & 1831 & 12 & 1826 & 0.500 & -1.304\\
+1849 & 1865 & 17 & 7 & 46.7 & 1852 & 13 & 1853 & 0.252 & -1.994\\
+1881 & 1895 & 15 & 8 & 53.3 & 1886 & 14 & 1885 & 0.262 & -1.945\\
+\addlinespace
+1959 & 1970 & 12 & 7 & 41.2 & 1960 & 15 & 1965 & 0.328 & -1.830\\
+1987 & 1996 & 10 & 9 & 52.9 & 1989 & 15 & 1995 & 0.378 & -1.640\\
+\bottomrule
+\end{tabular}
+\end{table}
+\end{landscape}
+
+\begin{figure}
+\centering
+\includegraphics{../paper_elsevier/Output/tree-plot-default.pdf}
+\caption{\label{fig:fig-defol}Default dfoliatR graphics for individual trees, produced by the `plot\_defol()' function.
+Dotted lines represent the series length for each tree and colored segments show periods of defoliation.
+Users can define the Cut-off values to determine severe-moderate-minor defoliation intensities, and use standard ggplot2 graphics parameters to adjust styles and themes.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{../paper_elsevier/Output/site-plot.pdf}
+\caption{\label{fig:fig-obr}Default graphic to show outbreak events at the DMJ site.
+Top panel shows the mean GSI through time, with inferred outbreak events filled.
+The middle panel shows the percent of trees defoliated, and the bottom panel shows the number of trees through time.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{../paper_elsevier/Output/plot-trees-df-obr.pdf}
+\caption{\label{fig:df-obr-1-1}Comparison of disturbance indices between dfoliatR and OUTBREAK. Diagonal lines show equal values (1:1).}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{../paper_elsevier/Output/plot-sites-df-obr.pdf}
+\caption{\label{fig:df-obr-site}Comparison of reconstructed western spruce budworm outbreaks computed by dfoliatR and OUTBREAK. Input parameters were identical between programs.}
+\end{figure}
 

### 2020-05-06 paper_elsevier/paper_elsevier.Rmd
more edits

@@
 abstract: |
-  We present a new R package to provide dendroecologists with tools to infer, quantify, analyze, and visualize growth suppression events in tree rings caused by insect defoliation, and to reconstruct defoliator outbreak chronologies.
-‘dfoliatR’ is based on the OUTBREAK program and builds on existing resources in the R and dplR computing environments.
-It performs an indexing procedure to remove climatic signals, represented in non-host chronologies or other reconstructed climate series, from the host-tree series.
+ We present a new R package to provide dendroecologists with tools to infer, quantify, analyze, and visualize growth suppression events in tree rings.
+`dfoliatR` is based on the OUTBREAK program and builds on existing resources in the `R` and `dplR` computing environments.
+It is designed wo aid research in the ecology of insect defoliation events and to reconstruct defoliator outbreak chronologies, but can be applied to other studies where host--non-host comparisons are useful.
+`It`dfoliatR` performs an indexing procedure to remove climatic signals in the host-tree series that are represented in the non-host chronologies, or other annually-resolved climate series.
 It then infers defoliation events in individual trees based on user-specified thresholds.
@@
 Functions are provided for summary statistics and graphics of tree- and site-level series.
-‘dfoliatR’ expands on OUTBREAK to provide greater control of suppression thresholds, additional output tables, and high-quality graphics.
+`dfoliatR` expands on OUTBREAK to provide greater control of suppression thresholds, additional output tables, and high-quality graphics.
 We provide two example data sets and script to enable users to gain familiarity with the package and its capabilities.
 The source code is available in the Comprehensive R Archive Network (CRAN) and on GitHub.
-
  \hfill\break
 keywords: |
-  Dendroecology, spruce budworm, Choristoneura, pandora moth, Coloradia pandora Blake, larch-bud-moth
+  Dendroecology, spruce budworm, *Choristoneura*, pandora moth, *Coloradia pandora Blake*, larch-bud-moth
   \newpage
@@
 ```
-
 Introduction
@@
 It was not until the 1980s, however, that precise dendrochronological techniques were applied for inferring defoliation events and reconstructing defoliator outbreak regimes [@Swetnam1985; @Speer2010; @Lynch2012].
-The first studies [@Swetnam1985; @Swetnam1989; @Swetnam1993] focused on developing historical reconstructions of western spruce budworm (WSBW; *Choristoneura freemani* (Lepidoptera: Tortricidae); previously known as *C.
-occidentalis*) outbreak regimes.
+The first studies [@Swetnam1985; @Swetnam1989; @Swetnam1993] focused on developing historical outbreak reconstructions of western spruce budworm (WSBW; *Choristoneura freemani* (Lepidoptera: Tortricidae); previously known as *C.
+occidentalis*).
 The methodology has since been successfully applied to a wide range of defoliator species, most of which are conifer herbivores, and has evolved in sophistication for a wide range of ecosystem situations [@Lynch2012].
@@
 It was written in FORTRAN V with inherently severe restrictions, as RAM and disk space were limited at that time (256 kb and 10 MB, respectively) and FORTRAN conventions imposed very strict formatting, file naming, and output conventions.
-The program lacks a graphical interface or capabilities, forcing users to import generated text files into spreadsheets or other software to assess results and performing analyses.
-Furthermore, OUTBREAK can only handle one test at a time, creating barriers to batch operation and a large burden for researchers with a datasets including multiple sites.
-We developed `dfoliatR` as an `R`- and `dplR`-based library to overcome these issues.
+The program lacks a graphical interface or capabilities, forcing users to import generated text files into spreadsheets or other software to assess results and perform analyses.
+Furthermore, OUTBREAK can only handle one test at a time, creating barriers to batch operation and a large burden for researchers with datasets including multiple sites.
+We developed `dfoliatR` [@dfoliatR] as an `R`- and `dplR`-based library to overcome these issues.
 
@@
 The key objective of `dfoliatR` is to provide tools to identify and analyze insect defoliation and outbreak events by building on the methods employed by OUTBREAK.
-It capitalizes on the robust software already available in `R` by using `dplR` data formats for incoming tree-ring series and providing output data formats embodied by the `tidyverse` [@Wickham2019] that include efficient data manipulation [@Wickham2020dplyr] and graphics [@wickham2016ggplot2].
+It capitalizes on the robust software already available in `R` by using `dplR` data formats for incoming tree-ring series and providing output data formats embodied by the `tidyverse` [@Wickham2019] that include efficient data manipulation [`dplyr`: @Wickham2020dplyr] and graphics [`ggplot2`: @wickham2016ggplot2].
 
@@
 Support documentation in addition to this paper is provided within the package via standard help menus and on the package website ([https://chguiterman.github.io/dfoliatR/](https://chguiterman.github.io/dfoliatR/)), which includes up-to-date vignettes that describe various routines.
-Code to generate a preprint of this manuscript, including the `R` scripts and tubular and graphical output is available from [https://github.com/chguiterman/dfoliatR_paper](https://github.com/chguiterman/dfoliatR_paper).
+Code to generate a preprint of this manuscript, including the `R` scripts and tabular and graphical output is available from [https://github.com/chguiterman/dfoliatR_paper](https://github.com/chguiterman/dfoliatR_paper).
 
@@
 
-Negative departures in the normalized GSI (NGSI; GSI converted to z-scores) that surpass user-specified thresholds in duration and magnitude are defined as *defoliation events*.
-As in OUTBREAK, NGSI must reach the magnitude threshold, assessed for each year within the departure sequence.
+Negative departures in the normalized GSI (NGSI, or GSI converted to z-scores) that surpass user-specified thresholds in duration and magnitude are defined as *defoliation events*.
+As in OUTBREAK, the lowest NGSI value in the particucar sequence being assessed must reach the magnitude threshold.
 The default setting is -1.28 (NGSI is in units of standard deviation), which was previously determined to be representative of WSBW effects (Swetnam and Lynch, 1989) and is commonly used for other species.
-Failure to reach the specified threshold results in sequences not being inferred as defoliation events.
+The year with the lowest value is termed the "year of maximum departure" and becomes a central point in time for assessing other thresholds before being included as a defoliation event.
+If the year of maximum departure is higher than the threshold (i.e.
+$\textrm{NGSI}_{lowest} > -1.28$), the sequence being assessed is omitted from the event results.
 
-Event duration is assessed by examining sequences of negative NGSI (that exceed the magnitude threshold) before and after the year of maximum departure.
+Event duration is assessed by examining sequences of negative NGSI (that exceed the magnitude threshold) before and after the year of maximum departure year.
 Each defoliation event is allowed one single-year positive excursion on each side of the year of maximum departure.
-Duration is computed across the entire negative sequence that allows these two positive excursions.
+Duration is computed across the entire sequence that may include these two positive excursions.
 As in OUTBREAK, the user specifies a duration threshold (minimum number of years) for a departure sequence to be inferred as a defoliation event.
 The default threshold is eight years, as is commonly used in WSBW studies (Swetnam and Lynch, 1989).
-Failure to reach the specified duration threshold results in sequences not being inferred as defoliation events, regardless of NGSI values (i.e., both thresholds must be met).
+If the sequence is shorter than the duration threshold, the sequence is omitted from the event results (i.e., both thresholds must be met).
 Researchers can, and should, adjust the duration and magnitude parameters accordingly and critically evaluate the results, as insect species vary in the length of their outbreaks and the degree to which they can suppress tree growth.
@@
 In cases where two defoliation events are separated by a single year, bridging will link them into a single event.
-This option was added to `dfoliatR` during the testing phase of development, when we realized that OUTBREAK deliberately omits sequential, or back-to-back events, even when these events surpass thresholds in duration and magnitude.
-Instead, OUTBREAK will select the one sequential event with the greatest negative departure year.
+This option was added to `dfoliatR` during the testing phase of development, when we realized that OUTBREAK deliberately omits sequential, or back-to-back events, even when these events surpass both the magnitude and duration thresholds.
+Instead, OUTBREAK will select the one sequential event with the lowest negative departure year.
 In every case we assessed (described below) we felt that the OUTBREAK-omitted defoliation events should have been maintained and recorded.
@@
 This is particularly relevant to WSBW and spruce budworm (*C.
-fumiferana*), for which regimes have been reconstructed for several geographic areas and for which researchers and forest health experts have gained considerable knowledge [@Schmitt1984; @Sanders1985; @Brookes1987, and many later publications].
+fumiferana*), for which outbreak regimes have been reconstructed for several geographic areas and for which researchers and forest health experts have gained considerable knowledge [@Schmitt1984; @Sanders1985; @Brookes1987, and many later publications].
 In many cases, the greatest growth suppression often occurs late in the outbreak due to cumulative effects accrued on a tree's resources [@Swetnam1989; @Swetnam1993].
@@
 
-We tested `dfoliatR` against OUTBREAK by comparing NGSI to OUTBREAK's normalized corrected indices for individual trees and years, defoliation status for individual trees and years, and percentage of trees recording outbreaks at the site level, using standardized ring-width data from eight sites in British Columbia, Colorado, and New Mexico, and author-provided non-host site chronologies.
+We tested `dfoliatR` against OUTBREAK by comparing NGSI to OUTBREAK's normalized corrected indices for individual trees and years, defoliation status for individual trees and years, and percentage of trees recording outbreaks at the site level.
+Our tests used standardized ring-width data from eight sites in British Columbia, Colorado, and New Mexico, and author-provided non-host site chronologies.
 
@@
 These seven sites were sampled during ongoing outbreak events [@Swetnam1993].
-We found it necessary to be consistent in how we detrended and in what software (e.g., ARSTAN vs `dplR`) we employed because subtle differences in standardized ring-width indices generated between the programs transferred into differences between `dfoliatR` and OUTBREAK.
+We found it necessary to be consistent in how we detrended and what software we employed (e.g., ARSTAN vs `dplR`) because subtle differences in standardized ring-width indices generated between the programs transferred into differences between `dfoliatR` and OUTBREAK.
 In the end, we chose to only use the standardization output files from ARSTAN, which are easily read into `R` (and then `dfoliatR`) using the `dplR` package.
@@
 In two cases, OUTBREAK included events at the beginning of the series where `dfoliatR` does not allow truncated events.
-In four cases, OUTBREAK omitted only the last year of the series because the index was positive.
-In two cases, `dfoliatR` omitted possible events because it had already included a positive NGSI excursion after the "max_reduction" year, and since it will only allow one excursion on either side of the max year, the events were omitted due to short duration.
+In four cases, OUTBREAK omitted only the last year of the series because the index was positive, but `dfoliatR` allowed this single positive excursion.
+Finally, there were two cases in which `dfoliatR` omitted possible events because it had already included a positive NGSI excursion after the "max_reduction" year, and since it will only allow one excursion on either side of the max year, the events were omitted due to short duration.
 
-* *Sequential events* (36%) in which OUTBREAK omitted events that occur one year prior or one year following an identified event.
-When this occurs, OUTBREAK selects the event with the greatest negative index year.
+* *Sequential events* (36%) in which OUTBREAK omitted back-to-back events that occur one year prior or one year following an identified event.
+When this occurs, OUTBREAK selects the one event sequence with the lowest negative index year.
 On two trees, OUTBREAK omitted two of three sequential events.
-While inspecting these differences, we added an option to `defoliate_trees()` that would "bridge" between sequential events (that surpass magnitude and duration thresholds)into single, long events.
-We felt that this was ecologically justified, especially for studies of WSBW, because the insect will occasionally defoliate individual trees for decades and reconstructed chronologies show outbreaks as long as **xxx (ANN CHECK)**.
+While inspecting these differences, we added an option to `defoliate_trees()` that would "bridge" between sequential events (that surpass magnitude and duration thresholds) into single, long events.
+We felt that this was ecologically justified, especially for studies of WSBW, because the insect will occasionally defoliate individual trees for decades and reconstructed chronologies have show outbreaks as long as **xxx (ANN CHECK)**.
 
@@
 At the site level, OUTBREAK and `dfoliatR` produce similar time series of percent trees defoliated (Figure \@ref(fig:df-obr-site)), which forms the basis for inferring outbreak occurrence, intensity, and duration.
-Nearly all of the differences between the programs include either more trees or a longer duration of inferred outbreak by `dfoliatR`.
-These arise from the inclusion of events omitted by OUTBREAK.
-These differences translate to greater numbers of trees in defoliation during certain events, or events recorded by single trees during periods of low sample depth.
+In nearly site comparison, `dfoliatR` included either more events or longer durations of inferred outbreak.
+These differences arise from the inclusion of tree-level events by `dfoliatR` that were omitted by OUTBREAK.
+The differences translate to greater numbers of trees in defoliation during certain events, or events recorded by single trees during periods of low sample depth.
 
@@
 In every one of the 102 cases we inspected, we felt that `dfoliatR` provided a more biologically and statistically appropriate assessment of defoliation.
-Therefore, `dfoliatR` also did a better job inferring OUTBREAK events and associated statistics at the site level.
+Therefore, `dfoliatR` also did a better job inferring outbreak events and associated statistics at the site level.
 
@@
 
-The `dfoliatR` library [@dfoliatR] is provided free and open source from the Comprehensive R Archive Network (CRAN; [https://cran.r-project.org/](https://cran.r-project.org/)).
-To install `dfoliatR` from CRAN use
+The `dfoliatR` library is provided free and open source from the Comprehensive R Archive Network (CRAN; [https://cran.r-project.org/](https://cran.r-project.org/)). To install `dfoliatR` from CRAN use
 
@@
 Additional support was provided by the LTRR and the United States Forest Service Rocky Mountain Research Station.
-Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S.
-Government.
+The findings and conclusions in this publication are those of the author(s) and should not be construed to represent any official USDA or U.S.
+Government determination or policy.
 

### 2020-05-06 paper_elsevier/paper_elsevier.tex
more edits

@@
   We present a new R package to provide dendroecologists with tools to infer, quantify, analyze, and visualize growth suppression events in tree rings caused by insect defoliation, and to reconstruct defoliator outbreak chronologies.
-`dfoliatR' is based on the OUTBREAK program and builds on existing resources in the R and dplR computing environments.
+\texttt{dfoliatR} is based on the OUTBREAK program and builds on existing resources in the \texttt{R} and \texttt{dplR} computing environments.
 It performs an indexing procedure to remove climatic signals, represented in non-host chronologies or other reconstructed climate series, from the host-tree series.
@@
 Functions are provided for summary statistics and graphics of tree- and site-level series.
-`dfoliatR' expands on OUTBREAK to provide greater control of suppression thresholds, additional output tables, and high-quality graphics.
+\texttt{dfoliatR} expands on OUTBREAK to provide greater control of suppression thresholds, additional output tables, and high-quality graphics.
 We provide two example data sets and script to enable users to gain familiarity with the package and its capabilities.
 The source code is available in the Comprehensive R Archive Network (CRAN) and on GitHub.
-
   \hfill\break
   \end{abstract}
-   \begin{keyword} Dendroecology, spruce budworm, Choristoneura, pandora moth, Coloradia pandora Blake, larch-bud-moth
+   \begin{keyword} Dendroecology, spruce budworm, \emph{Choristoneura}, pandora moth, \emph{Coloradia pandora Blake}, larch-bud-moth
 \newpage\end{keyword}
@@
 \leavevmode\hypertarget{ref-Carus2009}{}%
----------. 2009. ``Effects of Defoliation Caused by the Processionary Moth on Growth of Crimean Pines in Western Turkey.'' \emph{Phytoparasitica} 37 (2): 105--14.
+---------. 2009. ``Effects of defoliation caused by the processionary moth on growth of Crimean pines in western Turkey.'' \emph{Phytoparasitica} 37 (2): 105--14.
 
@@
 \leavevmode\hypertarget{ref-Wickham2020dplyr}{}%
-Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2020. \emph{Dplyr: A Grammar of Data Manipulation}. R package version 0.8.4. \url{https://CRAN.R-project.org/package=dplyr}.
+Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2020. \emph{dplyr: A Grammar of Data Manipulation}. R package version 0.8.4. \url{https://CRAN.R-project.org/package=dplyr}.
 

### 2020-05-07 paper_elsevier/R/debug.R
remove temp script for debugging

@@ -1,24 +1,0 @@
-
-df_run <- meta_in %>% 
-  transmute(site,
-            host = map(host_file, ~ read.compact(paste0(data_pth, .))),
-            nonhost = map(nonhost_file, ~ read.crn(paste0(data_pth, .)) %>% 
-                            rm_sample_depth()),
-            defol = pmap(list(host, nonhost, ongoing_defol),
-                         ~ defoliate_trees(host_tree = ..1, 
-                                           nonhost_chron = ..2,
-                                           series_end_event = ..3,
-                                           duration_years = 8,
-                                           max_reduction = -1.28,
-                                           bridge_events = FALSE,
-                                           list_output = TRUE))
-  )
-
-fnl <- df_run$defol[[4]]           
-input_series <- fnl[[11]] 
-
-dmj <- df_run$defol[[1]]
-input_series <- dmj[[12]]
-
-fcc <- df_run$defol[[3]]
-input_series <- fcc[[28]]

### 2020-05-07 paper_elsevier/paper_elsevier.Rmd
Update paper_elsevier.Rmd

@@
 library(magrittr)
+library(here)
 # library(kableExtra)
@@
 
-We detrended host data for both `dfoliatR` and OUTBREAK using ARSTAN [v6.1, @arstan] with default double detrending (128 year wavelength and a 50% smoothing spline).
+We detrended host data for both `dfoliatR` and OUTBREAK using ARSTAN [v6.1] with cubic smoothing splines (50% frequency response on 100-150 year wavelengths depending on the site).
 In both `dfoliatR` and OUTBREAK we used event thresholds of -1.28 normalized indices, 8 years duration, and allowed for events at the end of series in seven of eight sites.
@@
 Each data set consists of individual host-tree series and a local non-host chronology.
-The host-tree series were standardized using 50-year splines with a 50% frequency response, while the non-host ring-width data were standardized using 150-year splines with a 50% frequency response and then averaged via Tukey's biweight robust mean procedure.
+The host-tree series were standardized using 128-year splines with a 50% frequency response, while the non-host ring-width data were standardized using 150-year splines with a 50% frequency response and then averaged via Tukey's biweight robust mean procedure.
 Host trees from Demijohn Peak (DMJ; 2902 m asl) in the San Juan Mountains of southern Colorado include Douglas-fir (*Pseudotsuga menziesii*) compared against a local non-host ponderosa pine (*Pinus ponderosa*) chronology [@Ryerson2003].
@@
 It is built on the long-accepted host to nonhost comparison methodology used in the 1980s FORTRAN program OUTBREAK [@Swetnam1985; @Swetnam1989; @Swetnam1993].
-Key benefits are that `dfoliatR` provides easier control of suppression thresholds, provides additional output tables and high-quality graphics, defines defoliation events more consistently, and operates in the open-source `R` environment.
+Our evaluation of the two programs revelaed that `dfolaitR` excelled in identifying defoliation events on single trees, providing researchers with more consistent and biologically-justifiable results.
+Other key benefits are that `dfoliatR` provides easier control of suppression thresholds, additional output tables, high-quality and customizable graphics, and operates in the open-source `R` environment that is stable across computing platforms and under active development and maintence by a large and growing community.
 
@@
 Outbreak reconstructions of insects for which the interval is notably longer than typical outbreak duration, such as Douglas-fir tussock moth (*Orgyia pseudotsugata* (Lepidoptera: Erebidae)), are unlikely to be significantly affected.
-Species for which outbreak duration varies considerably, for which individual tree resilience to defoliation varies considerably, or which may chronically infest trees or sites, or alternate between chronic, outbreak, and minimal activity states, such as conifer-feeding *Choristoneura*, present more complicated challenges, and the researcher should use discretion with the bridge option.
+Species for which outbreak duration or individual tree resilience to defoliation varies considerably, or which may chronically infest trees or sites, or alternate between chronic, outbreak, and minimal activity states such as conifer-feeding *Choristoneura*, present more complicated challenges, and the researcher should use discretion with the bridge option.
 
@@
 Source code for `dfoliatR` is available in the Comprehensive R Archive Network (CRAN) and GitHub [https://github.com/chguiterman/dfoliatR](https://github.com/chguiterman/dfoliatR) with updated descriptions and helpful vignettes on the package website [https://chguiterman.github.io/dfoliatR/](https://chguiterman.github.io/dfoliatR/).
-Researchers wishing to contribute to further development of `dfoliatR` are encouraged to do so via the GitHub repository.
+Researchers wishing to contribute to the further development of `dfoliatR` are encouraged to do so via the GitHub repository.
 
@@
 
-We are grateful to late Richard Holmes of the Laboratory of Tree-Ring Research (LTRR) for his achievements in developing OUTBREAK and many other tree-ring software programs, and we thank Thomas Swetnam for his insights and encouragement.
-The `dfoliatR` project was supported by the Agnes Haury Visiting Scholars Fellowship at the LTRR, the University of California Berkeley, and the Western Wildlands Environmental Threat Center (WWETAC).
+We are grateful to late Richard Holmes of the Laboratory of Tree-Ring Research (LTRR) for his achievements in developing OUTBREAK and many other tree-ring software programs.
+We also thank Thomas Swetnam for his insights and encouragement, and David Frank for his support.
+The `dfoliatR` project is funded by the Agnes Haury Visiting Scholars Fellowship at the LTRR, the University of California, Berkeley, and the Western Wildlands Environmental Threat Center (WWETAC).
 Additional support was provided by the LTRR and the United States Forest Service Rocky Mountain Research Station.
-The findings and conclusions in this publication are those of the author(s) and should not be construed to represent any official USDA or U.S.
+The findings and conclusions in this publication are those of the authors and should not be construed to represent any official USDA or U.S.
 Government determination or policy.
@@
 ```{r tbl-defol, echo=FALSE}
-tree_dat <- read.csv("../paper_elsevier/Output/defol-stats.csv")
+tree_dat <- read.csv(here("paper_elsevier", "Output", "defol-stats.csv"))
 kable(tree_dat, caption = "Summary statistics for individual tree defoliation events.", booktabs = TRUE) %>% 
@@
 ```{r tbl-obr, echo=FALSE}
-obr_dat <- read.csv("../paper_elsevier/Output/obr-stats.csv")
+obr_dat <- read.csv(here("paper_elsevier", "Output", "obr-stats.csv"))
 kable(obr_dat, format = "latex", caption = "Summary statistics for inferred outbreak events.", booktabs = TRUE) %>% 
@@
 Users can define the Cut-off values to determine severe-moderate-minor defoliation intensities, and use standard ggplot2 graphics parameters to adjust styles and themes."}
-include_graphics("../paper_elsevier/Output/tree-plot-default.pdf")
+include_graphics(here("paper_elsevier", "Output", "tree-plot-default.pdf"))
 ```
@@
 The middle panel shows the percent of trees defoliated, and the bottom panel shows the number of trees through time."}
-include_graphics("../paper_elsevier/Output/site-plot.pdf")
-```
-
-``` {r df-obr-1-1, echo=FALSE, fig.cap="Comparison of disturbance indices between dfoliatR and OUTBREAK. Diagonal lines show equal values (1:1)."}
-include_graphics("../paper_elsevier/Output/plot-trees-df-obr.pdf")
+include_graphics(here("paper_elsevier", "Output", "site-plot.pdf"))
 ```
@@
 ``` {r df-obr-site, echo=FALSE, fig.cap="Comparison of reconstructed western spruce budworm outbreaks computed by dfoliatR and OUTBREAK. Input parameters were identical between programs."}
-include_graphics("../paper_elsevier/Output/plot-sites-df-obr.pdf")
+include_graphics(here("paper_elsevier", "Output", "plot-sites-df-obr.pdf"))
 ```

### 2020-05-11 paper_elsevier/paper_elsevier.Rmd
edits before send to Jodi

@@
 `dfoliatR` is based on the OUTBREAK program and builds on existing resources in the `R` and `dplR` computing environments.
-It is designed wo aid research in the ecology of insect defoliation events and to reconstruct defoliator outbreak chronologies, but can be applied to other studies where host--non-host comparisons are useful.
-`It`dfoliatR` performs an indexing procedure to remove climatic signals in the host-tree series that are represented in the non-host chronologies, or other annually-resolved climate series.
+It is designed to aid research in the ecology of insect defoliation events and to reconstruct defoliator outbreak chronologies, but can be applied to other studies where host--non-host comparisons are useful.
+`dfoliatR` performs an indexing procedure to remove climatic signals in the host-tree series that are represented in the non-host chronologies, or other annually-resolved climate series.
 It then infers defoliation events in individual trees based on user-specified thresholds.
@@
 Functions are provided for summary statistics and graphics of tree- and site-level series.
-`dfoliatR` expands on OUTBREAK to provide greater control of suppression thresholds, additional output tables, and high-quality graphics.
+We evaluated `dfoliatR` against OUTBREAK, using eight datasets including 222 host-trees, and found that `dfoliatR` improves on OUTBREAK with greater user control, identification of defoliation events, computing capacity, and both the statistical summary and graphical outputs.
 We provide two example data sets and script to enable users to gain familiarity with the package and its capabilities.
 The source code is available in the Comprehensive R Archive Network (CRAN) and on GitHub.
- \hfill\break
+  <!-- \hfill\break -->
 keywords: |
   Dendroecology, spruce budworm, *Choristoneura*, pandora moth, *Coloradia pandora Blake*, larch-bud-moth
-  \newpage
-journal: "Dendrochronologia"
-date: "`r Sys.Date()`"
+  <!-- \newpage -->
+# journal: "Dendrochronologia"
+# date: "`r Sys.Date()`"
 bibliography: References.bib
@@
 # csl: elsevier-vancouver.csl
+csl: dendrochronologia.csl
 # output: rticles::elsevier_article
-output:
-  bookdown::pdf_book:
-      base_format: rticles::elsevier_article
-#   bookdown::word_document2:
-#     reference_docx: "../common/word-styles-reference-01.docx"
-#     toc: false
+# output:
+#   bookdown::pdf_book:
+#       base_format: rticles::elsevier_article
+bookdown::word_document2:
+    reference_docx: "../common/word-styles-reference-01.docx"
+    toc: false
 #     toc_depth: 2
-#     fig_caption: yes
-#     df_print: kable
+    fig_caption: yes
+    df_print: kable
 # always_allow_html: true
@@
   \usepackage{booktabs}
+  \makeatletter
+  \def\ps@pprintTitle{%
+   \let\@oddhead\@empty
+   \let\@evenhead\@empty
+   \let\@oddfoot\@empty
+   \let\@evenfoot\@oddfoot
+  }
+  \makeatother
 ---
@@
 `dfoliatR` adds to a growing suite of dendrochronology packages in the `R` computing environment [@RCore].
-Stemming from the `dplR` library [@Bunn2008] that enables `R` users to read and write an array of tree-ring data formats, standardize ring width series, build and evaluate chronologies, and perform quality control (to name a few), one can now also measure ring widths from scanned images of prepared samples [@Shi2019; @Lara2015], conduct and check crossdating [@Bunn2010], analyze sub-annual anotomical features [@Campelo2016; @Rathgeber2011], and perform many analytical tests [@Jevsenak2018; @Zang2015].
+Stemming from the `dplR` library [@Bunn2008] that enables `R` users to read and write an array of tree-ring data formats, standardize ring width series, build and evaluate chronologies, and perform quality control (to name a few), one can now also measure ring widths from scanned images of prepared samples [@Shi2019; @Lara2015], conduct and check crossdating [@Bunn2010], analyze sub-annual anatomical features [@Campelo2016; @Rathgeber2011], and perform many analytical tests [@Jevsenak2018; @Zang2015].
 Tools for assessing stand dynamics and disturbance analyses are under rapid development, with new packages for assessing growth and release events [`TRADER`: @Altman2014], metrics of growth resilience [`pointRes`: @VanderMaaten-Theunissen2015], and fire history [`burnr`: @Malevich2018].
@@
 
-In this paper, we describe the statistical methods employed by `dfoliatR`, compare results to those produced by OUTBREAK, and present an example analysis inlcuding test data sets and script.
+In this paper, we describe the statistical methods employed by `dfoliatR`, compare results to those produced by OUTBREAK, and present an example analysis including test data sets and script.
 Users need not have much experience in `R` to replicate the analyses and graphics as presented.
@@
 
+(ref:defolCap) Default `dfoliatR` graphics for individual trees, produced by the `plot_defol()` function.
+Dotted lines represent the series length for each tree and colored segments show periods of defoliation.
+Users can define the cut-off values to determine severe--moderate--minor defoliation intensities, and use standard `ggplot2` graphics parameters to adjust styles and themes.
+
+```{r fig-defol, echo=FALSE, fig.cap='(ref:defolCap)'}
+include_graphics(here("paper_elsevier", "Output", "tree-plot-default.pdf"))
+```
+
+(ref:obrCap) Default graphics showing outbreak events at the DMJ site.
+Top panel shows the mean GSI through time, with inferred outbreak events filled.
+The middle panel shows the percent of trees defoliated, and the bottom panel shows the number of trees through time.
+
+```{r fig-obr, echo=FALSE, fig.cap='(ref:obrCap)'}
+include_graphics(here("paper_elsevier", "Output", "site-plot.pdf"))
+```
+
 ## Identifying Defoliation of Trees
@@
 Negative departures in the normalized GSI (NGSI, or GSI converted to z-scores) that surpass user-specified thresholds in duration and magnitude are defined as *defoliation events*.
-As in OUTBREAK, the lowest NGSI value in the particucar sequence being assessed must reach the magnitude threshold.
+As in OUTBREAK, the lowest NGSI value in the particular sequence being assessed must reach the magnitude threshold.
 The default setting is -1.28 (NGSI is in units of standard deviation), which was previously determined to be representative of WSBW effects (Swetnam and Lynch, 1989) and is commonly used for other species.
@@
 
-Diverging from OUTBREAK, `dfoliatR` allows users to extend defoliation events on individual trees by bridging between sequential events.
+Diverging from OUTBREAK, `dfoliatR` allows users to extend defoliation events on individual trees by bridging between sequential events (Figure \@ref(fig:fig-bridge)).
 In cases where two defoliation events are separated by a single year, bridging will link them into a single event.
@@
 In every case we assessed (described below) we felt that the OUTBREAK-omitted defoliation events should have been maintained and recorded.
-Due in large part to reconstructions using OUTBREAK [see papers cited by @Lynch2012], we now know considerably more about forest defoliator outbreak regimes than we did in the 1980s when OUTBREAK was written.
+Due in large part to reconstructions using OUTBREAK [see papers cited by @Lynch2012], we now know considerably more about forest defoliator outbreak regimes than we did in the 1980s when OUTBREAK was under development.
 We think that two or more prolonged events separated by a single year should in some situations be considered a single event.
@@
 We urge caution in using the bridging option, however, because it may not be appropriate for all studied insects, such as in situations where impacted stands barely recover from one outbreak before another begins, as with pine processionary caterpillars (*Thaumetopoea pityocampa* (Lepidoptera: Thaumetopoeidae)) [@Carus2004; @Carus2009].
+
+(ref:BridgeCap) Examples of sequential defoliation events (bars), wherein two or three events occur separated by single years.
+In each case OUTBREAK recorded only one of the events, omitting those with a higher "max_defol" year, or the year of the lowest NGSI value.
+`dfoliatR`, by contrast, recorded all of the events, including two sequential events on DMJ26 and EFK22, and three on FNL04.
+Events recorded by both programs have a darker color, while those recorded only by `dfoliatR` are lighter in color.
+Adding the bridging option in `dfoliatR` links these events ("bridge_defol") into single longer events that may more accurately capture the effects of long-term insect herbivory on individual trees.
+
+```{r fig-bridge, echo=FALSE, fig.cap='(ref:BridgeCap)'}
+include_graphics(here("paper_elsevier", "Output", "bridging.pdf"))
+```
 
@@
 * *Sequential events* (36%) in which OUTBREAK omitted back-to-back events that occur one year prior or one year following an identified event.
-When this occurs, OUTBREAK selects the one event sequence with the lowest negative index year.
+When this occurs, OUTBREAK selects the one event sequence with the lowest negative index year (e.g., Figure \@ref(fig:fig-bridge)).
 On two trees, OUTBREAK omitted two of three sequential events.
@@
 
-**(Bridging paragraph?)**
-
-At the site level, OUTBREAK and `dfoliatR` produce similar time series of percent trees defoliated (Figure \@ref(fig:df-obr-site)), which forms the basis for inferring outbreak occurrence, intensity, and duration.
+At the site level, OUTBREAK and `dfoliatR` produce similar time series of percent trees defoliated (Figure \@ref(fig:fig-df-obr)), which forms the basis for inferring outbreak occurrence, intensity, and duration.
 In nearly site comparison, `dfoliatR` included either more events or longer durations of inferred outbreak.
@@
 The differences translate to greater numbers of trees in defoliation during certain events, or events recorded by single trees during periods of low sample depth.
+
+(ref:df-obrCap) Comparison of reconstructed western spruce budworm outbreaks computed by `dfoliatR` and OUTBREAK.
+Input parameters were identical between programs.
+Difference arise because `dfoliatR` will identify and record more defoliation events on individual trees, and may more accurately capture herbivory dynamics.
+
+``` {r fig-df-obr, echo=FALSE, fig.cap='(ref:df-obrCap)'}
+include_graphics(here("paper_elsevier", "Output", "plot-sites-df-obr.pdf"))
+```
 
@@
 
+```{r tbl-defol, echo=FALSE}
+tree_dat <- read.csv(here("paper_elsevier", "Output", "defol-stats.csv"))
+kable(tree_dat, caption = "Summary statistics for individual tree defoliation events.", booktabs = TRUE) %>% 
+  kableExtra::kable_styling(latex_options = "hold_position")
+```
+
 The `plot_defol()` function produces a "ggplot" graphics object with line segments showing the measured sequence of each series and a filled segment for each identified defoliation event (Figure \@ref(fig:fig-defol)).
@@
 
+```{r tbl-obr, echo=FALSE}
+obr_dat <- read.csv(here("paper_elsevier", "Output", "obr-stats.csv"))
+kable(obr_dat, format = "latex", caption = "Summary statistics for inferred outbreak events.", booktabs = TRUE) %>% 
+  kable_styling(font_size = 7) %>% 
+  kableExtra::landscape()
+```
+
 The output table from `outbreak_stats()` shows the start and end years of each event, along with their corresponding duration, the number and percent of trees in defoliation at the start of the event, the the number of trees recording the outbreak event (in part or full), the minimums of the mean GSI and NGSI indices, and two points of "peak defoliation." The first, "peak_outbreak_year" is the year in which the greatest number of trees are recording the outbreak, and the second, "peak_defol_year" is the year with the lowest average NGSI.
@@
 The `dfoliatR` package provides dendroecologists with tools to infer, quantify, analyze, and visualize tree-ring growth suppression events and to reconstruct forest insect defoliator outbreak regimes.
-It is built on the long-accepted host to nonhost comparison methodology used in the 1980s FORTRAN program OUTBREAK [@Swetnam1985; @Swetnam1989; @Swetnam1993].
-Our evaluation of the two programs revelaed that `dfolaitR` excelled in identifying defoliation events on single trees, providing researchers with more consistent and biologically-justifiable results.
-Other key benefits are that `dfoliatR` provides easier control of suppression thresholds, additional output tables, high-quality and customizable graphics, and operates in the open-source `R` environment that is stable across computing platforms and under active development and maintence by a large and growing community.
+It is built on the long-accepted host to non-host comparison methodology used in the 1980s FORTRAN program OUTBREAK [@Swetnam1985; @Swetnam1989; @Swetnam1993].
+Our evaluation of the two programs revealed that `dfoliatR` excelled in identifying defoliation events on single trees, providing researchers with more consistent and biologically-justifiable results.
+Other key benefits are that `dfoliatR` provides easier control of suppression thresholds, additional output tables, high-quality and customizable graphics, and operates in the open-source `R` environment that is stable across computing platforms and under active development and maintenance by a large and growing community.
 
@@
 
-<div id="refs"></div>
-\newpage
+<!-- <div id="refs"></div> -->
+<!-- \newpage -->
 
-# Tables & Figures {-}
-
-```{r tbl-defol, echo=FALSE}
-tree_dat <- read.csv(here("paper_elsevier", "Output", "defol-stats.csv"))
-kable(tree_dat, caption = "Summary statistics for individual tree defoliation events.", booktabs = TRUE) %>% 
-  kableExtra::kable_styling(latex_options = "hold_position")
-```
-
-```{r tbl-obr, echo=FALSE}
-obr_dat <- read.csv(here("paper_elsevier", "Output", "obr-stats.csv"))
-kable(obr_dat, format = "latex", caption = "Summary statistics for inferred outbreak events.", booktabs = TRUE) %>% 
-  kable_styling(font_size = 7) %>% 
-  kableExtra::landscape()
-```
-
-```{r fig-defol, echo=FALSE, fig.cap="Default dfoliatR graphics for individual trees, produced by the 'plot_defol()' function.
-Dotted lines represent the series length for each tree and colored segments show periods of defoliation.
-Users can define the Cut-off values to determine severe-moderate-minor defoliation intensities, and use standard ggplot2 graphics parameters to adjust styles and themes."}
-include_graphics(here("paper_elsevier", "Output", "tree-plot-default.pdf"))
-```
-
-```{r fig-obr, echo=FALSE, fig.cap="Default graphic to show outbreak events at the DMJ site.
-Top panel shows the mean GSI through time, with inferred outbreak events filled.
-The middle panel shows the percent of trees defoliated, and the bottom panel shows the number of trees through time."}
-include_graphics(here("paper_elsevier", "Output", "site-plot.pdf"))
-```
-
-``` {r df-obr-site, echo=FALSE, fig.cap="Comparison of reconstructed western spruce budworm outbreaks computed by dfoliatR and OUTBREAK. Input parameters were identical between programs."}
-include_graphics(here("paper_elsevier", "Output", "plot-sites-df-obr.pdf"))
-```
+<!-- # Tables & Figures {-} -->

### 2020-05-11 paper_elsevier/paper_elsevier.tex
edits before send to Jodi

@@
 \usepackage[hyphens]{url}
-
-  \journal{Dendrochronologia} % Sets Journal name
 
@@
 \usepackage{booktabs}
+\makeatletter
+\def\ps@pprintTitle{%
+ \let\@oddhead\@empty
+ \let\@evenhead\@empty
+ \let\@oddfoot\@empty
+ \let\@evenfoot\@oddfoot
+}
+\makeatother
 
@@
   \begin{abstract}
-  We present a new R package to provide dendroecologists with tools to infer, quantify, analyze, and visualize growth suppression events in tree rings caused by insect defoliation, and to reconstruct defoliator outbreak chronologies.
+  We present a new R package to provide dendroecologists with tools to infer, quantify, analyze, and visualize growth suppression events in tree rings.
 \texttt{dfoliatR} is based on the OUTBREAK program and builds on existing resources in the \texttt{R} and \texttt{dplR} computing environments.
-It performs an indexing procedure to remove climatic signals, represented in non-host chronologies or other reconstructed climate series, from the host-tree series.
+It is designed to aid research in the ecology of insect defoliation events and to reconstruct defoliator outbreak chronologies, but can be applied to other studies where host--non-host comparisons are useful.
+\texttt{dfoliatR} performs an indexing procedure to remove climatic signals in the host-tree series that are represented in the non-host chronologies, or other annually-resolved climate series.
 It then infers defoliation events in individual trees based on user-specified thresholds.
@@
 Functions are provided for summary statistics and graphics of tree- and site-level series.
-\texttt{dfoliatR} expands on OUTBREAK to provide greater control of suppression thresholds, additional output tables, and high-quality graphics.
+We evaluated \texttt{dfoliatR} against OUTBREAK, using eight datasets including 222 host-trees, and found that \texttt{dfoliatR} improves on OUTBREAK with greater user control, identification of defoliation events, computing capacity, and both the statistical summary and graphical outputs.
 We provide two example data sets and script to enable users to gain familiarity with the package and its capabilities.
 The source code is available in the Comprehensive R Archive Network (CRAN) and on GitHub.
-  \hfill\break
   \end{abstract}
-   \begin{keyword} Dendroecology, spruce budworm, \emph{Choristoneura}, pandora moth, \emph{Coloradia pandora Blake}, larch-bud-moth
-\newpage\end{keyword}
+   \begin{keyword} Dendroecology, spruce budworm, \emph{Choristoneura}, pandora moth, \emph{Coloradia pandora Blake}, larch-bud-moth\end{keyword}
  \end{frontmatter}
@@
 Defoliation can be distinguished from climate- and other disturbance-related influences by comparing ring-width or other annually-resolved features in the wood of host species to that of non-host species or annually-resolved climate records.
-The effect of defoliation on radial growth of trees has been recognized since the 1860s, and used to reconstruct outbreak regimes since the 1950s (Blais 1954; Lynch 2012).
-It was not until the 1980s, however, that precise dendrochronological techniques were applied for inferring defoliation events and reconstructing defoliator outbreak regimes (Swetnam, Thompson, and Sutherland 1985; Speer 2010; Lynch 2012).
-The first studies (Swetnam, Thompson, and Sutherland 1985; Swetnam and Lynch 1989, 1993) focused on developing historical reconstructions of western spruce budworm (WSBW; \emph{Choristoneura freemani} (Lepidoptera: Tortricidae); previously known as \emph{C.
-occidentalis}) outbreak regimes.
-The methodology has since been successfully applied to a wide range of defoliator species, most of which are conifer herbivores, and has evolved in sophistication for a wide range of ecosystem situations (Lynch 2012).
+The effect of defoliation on radial growth of trees has been recognized since the 1860s, and used to reconstruct outbreak regimes since the 1950s (Blais, 1954; Lynch, 2012).
+It was not until the 1980s, however, that precise dendrochronological techniques were applied for inferring defoliation events and reconstructing defoliator outbreak regimes (Swetnam et al., 1985; Speer, 2010; Lynch, 2012).
+The first studies (Swetnam et al., 1985; Swetnam and Lynch, 1989, 1993) focused on developing historical outbreak reconstructions of western spruce budworm (WSBW; \emph{Choristoneura freemani} (Lepidoptera: Tortricidae); previously known as \emph{C.
+occidentalis}).
+The methodology has since been successfully applied to a wide range of defoliator species, most of which are conifer herbivores, and has evolved in sophistication for a wide range of ecosystem situations (Lynch, 2012).
 
-The main dendrochronological tool for inferring, dating, and characterizing defoliator outbreaks from tree-ring records has been the software routine OUTBREAK (Swetnam, Thompson, and Sutherland 1985; Holmes and Swetnam 1986; Swetnam and Lynch 1989).
+The main dendrochronological tool for inferring, dating, and characterizing defoliator outbreaks from tree-ring records has been the software routine OUTBREAK (Swetnam et al., 1985; Holmes and Swetnam, 1986; Swetnam and Lynch, 1989).
 OUTBREAK computes indices (described later in detail) of suppressed growth by subtracting a detrended and standardized climate series (a ``control'' chronology) from individual host-tree detrended and standardized radial growth series after the host and non-host series have been brought to a common variance.
-The non-host chronology usually consists of a site chronology developed from non-host tree species growing on a climate-sensitive site, but a gridded climate data point series, like the North American Drought Atlas (Cook and Krusic 2004) also suffices.
+The non-host chronology usually consists of a site chronology developed from non-host tree species growing on a climate-sensitive site, but a gridded climate data point series, like the North American Drought Atlas (Cook and Krusic, 2004) also suffices.
 If the host and non-host species respond similarly to climate (which can and should be tested), the derived series retains variability that the host and non-host series do not have in common, generally the insect signal and some unexplained variability (noise).
@@
 It was written in FORTRAN V with inherently severe restrictions, as RAM and disk space were limited at that time (256 kb and 10 MB, respectively) and FORTRAN conventions imposed very strict formatting, file naming, and output conventions.
-The program lacks a graphical interface or capabilities, forcing users to import generated text files into spreadsheets or other software to assess results and performing analyses.
-Furthermore, OUTBREAK can only handle one test at a time, creating barriers to batch operation and a large burden for researchers with a datasets including multiple sites.
-We developed \texttt{dfoliatR} as an \texttt{R}- and \texttt{dplR}-based library to overcome these issues.
+The program lacks a graphical interface or capabilities, forcing users to import generated text files into spreadsheets or other software to assess results and perform analyses.
+Furthermore, OUTBREAK can only handle one test at a time, creating barriers to batch operation and a large burden for researchers with datasets including multiple sites.
+We developed \texttt{dfoliatR} (Guiterman et al., 2020) as an \texttt{R}- and \texttt{dplR}-based library to overcome these issues.
 
-\texttt{dfoliatR} adds to a growing suite of dendrochronology packages in the \texttt{R} computing environment (R Core Team 2019).
-Stemming from the \texttt{dplR} library (Bunn 2008) that enables \texttt{R} users to read and write an array of tree-ring data formats, standardize ring width series, build and evaluate chronologies, and perform quality control (to name a few), one can now also measure ring widths from scanned images of prepared samples (Shi et al.
-2019; Lara, Bravo, and Sierra 2015), conduct and check crossdating (Bunn 2010), analyze sub-annual anotomical features (Campelo et al.
-2016; Rathgeber et al.
-2011), and perform many analytical tests (Jevšenak and Levanič 2018; Zang and Biondi 2015).
-Tools for assessing stand dynamics and disturbance analyses are under rapid development, with new packages for assessing growth and release events (\texttt{TRADER}: Altman et al.
-2014), metrics of growth resilience (\texttt{pointRes}: van der Maaten-Theunissen, van der Maaten, and Bouriaud 2015), and fire history (\texttt{burnr}: Malevich, Guiterman, and Margolis 2018).
+\texttt{dfoliatR} adds to a growing suite of dendrochronology packages in the \texttt{R} computing environment (R Core Team, 2019).
+Stemming from the \texttt{dplR} library (Bunn, 2008) that enables \texttt{R} users to read and write an array of tree-ring data formats, standardize ring width series, build and evaluate chronologies, and perform quality control (to name a few), one can now also measure ring widths from scanned images of prepared samples (Lara et al., 2015; Shi et al., 2019), conduct and check crossdating (Bunn, 2010), analyze sub-annual anotomical features (Rathgeber et al., 2011; Campelo et al., 2016), and perform many analytical tests (Zang and Biondi, 2015; Jevšenak and Levanič, 2018).
+Tools for assessing stand dynamics and disturbance analyses are under rapid development, with new packages for assessing growth and release events (\texttt{TRADER}: Altman et al., 2014), metrics of growth resilience (\texttt{pointRes}: van der Maaten-Theunissen et al., 2015), and fire history (\texttt{burnr}: Malevich et al., 2018).
 The key objective of \texttt{dfoliatR} is to provide tools to identify and analyze insect defoliation and outbreak events by building on the methods employed by OUTBREAK.
-It capitalizes on the robust software already available in \texttt{R} by using \texttt{dplR} data formats for incoming tree-ring series and providing output data formats embodied by the \texttt{tidyverse} (Wickham et al.
-2019) that include efficient data manipulation (Wickham et al.
-2020) and graphics (Wickham 2016).
+It capitalizes on the robust software already available in \texttt{R} by using \texttt{dplR} data formats for incoming tree-ring series and providing output data formats embodied by the \texttt{tidyverse} (Wickham et al., 2019) that include efficient data manipulation (\texttt{dplyr}: Wickham et al., 2020) and graphics (\texttt{ggplot2}: Wickham, 2016).
 
@@
 Support documentation in addition to this paper is provided within the package via standard help menus and on the package website (\url{https://chguiterman.github.io/dfoliatR/}), which includes up-to-date vignettes that describe various routines.
-Code to generate a preprint of this manuscript, including the \texttt{R} scripts and tubular and graphical output is available from \url{https://github.com/chguiterman/dfoliatR_paper}.
+Code to generate a preprint of this manuscript, including the \texttt{R} scripts and tabular and graphical output is available from \url{https://github.com/chguiterman/dfoliatR_paper}.
 
@@
 
-Users can develop these data sets in the software of their choosing, such as \texttt{dplR} or ARSTAN (Cook and Holmes 1996).
+Users can develop these data sets in the software of their choosing, such as \texttt{dplR} or ARSTAN (Cook and Holmes, 1996).
 It is important that the host-tree data include only one tree-ring series per tree.
@@
 These identify defoliation events on individual trees (Figure \ref{fig:fig-defol}) and then composite across multiple trees to identify outbreak events (Figure \ref{fig:fig-obr}).
+
+\begin{figure}
+\centering
+\includegraphics{D:/Dropbox/dfoliatR_paper/paper_elsevier/Output/tree-plot-default.pdf}
+\caption{\label{fig:fig-defol}Default \texttt{dfoliatR} graphics for individual trees, produced by the \texttt{plot\_defol()} function.
+Dotted lines represent the series length for each tree and colored segments show periods of defoliation.
+Users can define the cut-off values to determine severe--moderate--minor defoliation intensities, and use standard \texttt{ggplot2} graphics parameters to adjust styles and themes.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{D:/Dropbox/dfoliatR_paper/paper_elsevier/Output/site-plot.pdf}
+\caption{\label{fig:fig-obr}Default graphics showing outbreak events at the DMJ site.
+Top panel shows the mean GSI through time, with inferred outbreak events filled.
+The middle panel shows the percent of trees defoliated, and the bottom panel shows the number of trees through time.}
+\end{figure}
 
@@
 
-Negative departures in the normalized GSI (NGSI; GSI converted to z-scores) that surpass user-specified thresholds in duration and magnitude are defined as \emph{defoliation events}.
-As in OUTBREAK, NGSI must reach the magnitude threshold, assessed for each year within the departure sequence.
+Negative departures in the normalized GSI (NGSI, or GSI converted to z-scores) that surpass user-specified thresholds in duration and magnitude are defined as \emph{defoliation events}.
+As in OUTBREAK, the lowest NGSI value in the particucar sequence being assessed must reach the magnitude threshold.
 The default setting is -1.28 (NGSI is in units of standard deviation), which was previously determined to be representative of WSBW effects (Swetnam and Lynch, 1989) and is commonly used for other species.
-Failure to reach the specified threshold results in sequences not being inferred as defoliation events.
+The year with the lowest value is termed the ``year of maximum departure'' and becomes a central point in time for assessing other thresholds before being included as a defoliation event.
+If the year of maximum departure is higher than the threshold (i.e.~\(\textrm{NGSI}_{lowest} > -1.28\)), the sequence being assessed is omitted from the event results.
 
-Event duration is assessed by examining sequences of negative NGSI (that exceed the magnitude threshold) before and after the year of maximum departure.
+Event duration is assessed by examining sequences of negative NGSI (that exceed the magnitude threshold) before and after the year of maximum departure year.
 Each defoliation event is allowed one single-year positive excursion on each side of the year of maximum departure.
-Duration is computed across the entire negative sequence that allows these two positive excursions.
+Duration is computed across the entire sequence that may include these two positive excursions.
 As in OUTBREAK, the user specifies a duration threshold (minimum number of years) for a departure sequence to be inferred as a defoliation event.
 The default threshold is eight years, as is commonly used in WSBW studies (Swetnam and Lynch, 1989).
-Failure to reach the specified duration threshold results in sequences not being inferred as defoliation events, regardless of NGSI values (i.e., both thresholds must be met).
+If the sequence is shorter than the duration threshold, the sequence is omitted from the event results (i.e., both thresholds must be met).
 Researchers can, and should, adjust the duration and magnitude parameters accordingly and critically evaluate the results, as insect species vary in the length of their outbreaks and the degree to which they can suppress tree growth.
@@
 
-Diverging from OUTBREAK, \texttt{dfoliatR} allows users to extend defoliation events on individual trees by bridging between sequential events.
+Diverging from OUTBREAK, \texttt{dfoliatR} allows users to extend defoliation events on individual trees by bridging between sequential events (Figure \ref{fig:fig-bridge}).
 In cases where two defoliation events are separated by a single year, bridging will link them into a single event.
-This option was added to \texttt{dfoliatR} during the testing phase of development, when we realized that OUTBREAK deliberately omits sequential, or back-to-back events, even when these events surpass thresholds in duration and magnitude.
-Instead, OUTBREAK will select the one sequential event with the greatest negative departure year.
+This option was added to \texttt{dfoliatR} during the testing phase of development, when we realized that OUTBREAK deliberately omits sequential, or back-to-back events, even when these events surpass both the magnitude and duration thresholds.
+Instead, OUTBREAK will select the one sequential event with the lowest negative departure year.
 In every case we assessed (described below) we felt that the OUTBREAK-omitted defoliation events should have been maintained and recorded.
-Due in large part to reconstructions using OUTBREAK (see papers cited by Lynch 2012), we now know considerably more about forest defoliator outbreak regimes than we did in the 1980s when OUTBREAK was written.
+Due in large part to reconstructions using OUTBREAK (see papers cited by Lynch, 2012), we now know considerably more about forest defoliator outbreak regimes than we did in the 1980s when OUTBREAK was under development.
 We think that two or more prolonged events separated by a single year should in some situations be considered a single event.
 This is particularly relevant to WSBW and spruce budworm (\emph{C.
-fumiferana}), for which regimes have been reconstructed for several geographic areas and for which researchers and forest health experts have gained considerable knowledge (Schmitt, Grimble, and Searcy 1984; Sanders et al.
-1985; Brookes et al.
-1982, and many later publications).
-In many cases, the greatest growth suppression often occurs late in the outbreak due to cumulative effects accrued on a tree's resources (Swetnam and Lynch 1989, 1993).
-We urge caution in using the bridging option, however, because it may not be appropriate for all studied insects, such as in situations where impacted stands barely recover from one outbreak before another begins, as with pine processionary caterpillars (\emph{Thaumetopoea pityocampa} (Lepidoptera: Thaumetopoeidae)) (Carus 2004, 2009).
+fumiferana}), for which outbreak regimes have been reconstructed for several geographic areas and for which researchers and forest health experts have gained considerable knowledge (Brookes et al., 1982, and many later publications; Schmitt et al., 1984; Sanders et al., 1985).
+In many cases, the greatest growth suppression often occurs late in the outbreak due to cumulative effects accrued on a tree's resources (Swetnam and Lynch, 1989, 1993).
+We urge caution in using the bridging option, however, because it may not be appropriate for all studied insects, such as in situations where impacted stands barely recover from one outbreak before another begins, as with pine processionary caterpillars (\emph{Thaumetopoea pityocampa} (Lepidoptera: Thaumetopoeidae)) (Carus, 2004, 2009).
+
+\begin{figure}
+\centering
+\includegraphics{D:/Dropbox/dfoliatR_paper/paper_elsevier/Output/bridging.pdf}
+\caption{\label{fig:fig-bridge}Examples of sequential defoliation events (bars), wherein two or three events occur separated by single years.
+In each case OUTBREAK recorded only one of the events, ommitting those with a higher ``max\_defol'' year, or the year of the lowest NGSI value.
+\texttt{dfoliatR}, by contrast, recorded all of the events, including two sequential events on DMJ26 and EFK22, and three on FNL04.
+Events recorded by both programs have a darker color, while those recorded only by \texttt{dfoliatR} are lighter in color.
+Adding the bridging option in \texttt{dfoliatR} links these events (``bridge\_defol'') into single longer events that may more accurately capture the effects of long-term insect herbivory on individual trees.}
+\end{figure}
 
@@
 Three parameters control whether a defoliation event constitutes an outbreak: the minimum number of trees available, the minimum number of trees recording defoliation, and the percent of trees recording defoliation.
-The first allows the researcher to make a judgement call as to the confidence ascribed to reduced sample depth toward the ends of their chronologies, thus compensating for the ``fading record problem'' (Swetnam, Allen, and Betancourt 1999).
+The first allows the researcher to make a judgement call as to the confidence ascribed to reduced sample depth toward the ends of their chronologies, thus compensating for the ``fading record problem'' (Swetnam et al., 1999).
 The second two parameters adjust the scale of defoliation considered to be an outbreak.
-Absolute numbers of trees and percentages can be applied separately or in conjunction, following filtering conventions in tree-ring fire history studies (Malevich, Guiterman, and Margolis 2018).
+Absolute numbers of trees and percentages can be applied separately or in conjunction, following filtering conventions in tree-ring fire history studies (Malevich et al., 2018).
 We urge users to carefully consider the choice of absolute numbers in situations where the number of trees represented in the series varies with time, or the choice of percentages when sample size is small.
@@
 
-We tested \texttt{dfoliatR} against OUTBREAK by comparing NGSI to OUTBREAK's normalized corrected indices for individual trees and years, defoliation status for individual trees and years, and percentage of trees recording outbreaks at the site level, using standardized ring-width data from eight sites in British Columbia, Colorado, and New Mexico, and author-provided non-host site chronologies.
+We tested \texttt{dfoliatR} against OUTBREAK by comparing NGSI to OUTBREAK's normalized corrected indices for individual trees and years, defoliation status for individual trees and years, and percentage of trees recording outbreaks at the site level.
+Our tests used standardized ring-width data from eight sites in British Columbia, Colorado, and New Mexico, and author-provided non-host site chronologies.
 
-We detrended host data for both \texttt{dfoliatR} and OUTBREAK using ARSTAN (v6.1, Cook and Holmes 1996) with default double detrending (128 year wavelength and a 50\% smoothing spline).
+We detrended host data for both \texttt{dfoliatR} and OUTBREAK using ARSTAN {[}v6.1{]} with cubic smoothing splines (50\% frequency response on 100-150 year wavelengths depending on the site).
 In both \texttt{dfoliatR} and OUTBREAK we used event thresholds of -1.28 normalized indices, 8 years duration, and allowed for events at the end of series in seven of eight sites.
-These seven sites were sampled during ongoing outbreak events (Swetnam and Lynch 1993).
-We found it necessary to be consistent in how we detrended and in what software (e.g., ARSTAN vs \texttt{dplR}) we employed because subtle differences in standardized ring-width indices generated between the programs transferred into differences between \texttt{dfoliatR} and OUTBREAK.
+These seven sites were sampled during ongoing outbreak events (Swetnam and Lynch, 1993).
+We found it necessary to be consistent in how we detrended and what software we employed (e.g., ARSTAN vs \texttt{dplR}) because subtle differences in standardized ring-width indices generated between the programs transferred into differences between \texttt{dfoliatR} and OUTBREAK.
 In the end, we chose to only use the standardization output files from ARSTAN, which are easily read into \texttt{R} (and then \texttt{dfoliatR}) using the \texttt{dplR} package.
@@
 In two cases, OUTBREAK included events at the beginning of the series where \texttt{dfoliatR} does not allow truncated events.
-In four cases, OUTBREAK omitted only the last year of the series because the index was positive.
-In two cases, \texttt{dfoliatR} omitted possible events because it had already included a positive NGSI excursion after the ``max\_reduction'' year, and since it will only allow one excursion on either side of the max year, the events were omitted due to short duration.
+In four cases, OUTBREAK omitted only the last year of the series because the index was positive, but \texttt{dfoliatR} allowed this single positive excursion.
+Finally, there were two cases in which \texttt{dfoliatR} omitted possible events because it had already included a positive NGSI excursion after the ``max\_reduction'' year, and since it will only allow one excursion on either side of the max year, the events were omitted due to short duration.
 \item
-  \emph{Sequential events} (36\%) in which OUTBREAK omitted events that occur one year prior or one year following an identified event.
-When this occurs, OUTBREAK selects the event with the greatest negative index year.
+  \emph{Sequential events} (36\%) in which OUTBREAK omitted back-to-back events that occur one year prior or one year following an identified event.
+When this occurs, OUTBREAK selects the one event sequence with the lowest negative index year (e.g., Figure \ref{fig:fig-bridge}).
 On two trees, OUTBREAK omitted two of three sequential events.
-While inspecting these differences, we added an option to \texttt{defoliate\_trees()} that would ``bridge'' between sequential events (that surpass magnitude and duration thresholds)into single, long events.
-We felt that this was ecologically justified, especially for studies of WSBW, because the insect will occasionally defoliate individual trees for decades and reconstructed chronologies show outbreaks as long as \textbf{xxx (ANN CHECK)}.
+While inspecting these differences, we added an option to \texttt{defoliate\_trees()} that would ``bridge'' between sequential events (that surpass magnitude and duration thresholds) into single, long events.
+We felt that this was ecologically justified, especially for studies of WSBW, because the insect will occasionally defoliate individual trees for decades and reconstructed chronologies have show outbreaks as long as \textbf{xxx (ANN CHECK)}.
 \item
@@
 
-\textbf{(Bridging paragraph?)}
+At the site level, OUTBREAK and \texttt{dfoliatR} produce similar time series of percent trees defoliated (Figure \ref{fig:fig-df-obr}), which forms the basis for inferring outbreak occurrence, intensity, and duration.
+In nearly site comparison, \texttt{dfoliatR} included either more events or longer durations of inferred outbreak.
+These differences arise from the inclusion of tree-level events by \texttt{dfoliatR} that were omitted by OUTBREAK.
+The differences translate to greater numbers of trees in defoliation during certain events, or events recorded by single trees during periods of low sample depth.
 
-At the site level, OUTBREAK and \texttt{dfoliatR} produce similar time series of percent trees defoliated (Figure \ref{fig:df-obr-site}), which forms the basis for inferring outbreak occurrence, intensity, and duration.
-Nearly all of the differences between the programs include either more trees or a longer duration of inferred outbreak by \texttt{dfoliatR}.
-These arise from the inclusion of events omitted by OUTBREAK.
-These differences translate to greater numbers of trees in defoliation during certain events, or events recorded by single trees during periods of low sample depth.
+\begin{figure}
+\centering
+\includegraphics{D:/Dropbox/dfoliatR_paper/paper_elsevier/Output/plot-sites-df-obr.pdf}
+\caption{\label{fig:fig-df-obr}Comparison of reconstructed western spruce budworm outbreaks computed by \texttt{dfoliatR} and OUTBREAK.
+Input parameters were identical between programs.
+Difference arrise because \texttt{dfoliatR} will identify and record more defoliation events on idnividual trees, and may more accurately capture herbivory dynamics.}
+\end{figure}
 
@@
 In every one of the 102 cases we inspected, we felt that \texttt{dfoliatR} provided a more biologically and statistically appropriate assessment of defoliation.
-Therefore, \texttt{dfoliatR} also did a better job inferring OUTBREAK events and associated statistics at the site level.
+Therefore, \texttt{dfoliatR} also did a better job inferring outbreak events and associated statistics at the site level.
 
@@
 
-The \texttt{dfoliatR} library (Guiterman, Lynch, and Axelson 2020) is provided free and open source from the Comprehensive R Archive Network (CRAN; \url{https://cran.r-project.org/}).
-To install \texttt{dfoliatR} from CRAN use
+The \texttt{dfoliatR} library is provided free and open source from the Comprehensive R Archive Network (CRAN; \url{https://cran.r-project.org/}). To install \texttt{dfoliatR} from CRAN use
 
@@
 Each data set consists of individual host-tree series and a local non-host chronology.
-The host-tree series were standardized using 50-year splines with a 50\% frequency response, while the non-host ring-width data were standardized using 150-year splines with a 50\% frequency response and then averaged via Tukey's biweight robust mean procedure.
-Host trees from Demijohn Peak (DMJ; 2902 m asl) in the San Juan Mountains of southern Colorado include Douglas-fir (\emph{Pseudotsuga menziesii}) compared against a local non-host ponderosa pine (\emph{Pinus ponderosa}) chronology (Ryerson, Swetnam, and Lynch 2003).
-The East Fork site (EFK; 2580 m asl) in the Jemez Mountains of northcentral New Mexico includes Douglas-fir and white fir (\emph{Abies concolor}) host trees and a ponderosa pine non-host chronology (Swetnam and Lynch 1993).
+The host-tree series were standardized using 128-year splines with a 50\% frequency response, while the non-host ring-width data were standardized using 150-year splines with a 50\% frequency response and then averaged via Tukey's biweight robust mean procedure.
+Host trees from Demijohn Peak (DMJ; 2902 m asl) in the San Juan Mountains of southern Colorado include Douglas-fir (\emph{Pseudotsuga menziesii}) compared against a local non-host ponderosa pine (\emph{Pinus ponderosa}) chronology (Ryerson et al., 2003).
+The East Fork site (EFK; 2580 m asl) in the Jemez Mountains of northcentral New Mexico includes Douglas-fir and white fir (\emph{Abies concolor}) host trees and a ponderosa pine non-host chronology (Swetnam and Lynch, 1993).
 
@@
 \end{Shaded}
+
+\begin{table}[!h]
+
+\caption{\label{tab:tbl-defol}Summary statistics for individual tree defoliation events.}
+\centering
+\begin{tabular}[t]{lrrrrrr}
+\toprule
+series & first & last & years & num\_events & tot\_years & mean\_duration\\
+\midrule
+DMJ01 & 1810 & 1996 & 187 & 4 & 50 & 12\\
+DMJ02 & 1750 & 1996 & 247 & 6 & 75 & 12\\
+DMJ03 & 1830 & 1996 & 167 & 4 & 42 & 10\\
+DMJ04 & 1720 & 1996 & 277 & 8 & 90 & 11\\
+DMJ06 & 1700 & 1996 & 297 & 6 & 81 & 14\\
+\addlinespace
+DMJ07 & 1710 & 1996 & 287 & 7 & 96 & 14\\
+DMJ11 & 1900 & 1997 & 98 & 2 & 19 & 10\\
+DMJ14 & 1675 & 1996 & 322 & 9 & 112 & 12\\
+DMJ15 & 1730 & 1996 & 267 & 4 & 64 & 16\\
+DMJ16 & 1746 & 1996 & 251 & 7 & 86 & 12\\
+\addlinespace
+DMJ17 & 1733 & 1996 & 264 & 6 & 85 & 14\\
+DMJ22 & 1720 & 1996 & 277 & 5 & 56 & 11\\
+DMJ23 & 1675 & 1997 & 323 & 8 & 96 & 12\\
+DMJ24 & 1895 & 1996 & 102 & 3 & 28 & 9\\
+DMJ25 & 1680 & 1996 & 317 & 5 & 70 & 14\\
+\addlinespace
+DMJ26 & 1700 & 1996 & 297 & 6 & 79 & 13\\
+DMJ27 & 1710 & 1996 & 287 & 3 & 34 & 11\\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 
-The output table from \texttt{outbreak\_stats()} shows the start and end years of each event, along with their corresponding duration, the number and percent of trees in defoliation at the start of the event, the the number of trees recording the outbreak event (in part or full), the minimums of the mean GSI and NGSI indices, and two points of ``peak defoliation.'' The first, ``peak\_outbreak\_year'' is the year in which the greatest number of trees are recording the outbreak, and the second, ``peak\_defol\_year'' is the year with the lowest average NGSI.
-These output variables provide a array of options for assessing metrics of the insect defoliation regime, including interval analyses:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{## calculate mean return interval for outbreak start years}
-\NormalTok{intervals <-}\StringTok{ }\KeywordTok{diff}\NormalTok{(dmj_obr_stats}\OperatorTok{$}\NormalTok{start)}
-\KeywordTok{mean}\NormalTok{(intervals) }\CommentTok{## 51.2 year mean return interval}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{conclusions}{%
-\section{Conclusions}\label{conclusions}}
-
-The \texttt{dfoliatR} package provides dendroecologists with tools to infer, quantify, analyze, and visualize tree-ring growth suppression events and to reconstruct forest insect defoliator outbreak regimes.
-It is built on the long-accepted host to nonhost comparison methodology used in the 1980s FORTRAN program OUTBREAK (Swetnam, Thompson, and Sutherland 1985; Swetnam and Lynch 1989, 1993).
-Key benefits are that \texttt{dfoliatR} provides easier control of suppression thresholds, provides additional output tables and high-quality graphics, defines defoliation events more consistently, and operates in the open-source \texttt{R} environment.
-
-Using \texttt{dfoliatR} requires standardized ring-width measurements from insect host trees and either an indexed tree-ring chronology from local non-host trees or suitable climate chronology.
-It performs an indexing procedure to remove the climatic signal represented in the non-host chronology from the host-tree series.
-It then infers defoliation events in individual trees.
-Site-level analyses identify outbreak events that synchronously affect a user-defined number or proportion of the host trees.
-Functions are provided for summary statistics and graphics of tree- and site-level series.
-The package produces publication-quality plots, and tabulates growth suppression indices and tree- and site-level outbreak event statistics for user-defined post-processing needs, including those suitable for charting and tabulating landscape- and regional-level results.
-
-\texttt{dfoliatR} adds a new option for dendroentimology to combine, or ``bridge,'' sequential tree-level defoliation events into single events.
-In practice, we suggest that researchers carefully evaluate if bridging is ecologically applicable to study situations and insect ecology, and to carefully explore the data before deciding whether or not to use this option.
-It is probably not appropriate for insects with high-frequency, high--severity outbreaks, such as processionary caterpillars.
-Outbreak reconstructions of insects for which the interval is notably longer than typical outbreak duration, such as Douglas-fir tussock moth (\emph{Orgyia pseudotsugata} (Lepidoptera: Erebidae)), are unlikely to be significantly affected.
-Species for which outbreak duration varies considerably, for which individual tree resilience to defoliation varies considerably, or which may chronically infest trees or sites, or alternate between chronic, outbreak, and minimal activity states, such as conifer-feeding \emph{Choristoneura}, present more complicated challenges, and the researcher should use discretion with the bridge option.
-
-\texttt{dfoliatR} adds to the on-going open-source software development for dendrochronological methods (e.g., Bunn 2008; Brewer 2014; Brewer and Guiterman 2016).
-The \texttt{R} environment enables automation of analyses, allowing input/output processes to become routine, and empowering batch processing of large multi-site projects and sensitivity analyses.
-It also facilitates additional statistical analyses, such as spectral analyses and superposed epoch analyses (e.g., Malevich, Guiterman, and Margolis 2018), with easy transfer from \texttt{dfoliatR} and \texttt{dplR} to other libraries in \texttt{R}.
-Source code for \texttt{dfoliatR} is available in the Comprehensive R Archive Network (CRAN) and GitHub \url{https://github.com/chguiterman/dfoliatR} with updated descriptions and helpful vignettes on the package website \url{https://chguiterman.github.io/dfoliatR/}.
-Researchers wishing to contribute to further development of \texttt{dfoliatR} are encouraged to do so via the GitHub repository.
-
-\hypertarget{acknowledgments}{%
-\section*{Acknowledgments}\label{acknowledgments}}
-\addcontentsline{toc}{section}{Acknowledgments}
-
-We are grateful to late Richard Holmes of the Laboratory of Tree-Ring Research (LTRR) for his achievements in developing OUTBREAK and many other tree-ring software programs, and we thank Thomas Swetnam for his insights and encouragement.
-The \texttt{dfoliatR} project was supported by the Agnes Haury Visiting Scholars Fellowship at the LTRR, the University of California Berkeley, and the Western Wildlands Environmental Threat Center (WWETAC).
-Additional support was provided by the LTRR and the United States Forest Service Rocky Mountain Research Station.
-Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S.
-Government.
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Altman2014}{}%
-Altman, J., P.
-Fibich, J.
-Dolezal, and T.
-Aakala.
-2014.
-``TRADER: A package for Tree Ring Analysis of Disturbance Events in R.'' \emph{Dendrochronologia} 32: 107--12.
-\url{https://doi.org/10.1016/j.dendro.2014.01.004}.
-
-\leavevmode\hypertarget{ref-Blais1954}{}%
-Blais, JR. 1954. ``The Recurrence of Spruce Budworm Infestations in the Past Century in the Lac Seul Area of Northwestern Ontario.'' \emph{Ecology} 35 (1): 62--71.
-
-\leavevmode\hypertarget{ref-Brewer2014}{}%
-Brewer, Peter W. 2014. ``Data Management in Dendroarchaeology Using Tellervo.'' \emph{Radiocarbon} 56 (4): S79--S83. \url{https://doi.org/10.2458/azu_rc.56.18320}.
-
-\leavevmode\hypertarget{ref-Brewer2016}{}%
-Brewer, Peter W., and Christopher H.
-Guiterman.
-2016.
-``A New Digital Field Data Collection System for Dendrochronology.'' \emph{Dendrochronologia} 38: 131--35.
-\url{https://doi.org/10.1016/j.dendro.2016.04.005}.
-
-\leavevmode\hypertarget{ref-Brookes1987}{}%
-Brookes, M.
-H., Robert W.
-Campbell, J.
-J.
-Colbert, R.
-G.
-Mitchell, and R.
-W.
-Stark (Technical Coordinators).
-1982.
-``Western Spruce Budworm.'' USDA Forest Service, Cooperative State Research Service, Technical Bulletin No.
-1694.
-198 p., Washington D.C.
-
-\leavevmode\hypertarget{ref-Bunn2008}{}%
-Bunn, Andrew G. 2008. ``A dendrochronology program library in R (dplR).'' \emph{Dendrochronologia} 26: 115--24. \url{https://doi.org/10.1016/j.dendro.2008.01.002}.
-
-\leavevmode\hypertarget{ref-Bunn2010}{}%
----------. 2010. ``Statistical and visual crossdating in R using the dplR library.'' \emph{Dendrochronologia} 28: 251--58. \url{https://doi.org/10.1016/j.dendro.2009.12.001}.
-
-\leavevmode\hypertarget{ref-Campelo2016}{}%
-Campelo, Filipe, Cristina Nabais, Ana Carvalho, and Joana Vieira.
-2016.
-``tracheideR--An R package to standardize tracheidograms.'' \emph{Dendrochronologia} 37: 64--68.
-\url{https://doi.org/10.1016/j.dendro.2015.12.006}.
-
-\leavevmode\hypertarget{ref-Carus2004}{}%
-Carus, Serdar.
-2004.
-``Impact of defoliation by the pine processionary moth (\emph{Thaumetopoea pityocampa}) on radial, height and volume growth of Calabrian pine (\emph{Pinus brutia}) trees in Turkey.'' \emph{Phytoparasitica} 32 (5): 459--69.
-
-\leavevmode\hypertarget{ref-Carus2009}{}%
----------. 2009. ``Effects of defoliation caused by the processionary moth on growth of Crimean pines in western Turkey.'' \emph{Phytoparasitica} 37 (2): 105--14.
-
-\leavevmode\hypertarget{ref-arstan}{}%
-Cook, Edward R, and Richard L Holmes. 1996. ``Guide for computer program ARSTAN.'' \emph{The International Tree-Ring Data Bank Program Library Version 2}, 75--87.
-
-\leavevmode\hypertarget{ref-Cook2004}{}%
-Cook, Edward R., and Paul J.
-Krusic.
-2004.
-\emph{The North American Drought Atlas}.
-(Lamont- Doherty Earth Observatory; the National Science Foundation: New York) Available at https://www.ncdc.noaa.gov/data-access/paleoclimatology-data/datasets/tree-ring/drought-variability {[}Last accessed May 2020{]}.
-
-\leavevmode\hypertarget{ref-dfoliatR}{}%
-Guiterman, Christopher, Ann Lynch, and Jodi Axelson.
-2020.
-\emph{DfoliatR: Detection and Analysis of Insect Defoliation Signals in Tree Rings}.
-R package version 0.1.0.
-\url{https://doi.org/10.5281/zenodo.3698163}.
-
-\leavevmode\hypertarget{ref-outbreak}{}%
-Holmes, Richard L, and Thomas W Swetnam. 1986. ``Dendroecology program library: program OUTBREAK user's manual.'' \emph{Laboratory of Tree-Ring Research, University of Arizona, Tucson}.
-
-\leavevmode\hypertarget{ref-Jevsenak2018}{}%
-Jevšenak, Jernej, and Tom Levanič.
-2018.
-``dendroTools: R package for studying linear and nonlinear responses between tree-rings and daily environmental data.'' \emph{Dendrochronologia} 48: 32--39.
-\url{https://doi.org/10.1016/j.dendro.2018.01.005}.
-
-\leavevmode\hypertarget{ref-Lara2015}{}%
-Lara, W., F.
-Bravo, and C.
-A.
-Sierra.
-2015.
-``MeasuRing: An R package to measure tree-ring widths from scanned images.'' \emph{Dendrochronologia} 34: 43--50.
-\url{https://doi.org/10.1016/j.dendro.2015.04.002}.
-
-\leavevmode\hypertarget{ref-Lynch2012}{}%
-Lynch, Ann M.
-2012.
-``What Tree-Ring Reconstruction Tells Us about Conifer Defoliator Outbreaks.'' In \emph{Insect Outbreaks Revisited}, edited by Pedro Barbosa, Deborah K.
-Letourneau, and Anurag A.
-Agrawal, 126--54.
-Blackwell Publishing Ltd.
-\url{https://doi.org/10.1002/9781118295205.ch7}.
-
-\leavevmode\hypertarget{ref-Malevich2018}{}%
-Malevich, Steven B., Christopher H.
-Guiterman, and Ellis Q.
-Margolis.
-2018.
-``burnr: Fire history analysis and graphics in R.'' \emph{Dendrochronologia} 49: 9--15.
-\url{https://doi.org/10.1016/j.dendro.2018.02.005}.
-
-\leavevmode\hypertarget{ref-Rathgeber2011}{}%
-Rathgeber, Cyrille B.
-K., Fleur Longuetaud, Frédéric Mothe, Henri Cuny, and Gilles Le Moguédec.
-2011.
-``Phenology of wood formation: Data processing, analysis and visualisation using R (package CAVIAR).'' \emph{Dendrochronologia} 29 (3): 139--49.
-\url{https://doi.org/10.1016/j.dendro.2011.01.004}.
-
-\leavevmode\hypertarget{ref-RCore}{}%
-R Core Team. 2019. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.
-
-\leavevmode\hypertarget{ref-Ryerson2003}{}%
-Ryerson, Daniel E, Thomas W Swetnam, and Ann M Lynch.
-2003.
-``A tree-ring reconstruction of western spruce budworm outbreaks in the San Juan Mountains, Colorado, U.S.A.'' \emph{Canadian Journal of Forest Research} 33: 1010--28.
-\url{https://doi.org/10.1139/X03-026}.
-
-\leavevmode\hypertarget{ref-Sanders1985}{}%
-Sanders, C.
-J., R.
-W.
-Stark, E.
-J.
-Mullins, and J.
-Murphy (eds).
-1985.
-``Recent Advances in Spruce Budworms Research.'' In \emph{Proceedings of the CANUSA Spruce Budworms Research Symposium, 16-20 September 1984, Bangor, Maine.
-Canadian Forestry Service, Ottawa, Ontario.
-527p.}
-
-\leavevmode\hypertarget{ref-Schmitt1984}{}%
-Schmitt, D.
-M., D.
-G.
-Grimble, and JL Searcy.
-1984.
-\emph{Managing the spruce budworm in eastern North America}.
-USDA Forest Service, Cooperative State Research Service, Agriculture Handbook No.
-620.
-192p., Washington D.C.
-
-\leavevmode\hypertarget{ref-Shi2019}{}%
-Shi, Jingning, Wei Xiang, Qijing Liu, and Sher Shah.
-2019.
-``MtreeRing: An R Package with Graphical User Interface for Automatic Measurement of Tree Ring Widths Using Image Processing Techniques.'' \emph{Dendrochronologia} 58: 125644.
-\url{https://doi.org/10.1016/j.dendro.2019.125644}.
-
-\leavevmode\hypertarget{ref-Speer2010}{}%
-Speer, James H. 2010. \emph{Fundamentals of Tree-Ring Research}. The University of Arizona Press.
-
-\leavevmode\hypertarget{ref-Swetnam1989}{}%
-Swetnam, Thomas W., and Ann M.
-Lynch.
-1989.
-``A tree-ring reconstruction of western spruce budworm history in the southern Rocky Mountains.'' \emph{Forest Science} 35 (4): 962--86.
-\url{https://doi.org/10.1093/forestscience/35.4.962}.
-
-\leavevmode\hypertarget{ref-Swetnam1993}{}%
----------. 1993. ``Multicentury, Regional-Scale Patterns of Western Spruce Budworm Outbreaks.'' \emph{Ecological Monographs} 63 (4): 399--424. \url{https://doi.org/10.2307/2937153}.
-
-\leavevmode\hypertarget{ref-Swetnam1985}{}%
-Swetnam, Thomas W., Marna A.
-Thompson, and Elaine K.
-Sutherland.
-1985.
-\emph{Using dendrochronology to measure radial growth of defoliated trees}.
-Vol.
-639.
-United States Department of Agriculture handbook No.
-639.
-38 p., Washington D.C.
-\url{www.fs.usda.gov/treesearch/pubs/40898}.
-
-\leavevmode\hypertarget{ref-Swetnam1999}{}%
-Swetnam, TW, CD Allen, and JL Betancourt.
-1999.
-``Applied historical ecology: using the past to manage for the future.'' \emph{Ecological Applications} 9 (4): 1189--1206.
-\url{https://doi.org/10.1890/1051-0761(1999)009\%5B1189:AHEUTP\%5D2.0.CO;2}.
-
-\leavevmode\hypertarget{ref-VanderMaaten-Theunissen2015}{}%
-van der Maaten-Theunissen, Marieke, Ernst van der Maaten, and Olivier Bouriaud.
-2015.
-``PointRes: An R package to analyze pointer years and components of resilience.'' \emph{Dendrochronologia} 35: 34--38.
-\url{https://doi.org/10.1016/j.dendro.2015.05.006}.
-
-\leavevmode\hypertarget{ref-wickham2016ggplot2}{}%
-Wickham, Hadley. 2016. \emph{ggplot2: Elegant Graphics for Data Analysis}. Second edition. Springer-Verlag, New York. \url{https://ggplot2-book.org/}.
-
-\leavevmode\hypertarget{ref-Wickham2019}{}%
-Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy McGowan, Romain François, Garrett Grolemund, et al.
-2019.
-``Welcome to the Tidyverse.'' \emph{Journal of Open Source Software} 4 (43): 1686.
-\url{https://doi.org/10.21105/joss.01686}.
-
-\leavevmode\hypertarget{ref-Wickham2020dplyr}{}%
-Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2020. \emph{dplyr: A Grammar of Data Manipulation}. R package version 0.8.4. \url{https://CRAN.R-project.org/package=dplyr}.
-
-\leavevmode\hypertarget{ref-Zang2015}{}%
-Zang, Christian, and Franco Biondi.
-2015.
-``Treeclim: An R package for the numerical calibration of proxy-climate relationships.'' \emph{Ecography} 38 (4): 431--36.
-\url{https://doi.org/10.1111/ecog.01335}.
-
-\newpage
-
-\hypertarget{tables-figures}{%
-\section*{Tables \& Figures}\label{tables-figures}}
-\addcontentsline{toc}{section}{Tables \& Figures}
-
-\begin{table}[!h]
-
-\caption{\label{tab:tbl-defol}Summary statistics for individual tree defoliation events.}
-\centering
-\begin{tabular}[t]{lrrrrrr}
-\toprule
-series & first & last & years & num\_events & tot\_years & mean\_duration\\
-\midrule
-DMJ01 & 1810 & 1996 & 187 & 4 & 50 & 12\\
-DMJ02 & 1750 & 1996 & 247 & 6 & 75 & 12\\
-DMJ03 & 1830 & 1996 & 167 & 4 & 42 & 10\\
-DMJ04 & 1720 & 1996 & 277 & 8 & 90 & 11\\
-DMJ06 & 1700 & 1996 & 297 & 6 & 81 & 14\\
-\addlinespace
-DMJ07 & 1710 & 1996 & 287 & 7 & 96 & 14\\
-DMJ11 & 1900 & 1997 & 98 & 2 & 19 & 10\\
-DMJ14 & 1675 & 1996 & 322 & 9 & 112 & 12\\
-DMJ15 & 1730 & 1996 & 267 & 4 & 64 & 16\\
-DMJ16 & 1746 & 1996 & 251 & 7 & 86 & 12\\
-\addlinespace
-DMJ17 & 1733 & 1996 & 264 & 6 & 85 & 14\\
-DMJ22 & 1720 & 1996 & 277 & 5 & 56 & 11\\
-DMJ23 & 1675 & 1997 & 323 & 8 & 96 & 12\\
-DMJ24 & 1895 & 1996 & 102 & 3 & 28 & 9\\
-DMJ25 & 1680 & 1996 & 317 & 5 & 70 & 14\\
-\addlinespace
-DMJ26 & 1700 & 1996 & 297 & 6 & 79 & 13\\
-DMJ27 & 1710 & 1996 & 287 & 3 & 34 & 11\\
-\bottomrule
-\end{tabular}
-\end{table}
-
 \begin{landscape}\begin{table}
@@
 
-\begin{figure}
-\centering
-\includegraphics{../paper_elsevier/Output/tree-plot-default.pdf}
-\caption{\label{fig:fig-defol}Default dfoliatR graphics for individual trees, produced by the `plot\_defol()' function.
-Dotted lines represent the series length for each tree and colored segments show periods of defoliation.
-Users can define the Cut-off values to determine severe-moderate-minor defoliation intensities, and use standard ggplot2 graphics parameters to adjust styles and themes.}
-\end{figure}
+The output table from \texttt{outbreak\_stats()} shows the start and end years of each event, along with their corresponding duration, the number and percent of trees in defoliation at the start of the event, the the number of trees recording the outbreak event (in part or full), the minimums of the mean GSI and NGSI indices, and two points of ``peak defoliation.'' The first, ``peak\_outbreak\_year'' is the year in which the greatest number of trees are recording the outbreak, and the second, ``peak\_defol\_year'' is the year with the lowest average NGSI.
+These output variables provide a array of options for assessing metrics of the insect defoliation regime, including interval analyses:
 
-\begin{figure}
-\centering
-\includegraphics{../paper_elsevier/Output/site-plot.pdf}
-\caption{\label{fig:fig-obr}Default graphic to show outbreak events at the DMJ site.
-Top panel shows the mean GSI through time, with inferred outbreak events filled.
-The middle panel shows the percent of trees defoliated, and the bottom panel shows the number of trees through time.}
-\end{figure}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{## calculate mean return interval for outbreak start years}
+\NormalTok{intervals <-}\StringTok{ }\KeywordTok{diff}\NormalTok{(dmj_obr_stats}\OperatorTok{$}\NormalTok{start)}
+\KeywordTok{mean}\NormalTok{(intervals) }\CommentTok{## 51.2 year mean return interval}
+\end{Highlighting}
+\end{Shaded}
 
-\begin{figure}
-\centering
-\includegraphics{../paper_elsevier/Output/plot-trees-df-obr.pdf}
-\caption{\label{fig:df-obr-1-1}Comparison of disturbance indices between dfoliatR and OUTBREAK. Diagonal lines show equal values (1:1).}
-\end{figure}
+\hypertarget{conclusions}{%
+\section{Conclusions}\label{conclusions}}
 
-\begin{figure}
-\centering
-\includegraphics{../paper_elsevier/Output/plot-sites-df-obr.pdf}
-\caption{\label{fig:df-obr-site}Comparison of reconstructed western spruce budworm outbreaks computed by dfoliatR and OUTBREAK. Input parameters were identical between programs.}
-\end{figure}
+The \texttt{dfoliatR} package provides dendroecologists with tools to infer, quantify, analyze, and visualize tree-ring growth suppression events and to reconstruct forest insect defoliator outbreak regimes.
+It is built on the long-accepted host to nonhost comparison methodology used in the 1980s FORTRAN program OUTBREAK (Swetnam et al., 1985; Swetnam and Lynch, 1989, 1993).
+Our evaluation of the two programs revelaed that \texttt{dfolaitR} excelled in identifying defoliation events on single trees, providing researchers with more consistent and biologically-justifiable results.
+Other key benefits are that \texttt{dfoliatR} provides easier control of suppression thresholds, additional output tables, high-quality and customizable graphics, and operates in the open-source \texttt{R} environment that is stable across computing platforms and under active development and maintence by a large and growing community.
+
+Using \texttt{dfoliatR} requires standardized ring-width measurements from insect host trees and either an indexed tree-ring chronology from local non-host trees or suitable climate chronology.
+It performs an indexing procedure to remove the climatic signal represented in the non-host chronology from the host-tree series.
+It then infers defoliation events in individual trees.
+Site-level analyses identify outbreak events that synchronously affect a user-defined number or proportion of the host trees.
+Functions are provided for summary statistics and graphics of tree- and site-level series.
+The package produces publication-quality plots, and tabulates growth suppression indices and tree- and site-level outbreak event statistics for user-defined post-processing needs, including those suitable for charting and tabulating landscape- and regional-level results.
+
+\texttt{dfoliatR} adds a new option for dendroentimology to combine, or ``bridge,'' sequential tree-level defoliation events into single events.
+In practice, we suggest that researchers carefully evaluate if bridging is ecologically applicable to study situations and insect ecology, and to carefully explore the data before deciding whether or not to use this option.
+It is probably not appropriate for insects with high-frequency, high--severity outbreaks, such as processionary caterpillars.
+Outbreak reconstructions of insects for which the interval is notably longer than typical outbreak duration, such as Douglas-fir tussock moth (\emph{Orgyia pseudotsugata} (Lepidoptera: Erebidae)), are unlikely to be significantly affected.
+Species for which outbreak duration or individual tree resilience to defoliation varies considerably, or which may chronically infest trees or sites, or alternate between chronic, outbreak, and minimal activity states such as conifer-feeding \emph{Choristoneura}, present more complicated challenges, and the researcher should use discretion with the bridge option.
+
+\texttt{dfoliatR} adds to the on-going open-source software development for dendrochronological methods (e.g., Bunn, 2008; Brewer, 2014; Brewer and Guiterman, 2016).
+The \texttt{R} environment enables automation of analyses, allowing input/output processes to become routine, and empowering batch processing of large multi-site projects and sensitivity analyses.
+It also facilitates additional statistical analyses, such as spectral analyses and superposed epoch analyses (e.g., Malevich et al., 2018), with easy transfer from \texttt{dfoliatR} and \texttt{dplR} to other libraries in \texttt{R}.
+Source code for \texttt{dfoliatR} is available in the Comprehensive R Archive Network (CRAN) and GitHub \url{https://github.com/chguiterman/dfoliatR} with updated descriptions and helpful vignettes on the package website \url{https://chguiterman.github.io/dfoliatR/}.
+Researchers wishing to contribute to the further development of \texttt{dfoliatR} are encouraged to do so via the GitHub repository.
+
+\hypertarget{acknowledgments}{%
+\section*{Acknowledgments}\label{acknowledgments}}
+\addcontentsline{toc}{section}{Acknowledgments}
+
+We are grateful to late Richard Holmes of the Laboratory of Tree-Ring Research (LTRR) for his achievements in developing OUTBREAK and many other tree-ring software programs.
+We also thank Thomas Swetnam for his insights and encouragement, and David Frank for his support.
+The \texttt{dfoliatR} project is funded by the Agnes Haury Visiting Scholars Fellowship at the LTRR, the University of California, Berkeley, and the Western Wildlands Environmental Threat Center (WWETAC).
+Additional support was provided by the LTRR and the United States Forest Service Rocky Mountain Research Station.
+The findings and conclusions in this publication are those of the authors and should not be construed to represent any official USDA or U.S.
+Government determination or policy.
+
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-Altman2014}{}%
+Altman, J., Fibich, P., Dolezal, J., Aakala, T., 2014.
+TRADER: A package for Tree Ring Analysis of Disturbance Events in R.
+Dendrochronologia 32, 107--12, doi:\href{https://doi.org/10.1016/j.dendro.2014.01.004}{10.1016/j.dendro.2014.01.004}.
+
+\leavevmode\hypertarget{ref-Blais1954}{}%
+Blais, J., 1954. The recurrence of spruce budworm infestations in the past century in the lac seul area of northwestern ontario. Ecology 35, 62--71.
+
+\leavevmode\hypertarget{ref-Brewer2014}{}%
+Brewer, P.W., 2014.
+Data management in dendroarchaeology using tellervo.
+Radiocarbon Cambridge University Press 56, S79--S83, doi:\href{https://doi.org/10.2458/azu_rc.56.18320}{10.2458/azu\_rc.56.18320}.
+
+\leavevmode\hypertarget{ref-Brewer2016}{}%
+Brewer, P.W., Guiterman, C.H., 2016.
+A new digital field data collection system for dendrochronology.
+Dendrochronologia 38, 131--35, doi:\href{https://doi.org/10.1016/j.dendro.2016.04.005}{10.1016/j.dendro.2016.04.005}.
+
+\leavevmode\hypertarget{ref-Brookes1987}{}%
+Brookes, M.H., Campbell, R.W., Colbert, J.J., Mitchell, R.G., (Technical Coordinators), R.W.S., 1982.
+Western spruce budworm.
+USDA Forest Service, Cooperative State Research Service, Technical Bulletin No.
+1694.
+198 p., Washington D.C.
+
+\leavevmode\hypertarget{ref-Bunn2008}{}%
+Bunn, A.G., 2008. A dendrochronology program library in R (dplR). Dendrochronologia 26, 115--24, doi:\href{https://doi.org/10.1016/j.dendro.2008.01.002}{10.1016/j.dendro.2008.01.002}.
+
+\leavevmode\hypertarget{ref-Bunn2010}{}%
+Bunn, A.G., 2010. Statistical and visual crossdating in R using the dplR library. Dendrochronologia 28, 251--58, doi:\href{https://doi.org/10.1016/j.dendro.2009.12.001}{10.1016/j.dendro.2009.12.001}.
+
+\leavevmode\hypertarget{ref-Campelo2016}{}%
+Campelo, F., Nabais, C., Carvalho, A., Vieira, J., 2016.
+tracheideR--An R package to standardize tracheidograms.
+Dendrochronologia 37, 64--68, doi:\href{https://doi.org/10.1016/j.dendro.2015.12.006}{10.1016/j.dendro.2015.12.006}.
+
+\leavevmode\hypertarget{ref-Carus2004}{}%
+Carus, S., 2004.
+Impact of defoliation by the pine processionary moth (\emph{Thaumetopoea pityocampa}) on radial, height and volume growth of Calabrian pine (\emph{Pinus brutia}) trees in Turkey.
+Phytoparasitica 32, 459--69.
+
+\leavevmode\hypertarget{ref-Carus2009}{}%
+Carus, S., 2009. Effects of defoliation caused by the processionary moth on growth of Crimean pines in western Turkey. Phytoparasitica 37, 105--14.
+
+\leavevmode\hypertarget{ref-arstan}{}%
+Cook, E.R., Holmes, R.L., 1996. Guide for computer program ARSTAN. The international tree-ring data bank program library version 2 75--87.
+
+\leavevmode\hypertarget{ref-Cook2004}{}%
+Cook, E.R., Krusic, P.J., 2004.
+The North American Drought Atlas.
+(Lamont- Doherty Earth Observatory; the National Science Foundation: New York) Available at https://www.ncdc.noaa.gov/data-access/paleoclimatology-data/datasets/tree-ring/drought-variability {[}Last accessed May 2020{]}.
+
+\leavevmode\hypertarget{ref-dfoliatR}{}%
+Guiterman, C., Lynch, A., Axelson, J., 2020.
+DfoliatR: Detection and analysis of insect defoliation signals in tree rings., \url{https://chguiterman.github.io/dfoliatR/}, doi:\href{https://doi.org/10.5281/zenodo.3698163}{10.5281/zenodo.3698163}.
+
+\leavevmode\hypertarget{ref-outbreak}{}%
+Holmes, R.L., Swetnam, T.W., 1986. Dendroecology program library: program OUTBREAK user's manual. Laboratory of Tree-Ring Research, University of Arizona, Tucson.
+
+\leavevmode\hypertarget{ref-Jevsenak2018}{}%
+Jevšenak, J., Levanič, T., 2018.
+dendroTools: R package for studying linear and nonlinear responses between tree-rings and daily environmental data.
+Dendrochronologia 48, 32--39, doi:\href{https://doi.org/10.1016/j.dendro.2018.01.005}{10.1016/j.dendro.2018.01.005}.
+
+\leavevmode\hypertarget{ref-Lara2015}{}%
+Lara, W., Bravo, F., Sierra, C.A., 2015.
+MeasuRing: An R package to measure tree-ring widths from scanned images.
+Dendrochronologia 34, 43--50, doi:\href{https://doi.org/10.1016/j.dendro.2015.04.002}{10.1016/j.dendro.2015.04.002}.
+
+\leavevmode\hypertarget{ref-Lynch2012}{}%
+Lynch, A.M., 2012.
+What Tree-Ring Reconstruction Tells Us about Conifer Defoliator Outbreaks, in: Barbosa, P., Letourneau, D.K., Agrawal, A.A.
+(Eds.), Insect outbreaks revisited.
+Blackwell Publishing Ltd, pp.
+126--54, doi:\href{https://doi.org/10.1002/9781118295205.ch7}{10.1002/9781118295205.ch7}.
+
+\leavevmode\hypertarget{ref-Malevich2018}{}%
+Malevich, S.B., Guiterman, C.H., Margolis, E.Q., 2018.
+burnr: Fire history analysis and graphics in R.
+Dendrochronologia 49, 9--15, doi:\href{https://doi.org/10.1016/j.dendro.2018.02.005}{10.1016/j.dendro.2018.02.005}.
+
+\leavevmode\hypertarget{ref-Rathgeber2011}{}%
+Rathgeber, C.B.K., Longuetaud, F., Mothe, F., Cuny, H., Le Moguédec, G., 2011.
+Phenology of wood formation: Data processing, analysis and visualisation using R (package CAVIAR).
+Dendrochronologia 29, 139--49, doi:\href{https://doi.org/10.1016/j.dendro.2011.01.004}{10.1016/j.dendro.2011.01.004}.
+
+\leavevmode\hypertarget{ref-RCore}{}%
+R Core Team, 2019. R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria, \url{https://www.R-project.org/}.
+
+\leavevmode\hypertarget{ref-Ryerson2003}{}%
+Ryerson, D.E., Swetnam, T.W., Lynch, A.M., 2003.
+A tree-ring reconstruction of western spruce budworm outbreaks in the San Juan Mountains, Colorado, U.S.A.
+Canadian Journal of Forest Research 33, 1010--28, doi:\href{https://doi.org/10.1139/X03-026}{10.1139/X03-026}.
+
+\leavevmode\hypertarget{ref-Sanders1985}{}%
+Sanders, C.J., Stark, R.W., Mullins, E.J., (eds), J.M., 1985.
+Recent advances in spruce budworms research.
+in: Proceedings of the CANUSA spruce budworms research symposium, 16-20 september 1984, bangor, maine.
+Canadian forestry service, ottawa, ontario.
+527p..
+
+\leavevmode\hypertarget{ref-Schmitt1984}{}%
+Schmitt, D.M., Grimble, D.G., Searcy, J., 1984.
+Managing the spruce budworm in eastern North America.
+USDA Forest Service, Cooperative State Research Service, Agriculture Handbook No.
+620.
+192p., Washington D.C.
+
+\leavevmode\hypertarget{ref-Shi2019}{}%
+Shi, J., Xiang, W., Liu, Q., Shah, S., 2019.
+MtreeRing: An r package with graphical user interface for automatic measurement of tree ring widths using image processing techniques.
+Dendrochronologia 58, 125644, doi:\href{https://doi.org/10.1016/j.dendro.2019.125644}{10.1016/j.dendro.2019.125644}.
+
+\leavevmode\hypertarget{ref-Speer2010}{}%
+Speer, J.H., 2010. Fundamentals of Tree-Ring Research. The University of Arizona Press, 333 pp.
+
+\leavevmode\hypertarget{ref-Swetnam1999}{}%
+Swetnam, T., Allen, C., Betancourt, J., 1999.
+Applied historical ecology: using the past to manage for the future.
+Ecological applications 9, 1189--1206, doi:\href{https://doi.org/10.1890/1051-0761(1999)009\%5B1189:AHEUTP\%5D2.0.CO;2}{10.1890/1051-0761(1999)009{[}1189:AHEUTP{]}2.0.CO;2}.
+
+\leavevmode\hypertarget{ref-Swetnam1989}{}%
+Swetnam, T.W., Lynch, A.M., 1989.
+A tree-ring reconstruction of western spruce budworm history in the southern Rocky Mountains.
+Forest Science 35, 962--86, doi:\href{https://doi.org/10.1093/forestscience/35.4.962}{10.1093/forestscience/35.4.962}.
+
+\leavevmode\hypertarget{ref-Swetnam1993}{}%
+Swetnam, T.W., Lynch, A.M., 1993.
+Multicentury, Regional-Scale Patterns of Western Spruce Budworm Outbreaks.
+Ecological Monographs 63, 399--424, doi:\href{https://doi.org/10.2307/2937153}{10.2307/2937153}.
+
+\leavevmode\hypertarget{ref-Swetnam1985}{}%
+Swetnam, T.W., Thompson, M.A., Sutherland, E.K., 1985.
+Using dendrochronology to measure radial growth of defoliated trees.
+United States Department of Agriculture handbook No.
+639.
+38 p., Washington D.C, 38p.
+pp, \url{www.fs.usda.gov/treesearch/pubs/40898}.
+
+\leavevmode\hypertarget{ref-VanderMaaten-Theunissen2015}{}%
+van der Maaten-Theunissen, M., van der Maaten, E., Bouriaud, O., 2015.
+PointRes: An R package to analyze pointer years and components of resilience.
+Dendrochronologia 35, 34--38, \url{http://dx.doi.org/10.1016/j.dendro.2015.05.006}, doi:\href{https://doi.org/10.1016/j.dendro.2015.05.006}{10.1016/j.dendro.2015.05.006}.
+
+\leavevmode\hypertarget{ref-wickham2016ggplot2}{}%
+Wickham, H., 2016. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag, New York, \url{https://ggplot2-book.org/}.
+
+\leavevmode\hypertarget{ref-Wickham2019}{}%
+Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L., François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M., Pedersen, T., Miller, E., Bache, S., Müller, K., Ooms, J., Robinson, D., Seidel, D., Spinu, V., Takahashi, K., Vaughan, D., Wilke, C., Woo, K., Yutani, H., 2019.
+Welcome to the tidyverse.
+Journal of Open Source Software The Open Journal 4, 1686, \url{https://doi.org/10.21105/joss.01686}, doi:\href{https://doi.org/10.21105/joss.01686}{10.21105/joss.01686}.
+
+\leavevmode\hypertarget{ref-Wickham2020dplyr}{}%
+Wickham, H., François, R., Henry, L., Müller, K., 2020. dplyr: A Grammar of Data Manipulation., \url{https://CRAN.R-project.org/package=dplyr}.
+
+\leavevmode\hypertarget{ref-Zang2015}{}%
+Zang, C., Biondi, F., 2015.
+Treeclim: An R package for the numerical calibration of proxy-climate relationships.
+Ecography 38, 431--36, doi:\href{https://doi.org/10.1111/ecog.01335}{10.1111/ecog.01335}.
 

### 2020-05-11 paper_elsevier/R/bridging.R

@@ -1,0 +1,96 @@
+# Show bridging
+## CHG 2020-05-07
+
+library(dfoliatR)
+library(dplR)
+library(dplyr)
+library(purrr)
+library(tibble)
+library(tidyr)
+library(ggplot2)
+library(prismatic)
+library(paletteer)
+library(here)
+
+## Master metadata file
+meta_in <- read.csv(here("paper_elsevier", "Data", "Meta.csv"))
+# file of output from the comparison -- has the OBR results
+obr_results <- read.csv(here("paper_elsevier", "Output", "Tree-comps-data.csv")) %>% 
+  select(site, year, series, obr_type)
+
+## Helper function
+rm_sample_depth <- function(x){
+  if (any(names(x) %in% "samp.depth"))
+    df <- select(x, - samp.depth)
+  else df <- x
+  return(df)
+}
+
+df_run <- meta_in %>% 
+  transmute(site,
+            host = map(host_file, ~ read.compact(here("paper_elsevier", "Data", .))),
+            nonhost = map(nonhost_file, ~ read.crn(here("paper_elsevier", "Data", .)) %>% 
+                            rm_sample_depth()),
+            defol = pmap(list(host, nonhost, ongoing_defol),
+                         ~ defoliate_trees(host_tree = ..1, 
+                                           nonhost_chron = ..2,
+                                           series_end_event = ..3,
+                                           duration_years = 8,
+                                           max_reduction = -1.28,
+                                           bridge_events = TRUE,  ## Add bridges
+                                           list_output = FALSE))
+  )
+
+df_defol <- df_run %>% 
+  select(site, defol) %>% 
+  unnest(cols = "defol") %>% 
+  inner_join(obr_results)
+
+
+# Selected bridge example -------------------------------------------------
+
+fnl04 <- df_defol %>% 
+  filter(series == "FNL04",
+         year %in% 1881:1925)
+efk22 <- df_defol %>% 
+  filter(series == "EFK22",
+         year %in% 1940:1975)
+dmj26 <- df_defol %>% 
+  filter(series == "DMJ26",
+        year %in% 1850:1875)
+
+comps <- rbind(fnl04, efk22, dmj26)
+
+obr_events <- comps %>% 
+  filter(obr_type != "nd")
+df_events <- comps %>% 
+  filter(defol_status != "nd")
+
+cols <- dichromat::colorschemes$SteppedSequential.5[c(3, 8, 18)]
+
+ggplot(comps, aes(x=year)) +
[…]

### 2020-05-14 paper_elsevier/paper_elsevier.Rmd
merging Jodi and Ann's reviews

@@
 ---
-title: "*dfoliatR*: An R package for detection and analysis of insect defoliation signals in tree rings"
+title: "`dfoliatR`: An R package for detection and analysis of insect defoliation signals in tree rings"
 author:
@@
     email: jodi.axelson@berkeley.edu
-    affiliation: c
+    affiliation: d
     # footnote: 4
@@
 keywords: |
-  Dendroecology, spruce budworm, *Choristoneura*, pandora moth, *Coloradia pandora Blake*, larch-bud-moth
+  Dendroecology, dendroentomology, OUTBREAK, western spruce budworm, defoliator regimes, insect outbreak reconstruction
   <!-- \newpage -->
@@
 # output: rticles::elsevier_article
-# output:
-#   bookdown::pdf_book:
-#       base_format: rticles::elsevier_article
-bookdown::word_document2:
-    reference_docx: "../common/word-styles-reference-01.docx"
-    toc: false
-#     toc_depth: 2
-    fig_caption: yes
-    df_print: kable
+output:
+  bookdown::pdf_book:
+      base_format: rticles::elsevier_article
+# bookdown::word_document2:
+#     reference_docx: "../common/word-styles-reference-01.docx"
+#     toc: false
+# #     toc_depth: 2
+#     fig_caption: yes
+#     df_print: kable
 # always_allow_html: true
@@
 Defoliation can be distinguished from climate- and other disturbance-related influences by comparing ring-width or other annually-resolved features in the wood of host species to that of non-host species or annually-resolved climate records.
-The effect of defoliation on radial growth of trees has been recognized since the 1860s, and used to reconstruct outbreak regimes since the 1950s [@Blais1954; @Lynch2012].
+The effect of defoliation on radial growth of trees has been recognized since the 1860s, and used to reconstruct outbreak regimes since the 1950s [@Blais1954; @Alfaro1982; @Lynch2012].
 It was not until the 1980s, however, that precise dendrochronological techniques were applied for inferring defoliation events and reconstructing defoliator outbreak regimes [@Swetnam1985; @Speer2010; @Lynch2012].
@@
 It is important that the host-tree data include only one tree-ring series per tree.
-Both `dplR` and ARSTAN have options for averaging multiple sample series into a tree-level series.
+`dplR` (via the `dplR::treeMean()` function) and dpl versions of ARSTAN have options for averaging multiple sample series into a tree-level series.
 
 At the heart of `dfoliatR` lies two functions: `defoliate_trees()` and `outbreak()`.
-These identify defoliation events on individual trees (Figure \@ref(fig:fig-defol)) and then composite across multiple trees to identify outbreak events (Figure \@ref(fig:fig-obr)).
+These identify defoliation events on individual trees (Figure \@ref(fig:fig-defol)) and then composite across multiple trees to infer stand or site level outbreak events (Figure \@ref(fig:fig-obr)).
 
-(ref:defolCap) Default `dfoliatR` graphics for individual trees, produced by the `plot_defol()` function.
-Dotted lines represent the series length for each tree and colored segments show periods of defoliation.
-Users can define the cut-off values to determine severe--moderate--minor defoliation intensities, and use standard `ggplot2` graphics parameters to adjust styles and themes.
+(ref:defolCap) Default `dfoliatR` graphic for individual trees, produced by the `plot_defol()` function.
+The series names are listed on the Y-axis, dotted lines represent the series length for each tree, and colored segments show periods of defoliation.
+The colors of defoliation segments represent its severity, for which users can define cut-off values to determine severe--moderate--minor defoliation intensities.
+The default break points for severity classes are the mean and first quartile for event NGSI values.
+Colors and other features of the graphic can be adjusted using `ggplot2` parameters, as shown below.
 
@@
 (ref:obrCap) Default graphics showing outbreak events at the DMJ site.
-Top panel shows the mean GSI through time, with inferred outbreak events filled.
-The middle panel shows the percent of trees defoliated, and the bottom panel shows the number of trees through time.
+Top panel shows the sample depth, the middle panel shows the mean GSI with inferred outbreak events filled, and the bottom panel shows the percent of trees defoliated, used to identify time periods of inferred outbreaks.
 
@@
 As in OUTBREAK, the lowest NGSI value in the particular sequence being assessed must reach the magnitude threshold.
-The default setting is -1.28 (NGSI is in units of standard deviation), which was previously determined to be representative of WSBW effects (Swetnam and Lynch, 1989) and is commonly used for other species.
+The default setting is -1.28 (NGSI is in units of standard deviation), which was previously determined to be representative of WSBW effects (Swetnam and Lynch, 1989) and is commonly used for other species [see @Lynch2012].
 The year with the lowest value is termed the "year of maximum departure" and becomes a central point in time for assessing other thresholds before being included as a defoliation event.
@@
 
-Event duration is assessed by examining sequences of negative NGSI (that exceed the magnitude threshold) before and after the year of maximum departure year.
+Event duration is assessed by examining sequences of negative NGSI (that exceed the magnitude threshold) before and after the year of maximum departure.
 Each defoliation event is allowed one single-year positive excursion on each side of the year of maximum departure.
@@
 Researchers can, and should, adjust the duration and magnitude parameters accordingly and critically evaluate the results, as insect species vary in the length of their outbreaks and the degree to which they can suppress tree growth.
+OUTBREAK provides two sets of default values, those for WSBW, which typically has lengthy outbreaks, and ones for Douglas-fir tussock moth (*Orgyia pseudotsugata* (Lepidoptera: Tortricidae) of three years duration with -1.28 departure threshold that may be suitable for more eruptive species.
 
@@
 
-Diverging from OUTBREAK, `dfoliatR` allows users to extend defoliation events on individual trees by bridging between sequential events (Figure \@ref(fig:fig-bridge)).
+Diverging from OUTBREAK, `dfoliatR` includes an option allowing users to extend defoliation events on individual trees by bridging between sequential events (Figure \@ref(fig:fig-bridge)).
 In cases where two defoliation events are separated by a single year, bridging will link them into a single event.
-This option was added to `dfoliatR` during the testing phase of development, when we realized that OUTBREAK deliberately omits sequential, or back-to-back events, even when these events surpass both the magnitude and duration thresholds.
+This option was added to `dfoliatR` during the testing phase of development, when we realized that OUTBREAK deliberately omits sequential, or back-to-back events, even when both events surpass the magnitude and duration thresholds.
 Instead, OUTBREAK will select the one sequential event with the lowest negative departure year.
@@
 This is particularly relevant to WSBW and spruce budworm (*C.
-fumiferana*), for which outbreak regimes have been reconstructed for several geographic areas and for which researchers and forest health experts have gained considerable knowledge [@Schmitt1984; @Sanders1985; @Brookes1987, and many later publications].
-In many cases, the greatest growth suppression often occurs late in the outbreak due to cumulative effects accrued on a tree's resources [@Swetnam1989; @Swetnam1993].
-We urge caution in using the bridging option, however, because it may not be appropriate for all studied insects, such as in situations where impacted stands barely recover from one outbreak before another begins, as with pine processionary caterpillars (*Thaumetopoea pityocampa* (Lepidoptera: Thaumetopoeidae)) [@Carus2004; @Carus2009].
+fumiferana*), for which multiple outbreak regime reconstruction, as well as other research and forest health observations, show that outbreaks can be very long [@Schmitt1984; @Sanders1985; @Brookes1987, and many later publications].
+Often the greatest growth suppression occurs late in the outbreak due to lag effects between defoliation and radial growth, and to cumulative effects accrued on a tree's resources [@Brubaker1978; @Alfaro1982; @Mason1997; @Axelson2014; @Swetnam1989; @Wickman1986; @Swetnam1993].
+We urge caution in using the bridging option, however, because it may not be appropriate for all studied insects, such as in situations where impacted stands barely recover from one outbreak before another begins, as with pine processionary caterpillars (*Thaumetopoea pityocampa* (Lepidoptera: Thaumetopoeidae)) [@Carus2004; @Carus2009] or where outbreaks are known to be very short, such as larch budmoth (*Zeiraphera diniana* Gn.) in the European Alps [@Esper2007].
 
-(ref:BridgeCap) Examples of sequential defoliation events (bars), wherein two or three events occur separated by single years.
+(ref:BridgeCap) Examples of identified defoliation events on individual trees.
+For each tree (DMJ26 and EFK22), OUTBREAK and `dfoliatR` identify most of the same events, but there is one added event (blue colored) that was omitted by OUTBREAK.
+These were omitted because they were sequential, separated by a single year with possitive normalized growth suppression index (NGSI).
+`dfoliatR` provides an option to bridge these events into single long events that may better represent the duration of defoliation given the insect and sites under consideration.
+
+Examples of sequential defoliation events (bars), wherein two or three events occur separated by single years.
 In each case OUTBREAK recorded only one of the events, omitting those with a higher "max_defol" year, or the year of the lowest NGSI value.
@@
 ```{r fig-bridge, echo=FALSE, fig.cap='(ref:BridgeCap)'}
-include_graphics(here("paper_elsevier", "Output", "bridging.pdf"))
+include_graphics(here("paper_elsevier", "Output", "bridging_color.pdf"))
 ```
@@
 We tested `dfoliatR` against OUTBREAK by comparing NGSI to OUTBREAK's normalized corrected indices for individual trees and years, defoliation status for individual trees and years, and percentage of trees recording outbreaks at the site level.
-Our tests used standardized ring-width data from eight sites in British Columbia, Colorado, and New Mexico, and author-provided non-host site chronologies.
+Our tests used standardized ring-width data from eight host-tree sites spanning the range of WSBW.
+The sites were sampled in British Columbia [@Axelson2015], Wyoming [@NADEF2018], Colorado, and New Mexico [@Swetnam1993; @Ryerson2003].
+These host data were compared to non-host chronologies from the original studies, but we made no effort here to replicate the reconstructions or analyses of those studies.
 
 We detrended host data for both `dfoliatR` and OUTBREAK using ARSTAN [v6.1] with cubic smoothing splines (50% frequency response on 100-150 year wavelengths depending on the site).
-In both `dfoliatR` and OUTBREAK we used event thresholds of -1.28 normalized indices, 8 years duration, and allowed for events at the end of series in seven of eight sites.
-These seven sites were sampled during ongoing outbreak events [@Swetnam1993].
+In both `dfoliatR` and OUTBREAK we used event thresholds of -1.28 normalized indices, 8 years duration, and allowed for events at the end of series in seven of eight sites that had known outbreaks at the time of sampling.
 We found it necessary to be consistent in how we detrended and what software we employed (e.g., ARSTAN vs `dplR`) because subtle differences in standardized ring-width indices generated between the programs transferred into differences between `dfoliatR` and OUTBREAK.
@@
 
-* *Sequential events* (36%) in which OUTBREAK omitted back-to-back events that occur one year prior or one year following an identified event.
+* *Sequential events* (36%) in which OUTBREAK omitted back-to-back events that occur one year prior to, or one year following an identified event.
 When this occurs, OUTBREAK selects the one event sequence with the lowest negative index year (e.g., Figure \@ref(fig:fig-bridge)).
 On two trees, OUTBREAK omitted two of three sequential events.
-While inspecting these differences, we added an option to `defoliate_trees()` that would "bridge" between sequential events (that surpass magnitude and duration thresholds) into single, long events.
-We felt that this was ecologically justified, especially for studies of WSBW, because the insect will occasionally defoliate individual trees for decades and reconstructed chronologies have show outbreaks as long as **xxx (ANN CHECK)**.
+While inspecting these differences, we added an option to `defoliate_trees()` that would "bridge" between sequential events (that each surpass the magnitude and duration thresholds) into single, long events.
+We felt that this was ecologically justified, especially for studies of WSBW, because outbreaks are known to be of long duration and tree-ring reconstructions have shown that outbreaks may persist 30-50 at the site level.
 
@@
 At the site level, OUTBREAK and `dfoliatR` produce similar time series of percent trees defoliated (Figure \@ref(fig:fig-df-obr)), which forms the basis for inferring outbreak occurrence, intensity, and duration.
-In nearly site comparison, `dfoliatR` included either more events or longer durations of inferred outbreak.
-These differences arise from the inclusion of tree-level events by `dfoliatR` that were omitted by OUTBREAK.
-The differences translate to greater numbers of trees in defoliation during certain events, or events recorded by single trees during periods of low sample depth.
+In nearly all site comparison, `dfoliatR` included either more events or longer durations of inferred outbreaks.
+These differences arise from the inclusion of tree-level events by `dfoliatR` that were omitted by OUTBREAK (see note on sequential events above).
+Thus, in `dfoliatR`, there were a greater number of trees experiencing defoliation during outbreak periods, or outbreaks were represented by a single tree when
+there was low sample depth.
 
@@
 Input parameters were identical between programs.
-Difference arise because `dfoliatR` will identify and record more defoliation events on individual trees, and may more accurately capture herbivory dynamics.
+Differences arise because `dfoliatR` will identify and record more defoliation events on individual trees.
 
@@
 
-This evaluation revealed what we believe are short-comings in how OUTBREAK identifies defoliation events on individual trees.
-In every one of the 102 cases we inspected, we felt that `dfoliatR` provided a more biologically and statistically appropriate assessment of defoliation.
-Therefore, `dfoliatR` also did a better job inferring outbreak events and associated statistics at the site level.
+This comparison revealed what we believe are shortcomings in how OUTBREAK identifies defoliation events on individual trees.
+In every one of the 102 cases we inspected, we felt that `dfoliatR` provided a more biologically and statistically appropriate assessment of defoliation, translating to more robust inferences of outbreak events and associated statistics at the site level.
 
@@
 As an Open Source library, we welcome and encourage community involvement in future development.
-The best ways to contribute to `dfoliatR` are through standard GitHub procedures or by contacting the first author.
+The best ways to contribute to `dfoliatR` are through standard GitHub procedures or by contacting the corresponding author.
 
@@
 Host trees from Demijohn Peak (DMJ; 2902 m asl) in the San Juan Mountains of southern Colorado include Douglas-fir (*Pseudotsuga menziesii*) compared against a local non-host ponderosa pine (*Pinus ponderosa*) chronology [@Ryerson2003].
-The East Fork site (EFK; 2580 m asl) in the Jemez Mountains of northcentral New Mexico includes Douglas-fir and white fir (*Abies concolor*) host trees and a ponderosa pine non-host chronology [@Swetnam1993].
+The East Fork site (EFK; 2580 m asl) in the Jemez Mountains of north-central New Mexico includes Douglas-fir and white fir (*Abies concolor*) host trees and a ponderosa pine non-host chronology [@Swetnam1993].
 
@@
 In our example scripts below, `##` and `#` denote user comments, per standard `R` coding, which are colored in brown.
-Text in blue denote functions; black are loaded objects, and green are quoted variables and links.
+Text in blue denotes functions; black are loaded objects, and green are quoted variables and links.
+Values or other information provided after equal signs are filenames and parameters provided for this example, and in actual use would be replaced with user-specified information.
+In this example "dmj_h" and "dmj_nh" are the individual-tree host series and non-host site chronology files for the Demijohn site, thresholds are set at 8 years and -1.28 standard deviations, bridging is used, series-end events are included in the interval computations, and comprehensive results information is not included in the output.
 
@@
 
+(ref:tbl-defolCap) Tree-level tabular output provided by the `defol_stats()` function for the DMJ example site.
+
 ```{r tbl-defol, echo=FALSE}
 tree_dat <- read.csv(here("paper_elsevier", "Output", "defol-stats.csv"))
-kable(tree_dat, caption = "Summary statistics for individual tree defoliation events.", booktabs = TRUE) %>% 
+kable(tree_dat, caption = "(ref:tbl-defolCap)", booktabs = TRUE) %>% 
   kableExtra::kable_styling(latex_options = "hold_position")
@@
 The defoliation segments are colored by their relative severity based on their average NGSI value.
-The default cut-off values between "Severe" and "Moderate" is the overall mean across all events.
-The cut-off between "Moderate"  and "Minor" is the first quartile by default.
+By default, `plot_defol()` will calculate the average NGSI for all identified events, and assign severity based on the mean and first quartile of the averages.
+"Severe" events have a mean NGSI above the overall average event-period NGSI.
+"Moderate" events fall between the mean and first quartile.
+"Minor" events fall below the first quartile.
 Users can re-define the breaks to suit their needs via the "breaks" parameter in `plot_defol()`.
@@
 
-The default plotting function to visualize results from `outbreak()` creates a three-panel graph showing the mean site-level chronology, the percent of trees recording a defoliation, and the sample depth over time (Figure \@ref(fig:fig-obr)).
+The default plotting function to visualize results from `outbreak()` is `plot_outbreak()`.
+It creates a three-panel graph showing the sample depth, mean site-level chronology, and percent of trees recording a defoliation over time (Figure \@ref(fig:fig-obr)).
 
 ```{r echo=TRUE, eval=FALSE}
-plot_outbreak(dmj_obr, disp_index = "mean_gsi")
+plot_outbreak(dmj_obr, disp_index = "GSI")
 ```
 
-Inferred outbreak events are shown in the top panel of Figure \@ref(fig:fig-obr) as the filled-in spaces.
+Inferred outbreak events are shown in the middle panel of Figure \@ref(fig:fig-obr) as the filled-in spaces.
 Users can change the time series in this panel with the "disp_index" parameter, choosing between the mean NGSI (the default) or GSI.
@@
 
+(ref:tbl-obrCap) Site-level summary statistics for inferred outbreak events provided with the `outbreak_stats()` function for the DMJ example site.
+
 ```{r tbl-obr, echo=FALSE}
 obr_dat <- read.csv(here("paper_elsevier", "Output", "obr-stats.csv"))
-kable(obr_dat, format = "latex", caption = "Summary statistics for inferred outbreak events.", booktabs = TRUE) %>% 
+kable(obr_dat, format = "latex", caption = "(ref:tbl-obrCap)", booktabs = TRUE) %>% 
   kable_styling(font_size = 7) %>% 
@@
 The output table from `outbreak_stats()` shows the start and end years of each event, along with their corresponding duration, the number and percent of trees in defoliation at the start of the event, the the number of trees recording the outbreak event (in part or full), the minimums of the mean GSI and NGSI indices, and two points of "peak defoliation." The first, "peak_outbreak_year" is the year in which the greatest number of trees are recording the outbreak, and the second, "peak_defol_year" is the year with the lowest average NGSI.
-These output variables provide a array of options for assessing metrics of the insect defoliation regime, including interval analyses:
+
+Saving the results of `outbreak_stats()` (the `dmj_obr_stats` object above) provides an array of options for assessing metrics of the insect defoliation regime.
+Here we use the first year of each outbreak event to calculate the average duration of years between outbreaks, via the `diff()` function in `R`.
+The average of those differences, calculated via `mean()` is the mean return interval.
 

### 2020-05-14 paper_elsevier/paper_elsevier.tex
merging Jodi and Ann's reviews

@@
 
-  \title{\emph{dfoliatR}: An R package for detection and analysis of insect defoliation signals in tree rings}
+  \title{\texttt{dfoliatR}: An R package for detection and analysis of insect defoliation signals in tree rings}
     \author[a,b]{Christopher H. Guiterman\corref{1}}
@@
    \ead{ann.m.lynch@usda.gov} 
-    \author[c]{Jodi N. Axelson}
+    \author[d]{Jodi N. Axelson}
    \ead{jodi.axelson@berkeley.edu} 
@@
   \end{abstract}
-   \begin{keyword} Dendroecology, spruce budworm, \emph{Choristoneura}, pandora moth, \emph{Coloradia pandora Blake}, larch-bud-moth\end{keyword}
+   \begin{keyword} Dendroecology, dendroentomology, OUTBREAK, western spruce budworm, defoliator regimes, insect outbreak reconstruction\end{keyword}
  \end{frontmatter}
@@
 Defoliation can be distinguished from climate- and other disturbance-related influences by comparing ring-width or other annually-resolved features in the wood of host species to that of non-host species or annually-resolved climate records.
-The effect of defoliation on radial growth of trees has been recognized since the 1860s, and used to reconstruct outbreak regimes since the 1950s (Blais, 1954; Lynch, 2012).
+The effect of defoliation on radial growth of trees has been recognized since the 1860s, and used to reconstruct outbreak regimes since the 1950s (Blais, 1954; Alfaro et al., 1982; Lynch, 2012).
 It was not until the 1980s, however, that precise dendrochronological techniques were applied for inferring defoliation events and reconstructing defoliator outbreak regimes (Swetnam et al., 1985; Speer, 2010; Lynch, 2012).
@@
 \texttt{dfoliatR} adds to a growing suite of dendrochronology packages in the \texttt{R} computing environment (R Core Team, 2019).
-Stemming from the \texttt{dplR} library (Bunn, 2008) that enables \texttt{R} users to read and write an array of tree-ring data formats, standardize ring width series, build and evaluate chronologies, and perform quality control (to name a few), one can now also measure ring widths from scanned images of prepared samples (Lara et al., 2015; Shi et al., 2019), conduct and check crossdating (Bunn, 2010), analyze sub-annual anotomical features (Rathgeber et al., 2011; Campelo et al., 2016), and perform many analytical tests (Zang and Biondi, 2015; Jevšenak and Levanič, 2018).
+Stemming from the \texttt{dplR} library (Bunn, 2008) that enables \texttt{R} users to read and write an array of tree-ring data formats, standardize ring width series, build and evaluate chronologies, and perform quality control (to name a few), one can now also measure ring widths from scanned images of prepared samples (Lara et al., 2015; Shi et al., 2019), conduct and check crossdating (Bunn, 2010), analyze sub-annual anatomical features (Rathgeber et al., 2011; Campelo et al., 2016), and perform many analytical tests (Zang and Biondi, 2015; Jevšenak and Levanič, 2018).
 Tools for assessing stand dynamics and disturbance analyses are under rapid development, with new packages for assessing growth and release events (\texttt{TRADER}: Altman et al., 2014), metrics of growth resilience (\texttt{pointRes}: van der Maaten-Theunissen et al., 2015), and fire history (\texttt{burnr}: Malevich et al., 2018).
@@
 
-In this paper, we describe the statistical methods employed by \texttt{dfoliatR}, compare results to those produced by OUTBREAK, and present an example analysis inlcuding test data sets and script.
+In this paper, we describe the statistical methods employed by \texttt{dfoliatR}, compare results to those produced by OUTBREAK, and present an example analysis including test data sets and script.
 Users need not have much experience in \texttt{R} to replicate the analyses and graphics as presented.
@@
 It is important that the host-tree data include only one tree-ring series per tree.
-Both \texttt{dplR} and ARSTAN have options for averaging multiple sample series into a tree-level series.
+\texttt{dplR} (via the \texttt{dplR::treeMean()} function) and dpl versions of ARSTAN have options for averaging multiple sample series into a tree-level series.
 
 At the heart of \texttt{dfoliatR} lies two functions: \texttt{defoliate\_trees()} and \texttt{outbreak()}.
-These identify defoliation events on individual trees (Figure \ref{fig:fig-defol}) and then composite across multiple trees to identify outbreak events (Figure \ref{fig:fig-obr}).
+These identify defoliation events on individual trees (Figure \ref{fig:fig-defol}) and then composite across multiple trees to infer stand or site level outbreak events (Figure \ref{fig:fig-obr}).
 
@@
 \includegraphics{D:/Dropbox/dfoliatR_paper/paper_elsevier/Output/tree-plot-default.pdf}
-\caption{\label{fig:fig-defol}Default \texttt{dfoliatR} graphics for individual trees, produced by the \texttt{plot\_defol()} function.
-Dotted lines represent the series length for each tree and colored segments show periods of defoliation.
-Users can define the cut-off values to determine severe--moderate--minor defoliation intensities, and use standard \texttt{ggplot2} graphics parameters to adjust styles and themes.}
+\caption{\label{fig:fig-defol}Default \texttt{dfoliatR} graphic for individual trees, produced by the \texttt{plot\_defol()} function.
+The series names are listed on the Y-axis, dotted lines represent the series length for each tree, and colored segments show periods of defoliation.
+The colors of defoliation segments represent its severity, for which users can define cut-off values to determine severe--moderate--minor defoliation intensities.
+The default break points for severity classes are the mean and first quartile for event NGSI values.
+Colors and other features of the graphic can be adjusted using \texttt{ggplot2} parameters, as shown below.}
 \end{figure}
@@
 \caption{\label{fig:fig-obr}Default graphics showing outbreak events at the DMJ site.
-Top panel shows the mean GSI through time, with inferred outbreak events filled.
-The middle panel shows the percent of trees defoliated, and the bottom panel shows the number of trees through time.}
+Top panel shows the sample depth, the middle panel shows the mean GSI with inferred outbreak events filled, and the bottom panel shows the percent of trees defoliated, used to identify time periods of inferred outbreaks.}
 \end{figure}
@@
 Negative departures in the normalized GSI (NGSI, or GSI converted to z-scores) that surpass user-specified thresholds in duration and magnitude are defined as \emph{defoliation events}.
-As in OUTBREAK, the lowest NGSI value in the particucar sequence being assessed must reach the magnitude threshold.
-The default setting is -1.28 (NGSI is in units of standard deviation), which was previously determined to be representative of WSBW effects (Swetnam and Lynch, 1989) and is commonly used for other species.
+As in OUTBREAK, the lowest NGSI value in the particular sequence being assessed must reach the magnitude threshold.
+The default setting is -1.28 (NGSI is in units of standard deviation), which was previously determined to be representative of WSBW effects (Swetnam and Lynch, 1989) and is commonly used for other species (see Lynch, 2012).
 The year with the lowest value is termed the ``year of maximum departure'' and becomes a central point in time for assessing other thresholds before being included as a defoliation event.
@@
 Researchers can, and should, adjust the duration and magnitude parameters accordingly and critically evaluate the results, as insect species vary in the length of their outbreaks and the degree to which they can suppress tree growth.
+OUTBREAK provides two sets of default values, those for WSBW, which typically has lengthy outbreaks, and ones for Douglas-fir tussock moth (\emph{Orgyia pseudotsugata} (Lepidoptera: Tortricidae) of three years duration with -1.28 departure threshold that may be suitable for more eruptive species.
 
@@
 
-Diverging from OUTBREAK, \texttt{dfoliatR} allows users to extend defoliation events on individual trees by bridging between sequential events (Figure \ref{fig:fig-bridge}).
+Diverging from OUTBREAK, \texttt{dfoliatR} includes an option allowing users to extend defoliation events on individual trees by bridging between sequential events (Figure \ref{fig:fig-bridge}).
 In cases where two defoliation events are separated by a single year, bridging will link them into a single event.
-This option was added to \texttt{dfoliatR} during the testing phase of development, when we realized that OUTBREAK deliberately omits sequential, or back-to-back events, even when these events surpass both the magnitude and duration thresholds.
+This option was added to \texttt{dfoliatR} during the testing phase of development, when we realized that OUTBREAK deliberately omits sequential, or back-to-back events, even when both events surpass the magnitude and duration thresholds.
 Instead, OUTBREAK will select the one sequential event with the lowest negative departure year.
@@
 This is particularly relevant to WSBW and spruce budworm (\emph{C.
-fumiferana}), for which outbreak regimes have been reconstructed for several geographic areas and for which researchers and forest health experts have gained considerable knowledge (Brookes et al., 1982, and many later publications; Schmitt et al., 1984; Sanders et al., 1985).
-In many cases, the greatest growth suppression often occurs late in the outbreak due to cumulative effects accrued on a tree's resources (Swetnam and Lynch, 1989, 1993).
-We urge caution in using the bridging option, however, because it may not be appropriate for all studied insects, such as in situations where impacted stands barely recover from one outbreak before another begins, as with pine processionary caterpillars (\emph{Thaumetopoea pityocampa} (Lepidoptera: Thaumetopoeidae)) (Carus, 2004, 2009).
+fumiferana}), for which multiple outbreak regime reconstruction, as well as other research and forest health observations, show that outbreaks can be very long (Schmitt et al., 1984; Sanders et al., 1985; Brookes et al., 1987, and many later publications).
+Often the greatest growth suppression occurs late in the outbreak due to lag effects between defoliation and radial growth, and to cumulative effects accrued on a tree's resources (Brubaker, 1978; Alfaro et al., 1982; Wickman, 1986; Swetnam and Lynch, 1989, 1993; Mason et al., 1997; Axelson et al., 2014).
+We urge caution in using the bridging option, however, because it may not be appropriate for all studied insects, such as in situations where impacted stands barely recover from one outbreak before another begins, as with pine processionary caterpillars (\emph{Thaumetopoea pityocampa} (Lepidoptera: Thaumetopoeidae)) (Carus, 2004, 2009) or where outbreaks are known to be very short, such as larch budmoth (\emph{Zeiraphera diniana} Gn.) in the European Alps (Esper et al., 2007).
+
+Examples of sequential defoliation events (bars), wherein two or three events occur separated by single years.
+In each case OUTBREAK recorded only one of the events, omitting those with a higher ``max\_defol'' year, or the year of the lowest NGSI value.
+\texttt{dfoliatR}, by contrast, recorded all of the events, including two sequential events on DMJ26 and EFK22, and three on FNL04.
+Events recorded by both programs have a darker color, while those recorded only by \texttt{dfoliatR} are lighter in color.
+Adding the bridging option in \texttt{dfoliatR} links these events (``bridge\_defol'') into single longer events that may more accurately capture the effects of long-term insect herbivory on individual trees.
 
@@
 \centering
-\includegraphics{D:/Dropbox/dfoliatR_paper/paper_elsevier/Output/bridging.pdf}
-\caption{\label{fig:fig-bridge}Examples of sequential defoliation events (bars), wherein two or three events occur separated by single years.
-In each case OUTBREAK recorded only one of the events, ommitting those with a higher ``max\_defol'' year, or the year of the lowest NGSI value.
-\texttt{dfoliatR}, by contrast, recorded all of the events, including two sequential events on DMJ26 and EFK22, and three on FNL04.
-Events recorded by both programs have a darker color, while those recorded only by \texttt{dfoliatR} are lighter in color.
-Adding the bridging option in \texttt{dfoliatR} links these events (``bridge\_defol'') into single longer events that may more accurately capture the effects of long-term insect herbivory on individual trees.}
+\includegraphics{D:/Dropbox/dfoliatR_paper/paper_elsevier/Output/bridging_color.pdf}
+\caption{\label{fig:fig-bridge}Examples of identified defoliation events on individual trees.
+For each tree (DMJ26 and EFK22), OUTBREAK and \texttt{dfoliatR} identify most of the same events, but there is one added event (blue colored) that was omitted by OUTBREAK.
+These were omitted because they were sequential, separated by a single year with possitive normalized growth suppression index (NGSI).
+\texttt{dfoliatR} provides an option to bridge these events into single long events that may better represent the duration of defoliation given the insect and sites under consideration.}
 \end{figure}
@@
 We tested \texttt{dfoliatR} against OUTBREAK by comparing NGSI to OUTBREAK's normalized corrected indices for individual trees and years, defoliation status for individual trees and years, and percentage of trees recording outbreaks at the site level.
-Our tests used standardized ring-width data from eight sites in British Columbia, Colorado, and New Mexico, and author-provided non-host site chronologies.
+Our tests used standardized ring-width data from eight host-tree sites spanning the range of WSBW.
+The sites were sampled in British Columbia (Axelson et al., 2015), Wyoming (Axelson et al., 2018), Colorado, and New Mexico (Swetnam and Lynch, 1993; Ryerson et al., 2003).
+These host data were compared to non-host chronologies from the original studies, but we made no effort here to replicate the reconstructions or analyses of those studies.
 
 We detrended host data for both \texttt{dfoliatR} and OUTBREAK using ARSTAN {[}v6.1{]} with cubic smoothing splines (50\% frequency response on 100-150 year wavelengths depending on the site).
-In both \texttt{dfoliatR} and OUTBREAK we used event thresholds of -1.28 normalized indices, 8 years duration, and allowed for events at the end of series in seven of eight sites.
-These seven sites were sampled during ongoing outbreak events (Swetnam and Lynch, 1993).
+In both \texttt{dfoliatR} and OUTBREAK we used event thresholds of -1.28 normalized indices, 8 years duration, and allowed for events at the end of series in seven of eight sites that had known outbreaks at the time of sampling.
 We found it necessary to be consistent in how we detrended and what software we employed (e.g., ARSTAN vs \texttt{dplR}) because subtle differences in standardized ring-width indices generated between the programs transferred into differences between \texttt{dfoliatR} and OUTBREAK.
@@
 \item
-  \emph{Sequential events} (36\%) in which OUTBREAK omitted back-to-back events that occur one year prior or one year following an identified event.
+  \emph{Sequential events} (36\%) in which OUTBREAK omitted back-to-back events that occur one year prior to, or one year following an identified event.
 When this occurs, OUTBREAK selects the one event sequence with the lowest negative index year (e.g., Figure \ref{fig:fig-bridge}).
 On two trees, OUTBREAK omitted two of three sequential events.
-While inspecting these differences, we added an option to \texttt{defoliate\_trees()} that would ``bridge'' between sequential events (that surpass magnitude and duration thresholds) into single, long events.
-We felt that this was ecologically justified, especially for studies of WSBW, because the insect will occasionally defoliate individual trees for decades and reconstructed chronologies have show outbreaks as long as \textbf{xxx (ANN CHECK)}.
+While inspecting these differences, we added an option to \texttt{defoliate\_trees()} that would ``bridge'' between sequential events (that each surpass the magnitude and duration thresholds) into single, long events.
+We felt that this was ecologically justified, especially for studies of WSBW, because outbreaks are known to be of long duration and tree-ring reconstructions have shown that outbreaks may persist 30-50 at the site level.
 \item
@@
 At the site level, OUTBREAK and \texttt{dfoliatR} produce similar time series of percent trees defoliated (Figure \ref{fig:fig-df-obr}), which forms the basis for inferring outbreak occurrence, intensity, and duration.
-In nearly site comparison, \texttt{dfoliatR} included either more events or longer durations of inferred outbreak.
-These differences arise from the inclusion of tree-level events by \texttt{dfoliatR} that were omitted by OUTBREAK.
-The differences translate to greater numbers of trees in defoliation during certain events, or events recorded by single trees during periods of low sample depth.
+In nearly all site comparison, \texttt{dfoliatR} included either more events or longer durations of inferred outbreaks.
+These differences arise from the inclusion of tree-level events by \texttt{dfoliatR} that were omitted by OUTBREAK (see note on sequential events above).
+Thus, in \texttt{dfoliatR}, there were a greater number of trees experiencing defoliation during outbreak periods, or outbreaks were represented by a single tree when
+there was low sample depth.
 
@@
 Input parameters were identical between programs.
-Difference arrise because \texttt{dfoliatR} will identify and record more defoliation events on idnividual trees, and may more accurately capture herbivory dynamics.}
+Differences arise because \texttt{dfoliatR} will identify and record more defoliation events on individual trees.}
 \end{figure}
 
-This evaluation revealed what we believe are short-comings in how OUTBREAK identifies defoliation events on individual trees.
-In every one of the 102 cases we inspected, we felt that \texttt{dfoliatR} provided a more biologically and statistically appropriate assessment of defoliation.
-Therefore, \texttt{dfoliatR} also did a better job inferring outbreak events and associated statistics at the site level.
+This comparison revealed what we believe are shortcomings in how OUTBREAK identifies defoliation events on individual trees.
+In every one of the 102 cases we inspected, we felt that \texttt{dfoliatR} provided a more biologically and statistically appropriate assessment of defoliation, translating to more robust inferences of outbreak events and associated statistics at the site level.
 
@@
 As an Open Source library, we welcome and encourage community involvement in future development.
-The best ways to contribute to \texttt{dfoliatR} are through standard GitHub procedures or by contacting the first author.
+The best ways to contribute to \texttt{dfoliatR} are through standard GitHub procedures or by contacting the corresponding author.
 
@@
 Host trees from Demijohn Peak (DMJ; 2902 m asl) in the San Juan Mountains of southern Colorado include Douglas-fir (\emph{Pseudotsuga menziesii}) compared against a local non-host ponderosa pine (\emph{Pinus ponderosa}) chronology (Ryerson et al., 2003).
-The East Fork site (EFK; 2580 m asl) in the Jemez Mountains of northcentral New Mexico includes Douglas-fir and white fir (\emph{Abies concolor}) host trees and a ponderosa pine non-host chronology (Swetnam and Lynch, 1993).
+The East Fork site (EFK; 2580 m asl) in the Jemez Mountains of north-central New Mexico includes Douglas-fir and white fir (\emph{Abies concolor}) host trees and a ponderosa pine non-host chronology (Swetnam and Lynch, 1993).
 
@@
 In our example scripts below, \texttt{\#\#} and \texttt{\#} denote user comments, per standard \texttt{R} coding, which are colored in brown.
-Text in blue denote functions; black are loaded objects, and green are quoted variables and links.
+Text in blue denotes functions; black are loaded objects, and green are quoted variables and links.
+Values or other information provided after equal signs are filenames and parameters provided for this example, and in actual use would be replaced with user-specified information.
+In this example ``dmj\_h'' and ``dmj\_nh'' are the individual-tree host series and non-host site chronology files for the Demijohn site, thresholds are set at 8 years and -1.28 standard deviations, bridging is used, series-end events are included in the interval computations, and comprehensive results information is not included in the output.
 
@@
 
-\caption{\label{tab:tbl-defol}Summary statistics for individual tree defoliation events.}
+\caption{\label{tab:tbl-defol}Tree-level tabular output provided by the \texttt{defol\_stats()} function for the DMJ example site.}
 \centering
@@
 The defoliation segments are colored by their relative severity based on their average NGSI value.
-The default cut-off values between ``Severe'' and ``Moderate'' is the overall mean across all events.
-The cut-off between ``Moderate'' and ``Minor'' is the first quartile by default.
+By default, \texttt{plot\_defol()} will calculate the average NGSI for all identified events, and assign severity based on the mean and first quartile of the averages.
+``Severe'' events have a mean NGSI above the overall average event-period NGSI.
+``Moderate'' events fall between the mean and first quartile.
+``Minor'' events fall below the first quartile.
 Users can re-define the breaks to suit their needs via the ``breaks'' parameter in \texttt{plot\_defol()}.
@@
 
-The default plotting function to visualize results from \texttt{outbreak()} creates a three-panel graph showing the mean site-level chronology, the percent of trees recording a defoliation, and the sample depth over time (Figure \ref{fig:fig-obr}).
+The default plotting function to visualize results from \texttt{outbreak()} is \texttt{plot\_outbreak()}.
+It creates a three-panel graph showing the sample depth, mean site-level chronology, and percent of trees recording a defoliation over time (Figure \ref{fig:fig-obr}).
 
@@
 \begin{Highlighting}[]
-\KeywordTok{plot_outbreak}\NormalTok{(dmj_obr, }\DataTypeTok{disp_index =} \StringTok{"mean_gsi"}\NormalTok{)}
+\KeywordTok{plot_outbreak}\NormalTok{(dmj_obr, }\DataTypeTok{disp_index =} \StringTok{"GSI"}\NormalTok{)}
 \end{Highlighting}
@@
 
-Inferred outbreak events are shown in the top panel of Figure \ref{fig:fig-obr} as the filled-in spaces.
+Inferred outbreak events are shown in the middle panel of Figure \ref{fig:fig-obr} as the filled-in spaces.
 Users can change the time series in this panel with the ``disp\_index'' parameter, choosing between the mean NGSI (the default) or GSI.
@@
 
-\caption{\label{tab:tbl-obr}Summary statistics for inferred outbreak events.}
+\caption{\label{tab:tbl-obr}Site-level summary statistics for inferred outbreak events provided with the \texttt{outbreak\_stats()} function for the DMJ example site.}
 \centering
@@
 The output table from \texttt{outbreak\_stats()} shows the start and end years of each event, along with their corresponding duration, the number and percent of trees in defoliation at the start of the event, the the number of trees recording the outbreak event (in part or full), the minimums of the mean GSI and NGSI indices, and two points of ``peak defoliation.'' The first, ``peak\_outbreak\_year'' is the year in which the greatest number of trees are recording the outbreak, and the second, ``peak\_defol\_year'' is the year with the lowest average NGSI.
-These output variables provide a array of options for assessing metrics of the insect defoliation regime, including interval analyses:
+
+Saving the results of \texttt{outbreak\_stats()} (the \texttt{dmj\_obr\_stats} object above) provides an array of options for assessing metrics of the insect defoliation regime.
+Here we use the first year of each outbreak event to calculate the average duration of years between outbreaks, via the \texttt{diff()} function in \texttt{R}.
+The average of those differences, calculated via \texttt{mean()} is the mean return interval.
 
@@
 The \texttt{dfoliatR} package provides dendroecologists with tools to infer, quantify, analyze, and visualize tree-ring growth suppression events and to reconstruct forest insect defoliator outbreak regimes.
-It is built on the long-accepted host to nonhost comparison methodology used in the 1980s FORTRAN program OUTBREAK (Swetnam et al., 1985; Swetnam and Lynch, 1989, 1993).
-Our evaluation of the two programs revelaed that \texttt{dfolaitR} excelled in identifying defoliation events on single trees, providing researchers with more consistent and biologically-justifiable results.
-Other key benefits are that \texttt{dfoliatR} provides easier control of suppression thresholds, additional output tables, high-quality and customizable graphics, and operates in the open-source \texttt{R} environment that is stable across computing platforms and under active development and maintence by a large and growing community.
+It is built on the long-accepted host to non-host comparison methodology used in the 1980s FORTRAN program OUTBREAK (Swetnam et al., 1985; Swetnam and Lynch, 1989, 1993).
+Our evaluation of the two programs revealed that \texttt{dfoliatR} excelled in identifying defoliation events on single trees, providing researchers with more consistent and biologically-justifiable results.
+Other key benefits are that \texttt{dfoliatR} provides easier control of suppression thresholds, additional output tables, high-quality and customizable graphics, and operates in the open-source \texttt{R} environment that is stable across computing platforms and under active development and maintenance by a large and growing community.
 
@@
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-Alfaro1982}{}%
+Alfaro, R., Sickle, G.
+van, Thomson, A., Wegwitz, E., 1982.
+Tree mortality and radial growth losses caused by the western spruce budworm in a Douglas-fir stand in British Columbia.
+Canadian Journal of Forest Research 12, 780--87.
+
 \leavevmode\hypertarget{ref-Altman2014}{}%
 Altman, J., Fibich, P., Dolezal, J., Aakala, T., 2014.
-TRADER: A package for Tree Ring Analysis of Disturbance Events in R.
+\texttt{TRADER}: A package for Tree Ring Analysis of Disturbance Events in R.
 Dendrochronologia 32, 107--12, doi:\href{https://doi.org/10.1016/j.dendro.2014.01.004}{10.1016/j.dendro.2014.01.004}.
+
+\leavevmode\hypertarget{ref-Axelson2014}{}%
+Axelson, J., Bast, A., Alfaro, R., Smith, D., Gärtner, H., 2014.
+Variation in wood anatomical structure of Douglas-fir defoliated by the western spruce budworm: a case study in the coastal-transitional zone of British Columbia, Canada.
+Trees 28, 1837--46, doi:\href{https://doi.org/10.1007/s00468-014-1091-1}{10.1007/s00468-014-1091-1}.
+
+\leavevmode\hypertarget{ref-NADEF2018}{}%
+Axelson, J., Csank, A., Lee, R., McGee, J., McNiel, K., 2018.
+Reconstructing historical \emph{Choristoneura freemani} outbreaks using \emph{Pseudotsuga menziesii} in the greater Yellowstone region, Wyoming, U.S.A.
+Final Report of the 20th Annual North American Dendroecological Fieldweek (NADEF).
+
+\leavevmode\hypertarget{ref-Axelson2015}{}%
+Axelson, J.N., Smith, D.J., Daniels, L.D., Alfaro, R.I., 2015.
+Multicentury reconstruction of western spruce budworm outbreaks in central British Columbia, Canada.
+Forest Ecology and Management 335, 235--48, doi:\href{https://doi.org/10.1016/j.foreco.2014.10.002}{10.1016/j.foreco.2014.10.002}.
 
@@
 \leavevmode\hypertarget{ref-Brookes1987}{}%
-Brookes, M.H., Campbell, R.W., Colbert, J.J., Mitchell, R.G., (Technical Coordinators), R.W.S., 1982.
+Brookes, M.H., Campbell, R.W., Colbert, J.J., Mitchell, R.G., Stark, R.W., 1987.
 Western spruce budworm.
@@
 
+\leavevmode\hypertarget{ref-Brubaker1978}{}%
+Brubaker, L.B., 1978. Effects of defoliation by douglas-fir tussock moth on ring sequences of douglas-fir and grand fir. Tree-Ring Bulletin 38, 49--60.
+
 \leavevmode\hypertarget{ref-Bunn2008}{}%
-Bunn, A.G., 2008. A dendrochronology program library in R (dplR). Dendrochronologia 26, 115--24, doi:\href{https://doi.org/10.1016/j.dendro.2008.01.002}{10.1016/j.dendro.2008.01.002}.
+Bunn, A.G., 2008. A dendrochronology program library in R (\texttt{dplR}). Dendrochronologia 26, 115--24, doi:\href{https://doi.org/10.1016/j.dendro.2008.01.002}{10.1016/j.dendro.2008.01.002}.
 
 \leavevmode\hypertarget{ref-Bunn2010}{}%
-Bunn, A.G., 2010. Statistical and visual crossdating in R using the dplR library. Dendrochronologia 28, 251--58, doi:\href{https://doi.org/10.1016/j.dendro.2009.12.001}{10.1016/j.dendro.2009.12.001}.
+Bunn, A.G., 2010.
+Statistical and visual crossdating in \texttt{R} using the \texttt{dplR} library.
+Dendrochronologia 28, 251--58, doi:\href{https://doi.org/10.1016/j.dendro.2009.12.001}{10.1016/j.dendro.2009.12.001}.
 
@@
 Campelo, F., Nabais, C., Carvalho, A., Vieira, J., 2016.
-tracheideR--An R package to standardize tracheidograms.
+\texttt{tracheideR}--An R package to standardize tracheidograms.
 Dendrochronologia 37, 64--68, doi:\href{https://doi.org/10.1016/j.dendro.2015.12.006}{10.1016/j.dendro.2015.12.006}.
@@
 
+\leavevmode\hypertarget{ref-Esper2007}{}%
+Esper, J., Büntgen, U., Frank, D.C., Nievergelt, D., Liebhold, A., 2007.
+1200 years of regular outbreaks in alpine insects.
+Proceedings of the Royal Society B: Biological Sciences 274, 671--79, doi:\href{https://doi.org/10.1098/rspb.2006.0191}{10.1098/rspb.2006.0191}.
+
 \leavevmode\hypertarget{ref-dfoliatR}{}%
 Guiterman, C., Lynch, A., Axelson, J., 2020.
-DfoliatR: Detection and analysis of insect defoliation signals in tree rings., \url{https://chguiterman.github.io/dfoliatR/}, doi:\href{https://doi.org/10.5281/zenodo.3698163}{10.5281/zenodo.3698163}.
+\texttt{dfoliatR}: Detection and Analysis of Insect Defoliation Signals in Tree Rings., \url{https://chguiterman.github.io/dfoliatR/}, doi:\href{https://doi.org/10.5281/zenodo.3698163}{10.5281/zenodo.3698163}.
 
@@
 Jevšenak, J., Levanič, T., 2018.
-dendroTools: R package for studying linear and nonlinear responses between tree-rings and daily environmental data.
+\texttt{dendroTools}: R package for studying linear and nonlinear responses between tree-rings and daily environmental data.
 Dendrochronologia 48, 32--39, doi:\href{https://doi.org/10.1016/j.dendro.2018.01.005}{10.1016/j.dendro.2018.01.005}.
@@
 Lara, W., Bravo, F., Sierra, C.A., 2015.
-MeasuRing: An R package to measure tree-ring widths from scanned images.
+\texttt{MeasuRing}: An R package to measure tree-ring widths from scanned images.
 Dendrochronologia 34, 43--50, doi:\href{https://doi.org/10.1016/j.dendro.2015.04.002}{10.1016/j.dendro.2015.04.002}.
@@
 Malevich, S.B., Guiterman, C.H., Margolis, E.Q., 2018.
-burnr: Fire history analysis and graphics in R.
+\texttt{burnr}: Fire history analysis and graphics in R.
 Dendrochronologia 49, 9--15, doi:\href{https://doi.org/10.1016/j.dendro.2018.02.005}{10.1016/j.dendro.2018.02.005}.
+
+\leavevmode\hypertarget{ref-Mason1997}{}%
+Mason, R.R., Wickman, B.E., Paul, H.G., 1997.
+Radial growth response of douglas-fir and grand fir to larval densities of the douglas-fir tussock moth and the western spruce budworm.
+Forest Science 43, 194--205.
 
@@
 Rathgeber, C.B.K., Longuetaud, F., Mothe, F., Cuny, H., Le Moguédec, G., 2011.
-Phenology of wood formation: Data processing, analysis and visualisation using R (package CAVIAR).
+Phenology of wood formation: Data processing, analysis and visualisation using R (package \texttt{CAVIAR}).
 Dendrochronologia 29, 139--49, doi:\href{https://doi.org/10.1016/j.dendro.2011.01.004}{10.1016/j.dendro.2011.01.004}.
@@
 \leavevmode\hypertarget{ref-RCore}{}%
-R Core Team, 2019. R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria, \url{https://www.R-project.org/}.
+R Core Team, 2019. \texttt{R}: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria, \url{https://www.R-project.org/}.
 
@@
 \leavevmode\hypertarget{ref-Sanders1985}{}%
-Sanders, C.J., Stark, R.W., Mullins, E.J., (eds), J.M., 1985.
-Recent advances in spruce budworms research.
-in: Proceedings of the CANUSA spruce budworms research symposium, 16-20 september 1984, bangor, maine.
-Canadian forestry service, ottawa, ontario.
-527p..
+Sanders, C.J., Stark, R.W., Mullins, E.J., Murphy, J.
+(eds), 1985.
+Recent Advances in Spruce Budworms Research, \emph{Proceedings of the CANUSA Spruce Budworms Research Symposium}, 16-20 September, Bangor, ME.
+Canadian Forestry Service, Ottawa, Ontario.
+527p.
 
@@
 Shi, J., Xiang, W., Liu, Q., Shah, S., 2019.
-MtreeRing: An r package with graphical user interface for automatic measurement of tree ring widths using image processing techniques.
+\texttt{MtreeRing}: An R package with graphical user interface for automatic measurement of tree ring widths using image processing techniques.
 Dendrochronologia 58, 125644, doi:\href{https://doi.org/10.1016/j.dendro.2019.125644}{10.1016/j.dendro.2019.125644}.
@@
 van der Maaten-Theunissen, M., van der Maaten, E., Bouriaud, O., 2015.
-PointRes: An R package to analyze pointer years and components of resilience.
+\texttt{PointRes}: An R package to analyze pointer years and components of resilience.
 Dendrochronologia 35, 34--38, \url{http://dx.doi.org/10.1016/j.dendro.2015.05.006}, doi:\href{https://doi.org/10.1016/j.dendro.2015.05.006}{10.1016/j.dendro.2015.05.006}.
@@
 \leavevmode\hypertarget{ref-wickham2016ggplot2}{}%
-Wickham, H., 2016. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag, New York, \url{https://ggplot2-book.org/}.
+Wickham, H., 2016. \texttt{ggplot2}: Elegant Graphics for Data Analysis. Springer-Verlag, New York, \url{https://ggplot2-book.org/}.
 
@@
 Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L., François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M., Pedersen, T., Miller, E., Bache, S., Müller, K., Ooms, J., Robinson, D., Seidel, D., Spinu, V., Takahashi, K., Vaughan, D., Wilke, C., Woo, K., Yutani, H., 2019.
-Welcome to the tidyverse.
+Welcome to the \texttt{Tidyverse}.
 Journal of Open Source Software The Open Journal 4, 1686, \url{https://doi.org/10.21105/joss.01686}, doi:\href{https://doi.org/10.21105/joss.01686}{10.21105/joss.01686}.
@@
 \leavevmode\hypertarget{ref-Wickham2020dplyr}{}%
-Wickham, H., François, R., Henry, L., Müller, K., 2020. dplyr: A Grammar of Data Manipulation., \url{https://CRAN.R-project.org/package=dplyr}.
+Wickham, H., François, R., Henry, L., Müller, K., 2020. \texttt{dplyr}: A Grammar of Data Manipulation., \url{https://CRAN.R-project.org/package=dplyr}.
+
+\leavevmode\hypertarget{ref-Wickman1986}{}%
+Wickman, B.E., 1986.
+Radial growth of grand fir and douglas-fir 10 years after defoliation by the douglas-fir tussock moth in the blue mountains outbreak.
+USDA Forest Service, Pacific Northwest Research Station, Research Paper PNW-367.
+Portland, Oregon.
 
@@
 Zang, C., Biondi, F., 2015.
-Treeclim: An R package for the numerical calibration of proxy-climate relationships.
+\texttt{treeclim}: An R package for the numerical calibration of proxy-climate relationships.
 Ecography 38, 431--36, doi:\href{https://doi.org/10.1111/ecog.01335}{10.1111/ecog.01335}.

### 2020-05-14 paper_elsevier/R/Compare_OUTBREAK_dfoliatR.R

@@ -152,17 +152,13 @@ site_perc <- outbr_obr %>%
           select(site, year, perc_defol, software))
 
 ggplot(site_perc, aes(x = year, y = perc_defol)) +
-  geom_line(aes(color = software)) +
+  geom_line(aes(color = software, linetype = software)) +
+  scale_linetype_manual(values = c("solid", "longdash")) +
   ylab("Percent trees defoliated") +
   xlab("Year") +
   facet_wrap(~site, nrow = 4) +
-  # labs(tag = c("OUTBREAK", "dfoliatR")) +
-  # # coord_cartesian(clip = "off") +
-  # scale_color_identity(guide = "legend", name = "Software",
-  #                      # values = c("orange", "blue"),
-  #                      labels = c("OUTBREAK", "dfoliatR")) +
   theme_bw() +
-  # guides(color = guide_legend(nrow = 1)) +
+
   theme(legend.position = c(0.5, 1.1),
         legend.direction = "horizontal",
         legend.background = element_blank(),

### 2020-05-14 paper_elsevier/R/Example_Usage.R

@@ -1,5 +1,6 @@
 ## Create tables and graphics for dfoliatR manuscript
 
+devtools::install_github("chguiterman/dfoliatR@dev", force=TRUE)
 library(dfoliatR)
 library(ggplot2)
 library(patchwork)
@@ -45,7 +46,7 @@ dmj_obr <- outbreak(dmj_defol,
                     filter_min_defol = 1,
                     filter_perc = 25)
 
-plot_outbreak(dmj_obr, disp_index = "mean_gsi")
+plot_outbreak(dmj_obr, disp_index = "GSI")
 ggsave(paste0(out_pth, "site-plot.pdf"), width=5.75, dpi=300, units="in")
 
 dmj_obr_stats <- outbreak_stats(dmj_obr)

### 2020-05-14 paper_elsevier/R/bridging.R

@@ -49,48 +49,39 @@ df_defol <- df_run %>%
 
 # Selected bridge example -------------------------------------------------
 
-fnl04 <- df_defol %>% 
-  filter(series == "FNL04",
-         year %in% 1881:1925)
+# fnl04 <- df_defol %>% 
+#   filter(series == "FNL04",
+#          year %in% 1881:1925)
 efk22 <- df_defol %>% 
   filter(series == "EFK22",
-         year %in% 1940:1975)
+         year > 1850)
 dmj26 <- df_defol %>% 
   filter(series == "DMJ26",
-        year %in% 1850:1875)
+        year %in% 1750:1950)
 
-comps <- rbind(fnl04, efk22, dmj26)
+# comps <- rbind(fnl04, efk22, dmj26)
+comps <- rbind(efk22, dmj26)
 
 obr_events <- comps %>% 
-  filter(obr_type != "nd")
+  mutate(ngsi = replace(ngsi, obr_type == "nd", 0))
 df_events <- comps %>% 
-  filter(defol_status != "nd")
+  mutate(ngsi = replace(ngsi, defol_status == "nd", 0))
 
-cols <- dichromat::colorschemes$SteppedSequential.5[c(3, 8, 18)]
+# cols <- dichromat::colorschemes$SteppedSequential.5[c(3, 8, 18)]
 
+## COLOR
 ggplot(comps, aes(x=year)) +
   # scale_x_continuous(limits=c(1881, 1950)) +
   geom_hline(yintercept = 0, color = "grey80") +
-  geom_line(aes(y=ngsi), size=1.25) +
-  facet_wrap(~ series, scales = "free", ncol = 1) +
-  geom_segment(data = df_events,
-               aes(x=year, xend=year,
-                   y=0, yend=ngsi,
-                   color = stage(start = defol_status,
-                                 after_scale = clr_lighten(color, 0.25, space="combined"))),
-               size=2.5) +
-  geom_segment(data = obr_events,
-               aes(x=year, xend=year,
-                   y=0, yend=ngsi,
-                   color = stage(start = obr_type,
-                                 after_scale = clr_darken(color, 0.2))
-                   ),
-               size=2.5) +
+  facet_wrap(~ series, scales = "free_x", ncol = 1, strip.position = "right") +
+  geom_area(data=df_events, aes(x = year, y = ngsi, fill = "dfoliatR")) +
[…]

### 2020-05-21 paper_elsevier/paper_elsevier.Rmd
editing and re-running example script

@@
 abstract: |
- We present a new R package to provide dendroecologists with tools to infer, quantify, analyze, and visualize growth suppression events in tree rings.
-`dfoliatR` is based on the OUTBREAK program and builds on existing resources in the `R` and `dplR` computing environments.
+ We present a new `R` package to provide dendroecologists with tools to infer, quantify, analyze, and visualize growth suppression events in tree rings.
+`dfoliatR` is based on the OUTBREAK program and builds on existing resources in the `R` computing environment s the well-used `dplR` package.
 It is designed to aid research in the ecology of insect defoliation events and to reconstruct defoliator outbreak chronologies, but can be applied to other studies where host--non-host comparisons are useful.
-`dfoliatR` performs an indexing procedure to remove climatic signals in the host-tree series that are represented in the non-host chronologies, or other annually-resolved climate series.
+`dfoliatR` performs an indexing procedure to remove climatic signals in the host-tree series that are represented in the non-host chronology, or other annually-resolved climate series.
 It then infers defoliation events in individual trees based on user-specified thresholds.
@@
 
-(ref:obrCap) Default graphics showing outbreak events at the DMJ site.
+(ref:obrCap) Default graphic showing outbreak events at the DMJ site.
 Top panel shows the sample depth, the middle panel shows the mean GSI with inferred outbreak events filled, and the bottom panel shows the percent of trees defoliated, used to identify time periods of inferred outbreaks.
@@
 Researchers can, and should, adjust the duration and magnitude parameters accordingly and critically evaluate the results, as insect species vary in the length of their outbreaks and the degree to which they can suppress tree growth.
-OUTBREAK provides two sets of default values, those for WSBW, which typically has lengthy outbreaks, and ones for Douglas-fir tussock moth (*Orgyia pseudotsugata* (Lepidoptera: Tortricidae) of three years duration with -1.28 departure threshold that may be suitable for more eruptive species.
+OUTBREAK provides two sets of default values, those for WSBW, which typically has lengthy outbreaks, and ones for Douglas-fir tussock moth (*Orgyia pseudotsugata* (Lepidoptera: Tortricidae)) of three years duration with -1.28 departure threshold that may be suitable for more eruptive species.
 
@@
 (ref:BridgeCap) Examples of identified defoliation events on individual trees.
-For each tree (DMJ26 and EFK22), OUTBREAK and `dfoliatR` identify most of the same events, but there is one added event (blue colored) that was omitted by OUTBREAK.
-These were omitted because they were sequential, separated by a single year with possitive normalized growth suppression index (NGSI).
-`dfoliatR` provides an option to bridge these events into single long events that may better represent the duration of defoliation given the insect and sites under consideration.
-
-Examples of sequential defoliation events (bars), wherein two or three events occur separated by single years.
-In each case OUTBREAK recorded only one of the events, omitting those with a higher "max_defol" year, or the year of the lowest NGSI value.
-`dfoliatR`, by contrast, recorded all of the events, including two sequential events on DMJ26 and EFK22, and three on FNL04.
-Events recorded by both programs have a darker color, while those recorded only by `dfoliatR` are lighter in color.
-Adding the bridging option in `dfoliatR` links these events ("bridge_defol") into single longer events that may more accurately capture the effects of long-term insect herbivory on individual trees.
+For each tree (DMJ26 and EFK22), OUTBREAK and `dfoliatR` identify most of the same events, but there is one added event (in blue) that was omitted by OUTBREAK.
+These were omitted because they were separated by a single year of positive normalized growth suppression index (NGSI) and OUTBREAK selected the one event with the lowest maximum departure value.
+`dfoliatR` provides an option to bridge these sequential events into single long events that may better represent the duration of defoliation given the insect and sites under consideration.
 
@@
 Three parameters control whether a defoliation event constitutes an outbreak: the minimum number of trees available, the minimum number of trees recording defoliation, and the percent of trees recording defoliation.
-The first allows the researcher to make a judgement call as to the confidence ascribed to reduced sample depth toward the ends of their chronologies, thus compensating for the "fading record problem" [@Swetnam1999].
+The first allows the researcher to make a judgment call as to the confidence ascribed to reduced sample depth toward the ends of their chronologies, thus compensating for the "fading record problem" [@Swetnam1999].
 The second two parameters adjust the scale of defoliation considered to be an outbreak.
@@
 
-We detrended host data for both `dfoliatR` and OUTBREAK using ARSTAN [v6.1] with cubic smoothing splines (50% frequency response on 100-150 year wavelengths depending on the site).
+We detrended host data for both `dfoliatR` and OUTBREAK using ARSTAN (version 6.1) with cubic smoothing splines (50% frequency response on 100-150 year wavelengths depending on the site).
 In both `dfoliatR` and OUTBREAK we used event thresholds of -1.28 normalized indices, 8 years duration, and allowed for events at the end of series in seven of eight sites that had known outbreaks at the time of sampling.
@@
 
-Across the 43,280 ring-width indices from 222 trees included in our evaluation, we found that `dfoliatR` and OUTBREAK compute identical growth suppression indices at 0.000 precision.
-We expected this outcome because both programs employ Equation 1.
-At the tree-level, the programs identified 11,530 years of defoliation.
+Across the 43,280 ring-width indices from 222 trees included in our evaluation, we found that `dfoliatR` and OUTBREAK compute identical growth suppression indices at 0.00 precision.
+We expected this outcome because both programs apply Equation 1 to calcuate disturbance indices.
+At the tree-level, the programs identified 11,530 total index years with defoliation.
 The programs agreed on 97.9% of the years, leaving 927 "difference" years in which only one program identified defoliation on an individual tree.
@@
 In OUTBREAK, the option appears while changing the duration parameter (option 3).
-When selected, OUTBREAK will include any sequences of negative indices at the either the beginning or the end of each tree series as a defoliation event, without consideration of either duration or magnitude thresholds.
+When selected, OUTBREAK will include any sequences of negative indices at the beginning and the end of each tree series as a defoliation event, without consideration of either duration or magnitude thresholds.
 In `dfoliatR`, the duration threshold is omitted and the magnitude threshold is retained in series-end-events.
@@
 While inspecting these differences, we added an option to `defoliate_trees()` that would "bridge" between sequential events (that each surpass the magnitude and duration thresholds) into single, long events.
-We felt that this was ecologically justified, especially for studies of WSBW, because outbreaks are known to be of long duration and tree-ring reconstructions have shown that outbreaks may persist 30-50 at the site level.
+We felt that this was ecologically justified, especially for studies of WSBW, because outbreaks are known to be of long duration and tree-ring reconstructions have shown that outbreaks may persist for as long as 30-50 years at the site level.
 
@@
 At the site level, OUTBREAK and `dfoliatR` produce similar time series of percent trees defoliated (Figure \@ref(fig:fig-df-obr)), which forms the basis for inferring outbreak occurrence, intensity, and duration.
-In nearly all site comparison, `dfoliatR` included either more events or longer durations of inferred outbreaks.
+In nearly all site-level comparisons, `dfoliatR` included either more events or it inferred a longer duration outbreak.
 These differences arise from the inclusion of tree-level events by `dfoliatR` that were omitted by OUTBREAK (see note on sequential events above).
@@
 ```{r echo=TRUE, eval=FALSE}
-defol_stats(dmj_defol)
-View(dmj_defol)
+dmj_defol_stats <- defol_stats(dmj_defol)
+View(dmj_defol_stats)
 ```
@@
 Input parameters to `outbreak()` include "filter_min_series" to control the chronology cut-off points with regard to sample depth, "filter_min_defol" and "filter_perc" to control the minimum number and percent of trees recording a defoliation event in a given year.
-`Outbreak()` produces a new data frame with eight variables: "year", "num_defol", "percent_defol", "num_max_defol", "mean_gsi", "mean_ngsi", and "outbreak_status." All of these variables are populated regardless of an inferred outbreak event, providing a continuous disturbance chronology.
+`outbreak()` produces a new data frame with eight variables: "year", "num_defol", "percent_defol", "num_max_defol", "mean_gsi", "mean_ngsi", and "outbreak_status." All of these variables are populated regardless of an inferred outbreak event, providing a continuous outbreak reconstruction.
 The "num_max_defol" variable counts the number of trees recording their maximum defoliation in a given year.
@@
 
-Saving the results of `outbreak_stats()` (the `dmj_obr_stats` object above) provides an array of options for assessing metrics of the insect defoliation regime.
+Saving the results of `outbreak_stats()` (the `dmj_obr_stats` object above) provides an array of options for assessing metrics of the insect outbreak regime.
 Here we use the first year of each outbreak event to calculate the average duration of years between outbreaks, via the `diff()` function in `R`.
@@
 Our evaluation of the two programs revealed that `dfoliatR` excelled in identifying defoliation events on single trees, providing researchers with more consistent and biologically-justifiable results.
-Other key benefits are that `dfoliatR` provides easier control of suppression thresholds, additional output tables, high-quality and customizable graphics, and operates in the open-source `R` environment that is stable across computing platforms and under active development and maintenance by a large and growing community.
+Other key benefits are that `dfoliatR` provides easier control of suppression thresholds, additional output tables, high-quality and customizable graphics, and operates in the open-source `R` environment that is stable across computing platforms and is under active development and maintenance by a large and growing community.
 
@@
 It is probably not appropriate for insects with high-frequency, high–severity outbreaks, such as processionary caterpillars.
-Outbreak reconstructions of insects for which the interval is notably longer than typical outbreak duration, such as Douglas-fir tussock moth (*Orgyia pseudotsugata* (Lepidoptera: Erebidae)), are unlikely to be significantly affected.
+Outbreak reconstructions of insects for which the interval is notably longer than typical outbreak duration, such as Douglas-fir tussock moth, are unlikely to be significantly affected.
 Species for which outbreak duration or individual tree resilience to defoliation varies considerably, or which may chronically infest trees or sites, or alternate between chronic, outbreak, and minimal activity states such as conifer-feeding *Choristoneura*, present more complicated challenges, and the researcher should use discretion with the bridge option.
@@
 `dfoliatR` adds to the on-going open-source software development for dendrochronological methods [e.g., @Bunn2008; @Brewer2014; @Brewer2016].
-The `R` environment enables automation of analyses, allowing input/output processes to become routine, and empowering batch processing of large multi-site projects and sensitivity analyses.
+The `R` environment enables automation of analyses, allowing input/output processes to become routine, enables efficient sensitivity analyses, and empowers batch processing of large multi-site projects.
 It also facilitates additional statistical analyses, such as spectral analyses and superposed epoch analyses [e.g., @Malevich2018], with easy transfer from `dfoliatR` and `dplR` to other libraries in `R`.
@@
 We also thank Thomas Swetnam for his insights and encouragement, and David Frank for his support.
-The `dfoliatR` project is funded by the Agnes Haury Visiting Scholars Fellowship at the LTRR, the University of California, Berkeley, and the Western Wildlands Environmental Threat Center (WWETAC).
+The `dfoliatR` project is funded by the Agnes Haury Visiting Scholars Fellowship (to J.
+Axelson) at the LTRR, the University of California, Berkeley, and the Western Wildlands Environmental Threat Center (WWETAC).
 Additional support was provided by the LTRR and the United States Forest Service Rocky Mountain Research Station.

### 2020-05-21 paper_elsevier/paper_elsevier.tex
editing and re-running example script

@@
   \begin{abstract}
-  We present a new R package to provide dendroecologists with tools to infer, quantify, analyze, and visualize growth suppression events in tree rings.
-\texttt{dfoliatR} is based on the OUTBREAK program and builds on existing resources in the \texttt{R} and \texttt{dplR} computing environments.
+  We present a new \texttt{R} package to provide dendroecologists with tools to infer, quantify, analyze, and visualize growth suppression events in tree rings.
+\texttt{dfoliatR} is based on the OUTBREAK program and builds on existing resources in the \texttt{R} computing environment s the well-used \texttt{dplR} package.
 It is designed to aid research in the ecology of insect defoliation events and to reconstruct defoliator outbreak chronologies, but can be applied to other studies where host--non-host comparisons are useful.
-\texttt{dfoliatR} performs an indexing procedure to remove climatic signals in the host-tree series that are represented in the non-host chronologies, or other annually-resolved climate series.
+\texttt{dfoliatR} performs an indexing procedure to remove climatic signals in the host-tree series that are represented in the non-host chronology, or other annually-resolved climate series.
 It then infers defoliation events in individual trees based on user-specified thresholds.
@@
 \includegraphics{D:/Dropbox/dfoliatR_paper/paper_elsevier/Output/site-plot.pdf}
-\caption{\label{fig:fig-obr}Default graphics showing outbreak events at the DMJ site.
+\caption{\label{fig:fig-obr}Default graphic showing outbreak events at the DMJ site.
 Top panel shows the sample depth, the middle panel shows the mean GSI with inferred outbreak events filled, and the bottom panel shows the percent of trees defoliated, used to identify time periods of inferred outbreaks.}
@@
 Researchers can, and should, adjust the duration and magnitude parameters accordingly and critically evaluate the results, as insect species vary in the length of their outbreaks and the degree to which they can suppress tree growth.
-OUTBREAK provides two sets of default values, those for WSBW, which typically has lengthy outbreaks, and ones for Douglas-fir tussock moth (\emph{Orgyia pseudotsugata} (Lepidoptera: Tortricidae) of three years duration with -1.28 departure threshold that may be suitable for more eruptive species.
+OUTBREAK provides two sets of default values, those for WSBW, which typically has lengthy outbreaks, and ones for Douglas-fir tussock moth (\emph{Orgyia pseudotsugata} (Lepidoptera: Tortricidae)) of three years duration with -1.28 departure threshold that may be suitable for more eruptive species.
 
@@
 
-Examples of sequential defoliation events (bars), wherein two or three events occur separated by single years.
-In each case OUTBREAK recorded only one of the events, omitting those with a higher ``max\_defol'' year, or the year of the lowest NGSI value.
-\texttt{dfoliatR}, by contrast, recorded all of the events, including two sequential events on DMJ26 and EFK22, and three on FNL04.
-Events recorded by both programs have a darker color, while those recorded only by \texttt{dfoliatR} are lighter in color.
-Adding the bridging option in \texttt{dfoliatR} links these events (``bridge\_defol'') into single longer events that may more accurately capture the effects of long-term insect herbivory on individual trees.
-
 \begin{figure}
@@
 \caption{\label{fig:fig-bridge}Examples of identified defoliation events on individual trees.
-For each tree (DMJ26 and EFK22), OUTBREAK and \texttt{dfoliatR} identify most of the same events, but there is one added event (blue colored) that was omitted by OUTBREAK.
-These were omitted because they were sequential, separated by a single year with possitive normalized growth suppression index (NGSI).
-\texttt{dfoliatR} provides an option to bridge these events into single long events that may better represent the duration of defoliation given the insect and sites under consideration.}
+For each tree (DMJ26 and EFK22), OUTBREAK and \texttt{dfoliatR} identify most of the same events, but there is one added event (in blue) that was omitted by OUTBREAK.
+These were omitted because they were separated by a single year of positive normalized growth suppression index (NGSI) and OUTBREAK selected the one event with the lowest maximum departure value.
+\texttt{dfoliatR} provides an option to bridge these sequential events into single long events that may better represent the duration of defoliation given the insect and sites under consideration.}
 \end{figure}
@@
 Three parameters control whether a defoliation event constitutes an outbreak: the minimum number of trees available, the minimum number of trees recording defoliation, and the percent of trees recording defoliation.
-The first allows the researcher to make a judgement call as to the confidence ascribed to reduced sample depth toward the ends of their chronologies, thus compensating for the ``fading record problem'' (Swetnam et al., 1999).
+The first allows the researcher to make a judgment call as to the confidence ascribed to reduced sample depth toward the ends of their chronologies, thus compensating for the ``fading record problem'' (Swetnam et al., 1999).
 The second two parameters adjust the scale of defoliation considered to be an outbreak.
@@
 
-We detrended host data for both \texttt{dfoliatR} and OUTBREAK using ARSTAN {[}v6.1{]} with cubic smoothing splines (50\% frequency response on 100-150 year wavelengths depending on the site).
+We detrended host data for both \texttt{dfoliatR} and OUTBREAK using ARSTAN (version 6.1) with cubic smoothing splines (50\% frequency response on 100-150 year wavelengths depending on the site).
 In both \texttt{dfoliatR} and OUTBREAK we used event thresholds of -1.28 normalized indices, 8 years duration, and allowed for events at the end of series in seven of eight sites that had known outbreaks at the time of sampling.
@@
 
-Across the 43,280 ring-width indices from 222 trees included in our evaluation, we found that \texttt{dfoliatR} and OUTBREAK compute identical growth suppression indices at 0.000 precision.
-We expected this outcome because both programs employ Equation 1.
-At the tree-level, the programs identified 11,530 years of defoliation.
+Across the 43,280 ring-width indices from 222 trees included in our evaluation, we found that \texttt{dfoliatR} and OUTBREAK compute identical growth suppression indices at 0.00 precision.
+We expected this outcome because both programs apply Equation 1 to calcuate disturbance indices.
+At the tree-level, the programs identified 11,530 total index years with defoliation.
 The programs agreed on 97.9\% of the years, leaving 927 ``difference'' years in which only one program identified defoliation on an individual tree.
@@
 In OUTBREAK, the option appears while changing the duration parameter (option 3).
-When selected, OUTBREAK will include any sequences of negative indices at the either the beginning or the end of each tree series as a defoliation event, without consideration of either duration or magnitude thresholds.
+When selected, OUTBREAK will include any sequences of negative indices at the beginning and the end of each tree series as a defoliation event, without consideration of either duration or magnitude thresholds.
 In \texttt{dfoliatR}, the duration threshold is omitted and the magnitude threshold is retained in series-end-events.
@@
 While inspecting these differences, we added an option to \texttt{defoliate\_trees()} that would ``bridge'' between sequential events (that each surpass the magnitude and duration thresholds) into single, long events.
-We felt that this was ecologically justified, especially for studies of WSBW, because outbreaks are known to be of long duration and tree-ring reconstructions have shown that outbreaks may persist 30-50 at the site level.
+We felt that this was ecologically justified, especially for studies of WSBW, because outbreaks are known to be of long duration and tree-ring reconstructions have shown that outbreaks may persist for as long as 30-50 years at the site level.
 \item
@@
 At the site level, OUTBREAK and \texttt{dfoliatR} produce similar time series of percent trees defoliated (Figure \ref{fig:fig-df-obr}), which forms the basis for inferring outbreak occurrence, intensity, and duration.
-In nearly all site comparison, \texttt{dfoliatR} included either more events or longer durations of inferred outbreaks.
+In nearly all site-level comparisons, \texttt{dfoliatR} included either more events or it inferred a longer duration outbreak.
 These differences arise from the inclusion of tree-level events by \texttt{dfoliatR} that were omitted by OUTBREAK (see note on sequential events above).
@@
 \begin{Highlighting}[]
-\KeywordTok{defol_stats}\NormalTok{(dmj_defol)}
-\KeywordTok{View}\NormalTok{(dmj_defol)}
+\NormalTok{dmj_defol_stats <-}\StringTok{ }\KeywordTok{defol_stats}\NormalTok{(dmj_defol)}
+\KeywordTok{View}\NormalTok{(dmj_defol_stats)}
 \end{Highlighting}
@@
 Input parameters to \texttt{outbreak()} include ``filter\_min\_series'' to control the chronology cut-off points with regard to sample depth, ``filter\_min\_defol'' and ``filter\_perc'' to control the minimum number and percent of trees recording a defoliation event in a given year.
-\texttt{Outbreak()} produces a new data frame with eight variables: ``year'', ``num\_defol'', ``percent\_defol'', ``num\_max\_defol'', ``mean\_gsi'', ``mean\_ngsi'', and ``outbreak\_status.'' All of these variables are populated regardless of an inferred outbreak event, providing a continuous disturbance chronology.
+\texttt{outbreak()} produces a new data frame with eight variables: ``year'', ``num\_defol'', ``percent\_defol'', ``num\_max\_defol'', ``mean\_gsi'', ``mean\_ngsi'', and ``outbreak\_status.'' All of these variables are populated regardless of an inferred outbreak event, providing a continuous outbreak reconstruction.
 The ``num\_max\_defol'' variable counts the number of trees recording their maximum defoliation in a given year.
@@
 \toprule
-start & end & duration & num\_trees\_start & perc\_trees\_start & peak\_outbreak\_year & num\_trees\_outbreak & peak\_defol\_year & min\_gsi & mean\_ngsi\\
+start & end & duration & num\_trees\_start & perc\_trees\_start & num\_trees\_outbreak & peak\_outbreak\_year & peak\_defol\_year & min\_gsi & min\_ngsi\\
 \midrule
-1680 & 1699 & 20 & 1 & 33.3 & 1690 & 3 & 1692 & 0.121 & -2.407\\
-1753 & 1769 & 17 & 6 & 46.2 & 1754 & 7 & 1755 & 0.343 & -1.611\\
-1825 & 1840 & 16 & 11 & 78.6 & 1831 & 12 & 1826 & 0.500 & -1.304\\
-1849 & 1865 & 17 & 7 & 46.7 & 1852 & 13 & 1853 & 0.252 & -1.994\\
-1881 & 1895 & 15 & 8 & 53.3 & 1886 & 14 & 1885 & 0.262 & -1.945\\
+1680 & 1699 & 20 & 1 & 33.3 & 3 & 1690 & 1692 & 0.121 & -2.407\\
+1753 & 1769 & 17 & 6 & 46.2 & 7 & 1754 & 1755 & 0.343 & -1.611\\
+1825 & 1840 & 16 & 11 & 78.6 & 12 & 1831 & 1826 & 0.500 & -1.304\\
+1849 & 1865 & 17 & 7 & 46.7 & 13 & 1852 & 1853 & 0.252 & -1.994\\
+1881 & 1895 & 15 & 8 & 53.3 & 14 & 1886 & 1885 & 0.262 & -1.945\\
 \addlinespace
-1959 & 1970 & 12 & 7 & 41.2 & 1960 & 15 & 1965 & 0.328 & -1.830\\
-1987 & 1996 & 10 & 9 & 52.9 & 1989 & 15 & 1995 & 0.378 & -1.640\\
+1959 & 1970 & 12 & 7 & 41.2 & 15 & 1960 & 1965 & 0.328 & -1.830\\
+1987 & 1996 & 10 & 9 & 52.9 & 15 & 1989 & 1995 & 0.378 & -1.640\\
 \bottomrule
@@
 
-Saving the results of \texttt{outbreak\_stats()} (the \texttt{dmj\_obr\_stats} object above) provides an array of options for assessing metrics of the insect defoliation regime.
+Saving the results of \texttt{outbreak\_stats()} (the \texttt{dmj\_obr\_stats} object above) provides an array of options for assessing metrics of the insect outbreak regime.
 Here we use the first year of each outbreak event to calculate the average duration of years between outbreaks, via the \texttt{diff()} function in \texttt{R}.
@@
 Our evaluation of the two programs revealed that \texttt{dfoliatR} excelled in identifying defoliation events on single trees, providing researchers with more consistent and biologically-justifiable results.
-Other key benefits are that \texttt{dfoliatR} provides easier control of suppression thresholds, additional output tables, high-quality and customizable graphics, and operates in the open-source \texttt{R} environment that is stable across computing platforms and under active development and maintenance by a large and growing community.
+Other key benefits are that \texttt{dfoliatR} provides easier control of suppression thresholds, additional output tables, high-quality and customizable graphics, and operates in the open-source \texttt{R} environment that is stable across computing platforms and is under active development and maintenance by a large and growing community.
 
@@
 It is probably not appropriate for insects with high-frequency, high--severity outbreaks, such as processionary caterpillars.
-Outbreak reconstructions of insects for which the interval is notably longer than typical outbreak duration, such as Douglas-fir tussock moth (\emph{Orgyia pseudotsugata} (Lepidoptera: Erebidae)), are unlikely to be significantly affected.
+Outbreak reconstructions of insects for which the interval is notably longer than typical outbreak duration, such as Douglas-fir tussock moth, are unlikely to be significantly affected.
 Species for which outbreak duration or individual tree resilience to defoliation varies considerably, or which may chronically infest trees or sites, or alternate between chronic, outbreak, and minimal activity states such as conifer-feeding \emph{Choristoneura}, present more complicated challenges, and the researcher should use discretion with the bridge option.
@@
 \texttt{dfoliatR} adds to the on-going open-source software development for dendrochronological methods (e.g., Bunn, 2008; Brewer, 2014; Brewer and Guiterman, 2016).
-The \texttt{R} environment enables automation of analyses, allowing input/output processes to become routine, and empowering batch processing of large multi-site projects and sensitivity analyses.
+The \texttt{R} environment enables automation of analyses, allowing input/output processes to become routine, enables efficient sensitivity analyses, and empowers batch processing of large multi-site projects.
 It also facilitates additional statistical analyses, such as spectral analyses and superposed epoch analyses (e.g., Malevich et al., 2018), with easy transfer from \texttt{dfoliatR} and \texttt{dplR} to other libraries in \texttt{R}.
@@
 We also thank Thomas Swetnam for his insights and encouragement, and David Frank for his support.
-The \texttt{dfoliatR} project is funded by the Agnes Haury Visiting Scholars Fellowship at the LTRR, the University of California, Berkeley, and the Western Wildlands Environmental Threat Center (WWETAC).
+The \texttt{dfoliatR} project is funded by the Agnes Haury Visiting Scholars Fellowship (to J.
+Axelson) at the LTRR, the University of California, Berkeley, and the Western Wildlands Environmental Threat Center (WWETAC).
 Additional support was provided by the LTRR and the United States Forest Service Rocky Mountain Research Station.

### 2020-05-21 paper_elsevier/R/Example_Usage.R

@@ -1,11 +1,10 @@
 ## Create tables and graphics for dfoliatR manuscript
 
-devtools::install_github("chguiterman/dfoliatR@dev", force=TRUE)
+# devtools::install_github("chguiterman/dfoliatR", force=TRUE)
 library(dfoliatR)
 library(ggplot2)
-library(patchwork)
+library(here)
 
-out_pth <- "paper_elsevier/Output/" 
 # Tree-level outputs ------------------------------------------------------
 
 data("dmj_h")
@@ -20,24 +19,13 @@ dmj_defol <- defoliate_trees(host_tree = dmj_h,
                              list_output = FALSE)
 
 dmj_stats <- defol_stats(dmj_defol)
-write.csv(dmj_stats, paste0(out_pth, "defol-stats.csv"), row.names = FALSE)
+write.csv(dmj_stats, here("paper_elsevier", "Output", "defol-stats.csv"), row.names = FALSE)
 
 plot_defol(dmj_defol)
-ggsave(paste0(out_pth, "tree-plot-default.pdf"), width=5.75, dpi=300, units="in")
+ggsave(here("paper_elsevier", "Output", "tree-plot-default.pdf"), width=5.75, dpi=300, units="in")
 
-p1 <- plot_defol(dmj_defol) +
-  theme(legend.position = c(0.5, -0.15),
-        legend.direction = "horizontal",
-        legend.background = element_blank(),
-        legend.text = element_text(size=8),
-        plot.margin = unit(c(0, 0, 15, 0), units = "points"),
-        plot.background = element_blank())
-  
-p2 <- p1 + 
+plot_defol(dmj_defol) +
   scale_color_manual(values = c("red", "orange", "purple")) 
-
-p1 / p2 + plot_layout(guides = "keep") + plot_annotation(tag_levels = "A")
-ggsave(paste0(out_pth, "tree-plot-2panel.pdf"), width=5.75, dpi=300, units="in")
 
 # Site-level outputs ------------------------------------------------------
 
@@ -47,7 +35,7 @@ dmj_obr <- outbreak(dmj_defol,
                     filter_perc = 25)
 
 plot_outbreak(dmj_obr, disp_index = "GSI")
-ggsave(paste0(out_pth, "site-plot.pdf"), width=5.75, dpi=300, units="in")
[…]

### 2020-05-21 paper_elsevier/paper_elsevier.Rmd
more Ann-Jodi comments. Almost ready to submit

@@
   - code: d
-    address: "Dept of Environmental Science, Policy & Management, University of California, Berkeley, Berkeley, CA 94720"
+    address: "Dept of Environmental Science, Policy & Management, University of California, Berkeley, 130 Mulford Hall #3114, Berkeley, CA 94720"
 footnote:
@@
  We present a new `R` package to provide dendroecologists with tools to infer, quantify, analyze, and visualize growth suppression events in tree rings.
-`dfoliatR` is based on the OUTBREAK program and builds on existing resources in the `R` computing environment s the well-used `dplR` package.
+`dfoliatR` is based on the OUTBREAK program and builds on existing resources in the `R` computing environment and the well-used `dplR` package.
 It is designed to aid research in the ecology of insect defoliation events and to reconstruct defoliator outbreak chronologies, but can be applied to other studies where host--non-host comparisons are useful.
@@
 Stemming from the `dplR` library [@Bunn2008] that enables `R` users to read and write an array of tree-ring data formats, standardize ring width series, build and evaluate chronologies, and perform quality control (to name a few), one can now also measure ring widths from scanned images of prepared samples [@Shi2019; @Lara2015], conduct and check crossdating [@Bunn2010], analyze sub-annual anatomical features [@Campelo2016; @Rathgeber2011], and perform many analytical tests [@Jevsenak2018; @Zang2015].
-Tools for assessing stand dynamics and disturbance analyses are under rapid development, with new packages for assessing growth and release events [`TRADER`: @Altman2014], metrics of growth resilience [`pointRes`: @VanderMaaten-Theunissen2015], and fire history [`burnr`: @Malevich2018].
+Tools for assessing stand dynamics and disturbance analyses are under rapid development, with new packages for assessing release events [`TRADER`: @Altman2014], metrics of growth resilience [`pointRes`: @VanderMaaten-Theunissen2015], and fire history [`burnr`: @Malevich2018].
 The key objective of `dfoliatR` is to provide tools to identify and analyze insect defoliation and outbreak events by building on the methods employed by OUTBREAK.
@@
 (ref:defolCap) Default `dfoliatR` graphic for individual trees, produced by the `plot_defol()` function.
-The series names are listed on the Y-axis, dotted lines represent the series length for each tree, and colored segments show periods of defoliation.
+The series names from the DMJ site [@Ryerson2003] are listed on the Y-axis, dotted lines represent the series length for each tree, and colored segments show periods of defoliation.
 The colors of defoliation segments represent its severity, for which users can define cut-off values to determine severe--moderate--minor defoliation intensities.
@@
 
-(ref:obrCap) Default graphic showing outbreak events at the DMJ site.
+(ref:obrCap) Default `dfoliatR` graphic for site-level outbreak events, produced by the `plot_outbreak()` function.
+These plots summarize the data in Figure \@ref(fig:fig-defol) for the DMJ site.
 Top panel shows the sample depth, the middle panel shows the mean GSI with inferred outbreak events filled, and the bottom panel shows the percent of trees defoliated, used to identify time periods of inferred outbreaks.
@@
 
-Event duration is assessed by examining sequences of negative NGSI (that exceed the magnitude threshold) before and after the year of maximum departure.
+Event duration is assessed by examining sequences of negative NGSI (for which one or more values exceeded the magnitude threshold) before and after the year of maximum departure.
 Each defoliation event is allowed one single-year positive excursion on each side of the year of maximum departure.
@@
 This should be used if the user has direct knowledge of defoliation at the site during the sampling campaign.
-The advantage of allowing potentially short, series-end events is that it allows the current event to be included in return-interval estimates, and can aid in identifying the start-year for the current defoliation event or outbreak.
+The advantage of allowing potentially short, series-end events is that it allows a current event to be included in return-interval estimates, and can aid in identifying the start-year for the current defoliation event or outbreak.
 
@@
 Our tests used standardized ring-width data from eight host-tree sites spanning the range of WSBW.
-The sites were sampled in British Columbia [@Axelson2015], Wyoming [@NADEF2018], Colorado, and New Mexico [@Swetnam1993; @Ryerson2003].
+The sites were sampled in British Columbia [@Axelson2015], Wyoming [@NADEF2018], Colorado [@Ryerson2003], and New Mexico [@Swetnam1993].
 These host data were compared to non-host chronologies from the original studies, but we made no effort here to replicate the reconstructions or analyses of those studies.
@@
 (ref:tbl-defolCap) Tree-level tabular output provided by the `defol_stats()` function for the DMJ example site.
+Note that these calculations exclude the ongoing "series-end" events as selected in `defoliate_trees()`.
 
@@
 A summary table of the inferred outbreak events is generated by the `outbreak_stats()` function (Table \@ref(tab:tbl-obr)).
+The table provides a range of summary statistics, including the start and end years of each outbreak event, along with the corresponding duration, the number  and percent of trees in defoliation at the start of the event ("n_df_start" and "perc_df_start", respectively), the maximum number of trees recording the outbreak event during a single year ("max_df_obr"), the year corresponding to that peak ("yr_max_df"), the year with the lowest mean NGSI during the event ("yr_min_ngsi"), and the minima of mean GSI and mean NGSI indices during the event.
 
@@
 
-(ref:tbl-obrCap) Site-level summary statistics for inferred outbreak events provided with the `outbreak_stats()` function for the DMJ example site.
+(ref:tbl-obrCap) Site-level summary statistics for inferred outbreak events as generated by the `outbreak_stats()` function for the DMJ example site.
+Note that missing ("NA") values ocurr for certain statistics because the last outbreak event was defined earlier as ongoing (i.e.
+we set `series_end_event = TRUE` in the call to `defoliate_trees()`).
 
@@
 
-The output table from `outbreak_stats()` shows the start and end years of each event, along with their corresponding duration, the number and percent of trees in defoliation at the start of the event, the the number of trees recording the outbreak event (in part or full), the minimums of the mean GSI and NGSI indices, and two points of "peak defoliation." The first, "peak_outbreak_year" is the year in which the greatest number of trees are recording the outbreak, and the second, "peak_defol_year" is the year with the lowest average NGSI.
-
 Saving the results of `outbreak_stats()` (the `dmj_obr_stats` object above) provides an array of options for assessing metrics of the insect outbreak regime.
-Here we use the first year of each outbreak event to calculate the average duration of years between outbreaks, via the `diff()` function in `R`.
-The average of those differences, calculated via `mean()` is the mean return interval.
+For example, taking the first year of each outbreak event, we can calculate the duration of years between outbreaks, via the `diff()` function in `R`.
+The average of those differences, calculated via `mean()` is the mean return interval of reconstructed outbreak events at the DMJ site.
 
@@
 
-`dfoliatR` adds a new option for dendroentimology to combine, or "bridge," sequential tree-level defoliation events into single events.
+`dfoliatR` adds a new option for dendroentomology to combine, or "bridge," sequential tree-level defoliation events into single events.
 In practice, we suggest that researchers carefully evaluate if bridging is ecologically applicable to study situations and insect ecology, and to carefully explore the data before deciding whether or not to use this option.

### 2020-05-21 paper_elsevier/paper_elsevier.tex
more Ann-Jodi comments. Almost ready to submit

@@
     \address[c]{U.S. Forest Service, Rocky Mountain Research Station, 1215 E Lowell St.~Box 210045, Tucson, AZ, 85721}
-    \address[d]{Dept of Environmental Science, Policy \& Management, University of California, Berkeley, Berkeley, CA 94720}
+    \address[d]{Dept of Environmental Science, Policy \& Management, University of California, Berkeley, 130 Mulford Hall \#3114, Berkeley, CA 94720}
       \cortext[1]{Corresponding Author}
@@
   We present a new \texttt{R} package to provide dendroecologists with tools to infer, quantify, analyze, and visualize growth suppression events in tree rings.
-\texttt{dfoliatR} is based on the OUTBREAK program and builds on existing resources in the \texttt{R} computing environment s the well-used \texttt{dplR} package.
+\texttt{dfoliatR} is based on the OUTBREAK program and builds on existing resources in the \texttt{R} computing environment and the well-used \texttt{dplR} package.
 It is designed to aid research in the ecology of insect defoliation events and to reconstruct defoliator outbreak chronologies, but can be applied to other studies where host--non-host comparisons are useful.
@@
 Stemming from the \texttt{dplR} library (Bunn, 2008) that enables \texttt{R} users to read and write an array of tree-ring data formats, standardize ring width series, build and evaluate chronologies, and perform quality control (to name a few), one can now also measure ring widths from scanned images of prepared samples (Lara et al., 2015; Shi et al., 2019), conduct and check crossdating (Bunn, 2010), analyze sub-annual anatomical features (Rathgeber et al., 2011; Campelo et al., 2016), and perform many analytical tests (Zang and Biondi, 2015; Jevšenak and Levanič, 2018).
-Tools for assessing stand dynamics and disturbance analyses are under rapid development, with new packages for assessing growth and release events (\texttt{TRADER}: Altman et al., 2014), metrics of growth resilience (\texttt{pointRes}: van der Maaten-Theunissen et al., 2015), and fire history (\texttt{burnr}: Malevich et al., 2018).
+Tools for assessing stand dynamics and disturbance analyses are under rapid development, with new packages for assessing release events (\texttt{TRADER}: Altman et al., 2014), metrics of growth resilience (\texttt{pointRes}: van der Maaten-Theunissen et al., 2015), and fire history (\texttt{burnr}: Malevich et al., 2018).
 The key objective of \texttt{dfoliatR} is to provide tools to identify and analyze insect defoliation and outbreak events by building on the methods employed by OUTBREAK.
@@
 \caption{\label{fig:fig-defol}Default \texttt{dfoliatR} graphic for individual trees, produced by the \texttt{plot\_defol()} function.
-The series names are listed on the Y-axis, dotted lines represent the series length for each tree, and colored segments show periods of defoliation.
+The series names from the DMJ site (Ryerson et al., 2003) are listed on the Y-axis, dotted lines represent the series length for each tree, and colored segments show periods of defoliation.
 The colors of defoliation segments represent its severity, for which users can define cut-off values to determine severe--moderate--minor defoliation intensities.
@@
 \includegraphics{D:/Dropbox/dfoliatR_paper/paper_elsevier/Output/site-plot.pdf}
-\caption{\label{fig:fig-obr}Default graphic showing outbreak events at the DMJ site.
+\caption{\label{fig:fig-obr}Default \texttt{dfoliatR} graphic for site-level outbreak events, produced by the \texttt{plot\_outbreak()} function.
+These plots summarize the data in Figure \ref{fig:fig-defol} for the DMJ site.
 Top panel shows the sample depth, the middle panel shows the mean GSI with inferred outbreak events filled, and the bottom panel shows the percent of trees defoliated, used to identify time periods of inferred outbreaks.}
@@
 Our tests used standardized ring-width data from eight host-tree sites spanning the range of WSBW.
-The sites were sampled in British Columbia (Axelson et al., 2015), Wyoming (Axelson et al., 2018), Colorado, and New Mexico (Swetnam and Lynch, 1993; Ryerson et al., 2003).
+The sites were sampled in British Columbia (Axelson et al., 2015), Wyoming (Axelson et al., 2018), Colorado (Ryerson et al., 2003), and New Mexico (Swetnam and Lynch, 1993).
 These host data were compared to non-host chronologies from the original studies, but we made no effort here to replicate the reconstructions or analyses of those studies.
@@
 
-\caption{\label{tab:tbl-defol}Tree-level tabular output provided by the \texttt{defol\_stats()} function for the DMJ example site.}
+\caption{\label{tab:tbl-defol}Tree-level tabular output provided by the \texttt{defol\_stats()} function for the DMJ example site.
+Note that these calculations exclude the ongoing ``series-end'' events as selected in \texttt{defoliate\_trees()}.}
 \centering
@@
 \toprule
-series & first & last & years & num\_events & tot\_years & mean\_duration\\
+series & first & last & years & n\_events & tot\_years & mean\_duration\\
 \midrule
-DMJ01 & 1810 & 1996 & 187 & 4 & 50 & 12\\
-DMJ02 & 1750 & 1996 & 247 & 6 & 75 & 12\\
-DMJ03 & 1830 & 1996 & 167 & 4 & 42 & 10\\
-DMJ04 & 1720 & 1996 & 277 & 8 & 90 & 11\\
-DMJ06 & 1700 & 1996 & 297 & 6 & 81 & 14\\
+DMJ01 & 1810 & 1996 & 187 & 4 & 40 & 13\\
+DMJ02 & 1750 & 1996 & 247 & 6 & 66 & 13\\
+DMJ03 & 1830 & 1996 & 167 & 4 & 32 & 11\\
+DMJ04 & 1720 & 1996 & 277 & 8 & 82 & 12\\
+DMJ06 & 1700 & 1996 & 297 & 6 & 70 & 14\\
 \addlinespace
-DMJ07 & 1710 & 1996 & 287 & 7 & 96 & 14\\
-DMJ11 & 1900 & 1997 & 98 & 2 & 19 & 10\\
-DMJ14 & 1675 & 1996 & 322 & 9 & 112 & 12\\
+DMJ07 & 1710 & 1996 & 287 & 7 & 86 & 14\\
+DMJ11 & 1900 & 1997 & 98 & 2 & 10 & 10\\
+DMJ14 & 1675 & 1996 & 322 & 9 & 104 & 13\\
 DMJ15 & 1730 & 1996 & 267 & 4 & 64 & 16\\
-DMJ16 & 1746 & 1996 & 251 & 7 & 86 & 12\\
+DMJ16 & 1746 & 1996 & 251 & 7 & 78 & 13\\
 \addlinespace
-DMJ17 & 1733 & 1996 & 264 & 6 & 85 & 14\\
-DMJ22 & 1720 & 1996 & 277 & 5 & 56 & 11\\
-DMJ23 & 1675 & 1997 & 323 & 8 & 96 & 12\\
-DMJ24 & 1895 & 1996 & 102 & 3 & 28 & 9\\
+DMJ17 & 1733 & 1996 & 264 & 6 & 75 & 15\\
+DMJ22 & 1720 & 1996 & 277 & 5 & 47 & 12\\
+DMJ23 & 1675 & 1997 & 323 & 8 & 87 & 12\\
+DMJ24 & 1895 & 1996 & 102 & 3 & 19 & 10\\
 DMJ25 & 1680 & 1996 & 317 & 5 & 70 & 14\\
 \addlinespace
-DMJ26 & 1700 & 1996 & 297 & 6 & 79 & 13\\
-DMJ27 & 1710 & 1996 & 287 & 3 & 34 & 11\\
+DMJ26 & 1700 & 1996 & 297 & 6 & 70 & 14\\
+DMJ27 & 1710 & 1996 & 287 & 3 & 26 & 13\\
 \bottomrule
@@
 A summary table of the inferred outbreak events is generated by the \texttt{outbreak\_stats()} function (Table \ref{tab:tbl-obr}).
+The table provides a range of summary statistics, including the start and end years of each outbreak event, along with the corresponding duration, the number and percent of trees in defoliation at the start of the event (``n\_df\_start'' and ``perc\_df\_start'', respectively), the maximum number of trees recording the outbreak event during a single year (``max\_df\_obr''), the year corresponding to that peak (``yr\_max\_df''), the year with the lowest mean NGSI during the event (``yr\_min\_ngsi''), and the minima of mean GSI and mean NGSI indices during the event.
 
@@
 
-\caption{\label{tab:tbl-obr}Site-level summary statistics for inferred outbreak events provided with the \texttt{outbreak\_stats()} function for the DMJ example site.}
+\caption{\label{tab:tbl-obr}Site-level summary statistics for inferred outbreak events as generated by the \texttt{outbreak\_stats()} function for the DMJ example site.
+Note that missing (``NA'') values ocurr for certain statistics because the last outbreak event was defined earlier as ongoing (i.e.~we set \texttt{series\_end\_event\ =\ TRUE} in the call to \texttt{defoliate\_trees()}).}
 \centering
@@
 \toprule
-start & end & duration & num\_trees\_start & perc\_trees\_start & num\_trees\_outbreak & peak\_outbreak\_year & peak\_defol\_year & min\_gsi & min\_ngsi\\
+start & end & duration & n\_df\_start & perc\_df\_start & max\_df\_obr & yr\_max\_df & yr\_min\_ngsi & min\_gsi & min\_ngsi\\
 \midrule
@@
 1959 & 1970 & 12 & 7 & 41.2 & 15 & 1960 & 1965 & 0.328 & -1.830\\
-1987 & 1996 & 10 & 9 & 52.9 & 15 & 1989 & 1995 & 0.378 & -1.640\\
+1987 & NA & NA & 9 & 52.9 & 15 & NA & NA & 0.378 & -1.640\\
 \bottomrule
@@
 
-The output table from \texttt{outbreak\_stats()} shows the start and end years of each event, along with their corresponding duration, the number and percent of trees in defoliation at the start of the event, the the number of trees recording the outbreak event (in part or full), the minimums of the mean GSI and NGSI indices, and two points of ``peak defoliation.'' The first, ``peak\_outbreak\_year'' is the year in which the greatest number of trees are recording the outbreak, and the second, ``peak\_defol\_year'' is the year with the lowest average NGSI.
-
 Saving the results of \texttt{outbreak\_stats()} (the \texttt{dmj\_obr\_stats} object above) provides an array of options for assessing metrics of the insect outbreak regime.
-Here we use the first year of each outbreak event to calculate the average duration of years between outbreaks, via the \texttt{diff()} function in \texttt{R}.
-The average of those differences, calculated via \texttt{mean()} is the mean return interval.
+For example, taking the first year of each outbreak event, we can calculate the average duration of years between outbreaks, via the \texttt{diff()} function in \texttt{R}.
+The average of those differences, calculated via \texttt{mean()} is the mean return interval of reconstructed outbreak events at the DMJ site.
 
@@
 
-\texttt{dfoliatR} adds a new option for dendroentimology to combine, or ``bridge,'' sequential tree-level defoliation events into single events.
+\texttt{dfoliatR} adds a new option for dendroentomology to combine, or ``bridge,'' sequential tree-level defoliation events into single events.
 In practice, we suggest that researchers carefully evaluate if bridging is ecologically applicable to study situations and insect ecology, and to carefully explore the data before deciding whether or not to use this option.

### 2020-05-21 dfoliatR_ms.Rmd
clean house

--- 2020-05-21 paper_elsevier/paper_elsevier.Rmd
@@
     text: "Corresponding Author"
-#   - code: 2
-#     text: "Laboratory of Tree-Ring Research, University of Arizona, 1215 E Lowell St. Box 210045, Tucson, AZ, 85721"
-#   - code: 3
-#     text: "U.S. Forest Service, Rocky Mountain Research Station, 1215 E Lowell St. Box 210045, Tucson, AZ, 85721"
-#   - code: 4
-#     text: "Dept of Environmental Science, Policy & Management, University of California, Berkeley, Berkeley, CA 94720"
 abstract: |
@@
 #numbersections: true
-# csl: ecology.csl
-# csl: elsevier-harvard.csl
-# csl: elsevier-vancouver.csl
 csl: dendrochronologia.csl
@@
 ```{r fig-defol, echo=FALSE, fig.cap='(ref:defolCap)'}
-include_graphics(here("paper_elsevier", "Output", "tree-plot-default.pdf"))
+include_graphics(here( "Output", "tree-plot-default.pdf"))
 ```
@@
 ```{r fig-obr, echo=FALSE, fig.cap='(ref:obrCap)'}
-include_graphics(here("paper_elsevier", "Output", "site-plot.pdf"))
+include_graphics(here( "Output", "site-plot.pdf"))
 ```
@@
 ```{r fig-bridge, echo=FALSE, fig.cap='(ref:BridgeCap)'}
-include_graphics(here("paper_elsevier", "Output", "bridging_color.pdf"))
+include_graphics(here( "Output", "bridging_color.pdf"))
 ```
@@
 ``` {r fig-df-obr, echo=FALSE, fig.cap='(ref:df-obrCap)'}
-include_graphics(here("paper_elsevier", "Output", "plot-sites-df-obr.pdf"))
+include_graphics(here( "Output", "plot-sites-df-obr.pdf"))
 ```
@@
 ```{r tbl-defol, echo=FALSE}
-tree_dat <- read.csv(here("paper_elsevier", "Output", "defol-stats.csv"))
+tree_dat <- read.csv(here( "Output", "defol-stats.csv"))
 kable(tree_dat, caption = "(ref:tbl-defolCap)", booktabs = TRUE) %>% 
@@
 ```{r tbl-obr, echo=FALSE}
-obr_dat <- read.csv(here("paper_elsevier", "Output", "obr-stats.csv"))
+obr_dat <- read.csv(here( "Output", "obr-stats.csv"))
 kable(obr_dat, format = "latex", caption = "(ref:tbl-obrCap)", booktabs = TRUE) %>% 

### 2020-05-21 paper_elsevier/R/Compare_OUTBREAK_dfoliatR.R

@@ -1,171 +1,0 @@
-# Compare results from OUTBREAK to dfoliatR
-
-library(dfoliatR)
-library(dplR)
-library(dplyr)
-library(purrr)
-library(tibble)
-library(tidyr)
-library(ggplot2)
-library(cowplot)
-# Load read_OUTBREAK function
-devtools::source_gist("https://gist.github.com/chguiterman/c88e9601dd085eb4215f0dc133a0e21c", sha1 = "fdb3cfefcd7e6ed3f7df3e20057c815bbd91dd0c")
-
-data_pth <- "paper_elsevier/Data/"
-
-## Master metadata file
-meta_in <- read.csv(paste0(data_pth, "Meta.csv"))
-
-## Helper function
-rm_sample_depth <- function(x){
-  if (any(names(x) %in% "samp.depth"))
-    df <- select(x, - samp.depth)
-  else df <- x
-  return(df)
-}
-
-
-# OUTBREAK data -----------------------------------------------------------
-
-outbr_results <- meta_in %>% 
-  transmute(site, 
-            outbr_cor = map(outbr_corrected, ~ read.compact(paste0(data_pth, .))),
-            outbr_col = map(outbr_site, ~ read.delim(paste0(data_pth, .))),
-            outbr_out = map2(outbr_file, insect_code, 
-                             ~ read_OUTREAK(fname = paste0(data_pth, .x),
-                                            level = "tree",
-                                            insect_code = .y)))
-
-# Tree-level table
-outbr_defol <- outbr_results %>% 
-  transmute(site,
-            outbr_cor = map(outbr_cor, ~ .[] %>%  
-                              select(! ends_with("NormCorr")) %>%
-                              rownames_to_column("year") %>% 
-                              pivot_longer(-year, names_to = "series", values_to = "ncor"))
-  ) %>% 
-  unnest(cols = outbr_cor) %>% 
-  na.omit() %>% 
-  arrange(series, year)
-
-outbr_out <- outbr_results %>% 
-  select(site, outbr_out) %>% 
-  unnest(cols = outbr_out)
-
-outbr_defol <- merge(outbr_defol, outbr_out, by = c("site", "series", "year"))
-
-# Site-level table
-outbr_obr <- outbr_results %>% 
-  transmute(site,
-            outbr_col = map(outbr_col, ~ .[] %>% 
-                              select(Year, 
-                                     contains("Trees", ignore.case = TRUE), 
[…]

### 2020-05-21 paper_elsevier/R/bridging.R

@@ -1,87 +1,0 @@
-# Show bridging
-## CHG 2020-05-07
-
-library(dfoliatR)
-library(dplR)
-library(dplyr)
-library(purrr)
-library(tibble)
-library(tidyr)
-library(ggplot2)
-library(prismatic)
-library(paletteer)
-library(here)
-
-## Master metadata file
-meta_in <- read.csv(here("paper_elsevier", "Data", "Meta.csv"))
-# file of output from the comparison -- has the OBR results
-obr_results <- read.csv(here("paper_elsevier", "Output", "Tree-comps-data.csv")) %>% 
-  select(site, year, series, obr_type)
-
-## Helper function
-rm_sample_depth <- function(x){
-  if (any(names(x) %in% "samp.depth"))
-    df <- select(x, - samp.depth)
-  else df <- x
-  return(df)
-}
-
-df_run <- meta_in %>% 
-  transmute(site,
-            host = map(host_file, ~ read.compact(here("paper_elsevier", "Data", .))),
-            nonhost = map(nonhost_file, ~ read.crn(here("paper_elsevier", "Data", .)) %>% 
-                            rm_sample_depth()),
-            defol = pmap(list(host, nonhost, ongoing_defol),
-                         ~ defoliate_trees(host_tree = ..1, 
-                                           nonhost_chron = ..2,
-                                           series_end_event = ..3,
-                                           duration_years = 8,
-                                           max_reduction = -1.28,
-                                           bridge_events = TRUE,  ## Add bridges
-                                           list_output = FALSE))
-  )
-
-df_defol <- df_run %>% 
-  select(site, defol) %>% 
-  unnest(cols = "defol") %>% 
-  inner_join(obr_results)
-
-
-# Selected bridge example -------------------------------------------------
-
-# fnl04 <- df_defol %>% 
-#   filter(series == "FNL04",
-#          year %in% 1881:1925)
-efk22 <- df_defol %>% 
-  filter(series == "EFK22",
-         year > 1850)
-dmj26 <- df_defol %>% 
-  filter(series == "DMJ26",
-        year %in% 1750:1950)
-
-# comps <- rbind(fnl04, efk22, dmj26)
-comps <- rbind(efk22, dmj26)
-
-obr_events <- comps %>% 
[…]

### 2020-05-21 dfoliatR_ms.tex

--- 2020-05-21 paper_elsevier/paper_elsevier.tex
@@
 \centering
-\includegraphics{D:/Dropbox/dfoliatR_paper/paper_elsevier/Output/tree-plot-default.pdf}
+\includegraphics{D:/Dropbox/dfoliatR_paper/Output/tree-plot-default.pdf}
 \caption{\label{fig:fig-defol}Default \texttt{dfoliatR} graphic for individual trees, produced by the \texttt{plot\_defol()} function.
@@
 \centering
-\includegraphics{D:/Dropbox/dfoliatR_paper/paper_elsevier/Output/site-plot.pdf}
+\includegraphics{D:/Dropbox/dfoliatR_paper/Output/site-plot.pdf}
 \caption{\label{fig:fig-obr}Default \texttt{dfoliatR} graphic for site-level outbreak events, produced by the \texttt{plot\_outbreak()} function.
@@
 \centering
-\includegraphics{D:/Dropbox/dfoliatR_paper/paper_elsevier/Output/bridging_color.pdf}
+\includegraphics{D:/Dropbox/dfoliatR_paper/Output/bridging_color.pdf}
 \caption{\label{fig:fig-bridge}Examples of identified defoliation events on individual trees.
@@
 \centering
-\includegraphics{D:/Dropbox/dfoliatR_paper/paper_elsevier/Output/plot-sites-df-obr.pdf}
+\includegraphics{D:/Dropbox/dfoliatR_paper/Output/plot-sites-df-obr.pdf}
 \caption{\label{fig:fig-df-obr}Comparison of reconstructed western spruce budworm outbreaks computed by \texttt{dfoliatR} and OUTBREAK.
@@
 Saving the results of \texttt{outbreak\_stats()} (the \texttt{dmj\_obr\_stats} object above) provides an array of options for assessing metrics of the insect outbreak regime.
-For example, taking the first year of each outbreak event, we can calculate the average duration of years between outbreaks, via the \texttt{diff()} function in \texttt{R}.
+For example, taking the first year of each outbreak event, we can calculate the duration of years between outbreaks, via the \texttt{diff()} function in \texttt{R}.
 The average of those differences, calculated via \texttt{mean()} is the mean return interval of reconstructed outbreak events at the DMJ site.

### 2020-05-21 README.Rmd
add readme

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# `dfoliatR`: An R package for detection and analysis of insect defoliation signals in tree rings

<!-- badges: start -->
<!-- badges: end -->

A repository for the initial descriptive paper of the [`dfoliatR`](https://chguiterman.github.io/dfoliatR/) package in R. The paper is currently being prepared for submission to *Dendrochronologia*.

```{r, echo=FALSE}
knitr::include_graphics("dfoliatR_ms.pdf")
```

Please contact Chris Guiterman with questions or comments.

### 2020-08-04 R/Compare_OUTBREAK_dfoliatR.R
clean and update

@@ -6,15 +6,14 @@ library(dplyr)
 library(purrr)
 library(tibble)
 library(tidyr)
+library(here)
 library(ggplot2)
 library(cowplot)
 # Load read_OUTBREAK function
 devtools::source_gist("https://gist.github.com/chguiterman/c88e9601dd085eb4215f0dc133a0e21c", sha1 = "fdb3cfefcd7e6ed3f7df3e20057c815bbd91dd0c")
 
-data_pth <- "paper_elsevier/Data/"
-
 ## Master metadata file
-meta_in <- read.csv(paste0(data_pth, "Meta.csv"))
+meta_in <- read.csv(here("Data", "Meta.csv"))
 
 ## Helper function
 rm_sample_depth <- function(x){
@@ -29,10 +28,10 @@ rm_sample_depth <- function(x){
 
 outbr_results <- meta_in %>% 
   transmute(site, 
-            outbr_cor = map(outbr_corrected, ~ read.compact(paste0(data_pth, .))),
-            outbr_col = map(outbr_site, ~ read.delim(paste0(data_pth, .))),
+            outbr_cor = map(outbr_corrected, ~ read.compact(here("Data", .))),
+            outbr_col = map(outbr_site, ~ read.delim(here("Data", .))),
             outbr_out = map2(outbr_file, insect_code, 
-                             ~ read_OUTREAK(fname = paste0(data_pth, .x),
+                             ~ read_OUTREAK(fname = here("Data", .x),
                                             level = "tree",
                                             insect_code = .y)))
 
@@ -71,8 +70,8 @@ outbr_obr <- outbr_results %>%
 
 df_run <- meta_in %>% 
   transmute(site,
-            host = map(host_file, ~ read.compact(paste0(data_pth, .))),
-            nonhost = map(nonhost_file, ~ read.crn(paste0(data_pth, .)) %>% 
+            host = map(host_file, ~ read.compact(here("Data", .))),
+            nonhost = map(nonhost_file, ~ read.crn(here("Data", .)) %>% 
                             rm_sample_depth()),
             defol = pmap(list(host, nonhost, ongoing_defol),
                          ~ defoliate_trees(host_tree = ..1, 
@@ -110,7 +109,7 @@ ggplot(tree_data, aes(x = ncor, y = ngsi, group = series)) +
   ylab("dfoliatR NGSI") +
   theme_bw() +
[…]

### 2020-08-04 R/bridging.R

@@ -26,6 +26,9 @@ rm_sample_depth <- function(x){
   return(df)
 }
 
+
+# Perform reconstructions -------------------------------------------------
+
 df_run <- meta_in %>% 
   transmute(site,
             host = map(host_file, ~ read.compact(here("paper_elsevier", "Data", .))),

### 2020-08-04 R/bridging.R
update paths

@@ -13,9 +13,9 @@ library(paletteer)
 library(here)
 
 ## Master metadata file
-meta_in <- read.csv(here("paper_elsevier", "Data", "Meta.csv"))
+meta_in <- read.csv(here("Data", "Meta.csv"))
 # file of output from the comparison -- has the OBR results
-obr_results <- read.csv(here("paper_elsevier", "Output", "Tree-comps-data.csv")) %>% 
+obr_results <- read.csv(here("Output", "Tree-comps-data.csv")) %>% 
   select(site, year, series, obr_type)
 
 ## Helper function
@@ -31,8 +31,8 @@ rm_sample_depth <- function(x){
 
 df_run <- meta_in %>% 
   transmute(site,
-            host = map(host_file, ~ read.compact(here("paper_elsevier", "Data", .))),
-            nonhost = map(nonhost_file, ~ read.crn(here("paper_elsevier", "Data", .)) %>% 
+            host = map(host_file, ~ read.compact(here("Data", .))),
+            nonhost = map(nonhost_file, ~ read.crn(here("Data", .)) %>% 
                             rm_sample_depth()),
             defol = pmap(list(host, nonhost, ongoing_defol),
                          ~ defoliate_trees(host_tree = ..1, 
@@ -80,7 +80,8 @@ ggplot(comps, aes(x=year)) +
   geom_area(data=df_events, aes(x = year, y = ngsi, fill = "dfoliatR")) +
   geom_area(data=obr_events, aes(x = year, y = ngsi, fill = "OUTBREAK & dfoliatR")) +
   geom_line(aes(y=ngsi)) +
-  scale_fill_manual(values = c("grey60", "#1C86EE"), breaks = c("OUTBREAK & dfoliatR", "dfoliatR")) +
+  scale_fill_manual(values = c("grey60", "#1C86EE"), 
+                    breaks = c("OUTBREAK & dfoliatR", "dfoliatR")) +
   guides(fill = guide_legend(title = "Identified defoliation events: ")) +
   ylab("NGSI") + xlab("Year") +
   ggpubr::theme_pubr() +

### 2020-08-04 dfoliatR_ms.Rmd
add revision text. Take 1

@@
 The user defines a rule base specifying the magnitude and duration that a period of indexed growth suppression must meet or surpass for a period of suppressed growth to be inferred as a defoliation event at the tree level.
+Rule bases are derived from the user’s knowledge of insect and host ecologies, and from consideration of the likelihood and relative importance of Type I and II errors.
 
@@
 Our evaluation of the two programs revealed that `dfoliatR` excelled in identifying defoliation events on single trees, providing researchers with more consistent and biologically-justifiable results.
-Other key benefits are that `dfoliatR` provides easier control of suppression thresholds, additional output tables, high-quality and customizable graphics, and operates in the open-source `R` environment that is stable across computing platforms and is under active development and maintenance by a large and growing community.
+`dfoliatR` provides easier control of the rule base for suppression thresholds, additional output tables, and high-quality and customizable graphics.
+These features allow users to compare regimes of different tree species or geographic regions, evaluate sample-size considerations, examine a multitude of relevant insect disturbance questions, and more readily test the effects of different rule-base thresholds on the likelihood of Type I and II errors.
+Finally, `dfoliatR` operates in the open-source `R` environment that is stable across computing platforms and is under active development and maintenance by a large and growing community.
 

### 2020-08-24 dfoliatR_ms.Rmd
Last edits before resubmission/acceptance

@@
 It is built on the long-accepted host to non-host comparison methodology used in the 1980s FORTRAN program OUTBREAK [@Swetnam1985; @Swetnam1989; @Swetnam1993].
-Our evaluation of the two programs revealed that `dfoliatR` excelled in identifying defoliation events on single trees, providing researchers with more consistent and biologically-justifiable results.
+Our evaluation of the two programs revealed that `dfoliatR` excels in identifying defoliation events on single trees, providing researchers with more consistent and biologically-justifiable results.
 `dfoliatR` provides easier control of the rule base for suppression thresholds, additional output tables, and high-quality and customizable graphics.
-These features allow users to compare regimes of different tree species or geographic regions, evaluate sample-size considerations, examine a multitude of relevant insect disturbance questions, and more readily test the effects of different rule-base thresholds on the likelihood of Type I and II errors.
+These features allow users to compare insect outbreak regimes of different tree species or geographic regions, evaluate sample-size considerations, examine a multitude of relevant insect disturbance questions, and more readily evaluate the potential for Type I and II errors in their results.
 Finally, `dfoliatR` operates in the open-source `R` environment that is stable across computing platforms and is under active development and maintenance by a large and growing community.

### 2020-08-24 dfoliatR_ms.tex
Last edits before resubmission/acceptance

@@
 The user defines a rule base specifying the magnitude and duration that a period of indexed growth suppression must meet or surpass for a period of suppressed growth to be inferred as a defoliation event at the tree level.
+Rule bases are derived from the user's knowledge of insect and host ecologies, and from consideration of the likelihood and relative importance of Type I and II errors.
 
@@
 It is built on the long-accepted host to non-host comparison methodology used in the 1980s FORTRAN program OUTBREAK (Swetnam et al., 1985; Swetnam and Lynch, 1989, 1993).
-Our evaluation of the two programs revealed that \texttt{dfoliatR} excelled in identifying defoliation events on single trees, providing researchers with more consistent and biologically-justifiable results.
-Other key benefits are that \texttt{dfoliatR} provides easier control of suppression thresholds, additional output tables, high-quality and customizable graphics, and operates in the open-source \texttt{R} environment that is stable across computing platforms and is under active development and maintenance by a large and growing community.
+Our evaluation of the two programs revealed that \texttt{dfoliatR} excels in identifying defoliation events on single trees, providing researchers with more consistent and biologically-justifiable results.
+\texttt{dfoliatR} provides easier control of the rule base for suppression thresholds, additional output tables, and high-quality and customizable graphics.
+These features allow users to compare insect outbreak regimes of different tree species or geographic regions, evaluate sample-size considerations, examine a multitude of relevant insect disturbance questions, and more readily evaluate the potential for Type I and II errors in their results.
+Finally, \texttt{dfoliatR} operates in the open-source \texttt{R} environment that is stable across computing platforms and is under active development and maintenance by a large and growing community.
 
@@
 Guiterman, C., Lynch, A., Axelson, J., 2020.
-\texttt{dfoliatR}: Detection and Analysis of Insect Defoliation Signals in Tree Rings., \url{https://chguiterman.github.io/dfoliatR/}, doi:\href{https://doi.org/10.5281/zenodo.3698163}{10.5281/zenodo.3698163}.
+\texttt{dfoliatR}: Detection and Analysis of Insect Defoliation Signals in Tree Rings.
+R package version 0.2.0., \url{https://chguiterman.github.io/dfoliatR/}, doi:\href{https://doi.org/10.5281/zenodo.3626135}{10.5281/zenodo.3626135}.
 

### 2020-08-28 R/Compare_OUTBREAK_dfoliatR.R
output final versions

@@ -166,5 +166,6 @@ ggplot(site_perc, aes(x = year, y = perc_defol)) +
         plot.margin = unit(c(30, 5, 5, 5), units = "points"),
         plot.background = element_blank())
 ggsave(here("Output", "plot-sites-df-obr.pdf"), width=5.75, dpi=300, units='in')  
+ggsave(here("Output", "plot-sites-df-obr.tiff"), width=5.75, dpi=300, units='in')  
 
 

### 2020-08-28 R/Example_Usage.R

@@ -23,6 +23,8 @@ write.csv(dmj_stats, here("Output", "defol-stats.csv"), row.names = FALSE)
 
 plot_defol(dmj_defol)
 ggsave(here("Output", "tree-plot-default.pdf"), width=5.75, dpi=300, units="in")
+ggsave(here("Output", "tree-plot-default.tiff"), width=5.75, dpi=300, units="in")
+ggsave(here("Output", "tree-plot-default.eps"), width=5.75, dpi=300, units="in")
 
 plot_defol(dmj_defol) +
   scale_color_manual(values = c("red", "orange", "purple")) 
@@ -36,6 +38,7 @@ dmj_obr <- outbreak(dmj_defol,
 
 plot_outbreak(dmj_obr, disp_index = "GSI")
 ggsave(here("Output", "site-plot.pdf"), width=5.75, dpi=300, units="in")
+ggsave(here("Output", "site-plot.tiff"), width=5.75, dpi=300, units="in")
 
 dmj_obr_stats <- outbreak_stats(dmj_obr)
 write.csv(dmj_obr_stats, here("Output", "obr-stats.csv"), row.names = FALSE)

### 2020-08-28 R/bridging.R

@@ -88,4 +88,5 @@ ggplot(comps, aes(x=year)) +
   theme(legend.position = "bottom",
         strip.text = element_text(face="bold", size=12),
         axis.title = element_text(face="bold", size = 12))
-ggsave(here("paper_elsevier", "Output", "bridging_color.pdf"), width = 5.75)
+ggsave(here("Output", "bridging_color.pdf"), width = 5.75)
+ggsave(here("Output", "bridging_color.tiff"), width = 5.75)

### 2020-08-30 dfoliatR_ms.Rmd
Add reference

@@
     affiliation: a,c
-    # footnote: [2, 3]
   - name: Jodi N. Axelson
@@
     affiliation: d
-    # footnote: 4
 address:
@@
   <!-- \newpage -->
-# journal: "Dendrochronologia"
-# date: "`r Sys.Date()`"
 bibliography: References.bib
 linenumbers: true
-#numbersections: true
 csl: dendrochronologia.csl
-# output: rticles::elsevier_article
 output:
@@
       base_format: rticles::elsevier_article
-# bookdown::word_document2:
-#     reference_docx: "../common/word-styles-reference-01.docx"
-#     toc: false
-# #     toc_depth: 2
-#     fig_caption: yes
-#     df_print: kable
-# always_allow_html: true
 layout: review
@@
 ```
+> This article is published: Guiterman, CH, AM Lynch, and JN Axelson (2020) `dfoliatR`: An R package for detection and analysis of insect defoliation signals in tree rings.
+_Dendrochronologia_.
+DOI: 10.1016/j.dendro.2020.125750.
+> Available Open Access from [https://authors.elsevier.com/sd/article/S1125786520300898](https://authors.elsevier.com/sd/article/S1125786520300898)
+
 Introduction

### 2020-08-30 dfoliatR_ms.tex
Add reference

@@
 \usepackage{graphicx}
-% We will generate all images so they have a width \maxwidth. This means
-% that they will get their normal width if they fit onto the page, but
-% are scaled down if they would overflow the margins.
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
-\else\Gin@nat@width\fi}
-\makeatother
-\let\Oldincludegraphics\includegraphics
-\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
 \ifxetex
@@
 
+\begin{quote}
+This article is published: Guiterman, CH, AM Lynch, and JN Axelson (2020) \texttt{dfoliatR}: An R package for detection and analysis of insect defoliation signals in tree rings.
+\emph{Dendrochronologia}.
+DOI: 10.1016/j.dendro.2020.125750.
+Available Open Access from \url{https://authors.elsevier.com/sd/article/S1125786520300898}
+\end{quote}
+
 \hypertarget{introduction}{%
@@
 \centering
-\includegraphics{D:/Dropbox/dfoliatR_paper/Output/tree-plot-default.pdf}
+\includegraphics{C:/Users/chris/Dropbox/dfoliatR_paper/Output/tree-plot-default.pdf}
 \caption{\label{fig:fig-defol}Default \texttt{dfoliatR} graphic for individual trees, produced by the \texttt{plot\_defol()} function.
@@
 \centering
-\includegraphics{D:/Dropbox/dfoliatR_paper/Output/site-plot.pdf}
+\includegraphics{C:/Users/chris/Dropbox/dfoliatR_paper/Output/site-plot.pdf}
 \caption{\label{fig:fig-obr}Default \texttt{dfoliatR} graphic for site-level outbreak events, produced by the \texttt{plot\_outbreak()} function.
@@
 \centering
-\includegraphics{D:/Dropbox/dfoliatR_paper/Output/bridging_color.pdf}
+\includegraphics{C:/Users/chris/Dropbox/dfoliatR_paper/Output/bridging_color.pdf}
 \caption{\label{fig:fig-bridge}Examples of identified defoliation events on individual trees.
@@
 \centering
-\includegraphics{D:/Dropbox/dfoliatR_paper/Output/plot-sites-df-obr.pdf}
+\includegraphics{C:/Users/chris/Dropbox/dfoliatR_paper/Output/plot-sites-df-obr.pdf}
 \caption{\label{fig:fig-df-obr}Comparison of reconstructed western spruce budworm outbreaks computed by \texttt{dfoliatR} and OUTBREAK.

### 2020-09-02 README.Rmd
update readme
updates for proof and publication

@@
 
-# `dfoliatR`: An R package for detection and analysis of insect defoliation signals in tree rings
+### `dfoliatR`: An R package for detection and analysis of insect defoliation signals in tree rings
 
@@
 
-A repository for the initial descriptive paper of the [`dfoliatR`](https://chguiterman.github.io/dfoliatR/) package in R. The paper is currently being prepared for submission to *Dendrochronologia*.
+This repository includes the code and text to generate a preprint version of the published paper:
 
-```{r, echo=FALSE}
-knitr::include_graphics("dfoliatR_ms.pdf")
-```
+>Guiterman, CH, AM Lynch, and JN Axelson (2020) [`dfoliatR`: An R package for detection and analysis of insect defoliation signals in tree rings.](https://authors.elsevier.com/sd/article/S1125786520300898) _Dendrochronologia_.
+DOI: 10.1016/j.dendro.2020.125750.
 
-Please contact Chris Guiterman with questions or comments.
+The paper provides an overview of the `dfoliatR` R package. More information on using the software is availabe from [https://chguiterman.github.io/dfoliatR/](https://chguiterman.github.io/dfoliatR/).
+
+Please contact @chguiterman with questions or comments.

### 2020-09-02 dfoliatR_ms.Rmd
updates for proof and publication

@@
 ```
-> This article is published: Guiterman, CH, AM Lynch, and JN Axelson (2020) `dfoliatR`: An R package for detection and analysis of insect defoliation signals in tree rings.
+> This article is published as: Guiterman, CH, AM Lynch, and JN Axelson (2020) `dfoliatR`: An R package for detection and analysis of insect defoliation signals in tree rings.
 _Dendrochronologia_.
 DOI: 10.1016/j.dendro.2020.125750.
-> Available Open Access from [https://authors.elsevier.com/sd/article/S1125786520300898](https://authors.elsevier.com/sd/article/S1125786520300898)
+Available for free [(link)](https://authors.elsevier.com/sd/article/S1125786520300898)
 
@@
 dmj_defol_stats <- defol_stats(dmj_defol)
-View(dmj_defol_stats)
+View(dmj_defol_stats) # Table 1
 ```
@@
 ```{r echo=TRUE, eval=FALSE}
-plot_defol(dmj_defol)
+plot_defol(dmj_defol) # Fig. 1
 ## This creates a ggplot object. Additions can 
@@
 plot_defol(dmj_defol) +
-  scale_color_manual(values = c("red", "orange", "purple"))
+  scale_color_manual(values = c("red", "orange", 
+                                "purple"))
 ```
@@
 The "mean_gsi" and "mean_ngsi" variables provide averages of these indices across all available trees.
-Finally, the "outbreak_status" column shows whether an outbreak event is inferred ("outbreak") or not ("not_obr").
+Finally, the "outbreak_status" column shows if an outbreak event is inferred ("outbreak") or not ("not_obr"), and whether it represents an ongoing series-end event ("se_outbreak").
 
@@
 ```{r echo=TRUE, eval=FALSE}
-plot_outbreak(dmj_obr, disp_index = "GSI")
+plot_outbreak(dmj_obr, disp_index = "GSI") # Fig. 2
 ```
@@
 dmj_obr_stats <- outbreak_stats(dmj_obr)
-View(dmj_obr_stats)
+View(dmj_obr_stats) # Table 2
 ```
@@
 ```{r echo=TRUE, eval=FALSE}
-## calculate mean return interval for outbreak start years
+## calculate mean return interval for outbreak 
+## start years
 intervals <- diff(dmj_obr_stats$start)

### 2020-09-02 dfoliatR_ms.tex
updates for proof and publication

--- 2020-08-24 dfoliatR_ms.tex
@@
  \end{frontmatter}
+
+\begin{quote}
+This article is published as: Guiterman, CH, AM Lynch, and JN Axelson (2020) \texttt{dfoliatR}: An R package for detection and analysis of insect defoliation signals in tree rings.
+\emph{Dendrochronologia}.
+DOI: 10.1016/j.dendro.2020.125750.
+Available for free \href{https://authors.elsevier.com/sd/article/S1125786520300898}{(link)}
+\end{quote}
 
@@
 \NormalTok{dmj_defol_stats <-}\StringTok{ }\KeywordTok{defol_stats}\NormalTok{(dmj_defol)}
-\KeywordTok{View}\NormalTok{(dmj_defol_stats)}
+\KeywordTok{View}\NormalTok{(dmj_defol_stats) }\CommentTok{# Table 1}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\KeywordTok{plot_defol}\NormalTok{(dmj_defol)}
+\KeywordTok{plot_defol}\NormalTok{(dmj_defol) }\CommentTok{# Fig. 1}
 \CommentTok{## This creates a ggplot object. Additions can }
@@
 \KeywordTok{plot_defol}\NormalTok{(dmj_defol) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{scale_color_manual}\NormalTok{(}\DataTypeTok{values =} \KeywordTok{c}\NormalTok{(}\StringTok{"red"}\NormalTok{, }\StringTok{"orange"}\NormalTok{, }\StringTok{"purple"}\NormalTok{))}
+\StringTok{  }\KeywordTok{scale_color_manual}\NormalTok{(}\DataTypeTok{values =} \KeywordTok{c}\NormalTok{(}\StringTok{"red"}\NormalTok{, }\StringTok{"orange"}\NormalTok{, }
+                                \StringTok{"purple"}\NormalTok{))}
 \end{Highlighting}
@@
 The ``mean\_gsi'' and ``mean\_ngsi'' variables provide averages of these indices across all available trees.
-Finally, the ``outbreak\_status'' column shows whether an outbreak event is inferred (``outbreak'') or not (``not\_obr'').
+Finally, the ``outbreak\_status'' column shows if an outbreak event is inferred (``outbreak'') or not (``not\_obr''), and whether it represents an ongoing series-end event (``se\_outbreak'').
 
@@
 \begin{Highlighting}[]
-\KeywordTok{plot_outbreak}\NormalTok{(dmj_obr, }\DataTypeTok{disp_index =} \StringTok{"GSI"}\NormalTok{)}
+\KeywordTok{plot_outbreak}\NormalTok{(dmj_obr, }\DataTypeTok{disp_index =} \StringTok{"GSI"}\NormalTok{) }\CommentTok{# Fig. 2}
 \end{Highlighting}
@@
 \NormalTok{dmj_obr_stats <-}\StringTok{ }\KeywordTok{outbreak_stats}\NormalTok{(dmj_obr)}
-\KeywordTok{View}\NormalTok{(dmj_obr_stats)}
+\KeywordTok{View}\NormalTok{(dmj_obr_stats) }\CommentTok{# Table 2}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\CommentTok{## calculate mean return interval for outbreak start years}
+\CommentTok{## calculate mean return interval for outbreak }
+\CommentTok{## start years}
 \NormalTok{intervals <-}\StringTok{ }\KeywordTok{diff}\NormalTok{(dmj_obr_stats}\OperatorTok{$}\NormalTok{start)}

### 2020-09-02 dfoliatR_ms.tex

@@
 The North American Drought Atlas.
-(Lamont- Doherty Earth Observatory; the National Science Foundation: New York) Available at https://www.ncdc.noaa.gov/data-access/paleoclimatology-data/datasets/tree-ring/drought-variability {[}Last accessed May 2020{]}.
+(Lamont-Doherty Earth Observatory; the National Science Foundation: New York) Available at https://www.ncdc.noaa.gov/data-access/paleoclimatology-data/datasets/tree-ring/drought-variability {[}Last accessed May 2020{]}.
 

### 2020-09-14 dfoliatR_ms.tex
no real change

@@
 % Pandoc toggle for numbering sections (defaults to be off)
+
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newenvironment{cslreferences}%
+  {}%
+  {\par}
 
@@
 \begin{Highlighting}[]
-\NormalTok{devtools}\OperatorTok{::}\KeywordTok{install_github}\NormalTok{(}\StringTok{"chguiterman/dfoliatR"}\NormalTok{)}
+\NormalTok{devtools}\OperatorTok{::}\KeywordTok{install\_github}\NormalTok{(}\StringTok{"chguiterman/dfoliatR"}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\CommentTok{## Load the package}
+\CommentTok{\#\# Load the package}
 \KeywordTok{library}\NormalTok{(dfoliatR)}
 
-\CommentTok{## Load data}
-\KeywordTok{data}\NormalTok{(}\StringTok{"dmj_h"}\NormalTok{) }\CommentTok{# host tree series}
-\KeywordTok{data}\NormalTok{(}\StringTok{"dmj_nh"}\NormalTok{) }\CommentTok{# non-host chronology}
-\CommentTok{# View data structure for any data object via }
-\CommentTok{# "View(dmj_h)" or "head(dmj_h)"}
+\CommentTok{\#\# Load data}
+\KeywordTok{data}\NormalTok{(}\StringTok{"dmj\_h"}\NormalTok{) }\CommentTok{\# host tree series}
+\KeywordTok{data}\NormalTok{(}\StringTok{"dmj\_nh"}\NormalTok{) }\CommentTok{\# non{-}host chronology}
+\CommentTok{\# View data structure for any data object via }
+\CommentTok{\# "View(dmj\_h)" or "head(dmj\_h)"}
 
-\CommentTok{## Calculate indices, identify defoliation events}
-\NormalTok{dmj_defol <-}\StringTok{ }\KeywordTok{defoliate_trees}\NormalTok{(}\DataTypeTok{host_tree =}\NormalTok{ dmj_h,}
-                             \DataTypeTok{nonhost_chron =}\NormalTok{ dmj_nh,}
-                             \DataTypeTok{duration_years =} \DecValTok{8}\NormalTok{,  }
-                             \DataTypeTok{max_reduction =} \FloatTok{-1.28}\NormalTok{, }
-                             \DataTypeTok{bridge_events =} \OtherTok{TRUE}\NormalTok{, }
-                             \DataTypeTok{series_end_event =} \OtherTok{TRUE}\NormalTok{, }
-                             \DataTypeTok{list_output =} \OtherTok{FALSE}\NormalTok{) }
+\CommentTok{\#\# Calculate indices, identify defoliation events}
+\NormalTok{dmj\_defol \textless{}{-}}\StringTok{ }\KeywordTok{defoliate\_trees}\NormalTok{(}\DataTypeTok{host\_tree =}\NormalTok{ dmj\_h,}
+                             \DataTypeTok{nonhost\_chron =}\NormalTok{ dmj\_nh,}
+                             \DataTypeTok{duration\_years =} \DecValTok{8}\NormalTok{,  }
+                             \DataTypeTok{max\_reduction =} \FloatTok{{-}1.28}\NormalTok{, }
+                             \DataTypeTok{bridge\_events =} \OtherTok{TRUE}\NormalTok{, }
+                             \DataTypeTok{series\_end\_event =} \OtherTok{TRUE}\NormalTok{, }
+                             \DataTypeTok{list\_output =} \OtherTok{FALSE}\NormalTok{) }
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{dmj_defol_stats <-}\StringTok{ }\KeywordTok{defol_stats}\NormalTok{(dmj_defol)}
-\KeywordTok{View}\NormalTok{(dmj_defol_stats) }\CommentTok{# Table 1}
+\NormalTok{dmj\_defol\_stats \textless{}{-}}\StringTok{ }\KeywordTok{defol\_stats}\NormalTok{(dmj\_defol)}
+\KeywordTok{View}\NormalTok{(dmj\_defol\_stats) }\CommentTok{\# Table 1}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\KeywordTok{plot_defol}\NormalTok{(dmj_defol) }\CommentTok{# Fig. 1}
-\CommentTok{## This creates a ggplot object. Additions can }
-\CommentTok{## be made to adjust plotting themes and}
-\CommentTok{## aesthetics, like color.   }
-\CommentTok{## For more on the features of ggplot see}
-\CommentTok{## https://ggplot2.tidyverse.org/}
-\CommentTok{## Adding 'scale_color_manual()' will}
-\CommentTok{## change the colors that denote }
-\CommentTok{## the average NGSI value for each defoliation}
-\CommentTok{## event.}
+\KeywordTok{plot\_defol}\NormalTok{(dmj\_defol) }\CommentTok{\# Fig. 1}
+\CommentTok{\#\# This creates a ggplot object. Additions can }
+\CommentTok{\#\# be made to adjust plotting themes and}
+\CommentTok{\#\# aesthetics, like color.   }
+\CommentTok{\#\# For more on the features of ggplot see}
+\CommentTok{\#\# https://ggplot2.tidyverse.org/}
+\CommentTok{\#\# Adding \textquotesingle{}scale\_color\_manual()\textquotesingle{} will}
+\CommentTok{\#\# change the colors that denote }
+\CommentTok{\#\# the average NGSI value for each defoliation}
+\CommentTok{\#\# event.}
 \KeywordTok{library}\NormalTok{(ggplot2)}
-\KeywordTok{plot_defol}\NormalTok{(dmj_defol) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{scale_color_manual}\NormalTok{(}\DataTypeTok{values =} \KeywordTok{c}\NormalTok{(}\StringTok{"red"}\NormalTok{, }\StringTok{"orange"}\NormalTok{, }
+\KeywordTok{plot\_defol}\NormalTok{(dmj\_defol) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{scale\_color\_manual}\NormalTok{(}\DataTypeTok{values =} \KeywordTok{c}\NormalTok{(}\StringTok{"red"}\NormalTok{, }\StringTok{"orange"}\NormalTok{, }
                                 \StringTok{"purple"}\NormalTok{))}
@@
 \begin{Highlighting}[]
-\CommentTok{## Use the defol object "dmj_defol" created above}
-\NormalTok{dmj_obr <-}\StringTok{ }\KeywordTok{outbreak}\NormalTok{(dmj_defol,}
-                    \DataTypeTok{filter_min_series =} \DecValTok{3}\NormalTok{,}
-                    \DataTypeTok{filter_min_defol =} \DecValTok{1}\NormalTok{,}
-                    \DataTypeTok{filter_perc =} \DecValTok{25}\NormalTok{)}
+\CommentTok{\#\# Use the defol object "dmj\_defol" created above}
+\NormalTok{dmj\_obr \textless{}{-}}\StringTok{ }\KeywordTok{outbreak}\NormalTok{(dmj\_defol,}
+                    \DataTypeTok{filter\_min\_series =} \DecValTok{3}\NormalTok{,}
+                    \DataTypeTok{filter\_min\_defol =} \DecValTok{1}\NormalTok{,}
+                    \DataTypeTok{filter\_perc =} \DecValTok{25}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\KeywordTok{plot_outbreak}\NormalTok{(dmj_obr, }\DataTypeTok{disp_index =} \StringTok{"GSI"}\NormalTok{) }\CommentTok{# Fig. 2}
+\KeywordTok{plot\_outbreak}\NormalTok{(dmj\_obr, }\DataTypeTok{disp\_index =} \StringTok{"GSI"}\NormalTok{) }\CommentTok{\# Fig. 2}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{dmj_obr_stats <-}\StringTok{ }\KeywordTok{outbreak_stats}\NormalTok{(dmj_obr)}
-\KeywordTok{View}\NormalTok{(dmj_obr_stats) }\CommentTok{# Table 2}
+\NormalTok{dmj\_obr\_stats \textless{}{-}}\StringTok{ }\KeywordTok{outbreak\_stats}\NormalTok{(dmj\_obr)}
+\KeywordTok{View}\NormalTok{(dmj\_obr\_stats) }\CommentTok{\# Table 2}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\CommentTok{## calculate mean return interval for outbreak }
-\CommentTok{## start years}
-\NormalTok{intervals <-}\StringTok{ }\KeywordTok{diff}\NormalTok{(dmj_obr_stats}\OperatorTok{$}\NormalTok{start)}
-\KeywordTok{mean}\NormalTok{(intervals) }\CommentTok{## 51.2 year mean return interval}
+\CommentTok{\#\# calculate mean return interval for outbreak }
+\CommentTok{\#\# start years}
+\NormalTok{intervals \textless{}{-}}\StringTok{ }\KeywordTok{diff}\NormalTok{(dmj\_obr\_stats}\OperatorTok{$}\NormalTok{start)}
+\KeywordTok{mean}\NormalTok{(intervals) }\CommentTok{\#\# 51.2 year mean return interval}
 \end{Highlighting}
@@
 \hypertarget{refs}{}
+\begin{cslreferences}
 \leavevmode\hypertarget{ref-Alfaro1982}{}%
@@
 Ecography 38, 431--36, doi:\href{https://doi.org/10.1111/ecog.01335}{10.1111/ecog.01335}.
+\end{cslreferences}
 
